### 2016-02-15 additional_files/tex/file1_data_curation.tex
added distributions document from from mail; started drafting additional file 1; changed the read number filter; played with RUV

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font

%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{Curating RNAseq data for mouse and \textit{Eimeria falciformis}} % Title

\author{Emanuel \textsc{Heitlinger}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Objective}

In the differential expression analysis of RNAseq data we assume a
negative binomial distribution of transcript abundance. We analyse the
distributions in mouse and parasite data respectively. A cutoff is
applied to reads mapping per gene in raw data. For instance, a cutoff
of ten requires that at least ten reads from \textit{any sample} map
to a given gene for the gene to count as expressed. We evaluate the
effects of different cutoffs on the distribution.

%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION 2 
%----------------------------------------------------------------------------------------

\section{Results for mouse}

% figure 1
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for all
  mouse samples in analysis. All samples share the same bimodal
  distribution trend by visual inspection. The first density peak
  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
  detected between 1 and 100 times in that sample. The second density
  peak occurs in the range of $10^3 - 10^4$.}
\end{center}
\end{figure}

% figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for all mouse samples. All samples share the same bimodal
  distribution trend by visual inspection. The first density peak
  occurs at $<1$ transcripts, and in most samples it is more distinct
  than in the non-normalised data in Figure 1. For this normalised
  data, the second density peak occurs at roughly the same count value
  in all samples, close to $10^2$ transcripts.}
\end{center}
\end{figure}

% figure 3
\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.8\textwidth]{distributions_normmouse100} % Include the image placeholder.png
  \caption{Density distribution of normalised transcript counts
    (log10) for mouse samples with a cutoff of 100 (see Methods for
    details). All samples still have a bimodal distribution, however
    less pronounced compared to figures 1 and 2. The first density
    peak occurs at $<1$ transcripts as in Figure 2. The second density
    peak occurs at close to $10^2$ transcripts and did hence not
    change from the normalised data without cutoff (Figure 2).}
\end{center}
\end{figure}

% figure 4
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse2000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 2000 (see Methods for
  details). The bimodality seen at lower cutoff values is strongly
  weakened or not detectable visually in some samples. It is however
  visually still seen in a majority of the samples. The first tendency
  to a density peak has shifted somewhat towards higher transcript
  counts (right) but is still $<1$ transcripts as in Figure 3. The
  second density peak occurs at close to $10^2$ transcripts and did
  hence not change from the normalised data with a lower cutoff
  (Figure 3).}
\end{center}
\end{figure}

% figure 5
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse5000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 5000 (see Methods for
  details). The bimodality is not visually detectable in most samples,
  however it is still seen in \textit{e.g.} samples
  NMRI\_1stInf\_3dpi\_rep1 and NMRI\_2ndInf\_7dpi\_rep2. The major
  density peak remains at a value close to $10^2$ as in all the
  analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

% figure 6
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_mouse5000} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  mouse samples with a cutoff of 5000 (see Methods for details). The
  bimodality is not visually detectable in most samples, however it is
  still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2. The major density peak is seen at a value
  over $10^2$, shifting it almost one order of magnitude compared to
  the normalised data with the same cutoff (Figure 5).}
\end{center}
\end{figure}

% figure 7
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse10000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 10000 (see Methods for
  details). The bimodality is not visually detectable in any sample
  apart from a minor indication in sample
  NMRI\_1stInf\_5dpi\_rep2. The major density peak remains at a value
  close to $10^2$ as in all the analysed normalised datasets
  (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

%----------------------------------------------------------------------------------------
%	SECTION 2 PARASITE DATA
%----------------------------------------------------------------------------------------

\section{Results for parasite}

% figure 8 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef5} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 5 (see Methods for details). Upon
  visual inspection distributions in most samples appear smooth and
  not contradictory to a negative binomial distribution. Density peak
  positions with regards to transcript counts (x-axis) are however at
  different orders of magnitudes between samples. Samples
  NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2, however, do
  not.}
\end{center}
\end{figure}

% figure 9
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm5} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 5 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth also in the normalised data. Density peaks are also in
  the same order of magnitude with regards to transcript counts
  (x-axis). Non-negative binomial samples in Figure 8
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) have more
  pronounced spikes in this normalised data. Additionally, samples
  Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 also have one
  additional peak in the lower range of the distribution, as do some
  other samples, but less pronounced.}
\end{center}
\end{figure}

% figure 10
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef50000} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples in Figure 8 and 9
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2 have more than
  one density peak also when lowly expressed genes are removed with a
  high threshold (50000). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2, however, seem smoother with only one
  density peak with this threshold.}
\end{center}
\end{figure}

% figure 11
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm50000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples are pronounced bimodal
  in this normalised data (NMRI\_2ndInf\_3dpi\_rep1,
  NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2 are also bimodal, however less pronounced.}
\end{center}
\end{figure}

%----------------------------------------------------------------------------------------
%	SECTION 5
%----------------------------------------------------------------------------------------

\section{Conclusions}

Mouse\\
The assumption of a negative binomial distribution underlies most
differential gene expression analyses available. Similar bimodal
expression patterns as in our mouse dataset have been reported
(\textit{e.g.} ....) and suggested to be a result of a group of
non-coding RNAs being sequenced. Since we are not interested in such
RNA species in this project, different cutoff values were tested to
find the lowest possible threshold for removing transcripts and gain a
negative binomial distribution. By visual inspection (again: should we
formalise/quantify this somehow?) we chose 5000 as the cutoff for
further analysis.\\

Parasite\\
The parasite data follows the expected distribution in most samples
(although we have not formally tested it - check mean-var relationship
to do that?). Bimodal or multipeak samples do not show unimodal
distributions even with high cutoffs, \textit{i.e.} many transcripts
required to map to one gene for the reads to be included. The four
most obvious spiky samples are also the samples with fewest reads
sequenced overall, which can explain the uneven distribution (make
table to refer to?). We will exclude these four samples
(NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) from all
further analysis. It is not clear why there is such low coverage in
these samples. Possibilities include poor infection success because of
too low oocyst injection, mouse resistance due to \textit{e.g.}
unknown previous infections or other unknown variability in the
experiment. Because of this insecurity, these samples are also
exculded from the mouse data.\\

The fact that the density peak occurs at similar read counts after
normalisation in mouse and parasite data (respectively) is encouraging
and indicates that normalisation is successful in both datasets.

%----------------------------------------------------------------------------------------
%	SECTION 6
%----------------------------------------------------------------------------------------

%\section{Answers to Definitions}

%\begin{enumerate}
%\begin{item}
%  The \emph{atomic weight of an element} is the relative weight of
%  one of its atoms compared to C-12 with a weight of
%  12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with
%  a weight of 16.00. Atomic weight is also the average weight of all
%  the atoms of that element as they occur in nature.
%\end{item}
%\begin{item}
%  The \emph{units of atomic weight} are two-fold, with an identical
%  numerical value. They are g/mole of atoms (or just g/mol) or
%  amu/atom.
%\end{item}
%\begin{item}
%  \emph{Percentage discrepancy} between an accepted (literature)
%  value and an experimental value is
%\begin{equation*}
%\frac{\mathrm{experimental\;result} - \mathrm{accepted\;result}}{\mathrm{accepted\;result}}
%\end{equation*}
%\end{item}
%\end{enumerate}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

%\bibliographystyle{apalike}

%\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}

### 2016-02-15 2_edgeR_diff.R

@@ -10,30 +10,30 @@ if(!exists("mouse.bg")){
 }
 
 raw.counts.4.bg <- function(bg){
-exon.raw.count <- eexpr(bg, "ucount")
-## the linkage data
-e2t <- bg@indexes$e2t
-t2g <- bg@indexes$t2g 
-e2t2g <- merge(e2t, t2g)
-all.count <- merge(e2t2g, exon.raw.count,
-                   by.x = "e_id", by.y = 0)
-## sum up for transcripts
-transcript.raw.count <-
-    do.call("rbind",
-            by(all.count, all.count$t_id,
-               function(x) colSums(x[, 4:ncol(x)])))
-## sum up for genes
-gene.raw.count <-
-    do.call("rbind",
-            by(all.count, all.count$g_id,
-               function(x) colSums(x[, 4:ncol(x)])))
-colnames(transcript.raw.count) <-
-    as.character(pData(bg)$samples)
-colnames(gene.raw.count) <-
-    as.character(pData(bg)$samples)
-return(list(exon.raw.count,
-            transcript.raw.count,
-            gene.raw.count))
+    exon.raw.count <- eexpr(bg, "ucount")
+    ## the linkage data
+    e2t <- bg@indexes$e2t
+    t2g <- bg@indexes$t2g 
+    e2t2g <- merge(e2t, t2g)
+    all.count <- merge(e2t2g, exon.raw.count,
+                       by.x = "e_id", by.y = 0)
+    ## sum up for transcripts
+    transcript.raw.count <-
+        do.call("rbind",
+                by(all.count, all.count$t_id,
+                   function(x) colSums(x[, 4:ncol(x)])))
+    ## sum up for genes
+    gene.raw.count <-
+        do.call("rbind",
+                by(all.count, all.count$g_id,
+                   function(x) colSums(x[, 4:ncol(x)])))
+    colnames(transcript.raw.count) <-
+        as.character(pData(bg)$samples)
+    colnames(gene.raw.count) <-
+        as.character(pData(bg)$samples)
+    return(list(exon.raw.count,
+                transcript.raw.count,
+                gene.raw.count))
 }
 
 mouse.RC <- raw.counts.4.bg(mouse.bg)
@@ -51,9 +51,10 @@ colnames(design)  <- gsub("pData\\(mouse.bg\\)\\$grouped", "",
                              colnames(design))
 ## just to have the standard point delimitor
[…]

### 2016-02-15 4_compareArray.R

@@ -79,12 +79,6 @@ RG.g <- RG.MA(MA.g)
 ES.g <- build.eset(RG.g, targets, makePLOT=FALSE,
                    annotation.package="mgug4122a.db")
 
-biolrep <- targets$SampleNumber
-
-design <- modelMatrix(targets, ref="Uninf")
-cont.matrix <- makeContrasts("X24h" , "X144h", "X24h-X144h",
-                             levels=design)
-
 corfit.g <- duplicateCorrelation(MA.g, design=design, ndups=1, block=biolrep)
 
 fit.g <- lmFit(MA.g, design=design, block=biolrep, cor=corfit.g$consensus)
@@ -98,27 +92,47 @@ RNAseq.logFC <- ALL.top[, c("logFC.N3vs0", "logFC.N5vs0",
                             "logFC.N7vs0", "logFC.B5vs0", "logFC.R5vs0" )]
 RNAseq.logFC <- merge(RNAseq.logFC, id2name, by.x = 0, by.y = "gene_ids")
 
+RNAseq.logFC.ruved <- ALL.top.ruved[, c("logFC.N3vs0", "logFC.N5vs0",
+                                        "logFC.N7vs0", "logFC.B5vs0", "logFC.R5vs0" )]
+RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, id2name, by.x = 0, by.y = "gene_ids")
+
 RNAseq.Array.logFC <- merge(RNAseq.logFC, Array.logFC, by.x = "gene_names", by.y = 0)
 
+RNAseq.Array.logFC <- merge(RNAseq.Array.logFC, RNAseq.logFC.ruved, by = c("Row.names","gene_names"))
+
+names(RNAseq.Array.logFC) <- gsub(".x", ".plain", names(RNAseq.Array.logFC))
+
+names(RNAseq.Array.logFC) <- gsub(".y", ".ruved", names(RNAseq.Array.logFC))
 
 ## Very interesting...
-cor(RNAseq.Array.logFC[,3:9])
-cor(RNAseq.Array.logFC[,3:9], method="spearman")
-
+cor(RNAseq.Array.logFC[,3:14])
+cor(RNAseq.Array.logFC[,3:14], method="spearman")
 
 pdf("figures/Array_vs_RNAseq_pairs.pdf")
[…]

### 2016-02-15 5_enrichment.R

@@ -15,7 +15,6 @@ if(!exists("gene2GO")){
 ## the universe of genes which were tested at all:
 exp.universe <- get.annotation.for.xloc(gene.list[[2]])[[2]]
 
-
 ## The CHALLENGE in NMRI DAY 7 
 N7.1stvsN7.2nd.GID <-
     get.annotation.for.xloc(gene.list[["N7.1stvsN7.2nd"]])[[2]]

### 2016-02-15 DEG_Ef.R

@@ -41,11 +41,13 @@ tableEfsmall <- tableEf[!(str_detect(tableEf$Samples, "^NMRI_(oocysts|sporozoite
 tableEfsmall$percent.Ef.reads <- (tableEfsmall[,1]/tableMsmall[,1])*100
 
 library(scales)
+
 ggplot()+
-	geom_point(data = tableEf, aes(x = Samples, y = as.numeric(colSums(Ef.RC[[3]])))) +
+    geom_point(data = tableEf, aes(x = Samples,
+                   y = as.numeric(colSums(Ef.RC[[3]])))) +
 	scale_y_log10(labels = comma) +
-	ylab("Number of reads") +
-	theme(axis.text.x = element_text(angle = 45, hjust = 1))
+            ylab("Number of reads") +
+                theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 #geom_point(data = tableM, aes(x = , y = as.numeric(colSums(Ef.RC[[3]])))
 

### 2016-02-15 RUVseq.R

@@ -1,0 +1,88 @@
+library(RUVSeq)
+
+if(!exists("mouse.bg")){
+    source("2_edgeR_diff.R")
+}
+
+Mm.differences <- makeGroups(pData(mouse.bg)$group)
+
+## use the expression set and the filtering decided on before
+Mm.RUVset <- newSeqExpressionSet(mouse.RC[[3]][keep,])
+
+pdf("figures/ruv_eval/RLE_wo_NORM.pdf")
+plotRLE(Mm.RUVset)
+dev.off()
+
+pdf("figures/ruv_eval/PCA_wo_NORM.pdf")
+plotPCA(Mm.RUVset)
+dev.off()
+
+pdf("figures/ruv_eval/mds_wo_NORM.pdf")
+plotMDS(counts(Mm.RUVset.groups))
+dev.off()
+
+Mm.RUVset.BLnorm <- betweenLaneNormalization(Mm.RUVset, which="upper")
+
+pdf("figures/ruv_eval/RLE_BL_NORM.pdf")
+plotRLE(Mm.RUVset.BLnorm)
+dev.off()
+
+pdf("figures/ruv_eval/PCA_BL_NORM.pdf")
+plotPCA(Mm.RUVset.BLnorm)
+dev.off()
+
+pdf("figures/ruv_eval/mds_BL_NORM.pdf")
+plotMDS(counts(Mm.RUVset.BLnorm))
+dev.off()
+
+Mm.RUVset.groups <- RUVs(Mm.RUVset.BLnorm, rownames(Mm.RUVset.BLnorm), k=1, Mm.differences)
+
+pdf("figures/ruv_eval/RLE_group_NORM.pdf") ## not very telling?!
+plotRLE(Mm.RUVset.groups)
+dev.off()
+
+pdf("figures/ruv_eval/PCA_group_NORM.pdf") ## wow that seems to work a treat
+plotPCA(Mm.RUVset.groups)
+dev.off()
+
+pdf("figures/ruv_eval/hclust_group_NORM.pdf") ## to much variation removed?
+plot(hclust(dist(t(normCounts(Mm.RUVset.groups)))))
+dev.off()
+
+pdf("figures/ruv_eval/mds_group_NORM.pdf")
+plotMDS(normCounts(Mm.RUVset.groups))
+dev.off()
+
+## empirically non-significant genes
+Nsig.genes <- rownames(ALL.top$table[ALL.top$table$PValue>0.5, ])
+
[…]

### 2016-03-24 additional_files/tex/cluster1_BP_Mm_EarlyUp.tex
cleaned up status but as on poster

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 06:00:18 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0045087 & innate immune response & 340 &  61 & 7.54 & 0.00 & 0.00 \\ 
  2 & GO:0006955 & immune response & 674 &  79 & 14.95 & 0.00 & 0.00 \\ 
  3 & GO:0006952 & defense response & 724 &  75 & 16.06 & 0.00 & 0.00 \\ 
  4 & GO:0002376 & immune system process & 1326 &  98 & 29.41 & 0.00 & 0.00 \\ 
  5 & GO:0098542 & defense response to other organism & 282 &  48 & 6.25 & 0.00 & 0.00 \\ 
  6 & GO:0043207 & response to external biotic stimulus & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
  7 & GO:0051707 & response to other organism & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
  8 & GO:0009607 & response to biotic stimulus & 460 &  55 & 10.20 & 0.00 & 0.00 \\ 
  9 & GO:0002252 & immune effector process & 440 &  48 & 9.76 & 0.00 & 0.00 \\ 
  10 & GO:0009615 & response to virus & 205 &  34 & 4.55 & 0.00 & 0.00 \\ 
  11 & GO:0051607 & defense response to virus & 181 &  32 & 4.01 & 0.00 & 0.00 \\ 
  12 & GO:0034097 & response to cytokine & 346 &  42 & 7.67 & 0.00 & 0.00 \\ 
  13 & GO:0034341 & response to interferon-gamma &  47 &  18 & 1.04 & 0.00 & 0.00 \\ 
  14 & GO:0006950 & response to stress & 2042 &  98 & 45.29 & 0.00 & 0.00 \\ 
  15 & GO:0035456 & response to interferon-beta &  33 &  14 & 0.73 & 0.00 & 0.00 \\ 
  16 & GO:0071345 & cellular response to cytokine stimulus & 276 &  32 & 6.12 & 0.00 & 0.00 \\ 
  17 & GO:0045088 & regulation of innate immune response & 146 &  24 & 3.24 & 0.00 & 0.00 \\ 
  18 & GO:0051704 & multi-organism process & 959 &  61 & 21.27 & 0.00 & 0.00 \\ 
  19 & GO:0009605 & response to external stimulus & 1135 &  67 & 25.17 & 0.00 & 0.00 \\ 
  20 & GO:0002682 & regulation of immune system process & 732 &  52 & 16.23 & 0.00 & 0.00 \\ 
  21 & GO:0048525 & negative regulation of viral process &  66 &  17 & 1.46 & 0.00 & 0.00 \\ 
  22 & GO:0019882 & antigen processing and presentation &  69 &  17 & 1.53 & 0.00 & 0.00 \\ 
  23 & GO:0035458 & cellular response to interferon-beta &  27 &  12 & 0.60 & 0.00 & 0.00 \\ 
  24 & GO:1903901 & negative regulation of viral life cycle &  63 &  16 & 1.40 & 0.00 & 0.00 \\ 
  25 & GO:0043901 & negative regulation of multi-organism pr... & 103 &  19 & 2.28 & 0.00 & 0.00 \\ 
  26 & GO:0050776 & regulation of immune response & 366 &  34 & 8.12 & 0.00 & 0.00 \\ 
  27 & GO:0050792 & regulation of viral process & 118 &  20 & 2.62 & 0.00 & 0.00 \\ 
  28 & GO:0043900 & regulation of multi-organism process & 271 &  29 & 6.01 & 0.00 & 0.00 \\ 
  29 & GO:0071346 & cellular response to interferon-gamma &  32 &  12 & 0.71 & 0.00 & 0.00 \\ 
  30 & GO:0019884 & antigen processing and presentation of e... &  25 &  11 & 0.55 & 0.00 & 0.00 \\ 
  31 & GO:0044764 & multi-organism cellular process & 201 &  25 & 4.46 & 0.00 & 0.00 \\ 
  32 & GO:0043903 & regulation of symbiosis, encompassing mu... & 137 &  21 & 3.04 & 0.00 & 0.00 \\ 
  33 & GO:0045071 & negative regulation of viral genome repl... &  33 &  12 & 0.73 & 0.00 & 0.00 \\ 
  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
  35 & GO:0044419 & interspecies interaction between organis... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
  36 & GO:0031347 & regulation of defense response & 352 &  32 & 7.81 & 0.00 & 0.00 \\ 
  37 & GO:0048002 & antigen processing and presentation of p... &  46 &  13 & 1.02 & 0.00 & 0.00 \\ 
  38 & GO:0009617 & response to bacterium & 240 &  26 & 5.32 & 0.00 & 0.00 \\ 
  39 & GO:0045089 & positive regulation of innate immune res... & 120 &  18 & 2.66 & 0.00 & 0.00 \\ 
  40 & GO:0002478 & antigen processing and presentation of e... &  20 &   9 & 0.44 & 0.00 & 0.00 \\ 
  41 & GO:0050778 & positive regulation of immune response & 285 &  27 & 6.32 & 0.00 & 0.00 \\ 
  42 & GO:1903900 & regulation of viral life cycle & 109 &  17 & 2.42 & 0.00 & 0.00 \\ 
  43 & GO:0034340 & response to type I interferon &  22 &   9 & 0.49 & 0.00 & 0.00 \\ 
  44 & GO:0042742 & defense response to bacterium & 102 &  16 & 2.26 & 0.00 & 0.00 \\ 
  45 & GO:0002684 & positive regulation of immune system pro... & 471 &  34 & 10.45 & 0.00 & 0.00 \\ 
  46 & GO:0045069 & regulation of viral genome replication &  52 &  12 & 1.15 & 0.00 & 0.00 \\ 
  47 & GO:0031349 & positive regulation of defense response & 170 &  20 & 3.77 & 0.00 & 0.00 \\ 
  48 & GO:0019058 & viral life cycle & 153 &  19 & 3.39 & 0.00 & 0.00 \\ 
  49 & GO:0002221 & pattern recognition receptor signaling p... &  77 &  14 & 1.71 & 0.00 & 0.00 \\ 
  50 & GO:0016032 & viral process & 191 &  21 & 4.24 & 0.00 & 0.00 \\ 
  51 & GO:0002758 & innate immune response-activating signal... &  80 &  14 & 1.77 & 0.00 & 0.00 \\ 
  52 & GO:0002683 & negative regulation of immune system pro... & 257 &  24 & 5.70 & 0.00 & 0.00 \\ 
  53 & GO:0042832 & defense response to protozoan &  20 &   8 & 0.44 & 0.00 & 0.00 \\ 
  54 & GO:0002218 & activation of innate immune response &  86 &  14 & 1.91 & 0.00 & 0.00 \\ 
  55 & GO:0002474 & antigen processing and presentation of p... &  28 &   9 & 0.62 & 0.00 & 0.00 \\ 
  56 & GO:0019079 & viral genome replication &  60 &  12 & 1.33 & 0.00 & 0.00 \\ 
  57 & GO:0050896 & response to stimulus & 4167 & 136 & 92.42 & 0.00 & 0.00 \\ 
  58 & GO:0002483 & antigen processing and presentation of e... &   9 &   6 & 0.20 & 0.00 & 0.00 \\ 
  59 & GO:0010033 & response to organic substance & 1314 &  61 & 29.14 & 0.00 & 0.00 \\ 
  60 & GO:0001562 & response to protozoan &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
  61 & GO:0098586 & cellular response to virus &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
  62 & GO:0019883 & antigen processing and presentation of e... &  10 &   6 & 0.22 & 0.00 & 0.00 \\ 
  63 & GO:0001816 & cytokine production & 381 &  28 & 8.45 & 0.00 & 0.00 \\ 
  64 & GO:0035455 & response to interferon-alpha &  16 &   7 & 0.35 & 0.00 & 0.00 \\ 
  65 & GO:0001817 & regulation of cytokine production & 334 &  26 & 7.41 & 0.00 & 0.00 \\ 
  66 & GO:0002237 & response to molecule of bacterial origin & 154 &  17 & 3.42 & 0.00 & 0.00 \\ 
  67 & GO:0060337 & type I interferon signaling pathway &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
  68 & GO:0071357 & cellular response to type I interferon &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
  69 & GO:0051250 & negative regulation of lymphocyte activa... &  89 &  13 & 1.97 & 0.00 & 0.00 \\ 
  70 & GO:0080134 & regulation of response to stress & 790 &  42 & 17.52 & 0.00 & 0.00 \\ 
  71 & GO:0019885 & antigen processing and presentation of e... &   7 &   5 & 0.16 & 0.00 & 0.00 \\ 
  72 & GO:0039528 & cytoplasmic pattern recognition receptor... &  13 &   6 & 0.29 & 0.00 & 0.00 \\ 
  73 & GO:0002695 & negative regulation of leukocyte activat... &  99 &  13 & 2.20 & 0.00 & 0.00 \\ 
  74 & GO:0002753 & cytoplasmic pattern recognition receptor... &  22 &   7 & 0.49 & 0.00 & 0.00 \\ 
  75 & GO:0001819 & positive regulation of cytokine producti... & 222 &  19 & 4.92 & 0.00 & 0.00 \\ 
  76 & GO:0032606 & type I interferon production &  45 &   9 & 1.00 & 0.00 & 0.00 \\ 
  77 & GO:0002764 & immune response-regulating signaling pat... & 181 &  17 & 4.01 & 0.00 & 0.00 \\ 
  78 & GO:0048584 & positive regulation of response to stimu... & 1076 &  49 & 23.86 & 0.00 & 0.00 \\ 
  79 & GO:0050866 & negative regulation of cell activation & 108 &  13 & 2.40 & 0.00 & 0.00 \\ 
  80 & GO:0002821 & positive regulation of adaptive immune r... &  61 &  10 & 1.35 & 0.00 & 0.00 \\ 
  81 & GO:0002697 & regulation of immune effector process & 258 &  20 & 5.72 & 0.00 & 0.00 \\ 
  82 & GO:0045341 & MHC class I biosynthetic process &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
  83 & GO:0045343 & regulation of MHC class I biosynthetic p... &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
  84 & GO:0002757 & immune response-activating signal transd... & 171 &  16 & 3.79 & 0.00 & 0.00 \\ 
  85 & GO:0002699 & positive regulation of immune effector p... & 115 &  13 & 2.55 & 0.00 & 0.00 \\ 
  86 & GO:0019221 & cytokine-mediated signaling pathway & 175 &  16 & 3.88 & 0.00 & 0.00 \\ 
  87 & GO:0050868 & negative regulation of T cell activation &  66 &  10 & 1.46 & 0.00 & 0.00 \\ 
  88 & GO:0071310 & cellular response to organic substance & 981 &  45 & 21.76 & 0.00 & 0.00 \\ 
  89 & GO:0044406 & adhesion of symbiont to host &  11 &   5 & 0.24 & 0.00 & 0.00 \\ 
  90 & GO:0051249 & regulation of lymphocyte activation & 246 &  19 & 5.46 & 0.00 & 0.00 \\ 
  91 & GO:0032496 & response to lipopolysaccharide & 141 &  14 & 3.13 & 0.00 & 0.00 \\ 
  92 & GO:1903038 & negative regulation of leukocyte cell-ce... &  70 &  10 & 1.55 & 0.00 & 0.00 \\ 
  93 & GO:0002824 & positive regulation of adaptive immune r... &  56 &   9 & 1.24 & 0.00 & 0.00 \\ 
  94 & GO:0039529 & RIG-I signaling pathway &  12 &   5 & 0.27 & 0.00 & 0.00 \\ 
  95 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.69 & 0.00 & 0.00 \\ 
  96 & GO:0050830 & defense response to Gram-positive bacter... &  45 &   8 & 1.00 & 0.00 & 0.00 \\ 
  97 & GO:0002819 & regulation of adaptive immune response &  93 &  11 & 2.06 & 0.00 & 0.00 \\ 
  98 & GO:0002253 & activation of immune response & 194 &  16 & 4.30 & 0.00 & 0.00 \\ 
  99 & GO:0002250 & adaptive immune response & 217 &  17 & 4.81 & 0.00 & 0.00 \\ 
  100 & GO:1900246 & positive regulation of RIG-I signaling p... &   7 &   4 & 0.16 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-03-24 additional_files/tex/cluster2_BP_Mm_7dpiUp.tex
cleaned up status but as on poster

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 05:57:46 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0002376 & immune system process & 1326 & 136 & 44.91 & 0.00 & 0.00 \\ 
  2 & GO:0006955 & immune response & 674 &  91 & 22.83 & 0.00 & 0.00 \\ 
  3 & GO:0045321 & leukocyte activation & 487 &  74 & 16.50 & 0.00 & 0.00 \\ 
  4 & GO:0001775 & cell activation & 557 &  79 & 18.87 & 0.00 & 0.00 \\ 
  5 & GO:0022610 & biological adhesion & 769 &  90 & 26.05 & 0.00 & 0.00 \\ 
  6 & GO:0044700 & single organism signaling & 2837 & 190 & 96.09 & 0.00 & 0.00 \\ 
  7 & GO:0007155 & cell adhesion & 762 &  89 & 25.81 & 0.00 & 0.00 \\ 
  8 & GO:0023052 & signaling & 2844 & 190 & 96.33 & 0.00 & 0.00 \\ 
  9 & GO:0001816 & cytokine production & 381 &  62 & 12.91 & 0.00 & 0.00 \\ 
  10 & GO:0007154 & cell communication & 2986 & 194 & 101.14 & 0.00 & 0.00 \\ 
  11 & GO:0007165 & signal transduction & 2643 & 180 & 89.52 & 0.00 & 0.00 \\ 
  12 & GO:0002684 & positive regulation of immune system pro... & 471 &  68 & 15.95 & 0.00 & 0.00 \\ 
  13 & GO:0007159 & leukocyte cell-cell adhesion & 317 &  55 & 10.74 & 0.00 & 0.00 \\ 
  14 & GO:0002682 & regulation of immune system process & 732 &  84 & 24.79 & 0.00 & 0.00 \\ 
  15 & GO:0046649 & lymphocyte activation & 412 &  62 & 13.96 & 0.00 & 0.00 \\ 
  16 & GO:0048518 & positive regulation of biological proces... & 3218 & 200 & 109.00 & 0.00 & 0.00 \\ 
  17 & GO:0050896 & response to stimulus & 4167 & 234 & 141.14 & 0.00 & 0.00 \\ 
  18 & GO:0051239 & regulation of multicellular organismal p... & 1493 & 124 & 50.57 & 0.00 & 0.00 \\ 
  19 & GO:0098602 & single organism cell adhesion & 498 &  67 & 16.87 & 0.00 & 0.00 \\ 
  20 & GO:0016337 & single organismal cell-cell adhesion & 458 &  64 & 15.51 & 0.00 & 0.00 \\ 
  21 & GO:0001817 & regulation of cytokine production & 334 &  54 & 11.31 & 0.00 & 0.00 \\ 
  22 & GO:0006954 & inflammatory response & 316 &  52 & 10.70 & 0.00 & 0.00 \\ 
  23 & GO:0034109 & homotypic cell-cell adhesion & 346 &  54 & 11.72 & 0.00 & 0.00 \\ 
  24 & GO:0098609 & cell-cell adhesion & 485 &  64 & 16.43 & 0.00 & 0.00 \\ 
  25 & GO:0006952 & defense response & 724 &  79 & 24.52 & 0.00 & 0.00 \\ 
  26 & GO:0070486 & leukocyte aggregation & 301 &  50 & 10.20 & 0.00 & 0.00 \\ 
  27 & GO:0051716 & cellular response to stimulus & 3508 & 204 & 118.82 & 0.00 & 0.00 \\ 
  28 & GO:0040011 & locomotion & 869 &  85 & 29.43 & 0.00 & 0.00 \\ 
  29 & GO:0030155 & regulation of cell adhesion & 397 &  55 & 13.45 & 0.00 & 0.00 \\ 
  30 & GO:0042110 & T cell activation & 292 &  47 & 9.89 & 0.00 & 0.00 \\ 
  31 & GO:0070489 & T cell aggregation & 292 &  47 & 9.89 & 0.00 & 0.00 \\ 
  32 & GO:0048522 & positive regulation of cellular process & 2886 & 177 & 97.75 & 0.00 & 0.00 \\ 
  33 & GO:0071593 & lymphocyte aggregation & 294 &  47 & 9.96 & 0.00 & 0.00 \\ 
  34 & GO:0048870 & cell motility & 750 &  77 & 25.40 & 0.00 & 0.00 \\ 
  35 & GO:0051674 & localization of cell & 750 &  77 & 25.40 & 0.00 & 0.00 \\ 
  36 & GO:0002694 & regulation of leukocyte activation & 288 &  46 & 9.76 & 0.00 & 0.00 \\ 
  37 & GO:0016477 & cell migration & 712 &  74 & 24.12 & 0.00 & 0.00 \\ 
  38 & GO:0050865 & regulation of cell activation & 306 &  47 & 10.36 & 0.00 & 0.00 \\ 
  39 & GO:0044707 & single-multicellular organism process & 3254 & 188 & 110.22 & 0.00 & 0.00 \\ 
  40 & GO:0048583 & regulation of response to stimulus & 2012 & 137 & 68.15 & 0.00 & 0.00 \\ 
  41 & GO:0050900 & leukocyte migration & 157 &  33 & 5.32 & 0.00 & 0.00 \\ 
  42 & GO:0002521 & leukocyte differentiation & 328 &  47 & 11.11 & 0.00 & 0.00 \\ 
  43 & GO:0032501 & multicellular organismal process & 3326 & 189 & 112.66 & 0.00 & 0.00 \\ 
  44 & GO:0065007 & biological regulation & 6019 & 280 & 203.87 & 0.00 & 0.00 \\ 
  45 & GO:0050794 & regulation of cellular process & 5444 & 262 & 184.40 & 0.00 & 0.00 \\ 
  46 & GO:0051249 & regulation of lymphocyte activation & 246 &  40 & 8.33 & 0.00 & 0.00 \\ 
  47 & GO:0002520 & immune system development & 593 &  63 & 20.09 & 0.00 & 0.00 \\ 
  48 & GO:0001819 & positive regulation of cytokine producti... & 222 &  37 & 7.52 & 0.00 & 0.00 \\ 
  49 & GO:0006928 & movement of cell or subcellular componen... & 947 &  82 & 32.08 & 0.00 & 0.00 \\ 
  50 & GO:0050789 & regulation of biological process & 5734 & 268 & 194.22 & 0.00 & 0.00 \\ 
  51 & GO:0051240 & positive regulation of multicellular org... & 856 &  77 & 28.99 & 0.00 & 0.00 \\ 
  52 & GO:0030098 & lymphocyte differentiation & 225 &  37 & 7.62 & 0.00 & 0.00 \\ 
  53 & GO:0022407 & regulation of cell-cell adhesion & 238 &  38 & 8.06 & 0.00 & 0.00 \\ 
  54 & GO:0048534 & hematopoietic or lymphoid organ developm... & 557 &  59 & 18.87 & 0.00 & 0.00 \\ 
  55 & GO:0048584 & positive regulation of response to stimu... & 1076 &  87 & 36.45 & 0.00 & 0.00 \\ 
  56 & GO:0050776 & regulation of immune response & 366 &  46 & 12.40 & 0.00 & 0.00 \\ 
  57 & GO:0007166 & cell surface receptor signaling pathway & 1252 &  95 & 42.41 & 0.00 & 0.00 \\ 
  58 & GO:0002250 & adaptive immune response & 217 &  35 & 7.35 & 0.00 & 0.00 \\ 
  59 & GO:0009605 & response to external stimulus & 1135 &  89 & 38.44 & 0.00 & 0.00 \\ 
  60 & GO:0046651 & lymphocyte proliferation & 172 &  31 & 5.83 & 0.00 & 0.00 \\ 
  61 & GO:0032943 & mononuclear cell proliferation & 174 &  31 & 5.89 & 0.00 & 0.00 \\ 
  62 & GO:0060326 & cell chemotaxis & 141 &  28 & 4.78 & 0.00 & 0.00 \\ 
  63 & GO:0045785 & positive regulation of cell adhesion & 226 &  35 & 7.65 & 0.00 & 0.00 \\ 
  64 & GO:0070887 & cellular response to chemical stimulus & 1258 &  94 & 42.61 & 0.00 & 0.00 \\ 
  65 & GO:0070661 & leukocyte proliferation & 178 &  31 & 6.03 & 0.00 & 0.00 \\ 
  66 & GO:0035556 & intracellular signal transduction & 1498 & 105 & 50.74 & 0.00 & 0.00 \\ 
  67 & GO:1903037 & regulation of leukocyte cell-cell adhesi... & 179 &  31 & 6.06 & 0.00 & 0.00 \\ 
  68 & GO:0008219 & cell death & 1318 &  96 & 44.64 & 0.00 & 0.00 \\ 
  69 & GO:0016265 & death & 1321 &  96 & 44.74 & 0.00 & 0.00 \\ 
  70 & GO:0030595 & leukocyte chemotaxis &  97 &  23 & 3.29 & 0.00 & 0.00 \\ 
  71 & GO:0050867 & positive regulation of cell activation & 184 &  31 & 6.23 & 0.00 & 0.00 \\ 
  72 & GO:0040017 & positive regulation of locomotion & 262 &  37 & 8.87 & 0.00 & 0.00 \\ 
  73 & GO:0097530 & granulocyte migration &  63 &  19 & 2.13 & 0.00 & 0.00 \\ 
  74 & GO:0022409 & positive regulation of cell-cell adhesio... & 129 &  26 & 4.37 & 0.00 & 0.00 \\ 
  75 & GO:0097529 & myeloid leukocyte migration &  90 &  22 & 3.05 & 0.00 & 0.00 \\ 
  76 & GO:0006935 & chemotaxis & 266 &  37 & 9.01 & 0.00 & 0.00 \\ 
  77 & GO:0030097 & hemopoiesis & 528 &  54 & 17.88 & 0.00 & 0.00 \\ 
  78 & GO:0042330 & taxis & 267 &  37 & 9.04 & 0.00 & 0.00 \\ 
  79 & GO:0002696 & positive regulation of leukocyte activat... & 179 &  30 & 6.06 & 0.00 & 0.00 \\ 
  80 & GO:0030335 & positive regulation of cell migration & 245 &  35 & 8.30 & 0.00 & 0.00 \\ 
  81 & GO:0010646 & regulation of cell communication & 1709 & 112 & 57.89 & 0.00 & 0.00 \\ 
  82 & GO:0050793 & regulation of developmental process & 1301 &  93 & 44.07 & 0.00 & 0.00 \\ 
  83 & GO:2000147 & positive regulation of cell motility & 250 &  35 & 8.47 & 0.00 & 0.00 \\ 
  84 & GO:0023051 & regulation of signaling & 1678 & 110 & 56.84 & 0.00 & 0.00 \\ 
  85 & GO:0002460 & adaptive immune response based on somati... & 150 &  27 & 5.08 & 0.00 & 0.00 \\ 
  86 & GO:0001818 & negative regulation of cytokine producti... & 128 &  25 & 4.34 & 0.00 & 0.00 \\ 
  87 & GO:0071621 & granulocyte chemotaxis &  54 &  17 & 1.83 & 0.00 & 0.00 \\ 
  88 & GO:0034110 & regulation of homotypic cell-cell adhesi... & 189 &  30 & 6.40 & 0.00 & 0.00 \\ 
  89 & GO:1990266 & neutrophil migration &  55 &  17 & 1.86 & 0.00 & 0.00 \\ 
  90 & GO:0051272 & positive regulation of cellular componen... & 257 &  35 & 8.70 & 0.00 & 0.00 \\ 
  91 & GO:1903039 & positive regulation of leukocyte cell-ce... & 110 &  23 & 3.73 & 0.00 & 0.00 \\ 
  92 & GO:0009966 & regulation of signal transduction & 1495 & 101 & 50.64 & 0.00 & 0.00 \\ 
  93 & GO:0042221 & response to chemical & 1780 & 113 & 60.29 & 0.00 & 0.00 \\ 
  94 & GO:0002443 & leukocyte mediated immunity & 196 &  30 & 6.64 & 0.00 & 0.00 \\ 
  95 & GO:0050863 & regulation of T cell activation & 172 &  28 & 5.83 & 0.00 & 0.00 \\ 
  96 & GO:0051270 & regulation of cellular component movemen... & 467 &  48 & 15.82 & 0.00 & 0.00 \\ 
  97 & GO:0044763 & single-organism cellular process & 6624 & 285 & 224.37 & 0.00 & 0.00 \\ 
  98 & GO:0006915 & apoptotic process & 1219 &  87 & 41.29 & 0.00 & 0.00 \\ 
  99 & GO:0012501 & programmed cell death & 1241 &  88 & 42.03 & 0.00 & 0.00 \\ 
  100 & GO:0002252 & immune effector process & 440 &  46 & 14.90 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-03-24 additional_files/tex/cluster3_BP_Mm_7dpiDown.tex
cleaned up status but as on poster

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 05:59:04 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0055114 & oxidation-reduction process & 597 &  30 & 8.37 & 0.00 & 0.00 \\ 
  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  21 & 4.90 & 0.00 & 0.00 \\ 
  3 & GO:0019752 & carboxylic acid metabolic process & 578 &  27 & 8.10 & 0.00 & 0.00 \\ 
  4 & GO:0006631 & fatty acid metabolic process & 234 &  17 & 3.28 & 0.00 & 0.00 \\ 
  5 & GO:0043436 & oxoacid metabolic process & 617 &  27 & 8.65 & 0.00 & 0.00 \\ 
  6 & GO:0006082 & organic acid metabolic process & 625 &  27 & 8.76 & 0.00 & 0.00 \\ 
  7 & GO:0044281 & small molecule metabolic process & 1187 &  39 & 16.63 & 0.00 & 0.00 \\ 
  8 & GO:0006629 & lipid metabolic process & 767 &  29 & 10.75 & 0.00 & 0.00 \\ 
  9 & GO:0044710 & single-organism metabolic process & 3037 &  69 & 42.56 & 0.00 & 0.00 \\ 
  10 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 0.66 & 0.00 & 0.00 \\ 
  11 & GO:0006690 & icosanoid metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
  12 & GO:1901568 & fatty acid derivative metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
  13 & GO:0006805 & xenobiotic metabolic process &  38 &   6 & 0.53 & 0.00 & 0.00 \\ 
  14 & GO:0019369 & arachidonic acid metabolic process &  25 &   5 & 0.35 & 0.00 & 0.00 \\ 
  15 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &   6 & 0.60 & 0.00 & 0.00 \\ 
  16 & GO:0019395 & fatty acid oxidation &  65 &   7 & 0.91 & 0.00 & 0.00 \\ 
  17 & GO:0034440 & lipid oxidation &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
  18 & GO:0072329 & monocarboxylic acid catabolic process &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
  19 & GO:0042738 & exogenous drug catabolic process &  15 &   4 & 0.21 & 0.00 & 0.00 \\ 
  20 & GO:0009410 & response to xenobiotic stimulus &  47 &   6 & 0.66 & 0.00 & 0.00 \\ 
  21 & GO:0019373 & epoxygenase P450 pathway &   6 &   3 & 0.08 & 0.00 & 0.00 \\ 
  22 & GO:0044255 & cellular lipid metabolic process & 579 &  21 & 8.11 & 0.00 & 0.00 \\ 
  23 & GO:0042737 & drug catabolic process &  16 &   4 & 0.22 & 0.00 & 0.00 \\ 
  24 & GO:0042493 & response to drug & 157 &  10 & 2.20 & 0.00 & 0.00 \\ 
  25 & GO:0032099 & negative regulation of appetite &   8 &   3 & 0.11 & 0.00 & 0.00 \\ 
  26 & GO:0009062 & fatty acid catabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
  27 & GO:0033559 & unsaturated fatty acid metabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
  28 & GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
  29 & GO:0032096 & negative regulation of response to food &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
  30 & GO:0060979 & vasculogenesis involved in coronary vasc... &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
  31 & GO:0016054 & organic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
  32 & GO:0046395 & carboxylic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
  33 & GO:0017144 & drug metabolic process &  24 &   4 & 0.34 & 0.00 & 0.00 \\ 
  34 & GO:0006635 & fatty acid beta-oxidation &  45 &   5 & 0.63 & 0.00 & 0.00 \\ 
  35 & GO:0030258 & lipid modification & 127 &   8 & 1.78 & 0.00 & 0.00 \\ 
  36 & GO:0032095 & regulation of response to food &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
  37 & GO:0032098 & regulation of appetite &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
  38 & GO:2001023 & regulation of response to drug &   3 &   2 & 0.04 & 0.00 & 0.00 \\ 
  39 & GO:1901615 & organic hydroxy compound metabolic proce... & 287 &  12 & 4.02 & 0.00 & 0.00 \\ 
  40 & GO:0015701 & bicarbonate transport &  15 &   3 & 0.21 & 0.00 & 0.00 \\ 
  41 & GO:0001561 & fatty acid alpha-oxidation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  42 & GO:0030916 & otic vesicle formation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  43 & GO:0071600 & otic vesicle morphogenesis &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
  44 & GO:0032094 & response to food &  16 &   3 & 0.22 & 0.00 & 0.00 \\ 
  45 & GO:0048566 & embryonic digestive tract development &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
  46 & GO:0060977 & coronary vasculature morphogenesis &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
  47 & GO:0042760 & very long-chain fatty acid catabolic pro... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  48 & GO:0044539 & long-chain fatty acid import &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  49 & GO:0046135 & pyrimidine nucleoside catabolic process &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  50 & GO:0072366 & regulation of cellular ketone metabolic ... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
  51 & GO:0097164 & ammonium ion metabolic process &  94 &   6 & 1.32 & 0.00 & 0.00 \\ 
  52 & GO:0015711 & organic anion transport & 202 &   9 & 2.83 & 0.00 & 0.00 \\ 
  53 & GO:0001822 & kidney development & 132 &   7 & 1.85 & 0.00 & 0.00 \\ 
  54 & GO:0050892 & intestinal absorption &  20 &   3 & 0.28 & 0.00 & 0.00 \\ 
  55 & GO:0016042 & lipid catabolic process & 171 &   8 & 2.40 & 0.00 & 0.00 \\ 
  56 & GO:0019532 & oxalate transport &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  57 & GO:0060484 & lung-associated mesenchyme development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  58 & GO:0071599 & otic vesicle development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
  59 & GO:0043648 & dicarboxylic acid metabolic process &  70 &   5 & 0.98 & 0.00 & 0.01 \\ 
  60 & GO:0015718 & monocarboxylic acid transport &  71 &   5 & 0.99 & 0.00 & 0.01 \\ 
  61 & GO:0044282 & small molecule catabolic process & 178 &   8 & 2.49 & 0.00 & 0.01 \\ 
  62 & GO:0048754 & branching morphogenesis of an epithelial... & 106 &   6 & 1.49 & 0.00 & 0.01 \\ 
  63 & GO:0010817 & regulation of hormone levels & 262 &  10 & 3.67 & 0.00 & 0.01 \\ 
  64 & GO:0048002 & antigen processing and presentation of p... &  46 &   4 & 0.64 & 0.00 & 0.01 \\ 
  65 & GO:0032105 & negative regulation of response to extra... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
  66 & GO:0032108 & negative regulation of response to nutri... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
  67 & GO:0006560 & proline metabolic process &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  68 & GO:0090331 & negative regulation of platelet aggregat... &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  69 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
  70 & GO:0072001 & renal system development & 143 &   7 & 2.00 & 0.00 & 0.01 \\ 
  71 & GO:0018958 & phenol-containing compound metabolic pro... &  48 &   4 & 0.67 & 0.00 & 0.01 \\ 
  72 & GO:0015849 & organic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
  73 & GO:0046942 & carboxylic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
  74 & GO:0070374 & positive regulation of ERK1 and ERK2 cas... &  78 &   5 & 1.09 & 0.00 & 0.01 \\ 
  75 & GO:0006584 & catecholamine metabolic process &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  76 & GO:0007431 & salivary gland development &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  77 & GO:0009712 & catechol-containing compound metabolic p... &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
  78 & GO:0044242 & cellular lipid catabolic process & 114 &   6 & 1.60 & 0.01 & 0.01 \\ 
  79 & GO:0009948 & anterior/posterior axis specification &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  80 & GO:0060603 & mammary gland duct morphogenesis &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  81 & GO:0060976 & coronary vasculature development &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
  82 & GO:0030004 & cellular monovalent inorganic cation hom... &  51 &   4 & 0.71 & 0.01 & 0.01 \\ 
  83 & GO:0006790 & sulfur compound metabolic process & 194 &   8 & 2.72 & 0.01 & 0.01 \\ 
  84 & GO:0008272 & sulfate transport &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  85 & GO:0010544 & negative regulation of platelet activati... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  86 & GO:0072364 & regulation of cellular ketone metabolic ... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
  87 & GO:0002474 & antigen processing and presentation of p... &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
  88 & GO:0051196 & regulation of coenzyme metabolic process &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
  89 & GO:0006820 & anion transport & 288 &  10 & 4.04 & 0.01 & 0.01 \\ 
  90 & GO:0015909 & long-chain fatty acid transport &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  91 & GO:0045844 & positive regulation of striated muscle t... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  92 & GO:0048636 & positive regulation of muscle organ deve... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  93 & GO:0051193 & regulation of cofactor metabolic process &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
  94 & GO:0009056 & catabolic process & 1369 &  30 & 19.18 & 0.01 & 0.01 \\ 
  95 & GO:0051923 & sulfation &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  96 & GO:0060081 & membrane hyperpolarization &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  97 & GO:0072529 & pyrimidine-containing compound catabolic... &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
  98 & GO:0009880 & embryonic pattern specification &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
  99 & GO:1901863 & positive regulation of muscle tissue dev... &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
  100 & GO:0042060 & wound healing & 210 &   8 & 2.94 & 0.01 & 0.01 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-03-24 additional_files/tex/cluster4_BP_Mm_allInfDown.tex
cleaned up status but as on poster

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Mar 24 06:01:34 2016
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrrr}
  \hline
 & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
1 & GO:0044281 & small molecule metabolic process & 1187 &  94 & 44.61 & 0.00 & 0.00 \\ 
  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  40 & 13.15 & 0.00 & 0.00 \\ 
  3 & GO:0006629 & lipid metabolic process & 767 &  63 & 28.83 & 0.00 & 0.00 \\ 
  4 & GO:0006082 & organic acid metabolic process & 625 &  55 & 23.49 & 0.00 & 0.00 \\ 
  5 & GO:0043436 & oxoacid metabolic process & 617 &  53 & 23.19 & 0.00 & 0.00 \\ 
  6 & GO:0019752 & carboxylic acid metabolic process & 578 &  50 & 21.72 & 0.00 & 0.00 \\ 
  7 & GO:0042178 & xenobiotic catabolic process &   9 &   6 & 0.34 & 0.00 & 0.00 \\ 
  8 & GO:0006805 & xenobiotic metabolic process &  38 &  10 & 1.43 & 0.00 & 0.00 \\ 
  9 & GO:0044710 & single-organism metabolic process & 3037 & 158 & 114.14 & 0.00 & 0.00 \\ 
  10 & GO:0044255 & cellular lipid metabolic process & 579 &  46 & 21.76 & 0.00 & 0.00 \\ 
  11 & GO:0006631 & fatty acid metabolic process & 234 &  25 & 8.79 & 0.00 & 0.00 \\ 
  12 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &  10 & 1.62 & 0.00 & 0.00 \\ 
  13 & GO:0006790 & sulfur compound metabolic process & 194 &  22 & 7.29 & 0.00 & 0.00 \\ 
  14 & GO:0034754 & cellular hormone metabolic process &  45 &  10 & 1.69 & 0.00 & 0.00 \\ 
  15 & GO:0042445 & hormone metabolic process &  89 &  14 & 3.34 & 0.00 & 0.00 \\ 
  16 & GO:0009410 & response to xenobiotic stimulus &  47 &  10 & 1.77 & 0.00 & 0.00 \\ 
  17 & GO:0009719 & response to endogenous stimulus & 701 &  50 & 26.35 & 0.00 & 0.00 \\ 
  18 & GO:0042537 & benzene-containing compound metabolic pr... &  16 &   6 & 0.60 & 0.00 & 0.00 \\ 
  19 & GO:0016042 & lipid catabolic process & 171 &  19 & 6.43 & 0.00 & 0.00 \\ 
  20 & GO:0009813 & flavonoid biosynthetic process &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
  21 & GO:0052696 & flavonoid glucuronidation &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
  22 & GO:0055114 & oxidation-reduction process & 597 &  43 & 22.44 & 0.00 & 0.00 \\ 
  23 & GO:0071495 & cellular response to endogenous stimulus & 540 &  40 & 20.30 & 0.00 & 0.00 \\ 
  24 & GO:0072329 & monocarboxylic acid catabolic process &  66 &  11 & 2.48 & 0.00 & 0.00 \\ 
  25 & GO:0019626 & short-chain fatty acid catabolic process &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  26 & GO:0051410 & detoxification of nitrogen compound &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  27 & GO:0070458 & cellular detoxification of nitrogen comp... &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
  28 & GO:0006063 & uronic acid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  29 & GO:0009812 & flavonoid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  30 & GO:0019585 & glucuronate metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  31 & GO:0052695 & cellular glucuronidation &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
  32 & GO:0044242 & cellular lipid catabolic process & 114 &  14 & 4.28 & 0.00 & 0.00 \\ 
  33 & GO:0009725 & response to hormone & 365 &  29 & 13.72 & 0.00 & 0.00 \\ 
  34 & GO:0008202 & steroid metabolic process & 162 &  17 & 6.09 & 0.00 & 0.00 \\ 
  35 & GO:0044282 & small molecule catabolic process & 178 &  18 & 6.69 & 0.00 & 0.00 \\ 
  36 & GO:0006749 & glutathione metabolic process &  42 &   8 & 1.58 & 0.00 & 0.00 \\ 
  37 & GO:0009636 & response to toxic substance &  78 &  11 & 2.93 & 0.00 & 0.00 \\ 
  38 & GO:0085029 & extracellular matrix assembly &  15 &   5 & 0.56 & 0.00 & 0.00 \\ 
  39 & GO:0018916 & nitrobenzene metabolic process &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
  40 & GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
  41 & GO:0046459 & short-chain fatty acid metabolic process &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
  42 & GO:1901201 & regulation of extracellular matrix assem... &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
  43 & GO:0006164 & purine nucleotide biosynthetic process & 124 &  14 & 4.66 & 0.00 & 0.00 \\ 
  44 & GO:0009062 & fatty acid catabolic process &  59 &   9 & 2.22 & 0.00 & 0.00 \\ 
  45 & GO:1903055 & positive regulation of extracellular mat... &  10 &   4 & 0.38 & 0.00 & 0.00 \\ 
  46 & GO:0010817 & regulation of hormone levels & 262 &  22 & 9.85 & 0.00 & 0.00 \\ 
  47 & GO:0072522 & purine-containing compound biosynthetic ... & 130 &  14 & 4.89 & 0.00 & 0.00 \\ 
  48 & GO:0044699 & single-organism process & 7377 & 308 & 277.25 & 0.00 & 0.00 \\ 
  49 & GO:0035637 & multicellular organismal signaling &  38 &   7 & 1.43 & 0.00 & 0.00 \\ 
  50 & GO:0009165 & nucleotide biosynthetic process & 165 &  16 & 6.20 & 0.00 & 0.00 \\ 
  51 & GO:0042572 & retinol metabolic process &  11 &   4 & 0.41 & 0.00 & 0.00 \\ 
  52 & GO:0072521 & purine-containing compound metabolic pro... & 289 &  23 & 10.86 & 0.00 & 0.00 \\ 
  53 & GO:1900542 & regulation of purine nucleotide metaboli... &  92 &  11 & 3.46 & 0.00 & 0.00 \\ 
  54 & GO:0016054 & organic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
  55 & GO:0046395 & carboxylic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
  56 & GO:1901293 & nucleoside phosphate biosynthetic proces... & 170 &  16 & 6.39 & 0.00 & 0.00 \\ 
  57 & GO:0019395 & fatty acid oxidation &  65 &   9 & 2.44 & 0.00 & 0.00 \\ 
  58 & GO:0006081 & cellular aldehyde metabolic process &  41 &   7 & 1.54 & 0.00 & 0.00 \\ 
  59 & GO:0034440 & lipid oxidation &  66 &   9 & 2.48 & 0.00 & 0.00 \\ 
  60 & GO:0006000 & fructose metabolic process &   6 &   3 & 0.23 & 0.00 & 0.00 \\ 
  61 & GO:0006140 & regulation of nucleotide metabolic proce... &  97 &  11 & 3.65 & 0.00 & 0.00 \\ 
  62 & GO:0060688 & regulation of morphogenesis of a branchi... &  32 &   6 & 1.20 & 0.00 & 0.00 \\ 
  63 & GO:0051186 & cofactor metabolic process & 249 &  20 & 9.36 & 0.00 & 0.00 \\ 
  64 & GO:0007267 & cell-cell signaling & 440 &  30 & 16.54 & 0.00 & 0.00 \\ 
  65 & GO:0009117 & nucleotide metabolic process & 362 &  26 & 13.61 & 0.00 & 0.00 \\ 
  66 & GO:0034220 & ion transmembrane transport & 383 &  27 & 14.39 & 0.00 & 0.00 \\ 
  67 & GO:0006635 & fatty acid beta-oxidation &  45 &   7 & 1.69 & 0.00 & 0.00 \\ 
  68 & GO:0006811 & ion transport & 756 &  45 & 28.41 & 0.00 & 0.00 \\ 
  69 & GO:0055085 & transmembrane transport & 546 &  35 & 20.52 & 0.00 & 0.00 \\ 
  70 & GO:0006001 & fructose catabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  71 & GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  72 & GO:0034587 & piRNA metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  73 & GO:0061624 & fructose catabolic process to hydroxyace... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  74 & GO:0061625 & glycolytic process through fructose-1-ph... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
  75 & GO:1901135 & carbohydrate derivative metabolic proces... & 631 &  39 & 23.72 & 0.00 & 0.00 \\ 
  76 & GO:0009152 & purine ribonucleotide biosynthetic proce... & 117 &  12 & 4.40 & 0.00 & 0.00 \\ 
  77 & GO:0006753 & nucleoside phosphate metabolic process & 370 &  26 & 13.91 & 0.00 & 0.00 \\ 
  78 & GO:0003071 & renal system process involved in regulat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  79 & GO:0019755 & one-carbon compound transport &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  80 & GO:0051386 & regulation of neurotrophin TRK receptor ... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  81 & GO:1901203 & positive regulation of extracellular mat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
  82 & GO:0006163 & purine nucleotide metabolic process & 257 &  20 & 9.66 & 0.00 & 0.00 \\ 
  83 & GO:0044711 & single-organism biosynthetic process & 851 &  49 & 31.98 & 0.00 & 0.00 \\ 
  84 & GO:0006040 & amino sugar metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
  85 & GO:0017144 & drug metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
  86 & GO:0009118 & regulation of nucleoside metabolic proce... &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
  87 & GO:1903578 & regulation of ATP metabolic process &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
  88 & GO:0016266 & O-glycan processing &  15 &   4 & 0.56 & 0.00 & 0.00 \\ 
  89 & GO:0001523 & retinoid metabolic process &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
  90 & GO:0061098 & positive regulation of protein tyrosine ... &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
  91 & GO:0009260 & ribonucleotide biosynthetic process & 122 &  12 & 4.59 & 0.00 & 0.00 \\ 
  92 & GO:1901564 & organonitrogen compound metabolic proces... & 1198 &  64 & 45.03 & 0.00 & 0.00 \\ 
  93 & GO:0008285 & negative regulation of cell proliferatio... & 400 &  27 & 15.03 & 0.00 & 0.00 \\ 
  94 & GO:0035690 & cellular response to drug &  37 &   6 & 1.39 & 0.00 & 0.00 \\ 
  95 & GO:0009150 & purine ribonucleotide metabolic process & 246 &  19 & 9.25 & 0.00 & 0.00 \\ 
  96 & GO:0009755 & hormone-mediated signaling pathway &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
  97 & GO:0043467 & regulation of generation of precursor me... &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
  98 & GO:0046390 & ribose phosphate biosynthetic process & 124 &  12 & 4.66 & 0.00 & 0.00 \\ 
  99 & GO:1900371 & regulation of purine nucleotide biosynth... &  50 &   7 & 1.88 & 0.00 & 0.00 \\ 
  100 & GO:1990748 & cellular detoxification &  16 &   4 & 0.60 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-03-24 1_ballgown_import.R

@@ -5,7 +5,7 @@ setwd("~/Ef_RNAseq")
 
 ## IMPORT DATA USING BALLGOWN IMPORT
 if(!exists("All.bg")) {
-    All.bg = ballgown(dataDir="/data/Eimeria_Totta/tablemaker_060815/",
+    All.bg = ballgown(dataDir="/SAN/Eimeria_Totta/backup/tablemaker_060815.0/",
         samplePattern = "*")
 }
 
@@ -45,7 +45,7 @@ pData(All.bg) <- add.pdata(All.bg)
 ########################################
 ## Create mouse only object and remove sporozoite and oocyst samples
 ## from mouse data
-Mm.bg <- subset(All.bg, grepl('^XLOC.*', geneIDs(rnaseq.bg)),
+Mm.bg <- subset(All.bg, grepl('^XLOC.*', geneIDs(All.bg)),
                 genomesubset=TRUE)
 Mm.bg <- subset(Mm.bg, !is.na(pData(Mm.bg)$timepoint),
                 genomesubset=FALSE) # 

### 2016-03-24 2_edgeR_diff.R

@@ -8,6 +8,8 @@ if(!exists("Mm.RC")|!exists("Ef.RC")){
 
 library(edgeR)
 library(GGally)
+library(RUVSeq)
+library(reshape)
 
 ## Implement the sample exclusion (see A_data_curation.R)
 Mm.bg <- subset(Mm.bg,
@@ -27,6 +29,48 @@ Ef.RC <- raw.counts.4.bg(Ef.bg)
 ## use the alternate design like recommended by the EdgR manual (or
 ## eg. here: https://support.bioconductor.org/p/66952/) and specifiy
 ## interactions as contrasts!
+Ef.design <- model.matrix(~0+pData(Ef.bg)$grouped)
+colnames(Ef.design)  <- gsub("pData\\(Ef.bg\\)\\$grouped", "",
+                             colnames(Ef.design))
+## just to have the standard point delimitor
+colnames(Ef.design)  <- gsub("_", ".",
+                             colnames(Ef.design))
+
+Ef.contrasts <- makeContrasts(
+    ## just different stages in NMRI
+    N3vsN5 = NMRI.1stInf.3dpi-NMRI.1stInf.5dpi,
+    N3vsN7 = NMRI.1stInf.3dpi-NMRI.1stInf.7dpi,
+    N3vsSpo = NMRI.1stInf.3dpi-NMRI.sporozoites,
+    N3vsOoc = NMRI.1stInf.3dpi-NMRI.oocysts,
+    N5vsN7 = NMRI.1stInf.5dpi-NMRI.1stInf.7dpi,
+    N5vsSp = NMRI.1stInf.5dpi-NMRI.sporozoites,
+    N5vsOo = NMRI.1stInf.5dpi-NMRI.oocysts,
+    N7vsSp = NMRI.1stInf.7dpi-NMRI.sporozoites,
+    N7vsOo = NMRI.1stInf.7dpi-NMRI.oocysts,
+    SpvsOo=  NMRI.sporozoites-NMRI.oocysts,
+    ## differences in 1st vs 2nd infection
+    N3.1stvsN3.2nd = NMRI.1stInf.3dpi-NMRI.2ndInf.3dpi,
+    N5.1stvsN5.2nd = NMRI.1stInf.5dpi-NMRI.2ndInf.5dpi,
[…]

### 2016-03-24 3_annotations.R

@@ -2,14 +2,14 @@
 ## a TopTags (edgeR) object listing genes tested for differential
 ## expression
 
-if(!exists("ALL.top")){
+if(!exists("Mm.1st.pass.model")){
     source("2_edgeR_diff.R")
 }
 
 ## The Bioconductor mouse annotation libary
 library(Mus.musculus)
 
-get.cuff2name <- function(path="data/merged_mm9_Ef.gtf"){
+.get.cuff2name <- function(path="data/merged_mm9_Ef.gtf"){
     gene.gtf.info <- gffRead(path) 
     gff.attr.list <- strsplit(gene.gtf.info$attributes, " ")
     cuff_ids <- unlist(lapply(gff.attr.list, function (x) x[[4]]))
@@ -22,7 +22,7 @@ get.cuff2name <- function(path="data/merged_mm9_Ef.gtf"){
     return(unique(cuff2name))
 }
 
-cuff2name <- get.cuff2name()
+cuff2name <- .get.cuff2name()
 
 ## merge all this information in a data frame to combine mouse
 ## annotations with gene_id from Cufflinks get the information from
@@ -40,28 +40,30 @@ cuff2name <- get.cuff2name()
 annot.frame <- .get.annot.frame()
 
 ## cuff2GO contains the mapping of cufflink ids to GO terms
-cuff2GO.Mm <- by(annot.frame, annot.frame$cuff_ids,
-                 function(x) as.character(x$go_id))
+gene2GO <- list()
+
+gene2GO[["Mm"]] <- by(annot.frame, annot.frame$cuff_ids,
+                      function(x) as.character(x$go_id))
 ## define the universe of tested genes
-cuff2GO.Mm <- cuff2GO.Mm[names(cuff2GO.Mm)%in%rownames(ALL.top.Mm)]
+gene2GO[["Mm"]] <- gene2GO[["Mm"]][names(gene2GO[["Mm"]])%in%
+                                       rownames(Mm.1st.pass.model[[1]])]
 
 ## overwrite the cuff2name creator function to turn it into a subsetter function
-get.cuff2name <- function(cuff_ids){
-    cuffsubset <- cuff2name[cuff2name$cuff_ids%in%cuff_ids , "gene_names"]
+
+get.id4cuff <- function(cuff_ids){
+    cuffsubset <- annot.frame[annot.frame$cuff_ids%in%cuff_ids , "gene_id"]
[…]

### 2016-03-24 4_compareArray.R

@@ -8,12 +8,13 @@ library(Agi4x44PreProcess)
 library(Biobase)
 library(RSvgDevice)
 library(GGally)
+library(ggplot2)
 
-if(!exists("ALL.top.Mm")){
+if(!exists("Mm.1st.pass.model")){
     source("2_edgeR_diff.R")
 }
 
-if(!exists("cuff2GO.Mm")){
+if(!exists("gene2GO")){
     source("3_annotations.R")
 }
 
@@ -87,17 +88,24 @@ fit2.g <- eBayes(fit2.g)
 
 Array.logFC <- topTable(fit2.g, n=400000)[, c("X24h", "X144h")]
 
-RNAseq.logFC <- ALL.top.Mm[, c("logFC.N3vs0", "logFC.N5vs0",
-                            "logFC.N7vs0", "logFC.B5vs0", "logFC.R5vs0" )]
-RNAseq.logFC <- merge(RNAseq.logFC, id2name, by.x = 0, by.y = "gene_ids")
+RNAseq.logFC <- Mm.1st.pass.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
+                                           "logFC.N7vs0", "logFC.B5vs0",
+                                           "logFC.R5vs0" )]
 
-RNAseq.logFC.ruved <- Mm.ALL.top.ruved[, c("logFC.N3vs0", "logFC.N5vs0",
-                                        "logFC.N7vs0", "logFC.B5vs0", "logFC.R5vs0" )]
-RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, id2name, by.x = 0, by.y = "gene_ids")
+RNAseq.logFC <- merge(RNAseq.logFC, cuff2name, by.x = 0, by.y = "cuff_ids")
 
-RNAseq.Array.logFC <- merge(RNAseq.logFC, Array.logFC, by.x = "gene_names", by.y = 0)
+RNAseq.logFC.ruved <- Mm.RUVg.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
+                                             "logFC.N7vs0", "logFC.B5vs0",
+                                             "logFC.R5vs0" )]
 
-RNAseq.Array.logFC <- merge(RNAseq.Array.logFC, RNAseq.logFC.ruved, by = c("Row.names","gene_names"))
+RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, cuff2name,
+                            by.x = 0, by.y = "cuff_ids")
+
[…]

### 2016-03-24 5_enrichment.R

@@ -1,104 +1,50 @@
 ## still open question: How to implement the FCS(?) method
-
-
 library(topGO)
+library(xtable)
 
-if(!exists("chalGids")){
-    source("1_ballgown_import_diff.R")
+if(!exists("hcluster")){
+    source("clustering.R")
 }
 
-if(!exists("gene.list")){
-    source("2_edgeR_diff.R")
-}
-
-if(!exists("gene2GO.Mm")){
+if(!exists("gene2GO")){
     source("3_annotations.R")
 }
 
 ## the universe of genes which were tested at all:
-exp.universe <- names(gene2GO.Mm)
-
-## The CHALLENGE in NMRI DAY 7 
-MF.N7.chal <- TOGO.all.onto("MF", exp.universe,
-                            gene.list[["N7.1stvsN7.2nd"]], gene2GO.Mm)
-
-gene.table.topGO(MF.N7.chal)
-
-BP.N7.chal <- TOGO.all.onto("BP", exp.universe, 
-                            gene.list.ruved[["N7.1stvsN7.2nd"]], gene2GO.Mm)
-
-gene.table.topGO(BP.N7.chal)
-
-topKEGG(kegga(gene.list.ruved[["N7.1stvsN7.2nd"]], species = "Mm"), n=30)
-topGO(goana(N7.1stvsN7.2nd.GID, species = "Mm"), n=30)
+exp.universe <- list()
+exp.universe[["Mm"]] <- names(gene2GO[["Mm"]])
+exp.universe[["Ef"]] <- names(gene2GO[["Ef"]])
 
 
-## the genes not detected as DE  on the array but in RNA seq:
-Array.universe <-
-    get.annotation.for.xloc(RNAseq.Array.logFC$Row.names)[[2]]
+#### We search for enrichment of genes from clustering 
+######## Clustering #############
 
-A.R.fail <- get.annotation.for.xloc(A.low.R.high)[[2]]
+to.test <- as.data.frame(rbind(
+    ## Eimeria
+    cbind(cluster=2,type="oocysts", species="Ef"),
+    cbind(cluster=5,type="sporozoites", species="Ef"),
+    cbind(cluster=1,type="day7", species="Ef"),
+    ## Mouse
+    cbind(cluster=2,type="7dpiUp", species="Mm"),
+    cbind(cluster=3,type="7dpiDown", species="Mm"),
+    cbind(cluster=1,type="EarlyUp", species="Mm"),
+    cbind(cluster=4,type="allInfDown", species="Mm")
+    ))
 
[…]

### 2016-03-24 A_data_curataion.R

@@ -1,0 +1,161 @@
+## This is a script which is not part of the pipeline but needed for
+## evaluation purpose -> only "insight" gained from this is transfered
+## to the numberd scripts which are part of the processing pipeline
+
+## Goals:
+
+## EX- on IN-clusion of samples and genes ###################
+
+## a) decide for samples needed to be excluede because of uncertain
+## infection status
+
+## b) decice for a filter on minimal expression values over all
+## samples to exclude genes
+
+## c) decide on samples needed to be excluded because of bad
+## "unnormalizable" distribution of expression values
+
+
+### Normalization
+
+## d) decide on a general normalization strategy scaling the smaple
+## throughput
+
+## e) decide on a "removeal of unwanted variation" strategy
+
+## f) decide on a potential addional "batch effect" removal strategy
+
+
+## a) decide for samples needed to be excluede because of uncertain
+## infection status
+
+##object required from upstream scripts:
+
+## The raw counts:
+if(!exists("All.RC")|!exists("Mm.RC")|!exists("Ef.RC")){
+    source("1_ballgown_import.R")
+}
+
+library(ggplot2)
+library(reshape)
+
+## Overview for differential representation of host/parasite
+RC.table <-  as.data.frame(cbind(sample=as.character(pData(All.bg)$samples),
+                                 seq.method=as.character(pData(All.bg)$seq.method),
+                                 batch=pData(All.bg)$batch))
+
+RC.table$c.total.reads <- colSums(All.RC[[3]])
+
+RC.table$c.Mm.reads <- colSums(All.RC[[3]]
+                                [grepl('^XLOC.*', rownames(All.RC[[3]])),])
+
+RC.table$c.Ef.reads <- colSums(All.RC[[3]]
+                                [grepl('^EfaB.*', rownames(All.RC[[3]])),])
+
+RC.table$p.Ef.reads <- round(RC.table$c.Ef.reads/RC.table$c.total.reads*100, 4)
+
+## plotting parasite percentage of reads
+ef.percentage <- ggplot(RC.table, aes(x = sample, y = p.Ef.reads,
+                                      color=seq.method)) +
+    geom_point()+
+        scale_y_log10()+
+            ylab(label="Percentage of reads mapping to Eimeria genome") +
+                xlab(label="Sample") +
[…]

### 2016-03-24 clustering.R

@@ -1,63 +1,80 @@
 ## Making heatmaps and other clusters of gene expression data
 
+library(pheatmap)
 library(made4)
 library(RColorBrewer)
 
 
 ########## USE THIS #######################
 ########## PHEATMAP() specifying distance and method ############
-pheatmap(cpm(Ef.RC[[3]][unlist(gene.list.E),]),
-  color = brewer.pal(n = 11, name = "BrBG"),
-  scale = "row",
- # kmeans_k = NA,
- # cluster_rows = T,
- # cluster_cols = T,
- # clustering_distance_rows = "euclidean", --> has no influence on clustering
- # clustering_method = "complete",         --> has no influence on clustering
- #annotation_names_row = F,
-  annotation_row = NA,
+
+union.of.Ef.cycle.diff.top.100 <-
+    unique(unlist(lapply(Ef.1st.pass.model[[3]][1:10], head, n=500)))
+
+Ef.cycle.diff.top.100.data <-
+    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff.top.100,]
+
+get.scaled.and.clustered <- function(data){
+    rMeans <- rowMeans(data)
+    rSD <- apply(data, 1, sd)
+    scaled.data <- scale(t(data), rMeans, rSD)
+    hclustered <- hclust(dist(t(scaled.data)))
+    return(hclustered)
+}
+
+Ef.hclustered <- get.scaled.and.clustered(Ef.cycle.diff.top.100.data)
+
+hcluster <- list()
+hcluster[["Ef"]] <- cutree(Ef.hclustered, k= 5) ## h=7.2)
+
+Ef.hclustered.df <- as.data.frame(hcluster[["Ef"]])
+names(Ef.hclustered.df) <- "Cluster"
+Ef.hclustered.df$Cluster <- as.factor(Ef.hclustered.df$Cluster)
+
+pdf("/home/ele/Dropbox/Totta_tmp_Efposter/Ef_most_sig_lifecycle_heatmap.pdf",
+    height = 8, width = 8)
+pheatmap(Ef.cycle.diff.top.100.data,
+         color = brewer.pal(n = 11, name = "BrBG"), 
+         scale = "row",
+         cluster_rows = T, ## hc.high,
+         cluster_cols = T,
+         annotation_row = Ef.hclustered.df,
+         ## annotation_names_row = F,
+         cutree_rows = 5, 
   show_rownames = F,
[…]

### 2016-03-24 functionsEfRNAseq.R

@@ -14,17 +14,11 @@ gene.table.topGO <- function(TOGO.list, pval=0.01){
     all <- GenTable(TOGO.list[[1]], TOGO.list[[2]], topNodes=100)
     ##    all$fdr <- p.adjust(all$result1, method="BH")
     names(all)[names(all)%in%"result1"] <- "p.value"
+    all$p.value <- gsub(" ?< ?", "", all$p.value)
+    all$p.value <- as.numeric(all$p.value)
+    all$adj.p <- round(p.adjust(all$p.value, method="BH"), 4)
     return(all[all$p.value<pval,])
 }
-
-###### SHOULD THIS FUNCTION SIT HERE INSTEAD OF 1_ballgown...?
-#select.from.stats.results <- function (stats.results, bg.obj, qval=0.05){
-#    diff_id <- stats.results[which(stats.results$qval<qval), "id"]
-#    diff_id <- as.character(diff_id)
-#    diff_genes <- bg.obj@indexes$t2g[bg.obj@indexes$t2g$t_id%in%diff_id,
-#                                     "g_id"]
-#    return(as.character(diff_genes))
-#}
 
 
 ## adjusted functions from ballgown

### 2016-03-24 6_enrichment.R
renaming clustering script

@@ -3,7 +3,7 @@ library(topGO)
 library(xtable)
 
 if(!exists("hcluster")){
-    source("clustering.R")
+    source("5_clustering.R")
 }
 
 if(!exists("gene2GO")){

### 2016-04-08 additional_files/tex/cluster1_BP_Mm_EarlyUp.tex
Large changes in annotation script and lots of small adjustments in others (paths changed, typos, ggpair command adjustment... etc)

@@
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Thu Mar 24 06:00:18 2016
+% latex table generated in R 3.2.2 by xtable 1.8-2 package
+% Fri Apr  8 11:20:33 2016
 \begin{table}[ht]
@@
   \hline
-1 & GO:0045087 & innate immune response & 340 &  61 & 7.54 & 0.00 & 0.00 \\ 
-  2 & GO:0006955 & immune response & 674 &  79 & 14.95 & 0.00 & 0.00 \\ 
-  3 & GO:0006952 & defense response & 724 &  75 & 16.06 & 0.00 & 0.00 \\ 
-  4 & GO:0002376 & immune system process & 1326 &  98 & 29.41 & 0.00 & 0.00 \\ 
-  5 & GO:0098542 & defense response to other organism & 282 &  48 & 6.25 & 0.00 & 0.00 \\ 
-  6 & GO:0043207 & response to external biotic stimulus & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
-  7 & GO:0051707 & response to other organism & 438 &  55 & 9.71 & 0.00 & 0.00 \\ 
-  8 & GO:0009607 & response to biotic stimulus & 460 &  55 & 10.20 & 0.00 & 0.00 \\ 
-  9 & GO:0002252 & immune effector process & 440 &  48 & 9.76 & 0.00 & 0.00 \\ 
-  10 & GO:0009615 & response to virus & 205 &  34 & 4.55 & 0.00 & 0.00 \\ 
-  11 & GO:0051607 & defense response to virus & 181 &  32 & 4.01 & 0.00 & 0.00 \\ 
-  12 & GO:0034097 & response to cytokine & 346 &  42 & 7.67 & 0.00 & 0.00 \\ 
-  13 & GO:0034341 & response to interferon-gamma &  47 &  18 & 1.04 & 0.00 & 0.00 \\ 
-  14 & GO:0006950 & response to stress & 2042 &  98 & 45.29 & 0.00 & 0.00 \\ 
-  15 & GO:0035456 & response to interferon-beta &  33 &  14 & 0.73 & 0.00 & 0.00 \\ 
-  16 & GO:0071345 & cellular response to cytokine stimulus & 276 &  32 & 6.12 & 0.00 & 0.00 \\ 
-  17 & GO:0045088 & regulation of innate immune response & 146 &  24 & 3.24 & 0.00 & 0.00 \\ 
-  18 & GO:0051704 & multi-organism process & 959 &  61 & 21.27 & 0.00 & 0.00 \\ 
-  19 & GO:0009605 & response to external stimulus & 1135 &  67 & 25.17 & 0.00 & 0.00 \\ 
-  20 & GO:0002682 & regulation of immune system process & 732 &  52 & 16.23 & 0.00 & 0.00 \\ 
-  21 & GO:0048525 & negative regulation of viral process &  66 &  17 & 1.46 & 0.00 & 0.00 \\ 
-  22 & GO:0019882 & antigen processing and presentation &  69 &  17 & 1.53 & 0.00 & 0.00 \\ 
-  23 & GO:0035458 & cellular response to interferon-beta &  27 &  12 & 0.60 & 0.00 & 0.00 \\ 
-  24 & GO:1903901 & negative regulation of viral life cycle &  63 &  16 & 1.40 & 0.00 & 0.00 \\ 
-  25 & GO:0043901 & negative regulation of multi-organism pr... & 103 &  19 & 2.28 & 0.00 & 0.00 \\ 
-  26 & GO:0050776 & regulation of immune response & 366 &  34 & 8.12 & 0.00 & 0.00 \\ 
-  27 & GO:0050792 & regulation of viral process & 118 &  20 & 2.62 & 0.00 & 0.00 \\ 
-  28 & GO:0043900 & regulation of multi-organism process & 271 &  29 & 6.01 & 0.00 & 0.00 \\ 
-  29 & GO:0071346 & cellular response to interferon-gamma &  32 &  12 & 0.71 & 0.00 & 0.00 \\ 
-  30 & GO:0019884 & antigen processing and presentation of e... &  25 &  11 & 0.55 & 0.00 & 0.00 \\ 
-  31 & GO:0044764 & multi-organism cellular process & 201 &  25 & 4.46 & 0.00 & 0.00 \\ 
-  32 & GO:0043903 & regulation of symbiosis, encompassing mu... & 137 &  21 & 3.04 & 0.00 & 0.00 \\ 
-  33 & GO:0045071 & negative regulation of viral genome repl... &  33 &  12 & 0.73 & 0.00 & 0.00 \\ 
-  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
-  35 & GO:0044419 & interspecies interaction between organis... & 225 &  26 & 4.99 & 0.00 & 0.00 \\ 
-  36 & GO:0031347 & regulation of defense response & 352 &  32 & 7.81 & 0.00 & 0.00 \\ 
-  37 & GO:0048002 & antigen processing and presentation of p... &  46 &  13 & 1.02 & 0.00 & 0.00 \\ 
-  38 & GO:0009617 & response to bacterium & 240 &  26 & 5.32 & 0.00 & 0.00 \\ 
-  39 & GO:0045089 & positive regulation of innate immune res... & 120 &  18 & 2.66 & 0.00 & 0.00 \\ 
-  40 & GO:0002478 & antigen processing and presentation of e... &  20 &   9 & 0.44 & 0.00 & 0.00 \\ 
-  41 & GO:0050778 & positive regulation of immune response & 285 &  27 & 6.32 & 0.00 & 0.00 \\ 
-  42 & GO:1903900 & regulation of viral life cycle & 109 &  17 & 2.42 & 0.00 & 0.00 \\ 
-  43 & GO:0034340 & response to type I interferon &  22 &   9 & 0.49 & 0.00 & 0.00 \\ 
-  44 & GO:0042742 & defense response to bacterium & 102 &  16 & 2.26 & 0.00 & 0.00 \\ 
-  45 & GO:0002684 & positive regulation of immune system pro... & 471 &  34 & 10.45 & 0.00 & 0.00 \\ 
-  46 & GO:0045069 & regulation of viral genome replication &  52 &  12 & 1.15 & 0.00 & 0.00 \\ 
-  47 & GO:0031349 & positive regulation of defense response & 170 &  20 & 3.77 & 0.00 & 0.00 \\ 
-  48 & GO:0019058 & viral life cycle & 153 &  19 & 3.39 & 0.00 & 0.00 \\ 
-  49 & GO:0002221 & pattern recognition receptor signaling p... &  77 &  14 & 1.71 & 0.00 & 0.00 \\ 
-  50 & GO:0016032 & viral process & 191 &  21 & 4.24 & 0.00 & 0.00 \\ 
-  51 & GO:0002758 & innate immune response-activating signal... &  80 &  14 & 1.77 & 0.00 & 0.00 \\ 
-  52 & GO:0002683 & negative regulation of immune system pro... & 257 &  24 & 5.70 & 0.00 & 0.00 \\ 
-  53 & GO:0042832 & defense response to protozoan &  20 &   8 & 0.44 & 0.00 & 0.00 \\ 
-  54 & GO:0002218 & activation of innate immune response &  86 &  14 & 1.91 & 0.00 & 0.00 \\ 
-  55 & GO:0002474 & antigen processing and presentation of p... &  28 &   9 & 0.62 & 0.00 & 0.00 \\ 
-  56 & GO:0019079 & viral genome replication &  60 &  12 & 1.33 & 0.00 & 0.00 \\ 
-  57 & GO:0050896 & response to stimulus & 4167 & 136 & 92.42 & 0.00 & 0.00 \\ 
-  58 & GO:0002483 & antigen processing and presentation of e... &   9 &   6 & 0.20 & 0.00 & 0.00 \\ 
-  59 & GO:0010033 & response to organic substance & 1314 &  61 & 29.14 & 0.00 & 0.00 \\ 
-  60 & GO:0001562 & response to protozoan &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
-  61 & GO:0098586 & cellular response to virus &  22 &   8 & 0.49 & 0.00 & 0.00 \\ 
-  62 & GO:0019883 & antigen processing and presentation of e... &  10 &   6 & 0.22 & 0.00 & 0.00 \\ 
-  63 & GO:0001816 & cytokine production & 381 &  28 & 8.45 & 0.00 & 0.00 \\ 
-  64 & GO:0035455 & response to interferon-alpha &  16 &   7 & 0.35 & 0.00 & 0.00 \\ 
-  65 & GO:0001817 & regulation of cytokine production & 334 &  26 & 7.41 & 0.00 & 0.00 \\ 
-  66 & GO:0002237 & response to molecule of bacterial origin & 154 &  17 & 3.42 & 0.00 & 0.00 \\ 
-  67 & GO:0060337 & type I interferon signaling pathway &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
-  68 & GO:0071357 & cellular response to type I interferon &  18 &   7 & 0.40 & 0.00 & 0.00 \\ 
-  69 & GO:0051250 & negative regulation of lymphocyte activa... &  89 &  13 & 1.97 & 0.00 & 0.00 \\ 
-  70 & GO:0080134 & regulation of response to stress & 790 &  42 & 17.52 & 0.00 & 0.00 \\ 
-  71 & GO:0019885 & antigen processing and presentation of e... &   7 &   5 & 0.16 & 0.00 & 0.00 \\ 
-  72 & GO:0039528 & cytoplasmic pattern recognition receptor... &  13 &   6 & 0.29 & 0.00 & 0.00 \\ 
-  73 & GO:0002695 & negative regulation of leukocyte activat... &  99 &  13 & 2.20 & 0.00 & 0.00 \\ 
-  74 & GO:0002753 & cytoplasmic pattern recognition receptor... &  22 &   7 & 0.49 & 0.00 & 0.00 \\ 
-  75 & GO:0001819 & positive regulation of cytokine producti... & 222 &  19 & 4.92 & 0.00 & 0.00 \\ 
-  76 & GO:0032606 & type I interferon production &  45 &   9 & 1.00 & 0.00 & 0.00 \\ 
-  77 & GO:0002764 & immune response-regulating signaling pat... & 181 &  17 & 4.01 & 0.00 & 0.00 \\ 
-  78 & GO:0048584 & positive regulation of response to stimu... & 1076 &  49 & 23.86 & 0.00 & 0.00 \\ 
-  79 & GO:0050866 & negative regulation of cell activation & 108 &  13 & 2.40 & 0.00 & 0.00 \\ 
-  80 & GO:0002821 & positive regulation of adaptive immune r... &  61 &  10 & 1.35 & 0.00 & 0.00 \\ 
-  81 & GO:0002697 & regulation of immune effector process & 258 &  20 & 5.72 & 0.00 & 0.00 \\ 
-  82 & GO:0045341 & MHC class I biosynthetic process &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
-  83 & GO:0045343 & regulation of MHC class I biosynthetic p... &   5 &   4 & 0.11 & 0.00 & 0.00 \\ 
-  84 & GO:0002757 & immune response-activating signal transd... & 171 &  16 & 3.79 & 0.00 & 0.00 \\ 
-  85 & GO:0002699 & positive regulation of immune effector p... & 115 &  13 & 2.55 & 0.00 & 0.00 \\ 
-  86 & GO:0019221 & cytokine-mediated signaling pathway & 175 &  16 & 3.88 & 0.00 & 0.00 \\ 
-  87 & GO:0050868 & negative regulation of T cell activation &  66 &  10 & 1.46 & 0.00 & 0.00 \\ 
-  88 & GO:0071310 & cellular response to organic substance & 981 &  45 & 21.76 & 0.00 & 0.00 \\ 
-  89 & GO:0044406 & adhesion of symbiont to host &  11 &   5 & 0.24 & 0.00 & 0.00 \\ 
-  90 & GO:0051249 & regulation of lymphocyte activation & 246 &  19 & 5.46 & 0.00 & 0.00 \\ 
-  91 & GO:0032496 & response to lipopolysaccharide & 141 &  14 & 3.13 & 0.00 & 0.00 \\ 
-  92 & GO:1903038 & negative regulation of leukocyte cell-ce... &  70 &  10 & 1.55 & 0.00 & 0.00 \\ 
-  93 & GO:0002824 & positive regulation of adaptive immune r... &  56 &   9 & 1.24 & 0.00 & 0.00 \\ 
-  94 & GO:0039529 & RIG-I signaling pathway &  12 &   5 & 0.27 & 0.00 & 0.00 \\ 
-  95 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.69 & 0.00 & 0.00 \\ 
-  96 & GO:0050830 & defense response to Gram-positive bacter... &  45 &   8 & 1.00 & 0.00 & 0.00 \\ 
-  97 & GO:0002819 & regulation of adaptive immune response &  93 &  11 & 2.06 & 0.00 & 0.00 \\ 
-  98 & GO:0002253 & activation of immune response & 194 &  16 & 4.30 & 0.00 & 0.00 \\ 
-  99 & GO:0002250 & adaptive immune response & 217 &  17 & 4.81 & 0.00 & 0.00 \\ 
-  100 & GO:1900246 & positive regulation of RIG-I signaling p... &   7 &   4 & 0.16 & 0.00 & 0.00 \\ 
+1 & GO:0045087 & innate immune response & 362 &  51 & 7.18 & 0.00 & 0.00 \\ 
+  2 & GO:0006955 & immune response & 713 &  65 & 14.14 & 0.00 & 0.00 \\ 
+  3 & GO:0002376 & immune system process & 1406 &  85 & 27.88 & 0.00 & 0.00 \\ 
+  4 & GO:0006952 & defense response & 755 &  61 & 14.97 & 0.00 & 0.00 \\ 
+  5 & GO:0098542 & defense response to other organism & 295 &  38 & 5.85 & 0.00 & 0.00 \\ 
+  6 & GO:0043207 & response to external biotic stimulus & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
+  7 & GO:0051707 & response to other organism & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
+  8 & GO:0009607 & response to biotic stimulus & 488 &  45 & 9.68 & 0.00 & 0.00 \\ 
+  9 & GO:0034097 & response to cytokine & 360 &  38 & 7.14 & 0.00 & 0.00 \\ 
+  10 & GO:0034341 & response to interferon-gamma &  47 &  16 & 0.93 & 0.00 & 0.00 \\ 
+  11 & GO:0002252 & immune effector process & 461 &  39 & 9.14 & 0.00 & 0.00 \\ 
+  12 & GO:0009615 & response to virus & 220 &  27 & 4.36 & 0.00 & 0.00 \\ 
+  13 & GO:0051607 & defense response to virus & 192 &  25 & 3.81 & 0.00 & 0.00 \\ 
+  14 & GO:0071345 & cellular response to cytokine stimulus & 292 &  29 & 5.79 & 0.00 & 0.00 \\ 
+  15 & GO:0035456 & response to interferon-beta &  35 &  12 & 0.69 & 0.00 & 0.00 \\ 
+  16 & GO:0002682 & regulation of immune system process & 774 &  47 & 15.35 & 0.00 & 0.00 \\ 
+  17 & GO:0045088 & regulation of innate immune response & 152 &  20 & 3.01 & 0.00 & 0.00 \\ 
+  18 & GO:0071346 & cellular response to interferon-gamma &  34 &  11 & 0.67 & 0.00 & 0.00 \\ 
+  19 & GO:0035458 & cellular response to interferon-beta &  29 &  10 & 0.58 & 0.00 & 0.00 \\ 
+  20 & GO:0051704 & multi-organism process & 1006 &  51 & 19.95 & 0.00 & 0.00 \\ 
+  21 & GO:0050776 & regulation of immune response & 385 &  29 & 7.63 & 0.00 & 0.00 \\ 
+  22 & GO:0031347 & regulation of defense response & 367 &  28 & 7.28 & 0.00 & 0.00 \\ 
+  23 & GO:0034340 & response to type I interferon &  26 &   9 & 0.52 & 0.00 & 0.00 \\ 
+  24 & GO:0006950 & response to stress & 2191 &  81 & 43.44 & 0.00 & 0.00 \\ 
+  25 & GO:0042832 & defense response to protozoan &  21 &   8 & 0.42 & 0.00 & 0.00 \\ 
+  26 & GO:0009617 & response to bacterium & 253 &  22 & 5.02 & 0.00 & 0.00 \\ 
+  27 & GO:0001562 & response to protozoan &  23 &   8 & 0.46 & 0.00 & 0.00 \\ 
+  28 & GO:0002683 & negative regulation of immune system pro... & 261 &  22 & 5.18 & 0.00 & 0.00 \\ 
+  29 & GO:0019882 & antigen processing and presentation &  72 &  12 & 1.43 & 0.00 & 0.00 \\ 
+  30 & GO:0019884 & antigen processing and presentation of e... &  25 &   8 & 0.50 & 0.00 & 0.00 \\ 
+  31 & GO:0010033 & response to organic substance & 1390 &  58 & 27.56 & 0.00 & 0.00 \\ 
+  32 & GO:0009605 & response to external stimulus & 1176 &  52 & 23.32 & 0.00 & 0.00 \\ 
+  33 & GO:0043900 & regulation of multi-organism process & 281 &  22 & 5.57 & 0.00 & 0.00 \\ 
+  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
+  35 & GO:0044419 & interspecies interaction between organis... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
+  36 & GO:0060337 & type I interferon signaling pathway &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
+  37 & GO:0071357 & cellular response to type I interferon &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
+  38 & GO:0042742 & defense response to bacterium & 105 &  13 & 2.08 & 0.00 & 0.00 \\ 
+  39 & GO:0044764 & multi-organism cellular process & 212 &  18 & 4.20 & 0.00 & 0.00 \\ 
+  40 & GO:0035455 & response to interferon-alpha &  15 &   6 & 0.30 & 0.00 & 0.00 \\ 
+  41 & GO:0001816 & cytokine production & 397 &  25 & 7.87 & 0.00 & 0.00 \\ 
+  42 & GO:0043903 & regulation of symbiosis, encompassing mu... & 139 &  14 & 2.76 & 0.00 & 0.00 \\ 
+  43 & GO:0002237 & response to molecule of bacterial origin & 162 &  15 & 3.21 & 0.00 & 0.00 \\ 
+  44 & GO:0050792 & regulation of viral process & 122 &  13 & 2.42 & 0.00 & 0.00 \\ 
+  45 & GO:0048525 & negative regulation of viral process &  69 &  10 & 1.37 & 0.00 & 0.00 \\ 
+  46 & GO:0045089 & positive regulation of innate immune res... & 127 &  13 & 2.52 & 0.00 & 0.00 \\ 
+  47 & GO:0051250 & negative regulation of lymphocyte activa... &  90 &  11 & 1.78 & 0.00 & 0.00 \\ 
+  48 & GO:0002684 & positive regulation of immune system pro... & 494 &  27 & 9.79 & 0.00 & 0.00 \\ 
+  49 & GO:0002478 & antigen processing and presentation of e... &  20 &   6 & 0.40 & 0.00 & 0.00 \\ 
+  50 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.61 & 0.00 & 0.00 \\ 
+  51 & GO:0050778 & positive regulation of immune response & 298 &  20 & 5.91 & 0.00 & 0.00 \\ 
+  52 & GO:0032606 & type I interferon production &  45 &   8 & 0.89 & 0.00 & 0.00 \\ 
+  53 & GO:0080134 & regulation of response to stress & 830 &  37 & 16.46 & 0.00 & 0.00 \\ 
+  54 & GO:0031349 & positive regulation of defense response & 180 &  15 & 3.57 & 0.00 & 0.00 \\ 
+  55 & GO:0045071 & negative regulation of viral genome repl... &  33 &   7 & 0.65 & 0.00 & 0.00 \\ 
+  56 & GO:0048002 & antigen processing and presentation of p... &  47 &   8 & 0.93 & 0.00 & 0.00 \\ 
+  57 & GO:0002695 & negative regulation of leukocyte activat... & 100 &  11 & 1.98 & 0.00 & 0.00 \\ 
+  58 & GO:0001817 & regulation of cytokine production & 344 &  21 & 6.82 & 0.00 & 0.00 \\ 
+  59 & GO:0019221 & cytokine-mediated signaling pathway & 188 &  15 & 3.73 & 0.00 & 0.00 \\ 
+  60 & GO:0032607 & interferon-alpha production &  14 &   5 & 0.28 & 0.00 & 0.00 \\ 
+  61 & GO:1903901 & negative regulation of viral life cycle &  66 &   9 & 1.31 & 0.00 & 0.00 \\ 
+  62 & GO:0098586 & cellular response to virus &  24 &   6 & 0.48 & 0.00 & 0.00 \\ 
+  63 & GO:0071310 & cellular response to organic substance & 1039 &  42 & 20.60 & 0.00 & 0.00 \\ 
+  64 & GO:0002477 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
+  65 & GO:0002481 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
+  66 & GO:0043901 & negative regulation of multi-organism pr... & 108 &  11 & 2.14 & 0.00 & 0.00 \\ 
+  67 & GO:0050777 & negative regulation of immune response &  89 &  10 & 1.76 & 0.00 & 0.00 \\ 
+  68 & GO:0050866 & negative regulation of cell activation & 109 &  11 & 2.16 & 0.00 & 0.00 \\ 
+  69 & GO:0016032 & viral process & 200 &  15 & 3.97 & 0.00 & 0.00 \\ 
+  70 & GO:0002483 & antigen processing and presentation of e... &   9 &   4 & 0.18 & 0.00 & 0.00 \\ 
+  71 & GO:0002697 & regulation of immune effector process & 271 &  17 & 5.37 & 0.00 & 0.00 \\ 
+  72 & GO:0019883 & antigen processing and presentation of e... &  10 &   4 & 0.20 & 0.00 & 0.00 \\ 
+  73 & GO:0002428 & antigen processing and presentation of p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  74 & GO:0045341 & MHC class I biosynthetic process &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  75 & GO:0045343 & regulation of MHC class I biosynthetic p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  76 & GO:0002221 & pattern recognition receptor signaling p... &  81 &   9 & 1.61 & 0.00 & 0.00 \\ 
+  77 & GO:0042221 & response to chemical & 1868 &  61 & 37.04 & 0.00 & 0.00 \\ 
+  78 & GO:0032496 & response to lipopolysaccharide & 148 &  12 & 2.93 & 0.00 & 0.00 \\ 
+  79 & GO:0050830 & defense response to Gram-positive bacter... &  48 &   7 & 0.95 & 0.00 & 0.00 \\ 
+  80 & GO:0002758 & innate immune response-activating signal... &  84 &   9 & 1.67 & 0.00 & 0.00 \\ 
+  81 & GO:0002449 & lymphocyte mediated immunity & 150 &  12 & 2.97 & 0.00 & 0.00 \\ 
+  82 & GO:0050868 & negative regulation of T cell activation &  66 &   8 & 1.31 & 0.00 & 0.00 \\ 
+  83 & GO:0048584 & positive regulation of response to stimu... & 1130 &  42 & 22.41 & 0.00 & 0.00 \\ 
+  84 & GO:0070887 & cellular response to chemical stimulus & 1330 &  47 & 26.37 & 0.00 & 0.00 \\ 
+  85 & GO:0051249 & regulation of lymphocyte activation & 259 &  16 & 5.14 & 0.00 & 0.00 \\ 
+  86 & GO:0001819 & positive regulation of cytokine producti... & 232 &  15 & 4.60 & 0.00 & 0.00 \\ 
+  87 & GO:0002716 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
+  88 & GO:0045953 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
+  89 & GO:0050896 & response to stimulus & 4441 & 117 & 88.06 & 0.00 & 0.00 \\ 
+  90 & GO:1903038 & negative regulation of leukocyte cell-ce... &  71 &   8 & 1.41 & 0.00 & 0.00 \\ 
+  91 & GO:0002218 & activation of innate immune response &  91 &   9 & 1.80 & 0.00 & 0.00 \\ 
+  92 & GO:0031341 & regulation of cell killing &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
+  93 & GO:0045069 & regulation of viral genome replication &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
+  94 & GO:1903900 & regulation of viral life cycle & 113 &  10 & 2.24 & 0.00 & 0.00 \\ 
+  95 & GO:0002753 & cytoplasmic pattern recognition receptor... &  24 &   5 & 0.48 & 0.00 & 0.00 \\ 
+  96 & GO:0032647 & regulation of interferon-alpha productio... &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
+  97 & GO:0044406 & adhesion of symbiont to host &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
+  98 & GO:0001906 & cell killing &  73 &   8 & 1.45 & 0.00 & 0.00 \\ 
+  99 & GO:0019058 & viral life cycle & 163 &  12 & 3.23 & 0.00 & 0.00 \\ 
+  100 & GO:0042267 & natural killer cell mediated cytotoxicit... &  39 &   6 & 0.77 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-08 additional_files/tex/cluster2_BP_Mm_7dpiUp.tex
Large changes in annotation script and lots of small adjustments in others (paths changed, typos, ggpair command adjustment... etc)

@@
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Thu Mar 24 05:57:46 2016
+% latex table generated in R 3.2.2 by xtable 1.8-2 package
+% Fri Apr  8 11:17:23 2016
 \begin{table}[ht]
@@
   \hline
-1 & GO:0002376 & immune system process & 1326 & 136 & 44.91 & 0.00 & 0.00 \\ 
-  2 & GO:0006955 & immune response & 674 &  91 & 22.83 & 0.00 & 0.00 \\ 
-  3 & GO:0045321 & leukocyte activation & 487 &  74 & 16.50 & 0.00 & 0.00 \\ 
-  4 & GO:0001775 & cell activation & 557 &  79 & 18.87 & 0.00 & 0.00 \\ 
-  5 & GO:0022610 & biological adhesion & 769 &  90 & 26.05 & 0.00 & 0.00 \\ 
-  6 & GO:0044700 & single organism signaling & 2837 & 190 & 96.09 & 0.00 & 0.00 \\ 
-  7 & GO:0007155 & cell adhesion & 762 &  89 & 25.81 & 0.00 & 0.00 \\ 
-  8 & GO:0023052 & signaling & 2844 & 190 & 96.33 & 0.00 & 0.00 \\ 
-  9 & GO:0001816 & cytokine production & 381 &  62 & 12.91 & 0.00 & 0.00 \\ 
-  10 & GO:0007154 & cell communication & 2986 & 194 & 101.14 & 0.00 & 0.00 \\ 
-  11 & GO:0007165 & signal transduction & 2643 & 180 & 89.52 & 0.00 & 0.00 \\ 
-  12 & GO:0002684 & positive regulation of immune system pro... & 471 &  68 & 15.95 & 0.00 & 0.00 \\ 
-  13 & GO:0007159 & leukocyte cell-cell adhesion & 317 &  55 & 10.74 & 0.00 & 0.00 \\ 
-  14 & GO:0002682 & regulation of immune system process & 732 &  84 & 24.79 & 0.00 & 0.00 \\ 
-  15 & GO:0046649 & lymphocyte activation & 412 &  62 & 13.96 & 0.00 & 0.00 \\ 
-  16 & GO:0048518 & positive regulation of biological proces... & 3218 & 200 & 109.00 & 0.00 & 0.00 \\ 
-  17 & GO:0050896 & response to stimulus & 4167 & 234 & 141.14 & 0.00 & 0.00 \\ 
-  18 & GO:0051239 & regulation of multicellular organismal p... & 1493 & 124 & 50.57 & 0.00 & 0.00 \\ 
-  19 & GO:0098602 & single organism cell adhesion & 498 &  67 & 16.87 & 0.00 & 0.00 \\ 
-  20 & GO:0016337 & single organismal cell-cell adhesion & 458 &  64 & 15.51 & 0.00 & 0.00 \\ 
-  21 & GO:0001817 & regulation of cytokine production & 334 &  54 & 11.31 & 0.00 & 0.00 \\ 
-  22 & GO:0006954 & inflammatory response & 316 &  52 & 10.70 & 0.00 & 0.00 \\ 
-  23 & GO:0034109 & homotypic cell-cell adhesion & 346 &  54 & 11.72 & 0.00 & 0.00 \\ 
-  24 & GO:0098609 & cell-cell adhesion & 485 &  64 & 16.43 & 0.00 & 0.00 \\ 
-  25 & GO:0006952 & defense response & 724 &  79 & 24.52 & 0.00 & 0.00 \\ 
-  26 & GO:0070486 & leukocyte aggregation & 301 &  50 & 10.20 & 0.00 & 0.00 \\ 
-  27 & GO:0051716 & cellular response to stimulus & 3508 & 204 & 118.82 & 0.00 & 0.00 \\ 
-  28 & GO:0040011 & locomotion & 869 &  85 & 29.43 & 0.00 & 0.00 \\ 
-  29 & GO:0030155 & regulation of cell adhesion & 397 &  55 & 13.45 & 0.00 & 0.00 \\ 
-  30 & GO:0042110 & T cell activation & 292 &  47 & 9.89 & 0.00 & 0.00 \\ 
-  31 & GO:0070489 & T cell aggregation & 292 &  47 & 9.89 & 0.00 & 0.00 \\ 
-  32 & GO:0048522 & positive regulation of cellular process & 2886 & 177 & 97.75 & 0.00 & 0.00 \\ 
-  33 & GO:0071593 & lymphocyte aggregation & 294 &  47 & 9.96 & 0.00 & 0.00 \\ 
-  34 & GO:0048870 & cell motility & 750 &  77 & 25.40 & 0.00 & 0.00 \\ 
-  35 & GO:0051674 & localization of cell & 750 &  77 & 25.40 & 0.00 & 0.00 \\ 
-  36 & GO:0002694 & regulation of leukocyte activation & 288 &  46 & 9.76 & 0.00 & 0.00 \\ 
-  37 & GO:0016477 & cell migration & 712 &  74 & 24.12 & 0.00 & 0.00 \\ 
-  38 & GO:0050865 & regulation of cell activation & 306 &  47 & 10.36 & 0.00 & 0.00 \\ 
-  39 & GO:0044707 & single-multicellular organism process & 3254 & 188 & 110.22 & 0.00 & 0.00 \\ 
-  40 & GO:0048583 & regulation of response to stimulus & 2012 & 137 & 68.15 & 0.00 & 0.00 \\ 
-  41 & GO:0050900 & leukocyte migration & 157 &  33 & 5.32 & 0.00 & 0.00 \\ 
-  42 & GO:0002521 & leukocyte differentiation & 328 &  47 & 11.11 & 0.00 & 0.00 \\ 
-  43 & GO:0032501 & multicellular organismal process & 3326 & 189 & 112.66 & 0.00 & 0.00 \\ 
-  44 & GO:0065007 & biological regulation & 6019 & 280 & 203.87 & 0.00 & 0.00 \\ 
-  45 & GO:0050794 & regulation of cellular process & 5444 & 262 & 184.40 & 0.00 & 0.00 \\ 
-  46 & GO:0051249 & regulation of lymphocyte activation & 246 &  40 & 8.33 & 0.00 & 0.00 \\ 
-  47 & GO:0002520 & immune system development & 593 &  63 & 20.09 & 0.00 & 0.00 \\ 
-  48 & GO:0001819 & positive regulation of cytokine producti... & 222 &  37 & 7.52 & 0.00 & 0.00 \\ 
-  49 & GO:0006928 & movement of cell or subcellular componen... & 947 &  82 & 32.08 & 0.00 & 0.00 \\ 
-  50 & GO:0050789 & regulation of biological process & 5734 & 268 & 194.22 & 0.00 & 0.00 \\ 
-  51 & GO:0051240 & positive regulation of multicellular org... & 856 &  77 & 28.99 & 0.00 & 0.00 \\ 
-  52 & GO:0030098 & lymphocyte differentiation & 225 &  37 & 7.62 & 0.00 & 0.00 \\ 
-  53 & GO:0022407 & regulation of cell-cell adhesion & 238 &  38 & 8.06 & 0.00 & 0.00 \\ 
-  54 & GO:0048534 & hematopoietic or lymphoid organ developm... & 557 &  59 & 18.87 & 0.00 & 0.00 \\ 
-  55 & GO:0048584 & positive regulation of response to stimu... & 1076 &  87 & 36.45 & 0.00 & 0.00 \\ 
-  56 & GO:0050776 & regulation of immune response & 366 &  46 & 12.40 & 0.00 & 0.00 \\ 
-  57 & GO:0007166 & cell surface receptor signaling pathway & 1252 &  95 & 42.41 & 0.00 & 0.00 \\ 
-  58 & GO:0002250 & adaptive immune response & 217 &  35 & 7.35 & 0.00 & 0.00 \\ 
-  59 & GO:0009605 & response to external stimulus & 1135 &  89 & 38.44 & 0.00 & 0.00 \\ 
-  60 & GO:0046651 & lymphocyte proliferation & 172 &  31 & 5.83 & 0.00 & 0.00 \\ 
-  61 & GO:0032943 & mononuclear cell proliferation & 174 &  31 & 5.89 & 0.00 & 0.00 \\ 
-  62 & GO:0060326 & cell chemotaxis & 141 &  28 & 4.78 & 0.00 & 0.00 \\ 
-  63 & GO:0045785 & positive regulation of cell adhesion & 226 &  35 & 7.65 & 0.00 & 0.00 \\ 
-  64 & GO:0070887 & cellular response to chemical stimulus & 1258 &  94 & 42.61 & 0.00 & 0.00 \\ 
-  65 & GO:0070661 & leukocyte proliferation & 178 &  31 & 6.03 & 0.00 & 0.00 \\ 
-  66 & GO:0035556 & intracellular signal transduction & 1498 & 105 & 50.74 & 0.00 & 0.00 \\ 
-  67 & GO:1903037 & regulation of leukocyte cell-cell adhesi... & 179 &  31 & 6.06 & 0.00 & 0.00 \\ 
-  68 & GO:0008219 & cell death & 1318 &  96 & 44.64 & 0.00 & 0.00 \\ 
-  69 & GO:0016265 & death & 1321 &  96 & 44.74 & 0.00 & 0.00 \\ 
-  70 & GO:0030595 & leukocyte chemotaxis &  97 &  23 & 3.29 & 0.00 & 0.00 \\ 
-  71 & GO:0050867 & positive regulation of cell activation & 184 &  31 & 6.23 & 0.00 & 0.00 \\ 
-  72 & GO:0040017 & positive regulation of locomotion & 262 &  37 & 8.87 & 0.00 & 0.00 \\ 
-  73 & GO:0097530 & granulocyte migration &  63 &  19 & 2.13 & 0.00 & 0.00 \\ 
-  74 & GO:0022409 & positive regulation of cell-cell adhesio... & 129 &  26 & 4.37 & 0.00 & 0.00 \\ 
-  75 & GO:0097529 & myeloid leukocyte migration &  90 &  22 & 3.05 & 0.00 & 0.00 \\ 
-  76 & GO:0006935 & chemotaxis & 266 &  37 & 9.01 & 0.00 & 0.00 \\ 
-  77 & GO:0030097 & hemopoiesis & 528 &  54 & 17.88 & 0.00 & 0.00 \\ 
-  78 & GO:0042330 & taxis & 267 &  37 & 9.04 & 0.00 & 0.00 \\ 
-  79 & GO:0002696 & positive regulation of leukocyte activat... & 179 &  30 & 6.06 & 0.00 & 0.00 \\ 
-  80 & GO:0030335 & positive regulation of cell migration & 245 &  35 & 8.30 & 0.00 & 0.00 \\ 
-  81 & GO:0010646 & regulation of cell communication & 1709 & 112 & 57.89 & 0.00 & 0.00 \\ 
-  82 & GO:0050793 & regulation of developmental process & 1301 &  93 & 44.07 & 0.00 & 0.00 \\ 
-  83 & GO:2000147 & positive regulation of cell motility & 250 &  35 & 8.47 & 0.00 & 0.00 \\ 
-  84 & GO:0023051 & regulation of signaling & 1678 & 110 & 56.84 & 0.00 & 0.00 \\ 
-  85 & GO:0002460 & adaptive immune response based on somati... & 150 &  27 & 5.08 & 0.00 & 0.00 \\ 
-  86 & GO:0001818 & negative regulation of cytokine producti... & 128 &  25 & 4.34 & 0.00 & 0.00 \\ 
-  87 & GO:0071621 & granulocyte chemotaxis &  54 &  17 & 1.83 & 0.00 & 0.00 \\ 
-  88 & GO:0034110 & regulation of homotypic cell-cell adhesi... & 189 &  30 & 6.40 & 0.00 & 0.00 \\ 
-  89 & GO:1990266 & neutrophil migration &  55 &  17 & 1.86 & 0.00 & 0.00 \\ 
-  90 & GO:0051272 & positive regulation of cellular componen... & 257 &  35 & 8.70 & 0.00 & 0.00 \\ 
-  91 & GO:1903039 & positive regulation of leukocyte cell-ce... & 110 &  23 & 3.73 & 0.00 & 0.00 \\ 
-  92 & GO:0009966 & regulation of signal transduction & 1495 & 101 & 50.64 & 0.00 & 0.00 \\ 
-  93 & GO:0042221 & response to chemical & 1780 & 113 & 60.29 & 0.00 & 0.00 \\ 
-  94 & GO:0002443 & leukocyte mediated immunity & 196 &  30 & 6.64 & 0.00 & 0.00 \\ 
-  95 & GO:0050863 & regulation of T cell activation & 172 &  28 & 5.83 & 0.00 & 0.00 \\ 
-  96 & GO:0051270 & regulation of cellular component movemen... & 467 &  48 & 15.82 & 0.00 & 0.00 \\ 
-  97 & GO:0044763 & single-organism cellular process & 6624 & 285 & 224.37 & 0.00 & 0.00 \\ 
-  98 & GO:0006915 & apoptotic process & 1219 &  87 & 41.29 & 0.00 & 0.00 \\ 
-  99 & GO:0012501 & programmed cell death & 1241 &  88 & 42.03 & 0.00 & 0.00 \\ 
-  100 & GO:0002252 & immune effector process & 440 &  46 & 14.90 & 0.00 & 0.00 \\ 
+1 & GO:0055114 & oxidation-reduction process & 650 &  28 & 9.76 & 0.00 & 0.00 \\ 
+  2 & GO:0044281 & small molecule metabolic process & 1261 &  42 & 18.94 & 0.00 & 0.00 \\ 
+  3 & GO:0006629 & lipid metabolic process & 796 &  29 & 11.95 & 0.00 & 0.00 \\ 
+  4 & GO:0044710 & single-organism metabolic process & 3313 &  76 & 49.76 & 0.00 & 0.00 \\ 
+  5 & GO:0006805 & xenobiotic metabolic process &  42 &   6 & 0.63 & 0.00 & 0.00 \\ 
+  6 & GO:0071466 & cellular response to xenobiotic stimulus &  46 &   6 & 0.69 & 0.00 & 0.00 \\ 
+  7 & GO:0009410 & response to xenobiotic stimulus &  51 &   6 & 0.77 & 0.00 & 0.00 \\ 
+  8 & GO:0006140 & regulation of nucleotide metabolic proce... & 103 &   8 & 1.55 & 0.00 & 0.00 \\ 
+  9 & GO:1901387 & positive regulation of voltage-gated cal... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
+  10 & GO:0009117 & nucleotide metabolic process & 400 &  16 & 6.01 & 0.00 & 0.00 \\ 
+  11 & GO:0017144 & drug metabolic process &  23 &   4 & 0.35 & 0.00 & 0.00 \\ 
+  12 & GO:0019369 & arachidonic acid metabolic process &  24 &   4 & 0.36 & 0.00 & 0.00 \\ 
+  13 & GO:0006753 & nucleoside phosphate metabolic process & 408 &  16 & 6.13 & 0.00 & 0.00 \\ 
+  14 & GO:0008202 & steroid metabolic process & 169 &   9 & 2.54 & 0.00 & 0.01 \\ 
+  15 & GO:0055086 & nucleobase-containing small molecule met... & 443 &  16 & 6.65 & 0.00 & 0.01 \\ 
+  16 & GO:0015701 & bicarbonate transport &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
+  17 & GO:0042738 & exogenous drug catabolic process &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
+  18 & GO:0006631 & fatty acid metabolic process & 243 &  11 & 3.65 & 0.00 & 0.01 \\ 
+  19 & GO:0051289 & protein homotetramerization &  53 &   5 & 0.80 & 0.00 & 0.01 \\ 
+  20 & GO:0009992 & cellular water homeostasis &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
+  21 & GO:0032782 & bile acid secretion &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
+  22 & GO:0042737 & drug catabolic process &  15 &   3 & 0.23 & 0.00 & 0.01 \\ 
+  23 & GO:0006690 & icosanoid metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
+  24 & GO:1901568 & fatty acid derivative metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
+  25 & GO:0015850 & organic hydroxy compound transport & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
+  26 & GO:0043583 & ear development & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
+  27 & GO:0051262 & protein tetramerization &  87 &   6 & 1.31 & 0.00 & 0.01 \\ 
+  28 & GO:0055085 & transmembrane transport & 563 &  18 & 8.46 & 0.00 & 0.01 \\ 
+  29 & GO:0045989 & positive regulation of striated muscle c... &   5 &   2 & 0.08 & 0.00 & 0.01 \\ 
+  30 & GO:0042445 & hormone metabolic process &  89 &   6 & 1.34 & 0.00 & 0.01 \\ 
+  31 & GO:0051186 & cofactor metabolic process & 265 &  11 & 3.98 & 0.00 & 0.01 \\ 
+  32 & GO:0030104 & water homeostasis &  18 &   3 & 0.27 & 0.00 & 0.01 \\ 
+  33 & GO:0071333 & cellular response to glucose stimulus &  62 &   5 & 0.93 & 0.00 & 0.01 \\ 
+  34 & GO:0019637 & organophosphate metabolic process & 668 &  20 & 10.03 & 0.00 & 0.01 \\ 
+  35 & GO:0046903 & secretion & 625 &  19 & 9.39 & 0.00 & 0.01 \\ 
+  36 & GO:0071326 & cellular response to monosaccharide stim... &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
+  37 & GO:0071331 & cellular response to hexose stimulus &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
+  38 & GO:0019373 & epoxygenase P450 pathway &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  39 & GO:0019532 & oxalate transport &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  40 & GO:0048505 & regulation of timing of cell differentia... &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  41 & GO:0071322 & cellular response to carbohydrate stimul... &  68 &   5 & 1.02 & 0.00 & 0.01 \\ 
+  42 & GO:1900542 & regulation of purine nucleotide metaboli... &  98 &   6 & 1.47 & 0.00 & 0.01 \\ 
+  43 & GO:0050433 & regulation of catecholamine secretion &  21 &   3 & 0.32 & 0.00 & 0.01 \\ 
+  44 & GO:0015711 & organic anion transport & 207 &   9 & 3.11 & 0.00 & 0.01 \\ 
+  45 & GO:0016042 & lipid catabolic process & 172 &   8 & 2.58 & 0.00 & 0.01 \\ 
+  46 & GO:0003214 & cardiac left ventricle morphogenesis &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  47 & GO:0006833 & water transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  48 & GO:0032099 & negative regulation of appetite &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  49 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  50 & GO:0001678 & cellular glucose homeostasis &  72 &   5 & 1.08 & 0.00 & 0.01 \\ 
+  51 & GO:0032787 & monocarboxylic acid metabolic process & 377 &  13 & 5.66 & 0.00 & 0.01 \\ 
+  52 & GO:0007200 & phospholipase C-activating G-protein cou... &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
+  53 & GO:0050432 & catecholamine secretion &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
+  54 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   4 & 0.71 & 0.01 & 0.01 \\ 
+  55 & GO:0032105 & negative regulation of response to extra... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
+  56 & GO:0032108 & negative regulation of response to nutri... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
+  57 & GO:0055082 & cellular chemical homeostasis & 342 &  12 & 5.14 & 0.01 & 0.01 \\ 
+  58 & GO:0032612 & interleukin-1 production &  48 &   4 & 0.72 & 0.01 & 0.01 \\ 
+  59 & GO:0006820 & anion transport & 302 &  11 & 4.54 & 0.01 & 0.01 \\ 
+  60 & GO:0032096 & negative regulation of response to food &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  61 & GO:0033604 & negative regulation of catecholamine sec... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  62 & GO:0040034 & regulation of development, heterochronic &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  63 & GO:0060081 & membrane hyperpolarization &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  64 & GO:0060979 & vasculogenesis involved in coronary vasc... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  65 & GO:0006730 & one-carbon metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
+  66 & GO:1901661 & quinone metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
+  67 & GO:1901135 & carbohydrate derivative metabolic proces... & 677 &  19 & 10.17 & 0.01 & 0.01 \\ 
+  68 & GO:0034220 & ion transmembrane transport & 393 &  13 & 5.90 & 0.01 & 0.01 \\ 
+  69 & GO:0030004 & cellular monovalent inorganic cation hom... &  50 &   4 & 0.75 & 0.01 & 0.01 \\ 
+  70 & GO:0009187 & cyclic nucleotide metabolic process &  79 &   5 & 1.19 & 0.01 & 0.01 \\ 
+  71 & GO:0010817 & regulation of hormone levels & 267 &  10 & 4.01 & 0.01 & 0.01 \\ 
+  72 & GO:0008272 & sulfate transport &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  73 & GO:0032148 & activation of protein kinase B activity &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  74 & GO:0071875 & adrenergic receptor signaling pathway &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  75 & GO:0052652 & cyclic purine nucleotide metabolic proce... &  52 &   4 & 0.78 & 0.01 & 0.01 \\ 
+  76 & GO:0072521 & purine-containing compound metabolic pro... & 315 &  11 & 4.73 & 0.01 & 0.01 \\ 
+  77 & GO:0009190 & cyclic nucleotide biosynthetic process &  53 &   4 & 0.80 & 0.01 & 0.01 \\ 
+  78 & GO:0051937 & catecholamine transport &  28 &   3 & 0.42 & 0.01 & 0.01 \\ 
+  79 & GO:0019318 & hexose metabolic process & 154 &   7 & 2.31 & 0.01 & 0.01 \\ 
+  80 & GO:0045981 & positive regulation of nucleotide metabo... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
+  81 & GO:1900544 & positive regulation of purine nucleotide... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
+  82 & GO:0045980 & negative regulation of nucleotide metabo... &  29 &   3 & 0.44 & 0.01 & 0.01 \\ 
+  83 & GO:0002275 & myeloid cell activation involved in immu... &  55 &   4 & 0.83 & 0.01 & 0.01 \\ 
+  84 & GO:0032098 & regulation of appetite &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
+  85 & GO:0051923 & sulfation &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
+  86 & GO:0042493 & response to drug & 158 &   7 & 2.37 & 0.01 & 0.01 \\ 
    \hline

### 2016-04-08 additional_files/tex/cluster3_BP_Mm_7dpiDown.tex
Large changes in annotation script and lots of small adjustments in others (paths changed, typos, ggpair command adjustment... etc)

@@
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Thu Mar 24 05:59:04 2016
+% latex table generated in R 3.2.2 by xtable 1.8-2 package
+% Fri Apr  8 11:18:59 2016
 \begin{table}[ht]
@@
   \hline
-1 & GO:0055114 & oxidation-reduction process & 597 &  30 & 8.37 & 0.00 & 0.00 \\ 
-  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  21 & 4.90 & 0.00 & 0.00 \\ 
-  3 & GO:0019752 & carboxylic acid metabolic process & 578 &  27 & 8.10 & 0.00 & 0.00 \\ 
-  4 & GO:0006631 & fatty acid metabolic process & 234 &  17 & 3.28 & 0.00 & 0.00 \\ 
-  5 & GO:0043436 & oxoacid metabolic process & 617 &  27 & 8.65 & 0.00 & 0.00 \\ 
-  6 & GO:0006082 & organic acid metabolic process & 625 &  27 & 8.76 & 0.00 & 0.00 \\ 
-  7 & GO:0044281 & small molecule metabolic process & 1187 &  39 & 16.63 & 0.00 & 0.00 \\ 
-  8 & GO:0006629 & lipid metabolic process & 767 &  29 & 10.75 & 0.00 & 0.00 \\ 
-  9 & GO:0044710 & single-organism metabolic process & 3037 &  69 & 42.56 & 0.00 & 0.00 \\ 
-  10 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 0.66 & 0.00 & 0.00 \\ 
-  11 & GO:0006690 & icosanoid metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
-  12 & GO:1901568 & fatty acid derivative metabolic process &  56 &   7 & 0.78 & 0.00 & 0.00 \\ 
-  13 & GO:0006805 & xenobiotic metabolic process &  38 &   6 & 0.53 & 0.00 & 0.00 \\ 
-  14 & GO:0019369 & arachidonic acid metabolic process &  25 &   5 & 0.35 & 0.00 & 0.00 \\ 
-  15 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &   6 & 0.60 & 0.00 & 0.00 \\ 
-  16 & GO:0019395 & fatty acid oxidation &  65 &   7 & 0.91 & 0.00 & 0.00 \\ 
-  17 & GO:0034440 & lipid oxidation &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
-  18 & GO:0072329 & monocarboxylic acid catabolic process &  66 &   7 & 0.92 & 0.00 & 0.00 \\ 
-  19 & GO:0042738 & exogenous drug catabolic process &  15 &   4 & 0.21 & 0.00 & 0.00 \\ 
-  20 & GO:0009410 & response to xenobiotic stimulus &  47 &   6 & 0.66 & 0.00 & 0.00 \\ 
-  21 & GO:0019373 & epoxygenase P450 pathway &   6 &   3 & 0.08 & 0.00 & 0.00 \\ 
-  22 & GO:0044255 & cellular lipid metabolic process & 579 &  21 & 8.11 & 0.00 & 0.00 \\ 
-  23 & GO:0042737 & drug catabolic process &  16 &   4 & 0.22 & 0.00 & 0.00 \\ 
-  24 & GO:0042493 & response to drug & 157 &  10 & 2.20 & 0.00 & 0.00 \\ 
-  25 & GO:0032099 & negative regulation of appetite &   8 &   3 & 0.11 & 0.00 & 0.00 \\ 
-  26 & GO:0009062 & fatty acid catabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
-  27 & GO:0033559 & unsaturated fatty acid metabolic process &  59 &   6 & 0.83 & 0.00 & 0.00 \\ 
-  28 & GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
-  29 & GO:0032096 & negative regulation of response to food &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
-  30 & GO:0060979 & vasculogenesis involved in coronary vasc... &   9 &   3 & 0.13 & 0.00 & 0.00 \\ 
-  31 & GO:0016054 & organic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
-  32 & GO:0046395 & carboxylic acid catabolic process & 122 &   8 & 1.71 & 0.00 & 0.00 \\ 
-  33 & GO:0017144 & drug metabolic process &  24 &   4 & 0.34 & 0.00 & 0.00 \\ 
-  34 & GO:0006635 & fatty acid beta-oxidation &  45 &   5 & 0.63 & 0.00 & 0.00 \\ 
-  35 & GO:0030258 & lipid modification & 127 &   8 & 1.78 & 0.00 & 0.00 \\ 
-  36 & GO:0032095 & regulation of response to food &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
-  37 & GO:0032098 & regulation of appetite &  11 &   3 & 0.15 & 0.00 & 0.00 \\ 
-  38 & GO:2001023 & regulation of response to drug &   3 &   2 & 0.04 & 0.00 & 0.00 \\ 
-  39 & GO:1901615 & organic hydroxy compound metabolic proce... & 287 &  12 & 4.02 & 0.00 & 0.00 \\ 
-  40 & GO:0015701 & bicarbonate transport &  15 &   3 & 0.21 & 0.00 & 0.00 \\ 
-  41 & GO:0001561 & fatty acid alpha-oxidation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
-  42 & GO:0030916 & otic vesicle formation &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
-  43 & GO:0071600 & otic vesicle morphogenesis &   4 &   2 & 0.06 & 0.00 & 0.00 \\ 
-  44 & GO:0032094 & response to food &  16 &   3 & 0.22 & 0.00 & 0.00 \\ 
-  45 & GO:0048566 & embryonic digestive tract development &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
-  46 & GO:0060977 & coronary vasculature morphogenesis &  18 &   3 & 0.25 & 0.00 & 0.00 \\ 
-  47 & GO:0042760 & very long-chain fatty acid catabolic pro... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  48 & GO:0044539 & long-chain fatty acid import &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  49 & GO:0046135 & pyrimidine nucleoside catabolic process &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  50 & GO:0072366 & regulation of cellular ketone metabolic ... &   5 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  51 & GO:0097164 & ammonium ion metabolic process &  94 &   6 & 1.32 & 0.00 & 0.00 \\ 
-  52 & GO:0015711 & organic anion transport & 202 &   9 & 2.83 & 0.00 & 0.00 \\ 
-  53 & GO:0001822 & kidney development & 132 &   7 & 1.85 & 0.00 & 0.00 \\ 
-  54 & GO:0050892 & intestinal absorption &  20 &   3 & 0.28 & 0.00 & 0.00 \\ 
-  55 & GO:0016042 & lipid catabolic process & 171 &   8 & 2.40 & 0.00 & 0.00 \\ 
-  56 & GO:0019532 & oxalate transport &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  57 & GO:0060484 & lung-associated mesenchyme development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  58 & GO:0071599 & otic vesicle development &   6 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  59 & GO:0043648 & dicarboxylic acid metabolic process &  70 &   5 & 0.98 & 0.00 & 0.01 \\ 
-  60 & GO:0015718 & monocarboxylic acid transport &  71 &   5 & 0.99 & 0.00 & 0.01 \\ 
-  61 & GO:0044282 & small molecule catabolic process & 178 &   8 & 2.49 & 0.00 & 0.01 \\ 
-  62 & GO:0048754 & branching morphogenesis of an epithelial... & 106 &   6 & 1.49 & 0.00 & 0.01 \\ 
-  63 & GO:0010817 & regulation of hormone levels & 262 &  10 & 3.67 & 0.00 & 0.01 \\ 
-  64 & GO:0048002 & antigen processing and presentation of p... &  46 &   4 & 0.64 & 0.00 & 0.01 \\ 
-  65 & GO:0032105 & negative regulation of response to extra... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
-  66 & GO:0032108 & negative regulation of response to nutri... &  23 &   3 & 0.32 & 0.00 & 0.01 \\ 
-  67 & GO:0006560 & proline metabolic process &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
-  68 & GO:0090331 & negative regulation of platelet aggregat... &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
-  69 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.10 & 0.00 & 0.01 \\ 
-  70 & GO:0072001 & renal system development & 143 &   7 & 2.00 & 0.00 & 0.01 \\ 
-  71 & GO:0018958 & phenol-containing compound metabolic pro... &  48 &   4 & 0.67 & 0.00 & 0.01 \\ 
-  72 & GO:0015849 & organic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
-  73 & GO:0046942 & carboxylic acid transport & 147 &   7 & 2.06 & 0.00 & 0.01 \\ 
-  74 & GO:0070374 & positive regulation of ERK1 and ERK2 cas... &  78 &   5 & 1.09 & 0.00 & 0.01 \\ 
-  75 & GO:0006584 & catecholamine metabolic process &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
-  76 & GO:0007431 & salivary gland development &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
-  77 & GO:0009712 & catechol-containing compound metabolic p... &  25 &   3 & 0.35 & 0.00 & 0.01 \\ 
-  78 & GO:0044242 & cellular lipid catabolic process & 114 &   6 & 1.60 & 0.01 & 0.01 \\ 
-  79 & GO:0009948 & anterior/posterior axis specification &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
-  80 & GO:0060603 & mammary gland duct morphogenesis &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
-  81 & GO:0060976 & coronary vasculature development &  26 &   3 & 0.36 & 0.01 & 0.01 \\ 
-  82 & GO:0030004 & cellular monovalent inorganic cation hom... &  51 &   4 & 0.71 & 0.01 & 0.01 \\ 
-  83 & GO:0006790 & sulfur compound metabolic process & 194 &   8 & 2.72 & 0.01 & 0.01 \\ 
-  84 & GO:0008272 & sulfate transport &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
-  85 & GO:0010544 & negative regulation of platelet activati... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
-  86 & GO:0072364 & regulation of cellular ketone metabolic ... &   9 &   2 & 0.13 & 0.01 & 0.01 \\ 
-  87 & GO:0002474 & antigen processing and presentation of p... &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
-  88 & GO:0051196 & regulation of coenzyme metabolic process &  28 &   3 & 0.39 & 0.01 & 0.01 \\ 
-  89 & GO:0006820 & anion transport & 288 &  10 & 4.04 & 0.01 & 0.01 \\ 
-  90 & GO:0015909 & long-chain fatty acid transport &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
-  91 & GO:0045844 & positive regulation of striated muscle t... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
-  92 & GO:0048636 & positive regulation of muscle organ deve... &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
-  93 & GO:0051193 & regulation of cofactor metabolic process &  29 &   3 & 0.41 & 0.01 & 0.01 \\ 
-  94 & GO:0009056 & catabolic process & 1369 &  30 & 19.18 & 0.01 & 0.01 \\ 
-  95 & GO:0051923 & sulfation &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  96 & GO:0060081 & membrane hyperpolarization &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  97 & GO:0072529 & pyrimidine-containing compound catabolic... &  10 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  98 & GO:0009880 & embryonic pattern specification &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
-  99 & GO:1901863 & positive regulation of muscle tissue dev... &  30 &   3 & 0.42 & 0.01 & 0.01 \\ 
-  100 & GO:0042060 & wound healing & 210 &   8 & 2.94 & 0.01 & 0.01 \\ 
+1 & GO:0006805 & xenobiotic metabolic process &  42 &  18 & 1.51 & 0.00 & 0.00 \\ 
+  2 & GO:0006082 & organic acid metabolic process & 661 &  68 & 23.81 & 0.00 & 0.00 \\ 
+  3 & GO:0044281 & small molecule metabolic process & 1261 & 101 & 45.43 & 0.00 & 0.00 \\ 
+  4 & GO:0032787 & monocarboxylic acid metabolic process & 377 &  49 & 13.58 & 0.00 & 0.00 \\ 
+  5 & GO:0071466 & cellular response to xenobiotic stimulus &  46 &  18 & 1.66 & 0.00 & 0.00 \\ 
+  6 & GO:0009813 & flavonoid biosynthetic process &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
+  7 & GO:0052696 & flavonoid glucuronidation &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
+  8 & GO:0043436 & oxoacid metabolic process & 652 &  66 & 23.49 & 0.00 & 0.00 \\ 
+  9 & GO:0019752 & carboxylic acid metabolic process & 612 &  63 & 22.05 & 0.00 & 0.00 \\ 
+  10 & GO:0009812 & flavonoid metabolic process &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
+  11 & GO:0052695 & cellular glucuronidation &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
+  12 & GO:0009410 & response to xenobiotic stimulus &  51 &  18 & 1.84 & 0.00 & 0.00 \\ 
+  13 & GO:0006063 & uronic acid metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
+  14 & GO:0019585 & glucuronate metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
+  15 & GO:0052697 & xenobiotic glucuronidation &   8 &   8 & 0.29 & 0.00 & 0.00 \\ 
+  16 & GO:0006629 & lipid metabolic process & 796 &  66 & 28.68 & 0.00 & 0.00 \\ 
+  17 & GO:0009719 & response to endogenous stimulus & 729 &  60 & 26.26 & 0.00 & 0.00 \\ 
+  18 & GO:0071495 & cellular response to endogenous stimulus & 561 &  50 & 20.21 & 0.00 & 0.00 \\ 
+  19 & GO:0009725 & response to hormone & 385 &  37 & 13.87 & 0.00 & 0.00 \\ 
+  20 & GO:0006631 & fatty acid metabolic process & 243 &  28 & 8.75 & 0.00 & 0.00 \\ 
+  21 & GO:0044255 & cellular lipid metabolic process & 604 &  49 & 21.76 & 0.00 & 0.00 \\ 
+  22 & GO:0055114 & oxidation-reduction process & 650 &  50 & 23.42 & 0.00 & 0.00 \\ 
+  23 & GO:0042445 & hormone metabolic process &  89 &  15 & 3.21 & 0.00 & 0.00 \\ 
+  24 & GO:0032870 & cellular response to hormone stimulus & 278 &  28 & 10.02 & 0.00 & 0.00 \\ 
+  25 & GO:0044710 & single-organism metabolic process & 3313 & 163 & 119.36 & 0.00 & 0.00 \\ 
+  26 & GO:0034754 & cellular hormone metabolic process &  43 &  10 & 1.55 & 0.00 & 0.00 \\ 
+  27 & GO:0005996 & monosaccharide metabolic process & 177 &  20 & 6.38 & 0.00 & 0.00 \\ 
+  28 & GO:0042178 & xenobiotic catabolic process &   9 &   5 & 0.32 & 0.00 & 0.00 \\ 
+  29 & GO:0042537 & benzene-containing compound metabolic pr... &  15 &   6 & 0.54 & 0.00 & 0.00 \\ 
+  30 & GO:0006790 & sulfur compound metabolic process & 200 &  21 & 7.21 & 0.00 & 0.00 \\ 
+  31 & GO:0019395 & fatty acid oxidation &  66 &  11 & 2.38 & 0.00 & 0.00 \\ 
+  32 & GO:0034440 & lipid oxidation &  67 &  11 & 2.41 & 0.00 & 0.00 \\ 
+  33 & GO:0044699 & single-organism process & 7965 & 325 & 286.96 & 0.00 & 0.00 \\ 
+  34 & GO:0042221 & response to chemical & 1868 &  99 & 67.30 & 0.00 & 0.00 \\ 
+  35 & GO:1901615 & organic hydroxy compound metabolic proce... & 299 &  26 & 10.77 & 0.00 & 0.00 \\ 
+  36 & GO:0010817 & regulation of hormone levels & 267 &  24 & 9.62 & 0.00 & 0.00 \\ 
+  37 & GO:0016042 & lipid catabolic process & 172 &  18 & 6.20 & 0.00 & 0.00 \\ 
+  38 & GO:0006811 & ion transport & 785 &  50 & 28.28 & 0.00 & 0.00 \\ 
+  39 & GO:0022600 & digestive system process &  51 &   9 & 1.84 & 0.00 & 0.00 \\ 
+  40 & GO:0044242 & cellular lipid catabolic process & 117 &  14 & 4.22 & 0.00 & 0.00 \\ 
+  41 & GO:0042738 & exogenous drug catabolic process &  14 &   5 & 0.50 & 0.00 & 0.00 \\ 
+  42 & GO:0051386 & regulation of neurotrophin TRK receptor ... &   8 &   4 & 0.29 & 0.00 & 0.00 \\ 
+  43 & GO:0044344 & cellular response to fibroblast growth f... &  42 &   8 & 1.51 & 0.00 & 0.00 \\ 
+  44 & GO:0008202 & steroid metabolic process & 169 &  17 & 6.09 & 0.00 & 0.00 \\ 
+  45 & GO:0017144 & drug metabolic process &  23 &   6 & 0.83 & 0.00 & 0.00 \\ 
+  46 & GO:0042737 & drug catabolic process &  15 &   5 & 0.54 & 0.00 & 0.00 \\ 
+  47 & GO:0005975 & carbohydrate metabolic process & 481 &  34 & 17.33 & 0.00 & 0.00 \\ 
+  48 & GO:0072329 & monocarboxylic acid catabolic process &  68 &  10 & 2.45 & 0.00 & 0.00 \\ 
+  49 & GO:0071774 & response to fibroblast growth factor &  44 &   8 & 1.59 & 0.00 & 0.00 \\ 
+  50 & GO:0007586 & digestion &  56 &   9 & 2.02 & 0.00 & 0.00 \\ 
+  51 & GO:0072001 & renal system development & 141 &  15 & 5.08 & 0.00 & 0.00 \\ 
+  52 & GO:0044539 & long-chain fatty acid import &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
+  53 & GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
+  54 & GO:0042493 & response to drug & 158 &  16 & 5.69 & 0.00 & 0.00 \\ 
+  55 & GO:0006672 & ceramide metabolic process &  58 &   9 & 2.09 & 0.00 & 0.00 \\ 
+  56 & GO:0008543 & fibroblast growth factor receptor signal... &  35 &   7 & 1.26 & 0.00 & 0.00 \\ 
+  57 & GO:0044723 & single-organism carbohydrate metabolic p... & 435 &  31 & 15.67 & 0.00 & 0.00 \\ 
+  58 & GO:0007267 & cell-cell signaling & 456 &  32 & 16.43 & 0.00 & 0.00 \\ 
+  59 & GO:0070887 & cellular response to chemical stimulus & 1330 &  72 & 47.92 & 0.00 & 0.00 \\ 
+  60 & GO:0001822 & kidney development & 131 &  14 & 4.72 & 0.00 & 0.00 \\ 
+  61 & GO:0006040 & amino sugar metabolic process &  26 &   6 & 0.94 & 0.00 & 0.00 \\ 
+  62 & GO:0034762 & regulation of transmembrane transport & 180 &  17 & 6.48 & 0.00 & 0.00 \\ 
+  63 & GO:0042572 & retinol metabolic process &  10 &   4 & 0.36 & 0.00 & 0.00 \\ 
+  64 & GO:0043269 & regulation of ion transport & 306 &  24 & 11.02 & 0.00 & 0.00 \\ 
+  65 & GO:0009062 & fatty acid catabolic process &  61 &   9 & 2.20 & 0.00 & 0.00 \\ 
+  66 & GO:0044282 & small molecule catabolic process & 183 &  17 & 6.59 & 0.00 & 0.00 \\ 
+  67 & GO:0001655 & urogenital system development & 168 &  16 & 6.05 & 0.00 & 0.00 \\ 
+  68 & GO:0022603 & regulation of anatomical structure morph... & 572 &  37 & 20.61 & 0.00 & 0.00 \\ 
+  69 & GO:0044763 & single-organism cellular process & 7155 & 291 & 257.78 & 0.00 & 0.00 \\ 
+  70 & GO:0006637 & acyl-CoA metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
+  71 & GO:0035383 & thioester metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
+  72 & GO:0006732 & coenzyme metabolic process & 227 &  19 & 8.18 & 0.00 & 0.00 \\ 
+  73 & GO:0016054 & organic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
+  74 & GO:0046395 & carboxylic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
+  75 & GO:0015698 & inorganic anion transport &  67 &   9 & 2.41 & 0.00 & 0.00 \\ 
+  76 & GO:0007169 & transmembrane receptor protein tyrosine ... & 324 &  24 & 11.67 & 0.00 & 0.00 \\ 
+  77 & GO:0009308 & amine metabolic process &  82 &  10 & 2.95 & 0.00 & 0.00 \\ 
+  78 & GO:0048011 & neurotrophin TRK receptor signaling path... &  22 &   5 & 0.79 & 0.00 & 0.00 \\ 
+  79 & GO:0030258 & lipid modification & 134 &  13 & 4.83 & 0.00 & 0.00 \\ 
+  80 & GO:0034765 & regulation of ion transmembrane transpor... & 168 &  15 & 6.05 & 0.00 & 0.00 \\ 
+  81 & GO:0055085 & transmembrane transport & 563 &  35 & 20.28 & 0.00 & 0.00 \\ 
+  82 & GO:0006576 & cellular biogenic amine metabolic proces... &  59 &   8 & 2.13 & 0.00 & 0.00 \\ 
+  83 & GO:0032941 & secretion by tissue &  46 &   7 & 1.66 & 0.00 & 0.00 \\ 
+  84 & GO:0043537 & negative regulation of blood vessel endo... &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
+  85 & GO:0060602 & branch elongation of an epithelium &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
+  86 & GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  87 & GO:0046168 & glycerol-3-phosphate catabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  88 & GO:0061031 & endodermal digestive tract morphogenesis &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  89 & GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  90 & GO:0097623 & potassium ion export across plasma membr... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  91 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 1.69 & 0.00 & 0.00 \\ 
+  92 & GO:0007268 & synaptic transmission & 245 &  19 & 8.83 & 0.00 & 0.00 \\ 
+  93 & GO:0071310 & cellular response to organic substance & 1039 &  56 & 37.43 & 0.00 & 0.00 \\ 
+  94 & GO:0019755 & one-carbon compound transport &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  95 & GO:0060736 & prostate gland growth &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  96 & GO:0090331 & negative regulation of platelet aggregat... &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  97 & GO:0008610 & lipid biosynthetic process & 383 &  26 & 13.80 & 0.00 & 0.00 \\ 
+  98 & GO:0009755 & hormone-mediated signaling pathway &  61 &   8 & 2.20 & 0.00 & 0.00 \\ 
+  99 & GO:0007167 & enzyme linked receptor protein signaling... & 487 &  31 & 17.55 & 0.00 & 0.00 \\ 
+  100 & GO:0016266 & O-glycan processing &  15 &   4 & 0.54 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-08 additional_files/tex/cluster4_BP_Mm_allInfDown.tex
Large changes in annotation script and lots of small adjustments in others (paths changed, typos, ggpair command adjustment... etc)

@@
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Thu Mar 24 06:01:34 2016
+% latex table generated in R 3.2.2 by xtable 1.8-2 package
+% Fri Apr  8 11:22:12 2016
 \begin{table}[ht]
@@
   \hline
-1 & GO:0044281 & small molecule metabolic process & 1187 &  94 & 44.61 & 0.00 & 0.00 \\ 
-  2 & GO:0032787 & monocarboxylic acid metabolic process & 350 &  40 & 13.15 & 0.00 & 0.00 \\ 
-  3 & GO:0006629 & lipid metabolic process & 767 &  63 & 28.83 & 0.00 & 0.00 \\ 
-  4 & GO:0006082 & organic acid metabolic process & 625 &  55 & 23.49 & 0.00 & 0.00 \\ 
-  5 & GO:0043436 & oxoacid metabolic process & 617 &  53 & 23.19 & 0.00 & 0.00 \\ 
-  6 & GO:0019752 & carboxylic acid metabolic process & 578 &  50 & 21.72 & 0.00 & 0.00 \\ 
-  7 & GO:0042178 & xenobiotic catabolic process &   9 &   6 & 0.34 & 0.00 & 0.00 \\ 
-  8 & GO:0006805 & xenobiotic metabolic process &  38 &  10 & 1.43 & 0.00 & 0.00 \\ 
-  9 & GO:0044710 & single-organism metabolic process & 3037 & 158 & 114.14 & 0.00 & 0.00 \\ 
-  10 & GO:0044255 & cellular lipid metabolic process & 579 &  46 & 21.76 & 0.00 & 0.00 \\ 
-  11 & GO:0006631 & fatty acid metabolic process & 234 &  25 & 8.79 & 0.00 & 0.00 \\ 
-  12 & GO:0071466 & cellular response to xenobiotic stimulus &  43 &  10 & 1.62 & 0.00 & 0.00 \\ 
-  13 & GO:0006790 & sulfur compound metabolic process & 194 &  22 & 7.29 & 0.00 & 0.00 \\ 
-  14 & GO:0034754 & cellular hormone metabolic process &  45 &  10 & 1.69 & 0.00 & 0.00 \\ 
-  15 & GO:0042445 & hormone metabolic process &  89 &  14 & 3.34 & 0.00 & 0.00 \\ 
-  16 & GO:0009410 & response to xenobiotic stimulus &  47 &  10 & 1.77 & 0.00 & 0.00 \\ 
-  17 & GO:0009719 & response to endogenous stimulus & 701 &  50 & 26.35 & 0.00 & 0.00 \\ 
-  18 & GO:0042537 & benzene-containing compound metabolic pr... &  16 &   6 & 0.60 & 0.00 & 0.00 \\ 
-  19 & GO:0016042 & lipid catabolic process & 171 &  19 & 6.43 & 0.00 & 0.00 \\ 
-  20 & GO:0009813 & flavonoid biosynthetic process &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
-  21 & GO:0052696 & flavonoid glucuronidation &   6 &   4 & 0.23 & 0.00 & 0.00 \\ 
-  22 & GO:0055114 & oxidation-reduction process & 597 &  43 & 22.44 & 0.00 & 0.00 \\ 
-  23 & GO:0071495 & cellular response to endogenous stimulus & 540 &  40 & 20.30 & 0.00 & 0.00 \\ 
-  24 & GO:0072329 & monocarboxylic acid catabolic process &  66 &  11 & 2.48 & 0.00 & 0.00 \\ 
-  25 & GO:0019626 & short-chain fatty acid catabolic process &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
-  26 & GO:0051410 & detoxification of nitrogen compound &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
-  27 & GO:0070458 & cellular detoxification of nitrogen comp... &   3 &   3 & 0.11 & 0.00 & 0.00 \\ 
-  28 & GO:0006063 & uronic acid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  29 & GO:0009812 & flavonoid metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  30 & GO:0019585 & glucuronate metabolic process &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  31 & GO:0052695 & cellular glucuronidation &   7 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  32 & GO:0044242 & cellular lipid catabolic process & 114 &  14 & 4.28 & 0.00 & 0.00 \\ 
-  33 & GO:0009725 & response to hormone & 365 &  29 & 13.72 & 0.00 & 0.00 \\ 
-  34 & GO:0008202 & steroid metabolic process & 162 &  17 & 6.09 & 0.00 & 0.00 \\ 
-  35 & GO:0044282 & small molecule catabolic process & 178 &  18 & 6.69 & 0.00 & 0.00 \\ 
-  36 & GO:0006749 & glutathione metabolic process &  42 &   8 & 1.58 & 0.00 & 0.00 \\ 
-  37 & GO:0009636 & response to toxic substance &  78 &  11 & 2.93 & 0.00 & 0.00 \\ 
-  38 & GO:0085029 & extracellular matrix assembly &  15 &   5 & 0.56 & 0.00 & 0.00 \\ 
-  39 & GO:0018916 & nitrobenzene metabolic process &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
-  40 & GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.15 & 0.00 & 0.00 \\ 
-  41 & GO:0046459 & short-chain fatty acid metabolic process &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
-  42 & GO:1901201 & regulation of extracellular matrix assem... &   9 &   4 & 0.34 & 0.00 & 0.00 \\ 
-  43 & GO:0006164 & purine nucleotide biosynthetic process & 124 &  14 & 4.66 & 0.00 & 0.00 \\ 
-  44 & GO:0009062 & fatty acid catabolic process &  59 &   9 & 2.22 & 0.00 & 0.00 \\ 
-  45 & GO:1903055 & positive regulation of extracellular mat... &  10 &   4 & 0.38 & 0.00 & 0.00 \\ 
-  46 & GO:0010817 & regulation of hormone levels & 262 &  22 & 9.85 & 0.00 & 0.00 \\ 
-  47 & GO:0072522 & purine-containing compound biosynthetic ... & 130 &  14 & 4.89 & 0.00 & 0.00 \\ 
-  48 & GO:0044699 & single-organism process & 7377 & 308 & 277.25 & 0.00 & 0.00 \\ 
-  49 & GO:0035637 & multicellular organismal signaling &  38 &   7 & 1.43 & 0.00 & 0.00 \\ 
-  50 & GO:0009165 & nucleotide biosynthetic process & 165 &  16 & 6.20 & 0.00 & 0.00 \\ 
-  51 & GO:0042572 & retinol metabolic process &  11 &   4 & 0.41 & 0.00 & 0.00 \\ 
-  52 & GO:0072521 & purine-containing compound metabolic pro... & 289 &  23 & 10.86 & 0.00 & 0.00 \\ 
-  53 & GO:1900542 & regulation of purine nucleotide metaboli... &  92 &  11 & 3.46 & 0.00 & 0.00 \\ 
-  54 & GO:0016054 & organic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
-  55 & GO:0046395 & carboxylic acid catabolic process & 122 &  13 & 4.59 & 0.00 & 0.00 \\ 
-  56 & GO:1901293 & nucleoside phosphate biosynthetic proces... & 170 &  16 & 6.39 & 0.00 & 0.00 \\ 
-  57 & GO:0019395 & fatty acid oxidation &  65 &   9 & 2.44 & 0.00 & 0.00 \\ 
-  58 & GO:0006081 & cellular aldehyde metabolic process &  41 &   7 & 1.54 & 0.00 & 0.00 \\ 
-  59 & GO:0034440 & lipid oxidation &  66 &   9 & 2.48 & 0.00 & 0.00 \\ 
-  60 & GO:0006000 & fructose metabolic process &   6 &   3 & 0.23 & 0.00 & 0.00 \\ 
-  61 & GO:0006140 & regulation of nucleotide metabolic proce... &  97 &  11 & 3.65 & 0.00 & 0.00 \\ 
-  62 & GO:0060688 & regulation of morphogenesis of a branchi... &  32 &   6 & 1.20 & 0.00 & 0.00 \\ 
-  63 & GO:0051186 & cofactor metabolic process & 249 &  20 & 9.36 & 0.00 & 0.00 \\ 
-  64 & GO:0007267 & cell-cell signaling & 440 &  30 & 16.54 & 0.00 & 0.00 \\ 
-  65 & GO:0009117 & nucleotide metabolic process & 362 &  26 & 13.61 & 0.00 & 0.00 \\ 
-  66 & GO:0034220 & ion transmembrane transport & 383 &  27 & 14.39 & 0.00 & 0.00 \\ 
-  67 & GO:0006635 & fatty acid beta-oxidation &  45 &   7 & 1.69 & 0.00 & 0.00 \\ 
-  68 & GO:0006811 & ion transport & 756 &  45 & 28.41 & 0.00 & 0.00 \\ 
-  69 & GO:0055085 & transmembrane transport & 546 &  35 & 20.52 & 0.00 & 0.00 \\ 
-  70 & GO:0006001 & fructose catabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  71 & GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  72 & GO:0034587 & piRNA metabolic process &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  73 & GO:0061624 & fructose catabolic process to hydroxyace... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  74 & GO:0061625 & glycolytic process through fructose-1-ph... &   2 &   2 & 0.08 & 0.00 & 0.00 \\ 
-  75 & GO:1901135 & carbohydrate derivative metabolic proces... & 631 &  39 & 23.72 & 0.00 & 0.00 \\ 
-  76 & GO:0009152 & purine ribonucleotide biosynthetic proce... & 117 &  12 & 4.40 & 0.00 & 0.00 \\ 
-  77 & GO:0006753 & nucleoside phosphate metabolic process & 370 &  26 & 13.91 & 0.00 & 0.00 \\ 
-  78 & GO:0003071 & renal system process involved in regulat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
-  79 & GO:0019755 & one-carbon compound transport &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
-  80 & GO:0051386 & regulation of neurotrophin TRK receptor ... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
-  81 & GO:1901203 & positive regulation of extracellular mat... &   7 &   3 & 0.26 & 0.00 & 0.00 \\ 
-  82 & GO:0006163 & purine nucleotide metabolic process & 257 &  20 & 9.66 & 0.00 & 0.00 \\ 
-  83 & GO:0044711 & single-organism biosynthetic process & 851 &  49 & 31.98 & 0.00 & 0.00 \\ 
-  84 & GO:0006040 & amino sugar metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
-  85 & GO:0017144 & drug metabolic process &  24 &   5 & 0.90 & 0.00 & 0.00 \\ 
-  86 & GO:0009118 & regulation of nucleoside metabolic proce... &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
-  87 & GO:1903578 & regulation of ATP metabolic process &  35 &   6 & 1.32 & 0.00 & 0.00 \\ 
-  88 & GO:0016266 & O-glycan processing &  15 &   4 & 0.56 & 0.00 & 0.00 \\ 
-  89 & GO:0001523 & retinoid metabolic process &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
-  90 & GO:0061098 & positive regulation of protein tyrosine ... &  25 &   5 & 0.94 & 0.00 & 0.00 \\ 
-  91 & GO:0009260 & ribonucleotide biosynthetic process & 122 &  12 & 4.59 & 0.00 & 0.00 \\ 
-  92 & GO:1901564 & organonitrogen compound metabolic proces... & 1198 &  64 & 45.03 & 0.00 & 0.00 \\ 
-  93 & GO:0008285 & negative regulation of cell proliferatio... & 400 &  27 & 15.03 & 0.00 & 0.00 \\ 
-  94 & GO:0035690 & cellular response to drug &  37 &   6 & 1.39 & 0.00 & 0.00 \\ 
-  95 & GO:0009150 & purine ribonucleotide metabolic process & 246 &  19 & 9.25 & 0.00 & 0.00 \\ 
-  96 & GO:0009755 & hormone-mediated signaling pathway &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
-  97 & GO:0043467 & regulation of generation of precursor me... &  63 &   8 & 2.37 & 0.00 & 0.00 \\ 
-  98 & GO:0046390 & ribose phosphate biosynthetic process & 124 &  12 & 4.66 & 0.00 & 0.00 \\ 
-  99 & GO:1900371 & regulation of purine nucleotide biosynth... &  50 &   7 & 1.88 & 0.00 & 0.00 \\ 
-  100 & GO:1990748 & cellular detoxification &  16 &   4 & 0.60 & 0.00 & 0.00 \\ 
+1 & GO:0002376 & immune system process & 1406 & 126 & 41.64 & 0.00 & 0.00 \\ 
+  2 & GO:0006955 & immune response & 713 &  85 & 21.12 & 0.00 & 0.00 \\ 
+  3 & GO:0050896 & response to stimulus & 4441 & 229 & 131.52 & 0.00 & 0.00 \\ 
+  4 & GO:0006954 & inflammatory response & 327 &  53 & 9.68 & 0.00 & 0.00 \\ 
+  5 & GO:0044700 & single organism signaling & 2975 & 176 & 88.11 & 0.00 & 0.00 \\ 
+  6 & GO:0006952 & defense response & 755 &  80 & 22.36 & 0.00 & 0.00 \\ 
+  7 & GO:0023052 & signaling & 2981 & 176 & 88.28 & 0.00 & 0.00 \\ 
+  8 & GO:0007165 & signal transduction & 2773 & 167 & 82.12 & 0.00 & 0.00 \\ 
+  9 & GO:0001775 & cell activation & 589 &  69 & 17.44 & 0.00 & 0.00 \\ 
+  10 & GO:0007154 & cell communication & 3137 & 178 & 92.90 & 0.00 & 0.00 \\ 
+  11 & GO:0022610 & biological adhesion & 800 &  79 & 23.69 & 0.00 & 0.00 \\ 
+  12 & GO:0002684 & positive regulation of immune system pro... & 494 &  61 & 14.63 & 0.00 & 0.00 \\ 
+  13 & GO:0007155 & cell adhesion & 791 &  78 & 23.43 & 0.00 & 0.00 \\ 
+  14 & GO:0002682 & regulation of immune system process & 774 &  76 & 22.92 & 0.00 & 0.00 \\ 
+  15 & GO:0001816 & cytokine production & 397 &  53 & 11.76 & 0.00 & 0.00 \\ 
+  16 & GO:0065007 & biological regulation & 6468 & 274 & 191.55 & 0.00 & 0.00 \\ 
+  17 & GO:0048518 & positive regulation of biological proces... & 3404 & 182 & 100.81 & 0.00 & 0.00 \\ 
+  18 & GO:0045321 & leukocyte activation & 515 &  60 & 15.25 & 0.00 & 0.00 \\ 
+  19 & GO:0040011 & locomotion & 903 &  79 & 26.74 & 0.00 & 0.00 \\ 
+  20 & GO:0098602 & single organism cell adhesion & 516 &  58 & 15.28 & 0.00 & 0.00 \\ 
+  21 & GO:0051239 & regulation of multicellular organismal p... & 1538 & 108 & 45.55 & 0.00 & 0.00 \\ 
+  22 & GO:0007159 & leukocyte cell-cell adhesion & 332 &  46 & 9.83 & 0.00 & 0.00 \\ 
+  23 & GO:0016337 & single organismal cell-cell adhesion & 474 &  55 & 14.04 & 0.00 & 0.00 \\ 
+  24 & GO:0051716 & cellular response to stimulus & 3741 & 189 & 110.79 & 0.00 & 0.00 \\ 
+  25 & GO:0098609 & cell-cell adhesion & 507 &  56 & 15.01 & 0.00 & 0.00 \\ 
+  26 & GO:0044707 & single-multicellular organism process & 3409 & 176 & 100.96 & 0.00 & 0.00 \\ 
+  27 & GO:0050789 & regulation of biological process & 6161 & 259 & 182.46 & 0.00 & 0.00 \\ 
+  28 & GO:0048870 & cell motility & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
+  29 & GO:0051674 & localization of cell & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
+  30 & GO:0032501 & multicellular organismal process & 3487 & 177 & 103.27 & 0.00 & 0.00 \\ 
+  31 & GO:0050794 & regulation of cellular process & 5853 & 249 & 173.34 & 0.00 & 0.00 \\ 
+  32 & GO:0016477 & cell migration & 740 &  67 & 21.92 & 0.00 & 0.00 \\ 
+  33 & GO:0001817 & regulation of cytokine production & 344 &  44 & 10.19 & 0.00 & 0.00 \\ 
+  34 & GO:0034109 & homotypic cell-cell adhesion & 361 &  45 & 10.69 & 0.00 & 0.00 \\ 
+  35 & GO:0048583 & regulation of response to stimulus & 2109 & 126 & 62.46 & 0.00 & 0.00 \\ 
+  36 & GO:0048584 & positive regulation of response to stimu... & 1130 &  85 & 33.47 & 0.00 & 0.00 \\ 
+  37 & GO:0050865 & regulation of cell activation & 319 &  42 & 9.45 & 0.00 & 0.00 \\ 
+  38 & GO:0048522 & positive regulation of cellular process & 3048 & 160 & 90.27 & 0.00 & 0.00 \\ 
+  39 & GO:0046649 & lymphocyte activation & 435 &  49 & 12.88 & 0.00 & 0.00 \\ 
+  40 & GO:0035556 & intracellular signal transduction & 1570 & 103 & 46.50 & 0.00 & 0.00 \\ 
+  41 & GO:0002694 & regulation of leukocyte activation & 301 &  40 & 8.91 & 0.00 & 0.00 \\ 
+  42 & GO:0070486 & leukocyte aggregation & 314 &  40 & 9.30 & 0.00 & 0.00 \\ 
+  43 & GO:1902531 & regulation of intracellular signal trans... & 991 &  76 & 29.35 & 0.00 & 0.00 \\ 
+  44 & GO:0050900 & leukocyte migration & 164 &  29 & 4.86 & 0.00 & 0.00 \\ 
+  45 & GO:0009605 & response to external stimulus & 1176 &  83 & 34.83 & 0.00 & 0.00 \\ 
+  46 & GO:0030155 & regulation of cell adhesion & 414 &  45 & 12.26 & 0.00 & 0.00 \\ 
+  47 & GO:0051249 & regulation of lymphocyte activation & 259 &  35 & 7.67 & 0.00 & 0.00 \\ 
+  48 & GO:0006928 & movement of cell or subcellular componen... & 992 &  74 & 29.38 & 0.00 & 0.00 \\ 
+  49 & GO:0001819 & positive regulation of cytokine producti... & 232 &  33 & 6.87 & 0.00 & 0.00 \\ 
+  50 & GO:0006950 & response to stress & 2191 & 123 & 64.89 & 0.00 & 0.00 \\ 
+  51 & GO:0060326 & cell chemotaxis & 145 &  26 & 4.29 & 0.00 & 0.00 \\ 
+  52 & GO:0050793 & regulation of developmental process & 1367 &  89 & 40.48 & 0.00 & 0.00 \\ 
+  53 & GO:0023051 & regulation of signaling & 1746 & 104 & 51.71 & 0.00 & 0.00 \\ 
+  54 & GO:0044763 & single-organism cellular process & 7155 & 274 & 211.90 & 0.00 & 0.00 \\ 
+  55 & GO:0006935 & chemotaxis & 270 &  34 & 8.00 & 0.00 & 0.00 \\ 
+  56 & GO:0008219 & cell death & 1372 &  88 & 40.63 & 0.00 & 0.00 \\ 
+  57 & GO:0042330 & taxis & 271 &  34 & 8.03 & 0.00 & 0.00 \\ 
+  58 & GO:0065008 & regulation of biological quality & 1978 & 112 & 58.58 & 0.00 & 0.00 \\ 
+  59 & GO:0016265 & death & 1376 &  88 & 40.75 & 0.00 & 0.00 \\ 
+  60 & GO:0070887 & cellular response to chemical stimulus & 1330 &  86 & 39.39 & 0.00 & 0.00 \\ 
+  61 & GO:0042110 & T cell activation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
+  62 & GO:0070489 & T cell aggregation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
+  63 & GO:0010646 & regulation of cell communication & 1781 & 104 & 52.74 & 0.00 & 0.00 \\ 
+  64 & GO:0071593 & lymphocyte aggregation & 307 &  36 & 9.09 & 0.00 & 0.00 \\ 
+  65 & GO:0009966 & regulation of signal transduction & 1557 &  95 & 46.11 & 0.00 & 0.00 \\ 
+  66 & GO:0051240 & positive regulation of multicellular org... & 890 &  66 & 26.36 & 0.00 & 0.00 \\ 
+  67 & GO:0048731 & system development & 2399 & 126 & 71.05 & 0.00 & 0.00 \\ 
+  68 & GO:0048856 & anatomical structure development & 2919 & 144 & 86.45 & 0.00 & 0.00 \\ 
+  69 & GO:0007166 & cell surface receptor signaling pathway & 1300 &  83 & 38.50 & 0.00 & 0.00 \\ 
+  70 & GO:0044699 & single-organism process & 7965 & 292 & 235.89 & 0.00 & 0.00 \\ 
+  71 & GO:0040017 & positive regulation of locomotion & 276 &  33 & 8.17 & 0.00 & 0.00 \\ 
+  72 & GO:0030154 & cell differentiation & 2151 & 116 & 63.70 & 0.00 & 0.00 \\ 
+  73 & GO:0002520 & immune system development & 623 &  52 & 18.45 & 0.00 & 0.00 \\ 
+  74 & GO:0002440 & production of molecular mediator of immu... & 111 &  21 & 3.29 & 0.00 & 0.00 \\ 
+  75 & GO:0048534 & hematopoietic or lymphoid organ developm... & 586 &  50 & 17.35 & 0.00 & 0.00 \\ 
+  76 & GO:0002521 & leukocyte differentiation & 347 &  37 & 10.28 & 0.00 & 0.00 \\ 
+  77 & GO:0050663 & cytokine secretion & 101 &  20 & 2.99 & 0.00 & 0.00 \\ 
+  78 & GO:0030595 & leukocyte chemotaxis & 103 &  20 & 3.05 & 0.00 & 0.00 \\ 
+  79 & GO:1902533 & positive regulation of intracellular sig... & 556 &  48 & 16.47 & 0.00 & 0.00 \\ 
+  80 & GO:0050867 & positive regulation of cell activation & 195 &  27 & 5.77 & 0.00 & 0.00 \\ 
+  81 & GO:0042221 & response to chemical & 1868 & 104 & 55.32 & 0.00 & 0.00 \\ 
+  82 & GO:0002250 & adaptive immune response & 226 &  29 & 6.69 & 0.00 & 0.00 \\ 
+  83 & GO:0097529 & myeloid leukocyte migration &  94 &  19 & 2.78 & 0.00 & 0.00 \\ 
+  84 & GO:0030335 & positive regulation of cell migration & 258 &  31 & 7.64 & 0.00 & 0.00 \\ 
+  85 & GO:0044767 & single-organism developmental process & 3254 & 153 & 96.37 & 0.00 & 0.00 \\ 
+  86 & GO:0032879 & regulation of localization & 1499 &  89 & 44.39 & 0.00 & 0.00 \\ 
+  87 & GO:0009306 & protein secretion & 311 &  34 & 9.21 & 0.00 & 0.00 \\ 
+  88 & GO:0051270 & regulation of cellular component movemen... & 494 &  44 & 14.63 & 0.00 & 0.00 \\ 
+  89 & GO:2000147 & positive regulation of cell motility & 264 &  31 & 7.82 & 0.00 & 0.00 \\ 
+  90 & GO:0022407 & regulation of cell-cell adhesion & 249 &  30 & 7.37 & 0.00 & 0.00 \\ 
+  91 & GO:0097530 & granulocyte migration &  66 &  16 & 1.95 & 0.00 & 0.00 \\ 
+  92 & GO:0012501 & programmed cell death & 1290 &  80 & 38.20 & 0.00 & 0.00 \\ 
+  93 & GO:1990266 & neutrophil migration &  57 &  15 & 1.69 & 0.00 & 0.00 \\ 
+  94 & GO:0002696 & positive regulation of leukocyte activat... & 190 &  26 & 5.63 & 0.00 & 0.00 \\ 
+  95 & GO:0006915 & apoptotic process & 1267 &  79 & 37.52 & 0.00 & 0.00 \\ 
+  96 & GO:0002252 & immune effector process & 461 &  42 & 13.65 & 0.00 & 0.00 \\ 
+  97 & GO:0032502 & developmental process & 3284 & 153 & 97.26 & 0.00 & 0.00 \\ 
+  98 & GO:0032940 & secretion by cell & 559 &  47 & 16.55 & 0.00 & 0.00 \\ 
+  99 & GO:0009617 & response to bacterium & 253 &  30 & 7.49 & 0.00 & 0.00 \\ 
+  100 & GO:0051272 & positive regulation of cellular componen... & 270 &  31 & 8.00 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-08 1_ballgown_import.R

@@ -5,7 +5,7 @@ setwd("~/Ef_RNAseq")
 
 ## IMPORT DATA USING BALLGOWN IMPORT
 if(!exists("All.bg")) {
-    All.bg = ballgown(dataDir="/SAN/Eimeria_Totta/backup/tablemaker_060815.0/",
+    All.bg = ballgown(dataDir="/SAN/Eimeria_Totta/tablemaker_d_tophat_March_c/",
         samplePattern = "*")
 }
 
@@ -17,7 +17,7 @@ source("functionsEfRNAseq.R")
 
 ## the experimental meta-data to be stored in the pData slot of the object
 add.pdata <- function (ballgown.obj){
-    samples <- sub("^tablemakertophat_(.*)(_hiseq|_forw.fastq_paired)$",
+    samples <- sub("^(.*)(_hiseq.fastq|_forw.fastq_paired)$",
                    "\\1", sampleNames(ballgown.obj))
     grouped <- sub("_rep\\d+$" , "", samples)
     infection <- ifelse(grepl("1stInf", grouped), "1st",
@@ -28,12 +28,14 @@ add.pdata <- function (ballgown.obj){
     ## variables to summarize other variables sith mutliple levels into two levels ##
     immune.status <- ifelse(mouse.strain %in% "Rag", "Rag", "competent")
     late.early <- ifelse(timepoint < 7 , "early", "late")
-    batch <- c(3, 3, 3, 3, 3, 1, 1, 1, 1, 2, 1,
-               1, 2, 3, 3, 3, 2, 3, 3, 3, 1, 1, 3, 3, 3, 3, 3)
+    batch <- c(3, 3, 3, 3, 3, 1, 1, 100, 2, 2, 0, 1, 1, 2, 3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
     seq.method <- c("hiseq", "hiseq", "hiseq", "hiseq", "hiseq",
-                    "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
-                    "hiseq", "hiseq", "hiseq", "GAII", "hiseq", "hiseq", "hiseq",
-                    "GAII", "GAII", "hiseq", "hiseq", "hiseq", "hiseq", "hiseq")
+               "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
+               "hiseq", "hiseq", "hiseq",
[…]

### 2016-04-08 3_annotations.R

@@ -2,36 +2,22 @@
 ## a TopTags (edgeR) object listing genes tested for differential
 ## expression
 
-if(!exists("Mm.1st.pass.model")){
+if(!exists("Ef.1st.pass.model")){ #Totta changes Mm to Ef here
     source("2_edgeR_diff.R")
 }
 
 ## The Bioconductor mouse annotation libary
 library(Mus.musculus)
 
-.get.cuff2name <- function(path="data/merged_mm9_Ef.gtf"){
-    gene.gtf.info <- gffRead(path) 
-    gff.attr.list <- strsplit(gene.gtf.info$attributes, " ")
-    cuff_ids <- unlist(lapply(gff.attr.list, function (x) x[[4]]))
-    gene_names <- unlist(lapply(gff.attr.list, function (x) x[[6]]))
-    cuff_ids <- gsub("\"(.*)\";$", "\\1", cuff_ids)
-    gene_names <- gsub("\"(.*)\";$", "\\1", gene_names)
-    ## discard the cufflinks re-annotation information
-    gene_names <- gsub("_\\d$", "", gene_names) 
-    cuff2name <-as.data.frame(cbind(cuff_ids, gene_names))
-    return(unique(cuff2name))
-}
-
-cuff2name <- .get.cuff2name()
-
 ## merge all this information in a data frame to combine mouse
 ## annotations with gene_id from Cufflinks get the information from
 ## this annotation dbi into data frames
 .get.annot.frame <- function(){
+    ens2geneID <- as.data.frame(org.Mm.egENSEMBL)
     name2SY <- as.data.frame(org.Mm.egSYMBOL2EG) # get 5 digit "symbol" to "short gene name" mapping
     SY2Lname<- as.data.frame(org.Mm.egGENENAME) # get 5 digit "symbol" to "long name name" mapping
     SY2GO <- as.data.frame(org.Mm.egGO) # get 5 digit "symbol" to gene ontology mapping
-    annot.frame <- merge(name2SY, cuff2name, by.x = "symbol", by.y = "gene_names")
+    annot.frame <- merge(name2SY, ens2geneID, by = "gene_id")
     annot.frame <- merge(annot.frame, SY2GO, by = "gene_id")
     annot.frame <- merge(annot.frame, SY2Lname, by = "gene_id")
     return(annot.frame)
@@ -39,21 +25,15 @@ cuff2name <- .get.cuff2name()
 
 annot.frame <- .get.annot.frame()
 
[…]

### 2016-04-08 4_compareArray.R

@@ -1,3 +1,4 @@
+## RNAseq - Microarray Comparison
 ## read the array data as done for Schmidt et al.  "Eimeria
 ## falciformis infection of the mouse caecum identifies opposing roles
 ## of IFNg-regulated host pathways for the parasite development"
@@ -110,15 +111,21 @@ RNAseq.Array.logFC <- merge(RNAseq.Array.logFC, RNAseq.logFC.ruved,
 names(RNAseq.Array.logFC) <- gsub(".x", ".plain", names(RNAseq.Array.logFC))
 
 names(RNAseq.Array.logFC) <- gsub(".y", ".ruved", names(RNAseq.Array.logFC))
-
 ## Very interesting...
 cor(RNAseq.Array.logFC[,3:14])
 cor(RNAseq.Array.logFC[,3:14], method="spearman")
 
 pdf("figures/Array_vs_RNAseq_pairs.pdf")
-ggpairs(RNAseq.Array.logFC[, 3:14], alpha=0.2) + theme_bw()
+#ggpairs(RNAseq.Array.logFC[, 3:14], alpha=0.2) + theme_bw()
+ggpairs(RNAseq.Array.logFC[, 3:14], mapping=aes(alpha=0.2)) + theme_bw()
 dev.off()
 
+##Warning messages when script is run:
+##2: In ggpairs(RNAseq.Array.logFC[, 3:14], alpha = 0.2) :
+## Extra arguments: 'alpha' are being ignored.  
+## If these are meant to be aesthetics, submit them using the 'mapping' variable within 
+## ggpairs with ggplot2::aes or ggplot2::aes_string.
+##
 pdf("figures/Array144_vs_RNAseqN7.pdf")
 ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0.plain)) +
     geom_point(alpha=0.8) +

### 2016-04-08 5_clustering.R

@@ -4,7 +4,9 @@ library(pheatmap)
 library(made4)
 library(RColorBrewer)
 
-
+if(!exists("RNAseq.Array.logFC")){
+    source("4_compareArray.R")
+}
 ########## USE THIS #######################
 ########## PHEATMAP() specifying distance and method ############
 
@@ -25,14 +27,14 @@ get.scaled.and.clustered <- function(data){
 Ef.hclustered <- get.scaled.and.clustered(Ef.cycle.diff.top.100.data)
 
 hcluster <- list()
-hcluster[["Ef"]] <- cutree(Ef.hclustered, k= 5) ## h=7.2)
+hcluster[["Ef"]] <- cutree(Ef.hclustered, k= 7) ## TK changed k from 5 ## h=7.2)
 
 Ef.hclustered.df <- as.data.frame(hcluster[["Ef"]])
 names(Ef.hclustered.df) <- "Cluster"
 Ef.hclustered.df$Cluster <- as.factor(Ef.hclustered.df$Cluster)
 
-pdf("/home/ele/Dropbox/Totta_tmp_Efposter/Ef_most_sig_lifecycle_heatmap.pdf",
-    height = 8, width = 8)
+pdf("/figures/Ef_most_sig_lifecycle_heatmap.pdf")
+#    height = 8, width = 8)
 pheatmap(Ef.cycle.diff.top.100.data,
          color = brewer.pal(n = 11, name = "BrBG"), 
          scale = "row",
@@ -41,9 +43,11 @@ pheatmap(Ef.cycle.diff.top.100.data,
          annotation_row = Ef.hclustered.df,
          ## annotation_names_row = F,
          cutree_rows = 5, 
-  show_rownames = F,
+         show_rownames = F,
          main = expression(paste(italic("E. falciformis"),
-             " genes differentially expressed between lifecycle stages")))
+             " mRNAs differently abundant between lifecycle stages")),
+         width = 8,
+         height = 8)
 dev.off()
 
 
@@ -63,7 +67,7 @@ Mm.hclustered.df <- as.data.frame(hcluster[["Mm"]])
 names(Mm.hclustered.df) <- "Cluster"
 Mm.hclustered.df$Cluster <- as.factor(Mm.hclustered.df$Cluster)
 
-pdf("/home/ele/Dropbox/Totta_tmp_Efposter/Mm_most_sig_lifecycle_heatmap.pdf",
+pdf("/figures/Mm_most_sig_lifecycle_heatmap.pdf",
     height = 8, width = 8)
[…]

### 2016-04-08 6_enrichment.R

@@ -35,8 +35,9 @@ apply(to.test, 1, function (x){
           clus = x[[1]]
           type = x[[2]]
           species = x[[3]]
-          g2G = gene2GO[[species]]
-          hcl = hcluster[[species]]
+          g2G = gene2GO[[species]] # gene2GO object created in 3_annotations script
+          hcl = hcluster[[species]] # hcluster object created in 5_clustering script
+          ## creates first part of filename for each cluster
           file.path=("~/Ef_RNAseq/additional_files/tex/cluster")
           sapply(c("MF", "BP"), function (onto){
                      res <- TOGO.all.onto(onto, exp.universe[[species]],

### 2016-04-14 additional_files/tex/SI_clusters/clusterTablesCombined.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
%% Totta:
%% This collection aims to provide an overview of the enriched GO-terms in the
%% different clusters in the heatmaps. 
%% We include both mouse and Eimeria data

%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
\usepackage[hscale=0.8,vscale=0.8]{geometry}
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font

%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{GO-terms enriched in heatmap gene clusters for mouse and \textit{Eimeria falciformis}} % Title

\author{Totta \textsc{Kasemo}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Objective}
bla bla bla
\newpage
%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION 2 
%----------------------------------------------------------------------------------------

\section{Results for \textit{E. falciformis}}
  Cluster 1, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 2, which also have a high
  abundance on day 7, these genes cluster more closely with genes in cluster 4, 
  which are highly expressed in oocysts.\\
  %\newpage
  \input{../cluster1_BP_Ef_Day7Up_1.tex}

  Cluster 2, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 1, which is also upregulated 
  on day 7 p.i., these genes cluster more closely with cluster 6, which contains
  genes highly abundant in sporozoites.\\
  \input{../cluster2_BP_Ef_Day7Up_2.tex}

  Cluster 1, MF, Ef.\\
  Overrepresented GO molecular function terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 2, which also have a high
  abundance on day 7, these genes cluster more closely with genes in cluster 4, 
  which are highly expressed in oocysts.\\
  \input{../cluster1_MF_Ef_Day7Up_1.tex}

  Cluster 2, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are 
  highly abundant on day 7 p.i.. Compared to cluster 1, which is also upregulated 
  on day 7 p.i., these genes cluster more closely with cluster 6, which contains
  genes highly abundant in sporozoites.\\
  \input{../cluster2_MF_Ef_Day7Up_2.tex}

  Cluster 3, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have a low
  abundance in sporulated oocysts and sporozoites. In mouse stages, i.e., day 3, 5, and 7 samples
  there is a tendency to upregulation among these genes.\\
  \input{../cluster3_BP_Ef_OocSporDown.tex}

  Cluster 3, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have a low
  abundance in sporulated oocysts and sporozoites. In mouse stages, i.e., day 3, 5, and 7 samples
  there is a tendency to upregulation among these genes.\\
  \input{../cluster3_MF_Ef_OocSporDown.tex}

  Cluster 4, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are highly abundant in
  sporulated oocysts. These genes cluster most closely with cluster 1, in which genes are highly
  abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.\\
  \input{../cluster4_BP_Ef_OocystsUp.tex}

  Cluster 4, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are highly abundant in
  sporulated oocysts. These genes cluster most closely with cluster 1, in which genes are highly
  abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.\\
  \input{../cluster4_MF_Ef_OocystsUp.tex}

  Cluster 5, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have low abundance
  in sporulated oocysts, sporozoites and on day 7 p.i.. These genes seem to increase in abundance
  upon infection and decrease late in infection, i.e., on day 7.\\
  \input{../cluster5_BP_Ef_OoSpoDown_day5Up.tex}

  Cluster 5, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have low abundance
  in sporulated oocysts, sporozoites and on day 7 p.i.. These genes seem to increase in abundance
  upon infection and decrease late in infection, i.e., on day 7.\\
  \input{../cluster5_MF_Ef_OoSpoDown_day5Up.tex}

  Cluster 6, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which are highly abundant in sporozoites
  but mainly downregulated in sporulated oocysts, apart from a few genes in the cluster.
  This cluster seems to distinguish sporozoites from sporulated oocysts and have an average mRNA abundance in
  all other samples.\\
  \input{../cluster6_BP_Ef_SporozoitesUp.tex}

  Cluster 6, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which are highly abundant in sporozoites
  but mainly downregulated in sporulated oocysts, apart from a few genes in the cluster.
  This cluster seems to distinguish sporozoites from sporulated oocysts and have an average mRNA abundance in
  all other samples.\\
  \input{../cluster6_MF_Ef_SporozoitesUp.tex}

  Cluster 7, BP, Ef.
  Overrepresented GO biological process terms among parasite genes which have low abundance in sporulated oocysts.
  In sporozoites and early in infection, i.e., day 3 and 5 p.i., some of these genes are highly abundant. On day
  7 p.i. these mRNAs have a below average abundance.\\
  \input{../cluster7_BP_Ef_Day7OocyDown_SporoUP.tex}

  Cluster 7, MF, Ef.
  Overrepresented GO molecular function terms among parasite genes which have low abundance in sporulated oocysts.
  In sporozoites and early in infection, i.e., day 3 and 5 p.i., some of these genes are highly abundant. On day
  7 p.i. these mRNAs have a below average abundance.\\
  \input{../cluster7_MF_Ef_Day7OocyDown_SporoUP.tex}

\newpage

\section{Results for mouse}
  Cluster 1, BP, Mm
  Overrepresented GO biological process terms among mouse genes which have low abundance pre infection and 
  become more abundant in most day 3 and 5 samples from immunecompetent mice, but less in immune deficient Rag1-/- mice. 
  The upregulation trend is most clear in three first infection day 5 samples, which also form a separate
  sample cluster. On day 7, most of these genes are not differentially abundant, or below average, i.e., they
  seem to peak on day 5 p.i..\\
  \input{../cluster1_BP_Mm_EarlyUp.tex}

  Cluster 1, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which have low abundance pre infection and 
  become more abundant in most day 3 and 5 samples from immunecompetent mice, but less in immune deficient Rag1-/- mice. 
  The upregulation trend is most clear in three first infection day 5 samples, which also form a separate
  sample cluster. On day 7, most of these genes are not differentially abundant, or below average, i.e., they
  seem to peak on day 5 p.i..\\
  \input{../cluster1_MF_Mm_EarlyUp.tex}

  Cluster 2, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are downregulated on day 7 p.i. and
  upregulated pre infection, i.e., these genes become more and more downregulated during infection.\\
  \input{../cluster2_BP_Mm_7dpiUp.tex}

  Cluster 2, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are downregulated on day 7 p.i. and
  upregulated pre infection, i.e., these genes become more and more downregulated during infection.\\
  \input{../cluster2_MF_Mm_7dpiUp.tex}

  Cluster 3, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are upregulated pre infection and
  downregulated on day 7 p.i., except in one NMRI second infection sample, possibly indicating no or
  poor infection in this sample. As in cluster 2, these mRNAs are most highly abundant before infection and
  become less abundant over the course of infection.\\
  \input{../cluster3_BP_Mm_7dpiDown.tex}

  Cluster 3, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are upregulated pre infection and
  downregulated on day 7 p.i., except in one NMRI second infection sample, possibly indicating no or
  poor infection in this sample. As in cluster 2, these mRNAs are most highly abundant before infection and
  become less abundant over the course of infection.\\
  \input{../cluster3_MF_Mm_7dpiDown.tex}

  Cluster 4, BP, Mm.
  Overrepresented GO biological process terms among mouse genes which are highly abundant on day 7 p.i.. 
  There is overall no clear difference between pre infection and days 3 or 5 among these genesi. A slight
  tendency to higher abundance in second infection samples can be seen among these genes.\\
  \input{../cluster4_BP_Mm_allInfDown.tex}

  Cluster 4, MF, Mm.
  Overrepresented GO molecular function terms among mouse genes which are highly abundant on day 7 p.i.. 
  There is overall no clear difference between pre infection and days 3 or 5 among these genesi. A slight
  tendency to higher abundance in second infection samples can be seen among these genes.\\
  \input{../cluster4_MF_Mm_allInfDown.tex}
% figure 1
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
%\caption{Density distribution of raw transcript counts (log10) for all
%  mouse samples in analysis. All samples share the same bimodal
%  distribution trend by visual inspection. The first density peak
%  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
%  detected between 1 and 100 times in that sample. The second density
%  peak occurs in the range of $10^3 - 10^4$.}
%\end{center}
%\end{figure}

%----------------------------------------------------------------------------------------
%	SECTION 5
%----------------------------------------------------------------------------------------

\section{Conclusions}

Mouse\\

Parasite\\

%----------------------------------------------------------------------------------------
%	SECTION 6
%----------------------------------------------------------------------------------------

%\section{Answers to Definitions}

%\begin{enumerate}
%\begin{item}
%  The \emph{atomic weight of an element} is the relative weight of
%  one of its atoms compared to C-12 with a weight of
%  12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with
%  a weight of 16.00. Atomic weight is also the average weight of all
%  the atoms of that element as they occur in nature.
%\end{item}
%\begin{item}
%  The \emph{units of atomic weight} are two-fold, with an identical
%  numerical value. They are g/mole of atoms (or just g/mol) or
%  amu/atom.
%\end{item}
%\begin{item}
%  \emph{Percentage discrepancy} between an accepted (literature)
%  value and an experimental value is
%\begin{equation*}
%\frac{\mathrm{experimental\;result} - \mathrm{accepted\;result}}{\mathrm{accepted\;result}}
%\end{equation*}
%\end{item}
%\end{enumerate}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

%\bibliographystyle{apalike}

%\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}

### 2016-04-14 additional_files/tex/cluster1_BP_Mm_EarlyUp.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

@@
-% latex table generated in R 3.2.2 by xtable 1.8-2 package
-% Fri Apr  8 11:20:33 2016
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Thu Apr 14 15:02:54 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllrrrrr}
+\begin{tabular}{llrrrrr}
   \hline
- & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
+GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
   \hline
-1 & GO:0045087 & innate immune response & 362 &  51 & 7.18 & 0.00 & 0.00 \\ 
-  2 & GO:0006955 & immune response & 713 &  65 & 14.14 & 0.00 & 0.00 \\ 
-  3 & GO:0002376 & immune system process & 1406 &  85 & 27.88 & 0.00 & 0.00 \\ 
-  4 & GO:0006952 & defense response & 755 &  61 & 14.97 & 0.00 & 0.00 \\ 
-  5 & GO:0098542 & defense response to other organism & 295 &  38 & 5.85 & 0.00 & 0.00 \\ 
-  6 & GO:0043207 & response to external biotic stimulus & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
-  7 & GO:0051707 & response to other organism & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
-  8 & GO:0009607 & response to biotic stimulus & 488 &  45 & 9.68 & 0.00 & 0.00 \\ 
-  9 & GO:0034097 & response to cytokine & 360 &  38 & 7.14 & 0.00 & 0.00 \\ 
-  10 & GO:0034341 & response to interferon-gamma &  47 &  16 & 0.93 & 0.00 & 0.00 \\ 
-  11 & GO:0002252 & immune effector process & 461 &  39 & 9.14 & 0.00 & 0.00 \\ 
-  12 & GO:0009615 & response to virus & 220 &  27 & 4.36 & 0.00 & 0.00 \\ 
-  13 & GO:0051607 & defense response to virus & 192 &  25 & 3.81 & 0.00 & 0.00 \\ 
-  14 & GO:0071345 & cellular response to cytokine stimulus & 292 &  29 & 5.79 & 0.00 & 0.00 \\ 
-  15 & GO:0035456 & response to interferon-beta &  35 &  12 & 0.69 & 0.00 & 0.00 \\ 
-  16 & GO:0002682 & regulation of immune system process & 774 &  47 & 15.35 & 0.00 & 0.00 \\ 
-  17 & GO:0045088 & regulation of innate immune response & 152 &  20 & 3.01 & 0.00 & 0.00 \\ 
-  18 & GO:0071346 & cellular response to interferon-gamma &  34 &  11 & 0.67 & 0.00 & 0.00 \\ 
-  19 & GO:0035458 & cellular response to interferon-beta &  29 &  10 & 0.58 & 0.00 & 0.00 \\ 
-  20 & GO:0051704 & multi-organism process & 1006 &  51 & 19.95 & 0.00 & 0.00 \\ 
-  21 & GO:0050776 & regulation of immune response & 385 &  29 & 7.63 & 0.00 & 0.00 \\ 
-  22 & GO:0031347 & regulation of defense response & 367 &  28 & 7.28 & 0.00 & 0.00 \\ 
-  23 & GO:0034340 & response to type I interferon &  26 &   9 & 0.52 & 0.00 & 0.00 \\ 
-  24 & GO:0006950 & response to stress & 2191 &  81 & 43.44 & 0.00 & 0.00 \\ 
-  25 & GO:0042832 & defense response to protozoan &  21 &   8 & 0.42 & 0.00 & 0.00 \\ 
-  26 & GO:0009617 & response to bacterium & 253 &  22 & 5.02 & 0.00 & 0.00 \\ 
-  27 & GO:0001562 & response to protozoan &  23 &   8 & 0.46 & 0.00 & 0.00 \\ 
-  28 & GO:0002683 & negative regulation of immune system pro... & 261 &  22 & 5.18 & 0.00 & 0.00 \\ 
-  29 & GO:0019882 & antigen processing and presentation &  72 &  12 & 1.43 & 0.00 & 0.00 \\ 
-  30 & GO:0019884 & antigen processing and presentation of e... &  25 &   8 & 0.50 & 0.00 & 0.00 \\ 
-  31 & GO:0010033 & response to organic substance & 1390 &  58 & 27.56 & 0.00 & 0.00 \\ 
-  32 & GO:0009605 & response to external stimulus & 1176 &  52 & 23.32 & 0.00 & 0.00 \\ 
-  33 & GO:0043900 & regulation of multi-organism process & 281 &  22 & 5.57 & 0.00 & 0.00 \\ 
-  34 & GO:0044403 & symbiosis, encompassing mutualism throug... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
-  35 & GO:0044419 & interspecies interaction between organis... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
-  36 & GO:0060337 & type I interferon signaling pathway &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
-  37 & GO:0071357 & cellular response to type I interferon &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
-  38 & GO:0042742 & defense response to bacterium & 105 &  13 & 2.08 & 0.00 & 0.00 \\ 
-  39 & GO:0044764 & multi-organism cellular process & 212 &  18 & 4.20 & 0.00 & 0.00 \\ 
-  40 & GO:0035455 & response to interferon-alpha &  15 &   6 & 0.30 & 0.00 & 0.00 \\ 
-  41 & GO:0001816 & cytokine production & 397 &  25 & 7.87 & 0.00 & 0.00 \\ 
-  42 & GO:0043903 & regulation of symbiosis, encompassing mu... & 139 &  14 & 2.76 & 0.00 & 0.00 \\ 
-  43 & GO:0002237 & response to molecule of bacterial origin & 162 &  15 & 3.21 & 0.00 & 0.00 \\ 
-  44 & GO:0050792 & regulation of viral process & 122 &  13 & 2.42 & 0.00 & 0.00 \\ 
-  45 & GO:0048525 & negative regulation of viral process &  69 &  10 & 1.37 & 0.00 & 0.00 \\ 
-  46 & GO:0045089 & positive regulation of innate immune res... & 127 &  13 & 2.52 & 0.00 & 0.00 \\ 
-  47 & GO:0051250 & negative regulation of lymphocyte activa... &  90 &  11 & 1.78 & 0.00 & 0.00 \\ 
-  48 & GO:0002684 & positive regulation of immune system pro... & 494 &  27 & 9.79 & 0.00 & 0.00 \\ 
-  49 & GO:0002478 & antigen processing and presentation of e... &  20 &   6 & 0.40 & 0.00 & 0.00 \\ 
-  50 & GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.61 & 0.00 & 0.00 \\ 
-  51 & GO:0050778 & positive regulation of immune response & 298 &  20 & 5.91 & 0.00 & 0.00 \\ 
-  52 & GO:0032606 & type I interferon production &  45 &   8 & 0.89 & 0.00 & 0.00 \\ 
-  53 & GO:0080134 & regulation of response to stress & 830 &  37 & 16.46 & 0.00 & 0.00 \\ 
-  54 & GO:0031349 & positive regulation of defense response & 180 &  15 & 3.57 & 0.00 & 0.00 \\ 
-  55 & GO:0045071 & negative regulation of viral genome repl... &  33 &   7 & 0.65 & 0.00 & 0.00 \\ 
-  56 & GO:0048002 & antigen processing and presentation of p... &  47 &   8 & 0.93 & 0.00 & 0.00 \\ 
-  57 & GO:0002695 & negative regulation of leukocyte activat... & 100 &  11 & 1.98 & 0.00 & 0.00 \\ 
-  58 & GO:0001817 & regulation of cytokine production & 344 &  21 & 6.82 & 0.00 & 0.00 \\ 
-  59 & GO:0019221 & cytokine-mediated signaling pathway & 188 &  15 & 3.73 & 0.00 & 0.00 \\ 
-  60 & GO:0032607 & interferon-alpha production &  14 &   5 & 0.28 & 0.00 & 0.00 \\ 
-  61 & GO:1903901 & negative regulation of viral life cycle &  66 &   9 & 1.31 & 0.00 & 0.00 \\ 
-  62 & GO:0098586 & cellular response to virus &  24 &   6 & 0.48 & 0.00 & 0.00 \\ 
-  63 & GO:0071310 & cellular response to organic substance & 1039 &  42 & 20.60 & 0.00 & 0.00 \\ 
-  64 & GO:0002477 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
-  65 & GO:0002481 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
-  66 & GO:0043901 & negative regulation of multi-organism pr... & 108 &  11 & 2.14 & 0.00 & 0.00 \\ 
-  67 & GO:0050777 & negative regulation of immune response &  89 &  10 & 1.76 & 0.00 & 0.00 \\ 
-  68 & GO:0050866 & negative regulation of cell activation & 109 &  11 & 2.16 & 0.00 & 0.00 \\ 
-  69 & GO:0016032 & viral process & 200 &  15 & 3.97 & 0.00 & 0.00 \\ 
-  70 & GO:0002483 & antigen processing and presentation of e... &   9 &   4 & 0.18 & 0.00 & 0.00 \\ 
-  71 & GO:0002697 & regulation of immune effector process & 271 &  17 & 5.37 & 0.00 & 0.00 \\ 
-  72 & GO:0019883 & antigen processing and presentation of e... &  10 &   4 & 0.20 & 0.00 & 0.00 \\ 
-  73 & GO:0002428 & antigen processing and presentation of p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
-  74 & GO:0045341 & MHC class I biosynthetic process &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
-  75 & GO:0045343 & regulation of MHC class I biosynthetic p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
-  76 & GO:0002221 & pattern recognition receptor signaling p... &  81 &   9 & 1.61 & 0.00 & 0.00 \\ 
-  77 & GO:0042221 & response to chemical & 1868 &  61 & 37.04 & 0.00 & 0.00 \\ 
-  78 & GO:0032496 & response to lipopolysaccharide & 148 &  12 & 2.93 & 0.00 & 0.00 \\ 
-  79 & GO:0050830 & defense response to Gram-positive bacter... &  48 &   7 & 0.95 & 0.00 & 0.00 \\ 
-  80 & GO:0002758 & innate immune response-activating signal... &  84 &   9 & 1.67 & 0.00 & 0.00 \\ 
-  81 & GO:0002449 & lymphocyte mediated immunity & 150 &  12 & 2.97 & 0.00 & 0.00 \\ 
-  82 & GO:0050868 & negative regulation of T cell activation &  66 &   8 & 1.31 & 0.00 & 0.00 \\ 
-  83 & GO:0048584 & positive regulation of response to stimu... & 1130 &  42 & 22.41 & 0.00 & 0.00 \\ 
-  84 & GO:0070887 & cellular response to chemical stimulus & 1330 &  47 & 26.37 & 0.00 & 0.00 \\ 
-  85 & GO:0051249 & regulation of lymphocyte activation & 259 &  16 & 5.14 & 0.00 & 0.00 \\ 
-  86 & GO:0001819 & positive regulation of cytokine producti... & 232 &  15 & 4.60 & 0.00 & 0.00 \\ 
-  87 & GO:0002716 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
-  88 & GO:0045953 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
-  89 & GO:0050896 & response to stimulus & 4441 & 117 & 88.06 & 0.00 & 0.00 \\ 
-  90 & GO:1903038 & negative regulation of leukocyte cell-ce... &  71 &   8 & 1.41 & 0.00 & 0.00 \\ 
-  91 & GO:0002218 & activation of innate immune response &  91 &   9 & 1.80 & 0.00 & 0.00 \\ 
-  92 & GO:0031341 & regulation of cell killing &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
-  93 & GO:0045069 & regulation of viral genome replication &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
-  94 & GO:1903900 & regulation of viral life cycle & 113 &  10 & 2.24 & 0.00 & 0.00 \\ 
-  95 & GO:0002753 & cytoplasmic pattern recognition receptor... &  24 &   5 & 0.48 & 0.00 & 0.00 \\ 
-  96 & GO:0032647 & regulation of interferon-alpha productio... &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  97 & GO:0044406 & adhesion of symbiont to host &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
-  98 & GO:0001906 & cell killing &  73 &   8 & 1.45 & 0.00 & 0.00 \\ 
-  99 & GO:0019058 & viral life cycle & 163 &  12 & 3.23 & 0.00 & 0.00 \\ 
-  100 & GO:0042267 & natural killer cell mediated cytotoxicit... &  39 &   6 & 0.77 & 0.00 & 0.00 \\ 
+GO:0045087 & innate immune response & 362 &  51 & 7.18 & 0.00 & 0.00 \\ 
+  GO:0006955 & immune response & 713 &  65 & 14.14 & 0.00 & 0.00 \\ 
+  GO:0002376 & immune system process & 1406 &  85 & 27.88 & 0.00 & 0.00 \\ 
+  GO:0006952 & defense response & 755 &  61 & 14.97 & 0.00 & 0.00 \\ 
+  GO:0098542 & defense response to other organism & 295 &  38 & 5.85 & 0.00 & 0.00 \\ 
+  GO:0043207 & response to external biotic stimulus & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
+  GO:0051707 & response to other organism & 464 &  45 & 9.20 & 0.00 & 0.00 \\ 
+  GO:0009607 & response to biotic stimulus & 488 &  45 & 9.68 & 0.00 & 0.00 \\ 
+  GO:0034097 & response to cytokine & 360 &  38 & 7.14 & 0.00 & 0.00 \\ 
+  GO:0034341 & response to interferon-gamma &  47 &  16 & 0.93 & 0.00 & 0.00 \\ 
+  GO:0002252 & immune effector process & 461 &  39 & 9.14 & 0.00 & 0.00 \\ 
+  GO:0009615 & response to virus & 220 &  27 & 4.36 & 0.00 & 0.00 \\ 
+  GO:0051607 & defense response to virus & 192 &  25 & 3.81 & 0.00 & 0.00 \\ 
+  GO:0071345 & cellular response to cytokine stimulus & 292 &  29 & 5.79 & 0.00 & 0.00 \\ 
+  GO:0035456 & response to interferon-beta &  35 &  12 & 0.69 & 0.00 & 0.00 \\ 
+  GO:0002682 & regulation of immune system process & 774 &  47 & 15.35 & 0.00 & 0.00 \\ 
+  GO:0045088 & regulation of innate immune response & 152 &  20 & 3.01 & 0.00 & 0.00 \\ 
+  GO:0071346 & cellular response to interferon-gamma &  34 &  11 & 0.67 & 0.00 & 0.00 \\ 
+  GO:0035458 & cellular response to interferon-beta &  29 &  10 & 0.58 & 0.00 & 0.00 \\ 
+  GO:0051704 & multi-organism process & 1006 &  51 & 19.95 & 0.00 & 0.00 \\ 
+  GO:0050776 & regulation of immune response & 385 &  29 & 7.63 & 0.00 & 0.00 \\ 
+  GO:0031347 & regulation of defense response & 367 &  28 & 7.28 & 0.00 & 0.00 \\ 
+  GO:0034340 & response to type I interferon &  26 &   9 & 0.52 & 0.00 & 0.00 \\ 
+  GO:0006950 & response to stress & 2191 &  81 & 43.44 & 0.00 & 0.00 \\ 
+  GO:0042832 & defense response to protozoan &  21 &   8 & 0.42 & 0.00 & 0.00 \\ 
+  GO:0009617 & response to bacterium & 253 &  22 & 5.02 & 0.00 & 0.00 \\ 
+  GO:0001562 & response to protozoan &  23 &   8 & 0.46 & 0.00 & 0.00 \\ 
+  GO:0002683 & negative regulation of immune system pro... & 261 &  22 & 5.18 & 0.00 & 0.00 \\ 
+  GO:0019882 & antigen processing and presentation &  72 &  12 & 1.43 & 0.00 & 0.00 \\ 
+  GO:0019884 & antigen processing and presentation of e... &  25 &   8 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0010033 & response to organic substance & 1390 &  58 & 27.56 & 0.00 & 0.00 \\ 
+  GO:0009605 & response to external stimulus & 1176 &  52 & 23.32 & 0.00 & 0.00 \\ 
+  GO:0043900 & regulation of multi-organism process & 281 &  22 & 5.57 & 0.00 & 0.00 \\ 
+  GO:0044403 & symbiosis, encompassing mutualism throug... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
+  GO:0044419 & interspecies interaction between organis... & 234 &  20 & 4.64 & 0.00 & 0.00 \\ 
+  GO:0060337 & type I interferon signaling pathway &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
+  GO:0071357 & cellular response to type I interferon &  22 &   7 & 0.44 & 0.00 & 0.00 \\ 
+  GO:0042742 & defense response to bacterium & 105 &  13 & 2.08 & 0.00 & 0.00 \\ 
+  GO:0044764 & multi-organism cellular process & 212 &  18 & 4.20 & 0.00 & 0.00 \\ 
+  GO:0035455 & response to interferon-alpha &  15 &   6 & 0.30 & 0.00 & 0.00 \\ 
+  GO:0001816 & cytokine production & 397 &  25 & 7.87 & 0.00 & 0.00 \\ 
+  GO:0043903 & regulation of symbiosis, encompassing mu... & 139 &  14 & 2.76 & 0.00 & 0.00 \\ 
+  GO:0002237 & response to molecule of bacterial origin & 162 &  15 & 3.21 & 0.00 & 0.00 \\ 
+  GO:0050792 & regulation of viral process & 122 &  13 & 2.42 & 0.00 & 0.00 \\ 
+  GO:0048525 & negative regulation of viral process &  69 &  10 & 1.37 & 0.00 & 0.00 \\ 
+  GO:0045089 & positive regulation of innate immune res... & 127 &  13 & 2.52 & 0.00 & 0.00 \\ 
+  GO:0051250 & negative regulation of lymphocyte activa... &  90 &  11 & 1.78 & 0.00 & 0.00 \\ 
+  GO:0002684 & positive regulation of immune system pro... & 494 &  27 & 9.79 & 0.00 & 0.00 \\ 
+  GO:0002478 & antigen processing and presentation of e... &  20 &   6 & 0.40 & 0.00 & 0.00 \\ 
+  GO:0045824 & negative regulation of innate immune res... &  31 &   7 & 0.61 & 0.00 & 0.00 \\ 
+  GO:0050778 & positive regulation of immune response & 298 &  20 & 5.91 & 0.00 & 0.00 \\ 
+  GO:0032606 & type I interferon production &  45 &   8 & 0.89 & 0.00 & 0.00 \\ 
+  GO:0080134 & regulation of response to stress & 830 &  37 & 16.46 & 0.00 & 0.00 \\ 
+  GO:0031349 & positive regulation of defense response & 180 &  15 & 3.57 & 0.00 & 0.00 \\ 
+  GO:0045071 & negative regulation of viral genome repl... &  33 &   7 & 0.65 & 0.00 & 0.00 \\ 
+  GO:0048002 & antigen processing and presentation of p... &  47 &   8 & 0.93 & 0.00 & 0.00 \\ 
+  GO:0002695 & negative regulation of leukocyte activat... & 100 &  11 & 1.98 & 0.00 & 0.00 \\ 
+  GO:0001817 & regulation of cytokine production & 344 &  21 & 6.82 & 0.00 & 0.00 \\ 
+  GO:0019221 & cytokine-mediated signaling pathway & 188 &  15 & 3.73 & 0.00 & 0.00 \\ 
+  GO:0032607 & interferon-alpha production &  14 &   5 & 0.28 & 0.00 & 0.00 \\ 
+  GO:1903901 & negative regulation of viral life cycle &  66 &   9 & 1.31 & 0.00 & 0.00 \\ 
+  GO:0098586 & cellular response to virus &  24 &   6 & 0.48 & 0.00 & 0.00 \\ 
+  GO:0071310 & cellular response to organic substance & 1039 &  42 & 20.60 & 0.00 & 0.00 \\ 
+  GO:0002477 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
+  GO:0002481 & antigen processing and presentation of e... &   3 &   3 & 0.06 & 0.00 & 0.00 \\ 
+  GO:0043901 & negative regulation of multi-organism pr... & 108 &  11 & 2.14 & 0.00 & 0.00 \\ 
+  GO:0050777 & negative regulation of immune response &  89 &  10 & 1.76 & 0.00 & 0.00 \\ 
+  GO:0050866 & negative regulation of cell activation & 109 &  11 & 2.16 & 0.00 & 0.00 \\ 
+  GO:0016032 & viral process & 200 &  15 & 3.97 & 0.00 & 0.00 \\ 
+  GO:0002483 & antigen processing and presentation of e... &   9 &   4 & 0.18 & 0.00 & 0.00 \\ 
+  GO:0002697 & regulation of immune effector process & 271 &  17 & 5.37 & 0.00 & 0.00 \\ 
+  GO:0019883 & antigen processing and presentation of e... &  10 &   4 & 0.20 & 0.00 & 0.00 \\ 
+  GO:0002428 & antigen processing and presentation of p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  GO:0045341 & MHC class I biosynthetic process &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  GO:0045343 & regulation of MHC class I biosynthetic p... &   4 &   3 & 0.08 & 0.00 & 0.00 \\ 
+  GO:0002221 & pattern recognition receptor signaling p... &  81 &   9 & 1.61 & 0.00 & 0.00 \\ 
+  GO:0042221 & response to chemical & 1868 &  61 & 37.04 & 0.00 & 0.00 \\ 
+  GO:0032496 & response to lipopolysaccharide & 148 &  12 & 2.93 & 0.00 & 0.00 \\ 
+  GO:0050830 & defense response to Gram-positive bacter... &  48 &   7 & 0.95 & 0.00 & 0.00 \\ 
+  GO:0002758 & innate immune response-activating signal... &  84 &   9 & 1.67 & 0.00 & 0.00 \\ 
+  GO:0002449 & lymphocyte mediated immunity & 150 &  12 & 2.97 & 0.00 & 0.00 \\ 
+  GO:0050868 & negative regulation of T cell activation &  66 &   8 & 1.31 & 0.00 & 0.00 \\ 
+  GO:0048584 & positive regulation of response to stimu... & 1130 &  42 & 22.41 & 0.00 & 0.00 \\ 
+  GO:0070887 & cellular response to chemical stimulus & 1330 &  47 & 26.37 & 0.00 & 0.00 \\ 
+  GO:0051249 & regulation of lymphocyte activation & 259 &  16 & 5.14 & 0.00 & 0.00 \\ 
+  GO:0001819 & positive regulation of cytokine producti... & 232 &  15 & 4.60 & 0.00 & 0.00 \\ 
+  GO:0002716 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
+  GO:0045953 & negative regulation of natural killer ce... &  12 &   4 & 0.24 & 0.00 & 0.00 \\ 
+  GO:0050896 & response to stimulus & 4441 & 117 & 88.06 & 0.00 & 0.00 \\ 
+  GO:1903038 & negative regulation of leukocyte cell-ce... &  71 &   8 & 1.41 & 0.00 & 0.00 \\ 
+  GO:0002218 & activation of innate immune response &  91 &   9 & 1.80 & 0.00 & 0.00 \\ 
+  GO:0031341 & regulation of cell killing &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
+  GO:0045069 & regulation of viral genome replication &  53 &   7 & 1.05 & 0.00 & 0.00 \\ 
+  GO:1903900 & regulation of viral life cycle & 113 &  10 & 2.24 & 0.00 & 0.00 \\ 
+  GO:0002753 & cytoplasmic pattern recognition receptor... &  24 &   5 & 0.48 & 0.00 & 0.00 \\ 
+  GO:0032647 & regulation of interferon-alpha productio... &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
+  GO:0044406 & adhesion of symbiont to host &  13 &   4 & 0.26 & 0.00 & 0.00 \\ 
+  GO:0001906 & cell killing &  73 &   8 & 1.45 & 0.00 & 0.00 \\ 
+  GO:0019058 & viral life cycle & 163 &  12 & 3.23 & 0.00 & 0.00 \\ 
+  GO:0042267 & natural killer cell mediated cytotoxicit... &  39 &   6 & 0.77 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-14 additional_files/tex/cluster2_BP_Mm_7dpiUp.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

@@
-% latex table generated in R 3.2.2 by xtable 1.8-2 package
-% Fri Apr  8 11:17:23 2016
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Thu Apr 14 14:59:44 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllrrrrr}
+\begin{tabular}{llrrrrr}
   \hline
- & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
+GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
   \hline
-1 & GO:0055114 & oxidation-reduction process & 650 &  28 & 9.76 & 0.00 & 0.00 \\ 
-  2 & GO:0044281 & small molecule metabolic process & 1261 &  42 & 18.94 & 0.00 & 0.00 \\ 
-  3 & GO:0006629 & lipid metabolic process & 796 &  29 & 11.95 & 0.00 & 0.00 \\ 
-  4 & GO:0044710 & single-organism metabolic process & 3313 &  76 & 49.76 & 0.00 & 0.00 \\ 
-  5 & GO:0006805 & xenobiotic metabolic process &  42 &   6 & 0.63 & 0.00 & 0.00 \\ 
-  6 & GO:0071466 & cellular response to xenobiotic stimulus &  46 &   6 & 0.69 & 0.00 & 0.00 \\ 
-  7 & GO:0009410 & response to xenobiotic stimulus &  51 &   6 & 0.77 & 0.00 & 0.00 \\ 
-  8 & GO:0006140 & regulation of nucleotide metabolic proce... & 103 &   8 & 1.55 & 0.00 & 0.00 \\ 
-  9 & GO:1901387 & positive regulation of voltage-gated cal... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
-  10 & GO:0009117 & nucleotide metabolic process & 400 &  16 & 6.01 & 0.00 & 0.00 \\ 
-  11 & GO:0017144 & drug metabolic process &  23 &   4 & 0.35 & 0.00 & 0.00 \\ 
-  12 & GO:0019369 & arachidonic acid metabolic process &  24 &   4 & 0.36 & 0.00 & 0.00 \\ 
-  13 & GO:0006753 & nucleoside phosphate metabolic process & 408 &  16 & 6.13 & 0.00 & 0.00 \\ 
-  14 & GO:0008202 & steroid metabolic process & 169 &   9 & 2.54 & 0.00 & 0.01 \\ 
-  15 & GO:0055086 & nucleobase-containing small molecule met... & 443 &  16 & 6.65 & 0.00 & 0.01 \\ 
-  16 & GO:0015701 & bicarbonate transport &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
-  17 & GO:0042738 & exogenous drug catabolic process &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
-  18 & GO:0006631 & fatty acid metabolic process & 243 &  11 & 3.65 & 0.00 & 0.01 \\ 
-  19 & GO:0051289 & protein homotetramerization &  53 &   5 & 0.80 & 0.00 & 0.01 \\ 
-  20 & GO:0009992 & cellular water homeostasis &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
-  21 & GO:0032782 & bile acid secretion &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
-  22 & GO:0042737 & drug catabolic process &  15 &   3 & 0.23 & 0.00 & 0.01 \\ 
-  23 & GO:0006690 & icosanoid metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
-  24 & GO:1901568 & fatty acid derivative metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
-  25 & GO:0015850 & organic hydroxy compound transport & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
-  26 & GO:0043583 & ear development & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
-  27 & GO:0051262 & protein tetramerization &  87 &   6 & 1.31 & 0.00 & 0.01 \\ 
-  28 & GO:0055085 & transmembrane transport & 563 &  18 & 8.46 & 0.00 & 0.01 \\ 
-  29 & GO:0045989 & positive regulation of striated muscle c... &   5 &   2 & 0.08 & 0.00 & 0.01 \\ 
-  30 & GO:0042445 & hormone metabolic process &  89 &   6 & 1.34 & 0.00 & 0.01 \\ 
-  31 & GO:0051186 & cofactor metabolic process & 265 &  11 & 3.98 & 0.00 & 0.01 \\ 
-  32 & GO:0030104 & water homeostasis &  18 &   3 & 0.27 & 0.00 & 0.01 \\ 
-  33 & GO:0071333 & cellular response to glucose stimulus &  62 &   5 & 0.93 & 0.00 & 0.01 \\ 
-  34 & GO:0019637 & organophosphate metabolic process & 668 &  20 & 10.03 & 0.00 & 0.01 \\ 
-  35 & GO:0046903 & secretion & 625 &  19 & 9.39 & 0.00 & 0.01 \\ 
-  36 & GO:0071326 & cellular response to monosaccharide stim... &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
-  37 & GO:0071331 & cellular response to hexose stimulus &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
-  38 & GO:0019373 & epoxygenase P450 pathway &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
-  39 & GO:0019532 & oxalate transport &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
-  40 & GO:0048505 & regulation of timing of cell differentia... &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
-  41 & GO:0071322 & cellular response to carbohydrate stimul... &  68 &   5 & 1.02 & 0.00 & 0.01 \\ 
-  42 & GO:1900542 & regulation of purine nucleotide metaboli... &  98 &   6 & 1.47 & 0.00 & 0.01 \\ 
-  43 & GO:0050433 & regulation of catecholamine secretion &  21 &   3 & 0.32 & 0.00 & 0.01 \\ 
-  44 & GO:0015711 & organic anion transport & 207 &   9 & 3.11 & 0.00 & 0.01 \\ 
-  45 & GO:0016042 & lipid catabolic process & 172 &   8 & 2.58 & 0.00 & 0.01 \\ 
-  46 & GO:0003214 & cardiac left ventricle morphogenesis &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
-  47 & GO:0006833 & water transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
-  48 & GO:0032099 & negative regulation of appetite &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
-  49 & GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
-  50 & GO:0001678 & cellular glucose homeostasis &  72 &   5 & 1.08 & 0.00 & 0.01 \\ 
-  51 & GO:0032787 & monocarboxylic acid metabolic process & 377 &  13 & 5.66 & 0.00 & 0.01 \\ 
-  52 & GO:0007200 & phospholipase C-activating G-protein cou... &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
-  53 & GO:0050432 & catecholamine secretion &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
-  54 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   4 & 0.71 & 0.01 & 0.01 \\ 
-  55 & GO:0032105 & negative regulation of response to extra... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
-  56 & GO:0032108 & negative regulation of response to nutri... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
-  57 & GO:0055082 & cellular chemical homeostasis & 342 &  12 & 5.14 & 0.01 & 0.01 \\ 
-  58 & GO:0032612 & interleukin-1 production &  48 &   4 & 0.72 & 0.01 & 0.01 \\ 
-  59 & GO:0006820 & anion transport & 302 &  11 & 4.54 & 0.01 & 0.01 \\ 
-  60 & GO:0032096 & negative regulation of response to food &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
-  61 & GO:0033604 & negative regulation of catecholamine sec... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
-  62 & GO:0040034 & regulation of development, heterochronic &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
-  63 & GO:0060081 & membrane hyperpolarization &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
-  64 & GO:0060979 & vasculogenesis involved in coronary vasc... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
-  65 & GO:0006730 & one-carbon metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
-  66 & GO:1901661 & quinone metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
-  67 & GO:1901135 & carbohydrate derivative metabolic proces... & 677 &  19 & 10.17 & 0.01 & 0.01 \\ 
-  68 & GO:0034220 & ion transmembrane transport & 393 &  13 & 5.90 & 0.01 & 0.01 \\ 
-  69 & GO:0030004 & cellular monovalent inorganic cation hom... &  50 &   4 & 0.75 & 0.01 & 0.01 \\ 
-  70 & GO:0009187 & cyclic nucleotide metabolic process &  79 &   5 & 1.19 & 0.01 & 0.01 \\ 
-  71 & GO:0010817 & regulation of hormone levels & 267 &  10 & 4.01 & 0.01 & 0.01 \\ 
-  72 & GO:0008272 & sulfate transport &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  73 & GO:0032148 & activation of protein kinase B activity &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  74 & GO:0071875 & adrenergic receptor signaling pathway &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
-  75 & GO:0052652 & cyclic purine nucleotide metabolic proce... &  52 &   4 & 0.78 & 0.01 & 0.01 \\ 
-  76 & GO:0072521 & purine-containing compound metabolic pro... & 315 &  11 & 4.73 & 0.01 & 0.01 \\ 
-  77 & GO:0009190 & cyclic nucleotide biosynthetic process &  53 &   4 & 0.80 & 0.01 & 0.01 \\ 
-  78 & GO:0051937 & catecholamine transport &  28 &   3 & 0.42 & 0.01 & 0.01 \\ 
-  79 & GO:0019318 & hexose metabolic process & 154 &   7 & 2.31 & 0.01 & 0.01 \\ 
-  80 & GO:0045981 & positive regulation of nucleotide metabo... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
-  81 & GO:1900544 & positive regulation of purine nucleotide... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
-  82 & GO:0045980 & negative regulation of nucleotide metabo... &  29 &   3 & 0.44 & 0.01 & 0.01 \\ 
-  83 & GO:0002275 & myeloid cell activation involved in immu... &  55 &   4 & 0.83 & 0.01 & 0.01 \\ 
-  84 & GO:0032098 & regulation of appetite &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
-  85 & GO:0051923 & sulfation &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
-  86 & GO:0042493 & response to drug & 158 &   7 & 2.37 & 0.01 & 0.01 \\ 
+GO:0055114 & oxidation-reduction process & 650 &  28 & 9.76 & 0.00 & 0.00 \\ 
+  GO:0044281 & small molecule metabolic process & 1261 &  42 & 18.94 & 0.00 & 0.00 \\ 
+  GO:0006629 & lipid metabolic process & 796 &  29 & 11.95 & 0.00 & 0.00 \\ 
+  GO:0044710 & single-organism metabolic process & 3313 &  76 & 49.76 & 0.00 & 0.00 \\ 
+  GO:0006805 & xenobiotic metabolic process &  42 &   6 & 0.63 & 0.00 & 0.00 \\ 
+  GO:0071466 & cellular response to xenobiotic stimulus &  46 &   6 & 0.69 & 0.00 & 0.00 \\ 
+  GO:0009410 & response to xenobiotic stimulus &  51 &   6 & 0.77 & 0.00 & 0.00 \\ 
+  GO:0006140 & regulation of nucleotide metabolic proce... & 103 &   8 & 1.55 & 0.00 & 0.00 \\ 
+  GO:1901387 & positive regulation of voltage-gated cal... &   2 &   2 & 0.03 & 0.00 & 0.00 \\ 
+  GO:0009117 & nucleotide metabolic process & 400 &  16 & 6.01 & 0.00 & 0.00 \\ 
+  GO:0017144 & drug metabolic process &  23 &   4 & 0.35 & 0.00 & 0.00 \\ 
+  GO:0019369 & arachidonic acid metabolic process &  24 &   4 & 0.36 & 0.00 & 0.00 \\ 
+  GO:0006753 & nucleoside phosphate metabolic process & 408 &  16 & 6.13 & 0.00 & 0.00 \\ 
+  GO:0008202 & steroid metabolic process & 169 &   9 & 2.54 & 0.00 & 0.01 \\ 
+  GO:0055086 & nucleobase-containing small molecule met... & 443 &  16 & 6.65 & 0.00 & 0.01 \\ 
+  GO:0015701 & bicarbonate transport &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
+  GO:0042738 & exogenous drug catabolic process &  14 &   3 & 0.21 & 0.00 & 0.01 \\ 
+  GO:0006631 & fatty acid metabolic process & 243 &  11 & 3.65 & 0.00 & 0.01 \\ 
+  GO:0051289 & protein homotetramerization &  53 &   5 & 0.80 & 0.00 & 0.01 \\ 
+  GO:0009992 & cellular water homeostasis &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
+  GO:0032782 & bile acid secretion &   4 &   2 & 0.06 & 0.00 & 0.01 \\ 
+  GO:0042737 & drug catabolic process &  15 &   3 & 0.23 & 0.00 & 0.01 \\ 
+  GO:0006690 & icosanoid metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
+  GO:1901568 & fatty acid derivative metabolic process &  55 &   5 & 0.83 & 0.00 & 0.01 \\ 
+  GO:0015850 & organic hydroxy compound transport & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
+  GO:0043583 & ear development & 116 &   7 & 1.74 & 0.00 & 0.01 \\ 
+  GO:0051262 & protein tetramerization &  87 &   6 & 1.31 & 0.00 & 0.01 \\ 
+  GO:0055085 & transmembrane transport & 563 &  18 & 8.46 & 0.00 & 0.01 \\ 
+  GO:0045989 & positive regulation of striated muscle c... &   5 &   2 & 0.08 & 0.00 & 0.01 \\ 
+  GO:0042445 & hormone metabolic process &  89 &   6 & 1.34 & 0.00 & 0.01 \\ 
+  GO:0051186 & cofactor metabolic process & 265 &  11 & 3.98 & 0.00 & 0.01 \\ 
+  GO:0030104 & water homeostasis &  18 &   3 & 0.27 & 0.00 & 0.01 \\ 
+  GO:0071333 & cellular response to glucose stimulus &  62 &   5 & 0.93 & 0.00 & 0.01 \\ 
+  GO:0019637 & organophosphate metabolic process & 668 &  20 & 10.03 & 0.00 & 0.01 \\ 
+  GO:0046903 & secretion & 625 &  19 & 9.39 & 0.00 & 0.01 \\ 
+  GO:0071326 & cellular response to monosaccharide stim... &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
+  GO:0071331 & cellular response to hexose stimulus &  64 &   5 & 0.96 & 0.00 & 0.01 \\ 
+  GO:0019373 & epoxygenase P450 pathway &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  GO:0019532 & oxalate transport &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  GO:0048505 & regulation of timing of cell differentia... &   6 &   2 & 0.09 & 0.00 & 0.01 \\ 
+  GO:0071322 & cellular response to carbohydrate stimul... &  68 &   5 & 1.02 & 0.00 & 0.01 \\ 
+  GO:1900542 & regulation of purine nucleotide metaboli... &  98 &   6 & 1.47 & 0.00 & 0.01 \\ 
+  GO:0050433 & regulation of catecholamine secretion &  21 &   3 & 0.32 & 0.00 & 0.01 \\ 
+  GO:0015711 & organic anion transport & 207 &   9 & 3.11 & 0.00 & 0.01 \\ 
+  GO:0016042 & lipid catabolic process & 172 &   8 & 2.58 & 0.00 & 0.01 \\ 
+  GO:0003214 & cardiac left ventricle morphogenesis &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  GO:0006833 & water transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  GO:0032099 & negative regulation of appetite &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  GO:1902358 & sulfate transmembrane transport &   7 &   2 & 0.11 & 0.00 & 0.01 \\ 
+  GO:0001678 & cellular glucose homeostasis &  72 &   5 & 1.08 & 0.00 & 0.01 \\ 
+  GO:0032787 & monocarboxylic acid metabolic process & 377 &  13 & 5.66 & 0.00 & 0.01 \\ 
+  GO:0007200 & phospholipase C-activating G-protein cou... &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
+  GO:0050432 & catecholamine secretion &  23 &   3 & 0.35 & 0.00 & 0.01 \\ 
+  GO:0001676 & long-chain fatty acid metabolic process &  47 &   4 & 0.71 & 0.01 & 0.01 \\ 
+  GO:0032105 & negative regulation of response to extra... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
+  GO:0032108 & negative regulation of response to nutri... &  24 &   3 & 0.36 & 0.01 & 0.01 \\ 
+  GO:0055082 & cellular chemical homeostasis & 342 &  12 & 5.14 & 0.01 & 0.01 \\ 
+  GO:0032612 & interleukin-1 production &  48 &   4 & 0.72 & 0.01 & 0.01 \\ 
+  GO:0006820 & anion transport & 302 &  11 & 4.54 & 0.01 & 0.01 \\ 
+  GO:0032096 & negative regulation of response to food &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  GO:0033604 & negative regulation of catecholamine sec... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  GO:0040034 & regulation of development, heterochronic &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  GO:0060081 & membrane hyperpolarization &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  GO:0060979 & vasculogenesis involved in coronary vasc... &   8 &   2 & 0.12 & 0.01 & 0.01 \\ 
+  GO:0006730 & one-carbon metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
+  GO:1901661 & quinone metabolic process &  25 &   3 & 0.38 & 0.01 & 0.01 \\ 
+  GO:1901135 & carbohydrate derivative metabolic proces... & 677 &  19 & 10.17 & 0.01 & 0.01 \\ 
+  GO:0034220 & ion transmembrane transport & 393 &  13 & 5.90 & 0.01 & 0.01 \\ 
+  GO:0030004 & cellular monovalent inorganic cation hom... &  50 &   4 & 0.75 & 0.01 & 0.01 \\ 
+  GO:0009187 & cyclic nucleotide metabolic process &  79 &   5 & 1.19 & 0.01 & 0.01 \\ 
+  GO:0010817 & regulation of hormone levels & 267 &  10 & 4.01 & 0.01 & 0.01 \\ 
+  GO:0008272 & sulfate transport &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  GO:0032148 & activation of protein kinase B activity &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  GO:0071875 & adrenergic receptor signaling pathway &   9 &   2 & 0.14 & 0.01 & 0.01 \\ 
+  GO:0052652 & cyclic purine nucleotide metabolic proce... &  52 &   4 & 0.78 & 0.01 & 0.01 \\ 
+  GO:0072521 & purine-containing compound metabolic pro... & 315 &  11 & 4.73 & 0.01 & 0.01 \\ 
+  GO:0009190 & cyclic nucleotide biosynthetic process &  53 &   4 & 0.80 & 0.01 & 0.01 \\ 
+  GO:0051937 & catecholamine transport &  28 &   3 & 0.42 & 0.01 & 0.01 \\ 
+  GO:0019318 & hexose metabolic process & 154 &   7 & 2.31 & 0.01 & 0.01 \\ 
+  GO:0045981 & positive regulation of nucleotide metabo... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
+  GO:1900544 & positive regulation of purine nucleotide... &  54 &   4 & 0.81 & 0.01 & 0.01 \\ 
+  GO:0045980 & negative regulation of nucleotide metabo... &  29 &   3 & 0.44 & 0.01 & 0.01 \\ 
+  GO:0002275 & myeloid cell activation involved in immu... &  55 &   4 & 0.83 & 0.01 & 0.01 \\ 
+  GO:0032098 & regulation of appetite &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
+  GO:0051923 & sulfation &  10 &   2 & 0.15 & 0.01 & 0.01 \\ 
+  GO:0042493 & response to drug & 158 &   7 & 2.37 & 0.01 & 0.01 \\ 
    \hline

### 2016-04-14 additional_files/tex/cluster3_BP_Mm_7dpiDown.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

@@
-% latex table generated in R 3.2.2 by xtable 1.8-2 package
-% Fri Apr  8 11:18:59 2016
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Thu Apr 14 15:01:22 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllrrrrr}
+\begin{tabular}{llrrrrr}
   \hline
- & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
+GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
   \hline
-1 & GO:0006805 & xenobiotic metabolic process &  42 &  18 & 1.51 & 0.00 & 0.00 \\ 
-  2 & GO:0006082 & organic acid metabolic process & 661 &  68 & 23.81 & 0.00 & 0.00 \\ 
-  3 & GO:0044281 & small molecule metabolic process & 1261 & 101 & 45.43 & 0.00 & 0.00 \\ 
-  4 & GO:0032787 & monocarboxylic acid metabolic process & 377 &  49 & 13.58 & 0.00 & 0.00 \\ 
-  5 & GO:0071466 & cellular response to xenobiotic stimulus &  46 &  18 & 1.66 & 0.00 & 0.00 \\ 
-  6 & GO:0009813 & flavonoid biosynthetic process &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
-  7 & GO:0052696 & flavonoid glucuronidation &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
-  8 & GO:0043436 & oxoacid metabolic process & 652 &  66 & 23.49 & 0.00 & 0.00 \\ 
-  9 & GO:0019752 & carboxylic acid metabolic process & 612 &  63 & 22.05 & 0.00 & 0.00 \\ 
-  10 & GO:0009812 & flavonoid metabolic process &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
-  11 & GO:0052695 & cellular glucuronidation &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
-  12 & GO:0009410 & response to xenobiotic stimulus &  51 &  18 & 1.84 & 0.00 & 0.00 \\ 
-  13 & GO:0006063 & uronic acid metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
-  14 & GO:0019585 & glucuronate metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
-  15 & GO:0052697 & xenobiotic glucuronidation &   8 &   8 & 0.29 & 0.00 & 0.00 \\ 
-  16 & GO:0006629 & lipid metabolic process & 796 &  66 & 28.68 & 0.00 & 0.00 \\ 
-  17 & GO:0009719 & response to endogenous stimulus & 729 &  60 & 26.26 & 0.00 & 0.00 \\ 
-  18 & GO:0071495 & cellular response to endogenous stimulus & 561 &  50 & 20.21 & 0.00 & 0.00 \\ 
-  19 & GO:0009725 & response to hormone & 385 &  37 & 13.87 & 0.00 & 0.00 \\ 
-  20 & GO:0006631 & fatty acid metabolic process & 243 &  28 & 8.75 & 0.00 & 0.00 \\ 
-  21 & GO:0044255 & cellular lipid metabolic process & 604 &  49 & 21.76 & 0.00 & 0.00 \\ 
-  22 & GO:0055114 & oxidation-reduction process & 650 &  50 & 23.42 & 0.00 & 0.00 \\ 
-  23 & GO:0042445 & hormone metabolic process &  89 &  15 & 3.21 & 0.00 & 0.00 \\ 
-  24 & GO:0032870 & cellular response to hormone stimulus & 278 &  28 & 10.02 & 0.00 & 0.00 \\ 
-  25 & GO:0044710 & single-organism metabolic process & 3313 & 163 & 119.36 & 0.00 & 0.00 \\ 
-  26 & GO:0034754 & cellular hormone metabolic process &  43 &  10 & 1.55 & 0.00 & 0.00 \\ 
-  27 & GO:0005996 & monosaccharide metabolic process & 177 &  20 & 6.38 & 0.00 & 0.00 \\ 
-  28 & GO:0042178 & xenobiotic catabolic process &   9 &   5 & 0.32 & 0.00 & 0.00 \\ 
-  29 & GO:0042537 & benzene-containing compound metabolic pr... &  15 &   6 & 0.54 & 0.00 & 0.00 \\ 
-  30 & GO:0006790 & sulfur compound metabolic process & 200 &  21 & 7.21 & 0.00 & 0.00 \\ 
-  31 & GO:0019395 & fatty acid oxidation &  66 &  11 & 2.38 & 0.00 & 0.00 \\ 
-  32 & GO:0034440 & lipid oxidation &  67 &  11 & 2.41 & 0.00 & 0.00 \\ 
-  33 & GO:0044699 & single-organism process & 7965 & 325 & 286.96 & 0.00 & 0.00 \\ 
-  34 & GO:0042221 & response to chemical & 1868 &  99 & 67.30 & 0.00 & 0.00 \\ 
-  35 & GO:1901615 & organic hydroxy compound metabolic proce... & 299 &  26 & 10.77 & 0.00 & 0.00 \\ 
-  36 & GO:0010817 & regulation of hormone levels & 267 &  24 & 9.62 & 0.00 & 0.00 \\ 
-  37 & GO:0016042 & lipid catabolic process & 172 &  18 & 6.20 & 0.00 & 0.00 \\ 
-  38 & GO:0006811 & ion transport & 785 &  50 & 28.28 & 0.00 & 0.00 \\ 
-  39 & GO:0022600 & digestive system process &  51 &   9 & 1.84 & 0.00 & 0.00 \\ 
-  40 & GO:0044242 & cellular lipid catabolic process & 117 &  14 & 4.22 & 0.00 & 0.00 \\ 
-  41 & GO:0042738 & exogenous drug catabolic process &  14 &   5 & 0.50 & 0.00 & 0.00 \\ 
-  42 & GO:0051386 & regulation of neurotrophin TRK receptor ... &   8 &   4 & 0.29 & 0.00 & 0.00 \\ 
-  43 & GO:0044344 & cellular response to fibroblast growth f... &  42 &   8 & 1.51 & 0.00 & 0.00 \\ 
-  44 & GO:0008202 & steroid metabolic process & 169 &  17 & 6.09 & 0.00 & 0.00 \\ 
-  45 & GO:0017144 & drug metabolic process &  23 &   6 & 0.83 & 0.00 & 0.00 \\ 
-  46 & GO:0042737 & drug catabolic process &  15 &   5 & 0.54 & 0.00 & 0.00 \\ 
-  47 & GO:0005975 & carbohydrate metabolic process & 481 &  34 & 17.33 & 0.00 & 0.00 \\ 
-  48 & GO:0072329 & monocarboxylic acid catabolic process &  68 &  10 & 2.45 & 0.00 & 0.00 \\ 
-  49 & GO:0071774 & response to fibroblast growth factor &  44 &   8 & 1.59 & 0.00 & 0.00 \\ 
-  50 & GO:0007586 & digestion &  56 &   9 & 2.02 & 0.00 & 0.00 \\ 
-  51 & GO:0072001 & renal system development & 141 &  15 & 5.08 & 0.00 & 0.00 \\ 
-  52 & GO:0044539 & long-chain fatty acid import &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
-  53 & GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
-  54 & GO:0042493 & response to drug & 158 &  16 & 5.69 & 0.00 & 0.00 \\ 
-  55 & GO:0006672 & ceramide metabolic process &  58 &   9 & 2.09 & 0.00 & 0.00 \\ 
-  56 & GO:0008543 & fibroblast growth factor receptor signal... &  35 &   7 & 1.26 & 0.00 & 0.00 \\ 
-  57 & GO:0044723 & single-organism carbohydrate metabolic p... & 435 &  31 & 15.67 & 0.00 & 0.00 \\ 
-  58 & GO:0007267 & cell-cell signaling & 456 &  32 & 16.43 & 0.00 & 0.00 \\ 
-  59 & GO:0070887 & cellular response to chemical stimulus & 1330 &  72 & 47.92 & 0.00 & 0.00 \\ 
-  60 & GO:0001822 & kidney development & 131 &  14 & 4.72 & 0.00 & 0.00 \\ 
-  61 & GO:0006040 & amino sugar metabolic process &  26 &   6 & 0.94 & 0.00 & 0.00 \\ 
-  62 & GO:0034762 & regulation of transmembrane transport & 180 &  17 & 6.48 & 0.00 & 0.00 \\ 
-  63 & GO:0042572 & retinol metabolic process &  10 &   4 & 0.36 & 0.00 & 0.00 \\ 
-  64 & GO:0043269 & regulation of ion transport & 306 &  24 & 11.02 & 0.00 & 0.00 \\ 
-  65 & GO:0009062 & fatty acid catabolic process &  61 &   9 & 2.20 & 0.00 & 0.00 \\ 
-  66 & GO:0044282 & small molecule catabolic process & 183 &  17 & 6.59 & 0.00 & 0.00 \\ 
-  67 & GO:0001655 & urogenital system development & 168 &  16 & 6.05 & 0.00 & 0.00 \\ 
-  68 & GO:0022603 & regulation of anatomical structure morph... & 572 &  37 & 20.61 & 0.00 & 0.00 \\ 
-  69 & GO:0044763 & single-organism cellular process & 7155 & 291 & 257.78 & 0.00 & 0.00 \\ 
-  70 & GO:0006637 & acyl-CoA metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
-  71 & GO:0035383 & thioester metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
-  72 & GO:0006732 & coenzyme metabolic process & 227 &  19 & 8.18 & 0.00 & 0.00 \\ 
-  73 & GO:0016054 & organic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
-  74 & GO:0046395 & carboxylic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
-  75 & GO:0015698 & inorganic anion transport &  67 &   9 & 2.41 & 0.00 & 0.00 \\ 
-  76 & GO:0007169 & transmembrane receptor protein tyrosine ... & 324 &  24 & 11.67 & 0.00 & 0.00 \\ 
-  77 & GO:0009308 & amine metabolic process &  82 &  10 & 2.95 & 0.00 & 0.00 \\ 
-  78 & GO:0048011 & neurotrophin TRK receptor signaling path... &  22 &   5 & 0.79 & 0.00 & 0.00 \\ 
-  79 & GO:0030258 & lipid modification & 134 &  13 & 4.83 & 0.00 & 0.00 \\ 
-  80 & GO:0034765 & regulation of ion transmembrane transpor... & 168 &  15 & 6.05 & 0.00 & 0.00 \\ 
-  81 & GO:0055085 & transmembrane transport & 563 &  35 & 20.28 & 0.00 & 0.00 \\ 
-  82 & GO:0006576 & cellular biogenic amine metabolic proces... &  59 &   8 & 2.13 & 0.00 & 0.00 \\ 
-  83 & GO:0032941 & secretion by tissue &  46 &   7 & 1.66 & 0.00 & 0.00 \\ 
-  84 & GO:0043537 & negative regulation of blood vessel endo... &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
-  85 & GO:0060602 & branch elongation of an epithelium &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
-  86 & GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  87 & GO:0046168 & glycerol-3-phosphate catabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  88 & GO:0061031 & endodermal digestive tract morphogenesis &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  89 & GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  90 & GO:0097623 & potassium ion export across plasma membr... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
-  91 & GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 1.69 & 0.00 & 0.00 \\ 
-  92 & GO:0007268 & synaptic transmission & 245 &  19 & 8.83 & 0.00 & 0.00 \\ 
-  93 & GO:0071310 & cellular response to organic substance & 1039 &  56 & 37.43 & 0.00 & 0.00 \\ 
-  94 & GO:0019755 & one-carbon compound transport &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
-  95 & GO:0060736 & prostate gland growth &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
-  96 & GO:0090331 & negative regulation of platelet aggregat... &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
-  97 & GO:0008610 & lipid biosynthetic process & 383 &  26 & 13.80 & 0.00 & 0.00 \\ 
-  98 & GO:0009755 & hormone-mediated signaling pathway &  61 &   8 & 2.20 & 0.00 & 0.00 \\ 
-  99 & GO:0007167 & enzyme linked receptor protein signaling... & 487 &  31 & 17.55 & 0.00 & 0.00 \\ 
-  100 & GO:0016266 & O-glycan processing &  15 &   4 & 0.54 & 0.00 & 0.00 \\ 
+GO:0006805 & xenobiotic metabolic process &  42 &  18 & 1.51 & 0.00 & 0.00 \\ 
+  GO:0006082 & organic acid metabolic process & 661 &  68 & 23.81 & 0.00 & 0.00 \\ 
+  GO:0044281 & small molecule metabolic process & 1261 & 101 & 45.43 & 0.00 & 0.00 \\ 
+  GO:0032787 & monocarboxylic acid metabolic process & 377 &  49 & 13.58 & 0.00 & 0.00 \\ 
+  GO:0071466 & cellular response to xenobiotic stimulus &  46 &  18 & 1.66 & 0.00 & 0.00 \\ 
+  GO:0009813 & flavonoid biosynthetic process &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
+  GO:0052696 & flavonoid glucuronidation &  13 &  11 & 0.47 & 0.00 & 0.00 \\ 
+  GO:0043436 & oxoacid metabolic process & 652 &  66 & 23.49 & 0.00 & 0.00 \\ 
+  GO:0019752 & carboxylic acid metabolic process & 612 &  63 & 22.05 & 0.00 & 0.00 \\ 
+  GO:0009812 & flavonoid metabolic process &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0052695 & cellular glucuronidation &  14 &  11 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0009410 & response to xenobiotic stimulus &  51 &  18 & 1.84 & 0.00 & 0.00 \\ 
+  GO:0006063 & uronic acid metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
+  GO:0019585 & glucuronate metabolic process &  15 &  11 & 0.54 & 0.00 & 0.00 \\ 
+  GO:0052697 & xenobiotic glucuronidation &   8 &   8 & 0.29 & 0.00 & 0.00 \\ 
+  GO:0006629 & lipid metabolic process & 796 &  66 & 28.68 & 0.00 & 0.00 \\ 
+  GO:0009719 & response to endogenous stimulus & 729 &  60 & 26.26 & 0.00 & 0.00 \\ 
+  GO:0071495 & cellular response to endogenous stimulus & 561 &  50 & 20.21 & 0.00 & 0.00 \\ 
+  GO:0009725 & response to hormone & 385 &  37 & 13.87 & 0.00 & 0.00 \\ 
+  GO:0006631 & fatty acid metabolic process & 243 &  28 & 8.75 & 0.00 & 0.00 \\ 
+  GO:0044255 & cellular lipid metabolic process & 604 &  49 & 21.76 & 0.00 & 0.00 \\ 
+  GO:0055114 & oxidation-reduction process & 650 &  50 & 23.42 & 0.00 & 0.00 \\ 
+  GO:0042445 & hormone metabolic process &  89 &  15 & 3.21 & 0.00 & 0.00 \\ 
+  GO:0032870 & cellular response to hormone stimulus & 278 &  28 & 10.02 & 0.00 & 0.00 \\ 
+  GO:0044710 & single-organism metabolic process & 3313 & 163 & 119.36 & 0.00 & 0.00 \\ 
+  GO:0034754 & cellular hormone metabolic process &  43 &  10 & 1.55 & 0.00 & 0.00 \\ 
+  GO:0005996 & monosaccharide metabolic process & 177 &  20 & 6.38 & 0.00 & 0.00 \\ 
+  GO:0042178 & xenobiotic catabolic process &   9 &   5 & 0.32 & 0.00 & 0.00 \\ 
+  GO:0042537 & benzene-containing compound metabolic pr... &  15 &   6 & 0.54 & 0.00 & 0.00 \\ 
+  GO:0006790 & sulfur compound metabolic process & 200 &  21 & 7.21 & 0.00 & 0.00 \\ 
+  GO:0019395 & fatty acid oxidation &  66 &  11 & 2.38 & 0.00 & 0.00 \\ 
+  GO:0034440 & lipid oxidation &  67 &  11 & 2.41 & 0.00 & 0.00 \\ 
+  GO:0044699 & single-organism process & 7965 & 325 & 286.96 & 0.00 & 0.00 \\ 
+  GO:0042221 & response to chemical & 1868 &  99 & 67.30 & 0.00 & 0.00 \\ 
+  GO:1901615 & organic hydroxy compound metabolic proce... & 299 &  26 & 10.77 & 0.00 & 0.00 \\ 
+  GO:0010817 & regulation of hormone levels & 267 &  24 & 9.62 & 0.00 & 0.00 \\ 
+  GO:0016042 & lipid catabolic process & 172 &  18 & 6.20 & 0.00 & 0.00 \\ 
+  GO:0006811 & ion transport & 785 &  50 & 28.28 & 0.00 & 0.00 \\ 
+  GO:0022600 & digestive system process &  51 &   9 & 1.84 & 0.00 & 0.00 \\ 
+  GO:0044242 & cellular lipid catabolic process & 117 &  14 & 4.22 & 0.00 & 0.00 \\ 
+  GO:0042738 & exogenous drug catabolic process &  14 &   5 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0051386 & regulation of neurotrophin TRK receptor ... &   8 &   4 & 0.29 & 0.00 & 0.00 \\ 
+  GO:0044344 & cellular response to fibroblast growth f... &  42 &   8 & 1.51 & 0.00 & 0.00 \\ 
+  GO:0008202 & steroid metabolic process & 169 &  17 & 6.09 & 0.00 & 0.00 \\ 
+  GO:0017144 & drug metabolic process &  23 &   6 & 0.83 & 0.00 & 0.00 \\ 
+  GO:0042737 & drug catabolic process &  15 &   5 & 0.54 & 0.00 & 0.00 \\ 
+  GO:0005975 & carbohydrate metabolic process & 481 &  34 & 17.33 & 0.00 & 0.00 \\ 
+  GO:0072329 & monocarboxylic acid catabolic process &  68 &  10 & 2.45 & 0.00 & 0.00 \\ 
+  GO:0071774 & response to fibroblast growth factor &  44 &   8 & 1.59 & 0.00 & 0.00 \\ 
+  GO:0007586 & digestion &  56 &   9 & 2.02 & 0.00 & 0.00 \\ 
+  GO:0072001 & renal system development & 141 &  15 & 5.08 & 0.00 & 0.00 \\ 
+  GO:0044539 & long-chain fatty acid import &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
+  GO:0051387 & negative regulation of neurotrophin TRK ... &   4 &   3 & 0.14 & 0.00 & 0.00 \\ 
+  GO:0042493 & response to drug & 158 &  16 & 5.69 & 0.00 & 0.00 \\ 
+  GO:0006672 & ceramide metabolic process &  58 &   9 & 2.09 & 0.00 & 0.00 \\ 
+  GO:0008543 & fibroblast growth factor receptor signal... &  35 &   7 & 1.26 & 0.00 & 0.00 \\ 
+  GO:0044723 & single-organism carbohydrate metabolic p... & 435 &  31 & 15.67 & 0.00 & 0.00 \\ 
+  GO:0007267 & cell-cell signaling & 456 &  32 & 16.43 & 0.00 & 0.00 \\ 
+  GO:0070887 & cellular response to chemical stimulus & 1330 &  72 & 47.92 & 0.00 & 0.00 \\ 
+  GO:0001822 & kidney development & 131 &  14 & 4.72 & 0.00 & 0.00 \\ 
+  GO:0006040 & amino sugar metabolic process &  26 &   6 & 0.94 & 0.00 & 0.00 \\ 
+  GO:0034762 & regulation of transmembrane transport & 180 &  17 & 6.48 & 0.00 & 0.00 \\ 
+  GO:0042572 & retinol metabolic process &  10 &   4 & 0.36 & 0.00 & 0.00 \\ 
+  GO:0043269 & regulation of ion transport & 306 &  24 & 11.02 & 0.00 & 0.00 \\ 
+  GO:0009062 & fatty acid catabolic process &  61 &   9 & 2.20 & 0.00 & 0.00 \\ 
+  GO:0044282 & small molecule catabolic process & 183 &  17 & 6.59 & 0.00 & 0.00 \\ 
+  GO:0001655 & urogenital system development & 168 &  16 & 6.05 & 0.00 & 0.00 \\ 
+  GO:0022603 & regulation of anatomical structure morph... & 572 &  37 & 20.61 & 0.00 & 0.00 \\ 
+  GO:0044763 & single-organism cellular process & 7155 & 291 & 257.78 & 0.00 & 0.00 \\ 
+  GO:0006637 & acyl-CoA metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
+  GO:0035383 & thioester metabolic process &  52 &   8 & 1.87 & 0.00 & 0.00 \\ 
+  GO:0006732 & coenzyme metabolic process & 227 &  19 & 8.18 & 0.00 & 0.00 \\ 
+  GO:0016054 & organic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
+  GO:0046395 & carboxylic acid catabolic process & 126 &  13 & 4.54 & 0.00 & 0.00 \\ 
+  GO:0015698 & inorganic anion transport &  67 &   9 & 2.41 & 0.00 & 0.00 \\ 
+  GO:0007169 & transmembrane receptor protein tyrosine ... & 324 &  24 & 11.67 & 0.00 & 0.00 \\ 
+  GO:0009308 & amine metabolic process &  82 &  10 & 2.95 & 0.00 & 0.00 \\ 
+  GO:0048011 & neurotrophin TRK receptor signaling path... &  22 &   5 & 0.79 & 0.00 & 0.00 \\ 
+  GO:0030258 & lipid modification & 134 &  13 & 4.83 & 0.00 & 0.00 \\ 
+  GO:0034765 & regulation of ion transmembrane transpor... & 168 &  15 & 6.05 & 0.00 & 0.00 \\ 
+  GO:0055085 & transmembrane transport & 563 &  35 & 20.28 & 0.00 & 0.00 \\ 
+  GO:0006576 & cellular biogenic amine metabolic proces... &  59 &   8 & 2.13 & 0.00 & 0.00 \\ 
+  GO:0032941 & secretion by tissue &  46 &   7 & 1.66 & 0.00 & 0.00 \\ 
+  GO:0043537 & negative regulation of blood vessel endo... &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0060602 & branch elongation of an epithelium &  14 &   4 & 0.50 & 0.00 & 0.00 \\ 
+  GO:0018879 & biphenyl metabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  GO:0046168 & glycerol-3-phosphate catabolic process &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  GO:0061031 & endodermal digestive tract morphogenesis &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  GO:0097089 & methyl-branched fatty acid metabolic pro... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  GO:0097623 & potassium ion export across plasma membr... &   2 &   2 & 0.07 & 0.00 & 0.00 \\ 
+  GO:0001676 & long-chain fatty acid metabolic process &  47 &   7 & 1.69 & 0.00 & 0.00 \\ 
+  GO:0007268 & synaptic transmission & 245 &  19 & 8.83 & 0.00 & 0.00 \\ 
+  GO:0071310 & cellular response to organic substance & 1039 &  56 & 37.43 & 0.00 & 0.00 \\ 
+  GO:0019755 & one-carbon compound transport &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  GO:0060736 & prostate gland growth &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  GO:0090331 & negative regulation of platelet aggregat... &   7 &   3 & 0.25 & 0.00 & 0.00 \\ 
+  GO:0008610 & lipid biosynthetic process & 383 &  26 & 13.80 & 0.00 & 0.00 \\ 
+  GO:0009755 & hormone-mediated signaling pathway &  61 &   8 & 2.20 & 0.00 & 0.00 \\ 
+  GO:0007167 & enzyme linked receptor protein signaling... & 487 &  31 & 17.55 & 0.00 & 0.00 \\ 
+  GO:0016266 & O-glycan processing &  15 &   4 & 0.54 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-14 additional_files/tex/cluster3_MF_Ef_OocSporDown.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:26 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0004312 & fatty acid synthase activity &   2 &   2 & 0.09 & 0.00 & 0.07 \\ 
  GO:0051920 & peroxiredoxin activity &   2 &   2 & 0.09 & 0.00 & 0.07 \\ 
  GO:0003824 & catalytic activity & 1318 &  74 & 58.58 & 0.00 & 0.07 \\ 
  GO:0016746 & transferase activity, transferring acyl ... &  46 &   7 & 2.04 & 0.00 & 0.08 \\ 
  GO:0019205 & nucleobase-containing compound kinase ac... &   8 &   3 & 0.36 & 0.00 & 0.08 \\ 
  GO:0016615 & malate dehydrogenase activity &   3 &   2 & 0.13 & 0.01 & 0.09 \\ 
  GO:0003746 & translation elongation factor activity &  10 &   3 & 0.44 & 0.01 & 0.09 \\ 
  GO:0008135 & translation factor activity, RNA binding &  30 &   5 & 1.33 & 0.01 & 0.09 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-04-14 additional_files/tex/cluster4_BP_Mm_allInfDown.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

@@
-% latex table generated in R 3.2.2 by xtable 1.8-2 package
-% Fri Apr  8 11:22:12 2016
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Thu Apr 14 15:04:34 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllrrrrr}
+\begin{tabular}{llrrrrr}
   \hline
- & GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
+GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
   \hline
-1 & GO:0002376 & immune system process & 1406 & 126 & 41.64 & 0.00 & 0.00 \\ 
-  2 & GO:0006955 & immune response & 713 &  85 & 21.12 & 0.00 & 0.00 \\ 
-  3 & GO:0050896 & response to stimulus & 4441 & 229 & 131.52 & 0.00 & 0.00 \\ 
-  4 & GO:0006954 & inflammatory response & 327 &  53 & 9.68 & 0.00 & 0.00 \\ 
-  5 & GO:0044700 & single organism signaling & 2975 & 176 & 88.11 & 0.00 & 0.00 \\ 
-  6 & GO:0006952 & defense response & 755 &  80 & 22.36 & 0.00 & 0.00 \\ 
-  7 & GO:0023052 & signaling & 2981 & 176 & 88.28 & 0.00 & 0.00 \\ 
-  8 & GO:0007165 & signal transduction & 2773 & 167 & 82.12 & 0.00 & 0.00 \\ 
-  9 & GO:0001775 & cell activation & 589 &  69 & 17.44 & 0.00 & 0.00 \\ 
-  10 & GO:0007154 & cell communication & 3137 & 178 & 92.90 & 0.00 & 0.00 \\ 
-  11 & GO:0022610 & biological adhesion & 800 &  79 & 23.69 & 0.00 & 0.00 \\ 
-  12 & GO:0002684 & positive regulation of immune system pro... & 494 &  61 & 14.63 & 0.00 & 0.00 \\ 
-  13 & GO:0007155 & cell adhesion & 791 &  78 & 23.43 & 0.00 & 0.00 \\ 
-  14 & GO:0002682 & regulation of immune system process & 774 &  76 & 22.92 & 0.00 & 0.00 \\ 
-  15 & GO:0001816 & cytokine production & 397 &  53 & 11.76 & 0.00 & 0.00 \\ 
-  16 & GO:0065007 & biological regulation & 6468 & 274 & 191.55 & 0.00 & 0.00 \\ 
-  17 & GO:0048518 & positive regulation of biological proces... & 3404 & 182 & 100.81 & 0.00 & 0.00 \\ 
-  18 & GO:0045321 & leukocyte activation & 515 &  60 & 15.25 & 0.00 & 0.00 \\ 
-  19 & GO:0040011 & locomotion & 903 &  79 & 26.74 & 0.00 & 0.00 \\ 
-  20 & GO:0098602 & single organism cell adhesion & 516 &  58 & 15.28 & 0.00 & 0.00 \\ 
-  21 & GO:0051239 & regulation of multicellular organismal p... & 1538 & 108 & 45.55 & 0.00 & 0.00 \\ 
-  22 & GO:0007159 & leukocyte cell-cell adhesion & 332 &  46 & 9.83 & 0.00 & 0.00 \\ 
-  23 & GO:0016337 & single organismal cell-cell adhesion & 474 &  55 & 14.04 & 0.00 & 0.00 \\ 
-  24 & GO:0051716 & cellular response to stimulus & 3741 & 189 & 110.79 & 0.00 & 0.00 \\ 
-  25 & GO:0098609 & cell-cell adhesion & 507 &  56 & 15.01 & 0.00 & 0.00 \\ 
-  26 & GO:0044707 & single-multicellular organism process & 3409 & 176 & 100.96 & 0.00 & 0.00 \\ 
-  27 & GO:0050789 & regulation of biological process & 6161 & 259 & 182.46 & 0.00 & 0.00 \\ 
-  28 & GO:0048870 & cell motility & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
-  29 & GO:0051674 & localization of cell & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
-  30 & GO:0032501 & multicellular organismal process & 3487 & 177 & 103.27 & 0.00 & 0.00 \\ 
-  31 & GO:0050794 & regulation of cellular process & 5853 & 249 & 173.34 & 0.00 & 0.00 \\ 
-  32 & GO:0016477 & cell migration & 740 &  67 & 21.92 & 0.00 & 0.00 \\ 
-  33 & GO:0001817 & regulation of cytokine production & 344 &  44 & 10.19 & 0.00 & 0.00 \\ 
-  34 & GO:0034109 & homotypic cell-cell adhesion & 361 &  45 & 10.69 & 0.00 & 0.00 \\ 
-  35 & GO:0048583 & regulation of response to stimulus & 2109 & 126 & 62.46 & 0.00 & 0.00 \\ 
-  36 & GO:0048584 & positive regulation of response to stimu... & 1130 &  85 & 33.47 & 0.00 & 0.00 \\ 
-  37 & GO:0050865 & regulation of cell activation & 319 &  42 & 9.45 & 0.00 & 0.00 \\ 
-  38 & GO:0048522 & positive regulation of cellular process & 3048 & 160 & 90.27 & 0.00 & 0.00 \\ 
-  39 & GO:0046649 & lymphocyte activation & 435 &  49 & 12.88 & 0.00 & 0.00 \\ 
-  40 & GO:0035556 & intracellular signal transduction & 1570 & 103 & 46.50 & 0.00 & 0.00 \\ 
-  41 & GO:0002694 & regulation of leukocyte activation & 301 &  40 & 8.91 & 0.00 & 0.00 \\ 
-  42 & GO:0070486 & leukocyte aggregation & 314 &  40 & 9.30 & 0.00 & 0.00 \\ 
-  43 & GO:1902531 & regulation of intracellular signal trans... & 991 &  76 & 29.35 & 0.00 & 0.00 \\ 
-  44 & GO:0050900 & leukocyte migration & 164 &  29 & 4.86 & 0.00 & 0.00 \\ 
-  45 & GO:0009605 & response to external stimulus & 1176 &  83 & 34.83 & 0.00 & 0.00 \\ 
-  46 & GO:0030155 & regulation of cell adhesion & 414 &  45 & 12.26 & 0.00 & 0.00 \\ 
-  47 & GO:0051249 & regulation of lymphocyte activation & 259 &  35 & 7.67 & 0.00 & 0.00 \\ 
-  48 & GO:0006928 & movement of cell or subcellular componen... & 992 &  74 & 29.38 & 0.00 & 0.00 \\ 
-  49 & GO:0001819 & positive regulation of cytokine producti... & 232 &  33 & 6.87 & 0.00 & 0.00 \\ 
-  50 & GO:0006950 & response to stress & 2191 & 123 & 64.89 & 0.00 & 0.00 \\ 
-  51 & GO:0060326 & cell chemotaxis & 145 &  26 & 4.29 & 0.00 & 0.00 \\ 
-  52 & GO:0050793 & regulation of developmental process & 1367 &  89 & 40.48 & 0.00 & 0.00 \\ 
-  53 & GO:0023051 & regulation of signaling & 1746 & 104 & 51.71 & 0.00 & 0.00 \\ 
-  54 & GO:0044763 & single-organism cellular process & 7155 & 274 & 211.90 & 0.00 & 0.00 \\ 
-  55 & GO:0006935 & chemotaxis & 270 &  34 & 8.00 & 0.00 & 0.00 \\ 
-  56 & GO:0008219 & cell death & 1372 &  88 & 40.63 & 0.00 & 0.00 \\ 
-  57 & GO:0042330 & taxis & 271 &  34 & 8.03 & 0.00 & 0.00 \\ 
-  58 & GO:0065008 & regulation of biological quality & 1978 & 112 & 58.58 & 0.00 & 0.00 \\ 
-  59 & GO:0016265 & death & 1376 &  88 & 40.75 & 0.00 & 0.00 \\ 
-  60 & GO:0070887 & cellular response to chemical stimulus & 1330 &  86 & 39.39 & 0.00 & 0.00 \\ 
-  61 & GO:0042110 & T cell activation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
-  62 & GO:0070489 & T cell aggregation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
-  63 & GO:0010646 & regulation of cell communication & 1781 & 104 & 52.74 & 0.00 & 0.00 \\ 
-  64 & GO:0071593 & lymphocyte aggregation & 307 &  36 & 9.09 & 0.00 & 0.00 \\ 
-  65 & GO:0009966 & regulation of signal transduction & 1557 &  95 & 46.11 & 0.00 & 0.00 \\ 
-  66 & GO:0051240 & positive regulation of multicellular org... & 890 &  66 & 26.36 & 0.00 & 0.00 \\ 
-  67 & GO:0048731 & system development & 2399 & 126 & 71.05 & 0.00 & 0.00 \\ 
-  68 & GO:0048856 & anatomical structure development & 2919 & 144 & 86.45 & 0.00 & 0.00 \\ 
-  69 & GO:0007166 & cell surface receptor signaling pathway & 1300 &  83 & 38.50 & 0.00 & 0.00 \\ 
-  70 & GO:0044699 & single-organism process & 7965 & 292 & 235.89 & 0.00 & 0.00 \\ 
-  71 & GO:0040017 & positive regulation of locomotion & 276 &  33 & 8.17 & 0.00 & 0.00 \\ 
-  72 & GO:0030154 & cell differentiation & 2151 & 116 & 63.70 & 0.00 & 0.00 \\ 
-  73 & GO:0002520 & immune system development & 623 &  52 & 18.45 & 0.00 & 0.00 \\ 
-  74 & GO:0002440 & production of molecular mediator of immu... & 111 &  21 & 3.29 & 0.00 & 0.00 \\ 
-  75 & GO:0048534 & hematopoietic or lymphoid organ developm... & 586 &  50 & 17.35 & 0.00 & 0.00 \\ 
-  76 & GO:0002521 & leukocyte differentiation & 347 &  37 & 10.28 & 0.00 & 0.00 \\ 
-  77 & GO:0050663 & cytokine secretion & 101 &  20 & 2.99 & 0.00 & 0.00 \\ 
-  78 & GO:0030595 & leukocyte chemotaxis & 103 &  20 & 3.05 & 0.00 & 0.00 \\ 
-  79 & GO:1902533 & positive regulation of intracellular sig... & 556 &  48 & 16.47 & 0.00 & 0.00 \\ 
-  80 & GO:0050867 & positive regulation of cell activation & 195 &  27 & 5.77 & 0.00 & 0.00 \\ 
-  81 & GO:0042221 & response to chemical & 1868 & 104 & 55.32 & 0.00 & 0.00 \\ 
-  82 & GO:0002250 & adaptive immune response & 226 &  29 & 6.69 & 0.00 & 0.00 \\ 
-  83 & GO:0097529 & myeloid leukocyte migration &  94 &  19 & 2.78 & 0.00 & 0.00 \\ 
-  84 & GO:0030335 & positive regulation of cell migration & 258 &  31 & 7.64 & 0.00 & 0.00 \\ 
-  85 & GO:0044767 & single-organism developmental process & 3254 & 153 & 96.37 & 0.00 & 0.00 \\ 
-  86 & GO:0032879 & regulation of localization & 1499 &  89 & 44.39 & 0.00 & 0.00 \\ 
-  87 & GO:0009306 & protein secretion & 311 &  34 & 9.21 & 0.00 & 0.00 \\ 
-  88 & GO:0051270 & regulation of cellular component movemen... & 494 &  44 & 14.63 & 0.00 & 0.00 \\ 
-  89 & GO:2000147 & positive regulation of cell motility & 264 &  31 & 7.82 & 0.00 & 0.00 \\ 
-  90 & GO:0022407 & regulation of cell-cell adhesion & 249 &  30 & 7.37 & 0.00 & 0.00 \\ 
-  91 & GO:0097530 & granulocyte migration &  66 &  16 & 1.95 & 0.00 & 0.00 \\ 
-  92 & GO:0012501 & programmed cell death & 1290 &  80 & 38.20 & 0.00 & 0.00 \\ 
-  93 & GO:1990266 & neutrophil migration &  57 &  15 & 1.69 & 0.00 & 0.00 \\ 
-  94 & GO:0002696 & positive regulation of leukocyte activat... & 190 &  26 & 5.63 & 0.00 & 0.00 \\ 
-  95 & GO:0006915 & apoptotic process & 1267 &  79 & 37.52 & 0.00 & 0.00 \\ 
-  96 & GO:0002252 & immune effector process & 461 &  42 & 13.65 & 0.00 & 0.00 \\ 
-  97 & GO:0032502 & developmental process & 3284 & 153 & 97.26 & 0.00 & 0.00 \\ 
-  98 & GO:0032940 & secretion by cell & 559 &  47 & 16.55 & 0.00 & 0.00 \\ 
-  99 & GO:0009617 & response to bacterium & 253 &  30 & 7.49 & 0.00 & 0.00 \\ 
-  100 & GO:0051272 & positive regulation of cellular componen... & 270 &  31 & 8.00 & 0.00 & 0.00 \\ 
+GO:0002376 & immune system process & 1406 & 126 & 41.64 & 0.00 & 0.00 \\ 
+  GO:0006955 & immune response & 713 &  85 & 21.12 & 0.00 & 0.00 \\ 
+  GO:0050896 & response to stimulus & 4441 & 229 & 131.52 & 0.00 & 0.00 \\ 
+  GO:0006954 & inflammatory response & 327 &  53 & 9.68 & 0.00 & 0.00 \\ 
+  GO:0044700 & single organism signaling & 2975 & 176 & 88.11 & 0.00 & 0.00 \\ 
+  GO:0006952 & defense response & 755 &  80 & 22.36 & 0.00 & 0.00 \\ 
+  GO:0023052 & signaling & 2981 & 176 & 88.28 & 0.00 & 0.00 \\ 
+  GO:0007165 & signal transduction & 2773 & 167 & 82.12 & 0.00 & 0.00 \\ 
+  GO:0001775 & cell activation & 589 &  69 & 17.44 & 0.00 & 0.00 \\ 
+  GO:0007154 & cell communication & 3137 & 178 & 92.90 & 0.00 & 0.00 \\ 
+  GO:0022610 & biological adhesion & 800 &  79 & 23.69 & 0.00 & 0.00 \\ 
+  GO:0002684 & positive regulation of immune system pro... & 494 &  61 & 14.63 & 0.00 & 0.00 \\ 
+  GO:0007155 & cell adhesion & 791 &  78 & 23.43 & 0.00 & 0.00 \\ 
+  GO:0002682 & regulation of immune system process & 774 &  76 & 22.92 & 0.00 & 0.00 \\ 
+  GO:0001816 & cytokine production & 397 &  53 & 11.76 & 0.00 & 0.00 \\ 
+  GO:0065007 & biological regulation & 6468 & 274 & 191.55 & 0.00 & 0.00 \\ 
+  GO:0048518 & positive regulation of biological proces... & 3404 & 182 & 100.81 & 0.00 & 0.00 \\ 
+  GO:0045321 & leukocyte activation & 515 &  60 & 15.25 & 0.00 & 0.00 \\ 
+  GO:0040011 & locomotion & 903 &  79 & 26.74 & 0.00 & 0.00 \\ 
+  GO:0098602 & single organism cell adhesion & 516 &  58 & 15.28 & 0.00 & 0.00 \\ 
+  GO:0051239 & regulation of multicellular organismal p... & 1538 & 108 & 45.55 & 0.00 & 0.00 \\ 
+  GO:0007159 & leukocyte cell-cell adhesion & 332 &  46 & 9.83 & 0.00 & 0.00 \\ 
+  GO:0016337 & single organismal cell-cell adhesion & 474 &  55 & 14.04 & 0.00 & 0.00 \\ 
+  GO:0051716 & cellular response to stimulus & 3741 & 189 & 110.79 & 0.00 & 0.00 \\ 
+  GO:0098609 & cell-cell adhesion & 507 &  56 & 15.01 & 0.00 & 0.00 \\ 
+  GO:0044707 & single-multicellular organism process & 3409 & 176 & 100.96 & 0.00 & 0.00 \\ 
+  GO:0050789 & regulation of biological process & 6161 & 259 & 182.46 & 0.00 & 0.00 \\ 
+  GO:0048870 & cell motility & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
+  GO:0051674 & localization of cell & 784 &  70 & 23.22 & 0.00 & 0.00 \\ 
+  GO:0032501 & multicellular organismal process & 3487 & 177 & 103.27 & 0.00 & 0.00 \\ 
+  GO:0050794 & regulation of cellular process & 5853 & 249 & 173.34 & 0.00 & 0.00 \\ 
+  GO:0016477 & cell migration & 740 &  67 & 21.92 & 0.00 & 0.00 \\ 
+  GO:0001817 & regulation of cytokine production & 344 &  44 & 10.19 & 0.00 & 0.00 \\ 
+  GO:0034109 & homotypic cell-cell adhesion & 361 &  45 & 10.69 & 0.00 & 0.00 \\ 
+  GO:0048583 & regulation of response to stimulus & 2109 & 126 & 62.46 & 0.00 & 0.00 \\ 
+  GO:0048584 & positive regulation of response to stimu... & 1130 &  85 & 33.47 & 0.00 & 0.00 \\ 
+  GO:0050865 & regulation of cell activation & 319 &  42 & 9.45 & 0.00 & 0.00 \\ 
+  GO:0048522 & positive regulation of cellular process & 3048 & 160 & 90.27 & 0.00 & 0.00 \\ 
+  GO:0046649 & lymphocyte activation & 435 &  49 & 12.88 & 0.00 & 0.00 \\ 
+  GO:0035556 & intracellular signal transduction & 1570 & 103 & 46.50 & 0.00 & 0.00 \\ 
+  GO:0002694 & regulation of leukocyte activation & 301 &  40 & 8.91 & 0.00 & 0.00 \\ 
+  GO:0070486 & leukocyte aggregation & 314 &  40 & 9.30 & 0.00 & 0.00 \\ 
+  GO:1902531 & regulation of intracellular signal trans... & 991 &  76 & 29.35 & 0.00 & 0.00 \\ 
+  GO:0050900 & leukocyte migration & 164 &  29 & 4.86 & 0.00 & 0.00 \\ 
+  GO:0009605 & response to external stimulus & 1176 &  83 & 34.83 & 0.00 & 0.00 \\ 
+  GO:0030155 & regulation of cell adhesion & 414 &  45 & 12.26 & 0.00 & 0.00 \\ 
+  GO:0051249 & regulation of lymphocyte activation & 259 &  35 & 7.67 & 0.00 & 0.00 \\ 
+  GO:0006928 & movement of cell or subcellular componen... & 992 &  74 & 29.38 & 0.00 & 0.00 \\ 
+  GO:0001819 & positive regulation of cytokine producti... & 232 &  33 & 6.87 & 0.00 & 0.00 \\ 
+  GO:0006950 & response to stress & 2191 & 123 & 64.89 & 0.00 & 0.00 \\ 
+  GO:0060326 & cell chemotaxis & 145 &  26 & 4.29 & 0.00 & 0.00 \\ 
+  GO:0050793 & regulation of developmental process & 1367 &  89 & 40.48 & 0.00 & 0.00 \\ 
+  GO:0023051 & regulation of signaling & 1746 & 104 & 51.71 & 0.00 & 0.00 \\ 
+  GO:0044763 & single-organism cellular process & 7155 & 274 & 211.90 & 0.00 & 0.00 \\ 
+  GO:0006935 & chemotaxis & 270 &  34 & 8.00 & 0.00 & 0.00 \\ 
+  GO:0008219 & cell death & 1372 &  88 & 40.63 & 0.00 & 0.00 \\ 
+  GO:0042330 & taxis & 271 &  34 & 8.03 & 0.00 & 0.00 \\ 
+  GO:0065008 & regulation of biological quality & 1978 & 112 & 58.58 & 0.00 & 0.00 \\ 
+  GO:0016265 & death & 1376 &  88 & 40.75 & 0.00 & 0.00 \\ 
+  GO:0070887 & cellular response to chemical stimulus & 1330 &  86 & 39.39 & 0.00 & 0.00 \\ 
+  GO:0042110 & T cell activation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
+  GO:0070489 & T cell aggregation & 305 &  36 & 9.03 & 0.00 & 0.00 \\ 
+  GO:0010646 & regulation of cell communication & 1781 & 104 & 52.74 & 0.00 & 0.00 \\ 
+  GO:0071593 & lymphocyte aggregation & 307 &  36 & 9.09 & 0.00 & 0.00 \\ 
+  GO:0009966 & regulation of signal transduction & 1557 &  95 & 46.11 & 0.00 & 0.00 \\ 
+  GO:0051240 & positive regulation of multicellular org... & 890 &  66 & 26.36 & 0.00 & 0.00 \\ 
+  GO:0048731 & system development & 2399 & 126 & 71.05 & 0.00 & 0.00 \\ 
+  GO:0048856 & anatomical structure development & 2919 & 144 & 86.45 & 0.00 & 0.00 \\ 
+  GO:0007166 & cell surface receptor signaling pathway & 1300 &  83 & 38.50 & 0.00 & 0.00 \\ 
+  GO:0044699 & single-organism process & 7965 & 292 & 235.89 & 0.00 & 0.00 \\ 
+  GO:0040017 & positive regulation of locomotion & 276 &  33 & 8.17 & 0.00 & 0.00 \\ 
+  GO:0030154 & cell differentiation & 2151 & 116 & 63.70 & 0.00 & 0.00 \\ 
+  GO:0002520 & immune system development & 623 &  52 & 18.45 & 0.00 & 0.00 \\ 
+  GO:0002440 & production of molecular mediator of immu... & 111 &  21 & 3.29 & 0.00 & 0.00 \\ 
+  GO:0048534 & hematopoietic or lymphoid organ developm... & 586 &  50 & 17.35 & 0.00 & 0.00 \\ 
+  GO:0002521 & leukocyte differentiation & 347 &  37 & 10.28 & 0.00 & 0.00 \\ 
+  GO:0050663 & cytokine secretion & 101 &  20 & 2.99 & 0.00 & 0.00 \\ 
+  GO:0030595 & leukocyte chemotaxis & 103 &  20 & 3.05 & 0.00 & 0.00 \\ 
+  GO:1902533 & positive regulation of intracellular sig... & 556 &  48 & 16.47 & 0.00 & 0.00 \\ 
+  GO:0050867 & positive regulation of cell activation & 195 &  27 & 5.77 & 0.00 & 0.00 \\ 
+  GO:0042221 & response to chemical & 1868 & 104 & 55.32 & 0.00 & 0.00 \\ 
+  GO:0002250 & adaptive immune response & 226 &  29 & 6.69 & 0.00 & 0.00 \\ 
+  GO:0097529 & myeloid leukocyte migration &  94 &  19 & 2.78 & 0.00 & 0.00 \\ 
+  GO:0030335 & positive regulation of cell migration & 258 &  31 & 7.64 & 0.00 & 0.00 \\ 
+  GO:0044767 & single-organism developmental process & 3254 & 153 & 96.37 & 0.00 & 0.00 \\ 
+  GO:0032879 & regulation of localization & 1499 &  89 & 44.39 & 0.00 & 0.00 \\ 
+  GO:0009306 & protein secretion & 311 &  34 & 9.21 & 0.00 & 0.00 \\ 
+  GO:0051270 & regulation of cellular component movemen... & 494 &  44 & 14.63 & 0.00 & 0.00 \\ 
+  GO:2000147 & positive regulation of cell motility & 264 &  31 & 7.82 & 0.00 & 0.00 \\ 
+  GO:0022407 & regulation of cell-cell adhesion & 249 &  30 & 7.37 & 0.00 & 0.00 \\ 
+  GO:0097530 & granulocyte migration &  66 &  16 & 1.95 & 0.00 & 0.00 \\ 
+  GO:0012501 & programmed cell death & 1290 &  80 & 38.20 & 0.00 & 0.00 \\ 
+  GO:1990266 & neutrophil migration &  57 &  15 & 1.69 & 0.00 & 0.00 \\ 
+  GO:0002696 & positive regulation of leukocyte activat... & 190 &  26 & 5.63 & 0.00 & 0.00 \\ 
+  GO:0006915 & apoptotic process & 1267 &  79 & 37.52 & 0.00 & 0.00 \\ 
+  GO:0002252 & immune effector process & 461 &  42 & 13.65 & 0.00 & 0.00 \\ 
+  GO:0032502 & developmental process & 3284 & 153 & 97.26 & 0.00 & 0.00 \\ 
+  GO:0032940 & secretion by cell & 559 &  47 & 16.55 & 0.00 & 0.00 \\ 
+  GO:0009617 & response to bacterium & 253 &  30 & 7.49 & 0.00 & 0.00 \\ 
+  GO:0051272 & positive regulation of cellular componen... & 270 &  31 & 8.00 & 0.00 & 0.00 \\ 
    \hline

### 2016-04-14 additional_files/tex/cluster6_BP_Ef_SporozoitesUp.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:52 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0007154 & cell communication &  63 &   8 & 1.68 & 0.00 & 0.02 \\ 
  GO:0007165 & signal transduction &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0023052 & signaling &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0044700 & single organism signaling &  58 &   7 & 1.54 & 0.00 & 0.02 \\ 
  GO:0007186 & G-protein coupled receptor signaling pat... &   3 &   2 & 0.08 & 0.00 & 0.03 \\ 
  GO:0007205 & protein kinase C-activating G-protein co... &   3 &   2 & 0.08 & 0.00 & 0.03 \\ 
  GO:0051716 & cellular response to stimulus &  99 &   8 & 2.64 & 0.00 & 0.05 \\ 
  GO:0009405 & pathogenesis &   5 &   2 & 0.13 & 0.01 & 0.07 \\ 
  GO:0051704 & multi-organism process &   5 &   2 & 0.13 & 0.01 & 0.07 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-04-14 additional_files/tex/cluster7_MF_Ef_Day7OocyDown_SporoUP.tex
Changes to heatmaps (more cluster in Ef) and enrichment, to include new clusters. Also started a .tex to summarize all GO-tables with legends. This is work in progress

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Apr 14 14:57:55 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0052689 & carboxylic ester hydrolase activity &  12 &   5 & 0.56 & 0.00 & 0.01 \\ 
  GO:0016307 & phosphatidylinositol phosphate kinase ac... &   4 &   3 & 0.19 & 0.00 & 0.02 \\ 
  GO:0004725 & protein tyrosine phosphatase activity &   6 &   3 & 0.28 & 0.00 & 0.05 \\ 
  GO:0004812 & aminoacyl-tRNA ligase activity &  41 &   7 & 1.92 & 0.00 & 0.05 \\ 
  GO:0016875 & ligase activity, forming carbon-oxygen b... &  42 &   7 & 1.96 & 0.00 & 0.05 \\ 
  GO:0016876 & ligase activity, forming aminoacyl-tRNA ... &  42 &   7 & 1.96 & 0.00 & 0.05 \\ 
  GO:0008026 & ATP-dependent helicase activity &  60 &   8 & 2.81 & 0.01 & 0.08 \\ 
  GO:0070035 & purine NTP-dependent helicase activity &  60 &   8 & 2.81 & 0.01 & 0.08 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-05-19 manuscript/tex_bmc_format/manuscript_rnaseq.tex
working on GO clusters and genes and some text in draft manuscript

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
%\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%\usepackage{siunitx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphicx}
\graphicspath{ {../figures/} }
%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
%\affi1=Department of Molecular Parasitology, Humboldt University, Berlin, Germany
%\affi2=Max Planck Institute for Biology of Ageing, Joseph-Stelzmann-Strasse 9B, 50931 Cologne, Germany
%\affi3=Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research, Forschungsverbund Berlin e.V., Berlin, Germany
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A sample article title}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={jane.e.doe@cambridge.co.uk}   % email address
]{\inits{JE}\fnm{Jane E} \snm{Doe}}
\author[
   addressref={aff1,aff2},
   email={john.RS.Smith@cambridge.co.uk}
]{\inits{JRS}\fnm{John RS} \snm{Smith}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Department of Zoology, Cambridge}, % university, etc
  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
  \street{D\"{u}sternbrooker Weg 20},
  \postcode{24105}
  \city{Kiel},
  \cny{Germany}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\author[
%   addressref={affi1},                   % id's of addresses, e.g. {aff1,aff2}
%   corref={aff1},                       % id of corresponding address, if any
%  % noteref={n1},                        % id's of article notes, if any
%   email={kasemotq@hu-berlin.de}   % email address
%]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
%\author[
%   addressref={affi1}                   % id's of addresses, e.g. {aff1,aff2}
%  % corref={aff1},                       % id of corresponding address, if any
%  % noteref={n1},                        % id's of article notes, if any
%  % email={}   % email address
%]{\inits{SS}\fnm{Simone} \snm{Spork}}
%%\author[
%   addressref={affi2},                   % id's of addresses, e.g. {aff1,aff2}
%   %corref={aff3},                       % id of corresponding address, if any
%   noteref={n2},                        % id's of article notes, if any
%   %email={emanuel.heitlinger@hu-berlin.de}   % email address
%]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
%\author[
%   addressref={affi1, affi3},                   % id's of addresses, e.g. {aff1,aff2}
%   %corref={aff3},                       % id of corresponding address, if any
%   noteref={n1},                        % id's of article notes, if any
%   email={emanuel.heitlinger@hu-berlin.de}   % email address
%]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\address[id=aff1]{%                           % unique id
%  \orgname{Department of Zoology, Cambridge}, % university, etc
%  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
%  \city{London},                              % city
%  \cny{UK}                                    % country
%}
%\address[id=affi1]{%
%  \orgname{Humboldt-Universitat zu Berlin, Department of Molecular Parasitology},
%  \street{Philipstr. 13, House 14},
%  \postcode{10115}
%  \city{Berlin},
%  \cny{Germany}
%}
%\address[id=affi2]{%  
%  \orgname{Max Planck Institute for Biology of Ageing},
%  \street{Joseph-Stelzmann-Strasse 9B},
%  \postcode{50931}
%  \city{Cologne},
%  \cny{Germany}
%}
%\address[id=affi3]{%  
%  \orgname{Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research},
%  \street{Alfred-Kowalke-Str. 17},
%  \postcode{10315}
%  \city{Berlin},
%  \cny{Germany}
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Introduction}
Text and results for this section, as per the individual journal's instructions for authors. 
%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \section*{Results}
Text for this section \ldots
\subsection*{GO term enrichments in heatmap clusters}
The annotations referred to here are in many 
cases inferred from the existence of predicted protein domains and have in some cases been 
verified in e.g. \textit{E. tenella} or other \textit{Eimeria} spp.

\subsubsection*{Preparation for invasion in oocysts}
The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
Overrepresented GO-term in this cluster are largely determined by genes orthologous to 
peptidases, microneme localized proteins reported to be involved in invasion, 
genes associated with clotting and in protozoans, adhesion, and genes that are annotated 
to be involved in amino acid biosynthesis.
  Aminopeptidase N ('related' annotation) is the annotation for orthologs of three genes with 
abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
major histocompatibility complex, MHC, II (UniProt reference if we want to keep this... but does any
secretion happen from oocysts...? Or is this too far-fetched to be interesting?).

Three genes can be linked to adhesion and through localization the same genes
also are asociated with invasion in the literature. 
A protein domain which is associated with adhesion in protozoans, thrombospondin type 1 
domain-containing protein, was found in our oocyst cluster (cluster 1). 
Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme localizing 
proteins, MIC. Tomley et al. (Tomley01) describe MIC4 as a protein contatining thrombospondin 
type 1 repeats. In \textit{E. tenella} MIC4 is expressed in sporozoites where it localizes to the 
apical end, in late schizonts and late oocyst stages, when sporozoites are forming.
(Tomley01)
In the protozoan \textit{P. falciparum}, another domain occuring in our data, the sushi domain, 
is reported in the apical sushi protein, ASP. In \textit{P. falciparum} it localizes to the 
micronemes in merozoites but not other stages (OKeeffe05). To our knowledge, the role of sushi 
domain proteins in \textit{Eimeria} spp. oocysts has not yet been investigated. 
  Other protein domains detected via orthologs to mRNAs expressed in oocysts in our data are
Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, also known as F5/8 domain. 
LCCL domains have  been associated with discoidin lectin domains, which in turn are involved 
in cell adhesion (Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
In the slime mold \textit{Dictyostelium discoideum)} the LCCL domain
is part of the mold'd discoidin adhesion protein. In humans, the LCCL domain in annotated as a 
coagulation factor domain. (Pfam entry for 'discoidin domain', May 2016)

\subsubsection*{Amino acid biosynthesis in oocysts}
mRNA of a aminotransferases indicate amino acid biosynthesis in oocysts. 
We also identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase, which are enzymes 
contributing to L-serine and L-alanine production, respectively. A putative \textit{Eimeria} spp.
cystathionine beta-synthase, CBS, in this cluster also indicates de novo cysteine production. 
Alkyl sulfatase mRNA is another gene which contributes to overrepresentation of a GO-term in 
mRNAs highly abundant in oocysts. Generally, this enzyme enables an organism to exploit organic 
sulfur to produce and incorporate inorganic sulfur into the amino acids cysteine and methionine, 
when no inorganoc sulfur is available.

'Embryonic development'
Nicalin 1, patched family protein (hedgehog)

Propionate is one of two most abundant small-chain fatty acids 
in the gut along with butyrate. Both fatty acids are largely produced as degradation products from food
by commensal bacteria (Sun13). MmgE/PrpD is important for propionate catabolism in the 
2-methylcitric acid cycle and has been shown to be used by the intestinal intracellular bacterium 
\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{cluster 3..........}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
Two clusters contain genes with mRNAs highly abundant on day 7 p.i; cluster 1 and 2.
Dynein, kinesin and tubulin are annotations highly represented among orthologs of genes 
in both these clusters. The annotations indicate an important role for motility at this
timepoint, probably reflecting development of microgametes.
In addition, in cluster 2, there are two 'EF-hand domain containing proteins' annotations as well 
as caltractin, centrin-1, and troponin annotations. Caltractin and 
centrin-1 are associated with the centrosome and structure and function of microtubuli in mammals, 
and troponin is linked to muscle function (UniProt). 
Also potentially linked to motility is the occurence of growth arrest specific protein 8,
Gas8, which in the mouse has been reported to be highly 
expressed in the testes and important for mouse sperm function (Yeh02).

Other genes among the 38 indicate carbon fixation (glycolysis/gluconeogenesis) or 
conversions of nucleoside phosphates. In addition, a Ras family protein, RNA 
polymerase II transcription initiation factor and Sec23 and Sec24 were among orthologs 
identified in \textit{E. falciformis} cluster 2.

In cluster 1, carbon metabolism genes are represented by 6-phosphogluconate dehydrogenase 
and glycogen phosphorylase family protein 1. UDP-glucose 4-epimerase and amiloride-sensitive 
amine oxidase are reported as upregulated in gametocytes in \textit{E. tenella} by 
RNA-seq (Walker15) and suggested by those authors to play a role in cyst wall synthesis.

\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
Unintuitively for a protozoan organism, seven out of eight GO biological process terms
in cluster 1 are associated with wound healing and blood coagulation.
An explanation is offered by some of the orthologs to the three \textit{E. falciformis} genes
responsible for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and 
\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal protein MIC4,
related' (E. tenella) and more generally for several other protozoa, 'PAN domain containing proteins'.
The PAN domain is found in the plasminogen/hepatocyte growth factor family and in coagulation 
factor XI family (REF), explaining why terms related to blood coagulation are enriched by these genes.
Later publications on \textit{T. gondii} (Marchant12) also associate PAN domains and proteins in 
apicomplexan parasites with micronemes and therefore invasion. In our case, this is peculiar, since 
the enrichment appears on day 7 p.i.. A possible role at this timepoint is suggested by work on 
the fungi \textit{Sclerotinia sclerotiorum} where Yu et al. reported an important role for PAN domain
proteins in cell wall integrity (Yu12). This role for MIC proteins has to our knowledge 
not been investigated in apicomplexan parasites. 
The PAN domain domain has also been reported to be common in nematodes such as 
\textit{Caenorhabditis elegans}, however the function is not understood. (Thordai99)
The other two GO terms in the cluster of day seven upregulated genes are DNA replication and DNA 
replication initiation, which most likely reflects late stage schizogony or gamete formation.
Six genes contribute to this enrichment and orthologs are either annotated as DNA replcation
licencing factors, DNA polymerases or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.

\subsubsection*{Gene and sample patterns by hierarchical clustering}
Samples (columns) cluster into two major clusters where day 7 p.i. samples form one group distinct from
other samples. In the second group, oocsts and sporozoites are distinct and sporozoites cluster
most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples also cluster into two groups,
of which one contains all NMRI day 5 p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
clusters have no obvious patters.

For gene clusters (rows), the two groups with high mRNA abundance on day 7 p.i. (cluster 1 and 2)
do not cluster most closely with each other, but with the cluster for high mRNA abundance in oocysts
(cluster 1 association) and with the cluster for high mRNA abundance in sporozoites (cluster 2 
association).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
In our analysis we demonstrate which biological processes are dominant in different
life cycle stages of E. falciformis in the mouse. The RNAseq transcriptome provided here 
allows for detailed analysis of genes involved in those processes, providing candidates for 
life stage specific marker in \textit{Eimeria} spp. research.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods}
\subsection*{Mice and infection procedure}
Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
Animal procedures were performed according to the German Animal Protection Laws as 
directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
tapwater was used instead of PBS for administration of oocysts. Briefly, the indicated number of
sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate
was used for all infections and parasite samples. Oocysts were originally purchased from Bayer 
() and the strain is maintained through passage in mice  in our facilities as described 
elsewhere (Schmid12).
\ldots
\subsection*{RNA extraction}
 \ldots
\subsection*{Sequence quality assessment and alignment}
Fastq\_quality\_filter was applied to Illumina Hiseq 2000 sequenced samples. 
Since this is not easily applicable to pair-end sequencing data, a low threshold was 
used on the hiseq data. A phred score of 10 was used, i.e., the probability of false base 
calling is one in ten. We further set q = 60, i.e., nine out of ten bases
or more is required to be correct in at least 
60\% of the bases in each read for the read sequence to be kept for further analysis.
This resulted in.......
\subsubsection*{Alignment and reference genomes}
We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
files including mRNA sequences from both species were aligned against the dual reference genome
using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
(version 2.1.1, Trapnell10).

\ldots

\subsection*{Sub-heading for section}
Text for this sub-heading \ldots
\subsubsection*{Sub-sub heading for section}
Text for this sub-sub-heading \ldots
\paragraph*{Sub-sub-sub heading for section}
Text for this sub-sub-sub-heading \ldots
text text text
text text text
text text text
text text text
text text text
text text text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%\begin{figure}
 %\begin{picture}
 % \includegraphics[width=0.5\linewidth]{ef-heatmap}
 %\end{picture}
%\end{figure}

\section*{Figures}
\begin{figure}[h!]
  \includegraphics[width=0.5\linewidth]{efmm-ratio}
  \caption{\csentence{Parasite read sequences per mouse read sequences for each experimental condition.}
  The more advanced the infection is, the higher the ratio of parasite reads is, reflecting parasite replication between day zero and day seven. ......compare 1st and 2nd ....... patterns between Rag and NMRI/C57BL/6?  ........ Values are mean of replicates (n=2, if * n=1) on log10 scale. Each sample (replicate) consists of mRNA from three different mice.}
  \end{figure}

  \begin{figure}[h!]
  \includegraphics[width=0.7\linewidth]{ef-heatmap}
  \caption{\csentence{Parasite genes with different mRNA abundance between samples, sorted into five clusters by hierarchical clustering (method = complete, distance = euclidean).}
      \textit{E. falciformis} samples from seven days p.i. cluster together (NMRI mice only). These samples have distinct mRNA abundance patterns in all gene clusters, although more pronounced in clusters 1 (up) and 4 (down). Distinct groups of genes also define sporozoites (cluster 5, up) and oocysts (cluster 2, up). mRNA profiles on days three and five p.i. from all three mouse strains cluster together. These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites, however closest to the latter. On scale bar, 0 is mean mRNA abundance for each gene. Up and downregulation is standard deviations from mean.}
      \end{figure}

\begin{figure}[h!]
  \includegraphics[width=0.7\linewidth]{mm-heatmap}
  \caption{\csentence{Mouse genes with different mRNA abundance between samples, sorted into four clusters by hierarchical clustering (method = complete, distance = euclidean).}
Three out of four samples from day 7 p.i. (NMRI only) cluster together. These samples are characterized by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6 uninfected samples cluster together (left), defined by clusters 3, 4 (up), and 1 (down). Rag1-/- samples cluster together. These samples share a weak downregulation of most genes in cluster 2 as well as upregulation of a small group genes in the same cluster. Uninfected Rag1-/- samples are separated from infected ones with distinct profiles in clusters 1, 3, and 4. On scale bar, 0 is mean mRNA abundance for each gene. Up and downregulation is standard deviations from mean.}
      \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2016-06-03 additional_files/tex/SI_clusters/clusterTablesCombined.tex
Mainly gne_list changes --> orthologs from ToxoDB, and some text in manuscipt draft (including some obvious errors in results part)

@@
 \usepackage[section]{placeins}
-\usepackage[hscale=0.8,vscale=0.8]{geometry}
+\usepackage[hscale=0.8,vscale=0.85]{geometry}
 %\usepackage{natbib} % Required to change bibliography style to APA
@@
 Cluster 1, BP, Ef.
-  Overrepresented GO biological process terms among parasite genes which are 
-  highly abundant on day 7 p.i.. Compared to cluster 2, which also have a high
-  abundance on day 7, these genes cluster more closely with genes in cluster 4, 
-  which are highly expressed in oocysts.\\
+Overrepresented GO 'biological process' terms among parasite mRNAs which are 
+highly abundant on day 7 p.i. only. Cluster 1 groups closely with cluster 4,
+in which mRNAs are highly abundant in oocysts. The cluster contains 201 genes of which 
+96 have GO annotations.\\
+
+\newline
+The eight first GO terms in this list are supported by the same three genes. 
+Unintuitively for a protozoan organism, seven out of eight of these terms are associated with 
+wound healing and blood coagulation.
+However, orthologs in ToxoDB pointed towards \textit{Eimeria} spp. and \textit{Toxoplasma gondii} 
+proteins annotated as micronemal protein 4, MIC4, related proteins and more generally, PAN 
+domain containing proteins. The PAN domain is found in plasminogen
+/hepatocyte growth factor family and in coagulation factor XI family, explaining the GO terms. 
+Of interest here, the same domain has also been reported to be common in 
+nematodes such as \textit{Caenorhabditis elegans}, however the
+function is not understood. (Thordai99)
+In \textit{Eimeria} spp. and the apicomplexan parasite \textit{Hammondia hammondi}, orthologs
+are annotated as micronemal protein MIC4, a microneme protein reported by, e.g., the Tomley group
+(Brown01).
+Later publications on \textit{T. gondii} (Marchant12) also associate these domains and proteins with 
+micronemes and therefore invasion. In our case, this is a peculiar result, since the enrichment appears on
+day 7 p.i.. A possible role at this timepoint is suggested by work on the fungi 
+\textit{Sclerotinia sclerotiorum} where Yu et al. reported an important role for PAN domain
+proteins in cell wall integrity (Yu12). This role for MIC proteins has to our knowledge 
+not been investigated in apicomplexan parasites. 
+The other two GO terms in the cluster of day seven upregulated genes are DNA replication and DNA 
+replication initiation, which most likely reflects late stage schizogony or gamete formation.
 %\newpage
 \input{../cluster1_BP_Ef_Day7Up_1.tex}
-
-  Cluster 2, BP, Ef.
-  Overrepresented GO biological process terms among parasite genes which are 
-  highly abundant on day 7 p.i.. Compared to cluster 1, which is also upregulated 
-  on day 7 p.i., these genes cluster more closely with cluster 6, which contains
-  genes highly abundant in sporozoites.\\
-  \input{../cluster2_BP_Ef_Day7Up_2.tex}
 
@@
 which are highly expressed in oocysts.\\
+%%%
+Four GO molecular functions are enriched in cluster 1 (adj. p-value = 0.14), 
+supported by ten different genes, i.e., there is overlap in the genes. 
+Carbon metabolism genes are represented by 6-phosphogluconate dehydrogenase and glycogen phosphorylase 
+family protein 1. UDP-glucose 4-epimerase and amiloride-sensitive amine oxidase are 
+reported as upregulated in gametocytes in \textit{E. tenella} by RNA-seqi (Walker15) and 
+suggested by those authors to play a role in cyst wall synthesis.
 \input{../cluster1_MF_Ef_Day7Up_1.tex}
+
+Cluster 2, BP, Ef.
+Overrepresented GO biological process terms among parasite genes which are 
+highly abundant on day 7 p.i.. Compared to cluster 1, this cluster contains genes 
+which are highly abundant both on day 7 p.i. and in sporozoites (cluster 6).
+In total the cluster contains 271 genes of which 107 have GO annotations.\\
+\newline
+The term 'single-organism process' is enriched and is the GO-term containing the largest
+number of genes in cluster 2 (38). The two 'single-organism' terms have a complete overlap of genes.
+All genes contributing to other GO terms in this list, except those terms indicating motility 
+(top 3 GO terms), are also contained in the 38 'single-organism' genes.
+Therefore, although annotations such as purine/nucleoside/nucleotide metabolic process 
+dominate, the genes responsible for those enrichments are all contained in the two 'single organism'
+terms.
+A closer look at these 38 genes reveals 12 genes annotated as either dynein, kinesin or tubulin chains or 
+proteins, indicating activity related to parasite motility. (((((((((( COMMENT ))))))))))
+Also potentially linked to motility is the 
+occurence of growth arrest specific protein 8, Gas8, which in the mouse has been reported to be highly 
+expressed in the testes and important for mouse sperm function (Yeh02). Other genes among the 38 indicate
+carbon fixation (glycolysis/gluconeogenesis) or different conversions of nucleoside phosphates. 
+In addition, a Ras family protein, RNA polymerase II transcription initiation factor and 
+Sec23 and Sec24 were among orthologs identified in \textit{E. falciformis} cluster 2.
+%%%
+In the list of enriched GO 'molecular functions', 'microtubule motor activity' also indicate 
+parasite movement. Under the term 'protein binding', genes annotated as dynein and kinesin 
+are different from the genes annotated as dynein and kinesin in the list for 
+GO 'biological process'. 
+Orthologs to genes responsible for other GO term enrichments in cluster 2 also indicate motility
+with annotations including 'flagellar' or 'intraflagellar'.
+In addition, there are two 'EF-hand 
+domain containing proteins' annotations as well as direct naming of such proteins: caltractin and 
+centrin-1, and troponin. Caltractin and centrin-1 are associated with the centrosome 
+and structure and function of microtubuli in mammals, and troponin is linked to muscle function 
+(UniProt). 
+\input{../cluster2_BP_Ef_Day7Up_2.tex}
 
@@
 genes highly abundant in sporozoites.\\
+%%%
+
 \input{../cluster2_MF_Ef_Day7Up_2.tex}
@@
 abundance in sporulated oocysts and sporozoites. In mouse stages, i.e., day 3, 5, and 7 samples
-  there is a tendency to upregulation among these genes.\\
+there is a tendency to upregulation among these genes. In total the cluster contains 204 genes of which 
+121 have GO annotations, and 83 do not.\\
+ \\
 \input{../cluster3_BP_Ef_OocSporDown.tex}
@@
 sporulated oocysts. These genes cluster most closely with cluster 1, in which genes are highly
-  abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.\\
+abundant on day 7 p.i.. In cluster 4, genes have below average abundance in all day 3 and 5 samples.
+In total the cluster contains 316 genes of which 157 have GO annotations, and 159 do not.\\
+ \\
+
+  Serine hydroxymethyltransferase --> (L-)serine metabolism...(?)
+  Cystathionine beta-synthase -->  converts (L-)homocysteine to cystathionine using L-serine
+  Acyl-CoA dehydrogenase    --> beta-oxidation in mitochondria
+  MmgE/PrpD family domain -->  PrpD important for propionate catabolism in E. coli (see Sushi-domain)
+  Thrombospondin type 1 domain-containing... --> microneme protein implicated in adhesion
+                                                MIC proteins in Tomley et al (2001)
+  Sushi domain (scr repeat) domain
+
+  Sushi domain protein: complement binding in mammals
+In Pf microneme localised (OKeeffe05) and suggested GPI anchored. Expressed in merozoites (erythrocyte invasion
+stage) --> in Eimeria different, since oocyst high expression?
+MmgE/PrpD is important for propionate catabolism in the 2-methylcitric acid cycle mainly in E. coli and
+other bacteria (UniProt). Propionyl-CoA can be produced from propanoate and is reported to
+be toxic in... not sure about the statements in the databases here. Check these sites:
+http://www.ebi.ac.uk/intenz/query?cmd=SearchEC&ec=2.8.3.1
+https://en.wikipedia.org/wiki/Propionate_CoA-transferase
+http://www.llamp.net/?q=2-Methylcitrate%20cycle
+...and these publications:
+http://www.ncbi.nlm.nih.gov/pubmed/16151139
+http://www.ncbi.nlm.nih.gov/pubmed/10482501
+http://www.ncbi.nlm.nih.gov/pubmed/19767840
+
+Propionate is one of two most abundant small-chain fatty acids 
+in the gut along with butyrate. Both fatty acids are largely produced as degradation products from food
+by commensal bacteria (Sun13).
+
+## OLD CLUSTERING REFS???
 \input{../cluster4_BP_Ef_OocystsUp.tex}
@@
 
+ATPase or histidine kinase or DNA gyrase B or HSP90:
+Gyrase has also been found in Plasmodium (Wiki) but is commonly known from bacteria, 
+where it contributes to negative supercoiling of DNA. Many antibiotics act on gyrases
+according to Wiki.
+
+Patched family protein - patched is a membrane receptor which transduces hedgehog signals 
+in e.g. humans (Pfam). Conserved pathway, involved in embryonic development and 
+tissue maintenance. --> Not really able to find references for protozoa. Plasmodium
+paper (doi:  10.1186/s12936-015-0562-1) states possible crosstalk between parasite
+protein and host hedgehog...
+
+LCCL (Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1) -->
+In metazoa, this clotting factor has been associated with other protein domains,
+including discoidin lectin domains (Pfam entry for LCCL domain).
+  F5/8 type C domain - main annotation is coagulation factor. Perhaps more 
+relevant here, a similar domain has been found in slime molds (Dictyostelium discoideum)
+where it is part of the adhesion protein discoidin (Pfam entry for "Discoidin domain")
+
+Alkyl sulfatase --> cleaves organic sulfure: "Usually inorganic sulfate is 
+available and is transformed through a cascade of reactions to yield the 
+high-energy intermediate 3′-phosphoadenosine-5′-phosphosulfate (PAPS). 
+Ultimately, sulfur is incorporated into the essential amino acids 
+cysteine and methionine. In case inorganic sulfate is unavailable, organisms 
+are forced to express other sulfur metabolizing enzymes, such as alkyl sulfatases,
+which are able to hydrolytically cleave inorganic sulfate from organic sulfate 
+esters" (doi:  10.1007/s00253-013-5438-0)
+
+  Aminopeptidase N --> Peptide cleaving, in humans mainly after digestion by gastric
+and pancreatic proteases. Curious: "Found to cleave antigen peptides bound to major 
+histocompatibility complex class II molecules of presenting cells" (UniProt for
+Homo sapiens "amonipeptidase N").
+  X-Pro dipeptidyl-peptidase domain --> aminopeptidase activity, serine targeting
+(InterPro, IPR008252). The Toxo annotation for CocE/NonD are serine esterases specifically
+enabling using cocaine as carbon source (InterPro).
+
+  Aminotransferase --> most likely linked to amino acid biosynthesis (UniProt for 
+Drosophila)
+  Nicalin-1 --> binds Nomo. Important for embryonic development in vertebrates (PMID:
+15257293). Stated on Zebrafish DB that it's important for egg hatching, but it
+has the same reference as above. 
+  Toxo: M28 family peptidase --> contains aminopeptidases (Merops Sanger entry
+  Aspartyl protease --> specific protease, cleaves in one step, most commonly
+found in stomach of mammals (PMID:1455179, review on...).
+  Glucose-methanol-choline, GMC, oxidoreductase --> family of flavoenzymes, which
+contains, e.g., glucose dehydrogenases (DOI: 10.1016/0022-2836(92)90992-S)
+  Mannitol-1-phosphate dehydrogenase AKA fructose 6-phosphate reductase -->
+
+  Glucose-6-phosphate 1-dehydrogenase --> oxidizes glucose-6-phosphate and reduces
+NADP+ to NADPH. 
+  D-3-phosphoglycerate dehydrogenase --> PGDH catalyses an early step in the biosynthesis of L-serine by converting 
+D-3-phosphoglyceric acid to hydroxypyruvic acid phosphate (HPAP), utilising 
+NAD+ as a coenzyme [PMID: 19924905, PMID: 22023909].(InterPro)
+     transhydrogenase --> oxidoreductase family, NADH production
+OR   alanine dehydrogenase --> L-alanine + H2O + NAD+ <-> pyruvate + NH3 + NADH + H+
+    Protein NNT (nicotinamide nucleotide transhydrogenase) -->
+Identified in Plasmodium where it seems to be expressed in gametocytes.
+(Thesis: http://hdl.handle.net/1860/477). For human: "The enzyme couples hydride 
+transfer between NAD(H) and NADP(+) to proton translocation across the inner 
+mitochondrial membrane. Under most physiological conditions, the enzyme uses energy 
+from the mitochondrial proton gradient to produce high concentrations of NADPH. 
+The resulting NADPH is used for biosynthesis and in free radical detoxification."
+(NCBI, Gene ID: 23530).
+
 Cluster 5, BP, Ef.
@@
 in sporulated oocysts, sporozoites and on day 7 p.i.. These genes seem to increase in abundance
-  upon infection and decrease late in infection, i.e., on day 7.\\
+upon infection and decrease late in infection, i.e., on day 7. In total the cluster contains 183 genes
+of which 83 have GO annotations, and 100 do not.\\
+ \\
 \input{../cluster5_BP_Ef_OoSpoDown_day5Up.tex}
@@
 This cluster seems to distinguish sporozoites from sporulated oocysts and have an average mRNA abundance in
-  all other samples.\\
+all other samples. In total the cluster contains 180 genes of which 70 have GO annotations, and 110 do not.\\
+ \\
 \input{../cluster6_BP_Ef_SporozoitesUp.tex}
@@
 In sporozoites and early in infection, i.e., day 3 and 5 p.i., some of these genes are highly abundant. On day
-  7 p.i. these mRNAs have a below average abundance.\\
+7 p.i. these mRNAs have a below average abundance. In total the cluster contains 263 genes of which 131 have 
+GO annotations, and 132 do not.\\
+ \\
 \input{../cluster7_BP_Ef_Day7OocyDown_SporoUP.tex}
@@
   sample cluster. On day 7, most of these genes are not differentially abundant, or below average, i.e., they
-  seem to peak on day 5 p.i..\\
+  seem to peak on day 5 p.i.. In total the cluster contains 273 genes of which 244 have GO annotations, and 29 
+  do not.\\
+   \\
   \input{../cluster1_BP_Mm_EarlyUp.tex}
@@
   Overrepresented GO biological process terms among mouse genes which are downregulated on day 7 p.i. and
-  upregulated pre infection, i.e., these genes become more and more downregulated during infection.\\
+  upregulated pre infection, i.e., these genes become more and more downregulated during infection.
+  In total the cluster contains 198 genes of which 183 have GO annotations, and 15 do not. \\
   \input{../cluster2_BP_Mm_7dpiUp.tex}
@@
   poor infection in this sample. As in cluster 2, these mRNAs are most highly abundant before infection and
-  become less abundant over the course of infection.\\
+  become less abundant over the course of infection. In total the cluster contains 465 genes of which 440
+  have GO annotations, and 25 do not\\
   \input{../cluster3_BP_Mm_7dpiDown.tex}
@@
   There is overall no clear difference between pre infection and days 3 or 5 among these genesi. A slight
-  tendency to higher abundance in second infection samples can be seen among these genes.\\
+  tendency to higher abundance in second infection samples can be seen among these genes.
+  In total the cluster contains 377 genes of which 360 have GO annotations, and 17 do not\\
   \input{../cluster4_BP_Mm_allInfDown.tex}

### 2016-06-03 4_compareArray.R

@@ -93,20 +93,23 @@ RNAseq.logFC <- Mm.1st.pass.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
                                            "logFC.N7vs0", "logFC.B5vs0",
                                            "logFC.R5vs0" )]
 
-RNAseq.logFC <- merge(RNAseq.logFC, cuff2name, by.x = 0, by.y = "cuff_ids")
+RNAseq.logFC <- merge(RNAseq.logFC, annot.frame, by.x = 0, by.y = "ensembl_id")
+#RNAseq.logFC <- merge(RNAseq.logFC, cuff2name, by.x = 0, by.y = "cuff_ids")
 
 RNAseq.logFC.ruved <- Mm.RUVg.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
                                              "logFC.N7vs0", "logFC.B5vs0",
                                              "logFC.R5vs0" )]
 
-RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, cuff2name,
-                            by.x = 0, by.y = "cuff_ids")
+RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, annot.frame,
+                            by.x = 0, by.y = "ensembl_id")
+#RNAseq.logFC.ruved <- merge(RNAseq.logFC.ruved, cuff2name,
+#                            by.x = 0, by.y = "cuff_ids")
 
 RNAseq.Array.logFC <- merge(RNAseq.logFC, Array.logFC,
-                            by.x = "gene_names", by.y = 0)
+                            by.x = "gene_name", by.y = 0) # NOTE!!!! gene_name or gene_names
 
 RNAseq.Array.logFC <- merge(RNAseq.Array.logFC, RNAseq.logFC.ruved,
-                            by = c("Row.names","gene_names"))
+                            by = c("Row.names","gene_name"))
 
 names(RNAseq.Array.logFC) <- gsub(".x", ".plain", names(RNAseq.Array.logFC))
 

### 2016-06-03 5_clustering.R

@@ -33,8 +33,8 @@ Ef.hclustered.df <- as.data.frame(hcluster[["Ef"]])
 names(Ef.hclustered.df) <- "Cluster"
 Ef.hclustered.df$Cluster <- as.factor(Ef.hclustered.df$Cluster)
 
-pdf("figures/Ef_most_sig_lifecycle_heatmap.pdf")
-#    height = 8, width = 8)
+pdf("figures/Ef_most_sig_lifecycle_heatmap.pdf",
+    height = 8, width = 8)
 pheatmap(Ef.cycle.diff.top.100.data,
          color = brewer.pal(n = 11, name = "BrBG"), 
          scale = "row",
@@ -42,12 +42,10 @@ pheatmap(Ef.cycle.diff.top.100.data,
          cluster_cols = T,
          annotation_row = Ef.hclustered.df,
          ## annotation_names_row = F,
-         cutree_rows = 5, 
+         cutree_rows = 7, # 5 in EH script 
          show_rownames = F,
          main = expression(paste(italic("E. falciformis"),
-             " mRNAs differently abundant between lifecycle stages")),
-         width = 8,
-         height = 8)
+             " mRNAs differently abundant between lifecycle stages")))
 dev.off()
 
 

### 2016-06-03 6_enrichment.R

@@ -1,6 +1,7 @@
 ## still open question: How to implement the FCS(?) method
 library(topGO)
 library(xtable)
+library(AnnotationDbi)
 
 if(!exists("hcluster")){
     source("5_clustering.R")
@@ -21,9 +22,13 @@ exp.universe[["Ef"]] <- names(gene2GO[["Ef"]])
 
 to.test <- as.data.frame(rbind(
     ## Eimeria
-    cbind(cluster=2,type="oocysts", species="Ef"),
-    cbind(cluster=5,type="sporozoites", species="Ef"),
-    cbind(cluster=1,type="day7", species="Ef"),
+    cbind(cluster=1,type="Day7Up_1", species="Ef"),
+    cbind(cluster=2,type="Day7Up_2", species="Ef"),
+    cbind(cluster=3,type="OocSporDown", species="Ef"),
+    cbind(cluster=4,type="OocystsUp", species="Ef"),
+    cbind(cluster=5,type="OoSpoDown_day5Up", species="Ef"),
+    cbind(cluster=6,type="SporozoitesUp", species="Ef"),
+    cbind(cluster=7,type="Day7OocyDown_SporoUP", species="Ef"),
     ## Mouse
     cbind(cluster=2,type="7dpiUp", species="Mm"),
     cbind(cluster=3,type="7dpiDown", species="Mm"),
@@ -31,7 +36,10 @@ to.test <- as.data.frame(rbind(
     cbind(cluster=4,type="allInfDown", species="Mm")
     ))
 
-apply(to.test, 1, function (x){
+
+## Subset BPMF.ll by [[1]][1[]][1[]] to get GO-terms (third bracket)  per cluster (first No.) and 
+## either BP (1) or MF (2) (second number).
+BPMF.ll <- apply(to.test, 1, function (x){
           clus = x[[1]]
           type = x[[2]]
           species = x[[3]]
@@ -39,13 +47,41 @@ apply(to.test, 1, function (x){
           hcl = hcluster[[species]] # hcluster object created in 5_clustering script
           ## creates first part of filename for each cluster
           file.path=("~/Ef_RNAseq/additional_files/tex/cluster")
-          sapply(c("MF", "BP"), function (onto){
-                     res <- TOGO.all.onto(onto, exp.universe[[species]],
-                                          names(hcl[hcl%in%clus]),g2G)
-                     file.detail= paste(clus, "_", onto, "_",
-                         species, "_", type, ".tex", sep="")
-                     capture.output(print(xtable(gene.table.topGO(res)),
[…]

### 2016-06-14 additional_files/tex/distribution-notes.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

--- 2016-02-15 additional_files/tex/file1_data_curation.tex
@@
 
+%%\usepackage{pdfpages}
+
 \setlength\parindent{0pt} % Removes all indentation from paragraphs
@@
 
-\title{Curating RNAseq data for mouse and \textit{Eimeria falciformis}} % Title
+\title{Gene expression distribution in RNAseq data for mouse and \textit{Eimeria falciformis}} % Title
 
-\author{Emanuel \textsc{Heitlinger}} % Author name
+\author{Totta \textsc{Kasemo}} % Author name
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
-\caption{Density distribution of raw transcript counts (log10) for all
-  mouse samples in analysis. All samples share the same bimodal
+  \includegraphics[width=0.8\textwidth]{distributions_mouseNocutoffg} 
+  \caption{Density distribution of raw transcript counts (log10) for
+    all mouse samples in analysis. All samples share the same bimodal
     distribution trend by visual inspection. The first density peak
-  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
-  detected between 1 and 100 times in that sample. The second density
-  peak occurs in the range of $10^3 - 10^4$.}
+    occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which
+    are detected between 1 and 100 times in that sample. The second
+    density peak occurs in the range of $10^3 - 10^4$.}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} 
 \caption{Density distribution of normalised transcript counts (log10)
@@
 \begin{center}
-  \includegraphics[width=0.8\textwidth]{distributions_normmouse100} % Include the image placeholder.png
-  \caption{Density distribution of normalised transcript counts
-    (log10) for mouse samples with a cutoff of 100 (see Methods for
+\includegraphics[width=0.8\textwidth]{distributions_normmouse100} 
+\caption{Density distribution of normalised transcript counts (log10)
+  for mouse samples with a cutoff of 100 (see Methods for
   details). All samples still have a bimodal distribution, however
-    less pronounced compared to figures 1 and 2. The first density
-    peak occurs at $<1$ transcripts as in Figure 2. The second density
-    peak occurs at close to $10^2$ transcripts and did hence not
-    change from the normalised data without cutoff (Figure 2).}
+  less pronounced compared to figures 1 and 2. The first density peak
+  occurs at $<1$ transcripts as in Figure 2. The second density peak
+  occurs at close to $10^2$ transcripts and did hence not change from
+  the normalised data without cutoff (Figure 2).}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_normmouse2000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_normmouse2000} 
 \caption{Density distribution of normalised transcript counts (log10)
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_normmouse5000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_normmouse5000}
 \caption{Density distribution of normalised transcript counts (log10)
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_mouse5000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_mouse5000}
 \caption{Density distribution of raw transcript counts (log10) for
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_normmouse10000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_normmouse10000}
 \caption{Density distribution of normalised transcript counts (log10)
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_Ef5} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_Ef5}
 \caption{Density distribution of raw transcript counts (log10) for
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_Efnorm5} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_Efnorm5}
 \caption{Density distribution of normalised transcript counts (log10)
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_Ef50000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_Ef50000}
 \caption{Density distribution of raw transcript counts (log10) for
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{distributions_Efnorm50000} % Include the image placeholder.png
+\includegraphics[width=0.8\textwidth]{distributions_Efnorm50000}
 \caption{Density distribution of normalised transcript counts (log10)
@@
 %\begin{item}
-%  The \emph{atomic weight of an element} is the relative weight of
-%  one of its atoms compared to C-12 with a weight of
-%  12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with
-%  a weight of 16.00. Atomic weight is also the average weight of all
-%  the atoms of that element as they occur in nature.
+%The \emph{atomic weight of an element} is the relative weight of one of its atoms compared to C-12 with a weight of 12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with a weight of 16.00.
+Atomic weight is also the average weight of all the atoms of that element as they occur in nature.
 %\end{item}
 %\begin{item}
-%  The \emph{units of atomic weight} are two-fold, with an identical
-%  numerical value. They are g/mole of atoms (or just g/mol) or
-%  amu/atom.
+%The \emph{units of atomic weight} are two-fold, with an identical numerical value. They are g/mole of atoms (or just g/mol) or amu/atom.
 %\end{item}
 %\begin{item}
-%  \emph{Percentage discrepancy} between an accepted (literature)
-%  value and an experimental value is
+%\emph{Percentage discrepancy} between an accepted (literature) value and an experimental value is
 %\begin{equation*}

### 2016-06-14 distribution-notes.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

--- 2016-02-15 additional_files/tex/file1_data_curation.tex
@@
 
-In the differential expression analysis of RNAseq data we assume a
-negative binomial distribution of transcript abundance. We analyse the
-distributions in mouse and parasite data respectively. A cutoff is
-applied to reads mapping per gene in raw data. For instance, a cutoff
-of ten requires that at least ten reads from \textit{any sample} map
-to a given gene for the gene to count as expressed. We evaluate the
-effects of different cutoffs on the distribution.
+In the differential expression analysis of RNAseq data we assume a negative binomial distribution of transcript abundance.
+We analyse the distributions in mouse and parasite data respectively.
+A cutoff is applied to reads mapping per gene in raw data.
+For instance, a cutoff of ten requires that at least ten reads from \textit{any sample} map to a given gene for the gene to count as expressed.
+We evaluate the effects of different cutoffs on the distribution.
 
@@
 \includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
-\caption{Density distribution of raw transcript counts (log10) for all
-  mouse samples in analysis. All samples share the same bimodal
-  distribution trend by visual inspection. The first density peak
-  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
-  detected between 1 and 100 times in that sample. The second density
-  peak occurs in the range of $10^3 - 10^4$.}
+\caption{Density distribution of raw transcript counts (log10) for all mouse samples in analysis.
+All samples share the same bimodal distribution trend by visual inspection.
+The first density peak occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are detected between 1 and 100 times in that sample.
+The second density peak occurs in the range of $10^3 - 10^4$.}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for all mouse samples. All samples share the same bimodal
-  distribution trend by visual inspection. The first density peak
-  occurs at $<1$ transcripts, and in most samples it is more distinct
-  than in the non-normalised data in Figure 1. For this normalised
-  data, the second density peak occurs at roughly the same count value
-  in all samples, close to $10^2$ transcripts.}
+\caption{Density distribution of normalised transcript counts (log10) for all mouse samples.
+All samples share the same bimodal distribution trend by visual inspection.
+The first density peak occurs at $<1$ transcripts, and in most samples it is more distinct than in the non-normalised data in Figure 1.
+For this normalised data, the second density peak occurs at roughly the same count value in all samples, close to $10^2$ transcripts.}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_normmouse100} % Include the image placeholder.png
-  \caption{Density distribution of normalised transcript counts
-    (log10) for mouse samples with a cutoff of 100 (see Methods for
-    details). All samples still have a bimodal distribution, however
-    less pronounced compared to figures 1 and 2. The first density
-    peak occurs at $<1$ transcripts as in Figure 2. The second density
-    peak occurs at close to $10^2$ transcripts and did hence not
-    change from the normalised data without cutoff (Figure 2).}
+\caption{Density distribution of normalised transcript counts (log10) for mouse samples with a cutoff of 100 (see Methods for details).
+All samples still have a bimodal distribution, however less pronounced compared to figures 1 and 2.
+The first density peak occurs at $<1$ transcripts as in Figure 2.
+The second density peak occurs at close to $10^2$ transcripts and did hence not change from the normalised data without cutoff (Figure 2).}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_normmouse2000} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for mouse samples with a cutoff of 2000 (see Methods for
-  details). The bimodality seen at lower cutoff values is strongly
-  weakened or not detectable visually in some samples. It is however
-  visually still seen in a majority of the samples. The first tendency
-  to a density peak has shifted somewhat towards higher transcript
-  counts (right) but is still $<1$ transcripts as in Figure 3. The
-  second density peak occurs at close to $10^2$ transcripts and did
-  hence not change from the normalised data with a lower cutoff
-  (Figure 3).}
+\caption{Density distribution of normalised transcript counts (log10) for mouse samples with a cutoff of 2000 (see Methods for details).
+The bimodality seen at lower cutoff values is strongly weakened or not detectable visually in some samples.
+It is however visually still seen in a majority of the samples.
+The first tendency to a density peak has shifted somewhat towards higher transcript counts (right) but is still $<1$ transcripts as in Figure 3.
+The second density peak occurs at close to $10^2$ transcripts and did hence not change from the normalised data with a lower cutoff (Figure 3).}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_normmouse5000} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for mouse samples with a cutoff of 5000 (see Methods for
-  details). The bimodality is not visually detectable in most samples,
-  however it is still seen in \textit{e.g.} samples
-  NMRI\_1stInf\_3dpi\_rep1 and NMRI\_2ndInf\_7dpi\_rep2. The major
-  density peak remains at a value close to $10^2$ as in all the
-  analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
+\caption{Density distribution of normalised transcript counts (log10) for mouse samples with a cutoff of 5000 (see Methods for details).
+The bimodality is not visually detectable in most samples, however it is still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep1 and NMRI\_2ndInf\_7dpi\_rep2.
+The major density peak remains at a value close to $10^2$ as in all the analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_mouse5000} % Include the image placeholder.png
-\caption{Density distribution of raw transcript counts (log10) for
-  mouse samples with a cutoff of 5000 (see Methods for details). The
-  bimodality is not visually detectable in most samples, however it is
-  still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep2 and
-  NMRI\_2ndInf\_7dpi\_rep2. The major density peak is seen at a value
-  over $10^2$, shifting it almost one order of magnitude compared to
-  the normalised data with the same cutoff (Figure 5).}
+\caption{Density distribution of raw transcript counts (log10) for mouse samples with a cutoff of 5000 (see Methods for details).
+The bimodality is not visually detectable in most samples, however it is still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2.
+The major density peak is seen at a value over $10^2$, shifting it almost one order of magnitude compared to the normalised data with the same cutoff (Figure 5).}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_normmouse10000} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for mouse samples with a cutoff of 10000 (see Methods for
-  details). The bimodality is not visually detectable in any sample
-  apart from a minor indication in sample
-  NMRI\_1stInf\_5dpi\_rep2. The major density peak remains at a value
-  close to $10^2$ as in all the analysed normalised datasets
-  (\textit{e.g.} Figure 2 or 4).}
+\caption{Density distribution of normalised transcript counts (log10) for mouse samples with a cutoff of 10000 (see Methods for details).
+The bimodality is not visually detectable in any sample apart from a minor indication in sample NMRI\_1stInf\_5dpi\_rep2.
+The major density peak remains at a value close to $10^2$ as in all the analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_Ef5} % Include the image placeholder.png
-\caption{Density distribution of raw transcript counts (log10) for
-  parasite samples with a cutoff of 5 (see Methods for details). Upon
-  visual inspection distributions in most samples appear smooth and
-  not contradictory to a negative binomial distribution. Density peak
-  positions with regards to transcript counts (x-axis) are however at
-  different orders of magnitudes between samples. Samples
-  NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
-  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2, however, do
-  not.}
+\caption{Density distribution of raw transcript counts (log10) for parasite samples with a cutoff of 5 (see Methods for details).
+Upon visual inspection distributions in most samples appear smooth and not contradictory to a negative binomial distribution.
+Density peak positions with regards to transcript counts (x-axis) are however at different orders of magnitudes between samples.
+Samples NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2, however, do not.}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_Efnorm5} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for parasite samples with a cutoff of 5 (see Methods for
-  details). Upon visual inspection distributions in most samples
-  appear smooth also in the normalised data. Density peaks are also in
-  the same order of magnitude with regards to transcript counts
-  (x-axis). Non-negative binomial samples in Figure 8
-  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
-  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) have more
-  pronounced spikes in this normalised data. Additionally, samples
-  Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 also have one
-  additional peak in the lower range of the distribution, as do some
-  other samples, but less pronounced.}
+\caption{Density distribution of normalised transcript counts (log10) for parasite samples with a cutoff of 5 (see Methods for details).
+Upon visual inspection distributions in most samples appear smooth also in the normalised data.
+Density peaks are also in the same order of magnitude with regards to transcript counts (x-axis).
+Non-negative binomial samples in Figure 8 (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) have more pronounced spikes in this normalised data.
+Additionally, samples Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 also have one additional peak in the lower range of the distribution, as do some other samples, but less pronounced.}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_Ef50000} % Include the image placeholder.png
-\caption{Density distribution of raw transcript counts (log10) for
-  parasite samples with a cutoff of 50000 (see Methods for
-  details). Upon visual inspection distributions in most samples
-  appear smooth. Non-negative binomial samples in Figure 8 and 9
-  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
-  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2 have more than
-  one density peak also when lowly expressed genes are removed with a
-  high threshold (50000). Samples Rag\_1stInf\_5dpi\_rep1 and
-  Rag\_1stInf\_5dpi\_rep2, however, seem smoother with only one
-  density peak with this threshold.}
+\caption{Density distribution of raw transcript counts (log10) for parasite samples with a cutoff of 50000 (see Methods for details).
+Upon visual inspection distributions in most samples appear smooth.
+Non-negative binomial samples in Figure 8 and 9 (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2 have more than one density peak also when lowly expressed genes are removed with a high threshold (50000).
+Samples Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2, however, seem smoother with only one density peak with this threshold.}
 \end{center}
@@
 \includegraphics[width=0.8\textwidth]{distributions_Efnorm50000} % Include the image placeholder.png
-\caption{Density distribution of normalised transcript counts (log10)
-  for parasite samples with a cutoff of 50000 (see Methods for
-  details). Upon visual inspection distributions in most samples
-  appear smooth. Non-negative binomial samples are pronounced bimodal
-  in this normalised data (NMRI\_2ndInf\_3dpi\_rep1,
-  NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and
-  NMRI\_2ndInf\_7dpi\_rep2). Samples Rag\_1stInf\_5dpi\_rep1 and
-  Rag\_1stInf\_5dpi\_rep2 are also bimodal, however less pronounced.}
+\caption{Density distribution of normalised transcript counts (log10) for parasite samples with a cutoff of 50000 (see Methods for details).
+Upon visual inspection distributions in most samples appear smooth.
+Non-negative binomial samples are pronounced bimodal in this normalised data (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2).
+Samples Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 are also bimodal, however less pronounced.}
 \end{center}
@@
 Mouse\\
-The assumption of a negative binomial distribution underlies most
-differential gene expression analyses available. Similar bimodal
-expression patterns as in our mouse dataset have been reported
-(\textit{e.g.} ....) and suggested to be a result of a group of
-non-coding RNAs being sequenced. Since we are not interested in such
-RNA species in this project, different cutoff values were tested to
-find the lowest possible threshold for removing transcripts and gain a
-negative binomial distribution. By visual inspection (again: should we
-formalise/quantify this somehow?) we chose 5000 as the cutoff for
-further analysis.\\
+The assumption of a negative binomial distribution underlies most differential gene expression analyses available.
+Similar bimodal expression patterns as in our mouse dataset have been reported (\textit{e.g.} ....) and suggested to be a result of a group of non-coding RNAs being sequenced.
+Since we are not interested in such RNA species in this project, different cutoff values were tested to find the lowest possible threshold for removing transcripts and gain a negative binomial distribution.
+By visual inspection (again: should we formalise/quantify this somehow?) we chose 5000 as the cutoff for further analysis.
+\\
 
 Parasite\\
-The parasite data follows the expected distribution in most samples
-(although we have not formally tested it - check mean-var relationship
-to do that?). Bimodal or multipeak samples do not show unimodal
-distributions even with high cutoffs, \textit{i.e.} many transcripts
-required to map to one gene for the reads to be included. The four
-most obvious spiky samples are also the samples with fewest reads
-sequenced overall, which can explain the uneven distribution (make
-table to refer to?). We will exclude these four samples
-(NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
-NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) from all
-further analysis. It is not clear why there is such low coverage in
-these samples. Possibilities include poor infection success because of
-too low oocyst injection, mouse resistance due to \textit{e.g.}
-unknown previous infections or other unknown variability in the
-experiment. Because of this insecurity, these samples are also
-exculded from the mouse data.\\
+The parasite data follows the expected distribution in most samples (although we have not formally tested it - check mean-var relationship to do that?).
+Bimodal or multipeak samples do not show unimodal distributions even with high cutoffs, \textit{i.e.} many transcripts required to map to one gene for the reads to be included.
+The four most obvious spiky samples are also the samples with fewest reads sequenced overall, which can explain the uneven distribution (make table to refer to?).
+We will exclude these four samples (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) from all further analysis.
+It is not clear why there is such low coverage in these samples.
+Possibilities include poor infection success because of too low oocyst injection, mouse resistance due to \textit{e.g.} unknown previous infections or other unknown variability in the experiment.
+Because of this insecurity, these samples are also exculded from the mouse data.
+\\
 
-The fact that the density peak occurs at similar read counts after
-normalisation in mouse and parasite data (respectively) is encouraging
-and indicates that normalisation is successful in both datasets.
+The fact that the density peak occurs at similar read counts after normalisation in mouse and parasite data (respectively) is encouraging and indicates that normalisation is successful in both datasets.
 

### 2016-06-14 manuscript/tex_bmc_format/manuscript_rnaseq.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

@@
 \subsection*{GO term enrichments in heatmap clusters}
-The annotations referred to here are in many 
-cases inferred from the existence of predicted protein domains and have in some cases been 
-verified in e.g. \textit{E. tenella} or other \textit{Eimeria} spp.
+The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
+in \textit{T. gondii}
 
@@
 The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
-Overrepresented GO-term in this cluster are largely determined by genes orthologous to 
+Overrepresented GO-terms in this cluster are enriched by ortholog genes to 
 peptidases, microneme localized proteins reported to be involved in invasion, 
-genes associated with clotting and in protozoans, adhesion, and genes that are annotated 
-to be involved in amino acid biosynthesis.
-  Aminopeptidase N ('related' annotation) is the annotation for orthologs of three genes with 
+genes associated with adhesion in protozoans (and with clotting in higher eukaryotes),
+and genes that are annotated to be involved in amino acid biosynthesis.
+  Aminopeptidase N ('related' annotation) is the reported ortholog for three genes with 
 abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
@@
 
-Three genes can be linked to adhesion and through localization the same genes
-also are asociated with invasion in the literature. 
-A protein domain which is associated with adhesion in protozoans, thrombospondin type 1 
-domain-containing protein, was found in our oocyst cluster (cluster 1). 
-Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme localizing 
-proteins, MIC. Tomley et al. (Tomley01) describe MIC4 as a protein contatining thrombospondin 
-type 1 repeats. In \textit{E. tenella} MIC4 is expressed in sporozoites where it localizes to the 
-apical end, in late schizonts and late oocyst stages, when sporozoites are forming.
-(Tomley01)
-In the protozoan \textit{P. falciparum}, another domain occuring in our data, the sushi domain, 
-is reported in the apical sushi protein, ASP. In \textit{P. falciparum} it localizes to the 
-micronemes in merozoites but not other stages (OKeeffe05). To our knowledge, the role of sushi 
-domain proteins in \textit{Eimeria} spp. oocysts has not yet been investigated. 
-  Other protein domains detected via orthologs to mRNAs expressed in oocysts in our data are
-Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, also known as F5/8 domain. 
-LCCL domains have  been associated with discoidin lectin domains, which in turn are involved 
-in cell adhesion (Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
-In the slime mold \textit{Dictyostelium discoideum)} the LCCL domain
-is part of the mold'd discoidin adhesion protein. In humans, the LCCL domain in annotated as a 
-coagulation factor domain. (Pfam entry for 'discoidin domain', May 2016)
+A Thrombospondin type 1 domain-containing protein ortholog is highly abundant in cluster 4 (high 
+abundance in oocysts). Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme
+localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella} MIC4 mRNA was reported in 
+sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages,
+when sporozoites are forming. (Tomley01). For the same gene, the \textit{T. gondii} annotation is 
+Sushi domain-containing protein, which is also the ortholog annotation of another gene in this 
+cluster. In the related malaria parasite \textit{P. falciparum} the apical sushi 
+protein, ASP, (which has a sushi domain) localizes to micronemes in merozoites but not 
+other stages (OKeeffe05). 
+Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, 
+(syn. F5/8 domain) domains are associated discoidin lectin domains and thereby with adhesion.
+(Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
+Taken together, this indicates that the thrombospondin and sushi-domain genes 
+(EfaB_MINUS_4114.g412 and EfaB_PLUS_1425.g183) are involved in sporozoite invasion in 
+\textit{E. falciformis} and that the mRNAs are transcribed and available before excystation. 
+A role in merozoite re-invasion in \textit{E. falciformis} is not indicated by our data.
+The LCCL domain annotation and thrombospondins role in higher eukaryotes also indicates 
+that adhesion or preparation for adhesion is important in oocysts. 
+We suggest (speculate...?) that the thrombospondin annotated ortholog and the LCCL domain-containing protein 
+(EfaB_MINUS_11233.g986) are involved in cell adhesion in \textit{E. falciformis}.
+%One examples
+%is found in the slime mold \textit{Dictyostelium discoideum)} in which the LCCL domain
+%is part of the mold's discoidin adhesion protein. (Pfam entry for 'discoidin domain', May 2016)
 
 \subsubsection*{Amino acid biosynthesis in oocysts}
-mRNA of a aminotransferases indicate amino acid biosynthesis in oocysts. 
-We also identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase, which are enzymes 
-contributing to L-serine and L-alanine production, respectively. A putative \textit{Eimeria} spp.
-cystathionine beta-synthase, CBS, in this cluster also indicates de novo cysteine production. 
-Alkyl sulfatase mRNA is another gene which contributes to overrepresentation of a GO-term in 
-mRNAs highly abundant in oocysts. Generally, this enzyme enables an organism to exploit organic 
-sulfur to produce and incorporate inorganic sulfur into the amino acids cysteine and methionine, 
-when no inorganoc sulfur is available.
+High abundance of aminotransferase mRNAs indicate amino acid biosynthesis or preparation for 
+the same in oocysts (cluster 4). 
+We identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase orthologs, which 
+are enzymes contributing to L-serine and L-alanine production, respectively. A putative 
+\textit{Eimeria} spp. cystathionine beta-synthase, CBS, in this cluster also indicates 
+de novo cysteine production. Alkyl sulfatase mRNA is also abundant in oocysts. Generally, 
+this enzyme enables an organism to exploit organic sulfur to produce and incorporate 
+inorganic sulfur into the amino acids cysteine and methionine, when no inorganoc sulfur 
+is available.
 
@@
 
+\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
+MmgE/PrpD is overrepresented in oocysts. The enzyme is important for propionate catabolism 
+in the 2-methylcitric acid cycle and has been shown to be used by the 
+intestinal intracellular bacterium \textit{Salmonella typhimurium} to generate 
+pyruvate (Horswill99). 
 Propionate is one of two most abundant small-chain fatty acids 
-in the gut along with butyrate. Both fatty acids are largely produced as degradation products from food
-by commensal bacteria (Sun13). MmgE/PrpD is important for propionate catabolism in the 
-2-methylcitric acid cycle and has been shown to be used by the intestinal intracellular bacterium 
-\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).
+in the gut along with butyrate. Both fatty acids are largely produced as 
+degradation products from food by commensal bacteria (Sun13). Sharing the intestines as a niche
+with \textit{S. typhimurium} it is possible that also \textit{E. falciformis} uses Mmg/PrpD
+to exploit available propionate for pyruvate production.
+
+\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
+Interestingly, the genes described above which are thought to be involved in amino acid 
+biosynthesis and invasion are highly abundant in oocysts but are underrepresented in schizont
+stages (day 3 and day 5 samples) and even in sporozoites. An average abundance was detected on 
+day 7 for these genes, indicating a role in either gametes or early oocyst formation.
+This pattern supports the suggestion that these specific mRNAs (cluster 4) for invasion 
+and biosynthetic processes are prepared (and possibly expressed) in the oocyst stage but 
+are no longer detectable in the cell at the timepont when the protein is assumed to be 
+in use (sporozoites and merozoite stages). 
+Therefore, correlating mRNA prevalence with biological function at the timepoint when mRNAs 
+are detected must be done with care.
+
+\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
+Oocysts: profile for 3, 5, 6:
+
+Sporozoites: 3 and 5
+
+Day 7: 5 and 7
+
+Specific genes....
+Enolase 2, encoded by Eno2, is among the downregulated genes in oocysts and sporozoites.
+In \textit{T. gondii} the paralog Eno1 is strongly associated with the cyst (bradyzoite) stage
+and Eno2 is associated with tachyzoite stages (kibe05). It is therefore expected that this mRNA is
+underrepresented in oocysts and our data also show that the same is true in sporozoites for
+\textit{E. falciformis}. (TK: If important we could look specifically for Eno1 in cluster 4).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{cluster 3..........}
+\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
 

### 2016-06-14 5_clustering.R

@@ -9,12 +9,28 @@ if(!exists("RNAseq.Array.logFC")){
 }
 ########## USE THIS #######################
 ########## PHEATMAP() specifying distance and method ############
-
+## 1:10 refers to the Ef.contrast object and the first ten 
+## comparisons in it.
+############# LIFE CYCLE SELECTION ##################
 union.of.Ef.cycle.diff.top.100 <-
     unique(unlist(lapply(Ef.1st.pass.model[[3]][1:10], head, n=500)))
 
 Ef.cycle.diff.top.100.data <-
     cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff.top.100,]
+
+############# 1st VERSUS 2nd INFECTION ##################
+union.of.Ef.1st2nd.top <-
+    unique(unlist(lapply(Ef.1st.pass.model[[3]][11:15], head, n=500)))
+
+Ef.1st2nd.top.data <-
+    cpm(Ef.1st.pass.model[[4]])[union.of.1st2nd.top,]
+
+######## EIMERIA DEVELOPMENT IN DIFFERENT MOUSE STRAINS (day 5 only) ##################
+union.of.Ef.strain.depend.top <-
+    unique(unlist(lapply(Ef.1st.pass.model[[3]][16:18], head, n=500)))
+
+Ef.strain.depend.top.data <-
+    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.strain.depend.top,]
 
 get.scaled.and.clustered <- function(data){
     rMeans <- rowMeans(data)
@@ -25,6 +41,9 @@ get.scaled.and.clustered <- function(data){
 }
 
 Ef.hclustered <- get.scaled.and.clustered(Ef.cycle.diff.top.100.data)
+Ef.hclustered.first.second <- get.scaled.and.clustered(Ef.1st2nd.top.data) 
+Ef.hclustered.strain <- get.scaled.and.clustered(Ef.strain.depend.top.data )
+
 
 hcluster <- list()
 hcluster[["Ef"]] <- cutree(Ef.hclustered, k= 7) ## TK changed k from 5 ## h=7.2)

### 2016-06-16 additional_files/tex/SI_clusters/table_rawcounts.tex
table_rawcounts added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
\usepackage{booktabs}
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font

%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{Overview of RNA-seq reads for \textit{Eimeria falciformis} and \textit{Mus musculus}} % Title

\author{Totta \textsc{Kasemo}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Objective}

This summary table of reads generated in each sample.....

%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION 2 
%----------------------------------------------------------------------------------------

\section{Table of reads per sample}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Wed Mar  2 14:36:30 2016
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
Samples & Percent \emph{E. falciformis} & Mouse Transcripts & \emph{E. falciformis} Transcripts \\
  \hline
NMRI\_oocysts &  99.9854842 & 17516.000 & 120650726.000 \\
  NMRI\_sporozoites &  99.9805302 & 20860.000 & 107119299.000 \\
  NMRI\_1stInf\_7dpi\_rep1 &  83.5224277 & 18350269.000 & 93014856.000 \\
  NMRI\_1stInf\_7dpi\_rep2 &  54.9450408 & 145227236.000 & 177106284.000 \\
  NMRI\_2ndInf\_7dpi\_rep1 &   7.0575562 & 149369191.000 & 11342304.000 \\
  NMRI\_1stInf\_5dpi\_rep1 &   5.9692983 & 320787124.000 & 20364349.000 \\
  C57BL6\_1stInf\_5dpi\_rep2 &   0.5892196 & 46455447.000 & 275347.000 \\
  NMRI\_2ndInf\_5dpi\_rep1 &   0.4501036 & 917436287.000 & 4148084.000 \\
  NMRI\_1stInf\_5dpi\_rep2 &   0.3615460 & 490566909.000 & 1780061.000 \\
  C57BL6\_2ndInf\_5dpi\_rep1 &   0.3550692 & 95259948.000 & 339444.000 \\
  NMRI\_1stInf\_3dpi\_rep1 &   0.2721386 & 328870247.000 & 897425.000 \\
  C57BL6\_1stInf\_5dpi\_rep1 &   0.2347177 & 101231812.000 & 238168.000 \\
  Rag\_2ndInf\_5dpi\_rep1 &   0.1850507 & 132358804.000 & 245385.000 \\
  Rag\_1stInf\_5dpi\_rep2 &   0.1427648 & 88257677.000 & 126181.000 \\
  NMRI\_1stInf\_3dpi\_rep2 &   0.1216664 & 809232397.000 & 985763.000 \\
  Rag\_1stInf\_5dpi\_rep1 &   0.0712794 & 112174061.000 & 80014.000 \\
  NMRI\_2ndInf\_7dpi\_rep2 &   0.0271172 & 193993643.000 & 52620.000 \\
  NMRI\_2ndInf\_3dpi\_rep2 &   0.0235337 & 162464901.000 & 38243.000 \\
  NMRI\_1stInf\_0dpi\_rep1 &   0.0108286 & 347939504.000 & 37681.000 \\
  NMRI\_2ndInf\_5dpi\_rep2 &   0.0079426 & 178050386.000 & 14143.000 \\
  NMRI\_2ndInf\_3dpi\_rep1 &   0.0043015 & 142131560.000 & 6114.000 \\
  NMRI\_2ndInf\_0dpi\_rep2 &   0.0030853 & 140081531.000 & 4322.000 \\
  Rag\_1stInf\_0dpi\_rep1 &   0.0008362 & 67091135.000 & 561.000 \\
  Rag\_1stInf\_0dpi\_rep2 &   0.0007451 & 126286703.000 & 941.000 \\
  C57BL6\_1stInf\_0dpi\_rep1 &   0.0007381 & 82105133.000 & 606.000 \\
  C57BL6\_1stInf\_0dpi\_rep2 &   0.0006947 & 119614354.000 & 831.000 \\
  NMRI\_2ndInf\_0dpi\_rep1 &   0.0000834 & 441473116.000 & 368.000 \\

  \hline
\end{tabular}
\caption{Transcripts per sample listed for \textit{E. falciformis}, mouse and for the percentage of \textit{E. falciformis} transcripts out of the total (mouse plus parasite) transcripts. }
\end{table}

\section{Experimental overview with transcripts per sample}
	\setlength{\tabcolsep}{14pt}
\begin{table}
	\begin{center}
%\begin{tabular}{@{} *4l @{}}    \toprule
\begin{tabular}{*4l}    \toprule
\emph{Day, 1st infection}  	& NMRI  & C57BL/6  & Rag1-/- \\ \midrule
0 (control)     & $10^4$ / $10^8$  & $10^2$ / $10^8$  & $10^2$ / $10^7$  \\ %rep1/rep2, NA=no replicate
3  		& $10^5$ / $10^8$ & NA  & NA \\ 
5  		& $10^7$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^8$ \\
7  		& $10^8$ / $10^7$ & NA  & NA \\ 
Oocysts 	& $10^8$ / NA & NA  & NA \\ 
Sporozoites 	& $10^8$ / NA & NA  & NA \\ \midrule

\emph{Day, 2nd infection}  	\\ \midrule
0 (control)     & $10^x$ / $10^x$  & $10^x$ / $10^x$  & $10^x$ / $10^x$  \\ %rep1/rep2, NA=no replicate
3  		& $10^x$ / $10^x$ & NA  & NA \\ 
5  		& $10^x$ / $10^x$ & $10^x$ / $10^x$  & $10^x$ / $10^x$ \\
7  		& $10^x$ / $10^x$ & NA  & NA \\

	\bottomrule
 \hline
\end{tabular}
	\caption{Transcripts numbers as order of magnitude for \emph{E. falciformis} (first) and mouse (second). Replicate average shown. For exact values, see ... . }
\end{center}
\end{table}

\section{Mouse RNA-seq data compared with mouse microarray data}
% figure 1
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{~/Ef_RNAseq/figures/Array144_vs_RNAseqN7} % Merge 2 plots: RUV and non-RUV 
\caption{Comparison of mouse data from RNA-seq, day 7, (y-axis) and microarray data, day 6 (x-axis, (schmid14)). a. Data normalised with upperquartile method implemented in R package edgeR (version ...). $R^2$ = yy. b. Data adjusted with RUV method as implemented in R package RUVseq (version....). $R^2$ = xx. Each axis shows log fold changes on day 6 (microarray) and 7 (RNA-seq) post infection compared to control.}
\end{center}
\end{figure}

% figure 1
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{~/Ef_RNAseq/figures//old_figures/EfMm_2mds} % Include 
	\caption{Multidimensional scaling shows no clear sample pattern due to technical variation. Samples shown with a. batch and b. sequencing method visualized for mouse (upper) and \emph{E. falciformis} (lower).}
\end{center}
\end{figure}

\end{document}

### 2016-06-16 1_ballgown_import.R
added automatic check for and install packages if not on machine to all scripts + updated gitignore and made small manuscript changes

@@ -1,5 +1,9 @@
 ### ballgown for data import from the cufflinks/tablemaker pipeline
 
+# Check package for package control.
+source("https://bioconductor.org/biocLite.R")
+if(!require(ballgown)) biocLite("ballgown") # Imports package if user does not have it
+
 library(ballgown)
 setwd("~/Ef_RNAseq")
 

### 2016-06-16 2_edgeR_diff.R

@@ -6,6 +6,13 @@ if(!exists("Mm.RC")|!exists("Ef.RC")){
     source("1_ballgown_import.R")
 }
 
+# Check checkpoint package for package control.
+if(!require(edgeR)) biocLite("edgeR") # Imports package if user does not have it
+if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
+if(!require(RUVSeq)) biocLite("RUVSeq") # Imports package if user does not have it
+if(!require(reshape)) biocLite("reshape") # Imports package if user does not have it
+
+
 library(edgeR)
 library(GGally)
 library(RUVSeq)

### 2016-06-16 3_annotations.R

@@ -6,6 +6,9 @@ if(!exists("Ef.1st.pass.model")){ #Totta changes Mm to Ef here
     source("2_edgeR_diff.R")
 }
 
+# Check checkpoint package for package control.
+if(!require(Mus.musculus)) biocLite("Mus.musculus") # Imports package if user does not have it
+
 ## The Bioconductor mouse annotation libary
 library(Mus.musculus)
 

### 2016-06-16 4_compareArray.R

@@ -4,8 +4,16 @@
 ## of IFNg-regulated host pathways for the parasite development"
 ## DOI:10.1038/mi.2013.115
 
+# Check checkpoint package for package control.
+if(!require(Biobase)) biocLite("Biobase") # Imports package if user does not have it
+if(!require(RSvgDevice)) biocLite("RSvgDevice") # Imports package if user does not have it
+if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
+if(!require(ggplot2)) biocLite("ggplot2") # Imports package if user does not have it
+## For below package does not always work: then manual import is necessary
+if(!require(mgug4122a.db)) biocLite("mgug4122a.db") # Imports package if user does not have it
+
 library(mgug4122a.db)
-library(Agi4x44PreProcess)
+library(Agi4x44PreProcess) # !!!! NOT available for R 3.3.0 - NOT on Bioconductor anymore
 library(Biobase)
 library(RSvgDevice)
 library(GGally)

### 2016-06-16 5_clustering.R

@@ -1,5 +1,10 @@
 ## Making heatmaps and other clusters of gene expression data
 
+# Check checkpoint package for package control.
+if(!require(pheatmap)) biocLite("pheatmap") # Imports package if user does not have it
+if(!require(made4)) biocLite("made4") # Imports package if user does not have it
+if(!require(RColorBrewer)) biocLite("RColorBrewer") # Imports package if user does not have it
+
 library(pheatmap)
 library(made4)
 library(RColorBrewer)

### 2016-06-16 6_enrichment.R

@@ -1,4 +1,9 @@
 ## still open question: How to implement the FCS(?) method
+# Check checkpoint package for package control.
+if(!require(topGO)) biocLite("topGO") # Imports package if user does not have it
+if(!require(xtable)) biocLite("xtable") # Imports package if user does not have it
+if(!require(AnnotationDbi)) biocLite("AnnotationDbi") # Imports package if user does not have it
+
 library(topGO)
 library(xtable)
 library(AnnotationDbi)

### 2016-06-22 figures/table_rawcounts.tex
Totta update with SI material and trouble-shooting changes in scripts after harriet move

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
\usepackage{booktabs} % For prettier tables
\usepackage{float} % To control positions of e.g. tables or figures
\usepackage[colorlinks]{hyperref} % To color blocks of text 
\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
\usepackage{changepage}
%\usepackage[dvipdfmx]{graphicx}
%\usepackage{bmpsize} % Added at some point but now (June-16) it seems to do nothing
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\graphicspath{{home/totta/Ef\_RNAseq/figures/}}
\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font
%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{Overview of dual RNA-seq data from \textit{Eimeria falciformis} infected \textit{Mus musculus}} % Title

\author{Totta \textsc{Kasemo}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Objective}

This summary table of reads generated in each sample.....

%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION 2 
%----------------------------------------------------------------------------------------

\section{Table of reads per sample}
% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Wed Mar  2 14:36:30 2016
\setlength{\tabcolsep}{10pt}
\begin{table}[H]
\centering 
\caption{Transcripts per sample listed for \textit{E. falciformis} and mouse. Percentage of
	\textit{E. falciformis} transcripts and number of \textit{E. falciformis} genes are alo shown. 
	Gray indicates that samples are excluded from analysis (see Methods for details).}
\begin{tabular}{*5l}    \toprule
Samples  & Mouse 	& \textit{E. falciformis}  & Percent			& \#\textit{E.falciformis} \\
	& trancripts	& transcripts   	 & \textit{E. falciformis}	& genes \\ \midrule
NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
	{\color{Gray}NMRI\_2ndInf\_5dpi\_rep1} & {\color{Gray} 589643389.000} & {\color{Gray}3752923.000} & {\color{Gray} 0.632} &  {\color{Gray}5602.000} \\ 
NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
	{\color{Gray}NMRI\_1stInf\_3dpi\_rep1} & {\color{Gray}209723287.000} &  {\color{Gray}815820.000} &  {\color{Gray}0.388} &  {\color{Gray}5466.000} \\ 
C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
	 {\color{Gray}NMRI\_1stInf\_0dpi\_rep1} &  {\color{Gray}229165002.000} & {\color{Gray} 31459.000} &  {\color{Gray}0.014} & {\color{Gray} 1380.000} \\ 
NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
\bottomrule
\hline
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental overview with transcripts per sample}
	\setlength{\tabcolsep}{14pt}
	\begin{table}[H]
	\begin{center}
	\caption{Replicate average transcripts numbers as order of magnitude for \textit{E. falciformis}
		(first) and mouse (second) and experimental overview. Columns represent different mouse strains
		used (NMRI, C57BL/6 and Rag1-/-). Rows are different timepoints post infection or parasite life cycle 
		stages. The upper part shows data for first infection, and oocyst and sporozoite data. The lower part 
		shows data for challange infection. Averages were calculated after sample exclusions (see Methods). 
		For exact values, see Table 1.}
\begin{tabular}{*4l}    \toprule
\textit{Day, 1st infection}  	& NMRI  & C57BL/6  & Rag1-/- \\ \midrule
	0 (control)    & $10^4$ / $10^8$  & $10^2$ / $10^8$  & $10^2$ / $10^7$  \\ %rep1/rep2, NA=no replicate
3  		& $10^5$ / $10^8$ & NA  & NA \\ 
5  		& $10^7$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
7  		& $10^8$ / $10^7$ & NA  & NA \\ 
Oocysts 	& $10^8$ / NA & NA  & NA \\ 
Sporozoites 	& $10^7$ / NA & NA  & NA \\ \midrule

\textit{Day, 2nd infection}  	\\ \midrule
0 (control)     & $10^3$ / $10^8$  &  NA  & NA  \\ %rep1/rep2, NA=no replicate
3  		& $10^4$ / $10^8$ & NA  & NA \\ 
5  		& $10^4$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
7  		& $10^6$ / $10^8$ & NA  & NA \\

	\bottomrule
 \hline
\end{tabular}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MICROARRAY COMPARISON
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transcript distribution and excluding samples}
\subsection{Mouse transcript distributions}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{distributionsMm.pdf}
\caption{Density of transcripts (y-axis) with a specific number of reads mapping (x-axis). 
Density of raw transcripts (not normalised) for all mouse samples. Without any requirement 
(cutoff) for a minimum read coverage per transcript, all samples display a bimodal distribution 
trend and an additional peak at zero by visual inspection. A minimal number of total reads 
(sum for all samples) mapping to one transcript was set (cutoff). For further analysis, a cutoff of
2000 was applied with which for the edgeR anaylsis assumed negative binomial distribution is not 
contradicted. For the density distribution analysis, 0.1 was added to all read values to allow plotting 
density of absent transcripts on a log-scale.}
%\end{center}
\end{figure}

\clearpage
\subsection{\textit{E. falciformis} transcript distributions}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{distributionsEf.pdf}
\caption{Density of transcripts (y-axis) with a specific number of reads mapping (x-axis). 
Density of raw transcripts (not normalised) for all \textit{E. falciformis} samples. Without any
requirement (cutoff) for a minimum read coverage per transcript, most samples do not contradict
the assumed negative binomial distribution. However, two samples clearly do and a cutoff was applied 
to evaluate whether peaks can be removed by excluding transcripts with very low coverage.
The cutoff does not have a large effect on distributions of \textit{E. falciformis} data and therefore 
two samples were removed from further analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1.
	To all read values, 0.1 was added to allow plotting density of absent transcripts on a log-scale.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MICROARRAY COMPARISON
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mouse RNA-seq data compared with mouse microarray data}
%%% figure 1
\begin{figure}[H]
\begin{center}
%\includegraphics[width=\textwidth]{Array144_vs_RNAseqN7.pdf} % Merge 2 plots: RUV and non-RUV 
\caption{Comparison of mouse data from RNA-seq, day 7, (y-axis) and microarray data, day 6 (x-axis, (schmid14)). 
	a. Data normalised with upperquartile method implemented in R package edgeR (version ...). $R^2$ = yy. 
	b. Normalised data as i a. and adjusted with RUV method as implemented in R package RUVseq (version....). 
	$R^2$ = xx. Each axis shows log fold changes compared to control.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MDS ANALYSIS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysing variance in data by multidimensional scaling}
\begin{figure}[H]
\begin{center}
	\includegraphics[width=\textwidth]{EfMm_4-mds.pdf} % Include 
	\caption{Multidimensional scaling shows no clear sample pattern due to technical variation. Effect on scaling is visualized 
	for batch (left) and sequencing method (right). Upper panel displays \textit{E. falciformis} data and lower panel mouse data. 
	Distances on plot approximate log2 fold changes. Scaling was done with Euklidean distance and "pairwise" gene selection 
	method as implemented in R package limma (version ...).}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{mRNA abundance differences between different experimental groups}
\setlength{\tabcolsep}{14pt}
\begin{table}[H]
\begin{center}
\caption{\textit{E. falciformis} data overview of pairwise comparisons and number of genes with differently abundant mRNAs 
	per comparison. NMRI followed by number indicates day post infection (e.g. NMRI3 = \textit{E. falciformis} genes from
	NMRI mouse day 3 post infection). Genes with Benjamini-Hochberg corrected p-values \leq0.01 as implemented in edgeR are indluced. 
	ADD MOUSE DATA COLUMN.}
%\label{tab:table3}
\begin{tabular}{*2l}    \toprule
\textit{Comparisons with difference} & Genes different (FDR\leq0.01) \\ \midrule
	NMRI3 vs NMRI5     	& 111  \\
	NMRI3 vs NMRI7  	& 1385 \\ 
	NMRI5 vs NMRI7  	& 1895 \\
	Oocysts vs NMRI3  	& 3310 \\  
	Oocysts vs NMRI5	& 3605 \\ 
	Oocysts vs NMRI7	& 3085 \\ 
	Oocysts vs sporozoites & 3421 \\
	Sporozoites vs NMRI3 	& 1663 \\
	Sporozoites vs NMRI5 	& 1605 \\
	Sporozoites vs NMRI7 	& 2473 \\ 
	NMRI vs C57BL/6 	& 22 \\	\midrule	
\textit{Comparisons, no difference}  	\\ \midrule
	NMRI vs Rag1-/-     & 0 \\
	C57BL/6 vs Rag1-/-  & 0 \\
	NMRI3 1st vs NMRI3 2nd  	    & 0 \\
	NMRI5 1st vs NMRI5 2nd  	    & 0 \\
	NMRI7 1st vs NMRI7 2nd  	    & 0 \\
	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 \\
	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 \\ \bottomrule
\hline
\end{tabular}
\end{center}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	INPUT TO HEATMAP	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Differentially abundant mRNAs used for hierarchical clustering}
\setlength{\tabcolsep}{10pt}
\begin{table}[H]
\begin{adjustwidth}{-10}{}
\begin{center}
	\caption{A selection of differently abundant mRNAs are used for hierarchical clustering of \textit{E. falciformis}
	life cycle relevant genes. In each comparison (see Table 3), the 500 differentially abundant mRNAs with lowest FDR
	are selected. In the next step, the 500 mRNAs from each comparison (or less) are joined (4935 genes) and only 
	unique genes are selected (1618 genes). The 22 genes in the NMRI vs C57BL/6 
	comparison are not included in the \textit{E. falciformis} life cycle analysis.}
\begin{tabular}{*2l}    \toprule
	\textit{Data description} & Number of genes \\ \midrule
	Sum of 1st infection NMRI sample differences	& 4935  \\ 
	(including oocysts and sporozoites)	\\	
	Used in hierarchical clustering (heatmap)  	& 1618 \\ 	\bottomrule	
\hline
\end{tabular}
\end{center}
\end{adjustwidth}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	HEATMAPS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hierarchical clustering of genes different depending on time post infection}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{Ef_most_sig_lifecycle_heatmapi.pdf}  
	\caption{\textit{E. falciformis} with Euklidean distance .........as implemented in R package limma.}
\end{center}
\end{figure}

\end{document}

### 2016-06-22 1_ballgown_import.R

@@ -66,7 +66,7 @@ Ef.bg <- subset(Ef.bg,
                 pData(Ef.bg)$timepoint != 0 | is.na(pData(Ef.bg)$timepoint),
                 genomesubset=FALSE)
 
-## A function to et raw coungs for exons, transcripts and genes out of
+## A function to get raw coungs for exons, transcripts and genes out of
 ## the ballgown objects
 
 raw.counts.4.bg <- function(bg){

### 2016-06-22 2_edgeR_diff.R

@@ -2,17 +2,20 @@
 
 ## needed from data import script are two raw count objects 
 
+source("https://bioconductor.org/biocLite.R")
 if(!exists("Mm.RC")|!exists("Ef.RC")){
     source("1_ballgown_import.R")
 }
 
 # Check checkpoint package for package control.
-if(!require(edgeR)) biocLite("edgeR") # Imports package if user does not have it
+source("http://bioconductor.org/biocLite.R")
+if(!require(edgeR)) BiocInstaller::biocLite("edgeR") # Imports package if user does not have it
 if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
-if(!require(RUVSeq)) biocLite("RUVSeq") # Imports package if user does not have it
+if(!require(RUVSeq)) BiocInstaller::biocLite("RUVSeq") # Imports package if user does not have it
 if(!require(reshape)) biocLite("reshape") # Imports package if user does not have it
+if(!require(statmod)) biocLite("statmod") # Imports package if user does not have it
 
-
+library(statmod)
 library(edgeR)
 library(GGally)
 library(RUVSeq)

### 2016-06-22 4_compareArray.R

@@ -4,6 +4,7 @@
 ## of IFNg-regulated host pathways for the parasite development"
 ## DOI:10.1038/mi.2013.115
 
+source("https://bioconductor.org/biocLite.R")
 # Check checkpoint package for package control.
 if(!require(Biobase)) biocLite("Biobase") # Imports package if user does not have it
 if(!require(RSvgDevice)) biocLite("RSvgDevice") # Imports package if user does not have it

### 2016-06-22 A_data_curataion.R

@@ -38,6 +38,7 @@ if(!exists("All.RC")|!exists("Mm.RC")|!exists("Ef.RC")){
 
 library(ggplot2)
 library(reshape)
+library(gridExtra)
 
 ## Overview for differential representation of host/parasite
 RC.table <-  as.data.frame(cbind(sample=as.character(pData(All.bg)$samples),
@@ -46,8 +47,7 @@ RC.table <-  as.data.frame(cbind(sample=as.character(pData(All.bg)$samples),
 
 RC.table$c.total.reads <- colSums(All.RC[[3]])
 
-RC.table$c.Mm.reads <- colSums(All.RC[[3]]
-                                [grepl('^XLOC.*', rownames(All.RC[[3]])),])
+RC.table$c.Mm.reads <- colSums(All.RC[[3]][grepl('^ENSMUS.*', rownames(All.RC[[3]])),])
 
 RC.table$c.Ef.reads <- colSums(All.RC[[3]]
                                 [grepl('^EfaB.*', rownames(All.RC[[3]])),])
@@ -80,27 +80,51 @@ RC.table$c.Ef.genes <- colSums(All.RC[[3]]
 keep.val <- c(0, 100, 1000, 2000, 3000, 5000)
 names(keep.val) <- keep.val
 
+## Original function
+#r.c.s.l <-
+#    lapply(keep.val,
+#           function (x){
+#               kept.df <- Mm.RC[[3]][rowSums(Mm.RC[[3]])>x, ]
+#               kept.df <- melt(kept.df)
+#               return(kept.df)
+#           })
+
+## Totta's fix to have data plotted with all (not already removed) samples:
 r.c.s.l <-
     lapply(keep.val,
            function (x){
-               kept.df <- Mm.RC[[3]][rowSums(Mm.RC[[3]])>x, ]
-               kept.df <- melt(kept.df)
-               return(kept.df)
+		
+		Mm.RCtemp <- All.RC[[3]][grepl('^ENSMUS.*', rownames(All.RC[[3]])),]+ 0.1 # add tiny number to be able to plot transcripts with zero reads mapping
+        	kept.df <- Mm.RCtemp[rowSums(Mm.RCtemp)>x,]       
+		kept.df <- melt(kept.df)
+		kept.df <- kept.df[grep("^.*_(oocysts|sporozoites)_.*$", kept.df$X2, invert = T),] # remove ooc and sporo samples
+		return(kept.df)
            })
 
+#####################################################
 density.plots <-
[…]

### 2016-06-22 B_mds_sample_analysis.R

@@ -1,0 +1,165 @@
+## DIFFERENTIAL GENE EXPRESSION ANALYSIS ON EIMERIA SEQS #####
+
+#setwd("/home/totta/Ef_RNAseq")
+
+if(!exists("raw.counts.4.bg")){
+	source("2_edgeR_diff.R")
+        }
+
+library(limma)
+library(scales)
+library(stringr)
+
+#Ef.RC <- raw.counts.4.bg(Ef.bg)
+
+######################################
+
+######################################
+pdf("figures/rep_pairs_EfG.pdf", width = 70, height=70) # So far not working - because R interrupt?
+ggpairs(Ef.RC[[3]])
+dev.off()
+
+
+##################### For EIMERIA: #################################
+
+ggplot()+
+    geom_point(data = tableEf, aes(x = Samples,
+                   y = as.numeric(colSums(Ef.RC[[3]])))) +
+	scale_y_log10(labels = comma) +
+            ylab("Number of reads") +
+                theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+
+
+pdf("figures/rep_pairs_NORM_EfG.pdf", width = 70, height=70)
+ggpairs(cpm(GM.E)) #ggpairs makes matrix of plots
+dev.off()
+
+###################################################################
+## COLORS FOR pDATA() GROUPS 
+###################################################################
+
+### Show all the RColorBrewer colour schemes available
+display.brewer.all()
+####
+
+#day.colors = c("#1b9e77", "#1b9e77", "#1b9e77", 
+#               "#d95f02", "#d95f02", 
+#               "#1b9e77", "#1b9e77", 
+#               "#474747", "#474747",
+#               "#d95f02", "#d95f02", 
+#               "#1b9e77", "#1b9e77",
+#               "#474747", "#474747",
+#               "#1b9e77","#1b9e77", "#1b9e77")
+
+#strain.colors = c("#1b9e77", "#1b9e77", "#1b9e77",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#474747", "#474747",
+#                "#d95f02", "#d95f02", "#d95f02") 
[…]

### 2016-06-22 data_summaries.R

@@ -1,51 +1,59 @@
-controlsc## Summarize and create plots or tables for data
+#### WRONG
+## Something is wrong here: use "A_data_curation.R" instead!!!!!!! 
+
+#controlsc## Summarize and create plots or tables for data
 ## E.g. raw counts, percetage Eimeria reads... etc
 
 library(scales)
 library(xtable)
 library(stringr)
 
-<<<<<<< HEAD
+#<<<<<<< HEAD
 ####################################################################
 # TABLE WITH RAW DATA - OUTPUT IN LATEX
 ####################################################################
-=======
+#=======
 ## Summary table of transcript counts per sample
 ## MOUSE table
-tableM <- as.data.frame(colSums(mouse.RC[[3]]))
+tableM <- as.data.frame(colSums(Mm.RC[[3]]))
 #colnames(tableM)[,2] <- "Read.number"
 ## MOUSE w/o day 0 
 tableM$Samples <- rownames(tableM)
 
 tableMsmall <- tableM[!(str_detect(rownames(tableM),"^.*0dpi.*$")), ]
-row.names(tableMsmall) <- seq(1:18)
->>>>>>> Totta_working
+row.names(tableMsmall) <- seq(1:17)
+#>>>>>>> Totta_working
 
 ## table MOUSE and EIMERIA raw counts and percentage parasite reads 
 table.rc <- as.data.frame(colSums(All.RC[[3]][str_detect(rownames(All.RC[[3]]), "^EfaB_.*$"), ]))
 #tableEf <- as.data.frame(colSums(Ef.RC[[3]]))
 colnames(table.rc)[1] <- "EimeriaTranscripts"
-table.rc$MusTranscripts <- colSums(All.RC[[3]][str_detect(rownames(All.RC[[3]]), "^XLOC_.*$"), ])
+table.rc$MusTranscripts <- colSums(All.RC[[3]][str_detect(rownames(All.RC[[3]]), "^ENSMUS.*$"), ])
 table.rc$Samples <- rownames(table.rc)
-row.names(table.rc) <- seq(1:27)
+row.names(table.rc) <- seq(1:30)
 table.rc$PercentEimeria <- format((table.rc[,1]/colSums(All.RC[[3]]))*100, digits=3, scientific = F)
 
-<<<<<<< HEAD
+#<<<<<<< HEAD
 # sort and make rownames pretty for LaTeX use
[…]

### 2016-06-22 distribution_check.R

@@ -4,14 +4,14 @@
 library(reshape2)
 ############## MOUSE ##################
 ## rearrange GM.E for ggplot format
-long.Mcounts <- melt(GM$counts)
-long.counts$strain <- gsub("_.*", "", long.Mcounts$Var2)
-long.Mcounts$strain <- gsub("_.*", "", 
-		long.Mcounts$Var2)
+long.Mcounts <- melt(Mm.RC[[3]])
+#long.counts$strain <- gsub("_.*", "", long.Mcounts$Var2)
+#long.Mcounts$strain <- gsub("_.*", "", 
+#		long.Mcounts$Var2)
 
-long.normMcounts <- melt(cpm(GM))
-long.normMcounts$strain <- gsub("_.*", "", 
-		long.normMcounts$Var2)
+long.normMcounts <- melt(cpm(Mm.RC[[3]]))
+#long.normMcounts$strain <- gsub("_.*", "", 
+#		long.normMcounts$Var2)
 
 ## RAW MOUSE DISTRIBUTION
 pdf("figures/distributions_mouse5000.pdf", 
@@ -43,13 +43,13 @@ dev.off()
 
 ############## PARASITE ###################
 long.Ecounts <- melt(GM.E$counts)
-long.Ecounts$strain <- gsub("_.*", "", long.Ecounts$Var2)
-long.Ecounts$strain <- gsub("_.*", "", 
-		long.Ecounts$Var2)
+#long.Ecounts$strain <- gsub("_.*", "", long.Ecounts$Var2)
+#long.Ecounts$strain <- gsub("_.*", "", 
+#		long.Ecounts$Var2)
 
 long.normEcounts <- melt(cpm(GM.E))
-long.normEcounts$strain <- gsub("_.*", "", 
-		long.normEcounts$Var2)
+#long.normEcounts$strain <- gsub("_.*", "", 
+#		long.normEcounts$Var2)
 
 excludedE <- long.normEcounts[long.normEcounts$Var2%in%
 		c("NMRI_2ndInf_7dpi_rep2",

### 2016-06-22 functionsEfRNAseq.R

@@ -22,6 +22,7 @@ gene.table.topGO <- function(TOGO.list, pval=0.01){
 
 
 ## adjusted functions from ballgown
+if(!require(GenomicRanges)) biocLite("GenomicRanges")
 
 library(ballgown)
 library(GenomicRanges) # needed for elementMetadata()

### 2016-06-23 manuscript/tex_bmc_format/manuscript_rnaseq.tex
Commit before Serengeti trip. Lots of small fixes in scripts due to Mm10 annotations instead of Mm9, as well as addition of other heatmaps, improved MDS plots... etc. SI/initial data summary + legends (table_rawcounts.tex) also commited here.

@@
 Taken together, this indicates that the thrombospondin and sushi-domain genes 
-(EfaB_MINUS_4114.g412 and EfaB_PLUS_1425.g183) are involved in sporozoite invasion in 
+(EfaB\_MINUS\_4114.g412 and EfaB\_PLUS\_1425.g183) are involved in sporozoite invasion in 
 \textit{E. falciformis} and that the mRNAs are transcribed and available before excystation. 
@@
 We suggest (speculate...?) that the thrombospondin annotated ortholog and the LCCL domain-containing protein 
-(EfaB_MINUS_11233.g986) are involved in cell adhesion in \textit{E. falciformis}.
+(EfaB\_MINUS\_11233.g986) are involved in cell adhesion in \textit{E. falciformis}.
 %One examples
@@
 In our analysis we demonstrate which biological processes are dominant in different
-life cycle stages of E. falciformis in the mouse. The RNAseq transcriptome provided here 
+life cycle stages of \textit{E. falciformis} in the mouse. The RNAseq transcriptome provided here 
 allows for detailed analysis of genes involved in those processes, providing candidates for 
 life stage specific marker in \textit{Eimeria} spp. research.
+
+\ldots
+\ldots
+\ldots
 
@@
 \section*{Methods}
-\subsection*{Mice and infection procedure}
-Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
-Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
-Animal procedures were performed according to the German Animal Protection Laws as 
-directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
-(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
-tapwater was used instead of PBS for administration of oocysts. Briefly, the indicated number of
-sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
-administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate
-was used for all infections and parasite samples. Oocysts were originally purchased from Bayer 
-() and the strain is maintained through passage in mice  in our facilities as described 
-elsewhere (Schmid12).
-\ldots
-\subsection*{RNA extraction}
- \ldots
-\subsection*{Sequence quality assessment and alignment}
-Fastq\_quality\_filter was applied to Illumina Hiseq 2000 sequenced samples. 
-Since this is not easily applicable to pair-end sequencing data, a low threshold was 
-used on the hiseq data. A phred score of 10 was used, i.e., the probability of false base 
-calling is one in ten. We further set q = 60, i.e., nine out of ten bases
-or more is required to be correct in at least 
-60\% of the bases in each read for the read sequence to be kept for further analysis.
-This resulted in.......
-\subsubsection*{Alignment and reference genomes}
-We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
-Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
-\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
-alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
-files including mRNA sequences from both species were aligned against the dual reference genome
-using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
-pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
-for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
-files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
-(version 2.1.1, Trapnell10).
-
 \ldots

### 2016-06-23 1_ballgown_import.R

@@ -1,9 +1,5 @@
 ### ballgown for data import from the cufflinks/tablemaker pipeline
 
-# Check package for package control.
-source("https://bioconductor.org/biocLite.R")
-if(!require(ballgown)) biocLite("ballgown") # Imports package if user does not have it
-
 library(ballgown)
 setwd("~/Ef_RNAseq")
 

### 2016-06-23 2_edgeR_diff.R

@@ -2,18 +2,11 @@
 
 ## needed from data import script are two raw count objects 
 
-source("https://bioconductor.org/biocLite.R")
+#source("https://bioconductor.org/biocLite.R")
 if(!exists("Mm.RC")|!exists("Ef.RC")){
     source("1_ballgown_import.R")
 }
 
-# Check checkpoint package for package control.
-source("http://bioconductor.org/biocLite.R")
-if(!require(edgeR)) BiocInstaller::biocLite("edgeR") # Imports package if user does not have it
-if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
-if(!require(RUVSeq)) BiocInstaller::biocLite("RUVSeq") # Imports package if user does not have it
-if(!require(reshape)) biocLite("reshape") # Imports package if user does not have it
-if(!require(statmod)) biocLite("statmod") # Imports package if user does not have it
 
 library(statmod)
 library(edgeR)
@@ -48,7 +41,7 @@ colnames(Ef.design)  <- gsub("_", ".",
 
 Ef.contrasts <- makeContrasts(
     ## just different stages in NMRI
-    N3vsN5 = NMRI.1stInf.3dpi-NMRI.1stInf.5dpi,
+    N3vsN5 = NMRI.1stInf.3dpi-NMRI.1stInf.5dpi, #1
     N3vsN7 = NMRI.1stInf.3dpi-NMRI.1stInf.7dpi,
     N3vsSpo = NMRI.1stInf.3dpi-NMRI.sporozoites,
     N3vsOoc = NMRI.1stInf.3dpi-NMRI.oocysts,
@@ -57,24 +50,24 @@ Ef.contrasts <- makeContrasts(
     N5vsOo = NMRI.1stInf.5dpi-NMRI.oocysts,
     N7vsSp = NMRI.1stInf.7dpi-NMRI.sporozoites,
     N7vsOo = NMRI.1stInf.7dpi-NMRI.oocysts,
-    SpvsOo=  NMRI.sporozoites-NMRI.oocysts,
[…]

### 2016-06-23 3_annotations.R

@@ -6,9 +6,6 @@ if(!exists("Ef.1st.pass.model")){ #Totta changes Mm to Ef here
     source("2_edgeR_diff.R")
 }
 
-# Check checkpoint package for package control.
-if(!require(Mus.musculus)) biocLite("Mus.musculus") # Imports package if user does not have it
-
 ## The Bioconductor mouse annotation libary
 library(Mus.musculus)
 

### 2016-06-23 4_compareArray.R

@@ -4,14 +4,14 @@
 ## of IFNg-regulated host pathways for the parasite development"
 ## DOI:10.1038/mi.2013.115
 
-source("https://bioconductor.org/biocLite.R")
+#source("https://bioconductor.org/biocLite.R")
 # Check checkpoint package for package control.
-if(!require(Biobase)) biocLite("Biobase") # Imports package if user does not have it
-if(!require(RSvgDevice)) biocLite("RSvgDevice") # Imports package if user does not have it
-if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
-if(!require(ggplot2)) biocLite("ggplot2") # Imports package if user does not have it
+#if(!require(Biobase)) biocLite("Biobase") # Imports package if user does not have it
+#if(!require(RSvgDevice)) biocLite("RSvgDevice") # Imports package if user does not have it
+#if(!require(GGally)) biocLite("GGally") # Imports package if user does not have it
+#if(!require(ggplot2)) biocLite("ggplot2") # Imports package if user does not have it
 ## For below package does not always work: then manual import is necessary
-if(!require(mgug4122a.db)) biocLite("mgug4122a.db") # Imports package if user does not have it
+#if(!require(mgug4122a.db)) biocLite("mgug4122a.db") # Imports package if user does not have it
 
 library(mgug4122a.db)
 library(Agi4x44PreProcess) # !!!! NOT available for R 3.3.0 - NOT on Bioconductor anymore
@@ -28,9 +28,10 @@ if(!exists("gene2GO")){
     source("3_annotations.R")
 }
 
-targets <- readTargets("/data/Mouse_arrays/2012_arrays/targets.txt", sep=";")
+#targets <- readTargets("/data/Mouse_arrays/2012_arrays/targets.txt", sep=";")
+targets <- readTargets("/SAN/Mouse_arrays/2012_arrays/targets.txt", sep=";")
 
-RG <- read.maimages(targets, path= "/data/Mouse_arrays/2012_arrays/", source = "agilent", 
+RG <- read.maimages(targets, path= "/SAN/Mouse_arrays/2012_arrays/", source = "agilent", 
                     other.columns = list(IsFound = "gIsFound",
                       IsWellAboveBG = "gIsWellAboveBG", 
                       IsSaturated = "gIsSaturated",
@@ -138,7 +139,7 @@ dev.off()
[…]

### 2016-06-23 5_clustering.R

@@ -1,10 +1,5 @@
 ## Making heatmaps and other clusters of gene expression data
 
-# Check checkpoint package for package control.
-if(!require(pheatmap)) biocLite("pheatmap") # Imports package if user does not have it
-if(!require(made4)) biocLite("made4") # Imports package if user does not have it
-if(!require(RColorBrewer)) biocLite("RColorBrewer") # Imports package if user does not have it
-
 library(pheatmap)
 library(made4)
 library(RColorBrewer)
@@ -28,13 +23,13 @@ union.of.Ef.1st2nd.top <-
     unique(unlist(lapply(Ef.1st.pass.model[[3]][11:15], head, n=500)))
 
 Ef.1st2nd.top.data <-
-    cpm(Ef.1st.pass.model[[4]])[union.of.1st2nd.top,]
+    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.1st2nd.top,]
 
 ######## EIMERIA DEVELOPMENT IN DIFFERENT MOUSE STRAINS (day 5 only) ##################
 union.of.Ef.strain.depend.top <-
     unique(unlist(lapply(Ef.1st.pass.model[[3]][16:18], head, n=500)))
 
-Ef.strain.depend.top.data <-
+Ef.strain.top.data <-
     cpm(Ef.1st.pass.model[[4]])[union.of.Ef.strain.depend.top,]
 
 get.scaled.and.clustered <- function(data){
@@ -51,46 +46,102 @@ Ef.hclustered.strain <- get.scaled.and.clustered(Ef.strain.depend.top.data )
 
 
 hcluster <- list()
-hcluster[["Ef"]] <- cutree(Ef.hclustered, k= 7) ## TK changed k from 5 ## h=7.2)
+hcluster[["Ef.cyc"]] <- cutree(Ef.hclustered, k = 7) ## Life cycle
+hcluster[["Ef.1st2nd"]] <- cutree(Ef.hclustered, k = 7) ## 1st versus 2nd 
+hcluster[["Ef.strain"]] <- cutree(Ef.hclustered, k = 7) ## Strain comparison
 
-Ef.hclustered.df <- as.data.frame(hcluster[["Ef"]])
-names(Ef.hclustered.df) <- "Cluster"
-Ef.hclustered.df$Cluster <- as.factor(Ef.hclustered.df$Cluster)
+Ef.cyc.hclustered.df <- as.data.frame(hcluster[["Ef.cyc"]])
+names(Ef.cyc.hclustered.df) <- "Cluster"
[…]

### 2016-06-23 B_mds_sample_analysis.R

@@ -129,8 +129,9 @@ par(mfrow = c(2,2), mai = c(0.6, 0.5, 0.4, 0.4))
 plotMDS(Ef.RC[[3]], labels = pData(Ef.bg)$batch,
 	      col.axis = "#474747", col.lab = "#474747", col.main = "#474747", col.sub = "#474747",
               col = batch.colors.ef,
-              xlab = "Fold change, dimension 1", ylab = "Fold change, dimension 2",
-              mgp = c(2,1,0))
+              xlab = "Fold change, dimension 1", 
+	      ylab = "Fold change, dimension 2",
+	      mgp = c(2,1,0))
 #mai = c(1, 0.1, 0.1, 0.1)
 title("Batch, Ef", line = 0.7)
 
@@ -140,7 +141,7 @@ plotMDS(Ef.RC[[3]], labels = pData(Ef.bg)$seq.method,
               xlab = "Fold change, dimension 1", ylab = "Fold change, dimension 2",
               mgp = c(2,1,0))
 #mai = c(1, 0.1, 0.1, 0.1))
-title("Sequnecing method, Ef", line = 0.7)
+title("Sequencing method, Ef", line = 0.7)
 
 
 plotMDS(Mm.RC[[3]], labels = pData(Mm.bg)$batch,
@@ -157,7 +158,7 @@ plotMDS(Mm.RC[[3]], labels = pData(Mm.bg)$seq.method,
               xlab = "Fold change, dimension 1", ylab = "Fold change, dimension 2",
               mgp = c(2,1,0))
 #mai = c(1, 0.1, 0.1, 0.1))
-title("Sequnecing method, mouse", line = 0.7)
+title("Sequencing method, mouse", line = 0.7)
 
 dev.off()
 

### 2016-06-23 functionsEfRNAseq.R

@@ -22,7 +22,7 @@ gene.table.topGO <- function(TOGO.list, pval=0.01){
 
 
 ## adjusted functions from ballgown
-if(!require(GenomicRanges)) biocLite("GenomicRanges")
+#if(!require(GenomicRanges)) biocLite("GenomicRanges")
 
 library(ballgown)
 library(GenomicRanges) # needed for elementMetadata()

### 2016-07-13 5_clustering.R
Changes from 'before Serengeti trip'

@@ -71,7 +71,7 @@ pheatmap(Ef.strain.top.data,
          cluster_cols = T,
          annotation_row = Ef.strain.hclustered.df,
          ## annotation_names_row = F,
-         cutree_rows = 7, # 5 in EH script 
+         cutree_rows = 3, # 5 in EH script 
          show_rownames = F,
          main = expression(paste(italic("E. falciformis"),
              " mRNAs differently abundant between mouse strains")))
@@ -86,7 +86,7 @@ pheatmap(Ef.1st2nd.top.data,
          cluster_cols = T,
          annotation_row = Ef.1st2nd.hclustered.df,
          ## annotation_names_row = F,
-         cutree_rows = 7, # 5 in EH script 
+         cutree_rows = 2, # 5 in EH script 
          show_rownames = F,
          main = expression(paste(italic("E. falciformis"),
              " mRNAs differently abundant between 1st and 2nd infection")))
@@ -106,7 +106,11 @@ pheatmap(Ef.cycle.diff.top.100.data,
          main = expression(paste(italic("E. falciformis"),
              " mRNAs differently abundant between lifecycle stages")))
 dev.off()
-## mouse infection progression 
+
+###################################################
+## 	MOUSE
+###################################################
+##mouse infection progression 
 union.of.Mm.cycle.diff.top.100 <-
     unique(unlist(lapply(Mm.1st.pass.model[[3]][c(1:3, 12:14)], head, n=500)))
 
@@ -121,24 +125,65 @@ Mm.1st2nd.top.data <-
     cpm(Mm.1st.pass.model[[4]])[union.of.Mm.1st2nd.top,]
 
 ######## EIMERIA DEVELOPMENT IN DIFFERENT MOUSE STRAINS (day 5 only) ##################
-union.of.Mm.strain.depend.top <-
+union.of.Mm.strain.top <-
     unique(unlist(lapply(Mm.1st.pass.model[[3]][11], head, n=500)))
 
-Mm.strain.depend.top.data <-
-    cpm(Mm.1st.pass.model[[4]])[union.of.Mm.strain.depend.top,]
+Mm.strain.top.data <-
+    cpm(Mm.1st.pass.model[[4]])[union.of.Mm.strain.top,]
 
 ## clustering
 
[…]

### 2016-07-13 figures_and_manuscript/manuscript_bmc_format/manuscript_rnaseq.tex
try push 2

--- 2016-06-23 manuscript/tex_bmc_format/manuscript_rnaseq.tex
@@
 \usepackage{graphicx}
-\graphicspath{ {../figures/} }
+\graphicspath{ {../} }
 %\def\includegraphic{}
@@
 
-\title{A sample article title}
+\title{Dual host-parasite transcriptomes of apicomplexan \textit{Eimeria falciformis} and its natural mouse host}
 
@@
    noteref={n1},                        % id's of article notes, if any
-   email={jane.e.doe@cambridge.co.uk}   % email address
-]{\inits{JE}\fnm{Jane E} \snm{Doe}}
+   email={emanuel.heitlinger@hu-berlin.de}   % email address
+]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
 \author[
-   addressref={aff1,aff2},
-   email={john.RS.Smith@cambridge.co.uk}
-]{\inits{JRS}\fnm{John RS} \snm{Smith}}
+   addressref={aff1},
+   %email={john.RS.Smith@cambridge.co.uk}
+]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
 
@@
 \address[id=aff1]{%                           % unique id
-  \orgname{Department of Zoology, Cambridge}, % university, etc
-  \street{Waterloo Road},                     %
+  \orgname{Department of Biology, Humboldt-Universitat zu Berlin}, % university, etc
+  \street{Philippstr. 13, Haus 14},                     %
   %\postcode{}                                % post or zip code
-  \city{London},                              % city
-  \cny{UK}                                    % country
+  \city{Berlin},                              % city
+  \cny{Germany}                                    % country
 }
@@
 
+\subsubsection*{Data overview}
+In total, xx reads were sequenced. For each sample, mouse read numbers were 10⁷ - 10⁸
+Quality filters applied resulted in xx reads being removed from further analysis
+
 \subsubsection*{Preparation for invasion in oocysts}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Methods}
-\ldots
+\section{Methods}
+\subsection{Mice and infection procedure}
+Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
+Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
+Animal procedures were performed according to the German Animal Protection Laws as 
+directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
+(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
+tapwater was used instead of PBS for administration of oocysts. Briefly, NMRI mice were infected 
+two times, which will be referred to as first and second infection. For the first infection, 
+150 sporulated oocysts were administered in 100 $μ$L by oral gavage. During the first infection
+of 60 mice, all animals were weighed every day. On day zero, before infection, as well as on day three,
+five and seven days post infection, dpi, caeca from 3-4 sacrificed mice per time point were 
+collected. Epithelial cells were isolated as described in Schmid et al.(schmid12). For challenge 
+infection, mice recovered for four weeks before second infection. 
+Recovery was monitored by weighing and visual inspection of fur. For the second infection, 
+1500 sporulated oocysts were applied by oral gavage. Three mice were used as non-second infection control, 
+referred to as day 0, second infection.
 
-\subsection*{Sub-heading for section}
-Text for this sub-heading \ldots
-\subsubsection*{Sub-sub heading for section}
-Text for this sub-sub-heading \ldots
-\paragraph*{Sub-sub-sub heading for section}
-Text for this sub-sub-sub-heading \ldots
-text text text
-text text text
-text text text
-text text text
-text text text
-text text text
+\subsection{Oocyst purification for infection and sequencing}
+Sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
+administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate, 
+\textit{E. falciformis} Bayer Haberkorn 1970, was used for all infections and parasite 
+samples. The strain is maintained through passage in NMRI mice in our facilities as described 
+elsewhere (schmid12).
+
+\subsection{Sporozoite isolation}
+Sporozoites were isolated from sporocysts by excystment. For this, sporocysts were incubated at 
+37$°$C in DMEM containing 0.04\% tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) 
+for 30 min. Sporozoites were purified by the method of Schmatz et al (schmatz--).
+
+\subsection{RNA extraction}
+Total RNA was isolated from infected epithelial cells, sporozoites and sporulated oocysts 
+using Trizol according to the manufacturer’s protocol (Invitrogen). High quality  \emph{what is the meaning of 'high quality' here?}
+RNA was used to produce an mRNA library using the Illumina’s TruSeq RNA Sample Preparation guide.
+\emph{stolen from genome paper}
+Sporozoites were stored in 1 mL Trizol until RNA-isolation. 
+Total RNA was isolated using the PureLink RNA Mini Kit (Invitrogen).
+
+\subsection{Sequencing, quality assessment and alignment}
+cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as specified in SI xx (both unstranded). Fastq\_quality\_filter was applied to Illumina Hiseq 2000 samples. 
+A phred score of 10 was applied. We further set q = 60. These settings require that nine out of ten bases
+or more are correct in at least 60\% of the bases for each read.
+
+\subsubsection{Alignment and reference genomes}
+We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
+Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
+\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
+alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
+files including mRNA sequences from both species were aligned against the dual reference genome
+using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
+pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
+for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
+files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
+(version 2.1.1, Trapnell10).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-07-14 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
MS reshuffled and folder structure changed: MS nw i figures folder

--- 2016-07-13 figures_and_manuscript/manuscript_bmc_format/manuscript_rnaseq.tex
@@
 
+\usepackage[colorlinks]{hyperref} % To color blocks of text 
+%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
+\usepackage{changepage}
+\usepackage{booktabs} % For prettier tables
+\usepackage{float} % To control positions of e.g. tables or figures
+\usepackage[section]{placeins}
 \usepackage{graphicx}
@@
 \startlocaldefs
-%\affi1=Department of Molecular Parasitology, Humboldt University, Berlin, Germany
-%\affi2=Max Planck Institute for Biology of Ageing, Joseph-Stelzmann-Strasse 9B, 50931 Cologne, Germany
-%\affi3=Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research, Forschungsverbund Berlin e.V., Berlin, Germany
 \endlocaldefs
@@
 
-\title{Dual host-parasite transcriptomes of apicomplexan \textit{Eimeria falciformis} and its natural mouse host}
+\title{Dual host-parasite transcriptomes of apicomplexan Eimeria falciformis and its natural mouse host}
 
@@
 \author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   addressref={aff1},
+]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
+\author[
+   addressref={aff1},
+]{\inits{SS}\fnm{Simone} \snm{Spork}}
+\author[
+   addressref={aff2},
+]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
+\author[
+   addressref={aff1},
+]{\inits{RL}\fnm{Richard} \snm{Lucius}}
+\author[
+   addressref={aff1, aff3},                   % id's of addresses, e.g. {aff1,aff2}
    corref={aff1},                       % id of corresponding address, if any
@@
 ]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
-\author[
-   addressref={aff1},
-   %email={john.RS.Smith@cambridge.co.uk}
-]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
 
@@
 \address[id=aff2]{%  
-  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
-  \street{D\"{u}sternbrooker Weg 20},
-  \postcode{24105}
-  \city{Kiel},
+  \orgname{Max Planck Institute for Biology of Ageing},
+  \street{Joseph-Stelzmann-Strasse 9B},
+  \postcode{50931}
+  \city{Cologne},
+  \cny{Germany}
+}
+\address[id=aff3]{%  
+  \orgname{Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research},
+  \street{Alfred-Kowalke-Str. 17},
+  \postcode{10315}
+  \city{Berlin},
   \cny{Germany}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\author[
-%   addressref={affi1},                   % id's of addresses, e.g. {aff1,aff2}
-%   corref={aff1},                       % id of corresponding address, if any
-%  % noteref={n1},                        % id's of article notes, if any
-%   email={kasemotq@hu-berlin.de}   % email address
-%]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
-%\author[
-%   addressref={affi1}                   % id's of addresses, e.g. {aff1,aff2}
-%  % corref={aff1},                       % id of corresponding address, if any
-%  % noteref={n1},                        % id's of article notes, if any
-%  % email={}   % email address
-%]{\inits{SS}\fnm{Simone} \snm{Spork}}
-%%\author[
-%   addressref={affi2},                   % id's of addresses, e.g. {aff1,aff2}
-%   %corref={aff3},                       % id of corresponding address, if any
-%   noteref={n2},                        % id's of article notes, if any
-%   %email={emanuel.heitlinger@hu-berlin.de}   % email address
-%]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
-%\author[
-%   addressref={affi1, affi3},                   % id's of addresses, e.g. {aff1,aff2}
-%   %corref={aff3},                       % id of corresponding address, if any
-%   noteref={n1},                        % id's of article notes, if any
-%   email={emanuel.heitlinger@hu-berlin.de}   % email address
-%]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\address[id=aff1]{%                           % unique id
-%  \orgname{Department of Zoology, Cambridge}, % university, etc
-%  \street{Waterloo Road},                     %
-  %\postcode{}                                % post or zip code
-%  \city{London},                              % city
-%  \cny{UK}                                    % country
-%}
-%\address[id=affi1]{%
-%  \orgname{Humboldt-Universitat zu Berlin, Department of Molecular Parasitology},
-%  \street{Philipstr. 13, House 14},
-%  \postcode{10115}
-%  \city{Berlin},
-%  \cny{Germany}
-%}
-%\address[id=affi2]{%  
-%  \orgname{Max Planck Institute for Biology of Ageing},
-%  \street{Joseph-Stelzmann-Strasse 9B},
-%  \postcode{50931}
-%  \city{Cologne},
-%  \cny{Germany}
-%}
-%\address[id=affi3]{%  
-%  \orgname{Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research},
-%  \street{Alfred-Kowalke-Str. 17},
-%  \postcode{10315}
-%  \city{Berlin},
-%  \cny{Germany}
-%}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{artnotes}
+%\begin{artnotes}
 %\note{Sample of title note}     % note to the article
 %\note[id=n1]{Equal contributor} % note, connected to author
-\end{artnotes}
+%\end{artnotes}
 
-\end{fmbox}% comment this for two column layout
+%\end{fmbox}% comment this for two column layout
 
@@
 %
-%\end{fmbox}% uncomment this for twcolumn layout
+\end{fmbox}% uncomment this for twcolumn layout
 
@@
 Text for this section \ldots
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experimental overview}
+Replicate average of transcripts as order of magnitude for \textit{E.
+falciformis}/mouse are indicated in Table....
+Columns in table represent mouse strains used in infection experiments.
+Rows represent timepoints post infection plus oocyst and sporozoite samples.
+The upper part of the table shows data for first infection, and oocyst and sporozoite data.
+The lower part shows data for challenge infection.
+Averages were calculated after sample exclusions (see Methods).
+For exact values, see Table 1.
+
+	\setlength{\tabcolsep}{14pt}
+	\begin{table}[H]
+	\small
+	\begin{center}
+	\caption{Experimental overview with number of read for parasite/mouse indicated.}
+\begin{tabular}{*4l}    \toprule
+\textit{Day, 1st infection}  	& NMRI  & C57BL/6  & Rag1-/- \\ \midrule
+	0 (control)    & $10^4$ / $10^8$  & $10^2$ / $10^8$  & $10^2$ / $10^7$  \\ %rep1/rep2, NA=no replicate
+3  		& $10^5$ / $10^8$ & NA  & NA \\ 
+5  		& $10^7$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
+7  		& $10^8$ / $10^7$ & NA  & NA \\ 
+Oocysts 	& $10^8$ / NA & NA  & NA \\ 
+Sporozoites 	& $10^7$ / NA & NA  & NA \\ \midrule
+
+\textit{Day, 2nd infection}  	\\ \midrule
+0 (control)     & $10^3$ / $10^8$  &  NA  & NA  \\ %rep1/rep2, NA=no replicate
+3  		& $10^4$ / $10^8$ & NA  & NA \\ 
+5  		& $10^4$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
+7  		& $10^6$ / $10^8$ & NA  & NA \\
+
+	\bottomrule
+ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\section{mRNA abundance differences between different experimental groups}
+\setlength{\tabcolsep}{8pt}
+\begin{table}[H]
+\small
+\begin{center}
+\caption{mRNA abundance differences between different experimental groups.}
+%\label{tab:table3}
+\begin{tabular}{*3l}    \toprule
+\textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
+	\textit{comparisons} 	    & (FDR$\leq$0.01) &  (FDR$\leq$0.01/0.05) \\ \midrule
+	NMRI 0 vs NMRI 3		& NA   & 274 \\
+	NMRI 0 vs NMRI 5		& NA   & 1736 \\
+	NMRI 0 vs NMRI 7		& NA   & 2802 \\
+	NMRI 3 vs NMRI 5     		& 111  & 1 \\
+	NMRI 3 vs NMRI 7  		& 1385 & 1407 \\ 
+	NMRI 5 vs NMRI 7  		& 1895 & 873 \\ 
+	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
+	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
+\textit{First and second infection} & 		 & 	 \\ 
+\textit{comparisons} 	    & 		& 	\\ \midrule
+	NMRI 3 1st vs NMRI 3 2nd  	& 0  & 5 \\
+	NMRI 5 1st vs NMRI 5 2nd  	& 0  & 1 \\
+	NMRI 7 1st vs NMRI 7 2nd  	& 0  & 902 \\
+	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 &  mouse \\
+	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
+\textit{Mouse strain comparisons} & 		 & 	 \\ \midrule
+	NMRI vs C57BL/6 		& 22 	& NA* \\		
+	NMRI vs Rag1-/-     		& 0 & NA* 	\\
+	C57BL/6 vs Rag1-/-  		& 0   & 356 	\\ \midrule
+\textit{Day post infection,} & 		 & 		 \\ 
+\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
+	Oocysts vs NMRI 3  	& 3310 & NA \\  
+	Oocysts vs NMRI 5	& 3605 & NA \\ 
+	Oocysts vs NMRI 7	& 3085 & NA \\ 
+	Oocysts vs sporozoites  & 3421 & NA \\
+	Sporozoites vs NMRI 3 	& 1663 & NA \\
+	Sporozoites vs NMRI 5 	& 1605 & NA \\
+	Sporozoites vs NMRI 7 	& 2473 & NA \\ \bottomrule
+	\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\textit{E.
+falciformis} data overview of pairwise comparisons and number of genes with differently abundant mRNAs per comparison.
+NMRI followed by number indicates day post infection (e.g.
+NMRI3 = \textit{E.
+falciformis} genes from NMRI mouse day 3 post infection).
+Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are indluced.
+NAs are missing samples or not applicable for the species.
+NA* is due to missing NMRI day 0 sample from first infection.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\clearpage
+\subsection{Life cycle analysis... \textit{E. falciformis}}
+\begin{figure}[h!]
+%\begin{center}
+%\centering
+	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
+	\caption{\textit{E.
+falciformis} mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table ...).
+\textit{E.
+falciformis} samples from seven days p.i.
+cluster together (NMRI mice only).
+For \textit{E.
+falciformis} day 7, NMRI, the pattern is most pronounced in cluster 1 (up).
+Within this cluster, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 6).
+Distinct groups of genes also define sporozoites (cluster 6: up) and oocysts (cluster 3: up, cluster 7: down).
+mRNA profiles on days three and five p.i.
+from all three mouse strains cluster together.
+These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
+The latter cluster closer with days 3 and 5 than with oocysts or day 7.}
+%\end{center}
+\end{figure}
+
+\clearpage
+\subsection{Parasite development in mouse data}
+\begin{figure}[h!]
+%\begin{center}
+%\centering
+\includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
+\caption{Mouse mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table 3).
+Three out of four samples from day 7 p.i.
+(NMRI only) cluster together.
+These samples are characterized by genes in cluster 2 (up) as well as 3 and 4 (down).
+The fourth day 7 sample (challenge infection, second replicate) clusters with day 3 and 5 samples but upon visual inspection of cluster 4 diplays a profile similar to non-infected mice.
+The same sample is abnormal in the parasite profile (see Figure 5).
+NMRI and C57BL/6 uninfected samples cluster together, defined by clusters 3, 4 (up), and 1 (down).
+Day 3 and 5 samples cluster together, with Rag1-/- forming a separate group and Rag1-/- non-infected most distant in this sample cluster.
+All non-infected samples share a high mRNA abundance in cluster 3.}
+%\end{center}
+\end{figure}
+
+\clearpage
+\subsection{First versus second infection, mouse}
+\begin{figure}[h!]
+%\begin{center}
+\includegraphics[width=\linewidth]{Mm1st2ndHeatmap.pdf}  
+	\caption{Mouse mRNAs with significantly different abundance between first infection and challenge infection (see Table 3).
+Samples do not cluster into first and second infection clusters.
+The two major sample cluster each contain four challenge infection samples.
+}
+%\end{center}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection*{GO term enrichments in heatmap clusters}
@@
 \subsubsection*{Data overview}
-In total, xx reads were sequenced. For each sample, mouse read numbers were 10⁷ - 10⁸
+In total, xx reads were sequenced. For each sample, mouse read numbers were $10^7$ - $10^8$
 Quality filters applied resulted in xx reads being removed from further analysis
@@
 Sporozoites were stored in 1 mL Trizol until RNA-isolation. 
-Total RNA was isolated using the PureLink RNA Mini Kit (Invitrogen).
+Total RNA was isolated using the PureLink RNA Mini Kit (Invitrogen) and reverse transcribed into cDNA.
 
-\subsection{Sequencing, quality assessment and alignment}
-cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as specified in SI xx (both unstranded). Fastq\_quality\_filter was applied to Illumina Hiseq 2000 samples. 
+\subsection{Sequencing, sequence quality assessment and alignment}
+cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as specified in SI xx (both unstranded).
+A fastq\_quality\_filter (FASTQ-toolkit, version 0.0.14, available at https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina Hiseq 2000 samples after replacing "N" bases by "." annotation.
 A phred score of 10 was applied. We further set q = 60. These settings require that nine out of ten bases
@@
 files including mRNA sequences from both species were aligned against the dual reference genome
-using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
-pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
-for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
-files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
-(version 2.1.1, Trapnell10).
+using TopHat2 (version 2.0.14, Trapnell09) with -G specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual genome.
+Single-end and pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified for pair-end samples.
+Import into R was enabled by the R package Ballgown, which requires bam files to be processed by Tablemaker (Frazee15), in our case used with -qW -G specified.
+Tablemaker in turn makes use of Cufflinks (version 2.1.1, Trapnell10).
+
+\subsubsection{Data normalisation and sample exclusions}
+Count data was normalised in R, using the upperquartile method as implemented in edgeR (version 3.14.0).
+Briefly, normalisation assumes a negative binomial distribution of data and dispersed variance for genes with higher mean abundance.
+The normalisation shrinks the dispersion of variance towards a Poisson (linear) relationship between mean and variance.
+We evaluated the density distribution of our data and two samples contradicted the assumed negative binomial distribution and were excluded from further analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...).
+
+\subsection{Differentially abundant mRNAs used for hierarchical clustering}
+A selection of differently abundant mRNAs are used for hierarchical clustering of \textit{E.
+falciformis} life cycle relevant genes.
+In each comparison (see Table 3), the 500 differentially abundant mRNAs with lowest FDR are selected.
+In the next step, the 500 mRNAs from each comparison (or less) are joined (4935 genes) and only unique genes are selected (1618 genes).
+The 22 genes in the NMRI vs C57BL/6 comparison are not included in the \textit{E.
+falciformis} life cycle analysis.
+
+All parasite samples were included in heatmap of genes detected as differentially abundant in timepoint comparisons.
+
+On scale bar in heatmaps, 0 is mean mRNA abundance for each gene (row).
+Up (green) and down-regulation (brown) numbers denote number of standard deviations from row mean.
+Hierarchical clustering was performed using with Euclidean distances, method 'complete' (R package base).
+
+\setlength{\tabcolsep}{10pt}
+\begin{table}[H]
+\small
+\begin{center}
+\caption{Differentially abundant \textit{E. falciformis} mRNAs used for hierarchical clustering.}
+\begin{tabular}{*2l}    \toprule
+	\textit{Data description} & Number of genes \\ \midrule
+	Sum of 1st infection NMRI sample differences	& 4935  \\ 
+	(including oocysts and sporozoites)	\\	
+	Used in hierarchical clustering (heatmap)  	& 1618 \\ 	\bottomrule	
+\hline
+\end{tabular}
+\end{center}
+%\end{adjustwidth}
+\end{table}
 
@@
 \begin{figure}[h!]
-  \includegraphics[width=0.5\linewidth]{efmm-ratio}
+%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
   \caption{\csentence{Parasite read sequences per mouse read sequences for each experimental condition.}
@@
 
-  \begin{figure}[h!]
-  \includegraphics[width=0.7\linewidth]{ef-heatmap}
-  \caption{\csentence{Parasite genes with different mRNA abundance between samples, sorted into five clusters by hierarchical clustering (method = complete, distance = euclidean).}
-      \textit{E.
+ % \begin{figure}[h!]
+ % \includegraphics[width=0.7\linewidth]{ef-heatmap}
+ % \caption{\csentence{Parasite genes with different mRNA abundance between samples, sorted into five clusters by hierarchical clustering (method = complete, distance = euclidean).}
+ %     \textit{E.
 falciformis} samples from seven days p.i.
@@
 Up and downregulation is standard deviations from mean.}
-      \end{figure}
+ %     \end{figure}
 
-\begin{figure}[h!]
-  \includegraphics[width=0.7\linewidth]{mm-heatmap}
-  \caption{\csentence{Mouse genes with different mRNA abundance between samples, sorted into four clusters by hierarchical clustering (method = complete, distance = euclidean).}
-Three out of four samples from day 7 p.i.
+%\begin{figure}[h!]
+%  \includegraphics[width=0.7\linewidth]{mm-heatmap}
+%  \caption{\csentence{Mouse genes with different mRNA abundance between samples, sorted into four clusters by hierarchical clustering (method = complete, distance = euclidean).}
+%Three out of four samples from day 7 p.i.
 (NMRI only) cluster together.
@@
 Up and downregulation is standard deviations from mean.}
-      \end{figure}
+%      \end{figure}
 

### 2016-07-14 figuresANDmanuscript/table_rawcounts.tex
MS reshuffled and folder structure changed: MS nw i figures folder

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University/School Laboratory Report
% LaTeX Template
% Version 3.1 (25/3/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Linux and Unix Users Group at Virginia Tech Wiki 
% (https://vtluug.org/wiki/Example_LaTeX_chem_lab_report)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{article}

%\usepackage[version=3]{mhchem} % Package for chemical equation typesetting
%\usepackage{siunitx} % Provides the \SI{}{} and \si{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage[section]{placeins}
\usepackage{booktabs} % For prettier tables
\usepackage{float} % To control positions of e.g. tables or figures
\usepackage[colorlinks]{hyperref} % To color blocks of text 
\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
\usepackage{changepage}
%\usepackage[dvipdfmx]{graphicx}
%\usepackage{bmpsize} % Added at some point but now (June-16) it seems to do nothing
%\usepackage{natbib} % Required to change bibliography style to APA
%\usepackage{amsmath} % Required for some math elements

\graphicspath{{/home/totta/Ef\_RNAseq/figures\_and\_manuscript/}}
\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

%\usepackage{times} % Uncomment to use the Times New Roman font
%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\title{Data summary and analysis procedure: dual RNA-seq of \textit{Eimeria falciformis} infected \textit{Mus musculus}} % Title

\author{Totta \textsc{Kasemo}, Simone \textsc{Spork},\\ Christoph \textsc{Dietrich}, Richard \textsc{Lucius}, Emanuel \textsc{Heitlinger}} % Author name

\date{\today} % Date for the report

\begin{document}

\maketitle % Insert the title, author and date

\begin{center}
\begin{tabular}{l r}
%Date Performed: & January 1, 2012 \\ % Date the experiment was performed
%Partners: & James Smith \\ % Partner names
%& Mary Smith \\
%Instructor: & Professor Smith % Instructor/supervisor
\end{tabular}
\end{center}

% If you wish to include an abstract, uncomment the lines below
% \begin{abstract}
% Abstract text
% \end{abstract}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Overview}
This document contains data summaries and parts of analysis from the dual RNA-seq experiments of 
 \textit{Eimeria falciformis} infected mice carried our by Simone Spork. Basic read-outs
 such as number of transcripts detected in each sample from mouse and parasite are summarized, as
 well as compilations of the outcome of statistical tests for differential mRNA abundance from
 comparisons of different experimental groups. Mouse data was compared (correlated) with previously
 published (schmid12) microarray data from \textit{E. falciformis} infected mice.
 Hierarchical clustering illustrated by heatmaps is shown for different comparisons.\\
 The largest mRNA abundance differences on the parasite side are seen in comparisons between different
 time points post infection. Mouse strain, including Rag1-/- (adaptive immunity compromised mice) and 
 mouse immune status due to previous parasite exposure (second/challenge infection) seems to have less effects
 on mRNA abundance profiles in the parasite.
%\begin{center}\ce{}\end{center}

%----------------------------------------------------------------------------------------
%	SECTION  
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\subsection{Mice and infection procedure}
Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
Animal procedures were performed according to the German Animal Protection Laws as 
directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
tapwater was used instead of PBS for administration of oocysts. Briefly, NMRI mice were infected 
two times, which will be referred to as first and second infection. For the first infection, 
150 sporulated oocysts were administered in 100 $μ$L by oral gavage. During the first infection
of 60 mice, all animals were weighed every day. On day zero, before infection, as well as on day three,
five and seven days post infection, dpi, caeca from 3-4 sacrificed mice per time point were 
collected. Epithelial cells were isolated as described in Schmid et al.(schmid12). For challenge 
infection, mice recovered for four weeks before second infection. 
Recovery was monitored by weighing and visual inspection of fur. For the second infection, 
1500 sporulated oocysts were applied by oral gavage. Three mice were used as non-second infection control, 
referred to as day 0, second infection.

\subsection{Oocyst purification for infection and sequencing}
Sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate, 
\textit{E. falciformis} Bayer Haberkorn 1970, was used for all infections and parasite 
samples. The strain is maintained through passage in NMRI mice in our facilities as described 
elsewhere (schmid12).

\subsection{Sporozoite isolation}
Sporozoites were isolated from sporocysts by excystment. For this, sporocysts were incubated at 
37$°$C in DMEM containing 0.04\% tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) 
for 30 min. Sporozoites were purified by the method of Schmatz et al (schmatz--).

\subsection{RNA extraction}
Total RNA was isolated from infected epithelial cells, sporozoites and sporulated oocysts 
using Trizol according to the manufacturer’s protocol (Invitrogen). High quality  \emph{what is the meaning of 'high quality' here?}
RNA was used to produce an mRNA library using the Illumina’s TruSeq RNA Sample Preparation guide.
\emph{stolen from genome paper}
Sporozoites were stored in 1 mL Trizol until RNA-isolation. 
Total RNA was isolated using the PureLink RNA Mini Kit (Invitrogen).

\subsection{Sequence quality assessment and alignment}
Fastq\_quality\_filter was applied to Illumina Hiseq 2000 sequenced samples. 
Since this is not easily applicable to pair-end sequencing data, a low threshold was 
used on the hiseq data. A phred score of 10 was used, i.e., the probability of false base 
calling is one in ten. We further set q = 60, i.e., nine out of ten bases
or more is required to be correct in at least 
60\% of the bases in each read for the read sequence to be kept for further analysis.
This resulted in.......

\subsubsection{Alignment and reference genomes}
We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
files including mRNA sequences from both species were aligned against the dual reference genome
using TopHat2 (version 2.0.14, Trapnell09)/ Bowtie2 (version 1.1.2, Langmead12). Single-end and 
pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified 
for pair-end samples. Import into R was enabled by the R package Ballgown, which requires bam 
files to be processed by Tablemaker (Frazee15). Tablemaker in turn makes use of Cufflinks 
(version 2.1.1, Trapnell10).

%----------------------------------------------------------------------------------------
%	SECTION  
%----------------------------------------------------------------------------------------

\section{Table of reads per sample}
% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Wed Mar  2 14:36:30 2016
\setlength{\tabcolsep}{10pt}
\begin{table}[H]
\centering 
	\caption{Reads per sample listed.}
\begin{tabular}{*5l}    \toprule
Samples  & Mouse 	& \textit{E. falciformis}  & Percentage			& \#\textit{E.falciformis} \\
	& trancripts	& transcripts   	 & \textit{E. falciformis}	& genes \\ \midrule
NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
	{\color{Gray}NMRI\_2ndInf\_5dpi\_rep1} & {\color{Gray} 589643389.000} & {\color{Gray}3752923.000} & {\color{Gray} 0.632} &  {\color{Gray}5602.000} \\ 
NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
	{\color{Gray}NMRI\_1stInf\_3dpi\_rep1} & {\color{Gray}209723287.000} &  {\color{Gray}815820.000} &  {\color{Gray}0.388} &  {\color{Gray}5466.000} \\ 
C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
	 {\color{Gray}NMRI\_1stInf\_0dpi\_rep1} &  {\color{Gray}229165002.000} & {\color{Gray} 31459.000} &  {\color{Gray}0.014} & {\color{Gray} 1380.000} \\ 
NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
\bottomrule
\hline
\end{tabular}
\end{table}

Samples are listed from highest to lowest percentace of \textit{E. falciformis} transcripts per total number of transcripts (mouse plus \textit{E. falciformis}). Corresponding number of \textit{E. falciformis} genes are shown. Gray indicates that samples are excluded from analysis (see Methods for details)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%% in main MS %%%%%%%%%%%%%%%%%%

%\section{Experimental overview}
%	\setlength{\tabcolsep}{14pt}
%	\begin{table}[H]
%	\begin{center}
%	\caption{Experimental overview.}
%\begin{tabular}{*4l}    \toprule
%\textit{Day, 1st infection}  	& NMRI  & C57BL/6  & Rag1-/- \\ \midrule
%	0 (control)    & $10^4$ / $10^8$  & $10^2$ / $10^8$  & $10^2$ / $10^7$  \\ %rep1/rep2, NA=no replicate
%3  		& $10^5$ / $10^8$ & NA  & NA \\ 
%5  		& $10^7$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
%7  		& $10^8$ / $10^7$ & NA  & NA \\ 
%Oocysts 	& $10^8$ / NA & NA  & NA \\ 
%Sporozoites 	& $10^7$ / NA & NA  & NA \\ \midrule
%
%\textit{Day, 2nd infection}  	\\ \midrule
%0 (control)     & $10^3$ / $10^8$  &  NA  & NA  \\ %rep1/rep2, NA=no replicate
%3  		& $10^4$ / $10^8$ & NA  & NA \\ 
%5  		& $10^4$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
%7  		& $10^6$ / $10^8$ & NA  & NA \\ 
%	
%	
%	\bottomrule
% \hline
%\end{tabular}
%\end{center}
%\end{table}
%
%Replicate average of transcripts as order of magnitude for \textit{E. falciformis}/mouse. 
%Columns represent mouse strains used in infection experiments. Rows represent timepoints post 
%infection plus oocyst and sporozoite samples. The upper part of the table shows data for first 
%infection, and oocyst and sporozoite data. The lower part shows data for challenge infection. 
%Averages were calculated after sample exclusions (see Methods). For exact values, see Table 1.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MICROARRAY COMPARISON
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transcript coverage distribution and sample exclusions}
\subsection{Mouse transcript distributions}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{distributionsMm.pdf}
	\caption{Transcript coverage distribution (density) for mouse samples. 
 Without cutoff for a minimum read coverage over all samples, all samples display a bimodal distribution trend and an additional peak at zero by visual inspection. In our analyses, a cutoff of 3000 was applied.With this cutoff we can assume a negative binomial coverage distribution as required in the analysis pipline. Non-normalised read counts were used. For this visualization 0.1 was added to all read counts to allow to plot absent transcripts (zero reads) on a log-scale.}
%\end{center}
\end{figure}

\clearpage
\subsection{\textit{E. falciformis} transcript distributions}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{distributionsEf.pdf}
\caption{Transcript coverage density distribution density for \textit{E. falciformis} samples. Without cutoff for a minimum read coverage over all samples, most samples are in agreement with the assumed negative binomial distribution. However, two samples contradict this assumption and a cutoff was applied to evaluate whether peaks in these samples can be removed by excluding transcripts with very low coverage. The cutoff does not have a large effect on distributions and therefore two samples were removed from further 
analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1. On all kept samples, a cutoff of 100 was applied. Non-normalised read counts were used. To all read values, 0.1 was added to allow to plot density of absent transcripts on a log-scale.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MICROARRAY COMPARISON
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mouse RNA-seq data compared with mouse microarray data}
%%% figure 1
\begin{figure}[H]
\begin{center}
%\includegraphics[width=\textwidth]{Array144_vs_RNAseqN7.pdf} % Merge 2 plots: RUV and non-RUV 
\caption{Comparison of mouse data from RNA-seq, day 7, and microarray data, day 6 (schmid14). a. Data normalised with upperquartile method implemented in R package edgeR (version 3.14.0). $R^2$ = yy.	b. Normalised data as in a. and adjusted with RUV method as implemented in R package RUVseq (version....). $R^2$ = xx. Each axis shows log fold changes compared to control.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	MDS ANALYSIS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysing variance in data by multidimensional scaling}
\begin{figure}[H]
\begin{center}
	\includegraphics[width=\textwidth]{EfMm_4-mds.pdf} % Include 
	\caption{Multidimensional scaling shows no clear sample pattern due to technical variation. Effect on scaling is visualized for batch (left) and sequencing method (right). Upper panel displays \textit{E. falciformis} data and lower panel mouse data. Distances on plot approximate log2 fold changes. Scaling was done with Euclidean distance and "pairwise" gene selection method as implemented in R package limma (version 3.28.10).}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\clearpage
%\section{mRNA abundance differences between different experimental groups}
%\setlength{\tabcolsep}{8pt}
%\begin{table}[H]
%\begin{center}
%\caption{mRNA abundance differences between different experimental groups.}
%%\label{tab:table3}
%\begin{tabular}{*3l}    \toprule
%\textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
%	\textit{comparisons} 	    & (FDR$\leq$0.01) &  (FDR$\leq$0.01/0.05) \\ \midrule
%	NMRI 0 vs NMRI 3		& NA   & 274 \\
%	NMRI 0 vs NMRI 5		& NA   & 1736 \\
%	NMRI 0 vs NMRI 7		& NA   & 2802 \\
%	NMRI 3 vs NMRI 5     		& 111  & 1 \\
%	NMRI 3 vs NMRI 7  		& 1385 & 1407 \\ 
%	NMRI 5 vs NMRI 7  		& 1895 & 873 \\ 
%	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
%	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
%\textit{First and second infection} & 		 & 	 \\ 
%\textit{comparisons} 	    & 		& 	\\ \midrule
%	NMRI 3 1st vs NMRI 3 2nd  	& 0  & 5 \\
%	NMRI 5 1st vs NMRI 5 2nd  	& 0  & 1 \\
%	NMRI 7 1st vs NMRI 7 2nd  	& 0  & 902 \\
%	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 &  mouse \\
%	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
%\textit{Mouse strain comparisons} & 		 & 	 \\ \midrule
%	NMRI vs C57BL/6 		& 22 	& NA* \\		
%	NMRI vs Rag1-/-     		& 0 & NA* 	\\
%	C57BL/6 vs Rag1-/-  		& 0   & 356 	\\ \midrule
%\textit{Day post infection,} & 		 & 		 \\ 
%\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
%	Oocysts vs NMRI 3  	& 3310 & NA \\  
%	Oocysts vs NMRI 5	& 3605 & NA \\ 
%	Oocysts vs NMRI 7	& 3085 & NA \\ 
%	Oocysts vs sporozoites  & 3421 & NA \\
%	Sporozoites vs NMRI 3 	& 1663 & NA \\
%	Sporozoites vs NMRI 5 	& 1605 & NA \\
%	Sporozoites vs NMRI 7 	& 2473 & NA \\ \bottomrule
%	\hline
%\end{tabular}
%\end{center}
%\end{table}

%\textit{E. falciformis} data overview of pairwise comparisons and number of genes with differently abundant mRNAs per comparison. NMRI followed by number indicates day post infection (e.g. NMRI3 = \textit{E. falciformis} genes from NMRI mouse day 3 post infection). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are indluced. NAs are missing samples or not applicable for the species. NA* is due to missing NMRI day 0 sample from first infection.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	INPUT TO HEATMAP	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Differentially abundant mRNAs used for hierarchical clustering}

A selection of differently abundant mRNAs are used for hierarchical clustering of \textit{E. falciformis} life cycle relevant genes. In each comparison (see Table 3), the 500 differentially abundant mRNAs with lowest FDR are selected. In the next step, the 500 mRNAs from each comparison (or less) are joined (4935 genes) and only unique genes are selected (1618 genes). The 22 genes in the NMRI vs C57BL/6 comparison are not included in the \textit{E. falciformis} life cycle analysis.}

\setlength{\tabcolsep}{10pt}
\begin{table}[H]
%\begin{adjustwidth}{-10}{}
\small
\begin{center}
\caption{Differentially abundant mRNAs used for hierarchical clustering.}
\begin{tabular}{*2l}    \toprule
	\textit{Data description} & Number of genes \\ \midrule
	Sum of 1st infection NMRI sample differences	& 4935  \\ 
	(including oocysts and sporozoites)	\\	
	Used in hierarchical clustering (heatmap)  	& 1618 \\ 	\bottomrule	
\hline
\end{tabular}
\end{center}
%\end{adjustwidth}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	HEATMAPS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Hierarchical clustering...}
%All heatmaps shown are now with cutoff 0.05, even though we previously used 0.01 for the life cycle part. I will change it, but the difference seems small and I am really tired... Therefore, I also didn't write the legends for the 1st VS 2nd and mouse strain effect heatmaps. The 1st-2nd looks rediculous for Ef, for instance (it could only compile with cutree=2 or 1). So sorry abot the ugly end.

%\subsection{Life cycle analysis... \textit{E. falciformis}}
%\begin{figure}[H]
%\begin{center}
%\centering
%	\includegraphics[width=0.9\textwidth]{EfLifecycleHeatmap.pdf}  
%	\caption{\textit{E. falciformis} mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table ...).  \textit{E. falciformis} samples from seven days p.i. cluster together (NMRI mice only). For \textit{E. falciformis} day 7, NMRI, the pattern is most pronounced in cluster 1 (up). Within this cluster, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 6). Distinct groups of genes also define sporozoites (cluster 6: up) and oocysts (cluster 3: up, cluster 7: down). mRNA profiles on days three and five p.i. from all three mouse strains cluster together. These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites. The latter cluster closer with days 3 and 5 than with oocysts or day 7.}
%%\end{center}
%\end{figure}
%
%\subsection{Parasite development in mouse data}
%\begin{figure}[H]
%%\begin{center}
%\centering
%\includegraphics[width=0.9\textwidth]{MmLifecycleHeatmap.pdf}  
%\caption{Mouse mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table 3).
%Three out of four samples from day 7 p.i. (NMRI only) cluster together. These samples are characterized by genes in cluster 2 (up) as well as 3 and 4 (down). The fourth day 7 sample (challenge infection, second replicate) clusters with day 3 and 5 samples but upon visual inspection of cluster 4 diplays a profile similar to non-infected mice. The same sample is abnormal in the parasite profile (see Figure 5). NMRI and C57BL/6 uninfected samples cluster together, defined by clusters 3, 4 (up), and 1 (down). Day 3 and 5 samples cluster together, with Rag1-/- forming a separate group and Rag1-/- non-infected most distant in this sample cluster. All non-infected samples share a high mRNA abundance in cluster 3.}
%%\end{center}
%\end{figure}

\subsection{First versus second infection, \textit{E. falciformis}}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{Ef1st2ndHeatmap.pdf}  
	\caption{\textit{E. falciformis} with Euclidean distance .........as implemented in R package limma.}
\end{center}
\end{figure}

\subsection{First versus second infection, mouse}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{Mm1st2ndHeatmap.pdf}  
	\caption{Mouse.... with Euclidean distance .........as implemented in R package limma.}
\end{center}
\end{figure}

\subsection{Differences between mouse strains,\textit{E. falciformis}}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{EfStrainHeatmap.pdf}  
	\caption{\textit{E. falciformis} with Euclidean distance .........as implemented in R package limma.}
\end{center}
\end{figure}

\subsection{Differences between mouse strains, mouse}
\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth]{MmStrainHeatmap.pdf}  
	\caption{Mouse.... with Euclidean distance .........as implemented in R package limma.}
\end{center}
\end{figure}

\end{document}

### 2016-07-14 figuresANDmanuscript/table_rc.tex
MS reshuffled and folder structure changed: MS nw i figures folder

% latex table generated in R 3.3.0 by xtable 1.8-2 package
% Wed Jun 22 17:40:30 2016
\begin{table}[ht]
\centering
\begin{tabular}{lllll}
  \hline
Sample & Mouse transcripts & E. falciformis transcripts & Percentage E. falciformis & \#E. falciformis genes \\ 
  \hline
NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
  NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
  NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
  NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
  NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
  NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
  NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
  NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
  NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
  C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
  NMRI\_2ndInf\_5dpi\_rep1 & 589643389.000 & 3752923.000 & 0.632 & 5602.000 \\ 
  NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
  C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
  NMRI\_1stInf\_3dpi\_rep1 & 209723287.000 & 815820.000 & 0.388 & 5466.000 \\ 
  C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
  Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
  Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
  NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
  Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
  NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
  NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
  NMRI\_1stInf\_0dpi\_rep1 & 229165002.000 & 31459.000 & 0.014 & 1380.000 \\ 
  NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
  NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
  NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
  Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
  C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
  C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
  Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
  NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-07-14 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
changed figure path to figuresANDmanuscript/ in scripts

@@
 	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
+\textit{Day post infection,} & 		 & 		 \\ 
+\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
+	Oocysts vs NMRI 3  	& 3310 & NA \\  
+	Oocysts vs NMRI 5	& 3605 & NA \\ 
+	Oocysts vs NMRI 7	& 3085 & NA \\ 
+	Oocysts vs sporozoites  & 3421 & NA \\
+	Sporozoites vs NMRI 3 	& 1663 & NA \\
+	Sporozoites vs NMRI 5 	& 1605 & NA \\
+	Sporozoites vs NMRI 7 	& 2473 & NA \\ 
 \textit{First and second infection} & 		 & 	 \\ 
@@
 	C57BL/6 vs Rag1-/-  		& 0   & 356 	\\ \midrule
-\textit{Day post infection,} & 		 & 		 \\ 
-\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
-	Oocysts vs NMRI 3  	& 3310 & NA \\  
-	Oocysts vs NMRI 5	& 3605 & NA \\ 
-	Oocysts vs NMRI 7	& 3085 & NA \\ 
-	Oocysts vs sporozoites  & 3421 & NA \\
-	Sporozoites vs NMRI 3 	& 1663 & NA \\
-	Sporozoites vs NMRI 5 	& 1605 & NA \\
-	Sporozoites vs NMRI 7 	& 2473 & NA \\ \bottomrule
+\bottomrule
 	\hline
@@
 
-\textit{E.
-falciformis} data overview of pairwise comparisons and number of genes with differently abundant mRNAs per comparison.
-NMRI followed by number indicates day post infection (e.g.
-NMRI3 = \textit{E.
-falciformis} genes from NMRI mouse day 3 post infection).
-Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are indluced.
+Table 2 lists pairwise sample comparisons and number of genes with differently abundant mRNAs per comparison."NMRI" followed by number indicates day post infection (e.g.
+NMRI 3 = NMRI mouse on day 3 p.i.).
+Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are included.
 NAs are missing samples or not applicable for the species.
 NA* is due to missing NMRI day 0 sample from first infection.
+For \textit{E.
+falciformis} no mRNAs are significantly differently abundant between first and second infection, whereas in mouse there are some differences, especially on day 7 p.i.
+However, upon hierarchical clustering of these genes (union of 500 with lowest FDR), samples do not cluster according to first/second infection (Figure SX ?).
+In comparisons between mouse strains, 22 genes are detected to be different in the parasite data.
+The difference is however not between Rag1-/- mice as expected, but between NMRI and C57BL/6.
+For mouse....
 
@@
 	\caption{\textit{E.
-falciformis} mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table ...).
-\textit{E.
-falciformis} samples from seven days p.i.
-cluster together (NMRI mice only).
-For \textit{E.
-falciformis} day 7, NMRI, the pattern is most pronounced in cluster 1 (up).
-Within this cluster, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 6).
-Distinct groups of genes also define sporozoites (cluster 6: up) and oocysts (cluster 3: up, cluster 7: down).
-mRNA profiles on days three and five p.i.
-from all three mouse strains cluster together.
-These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
-The latter cluster closer with days 3 and 5 than with oocysts or day 7.}
+falciformis} mRNAs with significantly different abundance at different times p.i.
+in NMRI mice (see Table 2).
+Samples cluster according to expected developmental state of the parasite, without distinguishing between first and challenge infection.
+Hierarchical clustering was set to apply seven gene clusters.}
 %\end{center}
@@
 
+We applied hierarchical clustering on \textit{E.
+falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR).
+Samples from day 7 p.i.
+form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection.
+For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down).
+Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2).
+Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down).
+mRNA profiles on days 3 and 5 p.i.
+from all three mouse strains cluster together.
+These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
+Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.
+
 \clearpage
-\subsection{Parasite development in mouse data}
+\subsection{Parasite development reflected in mouse data}
 \begin{figure}[h!]
@@
 \includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
-\caption{Mouse mRNAs with significantly different abundance at different times post infection in NMRI mice (see Table 3).
-Three out of four samples from day 7 p.i.
-(NMRI only) cluster together.
-These samples are characterized by genes in cluster 2 (up) as well as 3 and 4 (down).
-The fourth day 7 sample (challenge infection, second replicate) clusters with day 3 and 5 samples but upon visual inspection of cluster 4 diplays a profile similar to non-infected mice.
-The same sample is abnormal in the parasite profile (see Figure 5).
-NMRI and C57BL/6 uninfected samples cluster together, defined by clusters 3, 4 (up), and 1 (down).
-Day 3 and 5 samples cluster together, with Rag1-/- forming a separate group and Rag1-/- non-infected most distant in this sample cluster.
-All non-infected samples share a high mRNA abundance in cluster 3.}
+\caption{Mouse mRNAs with significantly different abundance at different times p.i.
+in NMRI mice (see Table 2).
+Samples cluster according to expected developmental state of the parasite, without distinguishing between first and challenge infection.
+Rag1-/- mouse samples cluster together.
+Hierarchical clustering was set to apply four gene clusters.}
 %\end{center}
@@
 
-\clearpage
-\subsection{First versus second infection, mouse}
-\begin{figure}[h!]
-%\begin{center}
-\includegraphics[width=\linewidth]{Mm1st2ndHeatmap.pdf}  
-	\caption{Mouse mRNAs with significantly different abundance between first infection and challenge infection (see Table 3).
-Samples do not cluster into first and second infection clusters.
-The two major sample cluster each contain four challenge infection samples.
-}
-%\end{center}
-\end{figure}
+When clustering mouse mRNAs which were different in time p.i.
+comparisons in Table 2, three out of four samples from day 7 p.i.
+(NMRI only) cluster together.
+These samples are characterized by gene cluster 3 (up) as well as 2 and 4 (down).
+The fourth day 7-sample (challenge infection, second replicate) clusters with day 3 and 5 samples.
+Upon visual inspection of this sample in gene cluster 4, it diplays a profile similar to non-infected mice.
+The same sample is abnormal in the parasite profile (Figure 1).
+NMRI and C57BL/6 uninfected samples (0dpi) cluster together, defined by gene clusters 4 (up), and 1 (down).
+Rag1-/- day 0 samples are however less pronounced in gene cluster 1.
+Day 3 and 5 samples cluster together, with Rag1-/- forming a separate group and Rag1-/- non-infected most distant in this sample cluster.
 
@@
 
+\clearpage
 \subsection*{GO term enrichments in heatmap clusters}

### 2016-07-14 4_compareArray.R

@@ -128,7 +128,7 @@ names(RNAseq.Array.logFC) <- gsub(".y", ".ruved", names(RNAseq.Array.logFC))
 cor(RNAseq.Array.logFC[,3:14])
 cor(RNAseq.Array.logFC[,3:14], method="spearman")
 
-pdf("figures/Array_vs_RNAseq_pairs.pdf")
+pdf("figuresANDmanuscript/Array_vs_RNAseq_pairs.pdf")
 #ggpairs(RNAseq.Array.logFC[, 3:14], alpha=0.2) + theme_bw()
 ggpairs(RNAseq.Array.logFC[, 3:14], mapping=aes(alpha=0.2)) + theme_bw()
 dev.off()
@@ -139,7 +139,7 @@ dev.off()
 ## If these are meant to be aesthetics, submit them using the 'mapping' variable within 
 ## ggpairs with ggplot2::aes or ggplot2::aes_string.
 ##
-pdf("figures/Array144vsRNAseqN7.pdf")
+pdf("figuresANDmanuscript/Array144vsRNAseqN7.pdf")
 ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0.plain)) +
     geom_point(alpha=0.8) +
         stat_density2d(aes(alpha=..level.., fill=..level..), size=2,                                                                          geom="polygon") +
@@ -149,7 +149,7 @@ ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0.plain)) +
                         theme_bw()
 dev.off()
 
-pdf("figures/Array144vsRNAseqN7_ruved.pdf")
+pdf("figuresANDmanuscript/Array144vsRNAseqN7_ruved.pdf")
 ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0.ruved)) +
     geom_point(alpha=0.8) +
         stat_density2d(aes(alpha=..level.., fill=..level..), size=2,                                                                          geom="polygon") +

### 2016-07-14 7_omcl_comparison.R

@@ -70,7 +70,7 @@ scaling.f <- max(CS)/CS
 
 RC.ortho[, 3:ncol(RC.ortho)] <- t(t(RC.ortho[, 3:ncol(RC.ortho)])*scaling.f)
 
-pdf("figures/RC_correlation_Efal_Eten.pdf")
+pdf("figuresANDmanuscript/RC_correlation_Efal_Eten.pdf")
 pheatmap(cor(RC.ortho[,3:ncol(RC.ortho)], method="spearman")[-(1:3),1:3],
          display_numbers=TRUE)
 dev.off()
@@ -86,7 +86,7 @@ mean.columns <- function(x){
   
 RC.ortho.mean <- as.data.frame(cbind(RC.ortho[, 1:5], mean.columns(RC.ortho)))
 
-pdf("figures/RC_mean_correlation_Efal_Eten.pdf")
+pdf("figuresANDmanuscript/RC_mean_correlation_Efal_Eten.pdf")
 pheatmap(cor(RC.ortho.mean[,3:ncol(RC.ortho.mean)], method="spearman")[-(1:3),1:3],
          display_numbers=TRUE)
 dev.off()
@@ -113,14 +113,14 @@ FC.ortho[,3:ncol(FC.ortho)] <- apply(FC.ortho[,3:ncol(FC.ortho)],
                                      2, function(x) as.numeric(as.character(x)))
 
 
-pdf("figures/FC_correlation_Efal_Eten.pdf")
+pdf("figuresANDmanuscript/FC_correlation_Efal_Eten.pdf")
 pheatmap(cor(FC.ortho[, 3:ncol(FC.ortho)], method="spearman")[-c(1:2),1:2],
          display_numbers=TRUE)
 dev.off()
 
 
 
-pdf("figures/TenellaGamVsMer_vs_N3vsN7.pdf")
+pdf("figuresANDmanuscript/TenellaGamVsMer_vs_N3vsN7.pdf")
 ggplot(FC.ortho, aes(log2.fold.change.GamVsMer, logFC.N3vsN7*-1)) +
     geom_point(alpha=0.8) +
         stat_density2d(aes(alpha=..level.., fill=..level..), size=2,                                                                          geom="polygon") +
@@ -131,7 +131,7 @@ ggplot(FC.ortho, aes(log2.fold.change.GamVsMer, logFC.N3vsN7*-1)) +
 dev.off()
 
 
-pdf("figures/TenellaGamVsMer_vs_N5vsN7.pdf")
[…]

### 2016-07-20 7_omcl_comparison.R
deletions, .gitignore update and 7_omcl changes because script was not running

@@ -1,3 +1,5 @@
+## Description of script
+
 library(ggplot2)
 library(reshape)
 library(plyr)
@@ -62,7 +64,7 @@ RC.ortho <- merge(all.orthologs,
 
 RC.ortho <- merge(RC.ortho, cpm(Ef.1st.pass.model[[4]]),
                   by.x="Efa", by.y=0)
-RC.ortho[,3:ncol(RC.orhto)] <- apply(RC.ortho[,3:ncol(RC.orhto)], 2,
+RC.ortho[,3:ncol(RC.ortho)] <- apply(RC.ortho[,3:ncol(RC.ortho)], 2,
                                      function (x) as.numeric(as.character(x)))
 
 CS <- colSums(RC.ortho[,3:ncol(RC.ortho)])

### 2016-07-20 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex

@@
 %\RequirePackage{hyperref}
-%\usepackage[utf8]{inputenc} %unicode support
+\usepackage[utf8]{inputenc} %unicode support
 %\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
 %\usepackage{siunitx}
+\usepackage[colorlinks]{hyperref} % To color blocks of text 
+%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
+\usepackage{changepage}
+\usepackage{booktabs} % For prettier tables
+\usepackage{float} % To control positions of e.g. tables or figures
+%\usepackage[section]{placeins}
+\usepackage{graphicx}
+\graphicspath{ {../} }
 
@@
 
-\usepackage[colorlinks]{hyperref} % To color blocks of text 
-%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
-\usepackage{changepage}
-\usepackage{booktabs} % For prettier tables
-\usepackage{float} % To control positions of e.g. tables or figures
-\usepackage[section]{placeins}
-\usepackage{graphicx}
-\graphicspath{ {../} }
 %\def\includegraphic{}
@@
 \author[
-   addressref={aff1},
+%	addressref={aff1},
+%	noteref={n1},
 ]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
 \author[
-   addressref={aff1},
+%	addressref={aff1},
 ]{\inits{SS}\fnm{Simone} \snm{Spork}}
 \author[
-   addressref={aff2},
+%	addressref={aff2},
+%	noteref={n1},
 ]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
 \author[
-   addressref={aff1},
+%	addressref={aff1},
 ]{\inits{RL}\fnm{Richard} \snm{Lucius}}
 \author[
-   addressref={aff1, aff3},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},                       % id of corresponding address, if any
-   noteref={n1},                        % id's of article notes, if any
+%	addressref={aff1, aff3},                   % id's of addresses, e.g. {aff1,aff2}
+	corref={aff1, aff3},                       % id of corresponding address, if any
+%	noteref={n1},                        % id's of article notes, if any
 	email={emanuel.heitlinger@hu-berlin.de}   % email address
@@
 	\street{Philippstr. 13, Haus 14},                     %
-  %\postcode{}                                % post or zip code
+	\postcode{10115}                                % post or zip code
 	\city{Berlin},                              % city
@@
 
-%\begin{artnotes}
+\begin{artnotes}
 %\note{Sample of title note}     % note to the article
 %\note[id=n1]{Equal contributor} % note, connected to author
-%\end{artnotes}
+%\note[id=n2]
+\end{artnotes}
 
@@
 \kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
-\kwd{article}
-\kwd{author}
+%\kwd{article}
+%\kwd{author}
 \end{keyword}
@@
 \end{abstractbox}
-%
 \end{fmbox}% uncomment this for twcolumn layout
-
 \end{frontmatter}
@@
 	Sporozoites vs NMRI 5 	& 1605 & NA \\
-	Sporozoites vs NMRI 7 	& 2473 & NA \\ 
+	Sporozoites vs NMRI 7 	& 2473 & NA \\ \midrule
 \textit{First and second infection} & 		 & 	 \\ 
@@
 
-\subsubsection{Data normalisation and sample exclusions}
-Count data was normalised in R, using the upperquartile method as implemented in edgeR (version 3.14.0).
-Briefly, normalisation assumes a negative binomial distribution of data and dispersed variance for genes with higher mean abundance.
-The normalisation shrinks the dispersion of variance towards a Poisson (linear) relationship between mean and variance.
-We evaluated the density distribution of our data and two samples contradicted the assumed negative binomial distribution and were excluded from further analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...).
+\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
+Complete scripts are available at git@github.com:derele/Ef\_RNAseq.git.
+Count data was analysed in R, using the upperquartile normalisation method as implemented in edgeR (version 3.14.0).
+Briefly, genes with zero coverage in all samples (libraries) are removed and normalisation factors are calculated for the 75\% quantile for each library.
+This  normalisation assumes a negative binomial distribution of read densities.
+Two samples contradicted the assumption (parasite data) and both mouse and parasite data from these samples were excluded from further analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...).
+We used a the edgeR pipleine which estimates dispersion values for each gene (provided that a design matrix is defined).
+The method fits data for each gene to a negative binomial distribution using a generalized linear model (GLM) with the function glmFit() and glmLRT().
+
+........... include outcome of MDS analysis here?
 
@@
 falciformis} life cycle relevant genes.
-In each comparison (see Table 3), the 500 differentially abundant mRNAs with lowest FDR are selected.
-In the next step, the 500 mRNAs from each comparison (or less) are joined (4935 genes) and only unique genes are selected (1618 genes).
+In each comparison (see Table 3), the union of the 500 differentially abundant mRNAs with lowest FDR are selected.
+In the next step, the 500 mRNAs from each comparison (or less) are joined.
+For \textit{E.
+falciformis} time p.i.
+comparison, this resulted in 4935 genes of which only the 1618 unique genes were selected.
 The 22 genes in the NMRI vs C57BL/6 comparison are not included in the \textit{E.
 falciformis} life cycle analysis.
-
-All parasite samples were included in heatmap of genes detected as differentially abundant in timepoint comparisons.
-
-On scale bar in heatmaps, 0 is mean mRNA abundance for each gene (row).
-Up (green) and down-regulation (brown) numbers denote number of standard deviations from row mean.
-Hierarchical clustering was performed using with Euclidean distances, method 'complete' (R package base).
+The same comparisons for mouse yielded 8052 genes in total and 1313 unique ones.
+In heatmaps, all samples, i.e.
+also samples which themselves did not have any significantly different mRNAs according to our selection, were included in hierarchical clustering.
+Scale bar in heatmaps show 0 as mean mRNA abundance for each gene (row).
+Up (green) and down-regulation (brown) denote number of standard deviations from 0, i.e., row mean.
+Hierarchical clustering was performed using with Euclidean distances, using complete linkage ('complete', R package base).
 
@@
 \begin{center}
-\caption{Differentially abundant \textit{E. falciformis} mRNAs used for hierarchical clustering.}
-\begin{tabular}{*2l}    \toprule
-	\textit{Data description} & Number of genes \\ \midrule
-	Sum of 1st infection NMRI sample differences	& 4935  \\ 
-	(including oocysts and sporozoites)	\\	
-	Used in hierarchical clustering (heatmap)  	& 1618 \\ 	\bottomrule	
+\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
+\begin{tabular}{*3l}    \toprule
+	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
+	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
+	(including oocysts and sporozoites if appl.)	\\	
+	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
 \hline

### 2016-07-20 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
First additions to MS.tex and some gitignore

--- 2016-07-14 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
@@
 \section{Experimental overview}
-Replicate average of transcripts as order of magnitude for \textit{E.
-falciformis}/mouse are indicated in Table....
-Columns in table represent mouse strains used in infection experiments.
-Rows represent timepoints post infection plus oocyst and sporozoite samples.
-The upper part of the table shows data for first infection, and oocyst and sporozoite data.
-The lower part shows data for challenge infection.
-Averages were calculated after sample exclusions (see Methods).
-For exact values, see Table 1.
+We performed mRNA sequencing of caecum tissue of mice infected with
+the parasite \textit{E. falciformis} at different timepoints and for
+infections of different mouse strains.
+
+Replicate average of transcripts as order of magnitude for
+\textit{E. falciformis}/mouse are indicated in Table.... Columns in
+table represent mouse strains used in infection experiments. Rows
+represent timepoints post infection plus oocyst and sporozoite
+samples. The upper part of the table shows data for first infection,
+and oocyst and sporozoite data. The lower part shows data for
+challenge infection. Averages were calculated after sample exclusions
+(see Methods). For exact values, see Table 1.
 
@@
 
-Table 2 lists pairwise sample comparisons and number of genes with differently abundant mRNAs per comparison."NMRI" followed by number indicates day post infection (e.g.
-NMRI 3 = NMRI mouse on day 3 p.i.).
-Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are included.
-NAs are missing samples or not applicable for the species.
-NA* is due to missing NMRI day 0 sample from first infection.
-For \textit{E.
-falciformis} no mRNAs are significantly differently abundant between first and second infection, whereas in mouse there are some differences, especially on day 7 p.i.
-However, upon hierarchical clustering of these genes (union of 500 with lowest FDR), samples do not cluster according to first/second infection (Figure SX ?).
-In comparisons between mouse strains, 22 genes are detected to be different in the parasite data.
-The difference is however not between Rag1-/- mice as expected, but between NMRI and C57BL/6.
-For mouse....
+Table 2 lists pairwise sample comparisons and number of genes with
+differently abundant mRNAs per comparison."NMRI" followed by number
+indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
+p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
+implemented in edgeR are included. NAs are missing samples or not
+applicable for the species. NA* is due to missing NMRI day 0 sample
+from first infection. For \textit{E. falciformis} no mRNAs are
+significantly differently abundant between first and second infection,
+whereas in mouse there are some differences, especially on day 7
+p.i. However, upon hierarchical clustering of these genes (union of
+500 with lowest FDR), samples do not cluster according to first/second
+infection (Figure SX ?). In comparisons between mouse strains, 22
+genes are detected to be different in the parasite data. The
+difference is however not between Rag1-/- mice as expected, but
+between NMRI and C57BL/6. For mouse....
 
@@
 \section*{Additional Files}
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
+  \subsection*{Additional file 1 --- Raw and normalized counts }
+  Raw and normalized counts for individual samples in our
+  study (two compressed csv files).
 
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
+  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
+  Fold-changes, p-values and false discovery rates (adjusted p-values)
+  are given for all tested contrasts (one compressed csv file).
+
+  \subsection*{Additional file 3 --- Additional methods and resuts }
+  % this would be cool but requires special referencing in the text
+  % (hard to keep track and hard to fit in the journals referencing
+  % system). But we can try.
+  Document containing additonal figures and summary tables (pdf).
+
+\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
+  Tables listing all tested gene sets and resulting significatn GO
+  terms.
 

### 2016-07-20 7_omcl_comparison.R
removed scaling from ortholog scritp... not required for correlation analysis

@@ -1,7 +1,6 @@
 ## Description of script
 
 library(ggplot2)
-=======
 library(pheatmap)
 library(reshape)
 library(ggplot2)
@@ -154,16 +153,8 @@ get.ortholog.RC <- function(){
     return(foobar)
 }
 
-RC.ortho <- merge(RC.ortho, cpm(Ef.1st.pass.model[[4]]),
-                  by.x="Efa", by.y=0)
-RC.ortho[,3:ncol(RC.ortho)] <- apply(RC.ortho[,3:ncol(RC.ortho)], 2,
-                                     function (x) as.numeric(as.character(x)))
-=======
 RC.ortho <- get.ortholog.RC()
-
-CS <- colSums(RC.ortho[,4:ncol(RC.ortho)])
-scaling.f <- max(CS)/CS
-RC.ortho[, 4:ncol(RC.ortho)] <- t(t(RC.ortho[, 4:ncol(RC.ortho)])*scaling.f)
+RC.ortho[is.na(RC.ortho)] <- 0
 
 pdf("figuresANDmanuscript/RC_correlation_Efal_Ete_Tgo.pdf", heigh=15, width=15)
 pheatmap(cor(RC.ortho[,4:ncol(RC.ortho)], method="spearman"),

### 2016-07-20 7_omcl_comparison.R
further clean up in 7_*.R to remove all fold cahnge analysis

@@ -74,20 +74,18 @@ Walker.data <- read.csv("data/12864_2015_1298_MOESM3_ESM.csv", skip=1)
 names(Walker.data)[5:8] <- paste(names(Walker.data)[5:8], "GamVsMer", sep=".")
 names(Walker.data) <- gsub("\\.1", "\\.GamVsSpor", names(Walker.data))
 
-## make thins numeric
-Walker.data <- Walker.data[, c("gene.ID", "gametocyte", "merozoite", "sporozoite",
-                               "log2.fold.change.GamVsSpor", "log2.fold.change.GamVsMer")] 
+Walker.data <- Walker.data[, c("gene.ID", "gametocyte", "merozoite", "sporozoite")]
 
 Walker.data <- merge(all.orthologs, Walker.data, by.x="Ete", by.y="gene.ID")
 
 Walker.RC <- Walker.data[, c("Efa", "Ete", "Tgo", "gametocyte", "merozoite", "sporozoite")]
 names(Walker.RC) <-  c("Efa", "Ete", "Tgo", "Walker.Gam", "Walker.Mer", "Walker.Spo")
 
+## log2 transform to make it compatible with ToxoDB
 Walker.RC[,c("Walker.Gam", "Walker.Mer", "Walker.Spo")] <-
     log2(Walker.RC[,c("Walker.Gam", "Walker.Mer", "Walker.Spo")])
-Walker.RC[is.na(Walker.RC)] <- 0
 
-Walker.FC <-  Walker.data[, c("Efa", "Ete", "Tgo", "log2.fold.change.GamVsMer", "log2.fold.change.GamVsSpor")]
+Walker.RC[is.na(Walker.RC)] <- 0
 
 ## More from Reid et al via ToxoDB (tedious manual downloads)
 read.Reid.data <- function(){
@@ -110,7 +108,6 @@ Reid.data <- read.Reid.data()
 Reid.data <- merge(Reid.data, all.orthologs, by="Ete")
 
 Reid.RC <- Reid.data[, c("Efa", "Ete", "Tgo", "Reid.Mer","Reid.spOoc", "Reid.Sp", "Reid.Ooc")]
-Reid.FC <- Reid.data[, c("Efa", "Ete", "Tgo", "FC.OOvsMer","FC.OOvsspOO", "FC.OOvsSp")]
 
 ## Data from Hehl lab on toxo in the cat
 read.Hehl.data <- function(){
[…]

### 2016-07-20 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
A first day of work on the MS tex file

@@
   \section*{Results}
-Text for this section \ldots
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experimental overview}
-We performed mRNA sequencing of caecum tissue of mice infected with
-the parasite \textit{E. falciformis} at different timepoints and for
-infections of different mouse strains.
 
-Replicate average of transcripts as order of magnitude for
-\textit{E. falciformis}/mouse are indicated in Table.... Columns in
-table represent mouse strains used in infection experiments. Rows
-represent timepoints post infection plus oocyst and sporozoite
-samples. The upper part of the table shows data for first infection,
-and oocyst and sporozoite data. The lower part shows data for
-challenge infection. Averages were calculated after sample exclusions
-(see Methods). For exact values, see Table 1.
+  \section*{Experimental overview}
+We performed mRNA sequencing of caecum tissue from mice infected with
+the apicomplexan parasite \textit{E. falciformis}. For each replicate
+sample we enriched caecum tissue from three infected mice for
+epithelial cells. We extracted mRNA and constructed sequencing
+libraries for different timepoints after infection and mouse strains
+with assumed different immunocompetence.
+%% we have to bring in Simones data first with two sentences and a
+%% figure or write it like this
+Additonally we obtained data for challenge infections of mice
+immunized one month prior to the experiment. For the parasite oocysts
+and sporozoites were included as ``environmental'' lifecycle stages.
 
-	\setlength{\tabcolsep}{14pt}
-	\begin{table}[H]
-	\small
-	\begin{center}
-	\caption{Experimental overview with number of read for parasite/mouse indicated.}
-\begin{tabular}{*4l}    \toprule
-\textit{Day, 1st infection}  	& NMRI  & C57BL/6  & Rag1-/- \\ \midrule
-	0 (control)    & $10^4$ / $10^8$  & $10^2$ / $10^8$  & $10^2$ / $10^7$  \\ %rep1/rep2, NA=no replicate
-3  		& $10^5$ / $10^8$ & NA  & NA \\ 
-5  		& $10^7$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
-7  		& $10^8$ / $10^7$ & NA  & NA \\ 
-Oocysts 	& $10^8$ / NA & NA  & NA \\ 
-Sporozoites 	& $10^7$ / NA & NA  & NA \\ \midrule
+The number of sequencing reads obtained for individual replicates
+ranged between 111111 (samply NMRIxyz) and 222222 (NMRIwyv).
 
-\textit{Day, 2nd infection}  	\\ \midrule
-0 (control)     & $10^3$ / $10^8$  &  NA  & NA  \\ %rep1/rep2, NA=no replicate
-3  		& $10^4$ / $10^8$ & NA  & NA \\ 
-5  		& $10^4$ / $10^8$ & $10^5$ / $10^7$  & $10^5$ / $10^7$ \\
-7  		& $10^6$ / $10^8$ & NA  & NA \\
+As samples and idividual replicates were sequenced in batches to
+different depth and using different instrumentation (see additional
+file 1 %%{or include sequencing even in Table 1?
+) we performed quality controls confirming the abscence of batch
+effects influencing analysis and quality of our results (additional
+file xyc).
 
-	\bottomrule
- \hline
-\end{tabular}
-\end{center}
-\end{table}
+\section*{Parasite and host transcriptomes can be assessed in parallel}
+
+A maximum of 75\% (sample NMRIaaga) and a minimum of 0.01\% (sample
+NMRIagag) of mapped reads could be assigened to the
+\textit{E. falciformis} genome in samples considered infected. We
+excluded samples NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 due
+to an exptremely low parasite contribution to the overall
+transcriptome. Technically this made it possible to obtain read counts
+following negative bionmial distributions after normalization between
+samples (see adddtional file x). Judging from their mRNA output those
+samples can be considered non-infected.On the other end of the
+spectruem this means that at 7dpi the majority of mRNA originated from
+the parasite as opposed to the host transcriptome (Figure 1).
+
+%%%%%
+figure 1 label
+cotribution of mouse vs. parasite transcripts to the overall
+transcriptional output of the analysed tissue for all samples.
+%%%%
+
+The total number of sequencing reads obtained as well as those reads
+mapped to the transcriptome of either \textit{E. falciformis} and the
+mouse are indicated in Table 1 for all replicates.
+
+%% Before was not pretty. All those NA's. If people are like me they
+%% hate NAs in tables ;-). I think you sould make the above table from
+%% data with xtable.
+
+%% best use a straight 2d table with...
+
+% label & mouse strain & parasite dpi & naive/challenge & number of
+% reads & #mapped to mouse & #mapped to \textit{E. falciformis}
+
+%% column name "naive/challenge" can change
+
+%% as technical as it seems people will want to know what our labels
+%% mean (they appear in figures throughout) therfore it is good to
+%% have them in table 1
+
+\section*{The mouse transcriptome changes in the course of infection infection}
+
+Statistical testing for differences between infected and uninfected
+mice, revealed changes in mRNA abundance becoming more pronounced
+(involving more genes) at later timepoints of infection (Table 1).
+
+The observed fold-changes were in agreement with previously published
+microarry data. Fold-Change data obtained from mice at 6dpi on Agilent
+microarrays (Schmidt et al 2012) or 7dpi (our RNAseq data) shows a
+strong correlateion (Spearman's \rho = 333; $R^2$ from a linear model
+= 444 ; see also figure 2). Considering both and technical differences
+between the tow methods and the biological changes along the lifecycle
+this confirms the adequacy of our methods for assessing the host
+transcriptome.
+
+We selected mouse
+%%% This is from methods but all the results should be presented here:
+
+%% Selecting at most at most 500 genes differentially abundant with
+%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
+%% each comparison (or less) are joined. For \textit{E. falciformis} time
+%% p.i. comparison, this resulted in 1618 unique genes selected as
+%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
+%% are not included in the \textit{E. falciformis} life cycle
+%% analysis. The same comparisons for mouse yielded 8052 genes in total
+%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
+%% which themselves did not have any significantly different mRNAs
+%% according to our selection, were included in hierarchical
+%% clustering.
+
+Those changes are mainly related to the onset of an
+adaptive immune response.
+
+\section*{The lifecycle of Eimeria is characterized by large changes in mRNA abundance}
+
+%%% This is from methods but all the results should be presented here:
+
+%% Selecting at most at most 500 genes differentially abundant with
+%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
+%% each comparison (or less) are joined. For \textit{E. falciformis} time
+%% p.i. comparison, this resulted in 1618 unique genes selected as
+%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
+%% are not included in the \textit{E. falciformis} life cycle
+%% analysis. The same comparisons for mouse yielded 8052 genes in total
+%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
+%% which themselves did not have any significantly different mRNAs
+%% according to our selection, were included in hierarchical
+%% clustering.
+
+%% This might be figure label
+%% Scale bar in heatmaps show 0 as mean mRNA abundance for
+%% each gene (row). Up (green) and down-regulation (brown) denote number
+%% of standard deviations from 0, i.e., row mean.
 
@@
 \subsection{Mice and infection procedure}
-Three strains of mice were used in our experiments: NMRI (Charles River Laboratories, Sulzfeld,
-Germany), C57BL/6 (), and Rag1-/- on C57BL/6 background (gift from Susanne Hartmann, FU?). 
-Animal procedures were performed according to the German Animal Protection Laws as 
-directed and approved by the overseeing authority Landesamt fuer Gesundheit und Soziales 
-(Berlin, Germany). Animals where infected as described by Schmid et al., (Schmid12), but
-tapwater was used instead of PBS for administration of oocysts. Briefly, NMRI mice were infected 
-two times, which will be referred to as first and second infection. For the first infection, 
-150 sporulated oocysts were administered in 100 $μ$L by oral gavage. During the first infection
-of 60 mice, all animals were weighed every day. On day zero, before infection, as well as on day three,
-five and seven days post infection, dpi, caeca from 3-4 sacrificed mice per time point were 
-collected. Epithelial cells were isolated as described in Schmid et al.(schmid12). For challenge 
-infection, mice recovered for four weeks before second infection. 
-Recovery was monitored by weighing and visual inspection of fur. For the second infection, 
-1500 sporulated oocysts were applied by oral gavage. Three mice were used as non-second infection control, 
-referred to as day 0, second infection.
+Three strains of mice were used in our experiments: NMRI (Charles
+River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
+C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
+procedures were performed according to the German Animal Protection
+Laws as directed and approved by the overseeing authority Landesamt
+fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
+as described by Schmid et al., (Schmid12), but tapwater was used
+instead of PBS for administration of oocysts. Briefly, NMRI mice were
+infected two times, which will be referred to as first and second
+infection. For the first infection, 150 sporulated oocysts were
+administered in 100 $μ$L by oral gavage. During the first infection of
+60 mice, all animals were weighed every day. On day zero, before
+infection, as well as on day three, five and seven days post
+infection, dpi, caeca from 3-4 sacrificed mice per time point were
+collected. Epithelial cells were isolated as described in Schmid et
+al.(schmid12). For challenge infection, mice recovered for four weeks
+before second infection.  Recovery was monitored by weighing and
+visual inspection of fur. For the second infection, 1500 sporulated
+oocysts were applied by oral gavage. Three mice were used as
+non-second infection control, referred to as day 0, second infection.
 
 \subsection{Oocyst purification for infection and sequencing}
-Sporulated oocysts were purified by flotation from feces stored in potassium dichromate and 
-administered orally in 100 uL tapwater. One \textit{E. falciformis} isolate, 
-\textit{E. falciformis} Bayer Haberkorn 1970, was used for all infections and parasite 
-samples. The strain is maintained through passage in NMRI mice in our facilities as described 
-elsewhere (schmid12).
+Sporulated oocysts were purified by flotation from feces stored in
+potassium dichromate and administered orally in 100 uL tapwater. One
+\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
+Haberkorn 1970, was used for all infections and parasite samples. The
+strain is maintained through passage in NMRI mice in our facilities as
+described elsewhere (schmid12).
 
 \subsection{Sporozoite isolation}
-Sporozoites were isolated from sporocysts by excystment. For this, sporocysts were incubated at 
-37$°$C in DMEM containing 0.04\% tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) 
-for 30 min. Sporozoites were purified by the method of Schmatz et al (schmatz--).
+Sporozoites were isolated from sporocysts by excystment. For this,
+sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
+tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
+30 min. Sporozoites were purified by the method of Schmatz et al
+(schmatz--).
 
 \subsection{RNA extraction}
-Total RNA was isolated from infected epithelial cells, sporozoites and sporulated oocysts 
-using Trizol according to the manufacturer’s protocol (Invitrogen). High quality  \emph{what is the meaning of 'high quality' here?}
-RNA was used to produce an mRNA library using the Illumina’s TruSeq RNA Sample Preparation guide.
-\emph{stolen from genome paper}
-Sporozoites were stored in 1 mL Trizol until RNA-isolation. 
-Total RNA was isolated using the PureLink RNA Mini Kit (Invitrogen) and reverse transcribed into cDNA.
+Total RNA was isolated from infected epithelial cells, sporozoites and
+sporulated oocysts using Trizol according to the manufacturer’s
+protocol (Invitrogen). High quality \emph{what is the meaning of 'high
+  quality' here?}  RNA was used to produce an mRNA library using the
+Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
+  genome paper} Sporozoites were stored in 1 mL Trizol until
+RNA-isolation.  Total RNA was isolated using the PureLink RNA Mini Kit
+(Invitrogen) and reverse transcribed into cDNA.
 
 \subsection{Sequencing, sequence quality assessment and alignment}
-cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as specified in SI xx (both unstranded).
-A fastq\_quality\_filter (FASTQ-toolkit, version 0.0.14, available at https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina Hiseq 2000 samples after replacing "N" bases by "." annotation.
-A phred score of 10 was applied. We further set q = 60. These settings require that nine out of ten bases
-or more are correct in at least 60\% of the bases for each read.
+cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
+specified in SI xx (both unstranded). A fastq\_quality\_filter
+(FASTQ-toolkit, version 0.0.14, available at
+https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
+Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
+phred score of 10 was applied. We further set q = 60. These settings
+require that nine out of ten bases or more are correct in at least
+60\% of the bases for each read.
 
-\subsubsection{Alignment and reference genomes}
-We used the published \textit{Mus musculus} mm10 assembly (Genome Reference Consortium Mouse 
-Build 38, GCA\_000001635.2) as reference genome including annotations for mouse data. The
-\textit{E. falciformis} genome (Heitlinger14) was downloaded from ToxoDB (Gajria07). For the
-alignment, the mouse and parasite genome files were merged into a dual reference genome, and 
-files including mRNA sequences from both species were aligned against the dual reference genome
-using TopHat2 (version 2.0.14, Trapnell09) with -G specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual genome.
-Single-end and pair-end sequence samples were aligned separately with library type 'fr-unstranded' specified for pair-end samples.
-Import into R was enabled by the R package Ballgown, which requires bam files to be processed by Tablemaker (Frazee15), in our case used with -qW -G specified.
-Tablemaker in turn makes use of Cufflinks (version 2.1.1, Trapnell10).
+\subsection{Alignment and reference genomes}
+We used the published \textit{Mus musculus} mm10 assembly (Genome
+Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
+genome including annotations for mouse data. The
+\textit{E. falciformis} genome (Heitlinger14) was downloaded from
+ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
+files were merged into a dual reference genome, and files including
+mRNA sequences from both species were aligned against the dual
+reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
+specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
+genome. Single-end and pair-end sequence samples were aligned
+separately with library type 'fr-unstranded' specified for pair-end
+samples. Import into R was enabled by the R package Ballgown, which
+requires bam files to be processed by Tablemaker (Frazee15), in our
+case used with -qW -G specified. Tablemaker in turn makes use of
+Cufflinks (version 2.1.1, Trapnell10).
 
 \subsection{Differential mRNA abundance, data normalisation and sample exclusions}
-Complete scripts are available at git@github.com:derele/Ef\_RNAseq.git.
-Count data was analysed in R, using the upperquartile normalisation method as implemented in edgeR (version 3.14.0).
-Briefly, genes with zero coverage in all samples (libraries) are removed and normalisation factors are calculated for the 75\% quantile for each library.
-This  normalisation assumes a negative binomial distribution of read densities.
-Two samples contradicted the assumption (parasite data) and both mouse and parasite data from these samples were excluded from further analysis: NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...).
-We used a the edgeR pipleine which estimates dispersion values for each gene (provided that a design matrix is defined).
-The method fits data for each gene to a negative binomial distribution using a generalized linear model (GLM) with the function glmFit() and glmLRT().
+Count data was normalized using the R-package edgeR (version 3.14.0;
+cite) with the upperquartile normalisation method. Briefly, genes with
+zero coverage in all samples (libraries) are removed and normalisation
+factors are calculated for the 75\% quantile for each library. This
+normalisation is suitable for read densities following a negative
+binomial distribution. Two samples contradicted this assumption
+(parasite data) for later modelling and both mouse and parasite data
+from these samples were excluded from further analysis:
+NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
+then fits a generalized linear model (GLM with a negative binomial
+link function) for each gene (glmFit) and then performs likelihood
+ratio tests for models w or w/o focal factor (glmLRT).
 
-........... include outcome of MDS analysis here?
+%% ........... include outcome of MDS analysis here?
+%% No, results never are presented in methods.
+
+\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
+A selection of differently abundant mRNAs are used for hierarchical
+clustering of \textit{E. falciformis} life cycle relevant genes. In
+each comparison (see Table 3), the union of the at most 500 genes
+differentially abundant with lowest FDR (<0.05) are selected. In the
+next step, the 500 mRNAs from each comparison (or less) are
+joined. For \textit{E. falciformis} time p.i. comparison, this
+resulted in 1618 unique genes selected as differently expressed . The
+22 genes in the NMRI vs C57BL/6 comparison are not included in the
+\textit{E. falciformis} life cycle analysis. The same comparisons for
+mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
+all samples, i.e. also samples which themselves did not have any
+significantly different mRNAs according to our selection, were
+included in hierarchical clustering. Scale bar in heatmaps show 0 as
+mean mRNA abundance for each gene (row). Up (green) and
+down-regulation (brown) denote number of standard deviations from 0,
+i.e., row mean. Hierarchical clustering was performed using with
+Euclidean distances, using complete linkage ('complete', R package
+base).
+
+%% the paragraph above has results. All those number
 
 \subsection{Differentially abundant mRNAs used for hierarchical clustering}
-A selection of differently abundant mRNAs are used for hierarchical clustering of \textit{E.
-falciformis} life cycle relevant genes.
-In each comparison (see Table 3), the union of the 500 differentially abundant mRNAs with lowest FDR are selected.
-In the next step, the 500 mRNAs from each comparison (or less) are joined.
-For \textit{E.
-falciformis} time p.i.
-comparison, this resulted in 4935 genes of which only the 1618 unique genes were selected.
-The 22 genes in the NMRI vs C57BL/6 comparison are not included in the \textit{E.
-falciformis} life cycle analysis.
-The same comparisons for mouse yielded 8052 genes in total and 1313 unique ones.
-In heatmaps, all samples, i.e.
-also samples which themselves did not have any significantly different mRNAs according to our selection, were included in hierarchical clustering.
-Scale bar in heatmaps show 0 as mean mRNA abundance for each gene (row).
-Up (green) and down-regulation (brown) denote number of standard deviations from 0, i.e., row mean.
-Hierarchical clustering was performed using with Euclidean distances, using complete linkage ('complete', R package base).
+All analyses were performed in R (cite R-core). Complete scripts are
+available at git@github.com:derele/Ef\_RNAseq.git tagged as version
+1.0. 
+%% We will make this available as a git webpage probably
 

### 2016-07-21 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
some work on manuscript

--- 2016-07-20 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
@@
 
+We found xx E. falciformis genes of the total predicted yy, which were expressed in at leat one developmental stage measured.
+Of the total reported mouse genes aaa, hhh were expressed on at least one timepoint during E. falciformis infection and fff were expressed in controls.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Experimental overview}
-We performed mRNA sequencing of caecum tissue of mice infected with
-the parasite \textit{E. falciformis} at different timepoints and for
-infections of different mouse strains.
-
-Replicate average of transcripts as order of magnitude for
-\textit{E. falciformis}/mouse are indicated in Table.... Columns in
-table represent mouse strains used in infection experiments. Rows
-represent timepoints post infection plus oocyst and sporozoite
-samples. The upper part of the table shows data for first infection,
-and oocyst and sporozoite data. The lower part shows data for
-challenge infection. Averages were calculated after sample exclusions
-(see Methods). For exact values, see Table 1.
+Replicate average of transcripts as order of magnitude for \textit{E.
+falciformis}/mouse are indicated in Table....
+Columns in table represent mouse strains used in infection experiments.
+Rows represent timepoints post infection plus oocyst and sporozoite samples.
+The upper part of the table shows data for first infection, and oocyst and sporozoite data.
+The lower part shows data for challenge infection.
+Averages were calculated after sample exclusions (see Methods).
+For exact values, see Table 1.
 
@@
 
-Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison."NMRI" followed by number
-indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
-p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
-implemented in edgeR are included. NAs are missing samples or not
-applicable for the species. NA* is due to missing NMRI day 0 sample
-from first infection. For \textit{E. falciformis} no mRNAs are
-significantly differently abundant between first and second infection,
-whereas in mouse there are some differences, especially on day 7
-p.i. However, upon hierarchical clustering of these genes (union of
-500 with lowest FDR), samples do not cluster according to first/second
-infection (Figure SX ?). In comparisons between mouse strains, 22
-genes are detected to be different in the parasite data. The
-difference is however not between Rag1-/- mice as expected, but
-between NMRI and C57BL/6. For mouse....
-
+Table 2 lists pairwise sample comparisons and number of genes with differently abundant mRNAs per comparison."NMRI" followed by number indicates day post infection (e.g.
+NMRI 3 = NMRI mouse on day 3 p.i.).
+Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are included.
+NAs are missing samples or not applicable for the species.
+NA* is due to missing NMRI day 0 sample from first infection.
+For \textit{E.
+falciformis} no mRNAs are significantly differently abundant between first and second infection, whereas in mouse there are some differences, especially on day 7 p.i.
+However, upon hierarchical clustering of these mouse genes (union of 500 with lowest FDR), samples do not cluster according to first/second infection (Figure SX ?).
+In comparisons between mouse strains, 22 \textit{E.
+falciformis} genes are detected as different.
+The difference is however not between Rag1-/- immune incompetent and either WT mouse as expected, but between NMRI and C57BL/6.
+For mouse....
+The comparisons between timepoints p.i.
+are dominant in number of mRNAs being differently abundant and we were interested in finding patterns in this data.
+Unsupervised hierarchical clustering was applied on \textit{E.
+falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR, Figure 1).
+Samples from day 7 p.i.
+form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection.
+For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down).
+Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2).
+Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down).
+mRNA profiles on days 3 and 5 p.i.
+from all three mouse strains cluster together.
+These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
+Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure}
-
-We applied hierarchical clustering on \textit{E.
-falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR).
-Samples from day 7 p.i.
-form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection.
-For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down).
-Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2).
-Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down).
-mRNA profiles on days 3 and 5 p.i.
-from all three mouse strains cluster together.
-These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
-Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.
 
@@
 \section*{Additional Files}
-  \subsection*{Additional file 1 --- Raw and normalized counts }
-  Raw and normalized counts for individual samples in our
-  study (two compressed csv files).
+  \subsection*{Additional file 1 --- Sample additional file title}
+    Additional file descriptions text (including details of how to
+    view the file, if it is in a non-standard format or the file extension).  This might
+    refer to a multi-page table or a figure.
 
-  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
-  Fold-changes, p-values and false discovery rates (adjusted p-values)
-  are given for all tested contrasts (one compressed csv file).
-
-  \subsection*{Additional file 3 --- Additional methods and resuts }
-  % this would be cool but requires special referencing in the text
-  % (hard to keep track and hard to fit in the journals referencing
-  % system). But we can try.
-  Document containing additonal figures and summary tables (pdf).
-
-\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
-  Tables listing all tested gene sets and resulting significatn GO
-  terms.
+  \subsection*{Additional file 2 --- Sample additional file title}
+    Additional file descriptions text.
 

### 2016-07-22 5_clustering.R
steps towards a slim representation of code matching the manuscript sections

@@ -18,6 +18,17 @@ union.of.Ef.cycle.diff.top.100 <-
 Ef.cycle.diff.top.100.data <-
     cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff.top.100,]
 
+union.of.Ef.cycle.diff <-
+    unique(unlist(Ef.1st.pass.model[[3]][1:10]))
+
+Ef.cycle.diff.top.100.data <-
+    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff.top.100,]
+
+Ef.cycle.diff.data <-
+    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff,]
+
+
+
 ############# 1st VERSUS 2nd INFECTION ##################
 union.of.Ef.1st2nd.top <-
     unique(unlist(lapply(Ef.1st.pass.model[[3]][11:15], head, n=500)))
@@ -40,7 +51,7 @@ get.scaled.and.clustered <- function(data){
     return(hclustered)
 }
 
-Ef.hclustered <- get.scaled.and.clustered(Ef.cycle.diff.top.100.data)
+Ef.hclustered <- get.scaled.and.clustered(Ef.cycle.diff.data)
 Ef.hclustered.first.second <- get.scaled.and.clustered(Ef.1st2nd.top.data) 
 Ef.hclustered.strain <- get.scaled.and.clustered(Ef.strain.depend.top.data )
 
@@ -94,7 +105,7 @@ dev.off()
 
 pdf("figuresANDmanuscript/EfLifecycleHeatmap.pdf",
     height = 8, width = 8, onefile = FALSE) # onefile command to hack away empty page in pdf
-pheatmap(Ef.cycle.diff.top.100.data,
+pheatmap(Ef.cycle.diff.data,
          color = brewer.pal(n = 11, name = "BrBG"), 
          scale = "row",
          cluster_rows = T, ## hc.high,

### 2016-07-22 A_data_curataion.R

@@ -58,11 +58,11 @@ RC.table$p.Ef.reads <- round(RC.table$c.Ef.reads/RC.table$c.total.reads*100, 4)
 ef.percentage <- ggplot(RC.table, aes(x = sample, y = p.Ef.reads,
                                       color=seq.method)) +
     geom_point()+
-        scale_y_log10()+
-            ylab(label="Percentage of reads mapping to Eimeria genome") +
-                xlab(label="Sample") +
-                ggtitle("Eimeria fraction of total sequences per sample") +
-                    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+    scale_y_log10()+
+    ylab(label="Percentage of reads mapping to Eimeria genome") +
+    xlab(label="Sample") +
+    ggtitle("Eimeria fraction of total sequences per sample") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 ggsave("ef.percentage.png", ef.percentage, path = "figures", width = 10)
 
@@ -93,12 +93,13 @@ names(keep.val) <- keep.val
 r.c.s.l <-
     lapply(keep.val,
            function (x){
-		
-		Mm.RCtemp <- All.RC[[3]][grepl('^ENSMUS.*', rownames(All.RC[[3]])),]+ 0.1 # add tiny number to be able to plot transcripts with zero reads mapping
-        	kept.df <- Mm.RCtemp[rowSums(Mm.RCtemp)>x,]       
-		kept.df <- melt(kept.df)
-		kept.df <- kept.df[grep("^.*_(oocysts|sporozoites)_.*$", kept.df$X2, invert = T),] # remove ooc and sporo samples
-		return(kept.df)
+               ## add tiny number to be able to plot transcripts with zero reads mapping
+               Mm.RCtemp <- All.RC[[3]][grepl('^ENSMUS.*', rownames(All.RC[[3]])),]+ 0.1 
+               kept.df <- Mm.RCtemp[rowSums(Mm.RCtemp)>x,]       
+               kept.df <- melt(kept.df)
+               ## remove ooc and sporo samples
+               kept.df <- kept.df[grep("^.*_(oocysts|sporozoites)_.*$", kept.df$X2, invert = T),] 
+               return(kept.df)
            })
 
 #####################################################
@@ -106,7 +107,7 @@ density.plots <-
     lapply(seq_along(r.c.s.l),
            function(i){
                ggplot(r.c.s.l[[i]],
[…]

### 2016-07-22 DEG_Ef.R

@@ -1,224 +1,0 @@
-## DIFFERENTIAL GENE EXPRESSION ANALYSIS ON EIMERIA SEQS #####
-
-#setwd("/home/totta/Ef_RNAseq")
-
-if(!exists("raw.counts.4.bg")){
-	source("2_edgeR_diff.R")
-        }
-
-library(scales)
-library(stringr)
-
-Ef.RC <- raw.counts.4.bg(Ef.bg)
-
-pdf("figures/rep_pairs_EfG.pdf", width = 70, height=70) # So far not working - because R interrupt?
-ggpairs(Ef.RC[[3]])
-dev.off()
-
-## use the alternate designE like .... the EdgR manual (or
-## eg. here: https://support.bioconductor.org/p/66952/) and specifiy
-## interactions as contrasts!
-#MOUSE: design <- model.matrix(~0+pData(Ef.bg)$grouped)
-
-<<<<<<< HEAD
-##################### For EIMERIA: #################################
-
-## Summary table of transcript counts per sample
-## MOUSE table
-tableM <- as.data.frame(colSums(mouse.RC[[3]]))
-#colnames(tableM)[,2] <- "Read.number"
-## MOUSE w/o day 0 
-tableM$Samples <- rownames(tableM)
-library(stringr)
-tableMsmall <- tableM[!(str_detect(rownames(tableM),"^.*0dpi.*$")), ]
-
-## EIMERIA table
-tableEf <- as.data.frame(colSums(Ef.RC[[3]]))
-#colnames(tableEf)[,1] <- "Read.number" - doesnt work?
-#tableEf$Samples <- colnames(Ef.RC[[3]])
-#print.xtable(tableEf, type = "latex", file = "tableEf-reads.tex")
-tableEf$Samples <- rownames(tableEf)
-## EIMERIA table w/o oocysts and sporozoites, for mouse comparison
-tableEfsmall <- tableEf[!(str_detect(tableEf$Samples, "^NMRI_(oocysts|sporozoites)$")), ]
-
-## ADD EIMERIA % to EIMERIAsmall table
-tableEfsmall$percent.Ef.reads <- (tableEfsmall[,1]/tableMsmall[,1])*100
-
-library(scales)
-
-=======
->>>>>>> Totta_working
-ggplot()+
-    geom_point(data = tableEf, aes(x = Samples,
-                   y = as.numeric(colSums(Ef.RC[[3]])))) +
-	scale_y_log10(labels = comma) +
[…]

### 2016-07-22 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
Manuscript getting into results section - numbers added

--- 2016-07-20 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
@@
 The number of sequencing reads obtained for individual replicates
-ranged between 111111 (samply NMRIxyz) and 222222 (NMRIwyv).
+ranged from 21,517,055 (sample NMRI\_sporozoites\_rep2) to 593,396,312
+(NMRI\_2ndInf\_5dpi\_rep1).
 
 As samples and idividual replicates were sequenced in batches to
-different depth and using different instrumentation (see additional
-file 1 %%{or include sequencing even in Table 1?
-) we performed quality controls confirming the abscence of batch
-effects influencing analysis and quality of our results (additional
-file xyc).
+different depth and using different instrumentation (Table 1) we
+performed quality controls confirming the abscence of batch effects
+influencing analysis and quality of our results (additional file xyc).
 
@@
 
-A maximum of 75\% (sample NMRIaaga) and a minimum of 0.01\% (sample
-NMRIagag) of mapped reads could be assigened to the
-\textit{E. falciformis} genome in samples considered infected. We
-excluded samples NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 due
-to an exptremely low parasite contribution to the overall
-transcriptome. Technically this made it possible to obtain read counts
-following negative bionmial distributions after normalization between
-samples (see adddtional file x). Judging from their mRNA output those
-samples can be considered non-infected.On the other end of the
-spectruem this means that at 7dpi the majority of mRNA originated from
-the parasite as opposed to the host transcriptome (Figure 1).
+A maximum of 86\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of
+0.0326\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be
+assigened to the \textit{E. falciformis} genome in samples considered
+infected.
+
+We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
+due to an exptremely low parasite contribution to the overall
+transcriptome. Technically this exclusion made it possible to obtain
+read counts in agreement with a negative bionmial distributions (see
+adddtional file x). Judging from their mRNA output those samples can
+be considered non-infected, as both samples occured in a challenge
+infection we assume that the infection was cleard or reduced to a
+non-detectable output in those animmals.
+
+On the other end of the spectrum mice not previously immunized
+featured an mRNA ``holo-transcriptome'' at 7dpi with the majority of
+mRNA originating from the parasite as opposed to the host (Figure 1).
 
@@
 
-We selected mouse
+We selected all mouse genes differently expressed (FDR<1\%) between
+uinfected and any of the infected samples.
+
 %%% This is from methods but all the results should be presented here:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% I like this table. Looks good and transprorts the right kind of
+%% information
+
 \clearpage
@@
 \textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
-	\textit{comparisons} 	    & (FDR$\leq$0.01) &  (FDR$\leq$0.01/0.05) \\ \midrule
+	\textit{comparisons} 	    & (FDR$\leq$1\%) &  (FDR$\leq$1\%/5\%) \\ \midrule
 	NMRI 0 vs NMRI 3		& NA   & 274 \\

### 2016-07-29 1_ballgown_import.R
scripsts 1-3 don't source but write/read supplementary files

@@ -102,3 +102,25 @@ raw.counts.4.bg <- function(bg){
 All.RC <- raw.counts.4.bg(All.bg)
 Mm.RC <- raw.counts.4.bg(Mm.bg)
 Ef.RC <- raw.counts.4.bg(Ef.bg)
+
+## Create tables and analyse for sample exclusion: A_data_curation.R
+## Now exclude those samples
+Mm.bg <- subset(Mm.bg,
+                !pData(Mm.bg)$samples%in%c("NMRI_2ndInf_3dpi_rep1",
+                                           "NMRI_2ndInf_5dpi_rep2",
+                                           "NMRI_1stInf_0dpi_rep1"),
+                genomesubset=FALSE)
+Mm.RC <- raw.counts.4.bg(Mm.bg)
+
+Ef.bg <- subset(Ef.bg,
+                !pData(Ef.bg)$samples%in%c("NMRI_2ndInf_3dpi_rep1",
+                                           "NMRI_2ndInf_5dpi_rep2",
+                                           "NMRI_1stInf_0dpi_rep1"),
+                genomesubset=FALSE)
+Ef.RC <- raw.counts.4.bg(Ef.bg)
+
+write.table(Mm.RC[[3]], "output_data/RC_Mm_genes.csv", sep=",")
+write.table(Ef.RC[[3]], "output_data/RC_Ef_genes.csv", sep=",")
+
+write.table(pData(Ef.bg), "output_data/Ef_sample_pData.csv", sep=",")
+write.table(pData(Mm.bg), "output_data/Mm_sample_pData.csv", sep=",")

### 2016-07-29 2_edgeR_diff.R

@@ -1,39 +1,23 @@
 #### DIFFERENTIAL GENE EXPRESSION ANALYSIS
-
 ## needed from data import script are two raw count objects 
+Mm.RC <- as.matrix(read.table("output_data/RC_Mm_genes.csv", sep=","))
+Ef.RC <- as.matrix(read.table("output_data/RC_Ef_genes.csv", sep=","))
 
-#source("https://bioconductor.org/biocLite.R")
-if(!exists("Mm.RC")|!exists("Ef.RC")){
-    source("1_ballgown_import.R")
-}
-
+Ef.pData <- read.table("output_data/Ef_sample_pData.csv", sep=",")
+Mm.pData <- read.table("output_data/Mm_sample_pData.csv", sep=",")
 
 library(statmod)
 library(edgeR)
 library(GGally)
 library(RUVSeq)
 library(reshape)
-
-## Implement the sample exclusion (see A_data_curation.R)
-Mm.bg <- subset(Mm.bg,
-                !pData(Mm.bg)$samples%in%c("NMRI_2ndInf_3dpi_rep1",
-                                           "NMRI_2ndInf_5dpi_rep2",
-                                           "NMRI_1stInf_0dpi_rep1"),
-                genomesubset=FALSE)
-Mm.RC <- raw.counts.4.bg(Mm.bg)
-
-Ef.bg <- subset(Ef.bg,
-                !pData(Ef.bg)$samples%in%c("NMRI_2ndInf_3dpi_rep1",
-                                           "NMRI_2ndInf_5dpi_rep2",
-                                           "NMRI_1stInf_0dpi_rep1"),
-                genomesubset=FALSE)
-Ef.RC <- raw.counts.4.bg(Ef.bg)
+library(plyr)
 
 ## use the alternate design like recommended by the EdgR manual (or
 ## eg. here: https://support.bioconductor.org/p/66952/) and specifiy
 ## interactions as contrasts!
-Ef.design <- model.matrix(~0+pData(Ef.bg)$grouped)
-colnames(Ef.design)  <- gsub("pData\\(Ef.bg\\)\\$grouped", "",
+Ef.design <- model.matrix(~0+Ef.pData$grouped)
+colnames(Ef.design)  <- gsub("Ef.pData\\$grouped", "",
                              colnames(Ef.design))
 ## just to have the standard point delimitor
 colnames(Ef.design)  <- gsub("_", ".",
@@ -74,9 +58,9 @@ Ef.contrasts <- makeContrasts(
[…]

### 2016-07-29 3_annotations.R

@@ -2,9 +2,9 @@
 ## a TopTags (edgeR) object listing genes tested for differential
 ## expression
 
-if(!exists("Ef.1st.pass.model")){ #Totta changes Mm to Ef here
-    source("2_edgeR_diff.R")
-}
+Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
+Ef.DE.test <- read.table("output_data/Ef_DEtest.csv", sep=",")
+
 
 ## The Bioconductor mouse annotation libary
 library(Mus.musculus)
@@ -32,14 +32,14 @@ gene2GO[["Mm"]] <- by(annot.frame, annot.frame$ensembl_id,
                       function(x) as.character(x$go_id))
 ## define the universe of tested genes
 gene2GO[["Mm"]] <- gene2GO[["Mm"]][names(gene2GO[["Mm"]])%in%
-                                       rownames(Mm.1st.pass.model[[1]])]
+                                   unique(Mm.DE.test$gene)]
 
 
 go_Ef <- read.csv("data/12864_2014_6777_GO-annot.csv")
 
 gene2GO[["Ef"]] <- by(go_Ef, go_Ef$gene, function(x) as.character(x$go))
 gene2GO[["Ef"]] <- gene2GO[["Ef"]][names(gene2GO[["Ef"]])%in%
-                                       rownames(Ef.1st.pass.model[[1]])]
+                                   unique(Ef.DE.test$gene)]
 
 ### ORTHOMCL categories
 

### 2016-07-29 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
+1 merge of chnges in MS, removed ugly table 1...

@@
   \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
-  Fold-changes, p-values and false discovery rates (adjusted p-values)
-  are given for all tested contrasts (one compressed csv file).
+  Focal contrast, fold-changes, likelihood ratio in/excluding this
+  difference in models, p-values , and false discovery rates (adjusted
+  p-values) are given for all tested contrasts (one compressed csv
+  file).
 

### 2016-08-07 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
Before moving to feature counts and implementing the normalization correctly

@@
 \usepackage{graphicx}
+
+% to color table rows
+\usepackage{color, colortbl}
 \graphicspath{ {../} }
@@
 
-The number of sequencing reads obtained for individual replicates
-ranged from 21,517,055 (sample NMRI\_sporozoites\_rep2) to 593,396,312
-(NMRI\_2ndInf\_5dpi\_rep1).
+The number of sequence read mappings obtained for individual
+replicates ranged from 21,517,055 (sample NMRI\_sporozoites\_rep2) to
+593,396,312 (NMRI\_2ndInf\_5dpi\_rep1).
 
@@
 mRNA originating from the parasite as opposed to the host (Figure 1).
+
+\newcommand{\bcell}[2][c]{%
+  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+
+\definecolor{LightCyan}{rgb}{0.88,1,1}
+\definecolor{LightRed}{rgb}{1,0.88,1}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Fri Aug  5 13:00:01 2016
+\begin{table}[ht]
+\caption{Experimental design, sequencing output and mapping statistics.}
+\centering
+\hspace*{-1.5cm}\begin{tabular}{lllllllll}
+  \hline
+Sample & \bcell{Sequencing\\method} & batch & \bcell{read mappings\\Mouse} & \bcell{read mappings\\\textit{E. falciformis}} & \bcell{Percentage\\\textit{E. falciformis}} & dpi & infection  \\ 
+\hline
+  NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 285032128 & 326 & 0 & 0dpi & Challenge \\ 
+  Rag\_0dpi\_rep1 & hiseq & 3 & 43414004 & 474 & 0.001 & 0dpi & None \\ 
+  C57BL6\_0dpi\_rep1 & hiseq & 3 & 53877840 & 491 & 0.001 & 0dpi & None \\ 
+  C57BL6\_0dpi\_rep2 & hiseq & 3 & 76753491 & 657 & 0.001 & 0dpi & None \\ 
+  Rag\_0dpi\_rep2 & hiseq & 3 & 80702547 & 730 & 0.001 & 0dpi & None \\ 
+  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 90681785 & 3738 & 0.004 & 0dpi & Challenge \\ 
+  \rowcolor{LightCyan}
+  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 91352242 & 5355 & 0.006 & 3dpi & Challenge \\ 
+  \rowcolor{LightCyan}
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 113957667 & 12692 & 0.011 & 5dpi & Challenge \\ 
+  \rowcolor{LightRed}
+  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 229165002 & 31459 & 0.014 & 0dpi & None \\ 
+  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 106446839 & 34699 & 0.033 & 3dpi & Challenge \\ 
+  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 122999109 & 46700 & 0.038 & 7dpi & Challenge \\ 
+  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 71173382 & 73445 & 0.103 & 5dpi & First \\ 
+  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 57192380 & 113673 & 0.198 & 5dpi & First \\ 
+  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 65523435 & 217606 & 0.331 & 5dpi & First \\ 
+  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 85288323 & 224273 & 0.262 & 5dpi & Challenge \\ 
+  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 32887009 & 250954 & 0.757 & 5dpi & First \\ 
+  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 62053975 & 311900 & 0.500 & 5dpi & Challenge \\ 
+  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 209723287 & 815820 & 0.388 & 3dpi & First \\ 
+  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 515516785 & 903330 & 0.175 & 3dpi & First \\ 
+  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 316721928 & 1609009 & 0.505 & 5dpi & Firs \\ 
+  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 589643389 & 3752923 & 0.632 & 5dpi & Challenge \\ 
+  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 97221189 & 9927803 & 9.265 & 7dpi & Challenge \\ 
+  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 204647381 & 18549727 & 8.311 & 5dpi & First \\ 
+  NMRI\_sporozoites\_rep2 & GAII & 0 & 8702 & 21508353 & 99.960 & \textit{in vitro} & \textit{in vitro} \\ 
+  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 334671421 & 54022504 & 13.899 & 5dpi & First \\ 
+  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 12532238 & 79648900 & 86.405 & 7dpi & First \\ 
+  NMRI\_sporozoites\_rep1 & GAII & 1 & 13800 & 92259539 & 99.985 & \textit{in vitro} & \textit{in vitro} \\ 
+  NMRI\_oocysts\_rep1 & GAII & 1 & 11676 & 108477484 & 99.989 & envir. & envir. \\ 
+  NMRI\_oocysts\_rep2 & GAII & 0 & 19024 & 126543533 & 99.985 & envir. & envir. \\ 
+  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 94310278 & 154343046 & 62.072 & 7dpi & First \\ 
+   \hline
+\hspace*{+1.5cm}\end{tabular}
+\begin{tablenotes}
+  \small
+\item *envir. ``environmental'' oocysts sporulated in potassium dichromate. \\
+\item *\textit{in vitro} in vitro sporulated oocysts
+\end{tablenotes}
+\end{table}
 

### 2016-08-07 figuresANDmanuscript/table_rc.tex
Before moving to feature counts and implementing the normalization correctly

@@
-% latex table generated in R 3.3.0 by xtable 1.8-2 package
-% Wed Jun 22 17:40:30 2016
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Fri Aug  5 13:00:01 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}{lllllllll}
   \hline
-Sample & Mouse transcripts & E. falciformis transcripts & Percentage E. falciformis & \#E. falciformis genes \\ 
+Sample & Sequencing method & batch & read mappings Mouse & read mappings E. falciformis & Percentage E. falciformis & dpi & challenged & \# E. falciformis genes \\ 
   \hline
-NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
-  NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
-  NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
-  NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep1 & 589643389.000 & 3752923.000 & 0.632 & 5602.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
-  C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep1 & 209723287.000 & 815820.000 & 0.388 & 5466.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
-  Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
-  Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
-  Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
-  NMRI\_1stInf\_0dpi\_rep1 & 229165002.000 & 31459.000 & 0.014 & 1380.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
-  Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
-  Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
+NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 285032128.000 & 326.000 & 0.000 & 0dpi & Callenge & 2.000 \\ 
+  Rag\_0dpi\_rep1 & hiseq & 3 & 43414004.000 & 474.000 & 0.001 & 0dpi & None & 2.000 \\ 
+  C57BL6\_0dpi\_rep1 & hiseq & 3 & 53877840.000 & 491.000 & 0.001 & 0dpi & None & 2.000 \\ 
+  C57BL6\_0dpi\_rep2 & hiseq & 3 & 76753491.000 & 657.000 & 0.001 & 0dpi & None & 2.000 \\ 
+  Rag\_0dpi\_rep2 & hiseq & 3 & 80702547.000 & 730.000 & 0.001 & 0dpi & None & 3.000 \\ 
+  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 90681785.000 & 3738.000 & 0.004 & 0dpi & Callenge & 62.000 \\ 
+  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 91352242.000 & 5355.000 & 0.006 & 3dpi & Callenge & 121.000 \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 113957667.000 & 12692.000 & 0.011 & 5dpi & Callenge & 539.000 \\ 
+  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 229165002.000 & 31459.000 & 0.014 & 0dpi & None & 1380.000 \\ 
+  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 106446839.000 & 34699.000 & 0.033 & 3dpi & Callenge & 1901.000 \\ 
+  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 122999109.000 & 46700.000 & 0.038 & 7dpi & Callenge & 2174.000 \\ 
+  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 71173382.000 & 73445.000 & 0.103 & 5dpi & First & 2748.000 \\ 
+  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 57192380.000 & 113673.000 & 0.198 & 5dpi & First & 2969.000 \\ 
+  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 65523435.000 & 217606.000 & 0.331 & 5dpi & First & 4259.000 \\ 
+  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 85288323.000 & 224273.000 & 0.262 & 5dpi & Callenge & 4251.000 \\ 
+  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 32887009.000 & 250954.000 & 0.757 & 5dpi & First & 3946.000 \\ 
+  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 62053975.000 & 311900.000 & 0.500 & 5dpi & Callenge & 4610.000 \\ 
+  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 209723287.000 & 815820.000 & 0.388 & 3dpi & First & 5466.000 \\ 
+  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 515516785.000 & 903330.000 & 0.175 & 3dpi & First & 5101.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 316721928.000 & 1609009.000 & 0.505 & 5dpi & First & 5439.000 \\ 
+  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 589643389.000 & 3752923.000 & 0.632 & 5dpi & Callenge & 5602.000 \\ 
+  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 97221189.000 & 9927803.000 & 9.265 & 7dpi & Callenge & 5865.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 204647381.000 & 18549727.000 & 8.311 & 5dpi & First & 5739.000 \\ 
+  NMRI\_sporozoites\_rep2 & GAII & 0 & 8702.000 & 21508353.000 & 99.960 & environmental & environmental & 5564.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 334671421.000 & 54022504.000 & 13.899 & 5dpi & First & 5794.000 \\ 
+  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 12532238.000 & 79648900.000 & 86.405 & 7dpi & First & 5894.000 \\ 
+  NMRI\_sporozoites\_rep1 & GAII & 1 & 13800.000 & 92259539.000 & 99.985 & environmental & environmental & 5808.000 \\ 
+  NMRI\_oocysts\_rep1 & GAII & 1 & 11676.000 & 108477484.000 & 99.989 & environmental & environmental & 5734.000 \\ 
+  NMRI\_oocysts\_rep2 & GAII & 0 & 19024.000 & 126543533.000 & 99.985 & environmental & environmental & 5774.000 \\ 
+  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 94310278.000 & 154343046.000 & 62.072 & 7dpi & First & 5897.000 \\ 
    \hline

### 2016-08-07 1_ballgown_import.R

@@ -28,7 +28,8 @@ add.pdata <- function (ballgown.obj){
     ## variables to summarize other variables sith mutliple levels into two levels ##
     immune.status <- ifelse(mouse.strain %in% "Rag", "Rag", "competent")
     late.early <- ifelse(timepoint < 7 , "early", "late")
-    batch <- c(3, 3, 3, 3, 3, 1, 1, 100, 2, 2, 0, 1, 1, 2, 3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
+    batch <- c(3, 3, 3, 3, 3, 1, 1, 2, 2, 2, 0, 1, 1, 2,
+               3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
     seq.method <- c("hiseq", "hiseq", "hiseq", "hiseq", "hiseq",
                "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
                "hiseq", "hiseq", "hiseq",
@@ -66,7 +67,7 @@ Ef.bg <- subset(Ef.bg,
 ## the ballgown objects
 
 raw.counts.4.bg <- function(bg){
-    exon.raw.count <- eexpr(bg, "ucount")
+    exon.raw.count <- eexpr(bg, "mrcount") ## ucount
     ## the linkage data
     e2t <- bg@indexes$e2t
     t2g <- bg@indexes$t2g
@@ -121,6 +122,9 @@ Ef.RC <- raw.counts.4.bg(Ef.bg)
 
 write.table(Mm.RC[[3]], "output_data/RC_Mm_genes.csv", sep=",")
 write.table(Ef.RC[[3]], "output_data/RC_Ef_genes.csv", sep=",")
+write.table(All.RC[[3]], "output_data/RC_All_genes.csv", sep=",")
+
 
 write.table(pData(Ef.bg), "output_data/Ef_sample_pData.csv", sep=",")
 write.table(pData(Mm.bg), "output_data/Mm_sample_pData.csv", sep=",")
+write.table(pData(All.bg), "output_data/Sample_pData.csv", sep=",")

### 2016-08-07 5_clustering.R

@@ -4,44 +4,51 @@ library(pheatmap)
 library(made4)
 library(RColorBrewer)
 
-if(!exists("Ef.1st.pass.model")){
-    source("2_edgeR_diff.R")
+## raw count data
+Mm.RC <- read.table("output_data/RC_Mm_genes.csv", sep=",")
+Ef.RC <- read.table("output_data/RC_Ef_genes.csv", sep=",")
+
+## replace zeros and avoid zero variance in cluter scaling
+Mm.RC[Mm.RC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Mm.RC[Mm.RC==0]), replace =TRUE)
+Ef.RC[Ef.RC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Ef.RC[Ef.RC==0]), replace =TRUE)
+
+## data on statistical testing for selection
+Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
+Ef.DE.test <- read.table("output_data/Ef_DEtest.csv", sep=",")
+
+select.diff.data <- function (data, diff.data, columns, FDR.min=0.01){
+    DE.test.col <-
+        diff.data[diff.data$contrast%in%columns, ]
+    DE.diff.genes <-
+        unique(DE.test.col[DE.test.col$FDR<FDR.min, "gene"])
+    data[DE.diff.genes, ]
 }
-########## USE THIS #######################
-########## PHEATMAP() specifying distance and method ############
-## 1:10 refers to the Ef.contrast object and the first ten 
-## comparisons in it.
-############# LIFE CYCLE SELECTION ##################
-union.of.Ef.cycle.diff.top.100 <-
-    unique(unlist(lapply(Ef.1st.pass.model[[3]][1:10], head, n=500)))
 
-Ef.cycle.diff.top.100.data <-
-    cpm(Ef.1st.pass.model[[4]])[union.of.Ef.cycle.diff.top.100,]
+############# Eimeria LIFE CYCLE SELECTION ##################
+cycle.cols <- c("N3vsN5","N3vsN7", "N3vsOoc", "N3vsSpo",
+                "N5vsN7", "N5vsOo", "N5vsSp",
+                "N7vsOo", "N7vsSp",
+                "SpvsOo")
+Ef.cycle.diff.data <- select.diff.data(Ef.RC, Ef.DE.test, cycle.cols, 0.01)
 
-union.of.Ef.cycle.diff <-
[…]

### 2016-08-07 A_data_curataion.R

@@ -1,6 +1,5 @@
-## This is a script which is not part of the pipeline but needed for
-## evaluation purpose -> only "insight" gained from this is transfered
-## to the numberd scripts which are part of the processing pipeline
+## This is a script is 2nd the pipeline and needed for evaluation
+## purpose
 
 ## Goals:
 
@@ -29,46 +28,86 @@
 ## a) decide for samples needed to be excluede because of uncertain
 ## infection status
 
-##object required from upstream scripts:
+## The raw counts  required from upstream scripts:
+## GENES
+## All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
 
-## The raw counts:
-if(!exists("All.RC")|!exists("Mm.RC")|!exists("Ef.RC")){
-    source("1_ballgown_import.R")
-}
+## TRANSCRIPTS
+All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
+
+## The phenotype information
+All.pData <- read.table("output_data/Sample_pData.csv", sep=",")
 
 library(ggplot2)
 library(reshape)
 library(gridExtra)
+library(xtable)
 
 ## Overview for differential representation of host/parasite
-RC.table <-  as.data.frame(cbind(sample=as.character(pData(All.bg)$samples),
-                                 seq.method=as.character(pData(All.bg)$seq.method),
-                                 batch=pData(All.bg)$batch))
+RC.table <-  as.data.frame(cbind(sample=as.character(All.pData$samples),
+                                 seq.method=as.character(All.pData$seq.method),
+                                 batch=All.pData$batch))
 
-RC.table$c.total.reads <- colSums(All.RC[[3]])
+## get the total read count.
+## see README_fastq_linecount.txt on how this file was produced    
+## rawfastqRC <- readLines("/SAN/Eimeria_Totta/fastq_linecounts.txt")
 
-RC.table$c.Mm.reads <- colSums(All.RC[[3]][grepl('^ENSMUS.*', rownames(All.RC[[3]])),])
+## fastqRC <- rawfastqRC[seq(1, length(rawfastqRC), by=2)+1]
+## fastqRC <- as.numeric(as.character(fastqRC))/4
 
-RC.table$c.Ef.reads <- colSums(All.RC[[3]]
-                                [grepl('^EfaB.*', rownames(All.RC[[3]])),])
[…]

### 2016-08-08 1_ballgown_import.R
Replaced ballgown by featurecounts. Working up to annotation...

@@ -1,130 +1,0 @@
-### ballgown for data import from the cufflinks/tablemaker pipeline
-
-library(ballgown)
-setwd("~/Ef_RNAseq")
-
-## IMPORT DATA USING BALLGOWN IMPORT
-if(!exists("All.bg")) {
-    All.bg = ballgown(dataDir="/SAN/Eimeria_Totta/tablemaker_d_tophat_March_c/",
-        samplePattern = "*")
-}
-
-## some fundtions for RNAseq data, mainly a new subsetting function
-## for ballgown-objects
-source("functionsEfRNAseq.R")
-
-########################
-
-## the experimental meta-data to be stored in the pData slot of the object
-add.pdata <- function (ballgown.obj){
-    samples <- sub("^(.*)(_hiseq.fastq|_forw.fastq_paired)$",
-                   "\\1", sampleNames(ballgown.obj))
-    grouped <- sub("_rep\\d+$" , "", samples)
-    infection <- ifelse(grepl("1stInf", grouped), "1st",
-                        ifelse(grepl("2ndInf", grouped), "2nd", "NULLth"))
-    rep <- sub(".*_(rep\\d+)$" , "\\1", samples) # (name here) is taken "\\1" here
-    mouse.strain <- sub("^(.*?)_.*", "\\1", grouped)
-    timepoint <- as.numeric(as.character(sub("^.*_(.*)dpi", "\\1", grouped)))
-    ## variables to summarize other variables sith mutliple levels into two levels ##
-    immune.status <- ifelse(mouse.strain %in% "Rag", "Rag", "competent")
-    late.early <- ifelse(timepoint < 7 , "early", "late")
-    batch <- c(3, 3, 3, 3, 3, 1, 1, 2, 2, 2, 0, 1, 1, 2,
-               3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
-    seq.method <- c("hiseq", "hiseq", "hiseq", "hiseq", "hiseq",
-               "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
-               "hiseq", "hiseq", "hiseq",
-               "GAII",
-               "hiseq", "hiseq", "hiseq",
-               "GAII", "GAII", "GAII", "GAII",
-               "hiseq", "hiseq", "hiseq", "hiseq", "hiseq")
-    sample.seq <- paste(samples,seq.method)
-    data.frame(samples, grouped, infection, rep, mouse.strain,
[…]

### 2016-08-08 1_feature_count.R

@@ -1,0 +1,53 @@
+library(Rsubread)
+library(rtracklayer)
+library(Rsamtools)
+library(GenomicRanges)
+library(GenomicAlignments)
+
+dirs <- list.dirs("/SAN/Eimeria_Totta/tophat_March_c", recursive=FALSE)
+
+bams <- sapply(dirs, list.files, "accepted_hits.bam", full.names=TRUE)
+
+FC <- featureCounts(bams,
+                    annot.ext="/SAN/Eimeria_Totta/reference_genomes/mm10_eimeria_merge/indexes_bowtie2/mm10_GRCm38_eimeriaHaberkorn_fixed.gtf",
+                    isGTFAnnotationFile=TRUE,
+                    isPairedEnd=TRUE,
+                    nthreads=20)
+
+## the experimental meta-data as in the pData slot of all ballgown object
+create.pdata <- function (featurecountsTarget){
+    samples <- gsub(".*?tophat_March_c\\.(.*?_rep\\d).*", "\\1", 
+                   featurecountsTarget)
+    grouped <- sub("_rep\\d+$" , "", samples)
+    infection <- ifelse(grepl("1stInf", grouped), "1st",
+                        ifelse(grepl("2ndInf", grouped), "2nd", "NULLth"))
+    rep <- sub(".*_(rep\\d+)$" , "\\1", samples) # (name here) is taken "\\1" here
+    mouse.strain <- sub("^(.*?)_.*", "\\1", grouped)
+    timepoint <- as.numeric(as.character(sub("^.*_(.*)dpi", "\\1", grouped)))
+    ## variables to summarize other variables sith mutliple levels into two levels ##
+    immune.status <- ifelse(mouse.strain %in% "Rag", "Rag", "competent")
+    late.early <- ifelse(timepoint < 7 , "early", "late")
+    batch <- c(3, 3, 3, 3, 3, 1, 1, 2, 2, 2, 0, 1, 1, 2,
+               3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
+    seq.method <- c("hiseq", "hiseq", "hiseq", "hiseq", "hiseq",
+               "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
+               "hiseq", "hiseq", "hiseq",
+               "GAII",
+               "hiseq", "hiseq", "hiseq",
[…]

### 2016-08-08 2_data_curation.R

@@ -1,0 +1,226 @@
+## This is a script is 2nd the pipeline and needed for evaluation
+## purpose
+
+## Goals:
+
+## EX- on IN-clusion of samples and genes ###################
+
+## a) decide for samples needed to be excluede because of uncertain
+## infection status
+
+## b) decice for a filter on minimal expression values over all
+## samples to exclude genes
+
+## c) decide on samples needed to be excluded because of bad
+## "unnormalizable" distribution of expression values
+
+
+### Normalization
+
+## d) decide on a general normalization strategy scaling the smaple
+## throughput
+
+## e) decide on a "removeal of unwanted variation" strategy
+
+## f) decide on a potential addional "batch effect" removal strategy
+
+
+## a) decide for samples needed to be excluede because of uncertain
+## infection status
+
+## The raw counts  required from upstream scripts:
+## GENES
+## All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
+
+## TRANSCRIPTS
+All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
+
+## The phenotype information
+All.pData <- read.table("output_data/Sample_pData.csv", sep=",")
+
+library(ggplot2)
+library(reshape)
+library(gridExtra)
+library(xtable)
+
+## Overview for differential representation of host/parasite
+RC.table <-  as.data.frame(cbind(sample=as.character(All.pData$samples),
+                                 seq.method=as.character(All.pData$seq.method),
+                                 batch=All.pData$batch))
+
+## get the total read count.
+## see README_fastq_linecount.txt on how this file was produced    
+rawfastqRC <- readLines("/SAN/Eimeria_Totta/fastq_linecounts.txt")
+
+fastqRC <- rawfastqRC[seq(1, length(rawfastqRC), by=2)+1]
+fastqRC <- as.numeric(as.character(fastqRC))/4
+
+names(fastqRC) <- rawfastqRC[seq(1, length(rawfastqRC), by=2)]
+names(fastqRC) <- gsub("^.*?\\/(.*rep\\d).*", "\\1", names(fastqRC) )
+
+RC.table <- merge(RC.table, fastqRC, by.x="sample", by.y=0)
+names(RC.table)[ncol(RC.table)] <- "c.seq.reads"
+
+## this is total reads mapped not total reads!!!
+RC.table$c.mapping.counts <- colSums(All.RC)
+
[…]

### 2016-08-08 2_edgeR_diff.R

@@ -1,199 +1,0 @@
-#### DIFFERENTIAL GENE EXPRESSION ANALYSIS
-## needed from data import script are two raw count objects 
-Mm.RC <- as.matrix(read.table("output_data/RC_Mm_genes.csv", sep=","))
-Ef.RC <- as.matrix(read.table("output_data/RC_Ef_genes.csv", sep=","))
-
-Ef.pData <- read.table("output_data/Ef_sample_pData.csv", sep=",")
-Mm.pData <- read.table("output_data/Mm_sample_pData.csv", sep=",")
-
-library(statmod)
-library(edgeR)
-library(GGally)
-library(RUVSeq)
-library(reshape)
-library(plyr)
-
-## use the alternate design like recommended by the EdgR manual (or
-## eg. here: https://support.bioconductor.org/p/66952/) and specifiy
-## interactions as contrasts!
-Ef.design <- model.matrix(~0+Ef.pData$grouped)
-colnames(Ef.design)  <- gsub("Ef.pData\\$grouped", "",
-                             colnames(Ef.design))
-## just to have the standard point delimitor
-colnames(Ef.design)  <- gsub("_", ".",
-                             colnames(Ef.design))
-
-Ef.contrasts <- makeContrasts(
-    ## just different stages in NMRI
-    N3vsN5 = NMRI.1stInf.3dpi-NMRI.1stInf.5dpi, #1
-    N3vsN7 = NMRI.1stInf.3dpi-NMRI.1stInf.7dpi,
-    N3vsSpo = NMRI.1stInf.3dpi-NMRI.sporozoites,
-    N3vsOoc = NMRI.1stInf.3dpi-NMRI.oocysts,
-    N5vsN7 = NMRI.1stInf.5dpi-NMRI.1stInf.7dpi,
-    N5vsSp = NMRI.1stInf.5dpi-NMRI.sporozoites,
-    N5vsOo = NMRI.1stInf.5dpi-NMRI.oocysts,
-    N7vsSp = NMRI.1stInf.7dpi-NMRI.sporozoites,
-    N7vsOo = NMRI.1stInf.7dpi-NMRI.oocysts,
-    SpvsOo=  NMRI.sporozoites-NMRI.oocysts,	#10
-    ## differences in 1st vs 2nd infection
-    N3.1stvsN3.2nd = NMRI.1stInf.3dpi-NMRI.2ndInf.3dpi,	#11
-    N5.1stvsN5.2nd = NMRI.1stInf.5dpi-NMRI.2ndInf.5dpi,
[…]

### 2016-08-08 3_annotations.R

@@ -1,149 +1,0 @@
-## needed from previous script 2_edgeR_diff.R:
-## a TopTags (edgeR) object listing genes tested for differential
-## expression
-
-## The Bioconductor mouse annotation libary
-library(Mus.musculus)
-
-## merge all this information in a data frame to combine mouse
-## annotations with gene_id from Cufflinks get the information from
-## this annotation dbi into data frames
-.get.annot.frame <- function(){
-    ens2geneID <- as.data.frame(org.Mm.egENSEMBL)
-    name2SY <- as.data.frame(org.Mm.egSYMBOL2EG) # get 5 digit "symbol" to "short gene name" mapping
-    SY2Lname<- as.data.frame(org.Mm.egGENENAME) # get 5 digit "symbol" to "long name name" mapping
-    SY2GO <- as.data.frame(org.Mm.egGO) # get 5 digit "symbol" to gene ontology mapping
-    annot.frame <- merge(name2SY, ens2geneID, by = "gene_id")
-    annot.frame <- merge(annot.frame, SY2GO, by = "gene_id")
-    annot.frame <- merge(annot.frame, SY2Lname, by = "gene_id")
-    return(annot.frame)
-}
-
-annot.frame <- .get.annot.frame()
-
-### ORTHOMCL categories
-
-
-read.omcl <- function(path="data/Omcl_groups.txt"){
-  omcl <- readLines(path)
-  omcl <- lapply(omcl, strsplit, " ")
-  omcl <- lapply(omcl, function(x) x[[1]])
-  names(omcl) <- lapply(omcl, function(x) gsub("\\:", "", x[1]))
-  omcl <- lapply(omcl, function(x) x[-1])
-
-  ## make gene-ID identifiers used in the omcl clsters
-  omcl <- lapply(omcl, function (x)
-                 gsub("(Efa|)NODE_(.*_\\d+.g\\d+).t\\d+","\\1EfaB_\\2",  x))
-
-  omcl <- lapply(omcl, function (x) x[!duplicated(x)])
-  omcl <- omcl[!unlist(lapply(omcl, length))==1]
-  omcl <- omcl[!names(omcl)%in%"api10278"] ## exclude a cluster... wonder now why that ;-)
-
-}
-
-omcl <- read.omcl()
-
-get.omcl.categories <- function(ortho.obj){
-  ## Orthomcl
-  get.presence.omcl <- function(ortho.obj){
[…]

### 2016-08-08 3_edgeR_diff.R

@@ -1,0 +1,226 @@
+#### DIFFERENTIAL GENE EXPRESSION ANALYSIS
+## needed from data import script are two raw count objects 
+
+## We know from 2_data_curation.R that we want to exclude:
+excluded.samples <- c("NMRI_2ndInf_3dpi_rep1",
+                      "NMRI_2ndInf_5dpi_rep2",
+                      "NMRI_1stInf_0dpi_rep1")
+
+All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
+
+All.RC <- All.RC[, !colnames(All.RC)%in%excluded.samples]
+
+## Mouse and Eimeria specific
+Mm.RC <- All.RC[grepl('^ENSMUS.*', rownames(All.RC)),]
+Ef.RC <- All.RC[grepl('^EfaB.*', rownames(All.RC)),]
+
+pData <- read.table("output_data/Sample_pData.csv", sep=",")
+pData <- pData[!pData$samples%in%excluded.samples, ]
+
+## Mouse and Eimeria specific
+Ef.pData <- pData[!pData$timepoint%in%0, ]
+Mm.pData <- pData[!is.na(pData$timepoint), ]
+
+Ef.pData$grouped <- as.factor(as.character(Ef.pData$grouped))
+Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
+
+## Mouse and Eimeri spcific samples
+Mm.RC <- Mm.RC[, colnames(Mm.RC)%in%Mm.pData$samples]
+Ef.RC <- Ef.RC[, colnames(Ef.RC)%in%Ef.pData$samples] 
+
+library(statmod)
+library(edgeR)
+library(GGally)
+library(RUVSeq)
+library(reshape)
+library(plyr)
+
+## use the alternate design like recommended by the EdgR manual (or
+## eg. here: https://support.bioconductor.org/p/66952/) and specifiy
+## interactions as contrasts!
+Ef.design <- model.matrix(~0+Ef.pData$grouped)
+colnames(Ef.design)  <- gsub("Ef.pData\\$grouped", "",
+                             colnames(Ef.design))
+## just to have the standard point delimitor
+colnames(Ef.design)  <- gsub("_", ".",
+                             colnames(Ef.design))
+
[…]

### 2016-08-08 4_annotations.R

@@ -1,0 +1,149 @@
+## needed from previous script 2_edgeR_diff.R:
+## a TopTags (edgeR) object listing genes tested for differential
+## expression
+
+## The Bioconductor mouse annotation libary
+library(Mus.musculus)
+
+## merge all this information in a data frame to combine mouse
+## annotations with gene_id from Cufflinks get the information from
+## this annotation dbi into data frames
+.get.annot.frame <- function(){
+    ens2geneID <- as.data.frame(org.Mm.egENSEMBL)
+    name2SY <- as.data.frame(org.Mm.egSYMBOL2EG) # get 5 digit "symbol" to "short gene name" mapping
+    SY2Lname<- as.data.frame(org.Mm.egGENENAME) # get 5 digit "symbol" to "long name name" mapping
+    SY2GO <- as.data.frame(org.Mm.egGO) # get 5 digit "symbol" to gene ontology mapping
+    annot.frame <- merge(name2SY, ens2geneID, by = "gene_id")
+    annot.frame <- merge(annot.frame, SY2GO, by = "gene_id")
+    annot.frame <- merge(annot.frame, SY2Lname, by = "gene_id")
+    return(annot.frame)
+}
+
+annot.frame <- .get.annot.frame()
+
+### ORTHOMCL categories
+
+
+read.omcl <- function(path="data/Omcl_groups.txt"){
+  omcl <- readLines(path)
+  omcl <- lapply(omcl, strsplit, " ")
+  omcl <- lapply(omcl, function(x) x[[1]])
+  names(omcl) <- lapply(omcl, function(x) gsub("\\:", "", x[1]))
+  omcl <- lapply(omcl, function(x) x[-1])
+
+  ## make gene-ID identifiers used in the omcl clsters
+  omcl <- lapply(omcl, function (x)
+                 gsub("(Efa|)NODE_(.*_\\d+.g\\d+).t\\d+","\\1EfaB_\\2",  x))
+
+  omcl <- lapply(omcl, function (x) x[!duplicated(x)])
+  omcl <- omcl[!unlist(lapply(omcl, length))==1]
+  omcl <- omcl[!names(omcl)%in%"api10278"] ## exclude a cluster... wonder now why that ;-)
+
+}
+
+omcl <- read.omcl()
+
+get.omcl.categories <- function(ortho.obj){
+  ## Orthomcl
+  get.presence.omcl <- function(ortho.obj){
[…]

### 2016-08-08 4_compareArray.R

@@ -1,140 +1,0 @@
-## RNAseq - Microarray Comparison
-## read the array data as done for Schmidt et al.  "Eimeria
-## falciformis infection of the mouse caecum identifies opposing roles
-## of IFNg-regulated host pathways for the parasite development"
-## DOI:10.1038/mi.2013.115
-
-library(mgug4122a.db)
-library(Agi4x44PreProcess) # !!!! NOT available for R 3.3.0 - NOT on Bioconductor anymore
-library(Biobase)
-library(RSvgDevice)
-library(GGally)
-library(ggplot2)
-
-## read the fold-change data for mouse
-Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
-
-## make it wider format to merge with array results (bleow)
-Mm.DE.FC <- Mm.DE.test[, c("contrast", "gene", "logFC")]
-Mm.DE.FC <- reshape(Mm.DE.FC, idvar = "gene", timevar = "contrast", direction="wide")
-
-### getting all the annotation names to be able to merge based on them
-annot.frame <- read.table("output_data/annotation_data.csv", sep=",")
-
-targets <- readTargets("data/targets.txt", sep=";")
-
-RG <- read.maimages(targets, path= "/SAN/Mouse_arrays/2012_arrays/", source = "agilent", 
-                    other.columns = list(IsFound = "gIsFound",
-                      IsWellAboveBG = "gIsWellAboveBG", 
-                      IsSaturated = "gIsSaturated",
-                      IsFeatNonUnifOF = "gIsFeatNonUnifOL", 
-                      IsFeatPopnOL = "gIsFeatPopnOL",
-                      ChrCoord = "chr_coord"), 
-                    columns = list(G = "gMedianSignal", Gb = "gBGMedianSignal",
-                      R = "rMedianSignal",
-                      Rb = "rBGMedianSignal"), verbose = T, 
-                    sep = "\t", quote = "")
-
-array.names <- apply(targets, 1, function (x) paste(x[c(2, 4,5)], collapse="."))
-colnames(RG) <- array.names
-rownames(targets) <- array.names
-
-RG <- backgroundCorrect(RG, method="minimum", offset=1)
-MA <- normalizeWithinArrays(RG, method="loess", iterations=10)
[…]

### 2016-08-08 5_compareArray.R

@@ -1,0 +1,140 @@
+## RNAseq - Microarray Comparison
+## read the array data as done for Schmidt et al.  "Eimeria
+## falciformis infection of the mouse caecum identifies opposing roles
+## of IFNg-regulated host pathways for the parasite development"
+## DOI:10.1038/mi.2013.115
+
+library(mgug4122a.db)
+library(Agi4x44PreProcess) # !!!! NOT available for R 3.3.0 - NOT on Bioconductor anymore
+library(Biobase)
+library(RSvgDevice)
+library(GGally)
+library(ggplot2)
+
+## read the fold-change data for mouse
+Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
+
+## make it wider format to merge with array results (bleow)
+Mm.DE.FC <- Mm.DE.test[, c("contrast", "gene", "logFC")]
+Mm.DE.FC <- reshape(Mm.DE.FC, idvar = "gene", timevar = "contrast", direction="wide")
+
+### getting all the annotation names to be able to merge based on them
+annot.frame <- read.table("output_data/annotation_data.csv", sep=",")
+
+targets <- readTargets("data/targets.txt", sep=";")
+
+RG <- read.maimages(targets, path= "/SAN/Mouse_arrays/2012_arrays/", source = "agilent", 
+                    other.columns = list(IsFound = "gIsFound",
+                      IsWellAboveBG = "gIsWellAboveBG", 
+                      IsSaturated = "gIsSaturated",
+                      IsFeatNonUnifOF = "gIsFeatNonUnifOL", 
+                      IsFeatPopnOL = "gIsFeatPopnOL",
+                      ChrCoord = "chr_coord"), 
+                    columns = list(G = "gMedianSignal", Gb = "gBGMedianSignal",
+                      R = "rMedianSignal",
+                      Rb = "rBGMedianSignal"), verbose = T, 
+                    sep = "\t", quote = "")
+
+array.names <- apply(targets, 1, function (x) paste(x[c(2, 4,5)], collapse="."))
+colnames(RG) <- array.names
+rownames(targets) <- array.names
+
+RG <- backgroundCorrect(RG, method="minimum", offset=1)
+MA <- normalizeWithinArrays(RG, method="loess", iterations=10)
[…]

### 2016-08-08 A_data_curataion.R

@@ -1,243 +1,0 @@
-## This is a script is 2nd the pipeline and needed for evaluation
-## purpose
-
-## Goals:
-
-## EX- on IN-clusion of samples and genes ###################
-
-## a) decide for samples needed to be excluede because of uncertain
-## infection status
-
-## b) decice for a filter on minimal expression values over all
-## samples to exclude genes
-
-## c) decide on samples needed to be excluded because of bad
-## "unnormalizable" distribution of expression values
-
-
-### Normalization
-
-## d) decide on a general normalization strategy scaling the smaple
-## throughput
-
-## e) decide on a "removeal of unwanted variation" strategy
-
-## f) decide on a potential addional "batch effect" removal strategy
-
-
-## a) decide for samples needed to be excluede because of uncertain
-## infection status
-
-## The raw counts  required from upstream scripts:
-## GENES
-## All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
-
-## TRANSCRIPTS
-All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
-
-## The phenotype information
-All.pData <- read.table("output_data/Sample_pData.csv", sep=",")
-
-library(ggplot2)
-library(reshape)
-library(gridExtra)
-library(xtable)
-
-## Overview for differential representation of host/parasite
-RC.table <-  as.data.frame(cbind(sample=as.character(All.pData$samples),
-                                 seq.method=as.character(All.pData$seq.method),
-                                 batch=All.pData$batch))
-
-## get the total read count.
-## see README_fastq_linecount.txt on how this file was produced    
-## rawfastqRC <- readLines("/SAN/Eimeria_Totta/fastq_linecounts.txt")
-
-## fastqRC <- rawfastqRC[seq(1, length(rawfastqRC), by=2)+1]
-## fastqRC <- as.numeric(as.character(fastqRC))/4
-
-## names(fastqRC) <- rawfastqRC[seq(1, length(rawfastqRC), by=2)]
-## names(fastqRC) <- gsub("^.*?\\/(.*rep\\d).*", "\\1", names(fastqRC) )
-
-## RC.table <- merge(RC.table, fastqRC, by.x="sample", by.y=0)
-## names(RC.table)[ncol(RC.table)] <- "c.seq.reads"
-
-## RC.table$c.seq.reads <- ifelse(RC.table$seq.method%in%"GAIIX",
[…]

### 2016-08-08 fix_gff.R

@@ -1,0 +1,10 @@
+library(rtracklayer)
+
+messed.gff <- import.gff("/SAN/Eimeria_Totta/reference_genomes/mm10_eimeria_merge/indexes_bowtie2/mm10_GRCm38_eimeriaHaberkorn.gtf")
+
+messed.gff$gene_id <- gsub("_\\d+$", "", messed.gff$gene_id)
+messed.gff$gene_name <- gsub("_\\d+$", "", messed.gff$gene_name)
+
+good.gff <- messed.gff
+
+export.gff(good.gff, "/SAN/Eimeria_Totta/reference_genomes/mm10_eimeria_merge/indexes_bowtie2/mm10_GRCm38_eimeriaHaberkorn_fixed.gtf")

### 2016-08-08 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
Manuscript changes while working on feature counts integration and simple csv output to be handed over between scripts

@@
 
-\author[
-%	addressref={aff1},
-%	noteref={n1},
-]{\inits{TK}\fnm{Totta} \snm{Kasemo}}
-\author[
-%	addressref={aff1},
-]{\inits{SS}\fnm{Simone} \snm{Spork}}
-\author[
-%	addressref={aff2},
-%	noteref={n1},
-]{\inits{CD}\fnm{Christoph} \snm{Dieterich}}
-\author[
-%	addressref={aff1},
-]{\inits{RL}\fnm{Richard} \snm{Lucius}}
-\author[
-%	addressref={aff1, aff3},                   % id's of addresses, e.g. {aff1,aff2}
-	corref={aff1, aff3},                       % id of corresponding address, if any
-%	noteref={n1},                        % id's of article notes, if any
-	email={emanuel.heitlinger@hu-berlin.de}   % email address
-]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
+\author[ <addressref>={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
+\author[ <addressref>={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
+\author[ <addressref>={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
+\author[ <addressref>={aff1},]{\inits{RL}\fnm{Richard} \snm{Lucius}}
+\author[ <addressref>={aff1, aff3},              
+  corref={aff1},                   
+  email={emanuel.heitlinger@hu-berlin.de}]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
 
@@
 
-\address[id=aff1]{%                           % unique id
-	\orgname{Department of Biology, Humboldt-Universitat zu Berlin}, % university, etc
-	\street{Philippstr. 13, Haus 14},                     %
-	\postcode{10115}                                % post or zip code
-	\city{Berlin},                              % city
-	\cny{Germany}                                    % country
+\address[id=aff1]{
+	\orgname{Institute of Biology, Humboldt-Universitat zu Berlin}, 
+	\street{Philippstr. 13, Haus 14},
+	\postcode{10115}                 
+	\city{Berlin},                   
+	\cny{Germany}                    
 }
-\address[id=aff2]{%  
+\address[id=aff2]{  
 	\orgname{Max Planck Institute for Biology of Ageing},
@@
 }
-\address[id=aff3]{%  
-	\orgname{Department Reproduction Biology and Evolutionary Ecology, Leibniz Institute of Zoo and Wildlife Research},
+\address[id=aff3]{  
+	\orgname{Leibniz Institute of Zoo and Wildlife Research},
 	\street{Alfred-Kowalke-Str. 17},
@@
 
-  \section*{Experimental overview}
+  \section*{A dual transcriptomics experiment}
 We performed mRNA sequencing of caecum tissue from mice infected with
-the apicomplexan parasite \textit{E. falciformis}. For each replicate
-sample we enriched caecum tissue from three infected mice for
-epithelial cells. We extracted mRNA and constructed sequencing
-libraries for different timepoints after infection and mouse strains
-with assumed different immunocompetence.
-%% we have to bring in Simones data first with two sentences and a
-%% figure or write it like this
+the apicomplexan parasite \textit{E. falciformis}. We worked with two
+biological replicates for all but two conditions (with one and three
+replicates, respecitively). For each replicate sample we enriched
+caecum tissue from three infected mice for epithelial cells. We
+extracted mRNA and constructed sequencing libraries for different
+timepoints after infection and mouse strains which displayed different
+immunocompetence in infection trials (see figure x).
+
+Briefly RAG-1-/- knockout mice, in which mature B and T lymphocytes
+are absent showed ...
+
 Additonally we obtained data for challenge infections of mice
-immunized one month prior to the experiment. For the parasite oocysts
-and sporozoites were included as ``environmental'' lifecycle stages.
+immunized one month prior to the experiment.
+
+For the parasite oocysts and sporozoites were included as
+``environmental'' and ``\textit{in vitro}'' processed lifecycle
+stages.
 
@@
 influencing analysis and quality of our results (additional file xyc).
+
+The results of these multivariate analyses in relation to biological
+interpretation are described further below.
 
@@
 
+One sample (NMRI\_1stInf\_0dpi\_rep1) was additonally excluded because
+the uninfected control showed an unexpected mapping of reads to the
+\textit{E. falciformis} genome. We consider this together with the
+unsuccessfully infected samples above to display an undecided state of
+infection.
+
 On the other end of the spectrum mice not previously immunized
 featured an mRNA ``holo-transcriptome'' at 7dpi with the majority of
-mRNA originating from the parasite as opposed to the host (Figure 1).
+mRNA originating from the parasite as opposed to the host (Figure
+1). The overall mRNA output of the sampled caecum tissue at this point
+of infection is dominated by the parasite with proportional mRNA
+abundance between XXX and XXX %
+
+The total number of sequencing reads obtained as well as those reads
+mapped to the transcriptome of either \textit{E. falciformis} and the
+mouse are indicated in Table 1 for all replicates.
+
+\section*{The mouse transcriptome changes in the course of infection infection}
+
+Statistical testing for differetial expression between infected and
+uninfected mice, revealed changes in mRNA abundance becoming more
+pronounced (involving more genes) at later post infection (especially
+at 7dpi; Table 1).
+
+The observed changes between controls and 7dpi were in agreement with
+previously published microarry data. Fold-Change data obtained from
+mice at 6dpi on Agilent microarrays (Schmidt et al 2012) or 7dpi (our
+RNAseq data) agains uninfected controls shows a strong correlateion
+(Spearman's \rho = 333; $R^2$ from a linear model = 444 ; figure 2
+a). Considering both and technical differences between the tow methods
+and the biological changes along the lifecycle this confirms the
+adequacy of our methods for assessing the host transcriptome.
+
+XX mouse genes were considered differently expressed (DE; FDR<1\%)
+between uinfected and 3dpi, XX genes between controls and 5dpi, xxx
+genes between 0dpi and 7dpi. This lead to combined set of xxxx genes
+reacting to infection (see figure 2 b).
+
+We detected no distinct response early in infection.  When testing for
+differences between 3dpi and later infection (5dpi and 7dpi) DE genes
+were simply a subset of genes DE between respective infections and
+controls.
+
+To further validate these patterns we performed hierarchical
+clustering on mouse mRNAs DE at different dpi (union of DE
+genes).
+
+%% Samples from day 7 p.i. form a separate cluster (NMRI mice only), not
+%% distinguishing between first and challenge infection. For these
+%% samples, the pattern is most pronounced in gene cluster 1 (up), and 4
+%% and 6 (down). Within cluster 1, the second replicate from the
+%% challenge infection has a deviant profile (compare mouse data in
+%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
+%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
+%% 3 and 5 p.i. from all three mouse strains cluster together. These
+%% samples are distinct from oocysts, NMRI day 7 p.i., and
+%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
+%% oocysts or day 7.
+
+%%% Test _this_:
+The more ponounced changes later in infection are is likely reflecting
+a lag time in detection before the parasite reaches sufficient
+intensity during asexual expansion (schizogony). Those changes are
+mainly related to the onset of an adaptive immune response.
+
+\section*{The lifecycle of Eimeria is characterized by major changes in mRNA abundance}
+
+Table 2 lists pairwise sample comparisons and number of genes with
+differently abundant mRNAs per comparison."NMRI" followed by number
+indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
+p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
+implemented in edgeR are included. NAs are missing samples or not
+applicable for the species. NA* is due to missing NMRI day 0 sample
+from first infection. For \textit{E. falciformis} no mRNAs are
+significantly differently abundant between first and second infection,
+whereas in mouse there are some differences, especially on day 7
+p.i. However, upon hierarchical clustering of these genes (union of
+500 with lowest FDR), samples do not cluster according to first/second
+infection (Figure SX ?). In comparisons between mouse strains, 22
+genes are detected to be different in the parasite data. The
+difference is however not between Rag1-/- mice as expected, but
+between NMRI and C57BL/6. For mouse....
+
+We applied hierarchical clustering on \textit{E. falciformis} mRNAs
+which were differently abundant at different timepoints p.i.,
+including comparisons with oocysts and sporozoites (union of 500 with
+lowest FDR). Samples from day 7 p.i. form a separate cluster (NMRI
+mice only), not distinguishing between first and challenge
+infection. For these samples, the pattern is most pronounced in gene
+cluster 1 (up), and 4 and 6 (down). Within cluster 1, the second
+replicate from the challenge infection has a deviant profile (compare
+mouse data in Figure 2). Distinct groups of genes also define
+sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6
+down). mRNA profiles on days 3 and 5 p.i. from all three mouse strains
+cluster together. These samples are distinct from oocysts, NMRI day 7
+p.i., and sporozoites. Sporozoites cluster closer with days 3 and 5
+than with oocysts or day 7.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection*{GO term enrichments in heatmap clusters}
+The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
+in \textit{T. gondii}
+
+\subsubsection*{Data overview}
+In total, xx reads were sequenced. For each sample, mouse read numbers were $10^7$ - $10^8$
+Quality filters applied resulted in xx reads being removed from further analysis
+
+\subsubsection*{Preparation for invasion in oocysts}
+The mRNA profile in the oocyst stage is mainly determined by highly
+abundant genes in cluster 4.  Overrepresented GO-terms in this cluster
+are enriched by ortholog genes to peptidases, microneme localized
+proteins reported to be involved in invasion, genes associated with
+adhesion in protozoans (and with clotting in higher eukaryotes), and
+genes that are annotated to be involved in amino acid biosynthesis.
+Aminopeptidase N ('related' annotation) is the reported ortholog for
+three genes with abundant mRNAs in oocysts. In humans, this enzyme has
+been reported to cleave peptides bound to major histocompatibility
+complex, MHC, II (UniProt reference if we want to keep this... but
+does any secretion happen from oocysts...? Or is this too far-fetched
+to be interesting?).
+
+A Thrombospondin type 1 domain-containing protein ortholog is highly
+abundant in cluster 4 (high abundance in oocysts). Thrombospondin type
+1 domains have been reported in \textit{E. tenella} microneme
+localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella}
+MIC4 mRNA was reported in sporozoites where it localizes to the apical
+end, and in late schizonts and late oocyst stages, when sporozoites
+are forming. (Tomley01). For the same gene, the \textit{T. gondii}
+annotation is Sushi domain-containing protein, which is also the
+ortholog annotation of another gene in this cluster. In the related
+malaria parasite \textit{P. falciparum} the apical sushi protein, ASP,
+(which has a sushi domain) localizes to micronemes in merozoites but
+not other stages (OKeeffe05).  Limulus clotting factor C, Coch-5b2
+(Cochlin) and Lgl1, LCCL, (syn. F5/8 domain) domains are associated
+discoidin lectin domains and thereby with adhesion.  (Pfam entries for
+'LCCL domain' and 'Discoidin domain', May 2016).  Taken together, this
+indicates that the thrombospondin and sushi-domain genes
+(EfaB\_MINUS\_4114.g412 and EfaB\_PLUS\_1425.g183) are involved in
+sporozoite invasion in \textit{E. falciformis} and that the mRNAs are
+transcribed and available before excystation.  A role in merozoite
+re-invasion in \textit{E. falciformis} is not indicated by our data.
+The LCCL domain annotation and thrombospondins role in higher
+eukaryotes also indicates that adhesion or preparation for adhesion is
+important in oocysts.  We suggest (speculate...?) that the
+thrombospondin annotated ortholog and the LCCL domain-containing
+protein (EfaB\_MINUS\_11233.g986) are involved in cell adhesion in
+\textit{E. falciformis}.
+%One examples is found in the slime mold \textit{Dictyostelium
+%discoideum)} in which the LCCL domain is part of the mold's discoidin
+%adhesion protein. (Pfam entry for 'discoidin domain', May 2016)
+
+\subsubsection*{Amino acid biosynthesis in oocysts}
+High abundance of aminotransferase mRNAs indicate amino acid
+biosynthesis or preparation for the same in oocysts (cluster 4).  We
+identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase
+orthologs, which are enzymes contributing to L-serine and L-alanine
+production, respectively. A putative \textit{Eimeria}
+spp. cystathionine beta-synthase, CBS, in this cluster also indicates
+de novo cysteine production. Alkyl sulfatase mRNA is also abundant in
+oocysts. Generally, this enzyme enables an organism to exploit organic
+sulfur to produce and incorporate inorganic sulfur into the amino
+acids cysteine and methionine, when no inorganoc sulfur is available.
+
+'Embryonic development' Nicalin 1, patched family protein (hedgehog)
+
+\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
+MmgE/PrpD is overrepresented in oocysts. The enzyme is important for
+propionate catabolism in the 2-methylcitric acid cycle and has been
+shown to be used by the intestinal intracellular bacterium
+\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).
+Propionate is one of two most abundant small-chain fatty acids in the
+gut along with butyrate. Both fatty acids are largely produced as
+degradation products from food by commensal bacteria (Sun13). Sharing
+the intestines as a niche with \textit{S. typhimurium} it is possible
+that also \textit{E. falciformis} uses Mmg/PrpD to exploit available
+propionate for pyruvate production.
+
+\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
+Interestingly, the genes described above which are thought to be
+involved in amino acid biosynthesis and invasion are highly abundant
+in oocysts but are underrepresented in schizont stages (day 3 and day
+5 samples) and even in sporozoites. An average abundance was detected
+on day 7 for these genes, indicating a role in either gametes or early
+oocyst formation.  This pattern supports the suggestion that these
+specific mRNAs (cluster 4) for invasion and biosynthetic processes are
+prepared (and possibly expressed) in the oocyst stage but are no
+longer detectable in the cell at the timepont when the protein is
+assumed to be in use (sporozoites and merozoite stages).  Therefore,
+correlating mRNA prevalence with biological function at the timepoint
+when mRNAs are detected must be done with care.
+
+\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
+Oocysts: profile for 3, 5, 6:
+
+Sporozoites: 3 and 5
+
+Day 7: 5 and 7
+
+Specific genes....  Enolase 2, encoded by Eno2, is among the
+downregulated genes in oocysts and sporozoites.  In \textit{T. gondii}
+the paralog Eno1 is strongly associated with the cyst (bradyzoite)
+stage and Eno2 is associated with tachyzoite stages (kibe05). It is
+therefore expected that this mRNA is underrepresented in oocysts and
+our data also show that the same is true in sporozoites for
+\textit{E. falciformis}. (TK: If important we could look specifically
+for Eno1 in cluster 4).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
+Two clusters contain genes with mRNAs highly abundant on day 7 p.i;
+cluster 1 and 2.  Dynein, kinesin and tubulin are annotations highly
+represented among orthologs of genes in both these clusters. The
+annotations indicate an important role for motility at this timepoint,
+probably reflecting development of microgametes.  In addition, in
+cluster 2, there are two 'EF-hand domain containing proteins'
+annotations as well as caltractin, centrin-1, and troponin
+annotations. Caltractin and centrin-1 are associated with the
+centrosome and structure and function of microtubuli in mammals, and
+troponin is linked to muscle function (UniProt).  Also potentially
+linked to motility is the occurence of growth arrest specific protein
+8, Gas8, which in the mouse has been reported to be highly expressed
+in the testes and important for mouse sperm function (Yeh02).
+
+Other genes among the 38 indicate carbon fixation
+(glycolysis/gluconeogenesis) or conversions of nucleoside
+phosphates. In addition, a Ras family protein, RNA polymerase II
+transcription initiation factor and Sec23 and Sec24 were among
+orthologs identified in \textit{E. falciformis} cluster 2.
+
+In cluster 1, carbon metabolism genes are represented by
+6-phosphogluconate dehydrogenase and glycogen phosphorylase family
+protein 1. UDP-glucose 4-epimerase and amiloride-sensitive amine
+oxidase are reported as upregulated in gametocytes in
+\textit{E. tenella} by RNA-seq (Walker15) and suggested by those
+authors to play a role in cyst wall synthesis.
+
+\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
+Unintuitively for a protozoan organism, seven out of eight GO
+biological process terms in cluster 1 are associated with wound
+healing and blood coagulation.  An explanation is offered by some of
+the orthologs to the three \textit{E. falciformis} genes responsible
+for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and
+\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal
+protein MIC4, related' (E. tenella) and more generally for several
+other protozoa, 'PAN domain containing proteins'.  The PAN domain is
+found in the plasminogen/hepatocyte growth factor family and in
+coagulation factor XI family (REF), explaining why terms related to
+blood coagulation are enriched by these genes.  Later publications on
+\textit{T. gondii} (Marchant12) also associate PAN domains and
+proteins in apicomplexan parasites with micronemes and therefore
+invasion. In our case, this is peculiar, since the enrichment appears
+on day 7 p.i.. A possible role at this timepoint is suggested by work
+on the fungi \textit{Sclerotinia sclerotiorum} where Yu et
+al. reported an important role for PAN domain proteins in cell wall
+integrity (Yu12). This role for MIC proteins has to our knowledge not
+been investigated in apicomplexan parasites.  The PAN domain domain
+has also been reported to be common in nematodes such as
+\textit{Caenorhabditis elegans}, however the function is not
+understood. (Thordai99) The other two GO terms in the cluster of day
+seven upregulated genes are DNA replication and DNA replication
+initiation, which most likely reflects late stage schizogony or gamete
+formation.  Six genes contribute to this enrichment and orthologs are
+either annotated as DNA replcation licencing factors, DNA polymerases
+or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.
+
+\subsubsection*{Gene and sample patterns by hierarchical clustering}
+Samples (columns) cluster into two major clusters where day 7
+p.i. samples form one group distinct from other samples. In the second
+group, oocsts and sporozoites are distinct and sporozoites cluster
+most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples
+also cluster into two groups, of which one contains all NMRI day 5
+p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
+clusters have no obvious patters.
+
+For gene clusters (rows), the two groups with high mRNA abundance on
+day 7 p.i. (cluster 1 and 2) do not cluster most closely with each
+other, but with the cluster for high mRNA abundance in oocysts
+(cluster 1 association) and with the cluster for high mRNA abundance
+in sporozoites (cluster 2 association).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%	DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Discussion}
+In our analysis we demonstrate which biological processes are dominant
+in different life cycle stages of \textit{E. falciformis} in the
+mouse. The RNAseq transcriptome provided here allows for detailed
+analysis of genes involved in those processes, providing candidates
+for life stage specific marker in \textit{Eimeria} spp. research.
+
+\ldots
+\ldots
+\ldots
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%	METHODS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}
+\subsection{Mice and infection procedure}
+Three strains of mice were used in our experiments: NMRI (Charles
+River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
+C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
+procedures were performed according to the German Animal Protection
+Laws as directed and approved by the overseeing authority Landesamt
+fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
+as described by Schmid et al., (Schmid12), but tapwater was used
+instead of PBS for administration of oocysts. Briefly, NMRI mice were
+infected two times, which will be referred to as first and second
+infection. For the first infection, 150 sporulated oocysts were
+administered in 100 $μ$L by oral gavage. During the first infection of
+60 mice, all animals were weighed every day. On day zero, before
+infection, as well as on day three, five and seven days post
+infection, dpi, caeca from 3-4 sacrificed mice per time point were
+collected. Epithelial cells were isolated as described in Schmid et
+al.(schmid12). For challenge infection, mice recovered for four weeks
+before second infection.  Recovery was monitored by weighing and
+visual inspection of fur. For the second infection, 1500 sporulated
+oocysts were applied by oral gavage. Three mice were used as
+non-second infection control, referred to as day 0, second infection.
+
+\subsection{Oocyst purification for infection and sequencing}
+Sporulated oocysts were purified by flotation from feces stored in
+potassium dichromate and administered orally in 100 uL tapwater. One
+\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
+Haberkorn 1970, was used for all infections and parasite samples. The
+strain is maintained through passage in NMRI mice in our facilities as
+described elsewhere (schmid12).
+
+\subsection{Sporozoite isolation}
+Sporozoites were isolated from sporocysts by excystment. For this,
+sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
+tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
+30 min. Sporozoites were purified by the method of Schmatz et al
+(schmatz--).
+
+\subsection{RNA extraction}
+Total RNA was isolated from infected epithelial cells, sporozoites and
+sporulated oocysts using Trizol according to the manufacturer’s
+protocol (Invitrogen). High quality \emph{what is the meaning of 'high
+  quality' here?}  RNA was used to produce an mRNA library using the
+Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
+  genome paper} Sporozoites were stored in 1 mL Trizol until
+RNA-isolation.  Total RNA was isolated using the PureLink RNA Mini Kit
+(Invitrogen) and reverse transcribed into cDNA.
+
+\subsection{Sequencing, sequence quality assessment and alignment}
+cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
+specified in SI xx (both unstranded). A fastq\_quality\_filter
+(FASTQ-toolkit, version 0.0.14, available at
+https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
+Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
+phred score of 10 was applied. We further set q = 60. These settings
+require that nine out of ten bases or more are correct in at least
+60\% of the bases for each read.
+
+\subsection{Alignment and reference genomes}
+We used the published \textit{Mus musculus} mm10 assembly (Genome
+Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
+genome including annotations for mouse data. The
+\textit{E. falciformis} genome (Heitlinger14) was downloaded from
+ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
+files were merged into a dual reference genome, and files including
+mRNA sequences from both species were aligned against the dual
+reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
+specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
+genome. Single-end and pair-end sequence samples were aligned
+separately with library type 'fr-unstranded' specified for pair-end
+samples. Import into R was enabled by the R package Ballgown, which
+requires bam files to be processed by Tablemaker (Frazee15), in our
+case used with -qW -G specified. Tablemaker in turn makes use of
+Cufflinks (version 2.1.1, Trapnell10).
+
+\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
+Count data was normalized using the R-package edgeR (version 3.14.0;
+cite) with the upperquartile normalisation method. Briefly, genes with
+zero coverage in all samples (libraries) are removed and normalisation
+factors are calculated for the 75\% quantile for each library. This
+normalisation is suitable for read densities following a negative
+binomial distribution. Two samples contradicted this assumption
+(parasite data) for later modelling and both mouse and parasite data
+from these samples were excluded from further analysis:
+NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
+then fits a generalized linear model (GLM with a negative binomial
+link function) for each gene (glmFit) and then performs likelihood
+ratio tests for models w or w/o focal factor (glmLRT).
+
+%% ........... include outcome of MDS analysis here?
+%% No, results never are presented in methods.
+
+\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
+A selection of differently abundant mRNAs are used for hierarchical
+clustering of \textit{E. falciformis} life cycle relevant genes. In
+each comparison (see Table 3), the union of the at most 500 genes
+differentially abundant with lowest FDR (<0.05) are selected. In the
+next step, the 500 mRNAs from each comparison (or less) are
+joined. For \textit{E. falciformis} time p.i. comparison, this
+resulted in 1618 unique genes selected as differently expressed . The
+22 genes in the NMRI vs C57BL/6 comparison are not included in the
+\textit{E. falciformis} life cycle analysis. The same comparisons for
+mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
+all samples, i.e. also samples which themselves did not have any
+significantly different mRNAs according to our selection, were
+included in hierarchical clustering. Scale bar in heatmaps show 0 as
+mean mRNA abundance for each gene (row). Up (green) and
+down-regulation (brown) denote number of standard deviations from 0,
+i.e., row mean. Hierarchical clustering was performed using with
+Euclidean distances, using complete linkage ('complete', R package
+base).
+
+%% the paragraph above has results. All those number
+
+\subsection{Differentially abundant mRNAs used for hierarchical clustering}
+All analyses were performed in R (cite R-core). Complete scripts are
+available at https://github.com/derele/Ef\_RNAseq.git tagged as version
+1.0.
+%% We will make this available as a github webpage!!!
+
+\setlength{\tabcolsep}{10pt}
+\begin{table}[H]
+\small
+\begin{center}
+\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
+\begin{tabular}{*3l}    \toprule
+	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
+	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
+	(including oocysts and sporozoites if appl.)	\\	
+	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
+\hline
+\end{tabular}
+\end{center}
+%\end{adjustwidth}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Backmatter begins here                   %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{backmatter}
+
+\section*{Competing interests}
+  The authors declare that they have no competing interests.
+
+\section*{Author's contributions}
+    Text for this section \ldots
+
+\section*{Acknowledgements}
+  Text for this section \ldots
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                  The Bibliography                       %%
+%%                                                         %%
+%%  Bmc_mathpys.bst  will be used to                       %%
+%%  create a .BBL file for submission.                     %%
+%%  After submission of the .TEX file,                     %%
+%%  you will be prompted to submit your .BBL file.         %%
+%%                                                         %%
+%%                                                         %%
+%%  Note that the displayed Bibliography will not          %%
+%%  necessarily be rendered by Latex exactly as specified  %%
+%%  in the online Instructions for Authors.                %%
+%%                                                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% if your bibliography is in bibtex format, use those commands:
+\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
+% for author-year bibliography (bmc-mathphys or spbasic)
+% a) write to bib file (bmc-mathphys only)
+% @settings{label, options="nameyear"}
+% b) uncomment next line
+%\nocite{label}
+
+% or include bibliography directly:
+% \begin{thebibliography}
+% \bibitem{b1}
+% \end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Figures                       %%
+%%                               %%
+%% NB: this is for captions and  %%
+%% Titles. All graphics must be  %%
+%% submitted separately and NOT  %%
+%% included in the Tex document  %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% Do not use \listoffigures as most will included as separate files
+
+\section*{Figures}
+
+\begin{figure}[h!]
+%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
+  \caption{\csentence{Experimental design and outcome of a dual RNAseq experiment}
+    read sequences per mouse read sequences for each experimental
+    condition. The more advanced the infection is, the higher the
+    ratio of parasite reads is, reflecting parasite replication
+    between day zero and day seven. ......compare 1st and 2nd
+    ....... patterns between Rag and NMRI/C57BL/6?  ........ Values
+    are mean of replicates (n=2, if * n=1) on log10 scale. Each sample
+    (replicate) consists of mRNA from three different mice.}
+  \end{figure}
+
+%%%%%
+figure 1 label cotribution of mouse vs. parasite transcripts to the
+overall transcriptional output of the analysed tissue for all samples.
+%%%%
+
+%% Selecting at most at most 500 genes differentially abundant with
+%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
+%% each comparison (or less) are joined. For \textit{E. falciformis} time
+%% p.i. comparison, this resulted in 1618 unique genes selected as
+%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
+%% are not included in the \textit{E. falciformis} life cycle
+%% analysis. The same comparisons for mouse yielded 8052 genes in total
+%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
+%% which themselves did not have any significantly different mRNAs
+%% according to our selection, were included in hierarchical
+%% clustering.
+
+%% This might be figure label
+%% Scale bar in heatmaps show 0 as mean mRNA abundance for
+%% each gene (row). Up (green) and down-regulation (brown) denote number
+%% of standard deviations from 0, i.e., row mean.
+
+\subsection{Life cycle analysis... \textit{E. falciformis}}
+\begin{figure}[h!]
+%\begin{center}
+%\centering
+	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
+	\caption{\textit{E. falciformis} mRNAs with significantly
+          different abundance at different times p.i. in NMRI mice
+          (see Table 2). Samples cluster according to expected
+          developmental state of the parasite, without distinguishing
+          between first and challenge infection. Hierarchical
+          clustering was set to apply seven gene clusters.}
+%\end{center}
+\end{figure}
+
+\subsection{Parasite development reflected in mouse data}
+\begin{figure}[h!]
+%\begin{center}
+%\centering
+\includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
+\caption{Mouse mRNAs with significantly different abundance at
+  different times p.i. in NMRI mice (see Table 2). Samples cluster
+  according to expected developmental state of the parasite, without
+  distinguishing between first and challenge infection. Rag1-/- mouse
+  samples cluster together. Hierarchical clustering was set to apply
+  four gene clusters.}
+%\end{center}
+\end{figure}
+
+When clustering mouse mRNAs which were different in time
+p.i. comparisons in Table 2, three out of four samples from day 7
+p.i. (NMRI only) cluster together. These samples are characterized by
+gene cluster 3 (up) as well as 2 and 4 (down). The fourth day 7-sample
+(challenge infection, second replicate) clusters with day 3 and 5
+samples. Upon visual inspection of this sample in gene cluster 4, it
+diplays a profile similar to non-infected mice. The same sample is
+abnormal in the parasite profile (Figure 1). NMRI and C57BL/6
+uninfected samples (0dpi) cluster together, defined by gene clusters 4
+(up), and 1 (down). Rag1-/- day 0 samples are however less pronounced
+in gene cluster 1. Day 3 and 5 samples cluster together, with Rag1-/-
+forming a separate group and Rag1-/- non-infected most distant in this
+sample cluster.
+
+% \begin{figure}[h!]
+ % \includegraphics[width=0.7\linewidth]{ef-heatmap} \caption{\csentence{Parasite
+ % genes with different mRNA abundance between samples, sorted into
+ % five clusters by hierarchical clustering (method = complete,
+ % distance = euclidean).}  \textit{E. falciformis} samples from seven
+ % days p.i. cluster together (NMRI mice only). These samples have
+ % distinct mRNA abundance patterns in all gene clusters, although
+ % more pronounced in clusters 1 (up) and 4 (down). Distinct groups of
+ % genes also define sporozoites (cluster 5, up) and oocysts (cluster
+ % 2, up). mRNA profiles on days three and five p.i. from all three
+ % mouse strains cluster together. These samples are distinct from
+ % oocysts, NMRI day 7 p.i., and sporozoites, however closest to the
+ % latter. On scale bar, 0 is mean mRNA abundance for each gene. Up
+ % and downregulation is standard deviations from mean.}  \end{figure}
+
+%\begin{figure}[h!]
+%  \includegraphics[width=0.7\linewidth]{mm-heatmap} \caption{\csentence{Mouse
+%  genes with different mRNA abundance between samples, sorted into
+%  four clusters by hierarchical clustering (method = complete,
+%  distance = euclidean).}  Three out of four samples from day 7
+%  p.i. (NMRI only) cluster together. These samples are characterized
+%  by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6
+%  uninfected samples cluster together (left), defined by clusters 3,
+%  4 (up), and 1 (down). Rag1-/- samples cluster together. These
+%  samples share a weak downregulation of most genes in cluster 2 as
+%  well as upregulation of a small group genes in the same
+%  cluster. Uninfected Rag1-/- samples are separated from infected
+%  ones with distinct profiles in clusters 1, 3, and 4. On scale bar,
+%  0 is mean mRNA abundance for each gene. Up and downregulation is
+%  standard deviations from mean.}  \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Tables                        %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Use of \listoftables is discouraged.
+%%
 
@@
 
-% latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Fri Aug  5 13:00:01 2016
 \begin{table}[ht]
@@
 
-%%%%%
-figure 1 label
-cotribution of mouse vs. parasite transcripts to the overall
-transcriptional output of the analysed tissue for all samples.
-%%%%
-
-The total number of sequencing reads obtained as well as those reads
-mapped to the transcriptome of either \textit{E. falciformis} and the
-mouse are indicated in Table 1 for all replicates.
-
-%% Before was not pretty. All those NA's. If people are like me they
-%% hate NAs in tables ;-). I think you sould make the above table from
-%% data with xtable.
-
-%% best use a straight 2d table with...
-
-% label & mouse strain & parasite dpi & naive/challenge & number of
-% reads & #mapped to mouse & #mapped to \textit{E. falciformis}
-
-%% column name "naive/challenge" can change
-
-%% as technical as it seems people will want to know what our labels
-%% mean (they appear in figures throughout) therfore it is good to
-%% have them in table 1
-
-\section*{The mouse transcriptome changes in the course of infection infection}
-
-Statistical testing for differences between infected and uninfected
-mice, revealed changes in mRNA abundance becoming more pronounced
-(involving more genes) at later timepoints of infection (Table 1).
-
-The observed fold-changes were in agreement with previously published
-microarry data. Fold-Change data obtained from mice at 6dpi on Agilent
-microarrays (Schmidt et al 2012) or 7dpi (our RNAseq data) shows a
-strong correlateion (Spearman's \rho = 333; $R^2$ from a linear model
-= 444 ; see also figure 2). Considering both and technical differences
-between the tow methods and the biological changes along the lifecycle
-this confirms the adequacy of our methods for assessing the host
-transcriptome.
-
-We selected all mouse genes differently expressed (FDR<1\%) between
-uinfected and any of the infected samples.
-
-%%% This is from methods but all the results should be presented here:
-
-%% Selecting at most at most 500 genes differentially abundant with
-%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
-%% each comparison (or less) are joined. For \textit{E. falciformis} time
-%% p.i. comparison, this resulted in 1618 unique genes selected as
-%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
-%% are not included in the \textit{E. falciformis} life cycle
-%% analysis. The same comparisons for mouse yielded 8052 genes in total
-%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
-%% which themselves did not have any significantly different mRNAs
-%% according to our selection, were included in hierarchical
-%% clustering.
-
-Those changes are mainly related to the onset of an
-adaptive immune response.
-
-\section*{The lifecycle of Eimeria is characterized by large changes in mRNA abundance}
-
-%%% This is from methods but all the results should be presented here:
-
-%% Selecting at most at most 500 genes differentially abundant with
-%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
-%% each comparison (or less) are joined. For \textit{E. falciformis} time
-%% p.i. comparison, this resulted in 1618 unique genes selected as
-%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
-%% are not included in the \textit{E. falciformis} life cycle
-%% analysis. The same comparisons for mouse yielded 8052 genes in total
-%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
-%% which themselves did not have any significantly different mRNAs
-%% according to our selection, were included in hierarchical
-%% clustering.
-
-%% This might be figure label
-%% Scale bar in heatmaps show 0 as mean mRNA abundance for
-%% each gene (row). Up (green) and down-regulation (brown) denote number
-%% of standard deviations from 0, i.e., row mean.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
-\textit{Mouse strain comparisons} & 		 & 	 \\ \midrule
-	NMRI vs C57BL/6 		& 22 	& NA* \\		
-	NMRI vs Rag1-/-     		& 0 & NA* 	\\
-	C57BL/6 vs Rag1-/-  		& 0   & 356 	\\ \midrule
 \bottomrule
@@
 
-Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison."NMRI" followed by number
-indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
-p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
-implemented in edgeR are included. NAs are missing samples or not
-applicable for the species. NA* is due to missing NMRI day 0 sample
-from first infection. For \textit{E. falciformis} no mRNAs are
-significantly differently abundant between first and second infection,
-whereas in mouse there are some differences, especially on day 7
-p.i. However, upon hierarchical clustering of these genes (union of
-500 with lowest FDR), samples do not cluster according to first/second
-infection (Figure SX ?). In comparisons between mouse strains, 22
-genes are detected to be different in the parasite data. The
-difference is however not between Rag1-/- mice as expected, but
-between NMRI and C57BL/6. For mouse....
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-\subsection{Life cycle analysis... \textit{E. falciformis}}
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
-	\caption{\textit{E.
-falciformis} mRNAs with significantly different abundance at different times p.i.
-in NMRI mice (see Table 2).
-Samples cluster according to expected developmental state of the parasite, without distinguishing between first and challenge infection.
-Hierarchical clustering was set to apply seven gene clusters.}
-%\end{center}
-\end{figure}
-
-We applied hierarchical clustering on \textit{E.
-falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR).
-Samples from day 7 p.i.
-form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection.
-For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down).
-Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2).
-Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down).
-mRNA profiles on days 3 and 5 p.i.
-from all three mouse strains cluster together.
-These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites.
-Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.
-
-\clearpage
-\subsection{Parasite development reflected in mouse data}
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-\includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
-\caption{Mouse mRNAs with significantly different abundance at different times p.i.
-in NMRI mice (see Table 2).
-Samples cluster according to expected developmental state of the parasite, without distinguishing between first and challenge infection.
-Rag1-/- mouse samples cluster together.
-Hierarchical clustering was set to apply four gene clusters.}
-%\end{center}
-\end{figure}
-
-When clustering mouse mRNAs which were different in time p.i.
-comparisons in Table 2, three out of four samples from day 7 p.i.
-(NMRI only) cluster together.
-These samples are characterized by gene cluster 3 (up) as well as 2 and 4 (down).
-The fourth day 7-sample (challenge infection, second replicate) clusters with day 3 and 5 samples.
-Upon visual inspection of this sample in gene cluster 4, it diplays a profile similar to non-infected mice.
-The same sample is abnormal in the parasite profile (Figure 1).
-NMRI and C57BL/6 uninfected samples (0dpi) cluster together, defined by gene clusters 4 (up), and 1 (down).
-Rag1-/- day 0 samples are however less pronounced in gene cluster 1.
-Day 3 and 5 samples cluster together, with Rag1-/- forming a separate group and Rag1-/- non-infected most distant in this sample cluster.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-\subsection*{GO term enrichments in heatmap clusters}
-The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
-in \textit{T. gondii}
-
-\subsubsection*{Data overview}
-In total, xx reads were sequenced. For each sample, mouse read numbers were $10^7$ - $10^8$
-Quality filters applied resulted in xx reads being removed from further analysis
-
-\subsubsection*{Preparation for invasion in oocysts}
-The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
-Overrepresented GO-terms in this cluster are enriched by ortholog genes to 
-peptidases, microneme localized proteins reported to be involved in invasion, 
-genes associated with adhesion in protozoans (and with clotting in higher eukaryotes),
-and genes that are annotated to be involved in amino acid biosynthesis.
-  Aminopeptidase N ('related' annotation) is the reported ortholog for three genes with 
-abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
-major histocompatibility complex, MHC, II (UniProt reference if we want to keep this... but does any
-secretion happen from oocysts...? Or is this too far-fetched to be interesting?).
-
-A Thrombospondin type 1 domain-containing protein ortholog is highly abundant in cluster 4 (high 
-abundance in oocysts). Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme
-localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella} MIC4 mRNA was reported in 
-sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages,
-when sporozoites are forming. (Tomley01). For the same gene, the \textit{T. gondii} annotation is 
-Sushi domain-containing protein, which is also the ortholog annotation of another gene in this 
-cluster. In the related malaria parasite \textit{P. falciparum} the apical sushi 
-protein, ASP, (which has a sushi domain) localizes to micronemes in merozoites but not 
-other stages (OKeeffe05). 
-Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, 
-(syn. F5/8 domain) domains are associated discoidin lectin domains and thereby with adhesion.
-(Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
-Taken together, this indicates that the thrombospondin and sushi-domain genes 
-(EfaB\_MINUS\_4114.g412 and EfaB\_PLUS\_1425.g183) are involved in sporozoite invasion in 
-\textit{E. falciformis} and that the mRNAs are transcribed and available before excystation. 
-A role in merozoite re-invasion in \textit{E. falciformis} is not indicated by our data.
-The LCCL domain annotation and thrombospondins role in higher eukaryotes also indicates 
-that adhesion or preparation for adhesion is important in oocysts. 
-We suggest (speculate...?) that the thrombospondin annotated ortholog and the LCCL domain-containing protein 
-(EfaB\_MINUS\_11233.g986) are involved in cell adhesion in \textit{E. falciformis}.
-%One examples
-%is found in the slime mold \textit{Dictyostelium discoideum)} in which the LCCL domain
-%is part of the mold's discoidin adhesion protein. (Pfam entry for 'discoidin domain', May 2016)
-
-\subsubsection*{Amino acid biosynthesis in oocysts}
-High abundance of aminotransferase mRNAs indicate amino acid biosynthesis or preparation for 
-the same in oocysts (cluster 4). 
-We identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase orthologs, which 
-are enzymes contributing to L-serine and L-alanine production, respectively. A putative 
-\textit{Eimeria} spp. cystathionine beta-synthase, CBS, in this cluster also indicates 
-de novo cysteine production. Alkyl sulfatase mRNA is also abundant in oocysts. Generally, 
-this enzyme enables an organism to exploit organic sulfur to produce and incorporate 
-inorganic sulfur into the amino acids cysteine and methionine, when no inorganoc sulfur 
-is available.
-
-'Embryonic development'
-Nicalin 1, patched family protein (hedgehog)
-
-\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
-MmgE/PrpD is overrepresented in oocysts. The enzyme is important for propionate catabolism 
-in the 2-methylcitric acid cycle and has been shown to be used by the 
-intestinal intracellular bacterium \textit{Salmonella typhimurium} to generate 
-pyruvate (Horswill99). 
-Propionate is one of two most abundant small-chain fatty acids 
-in the gut along with butyrate. Both fatty acids are largely produced as 
-degradation products from food by commensal bacteria (Sun13). Sharing the intestines as a niche
-with \textit{S. typhimurium} it is possible that also \textit{E. falciformis} uses Mmg/PrpD
-to exploit available propionate for pyruvate production.
-
-\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
-Interestingly, the genes described above which are thought to be involved in amino acid 
-biosynthesis and invasion are highly abundant in oocysts but are underrepresented in schizont
-stages (day 3 and day 5 samples) and even in sporozoites. An average abundance was detected on 
-day 7 for these genes, indicating a role in either gametes or early oocyst formation.
-This pattern supports the suggestion that these specific mRNAs (cluster 4) for invasion 
-and biosynthetic processes are prepared (and possibly expressed) in the oocyst stage but 
-are no longer detectable in the cell at the timepont when the protein is assumed to be 
-in use (sporozoites and merozoite stages). 
-Therefore, correlating mRNA prevalence with biological function at the timepoint when mRNAs 
-are detected must be done with care.
-
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
-Oocysts: profile for 3, 5, 6:
-
-Sporozoites: 3 and 5
-
-Day 7: 5 and 7
-
-Specific genes....
-Enolase 2, encoded by Eno2, is among the downregulated genes in oocysts and sporozoites.
-In \textit{T. gondii} the paralog Eno1 is strongly associated with the cyst (bradyzoite) stage
-and Eno2 is associated with tachyzoite stages (kibe05). It is therefore expected that this mRNA is
-underrepresented in oocysts and our data also show that the same is true in sporozoites for
-\textit{E. falciformis}. (TK: If important we could look specifically for Eno1 in cluster 4).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
-Two clusters contain genes with mRNAs highly abundant on day 7 p.i; cluster 1 and 2.
-Dynein, kinesin and tubulin are annotations highly represented among orthologs of genes 
-in both these clusters. The annotations indicate an important role for motility at this
-timepoint, probably reflecting development of microgametes.
-In addition, in cluster 2, there are two 'EF-hand domain containing proteins' annotations as well 
-as caltractin, centrin-1, and troponin annotations. Caltractin and 
-centrin-1 are associated with the centrosome and structure and function of microtubuli in mammals, 
-and troponin is linked to muscle function (UniProt). 
-Also potentially linked to motility is the occurence of growth arrest specific protein 8,
-Gas8, which in the mouse has been reported to be highly 
-expressed in the testes and important for mouse sperm function (Yeh02).
-
-Other genes among the 38 indicate carbon fixation (glycolysis/gluconeogenesis) or 
-conversions of nucleoside phosphates. In addition, a Ras family protein, RNA 
-polymerase II transcription initiation factor and Sec23 and Sec24 were among orthologs 
-identified in \textit{E. falciformis} cluster 2.
-
-In cluster 1, carbon metabolism genes are represented by 6-phosphogluconate dehydrogenase 
-and glycogen phosphorylase family protein 1. UDP-glucose 4-epimerase and amiloride-sensitive 
-amine oxidase are reported as upregulated in gametocytes in \textit{E. tenella} by 
-RNA-seq (Walker15) and suggested by those authors to play a role in cyst wall synthesis.
-
-\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
-Unintuitively for a protozoan organism, seven out of eight GO biological process terms
-in cluster 1 are associated with wound healing and blood coagulation.
-An explanation is offered by some of the orthologs to the three \textit{E. falciformis} genes
-responsible for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and 
-\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal protein MIC4,
-related' (E. tenella) and more generally for several other protozoa, 'PAN domain containing proteins'.
-The PAN domain is found in the plasminogen/hepatocyte growth factor family and in coagulation 
-factor XI family (REF), explaining why terms related to blood coagulation are enriched by these genes.
-Later publications on \textit{T. gondii} (Marchant12) also associate PAN domains and proteins in 
-apicomplexan parasites with micronemes and therefore invasion. In our case, this is peculiar, since 
-the enrichment appears on day 7 p.i.. A possible role at this timepoint is suggested by work on 
-the fungi \textit{Sclerotinia sclerotiorum} where Yu et al. reported an important role for PAN domain
-proteins in cell wall integrity (Yu12). This role for MIC proteins has to our knowledge 
-not been investigated in apicomplexan parasites. 
-The PAN domain domain has also been reported to be common in nematodes such as 
-\textit{Caenorhabditis elegans}, however the function is not understood. (Thordai99)
-The other two GO terms in the cluster of day seven upregulated genes are DNA replication and DNA 
-replication initiation, which most likely reflects late stage schizogony or gamete formation.
-Six genes contribute to this enrichment and orthologs are either annotated as DNA replcation
-licencing factors, DNA polymerases or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.
-
-\subsubsection*{Gene and sample patterns by hierarchical clustering}
-Samples (columns) cluster into two major clusters where day 7 p.i. samples form one group distinct from
-other samples. In the second group, oocsts and sporozoites are distinct and sporozoites cluster
-most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples also cluster into two groups,
-of which one contains all NMRI day 5 p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
-clusters have no obvious patters.
-
-For gene clusters (rows), the two groups with high mRNA abundance on day 7 p.i. (cluster 1 and 2)
-do not cluster most closely with each other, but with the cluster for high mRNA abundance in oocysts
-(cluster 1 association) and with the cluster for high mRNA abundance in sporozoites (cluster 2 
-association).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
-In our analysis we demonstrate which biological processes are dominant in different
-life cycle stages of \textit{E. falciformis} in the mouse. The RNAseq transcriptome provided here 
-allows for detailed analysis of genes involved in those processes, providing candidates for 
-life stage specific marker in \textit{Eimeria} spp. research.
-
-\ldots
-\ldots
-\ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	METHODS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-\subsection{Mice and infection procedure}
-Three strains of mice were used in our experiments: NMRI (Charles
-River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
-C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
-procedures were performed according to the German Animal Protection
-Laws as directed and approved by the overseeing authority Landesamt
-fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
-as described by Schmid et al., (Schmid12), but tapwater was used
-instead of PBS for administration of oocysts. Briefly, NMRI mice were
-infected two times, which will be referred to as first and second
-infection. For the first infection, 150 sporulated oocysts were
-administered in 100 $μ$L by oral gavage. During the first infection of
-60 mice, all animals were weighed every day. On day zero, before
-infection, as well as on day three, five and seven days post
-infection, dpi, caeca from 3-4 sacrificed mice per time point were
-collected. Epithelial cells were isolated as described in Schmid et
-al.(schmid12). For challenge infection, mice recovered for four weeks
-before second infection.  Recovery was monitored by weighing and
-visual inspection of fur. For the second infection, 1500 sporulated
-oocysts were applied by oral gavage. Three mice were used as
-non-second infection control, referred to as day 0, second infection.
-
-\subsection{Oocyst purification for infection and sequencing}
-Sporulated oocysts were purified by flotation from feces stored in
-potassium dichromate and administered orally in 100 uL tapwater. One
-\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
-Haberkorn 1970, was used for all infections and parasite samples. The
-strain is maintained through passage in NMRI mice in our facilities as
-described elsewhere (schmid12).
-
-\subsection{Sporozoite isolation}
-Sporozoites were isolated from sporocysts by excystment. For this,
-sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
-tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
-30 min. Sporozoites were purified by the method of Schmatz et al
-(schmatz--).
-
-\subsection{RNA extraction}
-Total RNA was isolated from infected epithelial cells, sporozoites and
-sporulated oocysts using Trizol according to the manufacturer’s
-protocol (Invitrogen). High quality \emph{what is the meaning of 'high
-  quality' here?}  RNA was used to produce an mRNA library using the
-Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
-  genome paper} Sporozoites were stored in 1 mL Trizol until
-RNA-isolation.  Total RNA was isolated using the PureLink RNA Mini Kit
-(Invitrogen) and reverse transcribed into cDNA.
-
-\subsection{Sequencing, sequence quality assessment and alignment}
-cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
-specified in SI xx (both unstranded). A fastq\_quality\_filter
-(FASTQ-toolkit, version 0.0.14, available at
-https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
-Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
-phred score of 10 was applied. We further set q = 60. These settings
-require that nine out of ten bases or more are correct in at least
-60\% of the bases for each read.
-
-\subsection{Alignment and reference genomes}
-We used the published \textit{Mus musculus} mm10 assembly (Genome
-Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
-genome including annotations for mouse data. The
-\textit{E. falciformis} genome (Heitlinger14) was downloaded from
-ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
-files were merged into a dual reference genome, and files including
-mRNA sequences from both species were aligned against the dual
-reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
-specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
-genome. Single-end and pair-end sequence samples were aligned
-separately with library type 'fr-unstranded' specified for pair-end
-samples. Import into R was enabled by the R package Ballgown, which
-requires bam files to be processed by Tablemaker (Frazee15), in our
-case used with -qW -G specified. Tablemaker in turn makes use of
-Cufflinks (version 2.1.1, Trapnell10).
-
-\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
-Count data was normalized using the R-package edgeR (version 3.14.0;
-cite) with the upperquartile normalisation method. Briefly, genes with
-zero coverage in all samples (libraries) are removed and normalisation
-factors are calculated for the 75\% quantile for each library. This
-normalisation is suitable for read densities following a negative
-binomial distribution. Two samples contradicted this assumption
-(parasite data) for later modelling and both mouse and parasite data
-from these samples were excluded from further analysis:
-NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
-then fits a generalized linear model (GLM with a negative binomial
-link function) for each gene (glmFit) and then performs likelihood
-ratio tests for models w or w/o focal factor (glmLRT).
-
-%% ........... include outcome of MDS analysis here?
-%% No, results never are presented in methods.
-
-\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
-A selection of differently abundant mRNAs are used for hierarchical
-clustering of \textit{E. falciformis} life cycle relevant genes. In
-each comparison (see Table 3), the union of the at most 500 genes
-differentially abundant with lowest FDR (<0.05) are selected. In the
-next step, the 500 mRNAs from each comparison (or less) are
-joined. For \textit{E. falciformis} time p.i. comparison, this
-resulted in 1618 unique genes selected as differently expressed . The
-22 genes in the NMRI vs C57BL/6 comparison are not included in the
-\textit{E. falciformis} life cycle analysis. The same comparisons for
-mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
-all samples, i.e. also samples which themselves did not have any
-significantly different mRNAs according to our selection, were
-included in hierarchical clustering. Scale bar in heatmaps show 0 as
-mean mRNA abundance for each gene (row). Up (green) and
-down-regulation (brown) denote number of standard deviations from 0,
-i.e., row mean. Hierarchical clustering was performed using with
-Euclidean distances, using complete linkage ('complete', R package
-base).
-
-%% the paragraph above has results. All those number
-
-\subsection{Differentially abundant mRNAs used for hierarchical clustering}
-All analyses were performed in R (cite R-core). Complete scripts are
-available at git@github.com:derele/Ef\_RNAseq.git tagged as version
-1.0. 
-%% We will make this available as a git webpage probably
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}[H]
-\small
-\begin{center}
-\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
-\begin{tabular}{*3l}    \toprule
-	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
-	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
-	(including oocysts and sporozoites if appl.)	\\	
-	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
-\hline
-\end{tabular}
-\end{center}
-%\end{adjustwidth}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{Competing interests}
-  The authors declare that they have no competing interests.
-
-\section*{Author's contributions}
-    Text for this section \ldots
-
-\section*{Acknowledgements}
-  Text for this section \ldots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
-% for author-year bibliography (bmc-mathphys or spbasic)
-% a) write to bib file (bmc-mathphys only)
-% @settings{label, options="nameyear"}
-% b) uncomment next line
-%\nocite{label}
-
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
-
-%\begin{figure}
- %\begin{picture}
- % \includegraphics[width=0.5\linewidth]{ef-heatmap}
- %\end{picture}
-%\end{figure}
-
-\section*{Figures}
-\begin{figure}[h!]
-%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
-  \caption{\csentence{Parasite read sequences per mouse read sequences for each experimental condition.}
-  The more advanced the infection is, the higher the ratio of parasite reads is, reflecting parasite replication between day zero and day seven.
-......compare 1st and 2nd .......
-patterns between Rag and NMRI/C57BL/6?
-........
-Values are mean of replicates (n=2, if * n=1) on log10 scale.
-Each sample (replicate) consists of mRNA from three different mice.}
-  \end{figure}
-
- % \begin{figure}[h!]
- % \includegraphics[width=0.7\linewidth]{ef-heatmap}
- % \caption{\csentence{Parasite genes with different mRNA abundance between samples, sorted into five clusters by hierarchical clustering (method = complete, distance = euclidean).}
- %     \textit{E.
-falciformis} samples from seven days p.i.
-cluster together (NMRI mice only).
-These samples have distinct mRNA abundance patterns in all gene clusters, although more pronounced in clusters 1 (up) and 4 (down).
-Distinct groups of genes also define sporozoites (cluster 5, up) and oocysts (cluster 2, up).
-mRNA profiles on days three and five p.i.
-from all three mouse strains cluster together.
-These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites, however closest to the latter.
-On scale bar, 0 is mean mRNA abundance for each gene.
-Up and downregulation is standard deviations from mean.}
- %     \end{figure}
-
-%\begin{figure}[h!]
-%  \includegraphics[width=0.7\linewidth]{mm-heatmap}
-%  \caption{\csentence{Mouse genes with different mRNA abundance between samples, sorted into four clusters by hierarchical clustering (method = complete, distance = euclidean).}
-%Three out of four samples from day 7 p.i.
-(NMRI only) cluster together.
-These samples are characterized by genes in cluster 2 (up), 3, and 4 (down).
-NMRI and C57BL/6 uninfected samples cluster together (left), defined by clusters 3, 4 (up), and 1 (down).
-Rag1-/- samples cluster together.
-These samples share a weak downregulation of most genes in cluster 2 as well as upregulation of a small group genes in the same cluster.
-Uninfected Rag1-/- samples are separated from infected ones with distinct profiles in clusters 1, 3, and 4.
-On scale bar, 0 is mean mRNA abundance for each gene.
-Up and downregulation is standard deviations from mean.}
-%      \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
-\section*{Tables}
-\begin{table}[h!]
-\caption{Sample table title. This is where the description of the table should go.}
-      \begin{tabular}{cccc}
-        \hline
-           & B1  &B2   & B3\\ \hline
-        A1 & 0.1 & 0.2 & 0.3\\
-        A2 & ... & ..  & .\\
-        A3 & ..  & .   & .\\ \hline
-      \end{tabular}
-\end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \subsection*{Additional file 1 --- Raw and normalized counts }
-  Raw and normalized counts for individual samples in our
-  study (two compressed csv files).
+  Raw counts of reads mappins to the \textit{E. falciformis} and mouse
+  genome for individual samples in our study. Normalized counts for
+  seperately for the host and parasite mappings (three compressed csv
+  files).
 

### 2016-08-08 figuresANDmanuscript/table_rc.tex
Manuscript changes while working on feature counts integration and simple csv output to be handed over between scripts

--- 2016-07-14 figuresANDmanuscript/table_rc.tex
@@
-% latex table generated in R 3.3.0 by xtable 1.8-2 package
-% Wed Jun 22 17:40:30 2016
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Mon Aug  8 11:22:42 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}{lllllllll}
   \hline
-Sample & Mouse transcripts & E. falciformis transcripts & Percentage E. falciformis & \#E. falciformis genes \\ 
+Sample & Sequencing method & batch & read mappings Mouse & read mappings E. falciformis & Percentage E. falciformis & dpi & challenged & \# E. falciformis genes \\ 
   \hline
-NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
-  NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
-  NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
-  NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep1 & 589643389.000 & 3752923.000 & 0.632 & 5602.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
-  C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep1 & 209723287.000 & 815820.000 & 0.388 & 5466.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
-  Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
-  Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
-  Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
-  NMRI\_1stInf\_0dpi\_rep1 & 229165002.000 & 31459.000 & 0.014 & 1380.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
-  Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
-  Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
+NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 70489674.000 & 247.000 & 0.000 & 0dpi & Callenge & 1.000 \\ 
+  Rag\_0dpi\_rep1 & hiseq & 3 & 18853850.000 & 443.000 & 0.002 & 0dpi & None & 2.000 \\ 
+  C57BL6\_0dpi\_rep1 & hiseq & 3 & 25119348.000 & 457.000 & 0.002 & 0dpi & None & 2.000 \\ 
+  C57BL6\_0dpi\_rep2 & hiseq & 3 & 34377133.000 & 608.000 & 0.002 & 0dpi & None & 2.000 \\ 
+  Rag\_0dpi\_rep2 & hiseq & 3 & 35233327.000 & 676.000 & 0.002 & 0dpi & None & 2.000 \\ 
+  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 40794245.000 & 3406.000 & 0.008 & 0dpi & Callenge & 51.000 \\ 
+  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 40544287.000 & 4803.000 & 0.012 & 3dpi & Callenge & 95.000 \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 48289181.000 & 10941.000 & 0.023 & 5dpi & Callenge & 407.000 \\ 
+  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 55176243.000 & 17954.000 & 0.033 & 0dpi & None & 701.000 \\ 
+  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 46171909.000 & 29548.000 & 0.064 & 3dpi & Callenge & 1580.000 \\ 
+  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 51722265.000 & 40091.000 & 0.077 & 7dpi & Callenge & 1836.000 \\ 
+  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 29982453.000 & 63024.000 & 0.210 & 5dpi & First & 2548.000 \\ 
+  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 25297803.000 & 99000.000 & 0.390 & 5dpi & First & 2828.000 \\ 
+  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 29319604.000 & 185969.000 & 0.630 & 5dpi & First & 4173.000 \\ 
+  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 37093621.000 & 192856.000 & 0.517 & 5dpi & Callenge & 4167.000 \\ 
+  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 18062349.000 & 215696.000 & 1.180 & 5dpi & First & 3823.000 \\ 
+  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 25660184.000 & 262909.000 & 1.014 & 5dpi & Callenge & 4563.000 \\ 
+  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 49993358.000 & 394384.000 & 0.783 & 3dpi & First & 5220.000 \\ 
+  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 117791044.000 & 413051.000 & 0.349 & 3dpi & First & 4862.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 76419774.000 & 794570.000 & 1.029 & 5dpi & First & 5333.000 \\ 
+  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 143186486.000 & 1846840.000 & 1.273 & 5dpi & Callenge & 5533.000 \\ 
+  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 41467146.000 & 8634201.000 & 17.233 & 7dpi & Callenge & 5875.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 47037087.000 & 8669701.000 & 15.563 & 5dpi & First & 5700.000 \\ 
+  NMRI\_sporozoites\_rep2 & GAII & 0 & 8656.000 & 11470604.000 & 99.925 & environmental & environmental & 5513.000 \\ 
+  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 83735624.000 & 27839458.000 & 24.951 & 5dpi & First & 5784.000 \\ 
+  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 3310666.000 & 39400884.000 & 92.249 & 7dpi & First & 5932.000 \\ 
+  NMRI\_sporozoites\_rep1 & GAII & 1 & 4334.000 & 43774401.000 & 99.990 & environmental & environmental & 5825.000 \\ 
+  NMRI\_oocysts\_rep1 & GAII & 1 & 3805.000 & 49653065.000 & 99.992 & environmental & environmental & 5695.000 \\ 
+  NMRI\_oocysts\_rep2 & GAII & 0 & 18524.000 & 71019860.000 & 99.974 & environmental & environmental & 5777.000 \\ 
+  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 21699324.000 & 73539445.000 & 77.216 & 7dpi & First & 5943.000 \\ 
    \hline

### 2016-08-08 4_annotations.R
Annotation working without outside data

@@ -43,6 +43,13 @@ read.omcl <- function(path="data/Omcl_groups.txt"){
 
 omcl <- read.omcl()
 
+get.genes.4.clusters <- function (clusters) {
+  gsub("Efa\\|", "",   grep("Efa\\|",
+                            unlist(clusters),
+                            value=TRUE))
+}
+
+
 get.omcl.categories <- function(ortho.obj){
   ## Orthomcl
   get.presence.omcl <- function(ortho.obj){

### 2016-08-08 5_compareArray.R
Array comparison fixed

@@ -96,16 +96,13 @@ Array.logFC <- topTable(fit2.g, n=400000)[, c("X24h", "X144h")]
 ##                                                          "logFC.R5vs0" )])
 
 ## RuvSeq at the moment not tested
-RNAseq.logFC.ruved <- as.data.frame(Mm.RUVg.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
-                                                           "logFC.N7vs0", "logFC.B5vs0",
-
-## RuvSeq at the moment not tested
 ## RNAseq.logFC <- merge(RNAseq.logFC, RNAseq.logFC.ruved, by = 0)
 
 Array.logFC <- merge(annot.frame[,1:3], Array.logFC, 
                      by.x = "symbol", by.y = 0)
 
-RNAseq.Array.logFC <- merge(Array.logFC[,c("ensembl_id", "X24h", "X144h")], Mm.DE.FC, 
+RNAseq.Array.logFC <- merge(Array.logFC[,c("ensembl_id", "X24h", "X144h")],
+                            Mm.DE.FC, 
                             by.x = "ensembl_id", by.y = "gene")
 
 ## RuvSeq at the moment not tested

### 2016-08-08 2_data_curation.R
New order and first add of work on Simones data

@@ -85,7 +85,7 @@ RC.table$dpi[RC.table$challenged%in%"environmental"] <- "environmental"
 
 ## plotting parasite percentage of reads
 
-pdf("figuresANDmanuscript/Ef.percentage.pdf")
+pdf("figures/Figure1b_Ef.percentage.pdf")
 ggplot(RC.table, aes(x = dpi, y = p.Ef.reads,
                      color=challenged, shape=strain)) +
     geom_jitter(size=4, width=0.35, alpha=0.6)+
@@ -149,7 +149,7 @@ density.plots <-
                                names(r.c.s.l)[[i]]))
            })
 
-pdf("figuresANDmanuscript/distributionsMm.pdf", width = 27, height = 21)
+pdf("Supplement/distributionsMm.pdf", width = 27, height = 21)
 do.call(grid.arrange, c(density.plots, list(nrow=2)))
 dev.off()
 
@@ -192,7 +192,7 @@ density.plots <-
                                names(r.c.s.l)[[i]]))
            })
                
-pdf("figuresANDmanuscript/distributionsEf.pdf", width = 27, height = 21)
+pdf("Supplement/distributionsEf.pdf", width = 27, height = 21)
 do.call(grid.arrange, c(density.plots, list(nrow=2)))
 dev.off()
 
@@ -220,7 +220,7 @@ names(table.cleaned) <- sub("^c.Ef.genes$", "# E. falciformis genes", names(tabl
 
 ## EXPORT to Latex format
 table.tex <- xtable(table.cleaned, align = c("l", "l", "l", "l", "l", "l", "l", "l", "l", "l"), digits = 3)
-print(table.tex, type = "latex", file = "figuresANDmanuscript/table_rc.tex", include.rownames = F)
+print(table.tex, type = "latex", file = "tables/Table1_ReadCounts.tex", include.rownames = F)
 
 
 

### 2016-08-08 3_edgeR_diff.R

@@ -219,7 +219,6 @@ cbind(melt(lapply(Ef.1st.pass.model[[3]], length)),
       melt(lapply(Ef.RUVg.model[[3]], length)), by = "L1")[,c(1,3,4)]
 
 ## test for consistency with suppl file data:
-
 by(Mm.DE.test, Mm.DE.test$contrast, function (x) table(x$FDR<0.01))
 by(Ef.DE.test, Ef.DE.test$contrast, function (x) table(x$FDR<0.01))
 

### 2016-08-08 5_compareArray.R

@@ -25,14 +25,14 @@ targets <- readTargets("data/targets.txt", sep=";")
 
 RG <- read.maimages(targets, path= "/SAN/Mouse_arrays/2012_arrays/", source = "agilent", 
                     other.columns = list(IsFound = "gIsFound",
-                      IsWellAboveBG = "gIsWellAboveBG", 
-                      IsSaturated = "gIsSaturated",
-                      IsFeatNonUnifOF = "gIsFeatNonUnifOL", 
-                      IsFeatPopnOL = "gIsFeatPopnOL",
-                      ChrCoord = "chr_coord"), 
+                                         IsWellAboveBG = "gIsWellAboveBG", 
+                                         IsSaturated = "gIsSaturated",
+                                         IsFeatNonUnifOF = "gIsFeatNonUnifOL", 
+                                         IsFeatPopnOL = "gIsFeatPopnOL",
+                                         ChrCoord = "chr_coord"), 
                     columns = list(G = "gMedianSignal", Gb = "gBGMedianSignal",
-                      R = "rMedianSignal",
-                      Rb = "rBGMedianSignal"), verbose = T, 
+                                   R = "rMedianSignal",
+                                   Rb = "rBGMedianSignal"), verbose = T, 
                     sep = "\t", quote = "")
 
 array.names <- apply(targets, 1, function (x) paste(x[c(2, 4,5)], collapse="."))
@@ -91,13 +91,6 @@ fit2.g <- eBayes(fit2.g)
 
 Array.logFC <- topTable(fit2.g, n=400000)[, c("X24h", "X144h")]
 
-## RNAseq.logFC <- as.data.frame(Mm.1st.pass.model[[1]][, c("logFC.N3vs0", "logFC.N5vs0",
-##                                                          "logFC.N7vs0", "logFC.B5vs0",
-##                                                          "logFC.R5vs0" )])
-
-## RuvSeq at the moment not tested
-## RNAseq.logFC <- merge(RNAseq.logFC, RNAseq.logFC.ruved, by = 0)
-
 Array.logFC <- merge(annot.frame[,1:3], Array.logFC, 
                      by.x = "symbol", by.y = 0)
 
[…]

### 2016-08-08 RUVseq.R

@@ -1,99 +1,0 @@
-library(RUVSeq)
-
-## We now need a "DGEList" containing edgeR normalized (default TMM)
-## data
-if(!exists("Mm.RC")){
-    source("2_edgeR_diff.R")
-}
-
-Mm.differences <- makeGroups(pData(Mm.bg)$group)
-
-## use the expression set and the filtering decided on before
-Mm.RUVset <- newSeqExpressionSet(Mm.RC[[3]][keep,])
-
-pdf("figures/ruv_eval/RLE_wo_NORM.pdf")
-plotRLE(Mm.RUVset)
-dev.off()
-
-pdf("figures/ruv_eval/PCA_wo_NORM.pdf")
-plotPCA(Mm.RUVset)
-dev.off()
-
-pdf("figures/ruv_eval/mds_wo_NORM.pdf")
-plotMDS(counts(Mm.RUVset))
-dev.off()
-
-pdf("figures/ruv_eval/hclust_wo_NORM.pdf")
-plot(hclust(dist(t(counts(Mm.RUVset)))))
-dev.off()
-
-Mm.RUVset.BLnorm <- betweenLaneNormalization(Mm.RUVset, which="upper")
-
-pdf("figures/ruv_eval/RLE_BL_NORM.pdf")
-plotRLE(Mm.RUVset.BLnorm)
-dev.off()
-
-pdf("figures/ruv_eval/PCA_BL_NORM.pdf")
-plotPCA(Mm.RUVset.BLnorm)
-dev.off()
-
-pdf("figures/ruv_eval/mds_BL_NORM.pdf")
-plotMDS(counts(Mm.RUVset.BLnorm))
-dev.off()
-
-pdf("figures/ruv_eval/hclust_BL_NORM.pdf")
-plot(hclust(dist(t(normCounts(Mm.RUVset.BLnorm)))))
-dev.off()
-
-Mm.RUVset.groups <- RUVs(Mm.RUVset.BLnorm,
-                         rownames(Mm.RUVset.BLnorm), k=1, Mm.differences)
-
-pdf("figures/ruv_eval/RLE_group_NORM.pdf") ## not very telling?!
-plotRLE(Mm.RUVset.groups)
-dev.off()
-
-pdf("figures/ruv_eval/PCA_group_NORM.pdf") ## wow that seems to work a treat
-plotPCA(Mm.RUVset.groups)
-dev.off()
-
[…]

### 2016-08-08 distribution_check.R

@@ -1,126 +1,0 @@
-### Check distributions of transcript data
-
-###########################
-library(reshape2)
-############## MOUSE ##################
-## rearrange GM.E for ggplot format
-long.Mcounts <- melt(Mm.RC[[3]])
-#long.counts$strain <- gsub("_.*", "", long.Mcounts$Var2)
-#long.Mcounts$strain <- gsub("_.*", "", 
-#		long.Mcounts$Var2)
-
-long.normMcounts <- melt(cpm(Mm.RC[[3]]))
-#long.normMcounts$strain <- gsub("_.*", "", 
-#		long.normMcounts$Var2)
-
-## RAW MOUSE DISTRIBUTION
-pdf("figures/distributions_mouse5000.pdf", 
-	width = 12, height = 7)
-ggplot(long.Mcounts, 
-	aes(value)) + 
-	stat_density(geom="line") +
-	scale_x_log10("Read counts (log10)",
-			labels = trans_format("log10",
-			math_format(10^.x))) +
-	facet_wrap(~Var2) +
-	ggtitle("Mouse, non-norm., cutoff = 5000 reads") +
-	ylim(0, 0.6)
-dev.off()
-
-## NORMALISED MOUSE DISTRIBUTION
-pdf("figures/distributions_normmouse5000.pdf", 
-	width = 12, height = 7)
-ggplot(long.normMcounts, 
-	aes(value)) + 
-	stat_density(geom="line") +
-	scale_x_log10("Read counts (log10)",
-			labels = trans_format("log10",
-			math_format(10^.x))) +
-	facet_wrap(~Var2) +
-	ggtitle("Mouse, norm., cutoff = 5000 reads") +
-	ylim(0, 0.6)
-dev.off()
-
-############## PARASITE ###################
-long.Ecounts <- melt(GM.E$counts)
-#long.Ecounts$strain <- gsub("_.*", "", long.Ecounts$Var2)
-#long.Ecounts$strain <- gsub("_.*", "", 
-#		long.Ecounts$Var2)
-
-long.normEcounts <- melt(cpm(GM.E))
-#long.normEcounts$strain <- gsub("_.*", "", 
-#		long.normEcounts$Var2)
-
-excludedE <- long.normEcounts[long.normEcounts$Var2%in%
-		c("NMRI_2ndInf_7dpi_rep2",
-		"NMRI_2ndInf_5dpi_rep2",
-		"NMRI_2ndInf_3dpi_rep1",
-		"NMRI_2ndInf_3dpi_rep2"),]
[…]

### 2016-08-12 manuscript/bmc_article.tex
Streamlined pData and RC-summary table -> Table1

--- 2016-08-08 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
@@
 %%% Load packages
-%\usepackage{amsthm,amsmath}
-%\RequirePackage{natbib}
+\usepackage{amsthm,amsmath}
+\RequirePackage{natbib}
 %\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-%\RequirePackage{hyperref}
+\RequirePackage{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
@@
 \usepackage{graphicx}
+\usepackage[flushleft]{threeparttable}
 
@@
 
-\title{Dual host-parasite transcriptomes of apicomplexan Eimeria falciformis and its natural mouse host}
+\title{Dual host-parasite transcriptomes of apicomplexan Eimeria
+  falciformis and its natural mouse host}
 
@@
 
-\author[ <addressref>={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
-\author[ <addressref>={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
-\author[ <addressref>={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
-\author[ <addressref>={aff1},]{\inits{RL}\fnm{Richard} \snm{Lucius}}
-\author[ <addressref>={aff1, aff3},              
+\author[ addressref={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
+\author[ addressref={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
+\author[ addressref={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
+\author[ addressref={aff1}, ]{\inits{RL}\fnm{Richard} \snm{Lucius}}
+\author[ addressref={aff1, aff3},              
   corref={aff1},                   
@@
 
-Briefly RAG-1-/- knockout mice, in which mature B and T lymphocytes
-are absent showed ...
-
-Additonally we obtained data for challenge infections of mice
-immunized one month prior to the experiment.
+In these experiments we analysed RAG-1-/- mice and compared them to
+the parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
+response du to the absence of mature B and T lymphocytes. Differences
+in oocyst output were more pronounced between NMRI and both C57BL6 and
+RAG-1-/-. Surprisingly this finding was made both in primary and
+challenge infection. Similarly development of intestinal stages as
+assayed by quantitative reverse transcription PCR (RT qPCR) also ...
+
+This lead us to question the influence of the innate or cell
+autonomouse immune system on early stages of infection. And we thus
+settled on an experimental design comparing infections at 5dpi for all
 
@@
 
-\subsection{Life cycle analysis... \textit{E. falciformis}}
 \begin{figure}[h!]
@@
 %\centering
-	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
+%%	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
 	\caption{\textit{E. falciformis} mRNAs with significantly
@@
 %\centering
-\includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
+%% \includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
 \caption{Mouse mRNAs with significantly different abundance at
@@
 
+% Fri Aug 12 11:05:43 2016
 \begin{table}[ht]
-\caption{Experimental design, sequencing output and mapping statistics.}
 \centering
-\hspace*{-1.5cm}\begin{tabular}{lllllllll}
+\hspace*{-2.5cm}\begin{tabular}{lllllllll}
   \hline
-Sample & \bcell{Sequencing\\method} & batch & \bcell{read mappings\\Mouse} & \bcell{read mappings\\\textit{E. falciformis}} & \bcell{Percentage\\\textit{E. falciformis}} & dpi & infection  \\ 
+Sample* & \bcell{Sequencing\\method} & batch & \bcell{total\\reads} & \bcell{reads\\mapping\\Mouse} & \bcell{reads\\mapping\\\textit{E.
+falciformis}} & \bcell{Percentage\\\textit{E.
+falciformis}}** & \bcell{detected\\ \textit{E.
+falciformis}\\genes} \\
   \hline
-  NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 285032128 & 326 & 0 & 0dpi & Challenge \\ 
-  Rag\_0dpi\_rep1 & hiseq & 3 & 43414004 & 474 & 0.001 & 0dpi & None \\ 
-  C57BL6\_0dpi\_rep1 & hiseq & 3 & 53877840 & 491 & 0.001 & 0dpi & None \\ 
-  C57BL6\_0dpi\_rep2 & hiseq & 3 & 76753491 & 657 & 0.001 & 0dpi & None \\ 
-  Rag\_0dpi\_rep2 & hiseq & 3 & 80702547 & 730 & 0.001 & 0dpi & None \\ 
-  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 90681785 & 3738 & 0.004 & 0dpi & Challenge \\ 
+NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
+  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
+  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
+  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
+  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
+  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
   \rowcolor{LightCyan}
-  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 91352242 & 5355 & 0.006 & 3dpi & Challenge \\ 
+  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
   \rowcolor{LightCyan}
-  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 113957667 & 12692 & 0.011 & 5dpi & Challenge \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
   \rowcolor{LightRed}
-  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 229165002 & 31459 & 0.014 & 0dpi & None \\ 
-  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 106446839 & 34699 & 0.033 & 3dpi & Challenge \\ 
-  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 122999109 & 46700 & 0.038 & 7dpi & Challenge \\ 
-  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 71173382 & 73445 & 0.103 & 5dpi & First \\ 
-  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 57192380 & 113673 & 0.198 & 5dpi & First \\ 
-  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 65523435 & 217606 & 0.331 & 5dpi & First \\ 
-  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 85288323 & 224273 & 0.262 & 5dpi & Challenge \\ 
-  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 32887009 & 250954 & 0.757 & 5dpi & First \\ 
-  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 62053975 & 311900 & 0.500 & 5dpi & Challenge \\ 
-  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 209723287 & 815820 & 0.388 & 3dpi & First \\ 
-  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 515516785 & 903330 & 0.175 & 3dpi & First \\ 
-  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 316721928 & 1609009 & 0.505 & 5dpi & Firs \\ 
-  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 589643389 & 3752923 & 0.632 & 5dpi & Challenge \\ 
-  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 97221189 & 9927803 & 9.265 & 7dpi & Challenge \\ 
-  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 204647381 & 18549727 & 8.311 & 5dpi & First \\ 
-  NMRI\_sporozoites\_rep2 & GAII & 0 & 8702 & 21508353 & 99.960 & \textit{in vitro} & \textit{in vitro} \\ 
-  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 334671421 & 54022504 & 13.899 & 5dpi & First \\ 
-  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 12532238 & 79648900 & 86.405 & 7dpi & First \\ 
-  NMRI\_sporozoites\_rep1 & GAII & 1 & 13800 & 92259539 & 99.985 & \textit{in vitro} & \textit{in vitro} \\ 
-  NMRI\_oocysts\_rep1 & GAII & 1 & 11676 & 108477484 & 99.989 & envir. & envir. \\ 
-  NMRI\_oocysts\_rep2 & GAII & 0 & 19024 & 126543533 & 99.985 & envir. & envir. \\ 
-  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 94310278 & 154343046 & 62.072 & 7dpi & First \\ 
+  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
+  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
+  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
+  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
+  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
+  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
+  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
+  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
+  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
+  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
+  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
+  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
+  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
+  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
+  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
+  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
+  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
+  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
+  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
+  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
+  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
+  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
    \hline
-\hspace*{+1.5cm}\end{tabular}
-\begin{tablenotes}
-  \small
-\item *envir. ``environmental'' oocysts sporulated in potassium dichromate. \\
-\item *\textit{in vitro} in vitro sporulated oocysts
+\end{tabular}
+\begin{tablenotes}[flushleft]\footnotesize\singlespacing
+\item{*} sample names are given as a) mouse strain b) first or challenge
+  infection c) days post infection (dpi) and d) replicate number
+  seperated by undersocre . \\
+\item{**} percentag mapping \textit{E. falciformis} is given as percentage in total mapping reads
 \end{tablenotes}
 \end{table}
+\hspace*{+2.5cm}
 

### 2016-08-12 tables/Table1_ReadCounts.tex
New order and first add of work on Simones data
Streamlined pData and RC-summary table -> Table1

--- 2016-07-14 figuresANDmanuscript/table_rc.tex
@@
-% latex table generated in R 3.3.0 by xtable 1.8-2 package
-% Wed Jun 22 17:40:30 2016
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Fri Aug 12 11:39:34 2016
 \begin{table}[ht]
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}{llllllll}
   \hline
-Sample & Mouse transcripts & E. falciformis transcripts & Percentage E. falciformis & \#E. falciformis genes \\ 
+Sample & Sequencing method & batch & total reads & reads mapping Mouse & reads mapping E. falciformis & Percentage E. falciformis & \# E. falciformis genes \\ 
   \hline
-NMRI\_oocysts\_rep1 & 11676.000 & 108477484.000 & 99.989 & 5734.000 \\ 
-  NMRI\_oocysts\_rep2 & 19024.000 & 126543533.000 & 99.985 & 5774.000 \\ 
-  NMRI\_sporozoites\_rep1 & 13800.000 & 92259539.000 & 99.985 & 5808.000 \\ 
-  NMRI\_sporozoites\_rep2 & 8702.000 & 21508353.000 & 99.960 & 5564.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep1 & 12532238.000 & 79648900.000 & 86.405 & 5894.000 \\ 
-  NMRI\_1stInf\_7dpi\_rep2 & 94310278.000 & 154343046.000 & 62.072 & 5897.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep3 & 334671421.000 & 54022504.000 & 13.899 & 5794.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep1 & 97221189.000 & 9927803.000 & 9.265 & 5865.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep1 & 204647381.000 & 18549727.000 & 8.311 & 5739.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep2 & 32887009.000 & 250954.000 & 0.757 & 3946.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep1 & 589643389.000 & 3752923.000 & 0.632 & 5602.000 \\ 
-  NMRI\_1stInf\_5dpi\_rep2 & 316721928.000 & 1609009.000 & 0.505 & 5439.000 \\ 
-  C57BL6\_2ndInf\_5dpi\_rep1 & 62053975.000 & 311900.000 & 0.500 & 4610.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep1 & 209723287.000 & 815820.000 & 0.388 & 5466.000 \\ 
-  C57BL6\_1stInf\_5dpi\_rep1 & 65523435.000 & 217606.000 & 0.331 & 4259.000 \\ 
-  Rag\_2ndInf\_5dpi\_rep1 & 85288323.000 & 224273.000 & 0.262 & 4251.000 \\ 
-  Rag\_1stInf\_5dpi\_rep2 & 57192380.000 & 113673.000 & 0.198 & 2969.000 \\ 
-  NMRI\_1stInf\_3dpi\_rep2 & 515516785.000 & 903330.000 & 0.175 & 5101.000 \\ 
-  Rag\_1stInf\_5dpi\_rep1 & 71173382.000 & 73445.000 & 0.103 & 2748.000 \\ 
-  NMRI\_2ndInf\_7dpi\_rep2 & 122999109.000 & 46700.000 & 0.038 & 2174.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep2 & 106446839.000 & 34699.000 & 0.033 & 1901.000 \\ 
-  NMRI\_1stInf\_0dpi\_rep1 & 229165002.000 & 31459.000 & 0.014 & 1380.000 \\ 
-  NMRI\_2ndInf\_5dpi\_rep2 & 113957667.000 & 12692.000 & 0.011 & 539.000 \\ 
-  NMRI\_2ndInf\_3dpi\_rep1 & 91352242.000 & 5355.000 & 0.006 & 121.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep2 & 90681785.000 & 3738.000 & 0.004 & 62.000 \\ 
-  Rag\_0dpi\_rep1 & 43414004.000 & 474.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep1 & 53877840.000 & 491.000 & 0.001 & 2.000 \\ 
-  C57BL6\_0dpi\_rep2 & 76753491.000 & 657.000 & 0.001 & 2.000 \\ 
-  Rag\_0dpi\_rep2 & 80702547.000 & 730.000 & 0.001 & 3.000 \\ 
-  NMRI\_2ndInf\_0dpi\_rep1 & 285032128.000 & 326.000 & 0.000 & 2.000 \\ 
+NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
+  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
+  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
+  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
+  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
+  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
+  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
+  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
+  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
+  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
+  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
+  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
+  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
+  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
+  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
+  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
+  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
+  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
+  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
+  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
+  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
+  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
+  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
+  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
+  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
+  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
+  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
+  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
+  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
    \hline

### 2016-08-12 1_feature_count.R

@@ -14,40 +14,25 @@ FC <- featureCounts(bams,
                     isPairedEnd=TRUE,
                     nthreads=20)
 
-## the experimental meta-data as in the pData slot of all ballgown object
-create.pdata <- function (featurecountsTarget){
-    samples <- gsub(".*?tophat_March_c\\.(.*?_rep\\d).*", "\\1", 
-                   featurecountsTarget)
-    grouped <- sub("_rep\\d+$" , "", samples)
-    infection <- ifelse(grepl("1stInf", grouped), "1st",
-                        ifelse(grepl("2ndInf", grouped), "2nd", "NULLth"))
-    rep <- sub(".*_(rep\\d+)$" , "\\1", samples) # (name here) is taken "\\1" here
-    mouse.strain <- sub("^(.*?)_.*", "\\1", grouped)
-    timepoint <- as.numeric(as.character(sub("^.*_(.*)dpi", "\\1", grouped)))
-    ## variables to summarize other variables sith mutliple levels into two levels ##
-    immune.status <- ifelse(mouse.strain %in% "Rag", "Rag", "competent")
-    late.early <- ifelse(timepoint < 7 , "early", "late")
-    batch <- c(3, 3, 3, 3, 3, 1, 1, 2, 2, 2, 0, 1, 1, 2,
-               3, 3, 3, 2, 3, 3, 3, 1, 0, 1, 0, 3, 3, 3, 3, 3)
-    seq.method <- c("hiseq", "hiseq", "hiseq", "hiseq", "hiseq",
-               "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII", "GAII",
-               "hiseq", "hiseq", "hiseq",
-               "GAII",
-               "hiseq", "hiseq", "hiseq",
-               "GAII", "GAII", "GAII", "GAII",
-               "hiseq", "hiseq", "hiseq", "hiseq", "hiseq")
-    sample.seq <- paste(samples,seq.method)
-    data.frame(samples, grouped, infection, rep, mouse.strain,
-               timepoint, immune.status, late.early, batch, seq.method, sample.seq)
-    
-}
 
-pData <- create.pdata(FC$targets)
-
-FC$targets <- gsub(".*?tophat_March_c\\.(.*?_rep\\d).*", "\\1", FC$targets)
[…]

### 2016-08-12 2_data_curation.R

@@ -1,52 +1,42 @@
-## This is a script is 2nd the pipeline and needed for evaluation
-## purpose
-
-## Goals:
-
-## EX- on IN-clusion of samples and genes ###################
-
-## a) decide for samples needed to be excluede because of uncertain
-## infection status
-
-## b) decice for a filter on minimal expression values over all
-## samples to exclude genes
-
-## c) decide on samples needed to be excluded because of bad
-## "unnormalizable" distribution of expression values
-
-
-### Normalization
-
-## d) decide on a general normalization strategy scaling the smaple
-## throughput
-
-## e) decide on a "removeal of unwanted variation" strategy
-
-## f) decide on a potential addional "batch effect" removal strategy
-
-
-## a) decide for samples needed to be excluede because of uncertain
-## infection status
-
-## The raw counts  required from upstream scripts:
-## GENES
-## All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
-
-## TRANSCRIPTS
-All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
-
-## The phenotype information
-All.pData <- read.table("output_data/Sample_pData.csv", sep=",")
-
 library(ggplot2)
 library(reshape)
 library(gridExtra)
 library(xtable)
 
-## Overview for differential representation of host/parasite
-RC.table <-  as.data.frame(cbind(sample=as.character(All.pData$samples),
-                                 seq.method=as.character(All.pData$seq.method),
-                                 batch=All.pData$batch))
+## a) decide for samples needed to be excluede because of uncertain
+## infection status
+
+## b) create Table one summarizing metadata (previously pData) and
+## basic sequencing statistics read counts
+
+## Raw counst for tenes
+All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
+
+create.pdata <- function (featurecountsTarget){
+    sample <- featurecountsTarget
+    grouped <- sub("_rep\\d+$" , "", sample)
[…]

### 2016-08-12 3_edgeR_diff.R

@@ -15,18 +15,18 @@ Mm.RC <- All.RC[grepl('^ENSMUS.*', rownames(All.RC)),]
 Ef.RC <- All.RC[grepl('^EfaB.*', rownames(All.RC)),]
 
 pData <- read.table("output_data/Sample_pData.csv", sep=",")
-pData <- pData[!pData$samples%in%excluded.samples, ]
+pData <- pData[!pData$sample%in%excluded.samples, ]
 
 ## Mouse and Eimeria specific
-Ef.pData <- pData[!pData$timepoint%in%0, ]
-Mm.pData <- pData[!is.na(pData$timepoint), ]
+Ef.pData <- pData[!pData$dpi%in%0, ]
+Mm.pData <- pData[!pData$dpi%in%c("environmental", "in vitro"), ]
 
 Ef.pData$grouped <- as.factor(as.character(Ef.pData$grouped))
 Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
 
 ## Mouse and Eimeri spcific samples
-Mm.RC <- Mm.RC[, colnames(Mm.RC)%in%Mm.pData$samples]
-Ef.RC <- Ef.RC[, colnames(Ef.RC)%in%Ef.pData$samples] 
+Mm.RC <- Mm.RC[, colnames(Mm.RC)%in%Mm.pData$sample]
+Ef.RC <- Ef.RC[, colnames(Ef.RC)%in%Ef.pData$sample] 
 
 library(statmod)
 library(edgeR)
@@ -95,8 +95,8 @@ Mm.contrasts <- makeContrasts(
     N5vs0 = NMRI.1stInf.5dpi-NMRI.2ndInf.0dpi, #2
     N7vs0 = NMRI.1stInf.7dpi-NMRI.2ndInf.0dpi, #3
     ## just infections in Black and Rag
-    B5vs0 = C57BL6.1stInf.5dpi- C57BL6.0dpi , #4
-    R5vs0 = Rag.1stInf.5dpi - Rag.0dpi,		#5
+    B5vs0 = C57BL6.1stInf.5dpi- C57BL6.1stInf.0dpi , #4
+    R5vs0 = Rag.1stInf.5dpi - Rag.1stInf.0dpi,		#5
     ## differences in 1st vs 2nd infection
     N3.1stvsN3.2nd = NMRI.1stInf.3dpi-NMRI.2ndInf.3dpi, #6
[…]

### 2016-08-12 B_mds_sample_analysis.R

@@ -1,67 +1,41 @@
 ## DIFFERENTIAL GENE EXPRESSION ANALYSIS ON EIMERIA SEQS #####
 
-#setwd("/home/totta/Ef_RNAseq")
-
-if(!exists("raw.counts.4.bg")){
-	source("2_edgeR_diff.R")
-        }
-
 library(limma)
 library(scales)
 library(stringr)
 
-#Ef.RC <- raw.counts.4.bg(Ef.bg)
+library(edgeR)
 
-######################################
+excluded.samples <- c("NMRI_2ndInf_3dpi_rep1",
+                      "NMRI_2ndInf_5dpi_rep2",
+                      "NMRI_1stInf_0dpi_rep1")
 
-######################################
-pdf("figures/rep_pairs_EfG.pdf", width = 70, height=70) # So far not working - because R interrupt?
-ggpairs(Ef.RC[[3]])
-dev.off()
+All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
 
+All.RC <- All.RC[, !colnames(All.RC)%in%excluded.samples]
 
-##################### For EIMERIA: #################################
+## Mouse and Eimeria specific
+Mm.RC <- All.RC[grepl('^ENSMUS.*', rownames(All.RC)),]
+Ef.RC <- All.RC[grepl('^EfaB.*', rownames(All.RC)),]
 
-ggplot()+
-    geom_point(data = tableEf, aes(x = Samples,
-                   y = as.numeric(colSums(Ef.RC[[3]])))) +
-	scale_y_log10(labels = comma) +
-            ylab("Number of reads") +
-                theme(axis.text.x = element_text(angle = 45, hjust = 1))
+pData <- read.table("output_data/Sample_pData.csv", sep=",")
+pData <- pData[!pData$samples%in%excluded.samples, ]
 
+## Mouse and Eimeria specific
+Ef.pData <- pData[!pData$timepoint%in%0, ]
+Mm.pData <- pData[!is.na(pData$timepoint), ]
 
+Ef.pData$grouped <- as.factor(as.character(Ef.pData$grouped))
+Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
 
-pdf("figures/rep_pairs_NORM_EfG.pdf", width = 70, height=70)
-ggpairs(cpm(GM.E)) #ggpairs makes matrix of plots
-dev.off()
[…]

### 2016-08-15 manuscript/bmc_article.tex
Up to the enrichment tests

@@
   \section*{A dual transcriptomics experiment}
+
 We performed mRNA sequencing of caecum tissue from mice infected with
-the apicomplexan parasite \textit{E. falciformis}. We worked with two
-biological replicates for all but two conditions (with one and three
-replicates, respecitively). For each replicate sample we enriched
-caecum tissue from three infected mice for epithelial cells. We
-extracted mRNA and constructed sequencing libraries for different
-timepoints after infection and mouse strains which displayed different
-immunocompetence in infection trials (see figure x).
-
-In these experiments we analysed RAG-1-/- mice and compared them to
-the parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
-response du to the absence of mature B and T lymphocytes. Differences
-in oocyst output were more pronounced between NMRI and both C57BL6 and
-RAG-1-/-. Surprisingly this finding was made both in primary and
+the apicomplexan parasite \textit{E. falciformis}. Oocysts and
+sporozoites were included as ``environmental'' stages and processed
+``\textit{in vitro}''. To follow the lifecycle of the parsite, we
+compare different timepoints after infection.
+
+We additonally used different mouse strains which displayed different
+immunocompetence in infection trials to assess the influence host
+immunocompetence on parasite development (figure 1a,b). In these
+experiments we analysed RAG-1-/- mice and compared them to the
+parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
+response du to the absence of mature B and T lymphocytes.  All three
+mouse strains were used for infections of both naive and previously
+challenge infected animals.
+
+Concerning basic phenotyping, differences in oocyst output were
+visible both between the immunocompetent (NMRI and C57BL6) and
+immunodeficient (RAG-1-/-) host strain, but also between the two
+similarly immunocompetent (NMRI and C57BL6) strains (figure
+1a). Surprisingly this finding was consistent in both in primary and
 challenge infection. Similarly development of intestinal stages as
-assayed by quantitative reverse transcription PCR (RT qPCR) also ...
-
-This lead us to question the influence of the innate or cell
-autonomouse immune system on early stages of infection. And we thus
-settled on an experimental design comparing infections at 5dpi for all
-
-For the parasite oocysts and sporozoites were included as
-``environmental'' and ``\textit{in vitro}'' processed lifecycle
-stages.
+assayed by quantitative reverse transcription PCR (RT qPCR) also
+... ADD HERE (figure 1b).
+
+We thus aim to analyse the lifecycle of \textit{E. falciformis} under
+the influence of (most likely the innate or cell autonomouse) immune
+respons at on early stages of infection. We used an experimental
+design comparing infections at 5dpi for all possible experimental
+conditions (all three mouse strains in naive and challenge infection),
+and analyse additional infections at 3dpi (figure 1c).
+
+In our RNAseq experiment we worked with two biological replicates for
+all but two conditions (with one and three replicates,
+respecitively). For each replicate sample we enriched caecum tissue
+from three infected mice for epithelial cells, extracted mRNA and
+constructed sequencing libraries. These libraries were sequenced on
+several lanes of Illumina GAIIX and HiSeq machines and mapped to both
+mouse and parasite genome simulatneously to avoid spurious assignments
+of read in ultra conserved genomic regions.
 
 The number of sequence read mappings obtained for individual
-replicates ranged from 21,517,055 (sample NMRI\_sporozoites\_rep2) to
-593,396,312 (NMRI\_2ndInf\_5dpi\_rep1).
+replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to
+145,033,326 (NMRI\_2ndInf\_5dpi\_rep1).
 

### 2016-08-15 2_data_curation.R

@@ -17,7 +17,6 @@ create.pdata <- function (featurecountsTarget){
     grouped <- sub("_rep\\d+$" , "", sample)
     challenged <- ifelse(grepl("1stInf", grouped), "1st",
                   ifelse(grepl("2ndInf", grouped), "2nd",
-                  ifelse(grepl("oocysts", grouped), "environmental", "in vitro")))
     rep <- sub(".*_(rep\\d+)$" , "\\1", sample) # (name here) is taken "\\1" here
     mouse.strain <- sub("^(.*?)_.*", "\\1", grouped)
     dpi <- as.numeric(as.character(sub("^.*_(.*)dpi", "\\1", grouped)))
@@ -76,7 +75,6 @@ RC.table$p.Ef.reads <- round(RC.table$c.Ef.reads/RC.table$c.mapping.counts*100, 
 
 
 RC.table$dpi[RC.table$challenged%in%"environmental"] <- "environmental"
-RC.table$dpi[RC.table$challenged%in%"in vitro"] <- "in vitro"
 
 ## plotting parasite percentage of reads
 

### 2016-08-15 3_edgeR_diff.R

@@ -1,5 +1,6 @@
 #### DIFFERENTIAL GENE EXPRESSION ANALYSIS
 ## needed from data import script are two raw count objects 
+library(RSvgDevice)
 
 ## We know from 2_data_curation.R that we want to exclude:
 excluded.samples <- c("NMRI_2ndInf_3dpi_rep1",
@@ -19,7 +20,6 @@ pData <- pData[!pData$sample%in%excluded.samples, ]
 
 ## Mouse and Eimeria specific
 Ef.pData <- pData[!pData$dpi%in%0, ]
-Mm.pData <- pData[!pData$dpi%in%c("environmental", "in vitro"), ]
 
 Ef.pData$grouped <- as.factor(as.character(Ef.pData$grouped))
 Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
@@ -27,6 +27,14 @@ Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
 ## Mouse and Eimeri spcific samples
 Mm.RC <- Mm.RC[, colnames(Mm.RC)%in%Mm.pData$sample]
 Ef.RC <- Ef.RC[, colnames(Ef.RC)%in%Ef.pData$sample] 
+
+## Make sure the pData follows the same order as the sample count
+## columns!!!
+Mm.RC <- Mm.RC[, order(colnames(Mm.RC))]
+Ef.RC <- Ef.RC[, order(colnames(Ef.RC))]
+
+Mm.pData <- Mm.pData[order(Mm.pData$sample), ]
+Ef.pData <- Ef.pData[order(Ef.pData$sample), ]
 
 library(statmod)
 library(edgeR)
@@ -152,13 +160,13 @@ get.my.models <- function (RC, cutoff, group,
 
 ## filter: with 3000 as cutoff, bimodal distr. almost not visible (see
 ## A_data_curation.R)
-Mm.1st.pass.model <- get.my.models(Mm.RC, cutoff=3000,
+Mm.1st.pass.model <- get.my.models(Mm.RC, cutoff=1000,
                                    group = Mm.pData$grouped,
                                    contrasts=Mm.contrasts,
                                    design=Mm.design,
                                    norm="upperquartile")
 
-Ef.1st.pass.model <- get.my.models(Ef.RC, cutoff=100,
+Ef.1st.pass.model <- get.my.models(Ef.RC, cutoff=10,
                                    group = Ef.pData$grouped,
[…]

### 2016-08-15 5_compareArray.R

@@ -98,7 +98,9 @@ RNAseq.Array.logFC <- merge(Array.logFC[,c("ensembl_id", "X24h", "X144h")],
                             Mm.DE.FC, 
                             by.x = "ensembl_id", by.y = "gene")
 
-pdf("figures/Figure2a_Array144vsRNAseqN7.pdf")
+cor(RNAseq.Array.logFC[, -1])[,1:2]
+
+pdf("figures/Figure2a_Array144vsRNAseqN7.pdf", onefile = FALSE)
 ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0)) +
   geom_point(alpha=0.8) +
   stat_density2d(aes(fill=..level..), size=2, geom="polygon") +

### 2016-08-15 6_clustering_and_venns.R

@@ -3,14 +3,15 @@
 library(pheatmap)
 library(made4)
 library(RColorBrewer)
+library(VennDiagram)
 
 ## raw count data
-Mm.RC <- read.table("output_data/RC_Mm_genes.csv", sep=",")
-Ef.RC <- read.table("output_data/RC_Ef_genes.csv", sep=",")
+Mm.nRC <- read.table("output_data/Mm_norm_counts.csv", sep=",")
+Ef.nRC <- read.table("output_data/Ef_norm_counts.csv", sep=",")
 
 ## replace zeros and avoid zero variance in cluter scaling
-Mm.RC[Mm.RC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Mm.RC[Mm.RC==0]), replace =TRUE)
-Ef.RC[Ef.RC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Ef.RC[Ef.RC==0]), replace =TRUE)
+#Mm.nRC[Mm.nRC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Mm.nRC[Mm.nRC==0]), replace =TRUE)
+#Ef.nRC[Ef.nRC==0] <- sample(seq(0.01, 0.1, 0.0001), length(Ef.nRC[Ef.nRC==0]), replace =TRUE)
 
 ## data on statistical testing for selection
 Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
@@ -29,26 +30,26 @@ cycle.cols <- c("N3vsN5","N3vsN7", "N3vsOoc", "N3vsSpo",
                 "N5vsN7", "N5vsOo", "N5vsSp",
                 "N7vsOo", "N7vsSp",
                 "SpvsOo")
-Ef.cycle.diff.data <- select.diff.data(Ef.RC, Ef.DE.test, cycle.cols, 0.01)
+Ef.cycle.diff.data <- select.diff.data(Ef.nRC, Ef.DE.test, cycle.cols, 0.01)
 
 ############# Eimeria 1st VERSUS 2nd INFECTION ##################
 first.second.cols <- c("B5.1stvsB5.2nd", "N3.1stvsN3.2nd", "N5.1stvsN5.2nd",
                        "N7.1stvsN7.2nd","R5.1stvsR5.2nd")
-Ef.first2nd.diff.data <- select.diff.data(Ef.RC, Ef.DE.test, first.second.cols, 0.5)
[…]

### 2016-08-15 B_mds_sample_analysis.R

@@ -3,34 +3,30 @@
 library(limma)
 library(scales)
 library(stringr)
-
 library(edgeR)
+
 
 excluded.samples <- c("NMRI_2ndInf_3dpi_rep1",
                       "NMRI_2ndInf_5dpi_rep2",
                       "NMRI_1stInf_0dpi_rep1")
 
-All.RC <- as.matrix(read.table("output_data/RC_All_genes.csv", sep=","))
-
-All.RC <- All.RC[, !colnames(All.RC)%in%excluded.samples]
-
-## Mouse and Eimeria specific
-Mm.RC <- All.RC[grepl('^ENSMUS.*', rownames(All.RC)),]
-Ef.RC <- All.RC[grepl('^EfaB.*', rownames(All.RC)),]
+## Mouse and Eimeria specific NORMALIZED counts
+Mm.RC <- read.table("output_data/Mm_norm_counts.csv", sep=",")
+Ef.RC <- read.table("output_data/Ef_norm_counts.csv", sep=",")
 
 pData <- read.table("output_data/Sample_pData.csv", sep=",")
-pData <- pData[!pData$samples%in%excluded.samples, ]
+pData <- pData[!pData$sample%in%excluded.samples,]
 
 ## Mouse and Eimeria specific
-Ef.pData <- pData[!pData$timepoint%in%0, ]
-Mm.pData <- pData[!is.na(pData$timepoint), ]
+Ef.pData <- pData[!pData$dpi%in%0, ]
 
-Ef.pData$grouped <- as.factor(as.character(Ef.pData$grouped))
-Mm.pData$grouped <- as.factor(as.character(Mm.pData$grouped))
+## Make sure the pData follows the same order as the sample count
+## columns!!!
+Mm.RC <- Mm.RC[, order(colnames(Mm.RC))]
+Ef.RC <- Ef.RC[, order(colnames(Ef.RC))]
 
-## Mouse and Eimeri spcific samples
-Mm.RC <- Mm.RC[, colnames(Mm.RC)%in%Mm.pData$samples]
-Ef.RC <- Ef.RC[, colnames(Ef.RC)%in%Ef.pData$samples] 
+Mm.pData <- Mm.pData[order(Mm.pData$sample), ]
+Ef.pData <- Ef.pData[order(Ef.pData$sample), ]
 
 ###################################################################
 ## COLORS FOR pDATA() GROUPS 
@@ -96,77 +92,98 @@ seq.colors.mm = c("#474747", "#474747", "#474747", "#474747", "#474747",
[…]

### 2016-08-15 manuscript/bmc_article.tex
Spearman correlation estimates array<->RNAseq

@@
 
-The number of sequence read mappings obtained for individual
-replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to
-145,033,326 (NMRI\_2ndInf\_5dpi\_rep1).
-
-As samples and idividual replicates were sequenced in batches to
-different depth and using different instrumentation (Table 1) we
-performed quality controls confirming the abscence of batch effects
-influencing analysis and quality of our results (additional file xyc).
-
-The results of these multivariate analyses in relation to biological
-interpretation are described further below.
+The number of total read mappings obtained for individual replicates
+ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
+(NMRI\_2ndInf\_5dpi\_rep1). As samples and idividual replicates were
+sequenced in batches to different depth and using different
+instrumentation (Table 1) we performed quality controls confirming the
+abscence of batch effects influencing analysis and quality of our
+results (additional file xyc). The results of these multivariate
+analyses in relation to biological interpretation are described
+further below.
 
@@
 assigened to the \textit{E. falciformis} genome in samples considered
-infected.
+infected (table 1).
 
@@
 of infection is dominated by the parasite with proportional mRNA
-abundance between XXX and XXX %
+abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and even 92\% (in
+NMRI\_1stInf\_7dpi\_rep1).
 
@@
 mapped to the transcriptome of either \textit{E. falciformis} and the
-mouse are indicated in Table 1 for all replicates.
+mouse are indicated in Table 1 for all replicates. We additonally
+assessed the number of genes considered ``expressed'', further
+analysis and discussion considers the possibility that especially for
+\textit{E. falciformis} the number of reads might have been
+insufficient to fully normalize the dataset of the samples with the
+lowest representation against those with the hightest.
 
@@
 mice at 6dpi on Agilent microarrays (Schmidt et al 2012) or 7dpi (our
-RNAseq data) agains uninfected controls shows a strong correlateion
-(Spearman's \rho = 333; $R^2$ from a linear model = 444 ; figure 2
-a). Considering both and technical differences between the tow methods
-and the biological changes along the lifecycle this confirms the
-adequacy of our methods for assessing the host transcriptome.
-
-XX mouse genes were considered differently expressed (DE; FDR<1\%)
-between uinfected and 3dpi, XX genes between controls and 5dpi, xxx
-genes between 0dpi and 7dpi. This lead to combined set of xxxx genes
-reacting to infection (see figure 2 b).
-
-We detected no distinct response early in infection.  When testing for
-differences between 3dpi and later infection (5dpi and 7dpi) DE genes
-were simply a subset of genes DE between respective infections and
-controls.
-
-To further validate these patterns we performed hierarchical
-clustering on mouse mRNAs DE at different dpi (union of DE
-genes).
-
-%% Samples from day 7 p.i. form a separate cluster (NMRI mice only), not
-%% distinguishing between first and challenge infection. For these
-%% samples, the pattern is most pronounced in gene cluster 1 (up), and 4
-%% and 6 (down). Within cluster 1, the second replicate from the
-%% challenge infection has a deviant profile (compare mouse data in
-%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
-%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
-%% 3 and 5 p.i. from all three mouse strains cluster together. These
-%% samples are distinct from oocysts, NMRI day 7 p.i., and
-%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
-%% oocysts or day 7.
-
-%%% Test _this_:
-The more ponounced changes later in infection are is likely reflecting
-a lag time in detection before the parasite reaches sufficient
-intensity during asexual expansion (schizogony). Those changes are
-mainly related to the onset of an adaptive immune response.
+RNAseq data) against uninfected controls shows a strong correlateion
+(Spearman's \rho = 0.74; figure 2a). Considering both and technical
+differences between the tow methods and the biological changes along
+the lifecycle this confirms the adequacy of our methods for assessing
+the host transcriptome.
+
+325 mouse genes were considered differently expressed (DE; FDR<1\%)
+between uinfected and 3dpi, 1,804 between controls and 5dpi and 2,711
+genes between 0dpi and 7dpi. This lead to combined set of 3453 genes
+reacting to infection (figure 2bi).  DE genes early in infection (3dpi
+and and 5dpi) were not simply a subset of genes becoming DE later in
+infection (7dpi). To further validate this pattern of a distinct
+response in early infection we performed hierarchical clustering on
+(the union of) mouse genes DE between different dpi (figure 2c).
+
+The changes in early infection are ...
+
+Those changes are similarly pronounce in the immunocompetent strains
+and RAG-/- immunodefficient animals.
+
+The ponounced changes later in infection are is likely reflecting the
+onset of an adaptive immune response.
 

### 2016-08-15 5_compareArray.R

@@ -98,7 +98,7 @@ RNAseq.Array.logFC <- merge(Array.logFC[,c("ensembl_id", "X24h", "X144h")],
                             Mm.DE.FC, 
                             by.x = "ensembl_id", by.y = "gene")
 
-cor(RNAseq.Array.logFC[, -1])[,1:2]
+cor(RNAseq.Array.logFC[, -1], method="spearman")[,1:2]
 
 pdf("figures/Figure2a_Array144vsRNAseqN7.pdf", onefile = FALSE)
 ggplot(RNAseq.Array.logFC, aes(X144h, logFC.N7vs0)) +

### 2016-08-15 manuscript/bmc_article.tex
At the stage of merging patterns over samples and GO-Terms : Handing over to Totta!

@@
 pronounced (involving more genes) at later post infection (especially
-at 7dpi; Table 1).
+at 7dpi; Table 2).
 
@@
 
-The changes in early infection are ...
-
-Those changes are similarly pronounce in the immunocompetent strains
-and RAG-/- immunodefficient animals.
-
-The ponounced changes later in infection are is likely reflecting the
-onset of an adaptive immune response.
+The ponounced changes later in infection are reflecting the expected
+onset of an adaptive immune response. This is underlined by GO-terms
+such as ``antigen binding'' and ``immunoglobulin receptor binding''
+(molecular function) ``immune system process'' and ``adaptive immune
+response'' (biological process) highly enriched in Cluster 5 of figure
+2c.
+
+Particular ``regulation of cell death'' and ``regulation of apoptotic
+process'' (BP)
+
+``UDP-N-acetylglucosamine-lysosomal-enzyme'' (MF)
+
+%% This might be relevant: The DMAP interaction domain of
+%% UDP-GlcNAc:lysosomal enzyme
+%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
+%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
+%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
+%% N-acetyl-D-glucosaminyl-phospho-D-mannose
+
+General changes upon infection are a downregulation of genes involved
+``lipid modification'' or more detailed ``phosphatidylinositol
+phosphorylation''. These changes are similarly pronounce in the
+immunocompetent strains and RAG-/- immunodefficient animals (cluster 3
+of fibure 2c) (see supplementary table X).  The presence of these
+changes immunocompromized animals invite the speculation that those
+genes are independent of a potential later adaptive immune response.
+
+Changes not detectable in RAG-/- immunodefficient animals (cluster 3
+of fibure 2c) but otherwise observed in general infection are
+characterized by and enrichment to genes involved in ``positive
+regulation of calcineurin-NFAT'', ``blood coagulation'' and ``inositol
+phosphate-mediated signaling'' (see supplementary table X). The
+abscence of these changes in RAG-/- mice allow us to speculate that
+those genes might be involved in a very early onlset of adaptive
+immune response.
+
+%% Go on here if you can...
+
+We dedtected further differences between infections of naive and
+challenge infected mous hosts , especially at 7 dpi. These genes are
+characterized by...
+%% We should't limit ourselfs to testing the clusters only! But
+%% discuss the cluster first if you got difficulties coding the
+%% enrichment tests.
 
@@
 Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison."NMRI" followed by number
-indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
-p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
-implemented in edgeR are included. NAs are missing samples or not
-applicable for the species. NA* is due to missing NMRI day 0 sample
-from first infection. For \textit{E. falciformis} no mRNAs are
-significantly differently abundant between first and second infection,
-whereas in mouse there are some differences, especially on day 7
-p.i. However, upon hierarchical clustering of these genes (union of
-500 with lowest FDR), samples do not cluster according to first/second
-infection (Figure SX ?). In comparisons between mouse strains, 22
-genes are detected to be different in the parasite data. The
-difference is however not between Rag1-/- mice as expected, but
-between NMRI and C57BL/6. For mouse....
+differently abundant mRNAs per comparison.  For
+\textit{E. falciformis} no mRNAs are significantly differently
+abundant between first and second infection.
+
+Changes throughout the intr-host lifecycle can be clearly categorized
+in an early (3 and 5 dpi) and late (7dpi) phase. Sporozoites show
+expression profiles similar more similar to early infection stages
+while oocysts show distinctly different mRNA abundance profile (MDS
+plot .
 
 We applied hierarchical clustering on \textit{E. falciformis} mRNAs
-which were differently abundant at different timepoints p.i.,
-including comparisons with oocysts and sporozoites (union of 500 with
-lowest FDR). Samples from day 7 p.i. form a separate cluster (NMRI
-mice only), not distinguishing between first and challenge
-infection. For these samples, the pattern is most pronounced in gene
-cluster 1 (up), and 4 and 6 (down). Within cluster 1, the second
-replicate from the challenge infection has a deviant profile (compare
-mouse data in Figure 2). Distinct groups of genes also define
-sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6
-down). mRNA profiles on days 3 and 5 p.i. from all three mouse strains
-cluster together. These samples are distinct from oocysts, NMRI day 7
-p.i., and sporozoites. Sporozoites cluster closer with days 3 and 5
-than with oocysts or day 7.
+which were differently abundant at different dpi, including
+comparisons with oocysts and sporozoites.
+
+Samples from 7dpi form a separate cluster (NMRI mice only), not
+distinguishing between first and challenge infection.
+
+%% Mix it all well into the GO-term discussion! GO-term inferred
+%% biiology needs to be reported in connection to clutering-inferred
+%% ``sample biology''.
+
+%% For these samples, the pattern is most pronounced in gene cluster 1
+%% (up), and 4 and 6 (down). Within cluster 1, the second replicate from
+%% the challenge infection has a deviant profile (compare mouse data in
+%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
+%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
+%% 3 and 5 p.i. from all three mouse strains cluster together. These
+%% samples are distinct from oocysts, NMRI day 7 p.i., and
+%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
+%% oocysts or day 7.
 

### 2016-08-15 6_clustering_and_venns.R

@@ -68,10 +68,13 @@ get.cluster.tree.df <- function(data, k){
 Ef.hclustered.cycle <- get.cluster.tree.df(Ef.cycle.diff.data, 7)
 Ef.hclustered.first.second <- get.cluster.tree.df(Ef.first2nd.diff.data,  7)
 
+write.table(Ef.hclustered.cycle, "output_data/Ef_hclustered_cycle.csv", sep=",")
+
 Mm.hclustered.cycle <- get.cluster.tree.df(Mm.cycle.diff.data, 7)
 Mm.hclustered.first.second <- get.cluster.tree.df(Mm.first2nd.diff.data,  4)
 Mm.hclustered.strain <- get.cluster.tree.df(Mm.comp.rag.diff.data, 4)
 
+write.table(Mm.hclustered.cycle, "output_data/Mm_hclustered_cycle.csv", sep=",")
 
 pdf("Supplement/Ef1st2ndHeatmap.pdf",
     height = 8, width = 8, onefile = FALSE) # onefile command to hack away empty page in pdf

### 2016-08-15 7_enrichment.R

@@ -1,9 +1,3 @@
-## still open question: How to implement the FCS(?) method
-# Check checkpoint package for package control.
-if(!require(topGO)) biocLite("topGO") # Imports package if user does not have it
-if(!require(xtable)) biocLite("xtable") # Imports package if user does not have it
-if(!require(AnnotationDbi)) biocLite("AnnotationDbi") # Imports package if user does not have it
-
 library(topGO)
 library(xtable)
 library(AnnotationDbi)
@@ -11,8 +5,7 @@ library(AnnotationDbi)
 if(!exists("hcluster")){
     source("5_clustering.R")
 }
-
-annotation.frame <- read.table("output_data/annotation_data", sep=",")
+annotation.frame <- read.table("output_data/annotation_data.csv", sep=",")
 
 ## just to get the universe of genes which were tested at all:
 Mm.DE.test <- read.table("output_data/Mm_DEtest.csv", sep=",")
@@ -38,48 +31,71 @@ exp.universe[["Mm"]] <- names(gene2GO[["Mm"]])
 exp.universe[["Ef"]] <- names(gene2GO[["Ef"]])
 
 
+hcluster <- list()
+hcluster[["Mm"]] <- read.table("output_data/Mm_hclustered_cycle.csv", sep=",",
+                               header=TRUE)
+hcluster[["Ef"]] <- read.table("output_data/Ef_hclustered_cycle.csv", sep=",",
+                               header=TRUE)
 
 #### We search for enrichment of genes from clustering 
 ######## Clustering #############
-
 to.test <- as.data.frame(rbind(
     ## Eimeria
-    cbind(cluster=1,type="Day7Up_1", species="Ef"),
-    cbind(cluster=2,type="Day7Up_2", species="Ef"),
-    cbind(cluster=3,type="OocSporDown", species="Ef"),
-    cbind(cluster=4,type="OocystsUp", species="Ef"),
-    cbind(cluster=5,type="OoSpoDown_day5Up", species="Ef"),
-    cbind(cluster=6,type="SporozoitesUp", species="Ef"),
-    cbind(cluster=7,type="Day7OocyDown_SporoUP", species="Ef"),
+    ## cbind(cluster=1,type="Day7Up_1", species="Ef"),
+    ## cbind(cluster=2,type="Day7Up_2", species="Ef"),
+    ## cbind(cluster=3,type="OocSporDown", species="Ef"),
[…]

### 2016-08-15 manuscript/bmc_article.tex
A remider on the ortholog comparisons...

@@
 in sporozoites (cluster 2 association).
+
+\section*{Evolutionary conservation of expression changes in \textit{E. falciformis}}
+
+For comparison of changes in the lifecycle of \textit{E. falciformis}
+with the economically important chicken parasite \textit{E. tenella}
+and the model apicompomplexan \textit{Toxoplasma gondii} we downloaded
+expresson datasets from ...
+
+Oocysts, Sproozoites
 

### 2016-08-17 Supplement/tex/cluster3_BP_Mm_allInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:27:09 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0021527 & spinal cord association neuron different... &   2 &   2 & 0.04 & 0.00 & 0.02 \\ 
  GO:0071280 & cellular response to copper ion &   3 &   2 & 0.06 & 0.00 & 0.02 \\ 
  GO:0046854 & phosphatidylinositol phosphorylation &  25 &   4 & 0.50 & 0.00 & 0.02 \\ 
  GO:0061512 & protein localization to cilium &  25 &   4 & 0.50 & 0.00 & 0.02 \\ 
  GO:0030258 & lipid modification & 140 &   9 & 2.77 & 0.00 & 0.02 \\ 
  GO:0021516 & dorsal spinal cord development &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0060830 & ciliary receptor clustering involved in ... &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0060971 & embryonic heart tube left/right pattern ... &   4 &   2 & 0.08 & 0.00 & 0.02 \\ 
  GO:0046834 & lipid phosphorylation &  29 &   4 & 0.57 & 0.00 & 0.02 \\ 
  GO:0015936 & coenzyme A metabolic process &  16 &   3 & 0.32 & 0.00 & 0.02 \\ 
  GO:0021522 & spinal cord motor neuron differentiation &  16 &   3 & 0.32 & 0.00 & 0.02 \\ 
  GO:0019087 & transformation of host cell by virus &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0032484 & Ral protein signal transduction &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0048842 & positive regulation of axon extension in... &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:1902669 & positive regulation of axon guidance &   5 &   2 & 0.10 & 0.00 & 0.02 \\ 
  GO:0046486 & glycerolipid metabolic process & 218 &  11 & 4.32 & 0.00 & 0.02 \\ 
  GO:0007033 & vacuole organization & 130 &   8 & 2.58 & 0.00 & 0.02 \\ 
  GO:0051188 & cofactor biosynthetic process & 104 &   7 & 2.06 & 0.00 & 0.02 \\ 
  GO:0042073 & intraciliary transport &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0044088 & regulation of vacuole organization &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0098840 & protein transport along microtubule &  35 &   4 & 0.69 & 0.00 & 0.02 \\ 
  GO:0021514 & ventral spinal cord interneuron differen... &   6 &   2 & 0.12 & 0.01 & 0.02 \\ 
  GO:0021515 & cell differentiation in spinal cord &  19 &   3 & 0.38 & 0.01 & 0.02 \\ 
  GO:0021517 & ventral spinal cord development &  19 &   3 & 0.38 & 0.01 & 0.02 \\ 
  GO:0009410 & response to xenobiotic stimulus &  37 &   4 & 0.73 & 0.01 & 0.02 \\ 
  GO:0033865 & nucleoside bisphosphate metabolic proces... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0033875 & ribonucleoside bisphosphate metabolic pr... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0034032 & purine nucleoside bisphosphate metabolic... &  20 &   3 & 0.40 & 0.01 & 0.02 \\ 
  GO:0051186 & cofactor metabolic process & 269 &  12 & 5.33 & 0.01 & 0.02 \\ 
  GO:0044255 & cellular lipid metabolic process & 633 &  22 & 12.54 & 0.01 & 0.02 \\ 
  GO:0006729 & tetrahydrobiopterin biosynthetic process &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0014049 & positive regulation of glutamate secreti... &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0046146 & tetrahydrobiopterin metabolic process &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0070050 & neuron cellular homeostasis &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0072488 & ammonium transmembrane transport &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0097094 & craniofacial suture morphogenesis &   7 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0032892 & positive regulation of organic acid tran... &  21 &   3 & 0.42 & 0.01 & 0.02 \\ 
  GO:0021510 & spinal cord development &  40 &   4 & 0.79 & 0.01 & 0.02 \\ 
  GO:1903793 & positive regulation of anion transport &  40 &   4 & 0.79 & 0.01 & 0.02 \\ 
  GO:0031503 & protein complex localization &  91 &   6 & 1.80 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-08-17 Supplement/tex/cluster4_BP_Mm_CompInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:28:15 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0070886 & positive regulation of calcineurin-NFAT ... &   5 &   3 & 0.11 & 0.00 & 0.01 \\ 
  GO:0048016 & inositol phosphate-mediated signaling &  23 &   5 & 0.52 & 0.00 & 0.01 \\ 
  GO:0050817 & coagulation & 120 &  10 & 2.72 & 0.00 & 0.01 \\ 
  GO:0032102 & negative regulation of response to exter... & 184 &  12 & 4.17 & 0.00 & 0.02 \\ 
  GO:0007596 & blood coagulation & 118 &   9 & 2.67 & 0.00 & 0.02 \\ 
  GO:0007154 & cell communication & 3168 &  94 & 71.77 & 0.00 & 0.02 \\ 
  GO:0007599 & hemostasis & 120 &   9 & 2.72 & 0.00 & 0.02 \\ 
  GO:0007006 & mitochondrial membrane organization &  76 &   7 & 1.72 & 0.00 & 0.02 \\ 
  GO:0023052 & signaling & 3091 &  92 & 70.02 & 0.00 & 0.02 \\ 
  GO:0070884 & regulation of calcineurin-NFAT signaling... &  11 &   3 & 0.25 & 0.00 & 0.02 \\ 
  GO:0006120 & mitochondrial electron transport, NADH t... &  12 &   3 & 0.27 & 0.00 & 0.02 \\ 
  GO:0043588 & skin development & 150 &  10 & 3.40 & 0.00 & 0.02 \\ 
  GO:1903035 & negative regulation of response to wound... & 103 &   8 & 2.33 & 0.00 & 0.02 \\ 
  GO:0044700 & single organism signaling & 3084 &  91 & 69.86 & 0.00 & 0.02 \\ 
  GO:0007033 & vacuole organization & 130 &   9 & 2.94 & 0.00 & 0.02 \\ 
  GO:0035898 & parathyroid hormone secretion &   4 &   2 & 0.09 & 0.00 & 0.02 \\ 
  GO:0060684 & epithelial-mesenchymal cell signaling &   4 &   2 & 0.09 & 0.00 & 0.02 \\ 
  GO:0050701 & interleukin-1 secretion &  27 &   4 & 0.61 & 0.00 & 0.02 \\ 
  GO:0006914 & autophagy & 358 &  17 & 8.11 & 0.00 & 0.02 \\ 
  GO:0006937 & regulation of muscle contraction &  67 &   6 & 1.52 & 0.00 & 0.02 \\ 
  GO:0030168 & platelet activation &  67 &   6 & 1.52 & 0.00 & 0.02 \\ 
  GO:1902188 & positive regulation of viral release fro... &  15 &   3 & 0.34 & 0.00 & 0.02 \\ 
  GO:0071496 & cellular response to external stimulus & 139 &   9 & 3.15 & 0.00 & 0.02 \\ 
  GO:0032905 & transforming growth factor beta1 product... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0032908 & regulation of transforming growth factor... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0045991 & carbon catabolite activation of transcri... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:1900165 & negative regulation of interleukin-6 sec... &   5 &   2 & 0.11 & 0.00 & 0.02 \\ 
  GO:0032692 & negative regulation of interleukin-1 pro... &  16 &   3 & 0.36 & 0.01 & 0.02 \\ 
  GO:0050878 & regulation of body fluid levels & 197 &  11 & 4.46 & 0.01 & 0.02 \\ 
  GO:0007165 & signal transduction & 2897 &  84 & 65.63 & 0.01 & 0.02 \\ 
  GO:0007176 & regulation of epidermal growth factor-ac... &  17 &   3 & 0.39 & 0.01 & 0.02 \\ 
  GO:0033173 & calcineurin-NFAT signaling cascade &  17 &   3 & 0.39 & 0.01 & 0.02 \\ 
  GO:0050663 & cytokine secretion & 121 &   8 & 2.74 & 0.01 & 0.02 \\ 
  GO:0019076 & viral release from host cell &  33 &   4 & 0.75 & 0.01 & 0.02 \\ 
  GO:0006936 & muscle contraction & 123 &   8 & 2.79 & 0.01 & 0.02 \\ 
  GO:0050728 & negative regulation of inflammatory resp... &  75 &   6 & 1.70 & 0.01 & 0.02 \\ 
  GO:0007220 & Notch receptor processing &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0044828 & negative regulation by host of viral gen... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0045990 & carbon catabolite regulation of transcri... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0050711 & negative regulation of interleukin-1 sec... &   6 &   2 & 0.14 & 0.01 & 0.02 \\ 
  GO:0031670 & cellular response to nutrient &  18 &   3 & 0.41 & 0.01 & 0.02 \\ 
  GO:0050919 & negative chemotaxis &  18 &   3 & 0.41 & 0.01 & 0.02 \\ 
  GO:0032612 & interleukin-1 production &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0050818 & regulation of coagulation &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0060048 & cardiac muscle contraction &  54 &   5 & 1.22 & 0.01 & 0.02 \\ 
  GO:0042493 & response to drug & 151 &   9 & 3.42 & 0.01 & 0.02 \\ 
  GO:0031668 & cellular response to extracellular stimu... & 126 &   8 & 2.85 & 0.01 & 0.02 \\ 
  GO:0008053 & mitochondrial fusion &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0044803 & multi-organism membrane organization &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0050704 & regulation of interleukin-1 secretion &  19 &   3 & 0.43 & 0.01 & 0.02 \\ 
  GO:0050707 & regulation of cytokine secretion & 104 &   7 & 2.36 & 0.01 & 0.02 \\ 
  GO:0009991 & response to extracellular stimulus & 274 &  13 & 6.21 & 0.01 & 0.02 \\ 
  GO:0007618 & mating &  20 &   3 & 0.45 & 0.01 & 0.02 \\ 
  GO:0009612 & response to mechanical stimulus &  58 &   5 & 1.31 & 0.01 & 0.02 \\ 
  GO:0007620 & copulation &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:0044793 & negative regulation by host of viral pro... &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:1904467 & regulation of tumor necrosis factor secr... &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
  GO:1990774 & tumor necrosis factor secretion &   7 &   2 & 0.16 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-08-17 Supplement/tex/cluster4_MF_Mm_CompInfDown.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:27:20 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0019166 & trans-2-enoyl-CoA reductase (NADPH) acti... &   3 &   2 & 0.07 & 0.00 & 0.03 \\ 
  GO:0005102 & receptor binding & 889 &  34 & 20.21 & 0.00 & 0.03 \\ 
  GO:0047391 & alkylglycerophosphoethanolamine phosphod... &   4 &   2 & 0.09 & 0.00 & 0.03 \\ 
  GO:0004190 & aspartic-type endopeptidase activity &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0031683 & G-protein beta/gamma-subunit complex bin... &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0070001 & aspartic-type peptidase activity &  15 &   3 & 0.34 & 0.00 & 0.03 \\ 
  GO:0004197 & cysteine-type endopeptidase activity &  57 &   5 & 1.30 & 0.01 & 0.03 \\ 
  GO:0016922 & ligand-dependent nuclear receptor bindin... &  20 &   3 & 0.45 & 0.01 & 0.03 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-08-17 Supplement/tex/cluster5_BP_Mm_7dpiUp.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Aug 15 15:25:59 2016
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
  \hline
GO.ID & Term & Annotated & Significant & Expected & p.value & adj.p \\ 
  \hline
GO:0002376 & immune system process & 1490 &  62 & 25.03 & 0.00 & 0.00 \\ 
  GO:0006955 & immune response & 775 &  42 & 13.02 & 0.00 & 0.00 \\ 
  GO:0002250 & adaptive immune response & 241 &  21 & 4.05 & 0.00 & 0.00 \\ 
  GO:0002682 & regulation of immune system process & 821 &  38 & 13.79 & 0.00 & 0.00 \\ 
  GO:0042981 & regulation of apoptotic process & 1026 &  43 & 17.24 & 0.00 & 0.00 \\ 
  GO:0043067 & regulation of programmed cell death & 1040 &  43 & 17.47 & 0.00 & 0.00 \\ 
  GO:0001775 & cell activation & 620 &  31 & 10.42 & 0.00 & 0.00 \\ 
  GO:0010941 & regulation of cell death & 1120 &  44 & 18.82 & 0.00 & 0.00 \\ 
  GO:0006952 & defense response & 823 &  36 & 13.83 & 0.00 & 0.00 \\ 
  GO:0050776 & regulation of immune response & 410 &  24 & 6.89 & 0.00 & 0.00 \\ 
  GO:0045321 & leukocyte activation & 543 &  28 & 9.12 & 0.00 & 0.00 \\ 
  GO:0045087 & innate immune response & 395 &  23 & 6.64 & 0.00 & 0.00 \\ 
  GO:0006915 & apoptotic process & 1301 &  47 & 21.86 & 0.00 & 0.00 \\ 
  GO:0002684 & positive regulation of immune system pro... & 531 &  27 & 8.92 & 0.00 & 0.00 \\ 
  GO:0012501 & programmed cell death & 1323 &  47 & 22.23 & 0.00 & 0.00 \\ 
  GO:0002443 & leukocyte mediated immunity & 216 &  16 & 3.63 & 0.00 & 0.00 \\ 
  GO:0002460 & adaptive immune response based on somati... & 166 &  14 & 2.79 & 0.00 & 0.00 \\ 
  GO:0006954 & inflammatory response & 369 &  21 & 6.20 & 0.00 & 0.00 \\ 
  GO:0001816 & cytokine production & 434 &  23 & 7.29 & 0.00 & 0.00 \\ 
  GO:0050896 & response to stimulus & 4563 & 110 & 76.67 & 0.00 & 0.00 \\ 
  GO:0043065 & positive regulation of apoptotic process & 439 &  23 & 7.38 & 0.00 & 0.00 \\ 
  GO:0009617 & response to bacterium & 283 &  18 & 4.75 & 0.00 & 0.00 \\ 
  GO:0043068 & positive regulation of programmed cell d... & 443 &  23 & 7.44 & 0.00 & 0.00 \\ 
  GO:0002819 & regulation of adaptive immune response & 106 &  11 & 1.78 & 0.00 & 0.00 \\ 
  GO:0001817 & regulation of cytokine production & 385 &  21 & 6.47 & 0.00 & 0.00 \\ 
  GO:0006950 & response to stress & 2248 &  65 & 37.77 & 0.00 & 0.00 \\ 
  GO:0008219 & cell death & 1414 &  47 & 23.76 & 0.00 & 0.00 \\ 
  GO:0002703 & regulation of leukocyte mediated immunit... & 135 &  12 & 2.27 & 0.00 & 0.00 \\ 
  GO:0016265 & death & 1418 &  47 & 23.82 & 0.00 & 0.00 \\ 
  GO:0002366 & leukocyte activation involved in immune ... & 163 &  13 & 2.74 & 0.00 & 0.00 \\ 
  GO:0002263 & cell activation involved in immune respo... & 165 &  13 & 2.77 & 0.00 & 0.00 \\ 
  GO:0002822 & regulation of adaptive immune response b... &  95 &  10 & 1.60 & 0.00 & 0.00 \\ 
  GO:0010942 & positive regulation of cell death & 472 &  23 & 7.93 & 0.00 & 0.00 \\ 
  GO:0016337 & single organismal cell-cell adhesion & 508 &  24 & 8.54 & 0.00 & 0.00 \\ 
  GO:0050778 & positive regulation of immune response & 318 &  18 & 5.34 & 0.00 & 0.00 \\ 
  GO:0002252 & immune effector process & 485 &  23 & 8.15 & 0.00 & 0.00 \\ 
  GO:0007159 & leukocyte cell-cell adhesion & 357 &  19 & 6.00 & 0.00 & 0.00 \\ 
  GO:0001819 & positive regulation of cytokine producti... & 263 &  16 & 4.42 & 0.00 & 0.00 \\ 
  GO:0046649 & lymphocyte activation & 467 &  22 & 7.85 & 0.00 & 0.00 \\ 
  GO:0098609 & cell-cell adhesion & 545 &  24 & 9.16 & 0.00 & 0.00 \\ 
  GO:0002449 & lymphocyte mediated immunity & 160 &  12 & 2.69 & 0.00 & 0.00 \\ 
  GO:0002285 & lymphocyte activation involved in immune... & 111 &  10 & 1.87 & 0.00 & 0.00 \\ 
  GO:0098602 & single organism cell adhesion & 551 &  24 & 9.26 & 0.00 & 0.00 \\ 
  GO:0002683 & negative regulation of immune system pro... & 278 &  16 & 4.67 & 0.00 & 0.00 \\ 
  GO:0032623 & interleukin-2 production &  50 &   7 & 0.84 & 0.00 & 0.00 \\ 
  GO:0034109 & homotypic cell-cell adhesion & 386 &  19 & 6.49 & 0.00 & 0.00 \\ 
  GO:0002694 & regulation of leukocyte activation & 324 &  17 & 5.44 & 0.00 & 0.00 \\ 
  GO:0002706 & regulation of lymphocyte mediated immuni... &  98 &   9 & 1.65 & 0.00 & 0.00 \\ 
  GO:0042110 & T cell activation & 330 &  17 & 5.54 & 0.00 & 0.00 \\ 
  GO:0070489 & T cell aggregation & 330 &  17 & 5.54 & 0.00 & 0.00 \\ 
  GO:0007259 & JAK-STAT cascade &  99 &   9 & 1.66 & 0.00 & 0.00 \\ 
  GO:0097696 & STAT cascade &  99 &   9 & 1.66 & 0.00 & 0.00 \\ 
  GO:0071593 & lymphocyte aggregation & 331 &  17 & 5.56 & 0.00 & 0.00 \\ 
  GO:0002286 & T cell activation involved in immune res... &  57 &   7 & 0.96 & 0.00 & 0.00 \\ 
  GO:0097285 & cell-type specific apoptotic process & 367 &  18 & 6.17 & 0.00 & 0.00 \\ 
  GO:0002274 & myeloid leukocyte activation & 126 &  10 & 2.12 & 0.00 & 0.00 \\ 
  GO:0070486 & leukocyte aggregation & 337 &  17 & 5.66 & 0.00 & 0.00 \\ 
  GO:0042267 & natural killer cell mediated cytotoxicit... &  41 &   6 & 0.69 & 0.00 & 0.00 \\ 
  GO:0002705 & positive regulation of leukocyte mediate... &  81 &   8 & 1.36 & 0.00 & 0.00 \\ 
  GO:0007165 & signal transduction & 2897 &  73 & 48.67 & 0.00 & 0.00 \\ 
  GO:0046425 & regulation of JAK-STAT cascade &  83 &   8 & 1.39 & 0.00 & 0.00 \\ 
  GO:1904892 & regulation of STAT cascade &  83 &   8 & 1.39 & 0.00 & 0.00 \\ 
  GO:0002228 & natural killer cell mediated immunity &  43 &   6 & 0.72 & 0.00 & 0.00 \\ 
  GO:0050865 & regulation of cell activation & 348 &  17 & 5.85 & 0.00 & 0.00 \\ 
  GO:0046427 & positive regulation of JAK-STAT cascade &  44 &   6 & 0.74 & 0.00 & 0.00 \\ 
  GO:1904894 & positive regulation of STAT cascade &  44 &   6 & 0.74 & 0.00 & 0.00 \\ 
  GO:0002237 & response to molecule of bacterial origin & 190 &  12 & 3.19 & 0.00 & 0.00 \\ 
  GO:0051716 & cellular response to stimulus & 3830 &  90 & 64.35 & 0.00 & 0.00 \\ 
  GO:0042327 & positive regulation of phosphorylation & 613 &  24 & 10.30 & 0.00 & 0.00 \\ 
  GO:0002697 & regulation of immune effector process & 288 &  15 & 4.84 & 0.00 & 0.00 \\ 
  GO:0007155 & cell adhesion & 826 &  29 & 13.88 & 0.00 & 0.00 \\ 
  GO:0001934 & positive regulation of protein phosphory... & 583 &  23 & 9.80 & 0.00 & 0.00 \\ 
  GO:0042088 & T-helper 1 type immune response &  30 &   5 & 0.50 & 0.00 & 0.00 \\ 
  GO:0042269 & regulation of natural killer cell mediat... &  30 &   5 & 0.50 & 0.00 & 0.00 \\ 
  GO:0048518 & positive regulation of biological proces... & 3592 &  85 & 60.35 & 0.00 & 0.00 \\ 
  GO:0022610 & biological adhesion & 835 &  29 & 14.03 & 0.00 & 0.00 \\ 
  GO:0002715 & regulation of natural killer cell mediat... &  31 &   5 & 0.52 & 0.00 & 0.00 \\ 
  GO:0010562 & positive regulation of phosphorus metabo... & 672 &  25 & 11.29 & 0.00 & 0.00 \\ 
  GO:0045937 & positive regulation of phosphate metabol... & 672 &  25 & 11.29 & 0.00 & 0.00 \\ 
  GO:0001909 & leukocyte mediated cytotoxicity &  70 &   7 & 1.18 & 0.00 & 0.00 \\ 
  GO:0002707 & negative regulation of lymphocyte mediat... &  32 &   5 & 0.54 & 0.00 & 0.00 \\ 
  GO:0001910 & regulation of leukocyte mediated cytotox... &  50 &   6 & 0.84 & 0.00 & 0.00 \\ 
  GO:0044700 & single organism signaling & 3084 &  75 & 51.82 & 0.00 & 0.00 \\ 
  GO:0019221 & cytokine-mediated signaling pathway & 206 &  12 & 3.46 & 0.00 & 0.00 \\ 
  GO:0043403 & skeletal muscle tissue regeneration &  18 &   4 & 0.30 & 0.00 & 0.00 \\ 
  GO:0023052 & signaling & 3091 &  75 & 51.93 & 0.00 & 0.00 \\ 
  GO:0050777 & negative regulation of immune response &  96 &   8 & 1.61 & 0.00 & 0.00 \\ 
  GO:0007166 & cell surface receptor signaling pathway & 1389 &  41 & 23.34 & 0.00 & 0.00 \\ 
  GO:0048584 & positive regulation of response to stimu... & 1207 &  37 & 20.28 & 0.00 & 0.00 \\ 
  GO:0009605 & response to external stimulus & 1212 &  37 & 20.36 & 0.00 & 0.00 \\ 
  GO:0043312 & neutrophil degranulation &   8 &   3 & 0.13 & 0.00 & 0.00 \\ 
  GO:0007154 & cell communication & 3168 &  76 & 53.23 & 0.00 & 0.00 \\ 
  GO:0032612 & interleukin-1 production &  54 &   6 & 0.91 & 0.00 & 0.00 \\ 
  GO:0009611 & response to wounding & 424 &  18 & 7.12 & 0.00 & 0.00 \\ 
  GO:0048583 & regulation of response to stimulus & 2241 &  58 & 37.65 & 0.00 & 0.00 \\ 
  GO:0001906 & cell killing &  77 &   7 & 1.29 & 0.00 & 0.00 \\ 
  GO:0002685 & regulation of leukocyte migration & 102 &   8 & 1.71 & 0.00 & 0.00 \\ 
  GO:0035556 & intracellular signal transduction & 1613 &  45 & 27.10 & 0.00 & 0.00 \\ 
  GO:0050707 & regulation of cytokine secretion & 104 &   8 & 1.75 & 0.00 & 0.00 \\ 
  GO:0002699 & positive regulation of immune effector p... & 131 &   9 & 2.20 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

### 2016-08-18 0_oocystCounts_and_qPCR.R
qPCR plot and 0_oocyst... script updated. Plot needs small incscape-fix.

@@ -8,9 +8,7 @@ library(reshape2)
 library(plyr)
 library(doBy)
 library(stringr)
-library(svglite)
-
-
+library(RSvgDevice)
 
 setwd("~/Ef_RNAseq/")
 phen.data <- read.csv("data/Oocysts_output_weight_SS_longdata.csv")
@@ -44,10 +42,8 @@ all.oocysts.line +
     theme_bw()
 dev.off()
 
-
 ## qPCR data from Simone and Annica
-
-## Ef18S qPCR data for different timepoints, 1st and 2nd infection and 3 replicates.
+## Ef18S qPCR data for different timepoints, 1st and 2nd infection and 3 biological replicates.
 qpcr.df <- as.data.frame(read.csv("data/NMRI1st2ndqPCR_frSS_EfctRef_long.csv"))
 long.qpcr <- melt(qpcr.df, measure.vars = c("rep1_ct", "rep2_ct", "rep3_ct"))
 ## Normalise to highest ct among all replicates
@@ -57,7 +53,7 @@ long.qpcr$norm.value <- (long.qpcr$value - max(long.qpcr$value))*-1
 ## Calculate sd and means on normalised values
 stats.qpcr <- summaryBy(norm.value ~ dpi + inf,
           data = long.qpcr, 
-          FUN = list(mean, min, max, sd))
+          FUN = list(mean, min, max, sd, se))
 stats.qpcr <- merge(stats.qpcr, long.qpcr)
 
 ################ STATUS ####################
@@ -70,22 +66,28 @@ stats.qpcr <- merge(stats.qpcr, long.qpcr)
 ## Cool that we have so much qPCR data: find those genes in RNAseq data and 
 ## plot together
 
+devSVG("~/Ef_RNAseq/figures/Figure1c_qPCR18S.svg")
 qpcr18S <- ggplot(subset(stats.qpcr, stats.qpcr$gene %in% "Ef18S"), 
        aes(x = dpi, y = norm.value.mean, col = factor(inf))) +
-  geom_errorbar(aes(ymin = norm.value.min,
-                    ymax = norm.value.max,
+  ggtitle("Eimeria 18S by qPCR") +
+  labs(color = "Infection No.") +
+  ## to get rid of label, rm aes(), but then size is not controlled. How fix this? ################ HERE ####################
+  geom_errorbar(aes(ymin = norm.value.mean - norm.value.sd,
+                    ymax = norm.value.mean + norm.value.sd,
                     width = 0.2)) +
[…]

### 2016-08-18 0_oocystCounts_and_qPCR.R
working on qPCR and oocyst plots - harmonizing etc

@@ -53,7 +53,7 @@ long.qpcr$norm.value <- (long.qpcr$value - max(long.qpcr$value))*-1
 ## Calculate sd and means on normalised values
 stats.qpcr <- summaryBy(norm.value ~ dpi + inf,
           data = long.qpcr, 
-          FUN = list(mean, min, max, sd, se))
+          FUN = list(mean, min, max, sd))
 stats.qpcr <- merge(stats.qpcr, long.qpcr)
 
 ################ STATUS ####################

### 2016-08-18 0_oocystCounts_and_qPCR.R
fixed non-unique mouse-IDs in oocyst data

@@ -13,7 +13,6 @@ library(RSvgDevice)
 setwd("~/Ef_RNAseq/")
 phen.data <- read.csv("data/Oocysts_output_weight_SS_longdata.csv")
 
-## Mouse IDs are not unique so it is impossible to track individuals FIX THIS!!
 ## here just a hack to take the first occurence of a ID at a time point
 phen.data <- phen.data[!duplicated(phen.data[, c("Day_pi", "Mouse_ID", "Infection_No")]), ]
 

### 2016-09-05 7_enrichment.R
7_ : added Eimeria clusters of interest and 1 mouse cluster to enrichment output

@@ -1,5 +1,7 @@
 library(topGO)
 library(xtable)
+library(gridExtra) #  for color theme function
+library(grid)      # for color theme function
 
 annotation.frame <- read.table("output_data/annotation_data.csv", sep=",")
 
@@ -47,10 +49,21 @@ set.from.cluster <- function(hcluster, number){
 ######## Clustering #############
 to.test <- list(
     ## Eimeria
-    ## fill me in!!!
+    list(set=set.from.cluster(hcluster[["Ef"]], 1),
+         type = "oocystUp1", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 5),
+         type = "oocystUp2", species = "Ef"),
+    list(set=set.from.cluster(hcluster[["Ef"]], 7),
+         type = "lateWeak", species = "Ef"),
+    list(set=set.from.cluster(hcluster[["Ef"]], 2),
+         type = "lateStrong", species = "Ef"),
+    list(set=set.from.cluster(hcluster[["Ef"]], 4),
           type = "sporoz", species = "Ef"),
+    list(set=set.from.cluster(hcluster[["Ef"]], 6),
+         type = "earlyUp", species = "Ef"),
     ## Mouse
+    list(set=set.from.cluster(hcluster[["Mm"]], 6),
+         type="2ndInf", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 5),
          type="7dpiUp", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 3),

### 2016-09-05 manuscript/bmc_article.tex
manuscript: changes in results (adjustments) and addition of heatmap introduction. GO-discussion not included

@@
 \usepackage{amsthm,amsmath}
+%\usepackage{asmath}
 \RequirePackage{natbib}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 sporozoites were included as ``environmental'' stages and processed
-``\textit{in vitro}''. To follow the lifecycle of the parsite, we
-compare different timepoints after infection.
-
-We additonally used different mouse strains which displayed different
-immunocompetence in infection trials to assess the influence host
+``\textit{in vitro}''. To follow the life cycle of the parasite, we
+compare different time-points after infection.
+
+We additionally used different mouse strains which display different
+immunocompetence in infection trials measured by oocyst output to assess the influence of host
 immunocompetence on parasite development (figure 1a,b). In these
-experiments we analysed RAG-1-/- mice and compared them to the
-parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
-response du to the absence of mature B and T lymphocytes.  All three
-mouse strains were used for infections of both naive and previously
-challenge infected animals.
-
-Concerning basic phenotyping, differences in oocyst output were
-visible both between the immunocompetent (NMRI and C57BL6) and
-immunodeficient (RAG-1-/-) host strain, but also between the two
-similarly immunocompetent (NMRI and C57BL6) strains (figure
-1a). Surprisingly this finding was consistent in both in primary and
-challenge infection. Similarly development of intestinal stages as
-assayed by quantitative reverse transcription PCR (RT qPCR) also
-... ADD HERE (figure 1b).
-
-We thus aim to analyse the lifecycle of \textit{E. falciformis} under
-the influence of (most likely the innate or cell autonomouse) immune
-respons at on early stages of infection. We used an experimental
-design comparing infections at 5dpi for all possible experimental
-conditions (all three mouse strains in naive and challenge infection),
-and analyse additional infections at 3dpi (figure 1c).
-
-In our RNAseq experiment we worked with two biological replicates for
-all but two conditions (with one and three replicates,
-respecitively). For each replicate sample we enriched caecum tissue
-from three infected mice for epithelial cells, extracted mRNA and
-constructed sequencing libraries. These libraries were sequenced on
-several lanes of Illumina GAIIX and HiSeq machines and mapped to both
-mouse and parasite genome simulatneously to avoid spurious assignments
-of read in ultra conserved genomic regions.
-
-The number of total read mappings obtained for individual replicates
-ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
-(NMRI\_2ndInf\_5dpi\_rep1). As samples and idividual replicates were
-sequenced in batches to different depth and using different
-instrumentation (Table 1) we performed quality controls confirming the
-abscence of batch effects influencing analysis and quality of our
-results (additional file xyc). The results of these multivariate
-analyses in relation to biological interpretation are described
-further below.
-
-\section*{Parasite and host transcriptomes can be assessed in parallel}
-
+experiments we analysed mice lacking the Recombination activating gene 1, Rag1, and compared them to the
+parental C57BL/6 mouse line. Rag1 knock-out mice lack adaptive immune
+response due to the absence of mature B and T lymphocytes.  All three
+mouse strains were used for infections of both naive mice and previously
+infected animals (onward referred to as "challenged").
+
+Basic phenotyping showed differences in oocyst numbers
+between the immunocompetent (C57BL/6) and
+immunodeficient (Rag1-/-) host strain in infection of naive mice and 
+in challenge infections. Immunocompetent NMRI mice were infected with a higher dosis  
+of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice
+compared to naive mice (Figure 1a). Oocyst numbers in faeces peaked on days 8-9 and 
+all mice had cleared the infection by day 14. Development of \textit{E. falciformis} intestinal stages is reflected by 
+the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S  (Figure 1c).
+
+We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
+the influence of host immune responses at early stages of infection. We used an experimental
+design which allows to compare infections at 5 days post infection (dpi) for all experimental
+conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). In addition,
+additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).
+
+\section*{Parasite and host dual transcriptomes can be assessed in parallel}
+In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue which was pooled 
+from three mice. mRNA was extracted and
+sequencing libraries prepared. Two biological replicates were used for
+all but two conditions (with one and three replicates, respectively). These libraries were sequenced on
+several lanes of Illumina GAIIX (13 samples) and HiSeq machines (14 samples) and mapped to both
+mouse and parasite genomes simultaneously to avoid spurious assignments
+of reads in ultra conserved genomic regions. 
+As samples and individual replicates were sequenced in batches to different depth and using different
+instrumentation (Table 1) we performed quality controls (additional files xyz). These confirm
+the absence of batch effects influencing analysis and quality of 
+results.
+
+Total numbers of sequenced reads as well as reads
+mapped to either the \textit{E. falciformis} genome or the
+mouse genome are indicated in Table 1 for all replicates. The number of total read mappings 
+for individual replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
+(NMRI\_2ndInf\_5dpi\_rep1). At the latest time-point samples, 7 dpi, the overall mRNA
+output of the sampled caecum tissue is dominated by parasite material with proportional mRNA
+abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in
+NMRI\_1stInf\_7dpi\_rep1). (Figure 1b).
+
+\section*{Exclusion of samples with uncertain infection status}
 A maximum of 86\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of
 0.0326\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be
-assigened to the \textit{E. falciformis} genome in samples considered
-infected (table 1).
-
-We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
-due to an exptremely low parasite contribution to the overall
-transcriptome. Technically this exclusion made it possible to obtain
-read counts in agreement with a negative bionmial distributions (see
-adddtional file x). Judging from their mRNA output those samples can
-be considered non-infected, as both samples occured in a challenge
-infection we assume that the infection was cleard or reduced to a
-non-detectable output in those animmals.
-
-One sample (NMRI\_1stInf\_0dpi\_rep1) was additonally excluded because
-the uninfected control showed an unexpected mapping of reads to the
-\textit{E. falciformis} genome. We consider this together with the
-unsuccessfully infected samples above to display an undecided state of
-infection.
-
-On the other end of the spectrum mice not previously immunized
-featured an mRNA ``holo-transcriptome'' at 7dpi with the majority of
-mRNA originating from the parasite as opposed to the host (Figure
-1). The overall mRNA output of the sampled caecum tissue at this point
-of infection is dominated by the parasite with proportional mRNA
-abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and even 92\% (in
-NMRI\_1stInf\_7dpi\_rep1).
-
-The total number of sequencing reads obtained as well as those reads
-mapped to the transcriptome of either \textit{E. falciformis} and the
-mouse are indicated in Table 1 for all replicates. We additonally
-assessed the number of genes considered ``expressed'', further
-analysis and discussion considers the possibility that especially for
-\textit{E. falciformis} the number of reads might have been
-insufficient to fully normalize the dataset of the samples with the
-lowest representation against those with the hightest.
-
-\section*{The mouse transcriptome changes in the course of infection infection}
-
-Statistical testing for differetial expression between infected and
-uninfected mice, revealed changes in mRNA abundance becoming more
-pronounced (involving more genes) at later post infection (especially
-at 7dpi; Table 2).
-
-The observed changes between controls and 7dpi were in agreement with
+assigned to the \textit{E. falciformis} genome in samples considered
+infected (Table 1). We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
+due to low parasite contribution to the overall transcriptome. 
+Technically, this exclusion made it possible to obtain read counts in agreement with a 
+negative binomial distribution (see additional file x). It is also likely that the number of reads in the excluded samples would have been
+insufficient to fully normalise these datasets to those with the highest parasite contributions.
+From a biological point of view, both excluded are samples from challenge
+infection and we assume that the infection was cleared or reduced to a
+non-detectable level. One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because
+the uninfected control showed unexpected mapping of reads to the
+\textit{E. falciformis} genome. We consider the three excluded samples to display an uncertain 
+state of infection.
+
+\section*{The mouse transcriptome changes in the course of infection}
+Statistical testing for differential expression between infected and
+uninfected mice revealed changes in mRNA abundance becoming more
+pronounced (involving more genes) at later time-points post infection (especially
+at 7 dpi; Table 2).
+
+The observed changes between controls and 7 dpi were in agreement with
 previously published microarry data. Fold-Change data obtained from
-mice at 6dpi on Agilent microarrays (Schmidt et al 2012) or 7dpi (our
-RNAseq data) against uninfected controls shows a strong correlateion
-(Spearman's \rho = 0.74; figure 2a). Considering both and technical
-differences between the tow methods and the biological changes along
-the lifecycle this confirms the adequacy of our methods for assessing
-the host transcriptome.
-
-325 mouse genes were considered differently expressed (DE; FDR<1\%)
-between uinfected and 3dpi, 1,804 between controls and 5dpi and 2,711
-genes between 0dpi and 7dpi. This lead to combined set of 3453 genes
-reacting to infection (figure 2bi).  DE genes early in infection (3dpi
-and and 5dpi) were not simply a subset of genes becoming DE later in
-infection (7dpi). To further validate this pattern of a distinct
+\textit{E. falciformis} infected mice at 6 dpi on Agilent microarrays 
+(Schmidt et al 2012) or 7 dpi (our RNAseq data) against uninfected controls
+show a strong correlation (Spearman's \rho = 0.74; Figure 2a). Considering both biological 
+(e.g. exact time-points) and technical differences between the two methods this comparison 
+confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.
+
+For mouse, 325 mRNAs were considered differently abundant (DA; FDR<1\%)
+between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711
+mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 genes
+responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi
+and 5 dpi) were not simply a subset of genes DA later in
+infection (7 dpi), but instead, the transcriptional profile of the mouse 
+changes throughout the time-points tested here. To further validate this pattern of a distinct
 response in early infection we performed hierarchical clustering on
-(the union of) mouse genes DE between different dpi (figure 2c).
-
-The ponounced changes later in infection are reflecting the expected
-onset of an adaptive immune response. This is underlined by GO-terms
-such as ``antigen binding'' and ``immunoglobulin receptor binding''
-(molecular function) ``immune system process'' and ``adaptive immune
-response'' (biological process) highly enriched in Cluster 5 of figure
-2c.
-
-Particular ``regulation of cell death'' and ``regulation of apoptotic
-process'' (BP)
-
-``UDP-N-acetylglucosamine-lysosomal-enzyme'' (MF)
+the (union of) mouse genes DA between different time-points post infection (Figure 2c).
+
+By hierarchical clustering of mouse samples, three main sample clusters formed, “main clusters” 
+(dendrogram of columns at top of Figure 2c).
+Rag1-/- mice cluster with control samples, and there is no clear distinction between infected and non-infected Rag1-/- samples, confirming the immune deficiency phenotype in Rag1-/-.
+Control samples are defined by high abundance of transcripts in gene cluster 3 (yellow, clusters of rows on the left) and partly high abundance in cluster 4 (red).
+Immunodeficient Rag1-/- samples share this profile with control samples in gene cluster 4, and this cluster has low abundance in (most) other infection samples.
+This suggests that these are genes which become down-regulated in immune competent mice upon infection.
+%%Two naive 5 dpi samples to the left cluster separately and diverge from this and most other patterns. High in infection, but does not fit with third NMRI replicate....
+Three samples of challenged mice from 3, 5, and 7 dpi display high abundance of transcripts in gene cluster 6 (blue).
+The same genes have a low abundance in two naive, immune competent mouse samples from 5 dpi.
+Genes in cluster 6 are therefore particularly interesting for investigating the role of adaptive immunity in this infection (although sample NMRI\_2ndInf\_5dpi\_rep1 has a different profile).
+%% GO terms in cluster 6 discuss
+
+Transcripts in gene clusters 3 and 4 are lowly abundant among all samples in the middle main cluster.
+Among those samples, the NMRI mouse samples display a trend towards high abundance in clusters 1, 2, 5 (green) and 7.
+Especially in the two late infection samples (7 dpi) a high transcript abundance is seen in clusters 1 and 5.
+The leftmost main sample cluster containing immunocompetent 5 dpi samples is characterized by low transcript abundance in cluster 6, and high abundance in cluster 7.
+
+The pronounced changes later in infection reflect the expected
+onset of an adaptive immune response. This is underlined by enriched Gene Ontology, GO, terms
+in gene clusters with pronounced profiles late in infection (Figure 2c).
+Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process" and "adaptive immune
+response" (biological process, BP) are highly enriched in Cluster 5 and confirm an activated adaptive immune response
+at this time-point.
+
+Particular "regulation of cell death" and "regulation of apoptotic process" (BP)
+
+"UDP-N-acetylglucosamine-lysosomal-enzyme" (MF)
 
@@
 
-General changes upon infection are a downregulation of genes involved
-``lipid modification'' or more detailed ``phosphatidylinositol
-phosphorylation''. These changes are similarly pronounce in the
-immunocompetent strains and RAG-/- immunodefficient animals (cluster 3
+General changes upon infection are a down-regulation of genes involved
+"lipid modification" or more detailed "phosphatidylinositol
+phosphorylation". These changes are similarly pronounce in the
+immunocompetent strains and Rag1-/- immunodefficient animals (cluster 3
 of fibure 2c) (see supplementary table X).  The presence of these
-changes immunocompromized animals invite the speculation that those
+changes immunocompromised animals invite to the speculation that those
 genes are independent of a potential later adaptive immune response.
 
-Changes not detectable in RAG-/- immunodefficient animals (cluster 3
+Changes not detectable in Rag1-/- immunodefficient animals (cluster 3
 of fibure 2c) but otherwise observed in general infection are
-characterized by and enrichment to genes involved in ``positive
-regulation of calcineurin-NFAT'', ``blood coagulation'' and ``inositol
-phosphate-mediated signaling'' (see supplementary table X). The
-abscence of these changes in RAG-/- mice allow us to speculate that
-those genes might be involved in a very early onlset of adaptive
-immune response.
+characterised by and enrichment to genes involved in "positive
+regulation of calcineurin-NFAT", "blood coagulation" and "inositol
+phosphate-mediated signaling" (see supplementary table X). The
+absence of these changes in Rag1-/- mice allow us to speculate that
+those genes might be involved in a early onset of adaptive
+immune responses.
 
@@
 
-We dedtected further differences between infections of naive and
-challenge infected mous hosts , especially at 7 dpi. These genes are
+We detected further differences between infections of naive and
+challenge infected mouse hosts, especially at 7 dpi. These genes are
 characterized by...
@@
 
-\section*{The lifecycle of Eimeria is characterized by major changes in mRNA abundance}
+\section*{The life cycle of E. falciformis is characterized by major changes in mRNA abundance}
 
 Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison.  For
-\textit{E. falciformis} no mRNAs are significantly differently
-abundant between first and second infection.
-
-Changes throughout the intr-host lifecycle can be clearly categorized
-in an early (3 and 5 dpi) and late (7dpi) phase. Sporozoites show
-expression profiles similar more similar to early infection stages
-while oocysts show distinctly different mRNA abundance profile (MDS
-plot .
-
-We applied hierarchical clustering on \textit{E. falciformis} mRNAs
-which were differently abundant at different dpi, including
-comparisons with oocysts and sporozoites.
-
-Samples from 7dpi form a separate cluster (NMRI mice only), not
-distinguishing between first and challenge infection.
+differently abundant mRNAs per comparison. We applied hierarchical clustering on 
+\textit{E. falciformis} mRNAs which were differently abundant at different time-points 
+post infection, including comparisons with oocysts and sporozoites. 
+Upon hierarchical clustering (Figure 3c) changes throughout the parasite's intracellular life cycle 
+can be categorised into an early (3 and 5 dpi) and late (7dpi) phase. 
+Mouse immune status has no apparent influence on clustering of
+parasite samples. Sporozoites show expression profiles similar to early infection stages
+while oocysts show a distinct mRNA abundance profile (MDS plot...).
+
+Parasite gene clusters shown on the left side (colour-coded and numbered 1 - 7) can be used 
+to characterise clusters of samples.
+Late in infection, a profile of high mRNA abundance in clusters 2 and 7 (mint green and purple) is seen.
+Also parasite genes in cluster 6 (blue) appear characteristic  by displaying low mRNA abundance in late stages of infection.
+ One late challenge infection sample deviates from this pattern together with an early challenge infection sample.
+These two samples are, apart from controls and excluded samples, the ones with the least number of detected parasite reads (1836 and 1580, respectively), which taken together suggests that infections had been cleared before sampling in these challenge infected mice.
+
+The early infection sample cluster (12 samples in center of heatmap) is characterised by a high abundance of transcripts in clusters 3 (yellow) and 6, as well as a below average abundance in clusters 1 (pink), 2, and 5 (green).
+Sporozoite samples have highly abundant transcripts in cluster 4 (red) and low abundance of 
+transcripts in gene cluster 7. The latter is shared with oocyst samples. The oocyst samples, 
+in addition, have low abundance of transcripts in gene clusters 2, 3, 4, 6 (and 7) and highly 
+abundant transcripts in clusters 1 and 5.
+
+Taken together, sample and gene clustering indicates that genes in clusters 2, 6 and 7 might be 
+useful to characterise assumed gametocytes in 7 dpi samples. Genes in cluster 4 are candidates 
+for sporozoite specific genes, and clusters 1 and 5  for oocyst specific genes in E. falciformis.
 
@@
 in \textit{T. gondii}
-
-\subsubsection*{Data overview}
-In total, xx reads were sequenced. For each sample, mouse read numbers were $10^7$ - $10^8$
-Quality filters applied resulted in xx reads being removed from further analysis
 

### 2016-09-05 7_enrichment.R
changes to enrichment to export tables, plus GO-tables for interesting clusters as readable pdf:s

@@ -63,9 +63,9 @@ to.test <- list(
          type = "earlyUp", species = "Ef"),
     ## Mouse
     list(set=set.from.cluster(hcluster[["Mm"]], 6),
-         type="2ndInf", species="Mm"),
+         type="secondInfInf", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 5),
-         type="7dpiUp", species="Mm"),
+         type="SevendpiUp", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 3),
           type="allInfDown", species="Mm"),
     list (set=set.from.cluster(hcluster[["Mm"]], 4),
@@ -120,7 +120,7 @@ names(BPMF.ll) <- unlist(lapply(to.test, "[[", 2))
 ##colors for table
 myt <- ttheme_default(
   base_size = 18,
-  padding = unit(c(6, 6), "mm"),
+  padding = unit(c(2, 6), "mm"),
   # Use hjust and x to left justify the text
   # Alternate the row fill colours
   core = list(fg_params=list(col="dark green"), #, hjust = 1, x=1),
@@ -130,6 +130,18 @@ myt <- ttheme_default(
   colhead = list(fg_params=list(col="dark green"),
                  bg_params=list(fill="gray"))
 )
+
+## Change filename and part of BPMF.ll object to export tables 
+## (adjust size of PDF for better readability)
+pdf("~/Ef_RNAseq/additional_files/SI_GOMmSevendpiUp.pdf", width = 24, height = 48)
+oocysts1 <- grid.table(data.frame(BPMF.ll$SevendpiUp$BP),
+                               theme = myt,
+                               rows = NULL,
+                               cols = c("GO id", "Term", 
+                                        "Annot. genes", 
+                                        "Sign. genes", 
+                                        "Expected", "P-value", "Adj. P-value"))
+dev.off()
 
 
 ### NOT tested from here on...

### 2016-09-08 7_enrichment.R
added all clusters and renamed 'intuitive' names of clusters for enrichment - they were not so intuitive...

@@ -61,6 +61,8 @@ to.test <- list(
           type = "sporoz", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 6),
          type = "earlyUp", species = "Ef"),
+    list(set=set.from.cluster(hcluster[["Ef"]], 3),
+         type = "cluster3", species = "Ef"),
     ## Mouse
     list(set=set.from.cluster(hcluster[["Mm"]], 6),
          type="secondInfInf", species="Mm"),
@@ -70,6 +72,12 @@ to.test <- list(
           type="allInfDown", species="Mm"),
     list (set=set.from.cluster(hcluster[["Mm"]], 4),
           type="CompInfDown", species="Mm"),
+    list (set=set.from.cluster(hcluster[["Mm"]], 1),
+          type="cluster1", species="Mm"),
+    list (set=set.from.cluster(hcluster[["Mm"]], 2),
+          type="cluster2", species="Mm"),
+    list (set=set.from.cluster(hcluster[["Mm"]], 7),
+          type="cluster7", species="Mm"),
     list (set=set.from.DE(Mm.DE.test, "N5vsN7"),
           type="DE_N5vsN7", species="Mm")
     )

### 2016-09-08 7_enrichment.R
7_ again, renamed and added clusters (perhaps not correct version in previous commit)

@@ -50,34 +50,34 @@ set.from.cluster <- function(hcluster, number){
 to.test <- list(
     ## Eimeria
     list(set=set.from.cluster(hcluster[["Ef"]], 1),
-         type = "oocystUp1", species = "Ef"),
+         type = "cluster1ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 5),
-         type = "oocystUp2", species = "Ef"),
+         type = "cluster5ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 7),
-         type = "lateWeak", species = "Ef"),
+         type = "cluster7ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 2),
-         type = "lateStrong", species = "Ef"),
+         type = "cluster2ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 4),
-          type = "sporoz", species = "Ef"),
+          type = "cluster3ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 6),
-         type = "earlyUp", species = "Ef"),
+         type = "cluster6ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 3),
-         type = "cluster3", species = "Ef"),
+         type = "cluster3ef", species = "Ef"),
     ## Mouse
     list(set=set.from.cluster(hcluster[["Mm"]], 6),
-         type="secondInfInf", species="Mm"),
+         type="cluster6mm", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 5),
-         type="SevendpiUp", species="Mm"),
+         type="cluster5mm", species="Mm"),
     list(set=set.from.cluster(hcluster[["Mm"]], 3),
-          type="allInfDown", species="Mm"),
+          type="cluster3mm", species="Mm"),
     list (set=set.from.cluster(hcluster[["Mm"]], 4),
-          type="CompInfDown", species="Mm"),
-    list (set=set.from.cluster(hcluster[["Mm"]], 1),
-          type="cluster1", species="Mm"),
+          type="cluster4mm", species="Mm"),
+    #list (set=set.from.cluster(hcluster[["Mm"]], 1), # no enriched terms
+    #      type="cluster1mm", species="Mm"),
     list (set=set.from.cluster(hcluster[["Mm"]], 2),
[…]

### 2016-09-08 manuscript/bmc_article.tex
manuscript changes - Eimeria clustering and enrichment discussed. Still without ortholog and conservation discussion

@@
 
-\section*{The mouse transcriptome changes in the course of infection}
+\section*{The mouse transcriptome changes upon \textit{E. falciformis} infection}
 Statistical testing for differential expression between infected and
@@
 pronounced (involving more genes) at later time-points post infection (especially
-at 7 dpi; Table 2).
-
-The observed changes between controls and 7 dpi were in agreement with
-previously published microarry data. Fold-Change data obtained from
+at 7 dpi; Table 2). For mouse, 325 mRNAs were considered differently abundant (DA; FDR\<1\%)
+between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711
+mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 genes
+responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi
+and 5 dpi) were not a pure subset of genes DA later in
+infection (7 dpi). Instead, the transcriptional profile of the mouse 
+changes throughout the infection. Changes between controls and 7 dpi were in agreement with
+previously published microarry data. Fold-change data obtained from
 \textit{E. falciformis} infected mice at 6 dpi on Agilent microarrays 
@@
 
-For mouse, 325 mRNAs were considered differently abundant (DA; FDR<1\%)
-between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711
-mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 genes
-responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi
-and 5 dpi) were not simply a subset of genes DA later in
-infection (7 dpi), but instead, the transcriptional profile of the mouse 
-changes throughout the time-points tested here. To further validate this pattern of a distinct
+\section*{Epithelial responses depend on mouse immune status}
+To further validate the pattern of a distinct
 response in early infection we performed hierarchical clustering on
 the (union of) mouse genes DA between different time-points post infection (Figure 2c).
-
 By hierarchical clustering of mouse samples, three main sample clusters formed, “main clusters” 
-(dendrogram of columns at top of Figure 2c).
-Rag1-/- mice cluster with control samples, and there is no clear distinction between infected and non-infected Rag1-/- samples, confirming the immune deficiency phenotype in Rag1-/-.
-Control samples are defined by high abundance of transcripts in gene cluster 3 (yellow, clusters of rows on the left) and partly high abundance in cluster 4 (red).
-Immunodeficient Rag1-/- samples share this profile with control samples in gene cluster 4, and this cluster has low abundance in (most) other infection samples.
-This suggests that these are genes which become down-regulated in immune competent mice upon infection.
-%%Two naive 5 dpi samples to the left cluster separately and diverge from this and most other patterns. High in infection, but does not fit with third NMRI replicate....
-Three samples of challenged mice from 3, 5, and 7 dpi display high abundance of transcripts in gene cluster 6 (blue).
-The same genes have a low abundance in two naive, immune competent mouse samples from 5 dpi.
-Genes in cluster 6 are therefore particularly interesting for investigating the role of adaptive immunity in this infection (although sample NMRI\_2ndInf\_5dpi\_rep1 has a different profile).
-%% GO terms in cluster 6 discuss
-
-Transcripts in gene clusters 3 and 4 are lowly abundant among all samples in the middle main cluster.
-Among those samples, the NMRI mouse samples display a trend towards high abundance in clusters 1, 2, 5 (green) and 7.
-Especially in the two late infection samples (7 dpi) a high transcript abundance is seen in clusters 1 and 5.
-The leftmost main sample cluster containing immunocompetent 5 dpi samples is characterized by low transcript abundance in cluster 6, and high abundance in cluster 7.
-
-The pronounced changes later in infection reflect the expected
-onset of an adaptive immune response. This is underlined by enriched Gene Ontology, GO, terms
-in gene clusters with pronounced profiles late in infection (Figure 2c).
-Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process" and "adaptive immune
-response" (biological process, BP) are highly enriched in Cluster 5 and confirm an activated adaptive immune response
-at this time-point.
-
-Particular "regulation of cell death" and "regulation of apoptotic process" (BP)
-
+(dendrogram of columns at top of Figure 2c). Immune deficient Rag1-/- mice cluster with control samples. 
+There is no clear distinction between infected and non-infected Rag1-/- samples, which confirms the immune deficiency phenotype in these mice.
+
+We identify a group of genes (cluster 4, Figure 2c) which change the transcriptional profile upon infection (all time-points) in immune competent mice only.
+In immune compromised Rag1-/- another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in a T and B cell dependent manner.
+Hence, these genes in some way depend on functional T and B cells also in controls (cluster 3) at the earliest time-point post infection (3 dpi, cluster 4).
+Gene Ontology, GO, terms enriched and regulated in cluster 3 are, e.g., "lipid modification" and "lipid metabolic process" and "protein intraciliary transport".
+Other enriched categories in cluster 3 are terms for regulation of different metabolic processes, for blood coagulation as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
+(SI file x).
+We suggest that these are processes which are all intrinsically different in Rag1-/- mice compared to immune competent mice.
+The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/-.
+Infected immune competent mice have a lower abundance of these mRNAs, but in Rag1-/- there is no difference compared to controls.
+Therefore, these genes appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, surprisingly also at 3 dpi.
+Enriched GO terms in this gene cluster include 13 terms for cytokines, particularly interleukin-1, IL-1, ("negative regulation of" and "secretion of").
+Terms such as "negative regulation of viral (or inflammatory) response", "negative chemotaxis", "autophagy", "blood coagulation", "inositol phosphate-mediated signaling", and "positive regulation of calcineurin-NFAT" are also enriched in cluster 4.
+Inositol signaling can lead to release of calcium and calcineurin-dependent translocation of NFAT to the nucleus and activation of its target genes in T cells, but also many other cell types (reviewed by macian05).
+GO enrichment of cluster 4 also highlights particular regulation of transforming growth factor \beta, TGF\beta, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
+TGF\beta is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10).
+TNF appears to be dose-dependent and capable to suppress inflammatory responses (noti10) but is also reported to regulate proliferation of epithelial cells (kaiser97).
+Additionally, IL-1 and IL-6 are among the enriched GO terms in cluster 4.
+The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influence lymphocyte activity (dinarello09).
+IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14) and it is known to be important for development of Th17 responses (ref in kuhn intro).
+\\\\\\\\Th17 ....
+Stange's work (discuss Richard).
+IL-6 induces the Janus kinase, JAK, and signal transducer and activator of transcription STAT3, which, among other functions, is involved in wound healing in intestinal epithelial cells (pickert09).
+This analysis suggests that TGF\beta, TNF, EGF, IL-1 and IL-6 are main actors in the epithelial response to \textit{E.
+falciformis} infection and that the response is T and B cell dependent, also at early time-points.
+The pronounced changes late in infection (7 dpi) also reflect the expected onset of an adaptive immune response.
+This is underlined by enriched GO terms in gene clusters 1 and 5.
+Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process", "adaptive immune response"and "innate immune response" (biological process, BP) are highly enriched in cluster 5 and confirm an activated adaptive immune response at this time-point.
+Among the same genes, natural killer cell regulation, JAK-STAT signalling, and IL-1 and interleukin-2, IL-2 production are enriched biological processes.
+IL-2 is one target of NFAT signaling and as mentioned above, JAK-STAT signaling can be induced by IL-6.
+This indicates that the early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi).
+
+Taken together, clusters 3 and 4 seem to capture the difference between Rag1-/- mice and immune competent mice in this parasitic infection.
+Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3).
+These do not change upon infection in mice lacking mature T and B cells.
+Other genes (cluster 4) for which the transcriptional profile is similar in all controls, mRNA abundance also only changes upon infection only in animals with functional maturation of T and B cells.
+Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF\beta and EGF are regulated in both first and challenge infections and suggest wound healing as a common function of this regulation, especially at 7 dpi.
+
+\section{RNA processing is enriched in challenge infected immune competent mice} 
+Three challenge infected samples (3 dpi, 5 dpi and 7 dpi) from immune competent mice show a distinct profile in cluster 6.
+This cluster is highly enriched for RNA processing and splicing, as well as terms for histone and chromatin modification.
+% daunorubicin metabolic process  --> cancer drug which prevents DNA replication by stabilizing DNA
+% doxorubicin metabolic process --> cancer drug
+
+%%%% INPUT
 "UDP-N-acetylglucosamine-lysosomal-enzyme" (MF)
+%%%%% here %%%%%
 
@@
 
-General changes upon infection are a down-regulation of genes involved
-"lipid modification" or more detailed "phosphatidylinositol
-phosphorylation". These changes are similarly pronounce in the
-immunocompetent strains and Rag1-/- immunodefficient animals (cluster 3
-of fibure 2c) (see supplementary table X).  The presence of these
-changes immunocompromised animals invite to the speculation that those
-genes are independent of a potential later adaptive immune response.
-
-Changes not detectable in Rag1-/- immunodefficient animals (cluster 3
-of fibure 2c) but otherwise observed in general infection are
-characterised by and enrichment to genes involved in "positive
-regulation of calcineurin-NFAT", "blood coagulation" and "inositol
-phosphate-mediated signaling" (see supplementary table X). The
-absence of these changes in Rag1-/- mice allow us to speculate that
-those genes might be involved in a early onset of adaptive
-immune responses.
-
 %% Go on here if you can...
 
-We detected further differences between infections of naive and
-challenge infected mouse hosts, especially at 7 dpi. These genes are
-characterized by...
+%We detected further differences between infections of naive and
+%challenge infected mouse hosts, especially at 7 dpi. These genes are
+%characterized by...
 %% We should't limit ourselfs to testing the clusters only! But
@@
 
-\section*{The life cycle of E. falciformis is characterized by major changes in mRNA abundance}
-
-Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison. We applied hierarchical clustering on 
-\textit{E. falciformis} mRNAs which were differently abundant at different time-points 
-post infection, including comparisons with oocysts and sporozoites. 
-Upon hierarchical clustering (Figure 3c) changes throughout the parasite's intracellular life cycle 
-can be categorised into an early (3 and 5 dpi) and late (7dpi) phase. 
-Mouse immune status has no apparent influence on clustering of
-parasite samples. Sporozoites show expression profiles similar to early infection stages
-while oocysts show a distinct mRNA abundance profile (MDS plot...).
-
-Parasite gene clusters shown on the left side (colour-coded and numbered 1 - 7) can be used 
-to characterise clusters of samples.
-Late in infection, a profile of high mRNA abundance in clusters 2 and 7 (mint green and purple) is seen.
-Also parasite genes in cluster 6 (blue) appear characteristic  by displaying low mRNA abundance in late stages of infection.
- One late challenge infection sample deviates from this pattern together with an early challenge infection sample.
-These two samples are, apart from controls and excluded samples, the ones with the least number of detected parasite reads (1836 and 1580, respectively), which taken together suggests that infections had been cleared before sampling in these challenge infected mice.
-
-The early infection sample cluster (12 samples in center of heatmap) is characterised by a high abundance of transcripts in clusters 3 (yellow) and 6, as well as a below average abundance in clusters 1 (pink), 2, and 5 (green).
-Sporozoite samples have highly abundant transcripts in cluster 4 (red) and low abundance of 
-transcripts in gene cluster 7. The latter is shared with oocyst samples. The oocyst samples, 
-in addition, have low abundance of transcripts in gene clusters 2, 3, 4, 6 (and 7) and highly 
-abundant transcripts in clusters 1 and 5.
-
-Taken together, sample and gene clustering indicates that genes in clusters 2, 6 and 7 might be 
-useful to characterise assumed gametocytes in 7 dpi samples. Genes in cluster 4 are candidates 
-for sporozoite specific genes, and clusters 1 and 5  for oocyst specific genes in E. falciformis.
-
-%% Mix it all well into the GO-term discussion! GO-term inferred
-%% biiology needs to be reported in connection to clutering-inferred
-%% ``sample biology''.
-
-%% For these samples, the pattern is most pronounced in gene cluster 1
-%% (up), and 4 and 6 (down). Within cluster 1, the second replicate from
-%% the challenge infection has a deviant profile (compare mouse data in
-%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
-%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
-%% 3 and 5 p.i. from all three mouse strains cluster together. These
-%% samples are distinct from oocysts, NMRI day 7 p.i., and
-%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
-%% oocysts or day 7.
-
+\section*{mRNA abundance .........}
+We performed statistical tests to evaluate significant differences (FDR\<1\%) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2).
+Between early time-points, 3 dpi and 5 dpi, 103 mRNAs were different, whereas between 3 dpi and 7 dpi 1399 mRNAs were DA, and between 5 dpi and 7 dpi 2084 mRNAs were DA (Figure 3a).
+This indicates that the major changes take place between 5 dpi and 7 dpi, and that variation is smaller between 3 dpi and 5 dpi.
+This motivated us to define 3 dpi and 5 dpi as "early infection" and 7 dpi as "late infection".
+Early and late infection samples were tested for DA compared to sporozoites and sporulated oocysts, resulting in 1697 and 3919 DA mRNAs, respectively.
+%% Check venn vs table2 and change in ms + code
+To evaluate this outcome further, we performed hierarchical clustering of (the union of) the DA mRNAs in the comparisons described above.
+
+\section*{Major transcriptional differences in the \textit{E. falciformis} life cycle are independent of mouse immune status}
+In the parasite transcriptome, we see no difference between infection in immune competent mice, or T and B cell deficient Rag1-/- mice, or between naive and challenge infected mice.
+This is surprising considering the measured differences in oocyst output in the same comparisons (Figure 1a), and the fact that these differences are visible in the mouse transcriptomes.
+On the parasite side, major patters instead seem to be determined by life cycle stages, independent of the host immune status.
+Distinct clusters of genes define early infection (3 dpi and 5 dpi) in which schizogony takes place, and separately, late infection (7 dpi) in which it is assumed that gametocytes are present.
+Extracellular (and in our experiments, in vitro) samples of sporozoites and oocysts also cluster separately and are defined by distinct gene clusters.
+
+\section*{Early infection transcriptomes suggest parasite expansion}
+%%%% Include cluster 3!!!!!!!
+GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) proposes that different biosynthesis processes are important, as well as "maintenance of protein location in cell".
+Possibly, the latter is due to control of microneme protein localization as sporozoites prepare for invasion.
+Biosynthesis terms probably reflect preparation for asexual expansion and perhaps ATP production for motility in sporozoites, although no GO terms for motility are enriched in this cluster.
+In the following life cycle stages, measured 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
+Among these "early infection" genes, several GO terms for biosynthetic acitivity are enriched, e.g., "ribosome biogenesis" and "cellular biosynthetic process", as well as terms for "gene expression" and RNA processes, including tRNAs and ncRNAs.
+"Cellular amino acid catabolic process" is also enriched.
+mRNAs in cluster 6 are low in abundance in both oocysts and late infection, whereas the sporozoite profile is less defined in this cluster.
+Therefore, this cluster is distinct for early infection with some features shared with invasive sporozoites.
+The enriched terms indicate replication and growth, highlighting the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in sequences from the parasite (Figure 1b).
+
+\section*{Gametocyte presence likely determines transcriptome late in infection}
+Two gene clusters have a distinct profile with high mRNA abundance on 7 dpi (clusters 2 and 7).
+Both clusters display low mRNA abundance in other life cycle stages, especially in oocysts and sporozoites.
+Enriched GO terms such as "movement of cell or subcellular component" and "microtubule-based movement" along with terms suggesting ATP production ("ATP generation from ADP") indicate the presence of motile and energy demanding gametocytes in these samples.
+Other cluster 2 terms such as "chitin metabolic process" along with enriched metabolic and biosynthetic processes and "gene expression" in cluster 7 suggests start of or preparation for encystation at 7 dpi.
+In addition, cluster 2 is enriched for a number of GO terms for "blood coagulation" and reflect
+%%%(double check) 
+the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs. Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme
+localizing proteins, MIC, e.g. MIC4 (Tomley01). MIC4 mRNA was reported in sporozoites where it localizes to the apical
+end, and in late schizonts and late oocyst stages, when sporozoites are forming.
+This suggests that \textit{E.
+falciformis} prepares for invasion already in gametocytes or during oocyst formation, however this is speculative.
+
+\section*{Oocysts...}
+Oocysts have strong high or low mRNA abundance profiles and few intermediate abundance mRNAS.
+Clusters 1 and 5 contain mRNAS with high abundance in oocysts and all other clusters have low abundance in this stage.
+GO enrichment in cluster 1 contains only one term (adj.
+p-value 0.11) for "DNA-templated transcription, initiation".
+Cluster 5 is enriched for terms related to stress responses, "DNA repair", "protein modification process" and for "cell differentiation".
+Stress responses and DNA repair can be a result of storage in potassium dichromate of mouse faeces with oocysts before purification.
+Initiation of transcription and cell differentiation could reflect slow preparation for invasion when oocysts are taken out of storage and purified for RNA extraction.
+Overall, the oocyst profile in five of 7 gene clusters is characterized by below average abundance of mRNAs, as can be expected in this life cycle stage suited for long-term survival outside of the host.
+Taken together, sample and gene clustering indicates that genes in clusters 6 are abundant only early in infection and are candidates for merozoite specific genes, clusters 2 and 7 might be useful to characterise gametocytes.
+Genes in cluster 4 can be evaluated as sporozoite specific genes, and clusters 1 and 5 as oocyst specific genes in \textit{E.
+falciformis}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% comment on weird day 7 day 3 clustering
+\section*{Imperfect clustering might reflect true biological differences}
+One late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample.
+These two samples are, apart from controls and excluded samples (see above), the ones with the least number of detected parasite genes (1836 and 1580, respectively).
+This, along with the unclear transcriptional profile and the fact that they are both challenge infections in immune competent hosts, suggest that infections had been cleared before sampling.
+If this is the case, it raises the question of why it was cleared in these samples but not in other challenge infected immune competent mice.
+The answer can lie with technical problems in the experiment.
+Hierarchical clustering analysis in most cases does not cluster replicates together, which at a first glance suggests such problems.
+However, considering the strong overall patterns in the data and that these fit well with previous knowledge about this infection, it is worthwhile to consider the possibility that this imperfection reflects true biological variation between individual hosts and possible also parasites (oocysts used for infection).
+It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in light exposure or other factors which may vary also in a controlled animal facility.
+Parasite adjustment to such factors could explain the transcriptional profiles we see.
+We suggest that considering such possibilities might be useful for understanding the basic biology of the parasite, and as an extension, successfully fight infections in less homogeneous hosts than laboratory mice.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%		detailed gene descriptions etc below: use stuff from there
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%% used thrombosp. part above
 A Thrombospondin type 1 domain-containing protein ortholog is highly
@@
 end, and in late schizonts and late oocyst stages, when sporozoites
-are forming. (Tomley01). For the same gene, the \textit{T. gondii}
+are forming. (Tomley01). 
+%% until here
+For the same gene, the \textit{T. gondii}
 annotation is Sushi domain-containing protein, which is also the
@@
 (which has a sushi domain) localizes to micronemes in merozoites but
-not other stages (OKeeffe05).  Limulus clotting factor C, Coch-5b2
+not other stages (OKeeffe05).  
+Limulus clotting factor C, Coch-5b2
 (Cochlin) and Lgl1, LCCL, (syn. F5/8 domain) domains are associated

### 2016-09-08 manuscript/additional_GO_notes.tex
put all the old GO stuff into separate document in case we want to use details form it

--- 2016-08-15 manuscript/bmc_article.tex
@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
+\documentclass[a4paper, 11pt]{article}
+\usepackage{comment} % enables the use of multi-line comments (\ifx \fi) 
+\usepackage{fullpage} % changes the margin
 
-%%IMPORTANT: do not delete the first line of this template
-%%It must be present to enable the BMC Submission system to
-%%recognise this template!!
+\begin{document}
+%Header-Make sure you update this information!!!!
+\noindent
+\large\textbf{Below are (old) notes on GO and specific genes, which can be merged in discussion above some details might also have changed due to re-analysis} \hfill \textbf{Totta Kasemo} \\
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%          <8 June 2012>              %%
-%%                                     %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.html and the instructions for   %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% additional documentclass options:
-%  [doublespacing]
-%  [linenumbers]   - put the line numbers on margins
-
-%%% loading packages, author definitions
-
-%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}
-
-%%% Load packages
-\usepackage{amsthm,amsmath}
-\RequirePackage{natbib}
-%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-\RequirePackage{hyperref}
-\usepackage[utf8]{inputenc} %unicode support
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-%\usepackage{siunitx}
-\usepackage[colorlinks]{hyperref} % To color blocks of text 
-%\usepackage[dvipsnames]{xcolor} % More color options (see Wikibooks)  
-\usepackage{changepage}
-\usepackage{booktabs} % For prettier tables
-\usepackage{float} % To control positions of e.g. tables or figures
-%\usepackage[section]{placeins}
-\usepackage{graphicx}
-\usepackage[flushleft]{threeparttable}
-
-% to color table rows
-\usepackage{color, colortbl}
-\graphicspath{ {../} }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %%
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %%
-%%  submitted article.                         %%
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\def\includegraphic{}
-%\def\includegraphics{}
-
-%%% Put your definitions there:
-\startlocaldefs
-\endlocaldefs
-
-%%% Begin ...
-\begin{document}
-
-%%% Start of article front matter
-\begin{frontmatter}
-
-\begin{fmbox}
-\dochead{Research}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Dual host-parasite transcriptomes of apicomplexan Eimeria
-  falciformis and its natural mouse host}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Specify information, if available,       %%
-%% in the form:                             %%
-%%   <key>={<id1>,<id2>}                    %%
-%%   <key>=                                 %%
-%% Comment or delete the keys which are     %%
-%% not used. Repeat \author command as much %%
-%% as required.                             %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author[ addressref={aff1},] {\inits{TK}\fnm{Totta} \snm{Kasemo}}
-\author[ addressref={aff1},] {\inits{SS}\fnm{Simone} \snm{Spork}}
-\author[ addressref={aff2},] {\inits{CD}\fnm{Christoph} \snm{Dieterich}}
-\author[ addressref={aff1}, ]{\inits{RL}\fnm{Richard} \snm{Lucius}}
-\author[ addressref={aff1, aff3},              
-  corref={aff1},                   
-  email={emanuel.heitlinger@hu-berlin.de}]{\inits{EH}\fnm{Emanuel} \snm{Heitlinger}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address[id=aff1]{
-	\orgname{Institute of Biology, Humboldt-Universitat zu Berlin}, 
-	\street{Philippstr. 13, Haus 14},
-	\postcode{10115}                 
-	\city{Berlin},                   
-	\cny{Germany}                    
-}
-\address[id=aff2]{  
-	\orgname{Max Planck Institute for Biology of Ageing},
-	\street{Joseph-Stelzmann-Strasse 9B},
-	\postcode{50931}
-	\city{Cologne},
-	\cny{Germany}
-}
-\address[id=aff3]{  
-	\orgname{Leibniz Institute of Zoo and Wildlife Research},
-	\street{Alfred-Kowalke-Str. 17},
-	\postcode{10315}
-	\city{Berlin},
-	\cny{Germany}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{artnotes}
-%\note{Sample of title note}     % note to the article
-%\note[id=n1]{Equal contributor} % note, connected to author
-%\note[id=n2]
-\end{artnotes}
-
-%\end{fmbox}% comment this for two column layout
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% Please refer to the Instructions for     %%
-%% authors on http://www.biomedcentral.com  %%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
-
-\parttitle{Second part title} %if any
-Text for this section.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{Parasite, apicomplexa, RNA-seq, transcriptome, life-cycle, interaction}
-%\kwd{article}
-%\kwd{author}
-\end{keyword}
-
-% MSC classifications codes, if any
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\end{abstractbox}
-\end{fmbox}% uncomment this for twcolumn layout
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% Please refer to the instructions for     %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-% <put your article body there>
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	INTRODUCTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Introduction}
-Text and results for this section, as per the individual journal's instructions for authors. 
-%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	RESULTS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-  \section*{Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-  \section*{A dual transcriptomics experiment}
-
-We performed mRNA sequencing of caecum tissue from mice infected with
-the apicomplexan parasite \textit{E. falciformis}. Oocysts and
-sporozoites were included as ``environmental'' stages and processed
-``\textit{in vitro}''. To follow the lifecycle of the parsite, we
-compare different timepoints after infection.
-
-We additonally used different mouse strains which displayed different
-immunocompetence in infection trials to assess the influence host
-immunocompetence on parasite development (figure 1a,b). In these
-experiments we analysed RAG-1-/- mice and compared them to the
-parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
-response du to the absence of mature B and T lymphocytes.  All three
-mouse strains were used for infections of both naive and previously
-challenge infected animals.
-
-Concerning basic phenotyping, differences in oocyst output were
-visible both between the immunocompetent (NMRI and C57BL6) and
-immunodeficient (RAG-1-/-) host strain, but also between the two
-similarly immunocompetent (NMRI and C57BL6) strains (figure
-1a). Surprisingly this finding was consistent in both in primary and
-challenge infection. Similarly development of intestinal stages as
-assayed by quantitative reverse transcription PCR (RT qPCR) also
-... ADD HERE (figure 1b).
-
-We thus aim to analyse the lifecycle of \textit{E. falciformis} under
-the influence of (most likely the innate or cell autonomouse) immune
-respons at on early stages of infection. We used an experimental
-design comparing infections at 5dpi for all possible experimental
-conditions (all three mouse strains in naive and challenge infection),
-and analyse additional infections at 3dpi (figure 1c).
-
-In our RNAseq experiment we worked with two biological replicates for
-all but two conditions (with one and three replicates,
-respecitively). For each replicate sample we enriched caecum tissue
-from three infected mice for epithelial cells, extracted mRNA and
-constructed sequencing libraries. These libraries were sequenced on
-several lanes of Illumina GAIIX and HiSeq machines and mapped to both
-mouse and parasite genome simulatneously to avoid spurious assignments
-of read in ultra conserved genomic regions.
-
-The number of total read mappings obtained for individual replicates
-ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
-(NMRI\_2ndInf\_5dpi\_rep1). As samples and idividual replicates were
-sequenced in batches to different depth and using different
-instrumentation (Table 1) we performed quality controls confirming the
-abscence of batch effects influencing analysis and quality of our
-results (additional file xyc). The results of these multivariate
-analyses in relation to biological interpretation are described
-further below.
-
-\section*{Parasite and host transcriptomes can be assessed in parallel}
-
-A maximum of 86\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of
-0.0326\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be
-assigened to the \textit{E. falciformis} genome in samples considered
-infected (table 1).
-
-We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
-due to an exptremely low parasite contribution to the overall
-transcriptome. Technically this exclusion made it possible to obtain
-read counts in agreement with a negative bionmial distributions (see
-adddtional file x). Judging from their mRNA output those samples can
-be considered non-infected, as both samples occured in a challenge
-infection we assume that the infection was cleard or reduced to a
-non-detectable output in those animmals.
-
-One sample (NMRI\_1stInf\_0dpi\_rep1) was additonally excluded because
-the uninfected control showed an unexpected mapping of reads to the
-\textit{E. falciformis} genome. We consider this together with the
-unsuccessfully infected samples above to display an undecided state of
-infection.
-
-On the other end of the spectrum mice not previously immunized
-featured an mRNA ``holo-transcriptome'' at 7dpi with the majority of
-mRNA originating from the parasite as opposed to the host (Figure
-1). The overall mRNA output of the sampled caecum tissue at this point
-of infection is dominated by the parasite with proportional mRNA
-abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and even 92\% (in
-NMRI\_1stInf\_7dpi\_rep1).
-
-The total number of sequencing reads obtained as well as those reads
-mapped to the transcriptome of either \textit{E. falciformis} and the
-mouse are indicated in Table 1 for all replicates. We additonally
-assessed the number of genes considered ``expressed'', further
-analysis and discussion considers the possibility that especially for
-\textit{E. falciformis} the number of reads might have been
-insufficient to fully normalize the dataset of the samples with the
-lowest representation against those with the hightest.
-
-\section*{The mouse transcriptome changes in the course of infection infection}
-
-Statistical testing for differetial expression between infected and
-uninfected mice, revealed changes in mRNA abundance becoming more
-pronounced (involving more genes) at later post infection (especially
-at 7dpi; Table 2).
-
-The observed changes between controls and 7dpi were in agreement with
-previously published microarry data. Fold-Change data obtained from
-mice at 6dpi on Agilent microarrays (Schmidt et al 2012) or 7dpi (our
-RNAseq data) against uninfected controls shows a strong correlateion
-(Spearman's \rho = 0.74; figure 2a). Considering both and technical
-differences between the tow methods and the biological changes along
-the lifecycle this confirms the adequacy of our methods for assessing
-the host transcriptome.
-
-325 mouse genes were considered differently expressed (DE; FDR<1\%)
-between uinfected and 3dpi, 1,804 between controls and 5dpi and 2,711
-genes between 0dpi and 7dpi. This lead to combined set of 3453 genes
-reacting to infection (figure 2bi).  DE genes early in infection (3dpi
-and and 5dpi) were not simply a subset of genes becoming DE later in
-infection (7dpi). To further validate this pattern of a distinct
-response in early infection we performed hierarchical clustering on
-(the union of) mouse genes DE between different dpi (figure 2c).
-
-The ponounced changes later in infection are reflecting the expected
-onset of an adaptive immune response. This is underlined by GO-terms
-such as ``antigen binding'' and ``immunoglobulin receptor binding''
-(molecular function) ``immune system process'' and ``adaptive immune
-response'' (biological process) highly enriched in Cluster 5 of figure
-2c.
-
-Particular ``regulation of cell death'' and ``regulation of apoptotic
-process'' (BP)
-
-``UDP-N-acetylglucosamine-lysosomal-enzyme'' (MF)
-
-%% This might be relevant: The DMAP interaction domain of
-%% UDP-GlcNAc:lysosomal enzyme
-%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
-%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
-%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
-%% N-acetyl-D-glucosaminyl-phospho-D-mannose
-
-General changes upon infection are a downregulation of genes involved
-``lipid modification'' or more detailed ``phosphatidylinositol
-phosphorylation''. These changes are similarly pronounce in the
-immunocompetent strains and RAG-/- immunodefficient animals (cluster 3
-of fibure 2c) (see supplementary table X).  The presence of these
-changes immunocompromized animals invite the speculation that those
-genes are independent of a potential later adaptive immune response.
-
-Changes not detectable in RAG-/- immunodefficient animals (cluster 3
-of fibure 2c) but otherwise observed in general infection are
-characterized by and enrichment to genes involved in ``positive
-regulation of calcineurin-NFAT'', ``blood coagulation'' and ``inositol
-phosphate-mediated signaling'' (see supplementary table X). The
-abscence of these changes in RAG-/- mice allow us to speculate that
-those genes might be involved in a very early onlset of adaptive
-immune response.
-
-%% Go on here if you can...
-
-We dedtected further differences between infections of naive and
-challenge infected mous hosts , especially at 7 dpi. These genes are
-characterized by...
-%% We should't limit ourselfs to testing the clusters only! But
-%% discuss the cluster first if you got difficulties coding the
-%% enrichment tests.
-
-\section*{The lifecycle of Eimeria is characterized by major changes in mRNA abundance}
-
-Table 2 lists pairwise sample comparisons and number of genes with
-differently abundant mRNAs per comparison.  For
-\textit{E. falciformis} no mRNAs are significantly differently
-abundant between first and second infection.
-
-Changes throughout the intr-host lifecycle can be clearly categorized
-in an early (3 and 5 dpi) and late (7dpi) phase. Sporozoites show
-expression profiles similar more similar to early infection stages
-while oocysts show distinctly different mRNA abundance profile (MDS
-plot .
-
-We applied hierarchical clustering on \textit{E. falciformis} mRNAs
-which were differently abundant at different dpi, including
-comparisons with oocysts and sporozoites.
-
-Samples from 7dpi form a separate cluster (NMRI mice only), not
-distinguishing between first and challenge infection.
-
-%% Mix it all well into the GO-term discussion! GO-term inferred
-%% biiology needs to be reported in connection to clutering-inferred
-%% ``sample biology''.
-
-%% For these samples, the pattern is most pronounced in gene cluster 1
-%% (up), and 4 and 6 (down). Within cluster 1, the second replicate from
-%% the challenge infection has a deviant profile (compare mouse data in
-%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
-%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
-%% 3 and 5 p.i. from all three mouse strains cluster together. These
-%% samples are distinct from oocysts, NMRI day 7 p.i., and
-%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
-%% oocysts or day 7.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection*{GO term enrichments in heatmap clusters}
-The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
-in \textit{T. gondii}
-
-\subsubsection*{Data overview}
-In total, xx reads were sequenced. For each sample, mouse read numbers were $10^7$ - $10^8$
-Quality filters applied resulted in xx reads being removed from further analysis
-
-\subsubsection*{Preparation for invasion in oocysts}
+\section*{Preparation for invasion in oocysts}
 The mRNA profile in the oocyst stage is mainly determined by highly
@@
 A Thrombospondin type 1 domain-containing protein ortholog is highly
-abundant in cluster 4 (high abundance in oocysts). Thrombospondin type
-1 domains have been reported in \textit{E. tenella} microneme
-localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella}
-MIC4 mRNA was reported in sporozoites where it localizes to the apical
-end, and in late schizonts and late oocyst stages, when sporozoites
-are forming. (Tomley01). For the same gene, the \textit{T. gondii}
+abundant in cluster 4 (high abundance in oocysts). 
+%%% This sentence used above:
+%%%Thrombospondin type
+%%%1 domains have been reported in \textit{E. tenella} microneme
+%%%localizing proteins, MIC, e.g. MIC4 (Tomley01) 
+%In \textit{E. tenella}
+%MIC4 mRNA was reported in sporozoites where it localizes to the apical
+%end, and in late schizonts and late oocyst stages, when sporozoites
+%are forming. (Tomley01).
+
+For the same gene, the \textit{T. gondii}
 annotation is Sushi domain-containing protein, which is also the
@@
 (Cochlin) and Lgl1, LCCL, (syn. F5/8 domain) domains are associated
-discoidin lectin domains and thereby with adhesion.  (Pfam entries for
+discoidin lectin domains andhereby with adhesion.  (Pfam entries for
 'LCCL domain' and 'Discoidin domain', May 2016).  Taken together, this
@@
 
-\subsubsection*{Amino acid biosynthesis in oocysts}
+\section*{Amino acid biosynthesis in oocysts}
 High abundance of aminotransferase mRNAs indicate amino acid
@@
 
-\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
+\section*{Oocysts contain mRNA for fatty acid catabolism}
 MmgE/PrpD is overrepresented in oocysts. The enzyme is important for
@@
 
-\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
+\section*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
 Interestingly, the genes described above which are thought to be
@@
 
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
+\section*{Down in sporozoites and oocysts -->  cluster 3..........}
 Oocysts: profile for 3, 5, 6:
@@
 Specific genes....  Enolase 2, encoded by Eno2, is among the
-downregulated genes in oocysts and sporozoites.  In \textit{T. gondii}
+downreted genes in oocysts and sporozoites.  In \textit{T. gondii}
 the paralog Eno1 is strongly associated with the cyst (bradyzoite)
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
+\section*{Motility-related mRNAs indicate gamete development on day 7}
 Two clusters contain genes with mRNAs highly abundant on day 7 p.i;
@@
 
-\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
+\section*{Microneme proteins highly expressed on day 7 p.i.}
 Unintuitively for a protozoan organism, seven out of eight GO
@@
 has also been reported to be common in nematodes such as
-\textit{Caenorhabditis elegans}, however the function is not
+\textit{Cahabditis elegans}, however the function is not
 understood. (Thordai99) The other two GO terms in the cluster of day
@@
 
-\subsubsection*{Gene and sample patterns by hierarchical clustering}
-Samples (columns) cluster into two major clusters where day 7
-p.i. samples form one group distinct from other samples. In the second
-group, oocsts and sporozoites are distinct and sporozoites cluster
-most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples
-also cluster into two groups, of which one contains all NMRI day 5
-p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
-clusters have no obvious patters.
-
-For gene clusters (rows), the two groups with high mRNA abundance on
-day 7 p.i. (cluster 1 and 2) do not cluster most closely with each
-other, but with the cluster for high mRNA abundance in oocysts
-(cluster 1 association) and with the cluster for high mRNA abundance
-in sporozoites (cluster 2 association).
-
-\section*{Evolutionary conservation of expression changes in \textit{E. falciformis}}
-
-For comparison of changes in the lifecycle of \textit{E. falciformis}
-with the economically important chicken parasite \textit{E. tenella}
-and the model apicompomplexan \textit{Toxoplasma gondii} we downloaded
-expresson datasets from ...
-
-Oocysts, Sproozoites
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
-In our analysis we demonstrate which biological processes are dominant
-in different life cycle stages of \textit{E. falciformis} in the
-mouse. The RNAseq transcriptome provided here allows for detailed
-analysis of genes involved in those processes, providing candidates
-for life stage specific marker in \textit{Eimeria} spp. research.
-
-\ldots
-\ldots
-\ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	METHODS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-\subsection{Mice and infection procedure}
-Three strains of mice were used in our experiments: NMRI (Charles
-River Laboratories, Sulzfeld, Germany), C57BL/6 (), and Rag1-/- on
-C57BL/6 background (gift from Susanne Hartmann, FU?).  Animal
-procedures were performed according to the German Animal Protection
-Laws as directed and approved by the overseeing authority Landesamt
-fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
-as described by Schmid et al., (Schmid12), but tapwater was used
-instead of PBS for administration of oocysts. Briefly, NMRI mice were
-infected two times, which will be referred to as first and second
-infection. For the first infection, 150 sporulated oocysts were
-administered in 100 $μ$L by oral gavage. During the first infection of
-60 mice, all animals were weighed every day. On day zero, before
-infection, as well as on day three, five and seven days post
-infection, dpi, caeca from 3-4 sacrificed mice per time point were
-collected. Epithelial cells were isolated as described in Schmid et
-al.(schmid12). For challenge infection, mice recovered for four weeks
-before second infection.  Recovery was monitored by weighing and
-visual inspection of fur. For the second infection, 1500 sporulated
-oocysts were applied by oral gavage. Three mice were used as
-non-second infection control, referred to as day 0, second infection.
-
-\subsection{Oocyst purification for infection and sequencing}
-Sporulated oocysts were purified by flotation from feces stored in
-potassium dichromate and administered orally in 100 uL tapwater. One
-\textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
-Haberkorn 1970, was used for all infections and parasite samples. The
-strain is maintained through passage in NMRI mice in our facilities as
-described elsewhere (schmid12).
-
-\subsection{Sporozoite isolation}
-Sporozoites were isolated from sporocysts by excystment. For this,
-sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
-tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
-30 min. Sporozoites were purified by the method of Schmatz et al
-(schmatz--).
-
-\subsection{RNA extraction}
-Total RNA was isolated from infected epithelial cells, sporozoites and
-sporulated oocysts using Trizol according to the manufacturer’s
-protocol (Invitrogen). High quality \emph{what is the meaning of 'high
-  quality' here?}  RNA was used to produce an mRNA library using the
-Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
-  genome paper} Sporozoites were stored in 1 mL Trizol until
-RNA-isolation.  Total RNA was isolated using the PureLink RNA Mini Kit
-(Invitrogen) and reverse transcribed into cDNA.
-
-\subsection{Sequencing, sequence quality assessment and alignment}
-cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
-specified in SI xx (both unstranded). A fastq\_quality\_filter
-(FASTQ-toolkit, version 0.0.14, available at
-https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
-Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
-phred score of 10 was applied. We further set q = 60. These settings
-require that nine out of ten bases or more are correct in at least
-60\% of the bases for each read.
-
-\subsection{Alignment and reference genomes}
-We used the published \textit{Mus musculus} mm10 assembly (Genome
-Reference Consortium Mouse Build 38, GCA\_000001635.2) as reference
-genome including annotations for mouse data. The
-\textit{E. falciformis} genome (Heitlinger14) was downloaded from
-ToxoDB (Gajria07). For the alignment, the mouse and parasite genome
-files were merged into a dual reference genome, and files including
-mRNA sequences from both species were aligned against the dual
-reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
-specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
-genome. Single-end and pair-end sequence samples were aligned
-separately with library type 'fr-unstranded' specified for pair-end
-samples. Import into R was enabled by the R package Ballgown, which
-requires bam files to be processed by Tablemaker (Frazee15), in our
-case used with -qW -G specified. Tablemaker in turn makes use of
-Cufflinks (version 2.1.1, Trapnell10).
-
-\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
-Count data was normalized using the R-package edgeR (version 3.14.0;
-cite) with the upperquartile normalisation method. Briefly, genes with
-zero coverage in all samples (libraries) are removed and normalisation
-factors are calculated for the 75\% quantile for each library. This
-normalisation is suitable for read densities following a negative
-binomial distribution. Two samples contradicted this assumption
-(parasite data) for later modelling and both mouse and parasite data
-from these samples were excluded from further analysis:
-NMRI\_1st\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep1 (SI ...). The method
-then fits a generalized linear model (GLM with a negative binomial
-link function) for each gene (glmFit) and then performs likelihood
-ratio tests for models w or w/o focal factor (glmLRT).
-
-%% ........... include outcome of MDS analysis here?
-%% No, results never are presented in methods.
-
-\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
-A selection of differently abundant mRNAs are used for hierarchical
-clustering of \textit{E. falciformis} life cycle relevant genes. In
-each comparison (see Table 3), the union of the at most 500 genes
-differentially abundant with lowest FDR (<0.05) are selected. In the
-next step, the 500 mRNAs from each comparison (or less) are
-joined. For \textit{E. falciformis} time p.i. comparison, this
-resulted in 1618 unique genes selected as differently expressed . The
-22 genes in the NMRI vs C57BL/6 comparison are not included in the
-\textit{E. falciformis} life cycle analysis. The same comparisons for
-mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
-all samples, i.e. also samples which themselves did not have any
-significantly different mRNAs according to our selection, were
-included in hierarchical clustering. Scale bar in heatmaps show 0 as
-mean mRNA abundance for each gene (row). Up (green) and
-down-regulation (brown) denote number of standard deviations from 0,
-i.e., row mean. Hierarchical clustering was performed using with
-Euclidean distances, using complete linkage ('complete', R package
-base).
-
-%% the paragraph above has results. All those number
-
-\subsection{Differentially abundant mRNAs used for hierarchical clustering}
-All analyses were performed in R (cite R-core). Complete scripts are
-available at https://github.com/derele/Ef\_RNAseq.git tagged as version
-1.0.
-%% We will make this available as a github webpage!!!
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}[H]
-\small
-\begin{center}
-\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
-\begin{tabular}{*3l}    \toprule
-	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
-	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
-	(including oocysts and sporozoites if appl.)	\\	
-	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
-\hline
-\end{tabular}
-\end{center}
-%\end{adjustwidth}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{Competing interests}
-  The authors declare that they have no competing interests.
-
-\section*{Author's contributions}
-    Text for this section \ldots
-
-\section*{Acknowledgements}
-  Text for this section \ldots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
-% for author-year bibliography (bmc-mathphys or spbasic)
-% a) write to bib file (bmc-mathphys only)
-% @settings{label, options="nameyear"}
-% b) uncomment next line
-%\nocite{label}
-
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
-
-\section*{Figures}
-
-\begin{figure}[h!]
-%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
-  \caption{\csentence{Experimental design and outcome of a dual RNAseq experiment}
-    read sequences per mouse read sequences for each experimental
-    condition. The more advanced the infection is, the higher the
-    ratio of parasite reads is, reflecting parasite replication
-    between day zero and day seven. ......compare 1st and 2nd
-    ....... patterns between Rag and NMRI/C57BL/6?  ........ Values
-    are mean of replicates (n=2, if * n=1) on log10 scale. Each sample
-    (replicate) consists of mRNA from three different mice.}
-  \end{figure}
-
-%%%%%
-figure 1 label cotribution of mouse vs. parasite transcripts to the
-overall transcriptional output of the analysed tissue for all samples.
-%%%%
-
-%% Selecting at most at most 500 genes differentially abundant with
-%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
-%% each comparison (or less) are joined. For \textit{E. falciformis} time
-%% p.i. comparison, this resulted in 1618 unique genes selected as
-%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
-%% are not included in the \textit{E. falciformis} life cycle
-%% analysis. The same comparisons for mouse yielded 8052 genes in total
-%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
-%% which themselves did not have any significantly different mRNAs
-%% according to our selection, were included in hierarchical
-%% clustering.
-
-%% This might be figure label
-%% Scale bar in heatmaps show 0 as mean mRNA abundance for
-%% each gene (row). Up (green) and down-regulation (brown) denote number
-%% of standard deviations from 0, i.e., row mean.
-
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-%%	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
-	\caption{\textit{E. falciformis} mRNAs with significantly
-          different abundance at different times p.i. in NMRI mice
-          (see Table 2). Samples cluster according to expected
-          developmental state of the parasite, without distinguishing
-          between first and challenge infection. Hierarchical
-          clustering was set to apply seven gene clusters.}
-%\end{center}
-\end{figure}
-
-\subsection{Parasite development reflected in mouse data}
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-%% \includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
-\caption{Mouse mRNAs with significantly different abundance at
-  different times p.i. in NMRI mice (see Table 2). Samples cluster
-  according to expected developmental state of the parasite, without
-  distinguishing between first and challenge infection. Rag1-/- mouse
-  samples cluster together. Hierarchical clustering was set to apply
-  four gene clusters.}
-%\end{center}
-\end{figure}
-
-When clustering mouse mRNAs which were different in time
-p.i. comparisons in Table 2, three out of four samples from day 7
-p.i. (NMRI only) cluster together. These samples are characterized by
-gene cluster 3 (up) as well as 2 and 4 (down). The fourth day 7-sample
-(challenge infection, second replicate) clusters with day 3 and 5
-samples. Upon visual inspection of this sample in gene cluster 4, it
-diplays a profile similar to non-infected mice. The same sample is
-abnormal in the parasite profile (Figure 1). NMRI and C57BL/6
-uninfected samples (0dpi) cluster together, defined by gene clusters 4
-(up), and 1 (down). Rag1-/- day 0 samples are however less pronounced
-in gene cluster 1. Day 3 and 5 samples cluster together, with Rag1-/-
-forming a separate group and Rag1-/- non-infected most distant in this
-sample cluster.
-
-% \begin{figure}[h!]
- % \includegraphics[width=0.7\linewidth]{ef-heatmap} \caption{\csentence{Parasite
- % genes with different mRNA abundance between samples, sorted into
- % five clusters by hierarchical clustering (method = complete,
- % distance = euclidean).}  \textit{E. falciformis} samples from seven
- % days p.i. cluster together (NMRI mice only). These samples have
- % distinct mRNA abundance patterns in all gene clusters, although
- % more pronounced in clusters 1 (up) and 4 (down). Distinct groups of
- % genes also define sporozoites (cluster 5, up) and oocysts (cluster
- % 2, up). mRNA profiles on days three and five p.i. from all three
- % mouse strains cluster together. These samples are distinct from
- % oocysts, NMRI day 7 p.i., and sporozoites, however closest to the
- % latter. On scale bar, 0 is mean mRNA abundance for each gene. Up
- % and downregulation is standard deviations from mean.}  \end{figure}
-
-%\begin{figure}[h!]
-%  \includegraphics[width=0.7\linewidth]{mm-heatmap} \caption{\csentence{Mouse
-%  genes with different mRNA abundance between samples, sorted into
-%  four clusters by hierarchical clustering (method = complete,
-%  distance = euclidean).}  Three out of four samples from day 7
-%  p.i. (NMRI only) cluster together. These samples are characterized
-%  by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6
-%  uninfected samples cluster together (left), defined by clusters 3,
-%  4 (up), and 1 (down). Rag1-/- samples cluster together. These
-%  samples share a weak downregulation of most genes in cluster 2 as
-%  well as upregulation of a small group genes in the same
-%  cluster. Uninfected Rag1-/- samples are separated from infected
-%  ones with distinct profiles in clusters 1, 3, and 4. On scale bar,
-%  0 is mean mRNA abundance for each gene. Up and downregulation is
-%  standard deviations from mean.}  \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
-
-\newcommand{\bcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\definecolor{LightCyan}{rgb}{0.88,1,1}
-\definecolor{LightRed}{rgb}{1,0.88,1}
-
-% Fri Aug 12 11:05:43 2016
-\begin{table}[ht]
-\centering
-\hspace*{-2.5cm}\begin{tabular}{lllllllll}
-  \hline
-Sample* & \bcell{Sequencing\\method} & batch & \bcell{total\\reads} & \bcell{reads\\mapping\\Mouse} & \bcell{reads\\mapping\\\textit{E.
-falciformis}} & \bcell{Percentage\\\textit{E.
-falciformis}}** & \bcell{detected\\ \textit{E.
-falciformis}\\genes} \\
-  \hline
-NMRI\_2ndInf\_0dpi\_rep1 & GAII & 2 & 108,937,797 & 70,489,674 & 247 & 0.0004 & 1 \\ 
-  Rag\_1stInf\_0dpi\_rep1 & hiseq & 3 & 25,362,793 & 18,853,850 & 443 & 0.0023 & 2 \\ 
-  C57BL6\_1stInf\_0dpi\_rep1 & hiseq & 3 & 35,731,249 & 25,119,348 & 457 & 0.0018 & 2 \\ 
-  C57BL6\_1stInf\_0dpi\_rep2 & hiseq & 3 & 47,085,959 & 34,377,133 & 608 & 0.0018 & 2 \\ 
-  Rag\_1stInf\_0dpi\_rep2 & hiseq & 3 & 46,556,156 & 35,233,327 & 676 & 0.0019 & 2 \\ 
-  NMRI\_2ndInf\_0dpi\_rep2 & hiseq & 3 & 58,122,244 & 40,794,245 & 3,406 & 0.0083 & 51 \\ 
-  \rowcolor{LightCyan}
-  NMRI\_2ndInf\_3dpi\_rep1 & hiseq & 3 & 57,934,016 & 40,544,287 & 4,803 & 0.0118 & 95 \\ 
-  \rowcolor{LightCyan}
-  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
-  \rowcolor{LightRed}
-  NMRI\_1stInf\_0dpi\_rep1 & GAII & 1 & 82,364,585 & 55,176,243 & 17,954 & 0.0325 & 701 \\ 
-  NMRI\_2ndInf\_3dpi\_rep2 & hiseq & 3 & 65,548,826 & 46,171,909 & 29,548 & 0.0640 & 1,580 \\ 
-  NMRI\_2ndInf\_7dpi\_rep2 & hiseq & 3 & 67,487,466 & 51,722,265 & 40,091 & 0.0775 & 1,836 \\ 
-  Rag\_1stInf\_5dpi\_rep1 & hiseq & 3 & 38,651,359 & 29,982,453 & 63,024 & 0.2098 & 2,548 \\ 
-  Rag\_1stInf\_5dpi\_rep2 & hiseq & 3 & 34,779,832 & 25,297,803 & 99,000 & 0.3898 & 2,828 \\ 
-  C57BL6\_1stInf\_5dpi\_rep1 & hiseq & 3 & 40,904,388 & 29,319,604 & 185,969 & 0.6303 & 4,173 \\ 
-  Rag\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 50,049,848 & 37,093,621 & 192,856 & 0.5172 & 4,167 \\ 
-  C57BL6\_1stInf\_5dpi\_rep2 & hiseq & 3 & 29,511,368 & 18,062,349 & 215,696 & 1.1801 & 3,823 \\ 
-  C57BL6\_2ndInf\_5dpi\_rep1 & hiseq & 3 & 35,148,432 & 25,660,184 & 262,909 & 1.0142 & 4,563 \\ 
-  NMRI\_1stInf\_3dpi\_rep1 & GAII & 1 & 73,236,430 & 49,993,358 & 394,384 & 0.7827 & 5,220 \\ 
-  NMRI\_1stInf\_3dpi\_rep2 & GAII & 2 & 160,709,694 & 117,791,044 & 413,051 & 0.3494 & 4,862 \\ 
-  NMRI\_1stInf\_5dpi\_rep2 & GAII & 2 & 119,902,722 & 76,419,774 & 794,570 & 1.0290 & 5,333 \\ 
-  NMRI\_2ndInf\_5dpi\_rep1 & GAII & 2 & 230,773,955 & 143,186,486 & 1,846,840 & 1.2734 & 5,533 \\ 
-  NMRI\_2ndInf\_7dpi\_rep1 & hiseq & 3 & 70,366,762 & 41,467,146 & 8,634,201 & 17.2335 & 5,875 \\ 
-  NMRI\_1stInf\_5dpi\_rep1 & GAII & 2 & 76,702,168 & 47,037,087 & 8,669,701 & 15.5631 & 5,700 \\ 
-  NMRI\_sporozoites\_rep2 & GAII & 0 & 19,551,681 & 8,656 & 11,470,604 & 99.9246 & 5,513 \\ 
-  NMRI\_1stInf\_5dpi\_rep3 & GAII & 0 & 191,099,180 & 83,735,624 & 27,839,458 & 24.9513 & 5,784 \\ 
-  NMRI\_1stInf\_7dpi\_rep1 & GAII & 1 & 66,505,514 & 3,310,666 & 39,400,884 & 92.2488 & 5,932 \\ 
-  NMRI\_sporozoites\_rep1 & GAII & 1 & 67,325,397 & 4,334 & 43,774,401 & 99.9901 & 5,825 \\ 
-  NMRI\_oocysts\_rep1 & GAII & 1 & 68,859,802 & 3,805 & 49,653,065 & 99.9923 & 5,695 \\ 
-  NMRI\_oocysts\_rep2 & GAII & 0 & 151,090,783 & 18,524 & 71,019,860 & 99.9739 & 5,777 \\ 
-  NMRI\_1stInf\_7dpi\_rep2 & GAII & 1 & 139,749,046 & 21,699,324 & 73,539,445 & 77.2159 & 5,943 \\ 
-   \hline
-\end{tabular}
-\begin{tablenotes}[flushleft]\footnotesize\singlespacing
-\item{*} sample names are given as a) mouse strain b) first or challenge
-  infection c) days post infection (dpi) and d) replicate number
-  seperated by undersocre . \\
-\item{**} percentag mapping \textit{E. falciformis} is given as percentage in total mapping reads
-\end{tablenotes}
-\end{table}
-\hspace*{+2.5cm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% 	NUMBER OF GENES DIFFERENT IN DIFFERENT COMPARISONS	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% I like this table. Looks good and transprorts the right kind of
-%% information
-
-\clearpage
-\section{mRNA abundance differences between different experimental groups}
-\setlength{\tabcolsep}{8pt}
-\begin{table}[H]
-\small
-\begin{center}
-\caption{mRNA abundance differences between different experimental groups.}
-%\label{tab:table3}
-\begin{tabular}{*3l}    \toprule
-\textit{Day post infection} & \textit{Ef} genes different & Mouse genes different \\ 
-	\textit{comparisons} 	    & (FDR$\leq$1\%) &  (FDR$\leq$1\%/5\%) \\ \midrule
-	NMRI 0 vs NMRI 3		& NA   & 274 \\
-	NMRI 0 vs NMRI 5		& NA   & 1736 \\
-	NMRI 0 vs NMRI 7		& NA   & 2802 \\
-	NMRI 3 vs NMRI 5     		& 111  & 1 \\
-	NMRI 3 vs NMRI 7  		& 1385 & 1407 \\ 
-	NMRI 5 vs NMRI 7  		& 1895 & 873 \\ 
-	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
-	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
-\textit{Day post infection,} & 		 & 		 \\ 
-\textit{parasite relevant comparisons} 	    & 		&  \\ \midrule
-	Oocysts vs NMRI 3  	& 3310 & NA \\  
-	Oocysts vs NMRI 5	& 3605 & NA \\ 
-	Oocysts vs NMRI 7	& 3085 & NA \\ 
-	Oocysts vs sporozoites  & 3421 & NA \\
-	Sporozoites vs NMRI 3 	& 1663 & NA \\
-	Sporozoites vs NMRI 5 	& 1605 & NA \\
-	Sporozoites vs NMRI 7 	& 2473 & NA \\ \midrule
-\textit{First and second infection} & 		 & 	 \\ 
-\textit{comparisons} 	    & 		& 	\\ \midrule
-	NMRI 3 1st vs NMRI 3 2nd  	& 0  & 5 \\
-	NMRI 5 1st vs NMRI 5 2nd  	& 0  & 1 \\
-	NMRI 7 1st vs NMRI 7 2nd  	& 0  & 902 \\
-	C57BL/6 1st vs C57BL/6 2nd (day 5) & 0 &  mouse \\
-	Rag1-/- 1st vs Rag1-/- 2nd (day 5) & 0 & mouse  \\ \midrule
-\bottomrule
-	\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Additional Files}
-  \subsection*{Additional file 1 --- Raw and normalized counts }
-  Raw counts of reads mappins to the \textit{E. falciformis} and mouse
-  genome for individual samples in our study. Normalized counts for
-  seperately for the host and parasite mappings (three compressed csv
-  files).
-
-  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
-  Focal contrast, fold-changes, likelihood ratio in/excluding this
-  difference in models, p-values , and false discovery rates (adjusted
-  p-values) are given for all tested contrasts (one compressed csv
-  file).
-
-  \subsection*{Additional file 3 --- Additional methods and resuts }
-  % this would be cool but requires special referencing in the text
-  % (hard to keep track and hard to fit in the journals referencing
-  % system). But we can try.
-  Document containing additonal figures and summary tables (pdf).
-
-\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
-  Tables listing all tested gene sets and resulting significatn GO
-  terms.
-
-\end{backmatter}
 \end{document}

### 2016-09-08 manuscript/bmc_article.tex
commit manuscript before starting to mess with including bibliography

@@
 \usepackage{amsthm,amsmath}
-%\usepackage{asmath}
 \RequirePackage{natbib}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-We performed mRNA sequencing of caecum tissue from mice infected with
+We performed mRNA sequencing of caecum epithelial tissue from mice infected with
 the apicomplexan parasite \textit{E. falciformis}. Oocysts and
-sporozoites were included as ``environmental'' stages and processed
-``\textit{in vitro}''. To follow the life cycle of the parasite, we
+sporozoites were included as "environmental" stages and processed
+in vitro. To follow the life cycle of the parasite, we
 compare different time-points after infection.
-
-We additionally used different mouse strains which display different
-immunocompetence in infection trials measured by oocyst output to assess the influence of host
-immunocompetence on parasite development (figure 1a,b). In these
-experiments we analysed mice lacking the Recombination activating gene 1, Rag1, and compared them to the
-parental C57BL/6 mouse line. Rag1 knock-out mice lack adaptive immune
-response due to the absence of mature B and T lymphocytes.  All three
-mouse strains were used for infections of both naive mice and previously
-infected animals (onward referred to as "challenged").
-
-Basic phenotyping showed differences in oocyst numbers
-between the immunocompetent (C57BL/6) and
-immunodeficient (Rag1-/-) host strain in infection of naive mice and 
-in challenge infections. Immunocompetent NMRI mice were infected with a higher dosis  
-of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice
-compared to naive mice (Figure 1a). Oocyst numbers in faeces peaked on days 8-9 and 
-all mice had cleared the infection by day 14. Development of \textit{E. falciformis} intestinal stages is reflected by 
-the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S  (Figure 1c).
-
-We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
-the influence of host immune responses at early stages of infection. We used an experimental
+We additionally used different mouse strains which display different immunocompetence in infection trials, measured by oocyst output in order to assess the influence of host immunocompetence on parasite development (Figure 1a and 1b).
+Immune competent NMRI mice were infected and sampled at three time-points post infection, p.i.
+We also infected mice lacking the Recombination activating gene 1, Rag1, (Rag1-/-) and compared them to the parental C57BL/6 mouse line.
+Rag1-/- mice lack mature B and T lymphocytes, which is taken as a proxy for absent adaptive immunity.
+All three mouse strains were used for infections of both naive mice and previously infected (and recovered) animals (onward referred to as "challenged" or "challenge infection").
+\newline Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice.
+Immunocompetent NMRI mice were infected with a higher dosis of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice compared to naive mice.
+Oocyst numbers in faeces peaked on days 8-9 andall mice had cleared the infection by day 14.
+Development and replication of \textit{E.
+falciformis} intestinal stages is reflected by the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S (Figure 1c).
+\newline We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
+the influence of host immune responses at early and late stages of infection. We used an experimental
 design which allows to compare infections at 5 days post infection (dpi) for all experimental
-conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). In addition,
-additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).
+conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). Additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).
 
 \section*{Parasite and host dual transcriptomes can be assessed in parallel}
-In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue which was pooled 
-from three mice. mRNA was extracted and
-sequencing libraries prepared. Two biological replicates were used for
-all but two conditions (with one and three replicates, respectively). These libraries were sequenced on
-several lanes of Illumina GAIIX (13 samples) and HiSeq machines (14 samples) and mapped to both
-mouse and parasite genomes simultaneously to avoid spurious assignments
-of reads in ultra conserved genomic regions. 
-As samples and individual replicates were sequenced in batches to different depth and using different
-instrumentation (Table 1) we performed quality controls (additional files xyz). These confirm
-the absence of batch effects influencing analysis and quality of 
-results.
+In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue and pooled 
+from three mice.
+mRNA was extracted and sequencing libraries prepared.
+Two biological replicates were used for all but two conditions (with one and three replicates, respectively).
+Libraries were sequenced on several lanes of Illumina GAIIX (13 samples) and HiSeq machines (14 samples) and mapped to both mouse and parasite genomes simultaneously to avoid spurious assignments of reads in ultra conserved genomic regions.
+As samples and individual replicates were sequenced in batches to different depth and using different instrumentation (Table 1) we performed quality controls (additional files xyz).
+These confirm the absence of batch effects influencing analysis and quality of results.
 
-Total numbers of sequenced reads as well as reads
-mapped to either the \textit{E. falciformis} genome or the
-mouse genome are indicated in Table 1 for all replicates. The number of total read mappings 
-for individual replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to 145,033,326
-(NMRI\_2ndInf\_5dpi\_rep1). At the latest time-point samples, 7 dpi, the overall mRNA
-output of the sampled caecum tissue is dominated by parasite material with proportional mRNA
-abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in
-NMRI\_1stInf\_7dpi\_rep1). (Figure 1b).
+Total numbers of sequenced reads as well as reads mapped to either the \textit{E.
+falciformis} genome or the mouse genome are indicated in Table 1 for all replicates.
+The number of total read mappings for individual replicates ranged from 25,362,739 (sample Rag\_1stInf\_0dpi\_rep1) to 139,749,046 (NMRI\_1stInf\_7dpi\_rep2).
+At the latest time-point samples, 7 dpi, the overall mRNA output of the sampled caecum tissue is dominated by parasite material with proportional mRNA abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in NMRI\_1stInf\_7dpi\_rep1).
+(Figure 1b).
 
 \section*{Exclusion of samples with uncertain infection status}
-A maximum of 86\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of
-0.0326\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be
-assigned to the \textit{E. falciformis} genome in samples considered
-infected (Table 1). We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2
-due to low parasite contribution to the overall transcriptome. 
-Technically, this exclusion made it possible to obtain read counts in agreement with a 
-negative binomial distribution (see additional file x). It is also likely that the number of reads in the excluded samples would have been
-insufficient to fully normalise these datasets to those with the highest parasite contributions.
-From a biological point of view, both excluded are samples from challenge
-infection and we assume that the infection was cleared or reduced to a
-non-detectable level. One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because
-the uninfected control showed unexpected mapping of reads to the
-\textit{E. falciformis} genome. We consider the three excluded samples to display an uncertain 
-state of infection.
+A maximum of 92\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of 0.064\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be assigned to the \textit{E.
+falciformis} genome in samples considered infected (Table 1).
+We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2 due to low parasite contribution to the overall transcriptome.
+Technically, this exclusion made it possible to obtain read counts in agreement with a negative binomial distribution (see additional file x).
+It is also likely that the number of reads in the excluded samples would have been insufficient to fully normalise these datasets to those with the highest parasite contributions.
+From a biological point of view, both excluded samples are samples from challenge infection and we assume that the infection was cleared or reduced to a non-detectable level.
+One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because the uninfected control showed unexpected mapping of reads to the \textit{E.
+falciformis} genome.
+We consider the three excluded samples to display an uncertain state of infection.
 
-\section*{The mouse transcriptome changes upon \textit{E. falciformis} infection}
-Statistical testing for differential expression between infected and
-uninfected mice revealed changes in mRNA abundance becoming more
-pronounced (involving more genes) at later time-points post infection (especially
-at 7 dpi; Table 2). For mouse, 325 mRNAs were considered differently abundant (DA; FDR\<1\%)
-between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711
-mRNAs between controls and 7 dpi. This lead to a combined set of 3,453 genes
-responding to infection (Figure 2bi). DA mRNAs early in infection (3 dpi
-and 5 dpi) were not a pure subset of genes DA later in
-infection (7 dpi). Instead, the transcriptional profile of the mouse 
-changes throughout the infection. Changes between controls and 7 dpi were in agreement with
-previously published microarry data. Fold-change data obtained from
-\textit{E. falciformis} infected mice at 6 dpi on Agilent microarrays 
-(Schmidt et al 2012) or 7 dpi (our RNAseq data) against uninfected controls
-show a strong correlation (Spearman's \rho = 0.74; Figure 2a). Considering both biological 
-(e.g. exact time-points) and technical differences between the two methods this comparison 
-confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.
+\section*{The mouse transcriptome changes upon \emph{E. falciformis} infection}
+Statistical testing for differential expression between infected and uninfected mice revealed changes in mRNA abundance becoming more pronounced (containing more genes) at later time-points post infection (Table 2).
+For mouse, 325 mRNAs were considered differently abundant (DA; FDR < 0.01) between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711 mRNAs between controls and 7 dpi.
+This lead to a combined set of 3,453 unique genes responding to infection (Figure 2bi).
+DA mRNAs early in infection (3 dpi and 5 dpi) were not a strict subset of genes DA later in infection (7 dpi).
+Instead, the transcriptional profile of the mouse changes throughout the infection.
+Differences between controls and 7 dpi were in agreement with previously published microarry data.
+Fold-change data obtained from \textit{E.
+falciformis} infected mice at 6 dpi on Agilent microarrays (\cite{schmidt12}) or 7 dpi (our RNA-seq data) against uninfected controls show a strong correlation (Spearman's $\rho$ = 0.74; Figure 2a).
+Considering both biological differences in the experiments (e.g.
+exact time-points for samples), and technical differences between the two methods this comparison confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.
 
-\section*{Epithelial responses depend on mouse immune status}
-To further validate the pattern of a distinct
-response in early infection we performed hierarchical clustering on
-the (union of) mouse genes DA between different time-points post infection (Figure 2c).
-By hierarchical clustering of mouse samples, three main sample clusters formed, “main clusters” 
-(dendrogram of columns at top of Figure 2c). Immune deficient Rag1-/- mice cluster with control samples. 
+\section*{Epithelial responses vary with mouse immune status}
+A distinct response in early infection is indicated by the number of DA mRNAs between different time-points.
+To further validate the pattern we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2c).
+By hierarchical clustering of mouse samples, three main sample clusters formed (dendrogram of columns at top of Figure 2c).
+Immune deficient Rag1-/- mice cluster with control samples.
 There is no clear distinction between infected and non-infected Rag1-/- samples, which confirms the immune deficiency phenotype in these mice.
-
-We identify a group of genes (cluster 4, Figure 2c) which change the transcriptional profile upon infection (all time-points) in immune competent mice only.
-In immune compromised Rag1-/- another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in a T and B cell dependent manner.
-Hence, these genes in some way depend on functional T and B cells also in controls (cluster 3) at the earliest time-point post infection (3 dpi, cluster 4).
-Gene Ontology, GO, terms enriched and regulated in cluster 3 are, e.g., "lipid modification" and "lipid metabolic process" and "protein intraciliary transport".
-Other enriched categories in cluster 3 are terms for regulation of different metabolic processes, for blood coagulation as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
+We identify a group of genes (cluster 4, Figure 2c) which changes transcriptional profile upon infection (all time-points of infected samples) in immune competent mice only.
+In immune compromised Rag1-/-, another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in healthy animals in a T and B cell dependent manner.
+Hence, these genes in some way depend on functional T and B cells also in controls (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4).
+Gene Ontology, GO, terms enriched in cluster 3 are, e.g., "lipid modification" and "lipid metabolic process" and "protein intraciliary transport".
+Other enriched terms in cluster 3 are terms for regulation of other metabolic processes, for blood coagulation as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
 (SI file x).
@@
 The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/-.
-Infected immune competent mice have a lower abundance of these mRNAs, but in Rag1-/- there is no difference compared to controls.
+Infected immune competent mice have a lower abundance of these mRNAs.
  Therefore, these genes appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, surprisingly also at 3 dpi.
@@
 Inositol signaling can lead to release of calcium and calcineurin-dependent translocation of NFAT to the nucleus and activation of its target genes in T cells, but also many other cell types (reviewed by macian05).
-GO enrichment of cluster 4 also highlights particular regulation of transforming growth factor \beta, TGF\beta, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
-TGF\beta is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10).
+GO enrichment of cluster 4 also highlights particular regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
+TGF$\beta$ is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10).
 TNF appears to be dose-dependent and capable to suppress inflammatory responses (noti10) but is also reported to regulate proliferation of epithelial cells (kaiser97).
 Additionally, IL-1 and IL-6 are among the enriched GO terms in cluster 4.
-The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influence lymphocyte activity (dinarello09).
+The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influences lymphocyte activity (dinarello09).
 IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14) and it is known to be important for development of Th17 responses (ref in kuhn intro).
-\\\\\\\\Th17 ....
-Stange's work (discuss Richard).
+%Th17 .... Stange's work (discuss Richard). 
 IL-6 induces the Janus kinase, JAK, and signal transducer and activator of transcription STAT3, which, among other functions, is involved in wound healing in intestinal epithelial cells (pickert09).
-This analysis suggests that TGF\beta, TNF, EGF, IL-1 and IL-6 are main actors in the epithelial response to \textit{E.
+This analysis suggests that TGF$\beta$, TNF, EGF, IL-1 and IL-6 are main actors in the epithelial response to \textit{E.
 falciformis} infection and that the response is T and B cell dependent, also at early time-points.
-The pronounced changes late in infection (7 dpi) also reflect the expected onset of an adaptive immune response.
+Enrichment of these biological processes further indicates that hosts invest resources in intestinal healing also at early time-points of \textit{E.
+falciformis} infection.
+
+\newline The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response.
 This is underlined by enriched GO terms in gene clusters 1 and 5.
@@
 
-Taken together, clusters 3 and 4 seem to capture the difference between Rag1-/- mice and immune competent mice in this parasitic infection.
+Taken together, clusters 3 and 4 highlight the difference between Rag1-/- mice and immune competent mice in this parasitic infection.
 Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3).
 These do not change upon infection in mice lacking mature T and B cells.
-Other genes (cluster 4) for which the transcriptional profile is similar in all controls, mRNA abundance also only changes upon infection only in animals with functional maturation of T and B cells.
-Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF\beta and EGF are regulated in both first and challenge infections and suggest wound healing as a common function of this regulation, especially at 7 dpi.
+Other genes (cluster 4) for which the transcriptional profile is similar in all controls, mRNA abundance also changes upon infection only in animals with functional maturation of T and B cells.
+Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF$\beta$ and EGF are regulated in both first and challenge infections and suggest wound healing as a probably function of this regulation, especially at 7 dpi.
 
-\section{RNA processing is enriched in challenge infected immune competent mice} 
+\section*{RNA processing is enriched in challenge infected immune competent mice} 
 Three challenge infected samples (3 dpi, 5 dpi and 7 dpi) from immune competent mice show a distinct profile in cluster 6.
-This cluster is highly enriched for RNA processing and splicing, as well as terms for histone and chromatin modification.
-% daunorubicin metabolic process  --> cancer drug which prevents DNA replication by stabilizing DNA
-% doxorubicin metabolic process --> cancer drug
+This cluster is highly enriched for GO terms for RNA processing and splicing, as well as terms for histone and chromatin modification.
+\ldots
+\newline
+Include or not?
+- daunorubicin metabolic process > cancer drug which prevents DNA replication by stabilizing DNA
+- doxorubicin metabolic process  > cancer drug
+\newline
 
@@
 
-\section*{mRNA abundance .........}
-We performed statistical tests to evaluate significant differences (FDR\<1\%) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2).
+\section*{Transcriptional differences in the parasite life cycle are independent of mouse immune status}
+We performed statistical tests to evaluate significant differences (FDR$<$0.01) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2).
 Between early time-points, 3 dpi and 5 dpi, 103 mRNAs were different, whereas between 3 dpi and 7 dpi 1399 mRNAs were DA, and between 5 dpi and 7 dpi 2084 mRNAs were DA (Figure 3a).
@@
 %% Check venn vs table2 and change in ms + code
-To evaluate this outcome further, we performed hierarchical clustering of (the union of) the DA mRNAs in the comparisons described above.
-
-\section*{Major transcriptional differences in the \textit{E. falciformis} life cycle are independent of mouse immune status}
+To evaluate this outcome further, we performed hierarchical clustering of (the union of) the DA mRNAs in the comparisons described above and applied GO enrichment analysis of gene clusters (annotations are inferred from orthologs in other \textit{Eimeria} spp.
+or \textit{T.
+gondii}).
 In the parasite transcriptome, we see no difference between infection in immune competent mice, or T and B cell deficient Rag1-/- mice, or between naive and challenge infected mice.
@@
 On the parasite side, major patters instead seem to be determined by life cycle stages, independent of the host immune status.
-Distinct clusters of genes define early infection (3 dpi and 5 dpi) in which schizogony takes place, and separately, late infection (7 dpi) in which it is assumed that gametocytes are present.
-Extracellular (and in our experiments, in vitro) samples of sporozoites and oocysts also cluster separately and are defined by distinct gene clusters.
+Distinct clusters of genes define early infection (3 dpi and 5 dpi) in which schizogony (asexual reproduction) takes place, and separately, late infection (7 dpi) in which it is assumed that gametocytes are present.
+Extracellular samples of sporozoites and oocysts also cluster separately and are defined by distinct gene clusters.
 
-\section*{Early infection transcriptomes suggest parasite expansion}
-%%%% Include cluster 3!!!!!!!
+\section*{Early infection transcriptomes reflect parasite expansion}
 GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) proposes that different biosynthesis processes are important, as well as "maintenance of protein location in cell".
@@
 Biosynthesis terms probably reflect preparation for asexual expansion and perhaps ATP production for motility in sporozoites, although no GO terms for motility are enriched in this cluster.
-In the following life cycle stages, measured 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
-Among these "early infection" genes, several GO terms for biosynthetic acitivity are enriched, e.g., "ribosome biogenesis" and "cellular biosynthetic process", as well as terms for "gene expression" and RNA processes, including tRNAs and ncRNAs.
+In the following intracellular life cycle stages, measured at 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
+Among these "early infection" genes in cluster 6, several GO terms (biological process) for biosynthetic acitivity are enriched, e.g., "ribosome biogenesis" and "cellular biosynthetic process", as well as terms for "gene expression" and RNA processing, including terms for tRNAs and ncRNAs.
 "Cellular amino acid catabolic process" is also enriched.
-mRNAs in cluster 6 are low in abundance in both oocysts and late infection, whereas the sporozoite profile is less defined in this cluster.
-Therefore, this cluster is distinct for early infection with some features shared with invasive sporozoites.
-The enriched terms indicate replication and growth, highlighting the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in sequences from the parasite (Figure 1b).
+In cluster 3, with a similar profile as cluster 6, no GO terms are enriched.
+Cluster 6 is distinct in early infection with a biosynthesis profile shared with invasive sporozoites.
+Replication and growth-related processes being enriched highlights the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in sequences from the parasite (Figure 1b).
 
-\section*{Gametocyte presence likely determines transcriptome late in infection}
+\section*{Gametocytes likely determine transcriptome late in infection}
 Two gene clusters have a distinct profile with high mRNA abundance on 7 dpi (clusters 2 and 7).
@@
 Enriched GO terms such as "movement of cell or subcellular component" and "microtubule-based movement" along with terms suggesting ATP production ("ATP generation from ADP") indicate the presence of motile and energy demanding gametocytes in these samples.
-Other cluster 2 terms such as "chitin metabolic process" along with enriched metabolic and biosynthetic processes and "gene expression" in cluster 7 suggests start of or preparation for encystation at 7 dpi.
+Other cluster 2 terms such as "chitin metabolic process" along with enriched metabolic and biosynthetic processes and "gene expression" in cluster 7 suggests ongoing encystation at 7 dpi, fitting the timing of oocyst output which peaks at 8-9 dpi.
 In addition, cluster 2 is enriched for a number of GO terms for "blood coagulation" and reflect
 %%%(double check) 
-the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs. Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme
-localizing proteins, MIC, e.g. MIC4 (Tomley01). MIC4 mRNA was reported in sporozoites where it localizes to the apical
-end, and in late schizonts and late oocyst stages, when sporozoites are forming.
+the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs.
+Thrombospondin type 1 domains have been reported in \textit{E.
+tenella} microneme localizing proteins, MIC, e.g.
+MIC4 (Tomley01).
+MIC4 mRNA was reported in sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages, when sporozoites are forming.
 This suggests that \textit{E.
@@
 
-\section*{Oocysts...}
-Oocysts have strong high or low mRNA abundance profiles and few intermediate abundance mRNAS.
-Clusters 1 and 5 contain mRNAS with high abundance in oocysts and all other clusters have low abundance in this stage.
+\section*{Oocysts are characterized by stress responses and differentiation}
+Clusters 1 and 5 contain mRNAs with high abundance in oocysts and all other clusters have low abundance in this stage.
 GO enrichment in cluster 1 contains only one term (adj.
@@
 Stress responses and DNA repair can be a result of storage in potassium dichromate of mouse faeces with oocysts before purification.
-Initiation of transcription and cell differentiation could reflect slow preparation for invasion when oocysts are taken out of storage and purified for RNA extraction.
-Overall, the oocyst profile in five of 7 gene clusters is characterized by below average abundance of mRNAs, as can be expected in this life cycle stage suited for long-term survival outside of the host.
-Taken together, sample and gene clustering indicates that genes in clusters 6 are abundant only early in infection and are candidates for merozoite specific genes, clusters 2 and 7 might be useful to characterise gametocytes.
-Genes in cluster 4 can be evaluated as sporozoite specific genes, and clusters 1 and 5 as oocyst specific genes in \textit{E.
+Initiation of transcription and cell differentiation probably reflects (slow) preparation for invasion when oocysts are taken out of storage and purified for RNA extraction.
+Overall, the oocyst profile in five out of seven gene clusters is characterized by below average abundance of mRNAs, as can be expected in this life cycle stage which endures long-term survival outside the host.
+Taken together, sample and gene clustering indicates that genes in clusters 6 are abundant only early in infection and are candidates for merozoite specific genes, whereas clusters 2 and 7 might be useful to characterise gametocytes.
+Genes in cluster 4 can be considered sporozoite specific, and clusters 1 and 5 as oocyst specific genes in \textit{E.
 falciformis}.
@@
 \section*{Imperfect clustering might reflect true biological differences}
-One late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample.
+In parasite data, one late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample.
 These two samples are, apart from controls and excluded samples (see above), the ones with the least number of detected parasite genes (1836 and 1580, respectively).
@@
 If this is the case, it raises the question of why it was cleared in these samples but not in other challenge infected immune competent mice.
-The answer can lie with technical problems in the experiment.
+The answer can be technical problems in the experiment.
+However, other patterns suggest a different interpretation.
 Hierarchical clustering analysis in most cases does not cluster replicates together, which at a first glance suggests such problems.
-However, considering the strong overall patterns in the data and that these fit well with previous knowledge about this infection, it is worthwhile to consider the possibility that this imperfection reflects true biological variation between individual hosts and possible also parasites (oocysts used for infection).
-It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in light exposure or other factors which may vary also in a controlled animal facility.
-Parasite adjustment to such factors could explain the transcriptional profiles we see.
-We suggest that considering such possibilities might be useful for understanding the basic biology of the parasite, and as an extension, successfully fight infections in less homogeneous hosts than laboratory mice.
+On the other hand, very distinct samples such as oocysts and sporozoites do cluster as replicates.
+Also, considering the early/late infection patterns in the parasite data and that these fit well with previous knowledge about this infection (asynchronous schizogony and gametocyte formation at 7 dpi), it is worthwhile to consider whether replicate separation reflects true biological variation between hosts and possibly also parasites (oocysts used for infection).
+It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in ght exposure or other factors which may vary also in a controlled animal facility.
+Parasite adjustment to such factors could explain the transcriptional profiles we see, with overall patterns but replicate separation.
+We suggest that considering such possibilities is useful for interpreting results and understanding basic biology of the parasite, and, in extension better understand infections in less homogeneous hosts than laboratory mice.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%		detailed gene descriptions etc below: use stuff from there
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Evolutionary conservation in life cycle-characteristic gene groups}
+To gain insight about the conservation status of life stage-specific gene groups in 
+\textit{E. falciformis}, we tested gene clusters in Figure 3c for enrichment of gene orthologs which 
+are present in other species.
+Groups analyzed for orthologs are: i) one category with \textit{E.
+falciformis} only ("Efalci"), ii) one category of 10 apicomplexan parasites ("Api") (see xxx), iii) one category of Api but excluding \textit{Cryprosporidium hominis} ("Api minus C.h."), iv) one catogory with three \textit{Eimeria} species: \textit{E.
+falciformis, E.
+maxima} and \textit{E.
+tenella} ("Eimeria"), and v) one conserved group containing a broad range of species with, e.g., \textit{Saccharomyces cerevisiae} and \textit{Arabidopsis thaliana} ("conserved").
+Sporozoites are enriched for \textit{E.
+falciformis} specific genes, and the group "conserved" is underrepresented, indicating a strong species specificity in the sporozoite stage.
+Gene clusters characteristic for early infection are enriched for "Api" (cluster 6), probably reflecting that similar processes and genes are involved in asexual reproduction of the selected apicomplexan species.
+In late infection (cluster 2), characteristic genes are underrepresented for conserved and "Api minus C.h", whereas the other characteristic cluster for late infection is enriched for "conserved" orthologs.
+This clustering and difference in enrichment of conserved versus apicomplexan (minus \textit{C.
+hominis}) genes probably reflects activity of generic and phylum specific processes/genes in oocysts.
+One oocyst cluster (cluster 1) is underrepresented for "Eimeria", whereas the other oocyst genes (cluster 5) are enriched for "Api". 
+\ldots
 
-\subsection*{GO term enrichments in heatmap clusters}
-The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
-in \textit{T. gondii}
+In addition to analysing conservation by ortholog enrichment, we also performed Spearman correlation analysis between our RNA-seq transcriptomes and RNA-seq data from related parasites.
+Two datasets for the economically important chicken parasite \textit{E.
+tenella} (walker15 and reid14) and one dataset of the model apicompomplexan parasite \textit{Toxoplasma gondii} (hehl15) were included in the comparison.
+Interestingly, this analysis confirms the species specificity for the sporozoite transcriptome, by clustering \textit{E.
+tenella} sporozoite samples together, but \textit{E.
+falciformis} sporozoites with \textit{E.
+falciformis} early infection samples.
+\textit{E.
+falciformis} late infection samples correlate the most with \textit{E.
+tenella} gametocytes, indicating similarity also between the species in this stage and supporting the presence of gametocytes in our samples.
+\textit{E.
+tenella} merozoites from both independent studies are most similar to early \textit{E.
+falciformis} samples, indicating similarity also during asexual reproduction which is shown by the conservation analysis as well (Figure?/Table?).
+\textit{E.
+falciformis} oocysts cluster with unsporulated \textit{E.
+tenella} oocysts, whereas \textit{E.
+tenella} sporulated oocysts are most similar to \textit{E.
+tenella} sporozoites.
+%% SUM UP
 
-\subsubsection*{Preparation for invasion in oocysts}
-The mRNA profile in the oocyst stage is mainly determined by highly
-abundant genes in cluster 4.  Overrepresented GO-terms in this cluster
-are enriched by ortholog genes to peptidases, microneme localized
-proteins reported to be involved in invasion, genes associated with
-adhesion in protozoans (and with clotting in higher eukaryotes), and
-genes that are annotated to be involved in amino acid biosynthesis.
-Aminopeptidase N ('related' annotation) is the reported ortholog for
-three genes with abundant mRNAs in oocysts. In humans, this enzyme has
-been reported to cleave peptides bound to major histocompatibility
-complex, MHC, II (UniProt reference if we want to keep this... but
-does any secretion happen from oocysts...? Or is this too far-fetched
-to be interesting?).
-
-%% used thrombosp. part above
-A Thrombospondin type 1 domain-containing protein ortholog is highly
-abundant in cluster 4 (high abundance in oocysts). Thrombospondin type
-1 domains have been reported in \textit{E. tenella} microneme
-localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella}
-MIC4 mRNA was reported in sporozoites where it localizes to the apical
-end, and in late schizonts and late oocyst stages, when sporozoites
-are forming. (Tomley01). 
-%% until here
-For the same gene, the \textit{T. gondii}
-annotation is Sushi domain-containing protein, which is also the
-ortholog annotation of another gene in this cluster. In the related
-malaria parasite \textit{P. falciparum} the apical sushi protein, ASP,
-(which has a sushi domain) localizes to micronemes in merozoites but
-not other stages (OKeeffe05).  
-Limulus clotting factor C, Coch-5b2
-(Cochlin) and Lgl1, LCCL, (syn. F5/8 domain) domains are associated
-discoidin lectin domains and thereby with adhesion.  (Pfam entries for
-'LCCL domain' and 'Discoidin domain', May 2016).  Taken together, this
-indicates that the thrombospondin and sushi-domain genes
-(EfaB\_MINUS\_4114.g412 and EfaB\_PLUS\_1425.g183) are involved in
-sporozoite invasion in \textit{E. falciformis} and that the mRNAs are
-transcribed and available before excystation.  A role in merozoite
-re-invasion in \textit{E. falciformis} is not indicated by our data.
-The LCCL domain annotation and thrombospondins role in higher
-eukaryotes also indicates that adhesion or preparation for adhesion is
-important in oocysts.  We suggest (speculate...?) that the
-thrombospondin annotated ortholog and the LCCL domain-containing
-protein (EfaB\_MINUS\_11233.g986) are involved in cell adhesion in
-\textit{E. falciformis}.
-%One examples is found in the slime mold \textit{Dictyostelium
-%discoideum)} in which the LCCL domain is part of the mold's discoidin
-%adhesion protein. (Pfam entry for 'discoidin domain', May 2016)
-
-\subsubsection*{Amino acid biosynthesis in oocysts}
-High abundance of aminotransferase mRNAs indicate amino acid
-biosynthesis or preparation for the same in oocysts (cluster 4).  We
-identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase
-orthologs, which are enzymes contributing to L-serine and L-alanine
-production, respectively. A putative \textit{Eimeria}
-spp. cystathionine beta-synthase, CBS, in this cluster also indicates
-de novo cysteine production. Alkyl sulfatase mRNA is also abundant in
-oocysts. Generally, this enzyme enables an organism to exploit organic
-sulfur to produce and incorporate inorganic sulfur into the amino
-acids cysteine and methionine, when no inorganoc sulfur is available.
-
-'Embryonic development' Nicalin 1, patched family protein (hedgehog)
-
-\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
-MmgE/PrpD is overrepresented in oocysts. The enzyme is important for
-propionate catabolism in the 2-methylcitric acid cycle and has been
-shown to be used by the intestinal intracellular bacterium
-\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).
-Propionate is one of two most abundant small-chain fatty acids in the
-gut along with butyrate. Both fatty acids are largely produced as
-degradation products from food by commensal bacteria (Sun13). Sharing
-the intestines as a niche with \textit{S. typhimurium} it is possible
-that also \textit{E. falciformis} uses Mmg/PrpD to exploit available
-propionate for pyruvate production.
-
-\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
-Interestingly, the genes described above which are thought to be
-involved in amino acid biosynthesis and invasion are highly abundant
-in oocysts but are underrepresented in schizont stages (day 3 and day
-5 samples) and even in sporozoites. An average abundance was detected
-on day 7 for these genes, indicating a role in either gametes or early
-oocyst formation.  This pattern supports the suggestion that these
-specific mRNAs (cluster 4) for invasion and biosynthetic processes are
-prepared (and possibly expressed) in the oocyst stage but are no
-longer detectable in the cell at the timepont when the protein is
-assumed to be in use (sporozoites and merozoite stages).  Therefore,
-correlating mRNA prevalence with biological function at the timepoint
-when mRNAs are detected must be done with care.
-
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
-Oocysts: profile for 3, 5, 6:
-
-Sporozoites: 3 and 5
-
-Day 7: 5 and 7
-
-Specific genes....  Enolase 2, encoded by Eno2, is among the
-downregulated genes in oocysts and sporozoites.  In \textit{T. gondii}
-the paralog Eno1 is strongly associated with the cyst (bradyzoite)
-stage and Eno2 is associated with tachyzoite stages (kibe05). It is
-therefore expected that this mRNA is underrepresented in oocysts and
-our data also show that the same is true in sporozoites for
-\textit{E. falciformis}. (TK: If important we could look specifically
-for Eno1 in cluster 4).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%% DAY 7  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Motility-related mRNAs indicate gamete development on day 7}
-Two clusters contain genes with mRNAs highly abundant on day 7 p.i;
-cluster 1 and 2.  Dynein, kinesin and tubulin are annotations highly
-represented among orthologs of genes in both these clusters. The
-annotations indicate an important role for motility at this timepoint,
-probably reflecting development of microgametes.  In addition, in
-cluster 2, there are two 'EF-hand domain containing proteins'
-annotations as well as caltractin, centrin-1, and troponin
-annotations. Caltractin and centrin-1 are associated with the
-centrosome and structure and function of microtubuli in mammals, and
-troponin is linked to muscle function (UniProt).  Also potentially
-linked to motility is the occurence of growth arrest specific protein
-8, Gas8, which in the mouse has been reported to be highly expressed
-in the testes and important for mouse sperm function (Yeh02).
-
-Other genes among the 38 indicate carbon fixation
-(glycolysis/gluconeogenesis) or conversions of nucleoside
-phosphates. In addition, a Ras family protein, RNA polymerase II
-transcription initiation factor and Sec23 and Sec24 were among
-orthologs identified in \textit{E. falciformis} cluster 2.
-
-In cluster 1, carbon metabolism genes are represented by
-6-phosphogluconate dehydrogenase and glycogen phosphorylase family
-protein 1. UDP-glucose 4-epimerase and amiloride-sensitive amine
-oxidase are reported as upregulated in gametocytes in
-\textit{E. tenella} by RNA-seq (Walker15) and suggested by those
-authors to play a role in cyst wall synthesis.
-
-\subsubsection*{Microneme proteins highly expressed on day 7 p.i.}
-Unintuitively for a protozoan organism, seven out of eight GO
-biological process terms in cluster 1 are associated with wound
-healing and blood coagulation.  An explanation is offered by some of
-the orthologs to the three \textit{E. falciformis} genes responsible
-for these terms. In protozoa, e.g., other \textit{Eimeria} spp. and
-\textit{Toxoplasma gondii} orthologs are annotated as 'Micronemal
-protein MIC4, related' (E. tenella) and more generally for several
-other protozoa, 'PAN domain containing proteins'.  The PAN domain is
-found in the plasminogen/hepatocyte growth factor family and in
-coagulation factor XI family (REF), explaining why terms related to
-blood coagulation are enriched by these genes.  Later publications on
-\textit{T. gondii} (Marchant12) also associate PAN domains and
-proteins in apicomplexan parasites with micronemes and therefore
-invasion. In our case, this is peculiar, since the enrichment appears
-on day 7 p.i.. A possible role at this timepoint is suggested by work
-on the fungi \textit{Sclerotinia sclerotiorum} where Yu et
-al. reported an important role for PAN domain proteins in cell wall
-integrity (Yu12). This role for MIC proteins has to our knowledge not
-been investigated in apicomplexan parasites.  The PAN domain domain
-has also been reported to be common in nematodes such as
-\textit{Caenorhabditis elegans}, however the function is not
-understood. (Thordai99) The other two GO terms in the cluster of day
-seven upregulated genes are DNA replication and DNA replication
-initiation, which most likely reflects late stage schizogony or gamete
-formation.  Six genes contribute to this enrichment and orthologs are
-either annotated as DNA replcation licencing factors, DNA polymerases
-or minichromosome maintenance proteins 2/3/5/7, Mcm2/3/5/7.
-
-\subsubsection*{Gene and sample patterns by hierarchical clustering}
-Samples (columns) cluster into two major clusters where day 7
-p.i. samples form one group distinct from other samples. In the second
-group, oocsts and sporozoites are distinct and sporozoites cluster
-most closely with day 3 and 5 p.i. samples. Day 3 and 5 p.i. samples
-also cluster into two groups, of which one contains all NMRI day 5
-p.i. samples. Apart from this, the two day 3 and 5 p.i. sample
-clusters have no obvious patters.
-
-For gene clusters (rows), the two groups with high mRNA abundance on
-day 7 p.i. (cluster 1 and 2) do not cluster most closely with each
-other, but with the cluster for high mRNA abundance in oocysts
-(cluster 1 association) and with the cluster for high mRNA abundance
-in sporozoites (cluster 2 association).
-
-\section*{Evolutionary conservation of expression changes in \textit{E. falciformis}}
-
-For comparison of changes in the lifecycle of \textit{E. falciformis}
-with the economically important chicken parasite \textit{E. tenella}
-and the model apicompomplexan \textit{Toxoplasma gondii} we downloaded
-expresson datasets from ...
-
-Oocysts, Sproozoites
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
-In our analysis we demonstrate which biological processes are dominant
-in different life cycle stages of \textit{E. falciformis} in the
-mouse. The RNAseq transcriptome provided here allows for detailed
-analysis of genes involved in those processes, providing candidates
-for life stage specific marker in \textit{Eimeria} spp. research.
-
-\ldots
-\ldots
-\ldots
+In summary, we have performed a dual RNA-seq transcriptome of an apicomplexan parasite in its natural host.
+Our analysis of differentially abundant mRNAs at different time-points post infection highlights large groups of genes which characterize the different life stages of the parasite.
+The dual approach also allows insight into the host responses to this intracellular infection.
+Using GO enrichment tests and by analysing how species specific these life stage defining gene groups are, we show that the transcriptional profile of sporozoites is the most \textit{E.
+falciformis} specific stage of the time-points we have analyzed.
+The sporozoite defining gene group is characterized by ATP production, regulation of protein localization and biosynthetic processes.
+The analysis further demonstrates that early infection, 3 and 5 dpi, is not homogenous between samples, and that the conformity seen in sporozoites (and oocysts) is regained late in infection.
+We speculate that parasites during asexual replication are more asynchronous than late stage merozoites or gametocytes on 7 dpi, possibly due to an increasing influence from the host adaptive immune response.
+The asynchrone early parasite asexual replication stages are enriched for growth related processes, and a large expansion in parasite numbers is reflected in the percentage of parasite genes, read numbers as well asin RT-qPCR data.
+Gametocyte formation on 7 dpi is supported by GO enrichments of terms for motility and ATP production, as well as 7 dpi samples correlating strongly with \textit{E.
+tenella} gametocyte samples.
+Our analysis interestingly does not detect any influence of mouse immune status on the parasite transcriptome, even though a difference in oocyst output between Rag1-/- and its C57BL/6 control is clear.
+We sequenced enriched mRNA and it is possible that non-coding RNAs or other post transcriptional regulatory mechanisms in the parasite could explain differences in oocyst output.
+Another possibility is that immune competent mice mainly interfere with the parasite at a very late time-point before oocyst formation, which a different experimental design could elucidate.
 
@@
 fuer Gesundheit und Soziales (Berlin, Germany). Animals where infected
-as described by Schmid et al., (Schmid12), but tapwater was used
+as described by Schmid et al., (schmid12), but tapwater was used
 instead of PBS for administration of oocysts. Briefly, NMRI mice were
@@
 infection. For the first infection, 150 sporulated oocysts were
-administered in 100 $μ$L by oral gavage. During the first infection of
+administered in 100 $\mu$L by oral gavage. During the first infection of
 60 mice, all animals were weighed every day. On day zero, before
-infection, as well as on day three, five and seven days post
-infection, dpi, caeca from 3-4 sacrificed mice per time point were
+infection, as well as on 3 dpi, 5 dpi and 7 dpi, caeca from 3-4 sacrificed mice per time point were
 collected. Epithelial cells were isolated as described in Schmid et
@@
 Sporulated oocysts were purified by flotation from feces stored in
-potassium dichromate and administered orally in 100 uL tapwater. One
+potassium dichromate and administered orally in 100 $\mu$L tapwater. One
 \textit{E. falciformis} isolate, \textit{E. falciformis} Bayer
@@
 Sporozoites were isolated from sporocysts by excystment. For this,
-sporocysts were incubated at 37$°$C in DMEM containing 0.04\%
+sporocysts were incubated at 37\newcommand{\degree}C in DMEM containing 0.04\%
 tauroglycocholate (MP Biomedicals) and 0.25\% trypsin (Applichem) for
@@
   quality' here?}  RNA was used to produce an mRNA library using the
-Illumina’s TruSeq RNA Sample Preparation guide.  \emph{stolen from
-  genome paper} Sporozoites were stored in 1 mL Trizol until
+Illumina’s TruSeq RNA Sample Preparation guide. Sporozoites were stored in 1 mL Trizol until
 RNA-isolation. Total RNA was isolated using the PureLink RNA Mini Kit
@@
 https://github.com/agordon/fastx\_toolkit.git) was applied to Illumina
-Hiseq 2000 samples after replacing "N" bases by "." annotation.  A
+Hiseq 2000 samples after replacing "N" s by "." annotation.  A
 phred score of 10 was applied. We further set q = 60. These settings
@@
 
-%% ........... include outcome of MDS analysis here?
-%% No, results never are presented in methods.
-
-\subsection{Selection of differentially abundant mRNAs nad hierarchical clustering}
+\subsection{Selection of differentially abundant mRNAs and hierarchical clustering}
 A selection of differently abundant mRNAs are used for hierarchical
@@
 each comparison (see Table 3), the union of the at most 500 genes
-differentially abundant with lowest FDR (<0.05) are selected. In the
+differentially abundant with lowest FDR ($<$0.05) are selected. In the
 next step, the 500 mRNAs from each comparison (or less) are
-joined. For \textit{E. falciformis} time p.i. comparison, this
-resulted in 1618 unique genes selected as differently expressed . The
-22 genes in the NMRI vs C57BL/6 comparison are not included in the
-\textit{E. falciformis} life cycle analysis. The same comparisons for
-mouse yielded 8052 genes in total and 1313 unique ones. In heatmaps,
-all samples, i.e. also samples which themselves did not have any
+joined. In heatmaps, all samples, i.e., also samples which did not have any
 significantly different mRNAs according to our selection, were
-included in hierarchical clustering. Scale bar in heatmaps show 0 as
-mean mRNA abundance for each gene (row). Up (green) and
+included in hierarchical clustering. Scale bar in heatmaps shows 0 as
+mean mRNA abundance for each gene (row). Up- (green) and
 down-regulation (brown) denote number of standard deviations from 0,
 i.e., row mean. Hierarchical clustering was performed using with
-Euclidean distances, using complete linkage ('complete', R package
+Euclidean distances, by the complete linkage method ('complete', R package
 base).
 
-%% the paragraph above has results. All those number
-
-\subsection{Differentially abundant mRNAs used for hierarchical clustering}
 All analyses were performed in R (cite R-core). Complete scripts are
@@
 	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
-	Sum of 1st infection NMRI sample differences	& 4935 & 8052 \\ 
+	Sum of 1st infection NMRI sample differences	& 4 & 8052 \\ 
 	(including oocysts and sporozoites if appl.)	\\	
@@
 \section*{Author's contributions}
-    Text for this section \ldots
+Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, writing: Totta Kasemo, Emanuel Heilinger, Simone Spork (all?).
 
 \section*{Acknowledgements}
-  Text for this section \ldots
+This project was funded by..... Totta Kasemo is funded by the German Research Foundation (DFG) program GRK4026: Parasite Infections: From Experimental Models to Natural Systems. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
+\bibliographystyle{spbasic} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliographyi{ef-rnaseq-refs.bib} %{bmc_article}      % Bibliography file (usually '*.bib' )
 % for author-year bibliography (bmc-mathphys or spbasic)
@@
 \section*{Figures}
-
-\begin{figure}[h!]
-%  \includegraphics[width=0.5\linewidth]{efmm-ratio}
-  \caption{\csentence{Experimental design and outcome of a dual RNAseq experiment}
-    read sequences per mouse read sequences for each experimental
-    condition. The more advanced the infection is, the higher the
-    ratio of parasite reads is, reflecting parasite replication
-    between day zero and day seven. ......compare 1st and 2nd
-    ....... patterns between Rag and NMRI/C57BL/6?  ........ Values
-    are mean of replicates (n=2, if * n=1) on log10 scale. Each sample
-    (replicate) consists of mRNA from three different mice.}
-  \end{figure}
-
-%%%%%
-figure 1 label cotribution of mouse vs. parasite transcripts to the
-overall transcriptional output of the analysed tissue for all samples.
-%%%%
-
-%% Selecting at most at most 500 genes differentially abundant with
-%% lowest FDR (<0.05) are selected. In the next step, the 500 mRNAs from
-%% each comparison (or less) are joined. For \textit{E. falciformis} time
-%% p.i. comparison, this resulted in 1618 unique genes selected as
-%% differently expressed . The 22 genes in the NMRI vs C57BL/6 comparison
-%% are not included in the \textit{E. falciformis} life cycle
-%% analysis. The same comparisons for mouse yielded 8052 genes in total
-%% and 1313 unique ones. In heatmaps, all samples, i.e. also samples
-%% which themselves did not have any significantly different mRNAs
-%% according to our selection, were included in hierarchical
-%% clustering.
-
-%% This might be figure label
-%% Scale bar in heatmaps show 0 as mean mRNA abundance for
-%% each gene (row). Up (green) and down-regulation (brown) denote number
-%% of standard deviations from 0, i.e., row mean.
-
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-%%	\includegraphics[width=\linewidth]{EfLifecycleHeatmap.pdf}  
-	\caption{\textit{E. falciformis} mRNAs with significantly
-          different abundance at different times p.i. in NMRI mice
-          (see Table 2). Samples cluster according to expected
-          developmental state of the parasite, without distinguishing
-          between first and challenge infection. Hierarchical
-          clustering was set to apply seven gene clusters.}
-%\end{center}
-\end{figure}
-
-\subsection{Parasite development reflected in mouse data}
-\begin{figure}[h!]
-%\begin{center}
-%\centering
-%% \includegraphics[width=\linewidth]{MmLifecycleHeatmap.pdf}  
-\caption{Mouse mRNAs with significantly different abundance at
-  different times p.i. in NMRI mice (see Table 2). Samples cluster
-  according to expected developmental state of the parasite, without
-  distinguishing between first and challenge infection. Rag1-/- mouse
-  samples cluster together. Hierarchical clustering was set to apply
-  four gene clusters.}
-%\end{center}
-\end{figure}
-
-When clustering mouse mRNAs which were different in time
-p.i. comparisons in Table 2, three out of four samples from day 7
-p.i. (NMRI only) cluster together. These samples are characterized by
-gene cluster 3 (up) as well as 2 and 4 (down). The fourth day 7-sample
-(challenge infection, second replicate) clusters with day 3 and 5
-samples. Upon visual inspection of this sample in gene cluster 4, it
-diplays a profile similar to non-infected mice. The same sample is
-abnormal in the parasite profile (Figure 1). NMRI and C57BL/6
-uninfected samples (0dpi) cluster together, defined by gene clusters 4
-(up), and 1 (down). Rag1-/- day 0 samples are however less pronounced
-in gene cluster 1. Day 3 and 5 samples cluster together, with Rag1-/-
-forming a separate group and Rag1-/- non-infected most distant in this
-sample cluster.
-
-% \begin{figure}[h!]
- % \includegraphics[width=0.7\linewidth]{ef-heatmap} \caption{\csentence{Parasite
- % genes with different mRNA abundance between samples, sorted into
- % five clusters by hierarchical clustering (method = complete,
- % distance = euclidean).}  \textit{E. falciformis} samples from seven
- % days p.i. cluster together (NMRI mice only). These samples have
- % distinct mRNA abundance patterns in all gene clusters, although
- % more pronounced in clusters 1 (up) and 4 (down). Distinct groups of
- % genes also define sporozoites (cluster 5, up) and oocysts (cluster
- % 2, up). mRNA profiles on days three and five p.i. from all three
- % mouse strains cluster together. These samples are distinct from
- % oocysts, NMRI day 7 p.i., and sporozoites, however closest to the
- % latter. On scale bar, 0 is mean mRNA abundance for each gene. Up
- % and downregulation is standard deviations from mean.}  \end{figure}
-
-%\begin{figure}[h!]
-%  \includegraphics[width=0.7\linewidth]{mm-heatmap} \caption{\csentence{Mouse
-%  genes with different mRNA abundance between samples, sorted into
-%  four clusters by hierarchical clustering (method = complete,
-%  distance = euclidean).}  Three out of four samples from day 7
-%  p.i. (NMRI only) cluster together. These samples are characterized
-%  by genes in cluster 2 (up), 3, and 4 (down). NMRI and C57BL/6
-%  uninfected samples cluster together (left), defined by clusters 3,
-%  4 (up), and 1 (down). Rag1-/- samples cluster together. These
-%  samples share a weak downregulation of most genes in cluster 2 as
-%  well as upregulation of a small group genes in the same
-%  cluster. Uninfected Rag1-/- samples are separated from infected
-%  ones with distinct profiles in clusters 1, 3, and 4. On scale bar,
-%  0 is mean mRNA abundance for each gene. Up and downregulation is
-%  standard deviations from mean.}  \end{figure}
 
@@
   \rowcolor{LightCyan}
-  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & 3 & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq &  63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
   \rowcolor{LightRed}
@@
 \end{tabular}
-\begin{tablenotes}[flushleft]\footnotesize\singlespacing
-\item{*} sample names are given as a) mouse strain b) first or challenge
-  infection c) days post infection (dpi) and d) replicate number
+\begin{tablenotes}[flushleft]\footnotesize \singlespacing
+\item{*} sample names are given as a$)$ mouse strain b$)$ first or challenge
+  infection c$)$ days post infection (dpi) and d$)$ replicate number
   seperated by undersocre . \\
@@
 	C57BL/6 0 vs C57BL/6 5		& NA	& 914 \\
-	Rag1-/- 0 vs Rag1-/- 5		& NA	& 45 \\ \midrule
+	Rag1-/vs Rag1-/- 5		& NA	& 45 \\ \midrule
 \textit{Day post infection,} & 		 & 		 \\ 

### 2016-09-08 manuscript/bmc_article.tex
'references' now appear but key is not recognized

@@
 All three mouse strains were used for infections of both naive mice and previously infected (and recovered) animals (onward referred to as "challenged" or "challenge infection").
-\newline Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice.
+Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice.
 Immunocompetent NMRI mice were infected with a higher dosis of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice compared to naive mice.
@@
 falciformis} intestinal stages is reflected by the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S (Figure 1c).
-\newline We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
+We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
 the influence of host immune responses at early and late stages of infection. We used an experimental
@@
 
-\newline The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response.
+The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response.
 This is underlined by enriched GO terms in gene clusters 1 and 5.
@@
 \ldots
-\newline
+
 Include or not?
@@
 - doxorubicin metabolic process  > cancer drug
-\newline
 
@@
 \section*{Author's contributions}
-Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, writing: Totta Kasemo, Emanuel Heilinger, Simone Spork (all?).
+Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, text: Totta Kasemo, Emanuel Heilinger, Simone Spork (all?).
 
@@
 \bibliographystyle{spbasic} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliographyi{ef-rnaseq-refs.bib} %{bmc_article}      % Bibliography file (usually '*.bib' )
+\bibliography{ef-rnaseq-refs.bib} %{bmc_article}      % Bibliography file (usually '*.bib' )
 % for author-year bibliography (bmc-mathphys or spbasic)

### 2016-09-11 manuscript/bmc_article.tex
not sure whether something actually changed, but perhaps some little thing...

@@
 \begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
+%\parttitle{First part title} %if any
+Apicomplexan parasites such as Plasmodium spp., Toxoplasma gondii and Eimeria spp.
+cause disease in humans, livestock and wild animals.
+The genus Eimeria contains thousands of niche specific intracellular parasites, including several species which cause losses in poultry industries.
+Eimeria falciformis naturally infects the cecum of mice.
+Infecting one of the best studied and available animal models in biological research, E.
+falciformis constitutes a perfect model to investigate Eimeria parasites.
+However, much is still unknown about the parasite’s basic biology and no in vitro culture has been established for the full life cycle.
+We have performed a dual RNA-seq transcriptome study of the full life cycle in the mouse and of in vitro cultured sporozoites and oocysts.
+Drastic differences are seen in both parasite and host at three time-points post infection.
+Comparisons between immunocompetent and immunocompromised mice show differences in oocyst output as well as transcriptional differences indicated by Gene Ontology enrichments.
+In mouse, TGF-beta, EGF, TNF and IL-1 and IL-6 are examples of genes reacting differently depending on mouse immune status.
+Parasite transcriptomes have distinct profiles early and late in infection, characterized by biosynthesis and motility, respectively.
+Sporozoites and oocysts can also be identified by their respective transcriptional profiles.
+Taken together, the analysis highlights general patterns in the parasite’s life cycle and links them to biological processes.
+It also lays the ground for detailed analysis of specific parasitic stages and the genes relevant in them.
+The use of hosts with different immune competence highlights the role of adaptive and innate immunity and offers a source for in-depth analysis of these responses.
 
-\parttitle{Second part title} %if any
-Text for this section.
+%\parttitle{Second part title} %if any
+%Text for this section.
 \end{abstract}
@@
 \section*{Introduction}
-Text and results for this section, as per the individual journal's instructions for authors. 
+\item{Why Eimeria falciformis? Genome available}
+
+\item{Other apicomplexa, especially E. tenella, E. maxima, T. gondii and Plasmodium spp.}
+
+\item{Why is evolution of these parasites relevant/important}
+
+\item{Why mouse host (e.g. compared to chicken)}
+
+\item{Ef life cycle and timing}
+
+\item{Knowledge-gaps: basic biology not understood, no in vitro system, 
+no stable tranfection reported, unknown why so niche specific and Eimeria genus so numerous}
+
+\item{We approach some of these questions by using dual RNA-seq to produce simultaneous host and parasite 
+transcriptomes throughout in vivo infection infection}
+
+\item{Dual RNA-seq has been performed in bacteria... also in protozoans? Intracellular... Improved normalization methods 
+should make analysis of reproducing (increasing in numbers and biological material) possible... we show/evaluate this}
+
+\item{Transcriptome studies are a source of great amounts of data. Enrichment tests (in general) and GO enrichment are
+useful tools to elucidate patterns in such large data... some words about GO concept}
 %\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
@@
 
- \section*{A dual transcriptomics experiment}
+\subsection*{A dual transcriptomics experiment}
 
@@
 
-\section*{Parasite and host dual transcriptomes can be assessed in parallel}
+\subsection*{Parasite and host dual transcriptomes can be assessed in parallel}
 In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue and pooled 
@@
 
-\section*{Exclusion of samples with uncertain infection status}
+\subsection*{Exclusion of samples with uncertain infection status}
 A maximum of 92\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of 0.064\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be assigned to the \textit{E.
@@
 
-\section*{The mouse transcriptome changes upon \emph{E. falciformis} infection}
+\subsection*{The mouse transcriptome changes upon \emph{E. falciformis} infection}
 Statistical testing for differential expression between infected and uninfected mice revealed changes in mRNA abundance becoming more pronounced (containing more genes) at later time-points post infection (Table 2).
@@
 
-\section*{Epithelial responses vary with mouse immune status}
+\subsection*{Epithelial responses vary with mouse immune status}
 A distinct response in early infection is indicated by the number of DA mRNAs between different time-points.
-To further validate the pattern we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2c).
-By hierarchical clustering of mouse samples, three main sample clusters formed (dendrogram of columns at top of Figure 2c).
+To validate the pattern we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2c).
+Three main sample clusters formed (dendrogram of columns at top of Figure 2c).
 Immune deficient Rag1-/- mice cluster with control samples.
 There is no clear distinction between infected and non-infected Rag1-/- samples, which confirms the immune deficiency phenotype in these mice.
-We identify a group of genes (cluster 4, Figure 2c) which changes transcriptional profile upon infection (all time-points of infected samples) in immune competent mice only.
+We identify a group of genes (cluster 4, Figure 2c) which changes transcriptional profile upon infection  in immune competent mice only.
 In immune compromised Rag1-/-, another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in healthy animals in a T and B cell dependent manner.
-Hence, these genes in some way depend on functional T and B cells also in controls (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4).
-Gene Ontology, GO, terms enriched in cluster 3 are, e.g., "lipid modification" and "lipid metabolic process" and "protein intraciliary transport".
-Other enriched terms in cluster 3 are terms for regulation of other metabolic processes, for blood coagulation as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
+Hence, these genes surprisingly appear to depend on functional T and B cells also in controls (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4).
+Gene Ontology, GO, terms enriched in cluster 3 are, e.g., "lipid metabolic process" and "protein intraciliary transport".
+Other enriched terms in cluster 3 are regulation of other metabolic processes and "blood coagulation" as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
 (SI file x).
 We suggest that these are processes which are all intrinsically different in Rag1-/- mice compared to immune competent mice.
-The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/-.
-Infected immune competent mice have a lower abundance of these mRNAs.
- Therefore, these genes appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, surprisingly also at 3 dpi.
-Enriched GO terms in this gene cluster include 13 terms for cytokines, particularly interleukin-1, IL-1, ("negative regulation of" and "secretion of").
-Terms such as "negative regulation of viral (or inflammatory) response", "negative chemotaxis", "autophagy", "blood coagulation", "inositol phosphate-mediated signaling", and "positive regulation of calcineurin-NFAT" are also enriched in cluster 4.
+The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/- and appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, also as early as 3 dpi.
+
+\subsection*{Renewal of infected epithelium appears heavily regulated}
+Several terms enriched in cluster 4 are associated with wound healing and proliferation.
+13 terms for cytokines as well as "negative regulation of viral (or inflammatory) response", "negative chemotaxis", "autophagy", "blood coagulation", "inositol phosphate-mediated signaling", and "positive regulation of calcineurin-NFAT" are enriched.
+mRNAS supporting these terms are less abundant in infected samples.
+Although speculative, several of these processes can be linked.
 Inositol signaling can lead to release of calcium and calcineurin-dependent translocation of NFAT to the nucleus and activation of its target genes in T cells, but also many other cell types (reviewed by macian05).
-GO enrichment of cluster 4 also highlights particular regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
+Go enrichment also highlights regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
 TGF$\beta$ is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10).
-TNF appears to be dose-dependent and capable to suppress inflammatory responses (noti10) but is also reported to regulate proliferation of epithelial cells (kaiser97).
+TNF is dose-dependent and can suppress inflammatory responses (noti10) and is reported to regulate proliferation of epithelial cells (kaiser97).
 Additionally, IL-1 and IL-6 are among the enriched GO terms in cluster 4.
 The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influences lymphocyte activity (dinarello09).
-IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14) and it is known to be important for development of Th17 responses (ref in kuhn intro).
-%Th17 .... Stange's work (discuss Richard). 
-IL-6 induces the Janus kinase, JAK, and signal transducer and activator of transcription STAT3, which, among other functions, is involved in wound healing in intestinal epithelial cells (pickert09).
+IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14, probably through the Janus kinase, JAK, and signal transducer and activator of transcription STAT3 (pickert09).
+IL-6 is also known to be important for development of Th17 responses (ref in kuhn intro) which play an important role in responses to \textit{E.
+falciformis} (stange--).
 This analysis suggests that TGF$\beta$, TNF, EGF, IL-1 and IL-6 are main actors in the epithelial response to \textit{E.
 falciformis} infection and that the response is T and B cell dependent, also at early time-points.
-Enrichment of these biological processes further indicates that hosts invest resources in intestinal healing also at early time-points of \textit{E.
-falciformis} infection.
+Enrichment tests suggest that hosts invest resources in intestinal healing also at early time-points of \textit{E.
+falciformis} infection and that functional T and B cells are needed for these responses.
+An alternative interpretation is that pathology is lower in Rag1-/- mice and that these responses are therefore not triggered in them.
+Based on unpunlished data from colleagues (talk to Stange), severe pathology in infected Rag1-/- mice makes this scenario unlikely.
 
-The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response.
-This is underlined by enriched GO terms in gene clusters 1 and 5.
-Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process", "adaptive immune response"and "innate immune response" (biological process, BP) are highly enriched in cluster 5 and confirm an activated adaptive immune response at this time-point.
+\subsection*{Late infection is strongly enriched for adaptive immunity} 
+The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response, undelined by enriched GO terms in gene clusters 1 and 5.
+Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process", "adaptive immune response" and also "innate immune response" (biological process, BP) are highly enriched in cluster 5 and confirm an activated immune system and adaptive immune responses at this time-point.
 Among the same genes, natural killer cell regulation, JAK-STAT signalling, and IL-1 and interleukin-2, IL-2 production are enriched biological processes.
 IL-2 is one target of NFAT signaling and as mentioned above, JAK-STAT signaling can be induced by IL-6.
-This indicates that the early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi).
-
+It is likely that the enriched early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi) and it is encouraging that these links are detected by the methods applied here.
 Taken together, clusters 3 and 4 highlight the difference between Rag1-/- mice and immune competent mice in this parasitic infection.
-Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3).
-These do not change upon infection in mice lacking mature T and B cells.
-Other genes (cluster 4) for which the transcriptional profile is similar in all controls, mRNA abundance also changes upon infection only in animals with functional maturation of T and B cells.
+Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3), whereas in cluster 4 the profile is shared among all controls, including Rag1-/-.
+In both cases, the profile changes in immune competent mice but not in T and B cell deficient Rag1-/- mice.
 Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF$\beta$ and EGF are regulated in both first and challenge infections and suggest wound healing as a probably function of this regulation, especially at 7 dpi.
 
-\section*{RNA processing is enriched in challenge infected immune competent mice} 
+\subsection*{RNA processing is enriched in challenge infected immune competent mice} 
 Three challenge infected samples (3 dpi, 5 dpi and 7 dpi) from immune competent mice show a distinct profile in cluster 6.
@@
 
-\section*{Transcriptional differences in the parasite life cycle are independent of mouse immune status}
+\subsection*{Transcriptional differences in the parasite life cycle are independent of mouse immune status}
 We performed statistical tests to evaluate significant differences (FDR$<$0.01) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2).
@@
 
-\section*{Early infection transcriptomes reflect parasite expansion}
+\subsection*{Early infection transcriptomes reflect parasite expansion}
 GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) proposes that different biosynthesis processes are important, as well as "maintenance of protein location in cell".
 Possibly, the latter is due to control of microneme protein localization as sporozoites prepare for invasion.
-Biosynthesis terms probably reflect preparation for asexual expansion and perhaps ATP production for motility in sporozoites, although no GO terms for motility are enriched in this cluster.
+Biosynthesis terms could reflect preparation for asexual expansion.
+n reporting the genome of \textit{E.
+falciformis}, two fructose-bisphosphate aldolase, FBA, paralogs were inferred (heitlinger14).
+Based on previously reported data on \texytit{T.
+gondii}, the authors speculate that localization of FBA is controlled in sporozoites in order to direct energy supply during the supposedly costly invasion.
+% check if FBA is in cluster 4
+Our enrichment of control of protein localization support the idea that control over energy supply and probably other functions is important in the invasive stage.
 In the following intracellular life cycle stages, measured at 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
@@
 
-\section*{Gametocytes likely determine transcriptome late in infection}
+\subsection*{Gametocytes likely determine transcriptome late in infection}
 Two gene clusters have a distinct profile with high mRNA abundance on 7 dpi (clusters 2 and 7).
@@
 
-\section*{Oocysts are characterized by stress responses and differentiation}
+\subsection*{Oocysts are characterized by stress responses and differentiation}
 Clusters 1 and 5 contain mRNAs with high abundance in oocysts and all other clusters have low abundance in this stage.
@@
 %% comment on weird day 7 day 3 clustering
-\section*{Imperfect clustering might reflect true biological differences}
+\subsection*{Imperfect clustering might reflect true biological differences}
 In parasite data, one late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample.
@@
 
-\section*{Evolutionary conservation in life cycle-characteristic gene groups}
+\subsection*{Evolutionary conservation in life cycle-characteristic gene groups}
 To gain insight about the conservation status of life stage-specific gene groups in 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-\subsection{Mice and infection procedure}
+\section*{Methods}
+\subsection*{Mice and infection procedure}
 Three strains of mice were used in our experiments: NMRI (Charles
@@
 
-\subsection{Oocyst purification for infection and sequencing}
+\subsection*{Oocyst purification for infection and sequencing}
 Sporulated oocysts were purified by flotation from feces stored in
@@
 
-\subsection{Sporozoite isolation}
+\subsection*{Sporozoite isolation}
 Sporozoites were isolated from sporocysts by excystment. For this,
@@
 
-\subsection{RNA extraction}
+\subsection*{RNA extraction}
 Total RNA was isolated from infected epithelial cells, sporozoites and
@@
 
-\subsection{Sequencing, sequence quality assessment and alignment}
+\subsection*{Sequencing, sequence quality assessment and alignment}
 cDNA samples were sequenced by either GAIIX or Illumina Hiseq 2000 as
@@
 
-\subsection{Alignment and reference genomes}
+\subsection*{Alignment and reference genomes}
 We used the published \textit{Mus musculus} mm10 assembly (Genome
@@
 
-\subsection{Differential mRNA abundance, data normalisation and sample exclusions}
+\subsection*{Differential mRNA abundance, data normalisation and sample exclusions}
 Count data was normalized using the R-package edgeR (version 3.14.0;
@@
 
-\subsection{Selection of differentially abundant mRNAs and hierarchical clustering}
+\subsection*{Selection of differentially abundant mRNAs and hierarchical clustering}
 A selection of differently abundant mRNAs are used for hierarchical
@@
 %% We will make this available as a github webpage!!!
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}[H]
-\small
-\begin{center}
-\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
-\begin{tabular}{*3l}    \toprule
-	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
-	Sum of 1st infection NMRI sample differences	& 4 & 8052 \\ 
-	(including oocysts and sporozoites if appl.)	\\	
-	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
-\hline
-\end{tabular}
-\end{center}
-%\end{adjustwidth}
-\end{table}
 
@@
 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\setlength{\tabcolsep}{10pt}
+\begin{table}[H]
+\small
+\begin{center}
+\caption{Genes used for hierarchical clustering of mRNAs differently abundant depending on time p.i..}
+\begin{tabular}{*3l}    \toprule
+	\textit{Data description} & \emph{E. falciformis} genes & Mouse genes	\\ \midrule
+	Sum of 1st infection NMRI sample differences	& 4 & 8052 \\ 
+	(including oocysts and sporozoites if appl.)	\\	
+	Used in hierarchical clustering (heatmap)  	& 1618 & 1313 \\ 	\bottomrule	
+\hline
+\end{tabular}
+\end{center}
+%\end{adjustwidth}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \rowcolor{LightCyan}
-  NMRI\_2ndInf\_5dpi\_rep2 & hiseq &  63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
+  NMRI\_2ndInf\_5dpi\_rep2 & hiseq & x & 63,965,539 & 48,289,181 & 10,941 & 0.0227 & 407 \\ 
   \rowcolor{LightRed}
@@
 \end{tabular}
-\begin{tablenotes}[flushleft]\footnotesize \singlespacing
+\begin{tablenotes}[flushleft]\footnotesize\singlespacing
 \item{*} sample names are given as a$)$ mouse strain b$)$ first or challenge

### 2016-09-11 7_enrichment.R

@@ -58,7 +58,7 @@ to.test <- list(
     list(set=set.from.cluster(hcluster[["Ef"]], 2),
          type = "cluster2ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 4),
-          type = "cluster3ef", species = "Ef"),
+          type = "cluster4ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 6),
          type = "cluster6ef", species = "Ef"),
     list(set=set.from.cluster(hcluster[["Ef"]], 3),
@@ -141,8 +141,8 @@ myt <- ttheme_default(
 
 ## Change filename and part of BPMF.ll object to export tables 
 ## (adjust size of PDF for better readability)
-pdf("~/Ef_RNAseq/Supplement/SI_GOEfcluster2_mf.pdf", width = 24, height = 48)
-oocysts1 <- grid.table(data.frame(BPMF.ll$cluster2ef$MF),
+pdf("~/Ef_RNAseq/Supplement/SI_GOMmcluster7_mf.pdf", width = 24, height = 48)
+oocysts1 <- grid.table(data.frame(BPMF.ll$cluster7mm$MF),
                                theme = myt,
                                rows = NULL,
                                cols = c("GO id", "Term", 

### 2016-09-11 manuscript/bmc_article.tex
manuscript update with introduction

@@
 %\parttitle{First part title} %if any
-Apicomplexan parasites such as Plasmodium spp., Toxoplasma gondii and Eimeria spp.
+Apicomplexan parasites such as \textit{Plasmodium} spp., \textit{Toxoplasma gondii} and \textit{Eimeria} spp.
 cause disease in humans, livestock and wild animals.
-The genus Eimeria contains thousands of niche specific intracellular parasites, including several species which cause losses in poultry industries.
-Eimeria falciformis naturally infects the cecum of mice.
-Infecting one of the best studied and available animal models in biological research, E.
-falciformis constitutes a perfect model to investigate Eimeria parasites.
+The genus \textit{Eimeria} contains thousands of niche specific intracellular parasites, including several species which cause losses in poultry industries.
+\textit{Eimeria falciformis} naturally infects the cecum of mice.
+Infecting one of the best studied and available animal models in biological research, \textit{E.
+falciformis} constitutes a perfect model to investigate \textit{Eimeria} parasites.
 However, much is still unknown about the parasite’s basic biology and no in vitro culture has been established for the full life cycle.
@@
 Drastic differences are seen in both parasite and host at three time-points post infection.
-Comparisons between immunocompetent and immunocompromised mice show differences in oocyst output as well as transcriptional differences indicated by Gene Ontology enrichments.
+Comparisons between immunocompetent and immunocompromised mice show differences in oocyst output as well as transcriptional differences indicated by Gene Ontology, GO, enrichments.
 In mouse, TGF-beta, EGF, TNF and IL-1 and IL-6 are examples of genes reacting differently depending on mouse immune status.
@@
 \section*{Introduction}
-\item{Why Eimeria falciformis? Genome available}
+\textit{Eimeria falciformis} is an intracellular parasite in the phylum Apicomplexa.
+There are several thousands of apicomplexan species, including \textit{Toxoplasma gondii}, causative agent of toxoplasmosis, \textit{Plasmodium} spp., causing malaria and \textit{Cryptosporidium} spp, which cause cryptosporidosis.
+Coccidiosis is a disease in livestock and wildlife caused by coccidian parasites such as \textit{Eimeria} spp., including several economically problematic chicken parasites, e.g., \textit{E.
+tenella} (cite?).
+A useful model for studying \textit{Eimeria} spp.
+is \textit{E.
+falciformis}, which naturally infects wild mice, \textit{Mus musculus}, and laboratory mice.
 
-\item{Other apicomplexa, especially E. tenella, E. maxima, T. gondii and Plasmodium spp.}
+\textit{Eimeria} spp.
+infect the gut and are highly niche specific within parts and cell types of the intestines (cite?).
+There, these monoxenous parasites go through asexual (schizogony) and sexual reproduction which results in hosts releasing high numbers of resistant oocysts.
+(cite?) When a mouse ingests \textit{E.
+falciformis} oocysts, one sporulated oocyst releases eight infective sporozoites inside the host, which can infect epithelial crypt cells of the caecum and colon.
+Within the epithelium, so called merozoite stages form in several rounds of schizogony.
+(mesfin78) Parasite numbers increase drastically during schizogony.
+Oocysts are first detected in faeces on day six to seven after infection (cite?), and it peaks on day eight to nine p.i.
+Oocysts are not detectable after day 13-15 p.i.
+Oocysts form as merozoites differentiate into gametes, which fuse and form a zygote.
+Immature, unsporulated oocysts are shed into the environment where they mature into sporulated, infective oocysts..
+(cite?)
 
-\item{Why is evolution of these parasites relevant/important}
+Studies on host responses to this infection have shown that in \textit{E.
+falciformis} infection in laboratory mice, IFNy plays an important role and is upregulated.
+(schmid14, more...?).
+In IFNy deficient (or receptor deficient) mice, oocyst output is reduced but at the expense of the host, which displays larger weight losses and intestinal pathology in this model.
+In the IFNy defincient model, the phenotype was recovered by blocking IL-17A and IL-22 signalling, whereas only blocking IL-22 signalling recovered the oocyst shedding but had no effect on weight loss.
+(stange12) (should we bring up acquired immunity topic here?)
 
-\item{Why mouse host (e.g. compared to chicken)}
+These studies demonstrate the complex relationships between parasite and host, e.g., exemplifying that larger pathology in the host is not necessarily beneficial for the parasite, which is then less successful in producing oocysts.
+Several studies address host parasite interactions in apicomplexan parasites (e.g.
+walkerxx, hehlxx, reidxx, schmid14), but often with a strong focus on either host or parasite.
+In Plasmodium, reidberrimanxx have used large datasets to correlate host and parasite transciptional changes and protein domain information to infer host-parasite interactions at the molecular level.
+To our knowledge, in \textit{Eimeria} spp., no studies with simultaneous host and parasite data have been reported.
 
-\item{Ef life cycle and timing}
+By applying dual RNA-seq and producing host and parasite transcriptomes from the same samples and tissue, we provide a first dataset which allows analysis of host and parasite mRNA profiles at several time-points post infection.
+Not only can we for the first time provide transcriptomes for the full parasite life cycle.
+In addition, having host transcriptome data from the same samples, we analyse biological (Gene Ontology, GO) processes in both species and discuss possible interactions of interest.
+This is particularly useful since the mouse is such a well annotated organism.
+Using the mouse model, we also had the possibility to include a mouse strain which lack a gene for maturation of T and B receptors, Recombination-activating gene 1, Rag1.
+Both strains were evaluated for re-infection after recovery.
+This design allows a full comparison of infections between immune competent mice and mice with severely limited adaptive immunity, and analysis of how these differences affect the parasite transcriptomes and reproductive success.
+We contextualize the parasite data by looking for enrichment of \textit{E.
+falciformis} specific genes, apicomplexa generic genes or widely conserved genes at different life cycle stages.
+We also correlate our RNA-seq data with published transcriptomes from \textit{E.
+tenella} and \textit{T.
+gondii}.
 
-\item{Knowledge-gaps: basic biology not understood, no in vitro system, 
-no stable tranfection reported, unknown why so niche specific and Eimeria genus so numerous}
-
-\item{We approach some of these questions by using dual RNA-seq to produce simultaneous host and parasite 
-transcriptomes throughout in vivo infection infection}
-
-\item{Dual RNA-seq has been performed in bacteria... also in protozoans? Intracellular... Improved normalization methods 
-should make analysis of reproducing (increasing in numbers and biological material) possible... we show/evaluate this}
-
-\item{Transcriptome studies are a source of great amounts of data. Enrichment tests (in general) and GO enrichment are
-useful tools to elucidate patterns in such large data... some words about GO concept}
+Taken together, the present study provides the first full life cycle transcriptome of an Eimeria spp.
+which naturally infects the research model M.
+musculus.
+In addition to providing a description of changes in the life cycle, we present the matching host transcriptomes and the host biological processes which take place during infection.
+This data is a source for in depth analyses and hypothesis testing of apicomplexan parasite infections in one of the best annotated model animals at hand.
 %\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	RESULTS
+%	RESULTS & DISCUSSION
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section*{Results}
+\section*{Results and Discussion}
 
@@
 falciformis}, two fructose-bisphosphate aldolase, FBA, paralogs were inferred (heitlinger14).
-Based on previously reported data on \texytit{T.
+Based on previously reported data on \textit{T.
 gondii}, the authors speculate that localization of FBA is controlled in sporozoites in order to direct energy supply during the supposedly costly invasion.
@@
 \section*{Author's contributions}
-Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, text: Totta Kasemo, Emanuel Heilinger, Simone Spork (all?).
+Animal and parasite experiments: Simone Spork, experimental design: Richard Lucius, Simone Spork, Emanuel Heitlinger, RNA sequecing: Christoph Dieterich, data anylsis: Emanuel Heitlinger, Totta Kasemo, text: all?
 
 \section*{Acknowledgements}
-This project was funded by..... Totta Kasemo is funded by the German Research Foundation (DFG) program GRK4026: Parasite Infections: From Experimental Models to Natural Systems. 
+This project was funded by.....
+Totta Kasemo's PhD project is funded by the German Research Foundation (DFG) program GRK4026: \textit{Parasite Infections: From Experimental Models to Natural Systems}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \clearpage
-\section{mRNA abundance differences between different experimental groups}
+\section*{mRNA abundance differences between different experimental groups}
 \setlength{\tabcolsep}{8pt}

### 2016-10-07 manuscript/bmc_article.tex
manuscript updated

@@
 Within the epithelium, so called merozoite stages form in several rounds of schizogony.
-(mesfin78) Parasite numbers increase drastically during schizogony.
+(mesfin78) Parasite numbers increase drastically during schizogony, which is not completely synchronized and the exact number of schizogony rounds is not clear.
+Mesfin and Bellamy (1978) detected four schizont stages, but this has later been contradicted by our (?
+Richard?) unpublished data and others (???).
 Oocysts are first detected in faeces on day six to seven after infection (cite?), and it peaks on day eight to nine p.i.
@@
 Not only can we for the first time provide transcriptomes for the full parasite life cycle.
-In addition, having host transcriptome data from the same samples, we analyse biological (Gene Ontology, GO) processes in both species and discuss possible interactions of interest.
+In addition, having host transcriptome data from the same samples, we analyse biological processes in both species and discuss possible interactions of interest.
 This is particularly useful since the mouse is such a well annotated organism.
-Using the mouse model, we also had the possibility to include a mouse strain which lack a gene for maturation of T and B receptors, Recombination-activating gene 1, Rag1.
-Both strains were evaluated for re-infection after recovery.
+Using the mouse model, we also had the possibility to include a mouse strain which lack a gene for maturation of T and B cell receptors, Recombination-activating gene 1, Rag1.
+Both naive immune competent and T and B cell deficient mice were infected, and both groups were challenge infected after recovering from the first infection.
 This design allows a full comparison of infections between immune competent mice and mice with severely limited adaptive immunity, and analysis of how these differences affect the parasite transcriptomes and reproductive success.
-We contextualize the parasite data by looking for enrichment of \textit{E.
-falciformis} specific genes, apicomplexa generic genes or widely conserved genes at different life cycle stages.
-We also correlate our RNA-seq data with published transcriptomes from \textit{E.
+We contextualize the parasite data by testing for enrichment of \textit{E.
+falciformis} specific genes, apicomplexa specific genes or widely conserved genes in gene clusters which characterize different life cycle stages.
+We also correlate our RNA-seq data with published transcriptomes from the chicken parasite \textit{E.
 tenella} and \textit{T.
@@
 musculus.
-In addition to providing a description of changes in the life cycle, we present the matching host transcriptomes and the host biological processes which take place during infection.
-This data is a source for in depth analyses and hypothesis testing of apicomplexan parasite infections in one of the best annotated model animals at hand.
+In addition to providing a description of changes in the life cycle, we present matching host transcriptomes and analyze the host biological processes which take place during infection.
+This data is a source for in-depth analyses and hypothesis testing of apicomplexan parasite infections in one of the best annotated model animals at hand.
 %\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
@@
 compare different time-points after infection.
-We additionally used different mouse strains which display different immunocompetence in infection trials, measured by oocyst output in order to assess the influence of host immunocompetence on parasite development (Figure 1a and 1b).
+We additionally used different mouse strains which display different immunocompetence in infection trials, measured by oocyst output, in order to assess the influence of host immunocompetence on parasite development (Figure 1a and 1b).
 Immune competent NMRI mice were infected and sampled at three time-points post infection, p.i.
-We also infected mice lacking the Recombination activating gene 1, Rag1, (Rag1-/-) and compared them to the parental C57BL/6 mouse line.
+We also infected mice lacking Rag1 (Rag1-/-) and compared them to the parental C57BL/6 mouse line.
 Rag1-/- mice lack mature B and T lymphocytes, which is taken as a proxy for absent adaptive immunity.
-All three mouse strains were used for infections of both naive mice and previously infected (and recovered) animals (onward referred to as "challenged" or "challenge infection").
-Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice.
-Immunocompetent NMRI mice were infected with a higher dosis of sporulated oocysts, and a drastic reduction of oocysts in faeces was seen in challenged mice compared to naive mice.
-Oocyst numbers in faeces peaked on days 8-9 andall mice had cleared the infection by day 14.
-Development and replication of \textit{E.
-falciformis} intestinal stages is reflected by the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by quantitative reverse transcription PCR (RT qPCR) of parasitic 18S (Figure 1c).
-We thus use dual RNA-seq to analyse the life cycle of \textit{E. falciformis} under
-the influence of host immune responses at early and late stages of infection. We used an experimental
-design which allows to compare infections at 5 days post infection (dpi) for all experimental
-conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection). Additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).
+All three mouse strains were used for infections of both naive mice and previously infected (and recovered) animals (onward referred to as "challenged").
+Basic phenotyping (Figure 1a) showed differences in oocyst numbers between the immunocompetent (C57BL/6) and immunodeficient (Rag1-/-) host strains in naive and challenge infected mice, as well as differences in amount of parasite 18S sequences determined by quantitative reverse transcription PCR (RT qPCR).
+These data suggest reduced parasite numbers and reduced reproductive success in immunocompetent, challenge infected mice.
+In contrast, no difference was seen between first and challenge infection in immunodeficient Rag1-/- mice.
+Immunocompetent NMRI mice were infected with a higher dosis of sporulated oocysts (150 in first infection and 1500 in challenge infection, which was required to see a response.
+Data not shown), and a drastic reduction of oocysts in faeces was seen in challenged mice.
+Oocyst numbers in faeces peaked on days 8-9 and all mice had cleared the infection by day 14.
+Asexual replication of \textit{E.
+falciformis} intestinal stages is reflected by the percentage of parasite reads sequenced per time-point post infection (Figure 1b) and this was confirmed by RT qPCR of parasitic 18S (Figure 1c).
+We thus use dual RNA-seq to analyse the life cycle of \textit{E.
+falciformis} under the influence of different host immune capacity at early and late stages of infection.
+We used an experimental design which allows to compare infections at 5 days post infection (dpi) for all experimental conditions (NMRI, C57BL/6 and Rag1-/- mouse strains in naive and challenge infection).
+Additional time-points 3 dpi and 7 dpi were analysed from NMRI mice (Figure 1d).
 
 \subsection*{Parasite and host dual transcriptomes can be assessed in parallel}
-In this RNA-seq experiment, each replicate sample was enriched for caecum epithelial tissue and pooled 
-from three mice.
+We purified host and parasite RNA from infected tissues, and we here demonstrate that even early in infection there is sufficient parasite material for detection by RNA-seq.
+Each replicate sample was enriched for caecum epithelial tissue and pooled from three mice.
 mRNA was extracted and sequencing libraries prepared.
@@
 These confirm the absence of batch effects influencing analysis and quality of results.
-
 Total numbers of sequenced reads as well as reads mapped to either the \textit{E.
@@
 The number of total read mappings for individual replicates ranged from 25,362,739 (sample Rag\_1stInf\_0dpi\_rep1) to 139,749,046 (NMRI\_1stInf\_7dpi\_rep2).
-At the latest time-point samples, 7 dpi, the overall mRNA output of the sampled caecum tissue is dominated by parasite material with proportional mRNA abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in NMRI\_1stInf\_7dpi\_rep1).
-(Figure 1b).
+At the latest time-points, 7 dpi, the overall mRNA output of the sampled caecum tissue is dominated by parasite material with proportional parasite mRNA abundance of 77\% (in NMRI\_1stInf\_7dpi\_rep2) and 92\% (in NMRI\_1stInf\_7dpi\_rep1).
+(Figure 1b) suggesting heavily infected tissues and large parasite numbers at this time-point.
 
 \subsection*{Exclusion of samples with uncertain infection status}
-A maximum of 92\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of 0.064\% (sample NMRI\_2ndInf\_3dpi\_rep2) of mapped reads could be assigned to the \textit{E.
-falciformis} genome in samples considered infected (Table 1).
-We excluded samples NMRI\_2nd\_3dpi\_rep1 and NMRI\_2nd\_5dpi\_rep2 due to low parasite contribution to the overall transcriptome.
+Existing analysis methods of RNA-seq data (e.g.
+edgeR, DEseq) require certain distributions of data for reliable normalization and differential gene expression analysis.
+Low numbers of reads is a potential source of problematic, noisy distributions.
+We report a mapping to the parasite genome of maximum 92\% (sample NMRI\_1stInf\_7dpi\_rep1) and a minimum of 0.064\% (sample NMRI\_2ndInf\_3dpi\_rep2) in samples considered infected (Table 1).
+We excluded samples NMRI\_2nd\_3dpi\_rep1 (0.012\%) and NMRI\_2nd\_5dpi\_rep2 (0.023\%) due to low parasite contribution to the overall transcriptome.
 Technically, this exclusion made it possible to obtain read counts in agreement with a negative binomial distribution (see additional file x).
 It is also likely that the number of reads in the excluded samples would have been insufficient to fully normalise these datasets to those with the highest parasite contributions.
-From a biological point of view, both excluded samples are samples from challenge infection and we assume that the infection was cleared or reduced to a non-detectable level.
+From a biological point of view, both excluded samples are samples from challenge infection and it is likely that the infection had been cleared or reduced to a non-detectable level.
 One sample (NMRI\_1stInf\_0dpi\_rep1) was excluded because the uninfected control showed unexpected mapping of reads to the \textit{E.
@@
 \subsection*{The mouse transcriptome changes upon \emph{E. falciformis} infection}
+We here show that upon infection with \textit{E.
+falciformis}, which induces weight loss and intestinal pathology in mice, the host transcriptome undergoes drastic changes with more than 3000 genes changing their mRNA profile.
 Statistical testing for differential expression between infected and uninfected mice revealed changes in mRNA abundance becoming more pronounced (containing more genes) at later time-points post infection (Table 2).
-For mouse, 325 mRNAs were considered differently abundant (DA; FDR < 0.01) between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711 mRNAs between controls and 7 dpi.
+325 mouse mRNAs were considered differently abundant (DA; FDR < 0.01) between controls and 3 dpi, 1,804 mRNAs between controls and 5 dpi and 2,711 mRNAs between controls and 7 dpi.
 This lead to a combined set of 3,453 unique genes responding to infection (Figure 2bi).
-DA mRNAs early in infection (3 dpi and 5 dpi) were not a strict subset of genes DA later in infection (7 dpi).
-Instead, the transcriptional profile of the mouse changes throughout the infection.
-Differences between controls and 7 dpi were in agreement with previously published microarry data.
-Fold-change data obtained from \textit{E.
-falciformis} infected mice at 6 dpi on Agilent microarrays (\cite{schmidt12}) or 7 dpi (our RNA-seq data) against uninfected controls show a strong correlation (Spearman's $\rho$ = 0.74; Figure 2a).
-Considering both biological differences in the experiments (e.g.
-exact time-points for samples), and technical differences between the two methods this comparison confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.
+DA mRNAs early in infection (3 dpi and 5 dpi) were not a strict subset of genes DA later in infection (7 dpi), which would be the case if the same genes are regulated throughout infection.
+Instead, the transcriptional profile of the mouse changes more fundamentally with other genes regulated late compared to early in infection.
+Our results are in agreement with previously published microarray data from the same species.
+Differences between our controls and 7 dpi samples were correlated with fold-change data obtained from \textit{E.
+falciformis} infected mice at 6 dpi on Agilent microarrays (\cite{schmidt12}).
+The data-sets show a strong correlation (Spearman's $\rho$ = 0.74; Figure 2a).
+Considering both biological differences in the experiments such as exact time-points for sampling, and technical differences between the two methods, this comparison confirms the adequacy of using dual RNA-seq for assessing the host transcriptome.
 
 \subsection*{Epithelial responses vary with mouse immune status}
-A distinct response in early infection is indicated by the number of DA mRNAs between different time-points.
-To validate the pattern we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2c).
+A distinct response in late infection is indicated by over 1600 DA mRNAs on 7 dpi (Figure 2b).
+To analyze this result further, we performed hierarchical clustering on the (union of) mouse genes DA between different time-points post infection (Figure 2b and 2c).
 Three main sample clusters formed (dendrogram of columns at top of Figure 2c).
-Immune deficient Rag1-/- mice cluster with control samples.
-There is no clear distinction between infected and non-infected Rag1-/- samples, which confirms the immune deficiency phenotype in these mice.
-We identify a group of genes (cluster 4, Figure 2c) which changes transcriptional profile upon infection  in immune competent mice only.
-In immune compromised Rag1-/-, another group of genes (cluster 3, figure 2c) display the "infection profile" of immune competent mice also in control samples, indicating that these are mRNAs kept at a low abundance in healthy animals in a T and B cell dependent manner.
-Hence, these genes surprisingly appear to depend on functional T and B cells also in controls (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4).
+Immune deficient Rag1-/- mice, including infected samples, cluster with control samples.
+The failure to distinguish between infected and non-infected Rag1-/- samples, confirms the immune deficiency phenotype in these mice and suggest a strong influence of adaptive immune responses on sample and gene clustering.
+We identify a group of genes (cluster 4) which changes transcriptional profile upon infection in immune competent mice only, from high to low abundance.
+In immune compromised Rag1-/-, another group of genes (cluster 3) display the "infection profile" of immune competent mice also in control samples.
+These appear to be mRNAs which are kept at a low abundance in healthy animals in a T and B cell dependent manner.
+Hence, genes in cluster 3 surprisingly appear to depend on functional T and B cells also in uninfected mice (cluster 3) and at the earliest time-point post infection (3 dpi, cluster 4).
 Gene Ontology, GO, terms enriched in cluster 3 are, e.g., "lipid metabolic process" and "protein intraciliary transport".
-Other enriched terms in cluster 3 are regulation of other metabolic processes and "blood coagulation" as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
+Other enriched terms are regulation of other metabolic processes and "blood coagulation" as well as terms containing "spinal cord", "axon" or "neuronal" regulation or development.
 (SI file x).
@@
 The genes in cluster 4, which change only in immune competent mice, are highly abundant in controls including Rag1-/- and appear to be down-regulated upon infection by processes which depend on (mature) T and/or B cell activity, also as early as 3 dpi.
+Enriched GO biological processes include "signalling" and "cell communication" as well as terms containing "external stimuli" and "mating" and as described below, it appears that epithelial healing is induced .
 
-\subsection*{Renewal of infected epithelium appears heavily regulated}
+\subsection*{Epithelium renewal is different depending on immune status}
 Several terms enriched in cluster 4 are associated with wound healing and proliferation.
@@
 Inositol signaling can lead to release of calcium and calcineurin-dependent translocation of NFAT to the nucleus and activation of its target genes in T cells, but also many other cell types (reviewed by macian05).
-Go enrichment also highlights regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
+GO enrichment also highlights regulation of transforming growth factor$\beta$, TGF$\beta$, epidermal growth factor, EGF, and tumor necrosis factor, TNF.
 TGF$\beta$ is important for wound healing in intestinal epithelium (beck03), and EGF regulates proliferation of epithelial cells and inhibits apoptosis (suzuki10).
 TNF is dose-dependent and can suppress inflammatory responses (noti10) and is reported to regulate proliferation of epithelial cells (kaiser97).
-Additionally, IL-1 and IL-6 are among the enriched GO terms in cluster 4.
+Additionally, IL-1 and IL-6 regulation are among the enriched GO terms in cluster 4.
 The IL-1 receptor (type I) is similar to Toll-like receptors and IL-1 induces innate immune responses in many cell types, and influences lymphocyte activity (dinarello09).
+Both production and secretion appear to be downregulated by the genes is cluster 4.
 IL-6 has been shown to support repair and inhibit apoptosis after epithelial wounding (kuhn14, probably through the Janus kinase, JAK, and signal transducer and activator of transcription STAT3 (pickert09).
@@
 falciformis} infection and that the response is T and B cell dependent, also at early time-points.
-Enrichment tests suggest that hosts invest resources in intestinal healing also at early time-points of \textit{E.
-falciformis} infection and that functional T and B cells are needed for these responses.
+Enrichment tests suggest that hosts invest resources in intestinal healing and possible regulatory functions (IL-1 down) also at early time-points of \textit{E.
+falciformis} infection.
+The difference here between immunocompetent and immunodeficient Rag1-/- mice suggest that functional T and B cells are needed for these responses.
 An alternative interpretation is that pathology is lower in Rag1-/- mice and that these responses are therefore not triggered in them.
-Based on unpunlished data from colleagues (talk to Stange), severe pathology in infected Rag1-/- mice makes this scenario unlikely.
+However, based on unpublished data from colleagues (talk to Stange?), severe pathology seen in infected Rag1-/- mice makes this scenario unlikely.
 
 \subsection*{Late infection is strongly enriched for adaptive immunity} 
-The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response, undelined by enriched GO terms in gene clusters 1 and 5.
+The pronounced changes late in infection (7 dpi) reflect the expected onset of an adaptive immune response, undelined by enriched GO terms in gene clusters 1 (also abundant in early samples) and 5.
 Terms such as "antigen binding" and "immunoglobulin receptor binding" (molecular function, MF), and "immune system process", "adaptive immune response" and also "innate immune response" (biological process, BP) are highly enriched in cluster 5 and confirm an activated immune system and adaptive immune responses at this time-point.
 Among the same genes, natural killer cell regulation, JAK-STAT signalling, and IL-1 and interleukin-2, IL-2 production are enriched biological processes.
-IL-2 is one target of NFAT signaling and as mentioned above, JAK-STAT signaling can be induced by IL-6.
-It is likely that the enriched early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi) and it is encouraging that these links are detected by the methods applied here.
-Taken together, clusters 3 and 4 highlight the difference between Rag1-/- mice and immune competent mice in this parasitic infection.
-Some genes are different also between control animals (Rag1-/- and NMRI, cluster 3), whereas in cluster 4 the profile is shared among all controls, including Rag1-/-.
-In both cases, the profile changes in immune competent mice but not in T and B cell deficient Rag1-/- mice.
-Using GO enrichment analysis, we show that regulation of IL-1, IL-6, TNF,TGF$\beta$ and EGF are regulated in both first and challenge infections and suggest wound healing as a probably function of this regulation, especially at 7 dpi.
+IL-2 is one target of NFAT signalling and as mentioned above.
+JAK-STAT signalling can be induced by IL-6.
+It is likely that the enriched early responses in, e.g., NFAT and IL-6 regulation induce distinct mRNA abundace differences later in infection (7 dpi) and it is encouraging that these links are detected by the methods applied.
+Taken together, the clustering patterns of immunocompetent versus immunocompromised hosts suggest a strong influence of adaptive immunity on hierarchical clustering, and therefore on the major biological processes which characterize the infection.
+Using GO enrichment analysis, we additionally show that regulation of IL-1, IL-6, TNF, TGF$\beta$ and EGF are regulated in both first and challenge infections and suggest wound healing as a probable function of this regulation, especially at 7 dpi.
 
@@
 
-%% This might be relevant: The DMAP interaction domain of
-%% UDP-GlcNAc:lysosomal enzyme
-%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
-%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
-%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
-%% N-acetyl-D-glucosaminyl-phospho-D-mannose
-
-%% Go on here if you can...
-
-%We detected further differences between infections of naive and
-%challenge infected mouse hosts, especially at 7 dpi. These genes are
-%characterized by...
-%% We should't limit ourselfs to testing the clusters only! But
-%% discuss the cluster first if you got difficulties coding the
-%% enrichment tests.
-
 \subsection*{Transcriptional differences in the parasite life cycle are independent of mouse immune status}
+We present data which supports that \textit{E.
+falciformis} is "transcriptionally blind" to the host immune status, since comparisons between differently immune-competent mice return no significantly different mRNAs.
 We performed statistical tests to evaluate significant differences (FDR$<$0.01) in mRNA abundance between different parasite life cycle stages, approximated by time post infection (Table 2).
@@
 This is surprising considering the measured differences in oocyst output in the same comparisons (Figure 1a), and the fact that these differences are visible in the mouse transcriptomes.
-On the parasite side, major patters instead seem to be determined by life cycle stages, independent of the host immune status.
+Major patters in the parasite transcriptome instead seem to be determined by life cycle stages, independent of the host immune status.
 Distinct clusters of genes define early infection (3 dpi and 5 dpi) in which schizogony (asexual reproduction) takes place, and separately, late infection (7 dpi) in which it is assumed that gametocytes are present.
@@
 \subsection*{Early infection transcriptomes reflect parasite expansion}
-GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) proposes that different biosynthesis processes are important, as well as "maintenance of protein location in cell".
-Possibly, the latter is due to control of microneme protein localization as sporozoites prepare for invasion.
-Biosynthesis terms could reflect preparation for asexual expansion.
-n reporting the genome of \textit{E.
-falciformis}, two fructose-bisphosphate aldolase, FBA, paralogs were inferred (heitlinger14).
-Based on previously reported data on \textit{T.
-gondii}, the authors speculate that localization of FBA is controlled in sporozoites in order to direct energy supply during the supposedly costly invasion.
-% check if FBA is in cluster 4
-Our enrichment of control of protein localization support the idea that control over energy supply and probably other functions is important in the invasive stage.
-In the following intracellular life cycle stages, measured at 3 dpi and 5 dpi, in which several rounds of schizogony (asexual replication) take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
+Our data suggest ATP production as a dominant feature in sporozoites, followed by biosynthetic processes in the schizogony stages that follow upon invasion.
+GO enrichment of the major sporozoite defining cluster with high mRNA abundance in this stage (cluster 4) returns different biosynthesis processes.
+Metabolic pathway analysis (ToxoDB) of the same genes reveals several pathways which suggest that ATP production takes place, including fatty acid degradation, oxidative phosphorylation and  valine, leucine and isoleucine degradation (SI xx).
+The latter degradation of branched chain amino acids pathway is enriched by genes which have orthologs in \textit{T.
+gondii}, where they are used to utilize amino acids for energy supply (oppenheim14).
+In addition, the invasive stage is characterized by "maintenance of protein location in cell" and similar terms.
+Possibly, this is due to control of microneme protein localization as sporozoites prepare for invasion.
+In the following life cycle stages, measured at 3 dpi and 5 dpi, in which several rounds of schizogony take place, mRNAs in cluster 6 are abundant in all early samples (except one, discussed below).
 Among these "early infection" genes in cluster 6, several GO terms (biological process) for biosynthetic acitivity are enriched, e.g., "ribosome biogenesis" and "cellular biosynthetic process", as well as terms for "gene expression" and RNA processing, including terms for tRNAs and ncRNAs.
 "Cellular amino acid catabolic process" is also enriched.
-In cluster 3, with a similar profile as cluster 6, no GO terms are enriched.
-Cluster 6 is distinct in early infection with a biosynthesis profile shared with invasive sporozoites.
-Replication and growth-related processes being enriched highlights the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in sequences from the parasite (Figure 1b).
+In the small gene cluster 3, no GO terms are enriched.
+Cluster 6 is distinct in early infection with a biosynthesis profile similar to that of invasive sporozoites.
+Replication and growth-related processes being enriched highlights the parasite's expansion in numbers on 3 dpi and 5 dpi, as supported both by previous knowledge about the life cycle and our increase in parasite derived sequences (Figure 1b).
+The mRNAs supporting the processes described here all have a low abundance late in infection, on 7 dpi.
+This can be explained by a switch from the early asexual expansion towards sexual reproduction on the time-point one day before oocyst outputs peak.
 
@@
 Enriched GO terms such as "movement of cell or subcellular component" and "microtubule-based movement" along with terms suggesting ATP production ("ATP generation from ADP") indicate the presence of motile and energy demanding gametocytes in these samples.
-Other cluster 2 terms such as "chitin metabolic process" along with enriched metabolic and biosynthetic processes and "gene expression" in cluster 7 suggests ongoing encystation at 7 dpi, fitting the timing of oocyst output which peaks at 8-9 dpi.
-In addition, cluster 2 is enriched for a number of GO terms for "blood coagulation" and reflect
+Surprisingly, other genes seem to support ATP production in this life-cycle stage compared to sporozoites.
+Oocyst generation is probably also reflected in terms in cluster 2 and 7: different (peptide/nitrogen compound/cellular protein/macromolecule) "biosynthetic process" terms along with "chitin metabolic process" suggest that the parasite is producing building blocks for oocysts.
+This would fit the timing of oocyst output which peaks at 8-9 dpi.
 %%%(double check) 
-the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs.
+In addition, cluster 2 is enriched for a number of GO terms for "blood coagulation" and reflect the presence of Thrombospondin type I domains in the protein products of cluster 2 mRNAs.
 Thrombospondin type 1 domains have been reported in \textit{E.
@@
 MIC4 (Tomley01).
-MIC4 mRNA was reported in sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages, when sporozoites are forming.
-This suggests that \textit{E.
-falciformis} prepares for invasion already in gametocytes or during oocyst formation, however this is speculative.
+MIC4 mRNA was reported in \textit{E.
+tenella} sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages, when sporozoites are forming.
+This has so far not been demonstrated in \textit{E.
+falciformis}.
+Possibly, the mouse parasite prepares for invasion already in gametocytes stages or during oocyst formation, however this is speculative.
 
 \subsection*{Oocysts are characterized by stress responses and differentiation}
-Clusters 1 and 5 contain mRNAs with high abundance in oocysts and all other clusters have low abundance in this stage.
+Clusters 1 and 5 contain mRNAs with high abundance in oocysts and all other gene clusters have low mRNA abundance in this stage.
+Cluster 5 is enriched for terms related to stress responses, "DNA repair", "protein modification process" and for "cell differentiation".
 GO enrichment in cluster 1 contains only one term (adj.
 p-value 0.11) for "DNA-templated transcription, initiation".
-Cluster 5 is enriched for terms related to stress responses, "DNA repair", "protein modification process" and for "cell differentiation".
 Stress responses and DNA repair can be a result of storage in potassium dichromate of mouse faeces with oocysts before purification.
-Initiation of transcription and cell differentiation probably reflects (slow) preparation for invasion when oocysts are taken out of storage and purified for RNA extraction.
-Overall, the oocyst profile in five out of seven gene clusters is characterized by below average abundance of mRNAs, as can be expected in this life cycle stage which endures long-term survival outside the host.
+Initiation of transcription and cell differentiation probably reflects preparation for invasion when oocysts are taken out of storage and purified for RNA extraction.
+Overall, the oocyst profile in five out of seven gene clusters is characterized by below average abundance of mRNAs, as can be expected in this inactive life cycle stage which endures long-term survival outside the host.
 Taken together, sample and gene clustering indicates that genes in clusters 6 are abundant only early in infection and are candidates for merozoite specific genes, whereas clusters 2 and 7 might be useful to characterise gametocytes.
@@
 \subsection*{Imperfect clustering might reflect true biological differences}
-In parasite data, one late challenge infection sample deviates from the "late infection pattern" and clusters with an early challenge infection sample.
-These two samples are, apart from controls and excluded samples (see above), the ones with the least number of detected parasite genes (1836 and 1580, respectively).
-This, along with the unclear transcriptional profile and the fact that they are both challenge infections in immune competent hosts, suggest that infections had been cleared before sampling.
-If this is the case, it raises the question of why it was cleared in these samples but not in other challenge infected immune competent mice.
-The answer can be technical problems in the experiment.
-However, other patterns suggest a different interpretation.
-Hierarchical clustering analysis in most cases does not cluster replicates together, which at a first glance suggests such problems.
-On the other hand, very distinct samples such as oocysts and sporozoites do cluster as replicates.
-Also, considering the early/late infection patterns in the parasite data and that these fit well with previous knowledge about this infection (asynchronous schizogony and gametocyte formation at 7 dpi), it is worthwhile to consider whether replicate separation reflects true biological variation between hosts and possibly also parasites (oocysts used for infection).
-It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in ght exposure or other factors which may vary also in a controlled animal facility.
-Parasite adjustment to such factors could explain the transcriptional profiles we see, with overall patterns but replicate separation.
+Hierarchical clustering analysis in most cases does not cluster replicates together, which at a first glance might suggest technical problems.
+On the other hand, very distinct samples such as oocysts and sporozoites do cluster as replicates, and all 7 dpi samples form one cluster.
+Also, considering the early/late infection patterns in the parasite data and that these fit well with previous knowledge about asynchronous schizogony and gametocyte formation around 7 dpi, it is worthwhile to consider whether replicate separation reflects true biological variation.
+It is perceivable that the parasite accommodates to host variation with small differences in, e.g., host stress levels due to litter mates, draught, differences in light exposure or other factors which may vary also in a controlled animal facility.
+Parasite adjustment to such factors could explain the transcriptional profiles we see, with distinct overall patterns but replicate separation in the asexual phase.
 We suggest that considering such possibilities is useful for interpreting results and understanding basic biology of the parasite, and, in extension better understand infections in less homogeneous hosts than laboratory mice.
@@
 \subsection*{Evolutionary conservation in life cycle-characteristic gene groups}
-To gain insight about the conservation status of life stage-specific gene groups in 
-\textit{E. falciformis}, we tested gene clusters in Figure 3c for enrichment of gene orthologs which 
-are present in other species.
-Groups analyzed for orthologs are: i) one category with \textit{E.
-falciformis} only ("Efalci"), ii) one category of 10 apicomplexan parasites ("Api") (see xxx), iii) one category of Api but excluding \textit{Cryprosporidium hominis} ("Api minus C.h."), iv) one catogory with three \textit{Eimeria} species: \textit{E.
+A good understanding for similarities and differences between closely related parasites will be useful to draw parallels between the species, and importantly also to avoid incorrect assumptions about, e.g., gene function.
+One way to achieve such understanding is to identify which genes are species-specific, phylum specific or broadly conserved.
+To this aim, we tested gene clusters in Figure 3c for enrichment of gene orthologs which are present in other selected species.
+Groups analyzed for orthologs are: i) \textit{E.
+falciformis} only ("Efalci"), ii) 10 apicomplexan parasites ("Api") (see additional file xxx), iii) Api, but excluding \textit{Cryprosporidium hominis} ("Api minus C.h."), iv) three \textit{Eimeria} species: \textit{E.
 falciformis, E.
@@
 tenella} ("Eimeria"), and v) one conserved group containing a broad range of species with, e.g., \textit{Saccharomyces cerevisiae} and \textit{Arabidopsis thaliana} ("conserved").
-Sporozoites are enriched for \textit{E.
-falciformis} specific genes, and the group "conserved" is underrepresented, indicating a strong species specificity in the sporozoite stage.
+We identify sporozoites as the most \textit{E.
+falciformis} specific life cycle stage.
+Sporozoite characteristic genes are enriched for \textit{E.
+falciformis} unique genes, and the group "conserved" is underrepresented among these genes, further supporting \textit{E.
+falciformis}-specificity in the sporozoite stage.
 Gene clusters characteristic for early infection are enriched for "Api" (cluster 6), probably reflecting that similar processes and genes are involved in asexual reproduction of the selected apicomplexan species.
-In late infection (cluster 2), characteristic genes are underrepresented for conserved and "Api minus C.h", whereas the other characteristic cluster for late infection is enriched for "conserved" orthologs.
-This clustering and difference in enrichment of conserved versus apicomplexan (minus \textit{C.
-hominis}) genes probably reflects activity of generic and phylum specific processes/genes in oocysts.
-One oocyst cluster (cluster 1) is underrepresented for "Eimeria", whereas the other oocyst genes (cluster 5) are enriched for "Api". 
-\ldots
+This could have implications for, e.g., the use of stage specific markers or drugs which target the asexual stage of the parasite (examples?....).
+Late infection (cluster 2) genes are underrepresented for conserved and "Api minus C.h", indicating that the Eimeria and Efalci categories dominate among these genes.
+The other characteristic cluster for late infection (cluster 7) is enriched for "conserved" orthologs.
+This clustering and difference in enrichment of conserved versus apicomplexan genes (minus \textit{C.
+hominis}) probably reflects simultaneous activity of generic and phylum specific processes in gametes or oocyst formation.
+One oocyst cluster (cluster 1) is underrepresented for "Eimeria".
+It is possible that this result is driven by high presence of Efalci and Api or Api minus C.h.
+genes, which are not orthologs with the chicken \textit{Eimeria} species.
+The other oocyst genes (cluster 5) are enriched for "Api" and are therefore likely oocyst genes which are important for all apicomplexans analyzed here.
 
-In addition to analysing conservation by ortholog enrichment, we also performed Spearman correlation analysis between our RNA-seq transcriptomes and RNA-seq data from related parasites.
+In addition to analysing conservation by ortholog enrichment, we also performed Spearman's correlation analysis between our RNA-seq transcriptomes and RNA-seq data from related parasites.
 Two datasets for the economically important chicken parasite \textit{E.
@@
 tenella} merozoites from both independent studies are most similar to early \textit{E.
-falciformis} samples, indicating similarity also during asexual reproduction which is shown by the conservation analysis as well (Figure?/Table?).
+falciformis} samples, indicating similarity also during asexual reproduction which is also shown by the conservation analysis (Figure?/Table?).
+We speculate that this clustering is driven by genes in cluster 6 (Figure 3c) since they determine early stages and are enriched for shared apicomplexa orthologs.
 \textit{E.
@@
 tenella} sporozoites.
-%% SUM UP
+We have identified groups of genes which are shared among apicomplexa and linked them to the asexual reproduction phase of \textit{E.
+falciformis}.
+The analysis also returns \textit{E.
+falciformis} specific genes, which are abundant in sporozoites.
+One group of \textit{E.
+falciformis} oocyst genes (cluster 1, Figure 3c) are underrepresented for \textit{E.
+tenella} orthologs and can perhaps be used to understand differences between mammalian and avian \textit{Eimeria} spp.
+oocysts.
 
-In summary, we have performed a dual RNA-seq transcriptome of an apicomplexan parasite in its natural host.
+\subsection*{Conclusions}
+In summary, we have performed the first dual RNA-seq transcriptome of the full life-cycle of an apicomplexan parasite in its natural host.
 Our analysis of differentially abundant mRNAs at different time-points post infection highlights large groups of genes which characterize the different life stages of the parasite.
-The dual approach also allows insight into the host responses to this intracellular infection.
-Using GO enrichment tests and by analysing how species specific these life stage defining gene groups are, we show that the transcriptional profile of sporozoites is the most \textit{E.
-falciformis} specific stage of the time-points we have analyzed.
+The dual approach adds insight into the host responses to this intracellular infection, e.g., demonstrating that the host transcriptome is different in a first versus challenge infection, whereas the parasite seems to be transcriptionally "unaware" of the differences in mouse immune competence.
+We show that the transcriptional profile of sporozoites is the most \textit{E.
+falciformis}-specific stage.
 The sporozoite defining gene group is characterized by ATP production, regulation of protein localization and biosynthetic processes.
-The analysis further demonstrates that early infection, 3 and 5 dpi, is not homogenous between samples, and that the conformity seen in sporozoites (and oocysts) is regained late in infection.
-We speculate that parasites during asexual replication are more asynchronous than late stage merozoites or gametocytes on 7 dpi, possibly due to an increasing influence from the host adaptive immune response.
-The asynchrone early parasite asexual replication stages are enriched for growth related processes, and a large expansion in parasite numbers is reflected in the percentage of parasite genes, read numbers as well asin RT-qPCR data.
-Gametocyte formation on 7 dpi is supported by GO enrichments of terms for motility and ATP production, as well as 7 dpi samples correlating strongly with \textit{E.
-tenella} gametocyte samples.
-Our analysis interestingly does not detect any influence of mouse immune status on the parasite transcriptome, even though a difference in oocyst output between Rag1-/- and its C57BL/6 control is clear.
-We sequenced enriched mRNA and it is possible that non-coding RNAs or other post transcriptional regulatory mechanisms in the parasite could explain differences in oocyst output.
-Another possibility is that immune competent mice mainly interfere with the parasite at a very late time-point before oocyst formation, which a different experimental design could elucidate.
+The analysis further demonstrates that early infection is not synchronized between samples, neither in host nor parasite.
+For the parasite, the conformity which is seen in sporozoites and oocysts is however regained late in infection, where both parasite and host transcriptomes are synchronized between samples.
+We speculate that parasites during asexual replication are more asynchronous than developing gametes and oocysts on 7 dpi.
+The distinct profile on 7 dpi in both host and parasite suggests that synchronization is driven by interactive processes between the species.
+Whether gamete formation triggers the adaptive immune responses seen in the host, or previously activated immune responses lead to gamete and oocyst formation remains to be investigated.
 
@@
 \subsection*{Sporozoite isolation}
-Sporozoites were isolated from sporocysts by excystment. For this,
+Sporozoites were isolated from sporocysts by in vitro excystation. For this,
 sporocysts were incubated at 37\newcommand{\degree}C in DMEM containing 0.04\%
@@
 mRNA sequences from both species were aligned against the dual
-reference genome using TopHat2 (version 2.0.14, Trapnell09) with -G
-specified, and a Bowtie2 (version 1.1.2, Langmead12) index of the dual
+reference genome using
+
+###adjust### 
+TopHat2 (version 2.0.14, Trapnell09) with -G
+specified, and a ####Bowtie2##### (version 1.1.2, Langmead12) index of the dual
 genome. Single-end and pair-end sequence samples were aligned
@@
 Cufflinks (version 2.1.1, Trapnell10).
+###########################################
 
@@
 clustering of \textit{E. falciformis} life cycle relevant genes. In
-each comparison (see Table 3), the union of the at most 500 genes
-differentially abundant with lowest FDR ($<$0.05) are selected. In the
-next step, the 500 mRNAs from each comparison (or less) are
+each comparison (see Table 3), the union of significantly different genes (mRNAs)
+were selected. In the next step, the mRNAs from each comparison are
 joined. In heatmaps, all samples, i.e., also samples which did not have any
@@
 down-regulation (brown) denote number of standard deviations from 0,
-i.e., row mean. Hierarchical clustering was performed using with
+i.e., row mean. Hierarchical clustering was performed using
 Euclidean distances, by the complete linkage method ('complete', R package
@@
 
-All analyses were performed in R (cite R-core). Complete scripts are
+All (### alignments too?) ### analyses were performed in R (cite R-core). Complete scripts are
 available at https://github.com/derele/Ef\_RNAseq.git tagged as version

### 2016-10-20 10_combining_mouseEf.R
RnB finally running.

@@ -66,18 +66,23 @@ registerDoParallel(cores=20)
 outer.reps <- 5000
 inner.reps <- 100
 
+## uncomment for testing
+## outer.reps <- 10
+## inner.reps <- 20
 
 ## sum up parallel the numbers the random correlation values are
 ## bigger than the real correlations
 parallel.RnB <- function(inner.reps){
     randCors <-
         foreach(icount(inner.reps), .combine="+") %dopar% {
-            rm(list=ls())
             rand.col <- sample(both.col, length(both.col))
             RC <- cor(t(Mm.nRC[, rand.col]),
                       t(Ef.nRC[, both.col]))        
             res <- abs(RC)>abs(Cors)
-            gc()
+            ## avoid an accumulation of NAs for In cor(t(Mm.nRC[,
+            ## both.col]), t(Ef.nRC[, both.col])) : the standard
+            ## deviation is zero
+            res[is.na(res)] <- TRUE
             return(res)
         }
     return(randCors)
@@ -86,13 +91,14 @@ parallel.RnB <- function(inner.reps){
 
 ## a large number of non-parallelized chunks. Putting everything in
 ## the parallel function would eat up memory
-RnB <- matrix()
+RnB <- matrix(0, nrow=nrow(Cors), ncol=ncol(Cors))
 for(i in 1:outer.reps){
     ## sum each result with what has been there before
-    RnB <- do.call(sum, list(RnB, parallel.RnB(inner.reps)))
+    RnB.new <- parallel.RnB(inner.reps)
     print(paste("finished outer rep", i))
+    RnB <- RnB + RnB.new
 }
 
-RnB.final <- RnB.sum / (outer.reps*inner.reps)
+RnB.final <- RnB / (outer.reps*inner.reps)
 
-all.x <- apply(RnB.final, 2, function(x) sum(x<0.01))
+all.x <- apply(RnB.final, 2, function(x) sum(x==0))

### 2016-10-20 2_data_curation.R

@@ -220,5 +220,8 @@ table.tex <- xtable(table.cleaned, align = c("l", "l", "l", "l", "l", "l", "l", 
 print(table.tex, type = "latex", file = "tables/Table1_ReadCounts.tex", include.rownames = F,
       format.args = list(big.mark = ",", decimal.mark = "."))
 
+print(table.tex, type = "html", file = "tables/Table1_ReadCounts.html", include.rownames = F,
+      format.args = list(big.mark = ",", decimal.mark = "."))
+
 
 
